---
title: "20250908_workflow_MastCellSeurat"
output: html_document
date: "2025-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Prerequisites -------------------------

## [0.0.] Load necessary packages
#### Analysis initialised with Seurat package version 4.3.0.1
```{r message=FALSE}
base::library(tidyverse)
base::library(magrittr)
base::library(ggraph)
base::library(Seurat)
base::library(reticulate)

utils::packageVersion('Seurat') #4.3.0.1
```

## [0.1.] Specify conda environment to use leidenalg module
#### reticulate::use_python(python = "/home/pbr316/.conda/envs/20220620_placenta/bin/python3.8", required = TRUE)
#### see https://www.reddit.com/r/bioinformatics/comments/1jbgjt7/how_to_correctly_install_leidenalg_for_seurat/
```{r}
reticulate::use_condaenv(
    condaenv = "/home/pbr316/.conda/envs/20220620_placenta/",
    conda = "auto",
    required = TRUE
  )
reticulate::py_config()
```

## [0.2.] Identify doublets via Scrublet
#### Use py$dbls and py$predicted_doublets to make the array availabel in R
```{python}
from IPython import get_ipython
get_ipython().run_line_magic('matplotlib', 'inline')
import scrublet as scr
import scipy.io
import matplotlib.pyplot as plt
import numpy as np
import os

input_dir = '/maps/projects/ralab/data/projects/nucleiCAGEproject/12.BloodAnalysis/data/GSE184351/GSM5584896_Sample1_P15054_filtered_feature_bc_matrix/'
counts_matrix = scipy.io.mmread(input_dir + 'matrix.mtx').T.tocsc()
genes = np.array(scr.load_genes(input_dir + 'features.tsv', delimiter='\t', column=2))

print('Counts matrix shape: {} rows, {} columns'.format(counts_matrix.shape[0], counts_matrix.shape[1]))
print('Number of genes in gene list: {}'.format(len(genes)))

scrub = scr.Scrublet(counts_matrix , expected_doublet_rate = 0.041)
doublet_scores, predicted_doublets = scrub.scrub_doublets(
    min_counts = 2,
    min_cells = 3,
    min_gene_variability_pctl = 85,
    n_prin_comps = 30
  )
predicted_doublets
len(predicted_doublets)
len(np.where(predicted_doublets)[0])
dbls = np.where(predicted_doublets)[0] + 1
dbls
```

## [0.2.] Save & load environment
```{r}
# base::save.image(
#   file = base::paste0(
#         here::here("12.BloodAnalysis/data/"),
#         "20250908_Seurat.RData"
#       ),
#     compress = TRUE
#   )
```

```{r}
base::load(
  file = base::paste0(
      here::here("12.BloodAnalysis/data/"),
      "20250908_Seurat.RData"
    )
  )
```

## [0.3.] Read in cellranger data
#### Lin(-) c-Kit(+) 33,538 genes x 7,336 cells
```{r}
LincKit_mat <- Seurat::Read10X(
  data.dir = base::paste0(
      here::here("12.BloodAnalysis/data/GSE184351/"),
      "GSM5584896_Sample1_P15054_filtered_feature_bc_matrix"
    )
  )
```

#### Lin(-) CD34(+) c-Kit(+) 36,601 genes x 6,794,880 cells
```{r}
LinCD34cKit_mat <- Seurat::Read10X(
  data.dir = base::paste0(
      here::here("12.BloodAnalysis/data/GSE184351/"),
      "GSM5584897_Sample2_P20409_1003-4_filtered_feature_bc_matrix"
    )
  )
```

# [1.] Seurat filtering -------------------------

## [1.0.] Generate Seurat object(s) & add doublets information
```{r}
LincKit <- Seurat::CreateSeuratObject(
    counts = LincKit_mat,
    project = "Lin(-)cKit(+)",
    names.field = 1, ## Initial identity for the cell is the barcode,
    names.delim = "-",
    assay = "RNA",
    min.cells = 3,
    min.features = 500
  )
```

## Add doublet information
#### Evaluate with first doublet barcode: 'AAAGATGCAACACCTA-1'
```{r}
LincKit_all <- Seurat::CreateSeuratObject(
    counts = LincKit_mat,
    project = "Lin(-)cKit(+)",
    names.field = 1, ## Initial identity for the cell is the barcode,
    names.delim = "-",
    assay = "RNA",
    min.cells = 0,
    min.features = 0
  )

meta_all <- LincKit_all@meta.data %>%
  dplyr::mutate(
      dbl = py$predicted_doublets
    ) %>%
  tibble::rownames_to_column(var = "barcode") %>%
  dplyr::select(barcode, dbl)
  
  
LincKit[["dbl"]] <- LincKit@meta.data %>%
  tibble::rownames_to_column(var = "barcode") %>% 
  dplyr::left_join(
      x = .,
      y = meta_all,
      by = "barcode"
    ) %>%
  dplyr::pull(dbl)
```

## [1.1.] Identify mitochondrial genes
####  13 'MT-' genes in the dataset
```{r}
stringr::str_subset(
    string = LincKit@assays$RNA@counts@Dimnames[[1]],
    pattern = "^MT-|^mt-"
  )

LincKit[["percent.mt"]] <- Seurat::PercentageFeatureSet(
    object = LincKit,
    pattern = "^MT-|^mt-"
  )
```

## [1.2.] Subset data
```{r}
LincKit <- base::subset(
    x = LincKit,
    subset = percent.mt < 5 & nFeature_RNA < 4000
  )
```

#### Evaluation
#### 7074 cells left specifically 'including' the 200 doublets identified by Wu C et al., 2022
```{r}
LincKit@meta.data
```

## [1.3.] Violin plot for QC features
```{r}
Seurat::VlnPlot(
  object  = LincKit,
  features = base::c(
        "nFeature_RNA","nCount_RNA","percent.mt"
        #"G2M.Score", "S.Score"
      ),
    ncol = 5
  )
```

## [1.4.] Scatterplot number of features and mitochondrial reads per cell
```{r}
ggplot2::ggplot(
      data = LincKit@meta.data,
      mapping = aes(
        x = nFeature_RNA,
        y = percent.mt,
        colour = dbl
      )
    ) +
  ggrastr::geom_point_rast(
      size = 0.1,
      alpha = 0.5
    ) +
  # geom_density2d(
  #     binwidth = 0.003
  #   ) +
  geom_vline(
      xintercept = base::c(200,2500),
      linetype = "dashed"
    ) +
  geom_hline(
      yintercept = base::c(5,10),
      linetype = "dashed"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"line")
    )
```

## [1.5.] Scatterplot number of features and number of counts oer cell
```{r}
ggplot2::ggplot(
      data = LincKit@meta.data,
      mapping = aes(
        x = nCount_RNA,
        y = nFeature_RNA,
        colour = dbl
      )
    ) +
  ggrastr::geom_point_rast(
      size = 0.1,
      alpha = 0.5
    ) +
  # geom_density2d(
  #     binwidth = 0.003
  #   ) +
  geom_hline(
      yintercept = base::c(1500),
      linetype = "dashed"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"line")
    )
```

# [2.] Data processing -------------------------

## [2.0.] Normalization
#### the 'data' bucket of LincKit@assays$RNA@data
```{r}
LincKit <- Seurat::NormalizeData(
    object = LincKit,
    normalization.method = "LogNormalize",
    scale.factor = 10000
  )
```

## [2.1.] Find variable features
```{r}
LincKit <- Seurat::FindVariableFeatures(
    object = LincKit,
    selection.method = "vst",
    nfeatures = 2000
  )
```

#### Evaluations
```{r}
varFeat <- Seurat::VariableFeatures(
    object = LincKit
  )[1:10]
  
plot <- Seurat::VariableFeaturePlot(
    object = LincKit
  )
Seurat::LabelPoints(
    plot = plot,
    points = varFeat,
    repel = TRUE,
    xnudge = 0,
    ynudge = 0
  )
```

## [2.2] Data scaling
#### results are stored in 'LincKit@assays$RNA@scale.data', evaluate with LincKit@assays$RNA@scale.data[1:5,1:5]
```{r}
all.genes <- base::row.names(x = LincKit)
LincKit <- Seurat::ScaleData(
    object = LincKit,
    features = all.genes
  )
```

## [2.3] Perform PCA
```{r}
LincKit <- Seurat::RunPCA(
  object = LincKit,
  npcs = 50,
  features = Seurat::VariableFeatures(
      object = LincKit
    )
  )
```

#### Evaluate
```{r}
Seurat::DimPlot(
    object = LincKit
  )
Seurat::ElbowPlot(
    object = LincKit
  )
```

## [2.3.] Compute nearest neighbour graph through PCA space
#### Leiden clustering based on leidenalg and igraph python modules
#### see [0.1.] for coupling with conda enviornment including these modules
```{r}
LincKit <- Seurat::FindNeighbors(
    object = LincKit,
    reduction = "pca",
    dims = 1:15
  )
# base::debugonce(Seurat::FindClusters)
LincKit <- Seurat::FindClusters(
    object = LincKit,
    resolution = 0.5,
    algorithm = 4 # Leiden based on 'leidenalg' module
  )

dplyr::distinct(LincKit@meta.data, seurat_clusters)
```

## [2.4] Embedd data via UMAP
```{r}
LincKit <- Seurat::RunUMAP(
    object = LincKit,
    dims = 1:15
  )
```

```{r}
Seurat::DimPlot(
    object = LincKit,
    reduction = "umap"
  )
```

