---
title: "20250908_workflow_MastCellSeurat"
output: html_document
date: "2025-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [0.0.] Load necessary packages
#### Analysis initialised with Seurat package version 4.3.0.1
```{r message=FALSE}
base::library(tidyverse)
base::library(magrittr)
base::library(ggraph)
base::library(Seurat)

utils::packageVersion('Seurat') #4.3.0.1
```

## [0.1.] Save & load environment
```{r}
# base::save.image(
#   file = base::paste0(
#         here::here("12.BloodAnalysis/data/"),
#         "20250908_Seurat.RData"
#       ),
#     compress = TRUE
#   )
```

```{r}
base::load(
  file = base::paste0(
      here::here("12.BloodAnalysis/data/"),
      "20250908_Seurat.RData"
    )
  )
```

## [0.2.] Read in cellranger data
#### Lin(-) c-Kit(+) 33,538 genes x 7,336 cells
```{r}
LincKit_mat <- Seurat::Read10X(
  data.dir = base::paste0(
      here::here("12.BloodAnalysis/data/GSE184351/"),
      "GSM5584896_Sample1_P15054_filtered_feature_bc_matrix"
    )
  )
```

#### Lin(-) CD34(+) c-Kit(+) 36,601 genes x 6,794,880 cells
```{r}
LinCD34cKit_mat <- Seurat::Read10X(
  data.dir = base::paste0(
      here::here("12.BloodAnalysis/data/GSE184351/"),
      "GSM5584897_Sample2_P20409_1003-4_filtered_feature_bc_matrix"
    )
  )
```

# [1.] Seurat filtering -------------------------

## [1.0.] Generate Seurat object(s)
```{r}
LincKit <- Seurat::CreateSeuratObject(
    counts = LincKit_mat,
    project = "Lin(-)cKit(+)",
    names.field = 1, ## Initial identity for the cell is the barcode,
    names.delim = "-",
    assay = "RNA",
    min.cells = 3,
    min.features = 500
  )
```

## [1.1.] Identify mitochondrial genes
####  13 'MT-' genes in the dataset
```{r}
stringr::str_subset(
    string = LincKit@assays$RNA@counts@Dimnames[[1]],
    pattern = "^MT-|^mt-"
  )

LincKit[["percent.mt"]] <- Seurat::PercentageFeatureSet(
    object = LincKit,
    pattern = "^MT-|^mt-"
  )
```

## [1.2.] Subset data
```{r}
LincKit <- base::subset(
    x = LincKit,
    subset = percent.mt < 5 & nFeature_RNA < 4000
  )
```

#### Evaluation
#### 7074 cells left specifically 'including' the 200 doublets identified by Wu C et al., 2022
```{r}
LincKit@meta.data
```


## [1.3.] Violin plot for QC features
```{r}
Seurat::VlnPlot(
  object  = LincKit,
  features = base::c(
        "nFeature_RNA","nCount_RNA","percent.mt"
        #"G2M.Score", "S.Score"
      ),
    ncol = 5
  )
```

## [1.4.] Scatterplot number of features and mitochondrial reads per cell
```{r}
ggplot2::ggplot(
      data = LincKit@meta.data,
      mapping = aes(
        x = nFeature_RNA,
        y = percent.mt
      )
    ) +
  ggrastr::geom_point_rast(
      size = 0.1,
      alpha = 0.05
    ) +
  # geom_density2d(
  #     binwidth = 0.003
  #   ) +
  geom_vline(
      xintercept = base::c(200,2500),
      linetype = "dashed"
    ) +
  geom_hline(
      yintercept = base::c(5,10),
      linetype = "dashed"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"line")
    )
```

## [1.5.] Scatterplot number of features and number of counts oer cell
```{r}
ggplot2::ggplot(
      data = LincKit@meta.data,
      mapping = aes(
        x = nCount_RNA,
        y = nFeature_RNA
      )
    ) +
  ggrastr::geom_point_rast(
      size = 0.1,
      alpha = 0.05
    ) +
  # geom_density2d(
  #     binwidth = 0.003
  #   ) +
  geom_hline(
      yintercept = base::c(1500),
      linetype = "dashed"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"line")
    )
```

# [2.] Data processing -------------------------

## [2.0.] Normalization
#### the 'data' bucket of LincKit@assays$RNA@data
```{r}
LincKit <- Seurat::NormalizeData(
    object = LincKit,
    normalization.method = "LogNormalize",
    scale.factor = 10000
  )
```

## [2.1.] Find variable features
```{r}
LincKit <- Seurat::FindVariableFeatures(
    object = LincKit,
    selection.method = "vst",
    nfeatures = 2000
  )
```

#### Evaluations
```{r}
varFeat <- Seurat::VariableFeatures(
    object = LincKit
  )[1:10]
  
plot <- Seurat::VariableFeaturePlot(
    object = LincKit
  )
Seurat::LabelPoints(
    plot = plot,
    points = varFeat,
    repel = TRUE,
    xnudge = 0,
    ynudge = 0
  )
```

## [2.2] Data scaling
#### results are stored in 'LincKit@assays$RNA@scale.data', evaluate with LincKit@assays$RNA@scale.data[1:5,1:5]
```{r}
all.genes <- base::row.names(x = LincKit)
LincKit <- Seurat::ScaleData(
    object = LincKit,
    features = all.genes
  )
```

## [2.3] Perform PCA
```{r}
LincKit <- Seurat::RunPCA(
  object = LincKit,
  npcs = 50,
  features = Seurat::VariableFeatures(
      object = LincKit
    )
  )
```

#### Evaluate
```{r}
Seurat::DimPlot(
    object = LincKit
  )
Seurat::ElbowPlot(
    object = LincKit
  )
```

## [2.3.] Compute nearest neighbour graph through PCA space
```{r}
LincKit <- Seurat::FindNeighbors(
    object = LincKit,
    reduction = "pca",
    dims = 1:15
  )
base::debugonce(Seurat::FindClusters)
LincKit <- Seurat::FindClusters(
    object = LincKit,
    resolution = 0.5,
    algorithm = 4
  )
```

```{r}
ids <- RunLeiden(object = object, method = method, 
            partition.type = "RBConfigurationVertexPartition", 
            initial.membership = initial.membership, 
            node.sizes = node.sizes, resolution.parameter = r, 
            random.seed = random.seed, n.iter = n.iter)

reticulate::py_module_available(module = "leidenalg")

base::debugonce(reticulate::import_from_path)
reticulate::import_from_path(module = "leidenalg", path = "/home/pbr316/.local/lib/python3.10/site-packages/")
reticulate::py_last_error()
reticulate:::py_module_import(module = "leidenalg", convert = convert)

```

```{r}
path <- base::normalizePath(path)
sys <- reticulate::import("sys", convert = FALSE)
```


```{python}
sys$path$insert(0L, path)
on.exit(sys$path$remove(path), add = TRUE)
import(module, convert = convert)
```

```{r}
reticulate::use_python(python = "/home/pbr316/.conda/envs/20220620_placenta", required = TRUE)
```

