---
title: "1.1. CTSSs preprocessing"
output: html_document
date: "2024-05-17"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
base::library(PRIME)
```

## [0.1.] Register server cores / threads
```{r}
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 10
    )
  )
```

## [0.2.] Set working directory
```{r eval=FALSE}
#base::setwd("/maps/projects/ralab/data/projects/nucCAGE_PRIME_paper/1.CTSSs")
```

# [1.] Process K562 datasets -------------------------

## [1.0.] Read in K562 design matrix
```{r}
design_ <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "K562.design.txt"
      ),
    sep = "\t",
    header = T
  )
```

## [1.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::filter(use == "yes") %>%
  tibble::column_to_rownames(var = "ID")
```

## [1.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [1.3.] Quantify CTSSs with no reference genome set
```{r}
k562.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )
k562.CTSSs <- CAGEfightR::calcTotalTags(k562.CTSSs)
```

## [1.4.] Visualize amount of tags across libraries
```{r}
SummarizedExperiment::colData(k562.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      ID = base::paste(Cell.line,Type,replicate,sep = "_"),
      ID = forcats::as_factor(x = ID)
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = totalTags/1000000,
        y = ID #base::paste(Type,replicate,sep="_")
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_y_discrete(
      limits = rev
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank()
    )
```



## [1.5.] Keep only standard chromosomes and pool replicates
#### Evaluate with SummarizedExperiment::colData(k562.p.CTSSs), SummarizedExperiment::colData(k562.CTSSs)
```{r}
keep <- GenomeInfoDb::seqlevels(k562.CTSSs)[1:23]

k562.CTSSs <- GenomeInfoDb::keepSeqlevels(k562.CTSSs, keep, pruning.mode = "coarse")


k562.CTSSs <- k562.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()

k562.p.CTSSs <- PRIME::poolReplicates(
  object = k562.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(k562.CTSSs)$Cell.line,
      SummarizedExperiment::colData(k562.CTSSs)$Type,
      sep="_"
    )
  )

k562.p.CTSSs <- CAGEfightR::calcTPM(k562.p.CTSSs)
```




## [1.6.] Combine into one object
```{r}
SummarizedExperiment::colData(k562.CTSSs)$pooled <- "no"
SummarizedExperiment::colData(k562.p.CTSSs)$pooled <- "yes"

k562.combined.CTSSs <- BiocGenerics::cbind(k562.CTSSs,k562.p.CTSSs)
```


## [1.7.] Subsample to minimum seq depth of CAGE across replicates and across pooled samples
```{r}
#Subsample both original and pooled to minimum common depth
set.seed(1337)

#subsample combined. Three sets, lowest CAGE rep and lowest pooled (excluding proCap and groCap)
rep.min <- min(colData(k562.combined.CTSSs)$totalTags[colData(k562.combined.CTSSs)$pooled=="no"& !colData(k562.combined.CTSSs)$method %in% c("ProCap","GroCap")]) #28619974
pool.min <- min(colData(k562.combined.CTSSs)$totalTags[colData(k562.combined.CTSSs)$pooled=="yes" & !colData(k562.combined.CTSSs)$method %in% c("ProCap,ProCap","GroCap")]) #96351121

k562.combined.repSub.CTSSs<- PRIME::subsampleTarget(k562.combined.CTSSs, target = rep.min)
k562.combined.poolSub.CTSSs<- PRIME::subsampleTarget(k562.combined.CTSSs, target = pool.min)


k562.combined.repSub.CTSSs <- CAGEfightR::calcTPM(k562.combined.repSub.CTSSs)
k562.combined.poolSub.CTSSs <- CAGEfightR::calcTPM(k562.combined.poolSub.CTSSs)


k562.combined.repSub.rmS.CTSSs <- PRIME::rmSingletons(k562.combined.repSub.CTSSs)
k562.combined.poolSub.rmS.CTSSs <- PRIME::rmSingletons(k562.combined.poolSub.CTSSs)
```

# [1.8.] Save CTSS objects
```{r}
base::saveRDS(
  object = k562.combined.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.combined.repSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.combined.poolSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.combined.repSub.rmS.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
base::saveRDS(
  object = k562.combined.poolSub.rmS.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
```


## [1.9.] TMM normalise data
```{r}
facs <- edgeR::calcNormFactors(
    object = k562.CTSSs,
    method = "TMM",
    refColumn = NULL
  )
tmm.CTSSs <- PRIME::TPMnormalizeBySizeFactors(
    object = k562.CTSSs,
    sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
    inputAssay = "counts",
    outputAssay = "TMM"
  )
```

## [1.10.] Write out the TMM normalized replicates as bigwig files
```{r}
dir <- base::paste0(here::here("1.CTSSs/"),"bw_files_tmm/")
if (!dir.exists(dir))
  dir.create(dir)

library(doParallel)
registerDoParallel(cores = 10)

PRIME::writeBw(
    object = tmm.CTSSs,
    replicates = base::rownames(SummarizedExperiment::colData(tmm.CTSSs)),
    dir
  )
```

## [1.10.1.] Save TMM normalized CTSSs
```{r}
base::saveRDS(
  object = tmm.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_tmm.CTSSs.rds"
    )
  )
```

## [1.10.2.] Load saved rds files
```{r}
fn <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("tmm.CTSSs")

tmm.CTSSs <- base::readRDS(file = fn)
```


#Downstream needs modifying
# [2.] Process GM12878 datasets -------------------------

## [2.0.] Read in GM12878 design matrix
```{r}
design_ <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "GM12878.design.txt"
      ),
    sep = "\t",
    header = TRUE
  )
```

## [2.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::filter(use == "yes") %>%
  dplyr::mutate(
      rn = base::paste(ID,Seq_run,sep="_")
    ) %>% 
  tibble::column_to_rownames(var = "rn")
```

## [2.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [2.3.] Quantify CTSSs with no reference genome set
```{r}
GM12878.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )
GM12878.CTSSs <- CAGEfightR::calcTotalTags(GM12878.CTSSs)
```

## [2.4.] Visualize amount of tags across libraries
```{r}
SummarizedExperiment::colData(GM12878.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      ID = base::paste(Cell.line,Type,replicate,sep = "_"),
      ID = forcats::as_factor(x = ID)
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = totalTags/1000000,
        y = ID #base::paste(Type,replicate,sep="_")
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_y_discrete(
      limits = rev
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank()
    )
```

## [2.5.] Pool replicates
#### Evaluate with SummarizedExperiment::colData(k562.p.CTSSs), SummarizedExperiment::colData(k562.CTSSs)
```{r}
keep <- GenomeInfoDb::seqlevels(GM12878.CTSSs)[1:23]

GM12878.CTSSs <- GenomeInfoDb::keepSeqlevels(GM12878.CTSSs, keep, pruning.mode = "coarse")

GM12878.CTSSs <- GM12878.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()

GM12878.p.CTSSs <- PRIME::poolReplicates(
  object = GM12878.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(GM12878.CTSSs)$Cell.line,
      SummarizedExperiment::colData(GM12878.CTSSs)$cell_count,
      SummarizedExperiment::colData(GM12878.CTSSs)$Type,
      sep = "_"
    )
  )

GM12878.p.CTSSs <- CAGEfightR::calcTPM(GM12878.p.CTSSs)
```

## [2.6.] Pool only based on reqsequencing
```{r}
GM12878.pID.CTSSs <- PRIME::poolReplicates(
    object = GM12878.CTSSs,
    replicates = SummarizedExperiment::colData(GM12878.CTSSs)$ID
  )

GM12878.pID.CTSSs <- CAGEfightR::calcTPM(GM12878.pID.CTSSs)
```

## [2.7.] Create combined objects
# Create object that also contains fully pooled samples.
#### Main indicates if part of main analysis. Removing Triptolide and low input
```{r}
SummarizedExperiment::colData(GM12878.pID.CTSSs)$main <- "yes"
SummarizedExperiment::colData(GM12878.pID.CTSSs)$main[
    SummarizedExperiment::colData(GM12878.pID.CTSSs)$Type %in% base::c("Cap_align","ctrl","trip") |
    SummarizedExperiment::colData(GM12878.pID.CTSSs)$cell_count %in% base::c("50K","250K,250K")
  ] <- "no"
SummarizedExperiment::colData(GM12878.pID.CTSSs)$pooled <- "no"
```

```{r}
SummarizedExperiment::colData(GM12878.p.CTSSs)$main <- "yes"
SummarizedExperiment::colData(GM12878.p.CTSSs)$main[
    SummarizedExperiment::colData(GM12878.p.CTSSs)$Type %in% base::c("Cap_align","ctrl,ctrl,ctrl","trip,trip,trip,trip") |
    SummarizedExperiment::colData(GM12878.p.CTSSs)$cell_count %in% base::c("50K,50K,50K","250K,250K,250K,250K,250K,250K")
  ] <- "no"
SummarizedExperiment::colData(GM12878.p.CTSSs)$pooled <- "yes"
```

```{r}
GM12878.combined.CTSSs <- BiocGenerics::cbind(GM12878.p.CTSSs,GM12878.pID.CTSSs)
GM12878.combined.CTSSs <- CAGEfightR::calcTotalTags(GM12878.combined.CTSSs)
```

## [2.8.] Create two subsamples one to the min replicate depth and one to min pool depth and keep original
```{r}
rep.min <- base::min(
  SummarizedExperiment::colData(GM12878.combined.CTSSs)$totalTags[
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$main == "yes" &
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$pooled == "no"
    ]
  ) #29027358
pool.min <- base::min(
  SummarizedExperiment::colData(GM12878.combined.CTSSs)$totalTags[
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$main == "yes" &
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$pooled == "yes"
    ]
  ) #106618677

GM12878.combined.repSub.CTSSs<- PRIME::subsampleTarget(
    object = GM12878.combined.CTSSs,
    target = rep.min
  )
GM12878.combined.poolSub.CTSSs<- PRIME::subsampleTarget(
      object = GM12878.combined.CTSSs,
      target = pool.min
    )

GM12878.combined.repSub.CTSSs <- CAGEfightR::calcTPM(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- CAGEfightR::calcTPM(GM12878.combined.poolSub.CTSSs)
```

```{r}
GM12878.combined.repSub.rmS.CTSSs <- PRIME::rmSingletons(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.rmS.CTSSs <- PRIME::rmSingletons(GM12878.combined.poolSub.CTSSs)
```

## [2.9.] Save CTSS objects
```{r}
base::saveRDS(
    object = GM12878.combined.CTSSs,
    file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )
base::saveRDS(
  object = GM12878.combined.repSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rds"
    )
  )
base::saveRDS(
  object = GM12878.combined.poolSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rds"
    )
  )
base::saveRDS(
  object = GM12878.combined.repSub.rmS.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
base::saveRDS(
  object = GM12878.combined.poolSub.rmS.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```



# [3.] Process HepG2, HCT116 and A549  datasets -------------------------

## [3.0.] Read in design matrixes
```{r}
a549.design <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "A549.design.txt"
      ),
    sep = "\t",
    header = T
  )

hct.design <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "HCT116.design.txt"
      ),
    sep = "\t",
    header = T
  )

hep.design <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "HepG2.design.txt"
      ),
    sep = "\t",
    header = T
  )
```


## [3.1.0] Prepare and quantify A549
```{r}
base::row.names(a549.design) <- a549.design$ID
plus_files <- a549.design$bw_plus
minus_files <- a549.design$bw_minus

base::names(plus_files) <- base::rownames(a549.design)
base::names(minus_files) <- base::rownames(a549.design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)

a549.CTSSs <- CAGEfightR::quantifyCTSSs(plusStrand = bw_plus,
                       minusStrand = bw_minus,
                       design = a549.design)


a549.CTSSs <- a549.CTSSs %>% CAGEfightR::calcTPM() %>% CAGEfightR::calcPooled()
```


## [3.1.1] Prepare and quantify HCT116
```{r}
base::row.names(hct.design) <- hct.design$ID
plus_files <- hct.design$bw_plus
minus_files <- hct.design$bw_minus

base::names(plus_files) <- base::rownames(hct.design)
base::names(minus_files) <- base::rownames(hct.design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)

hct.CTSSs <- CAGEfightR::quantifyCTSSs(plusStrand = bw_plus,
                       minusStrand = bw_minus,
                       design = hct.design)


hct.CTSSs <- hct.CTSSs %>% CAGEfightR::calcTPM() %>% CAGEfightR::calcPooled()
```


## [3.1.2] Prepare and quantify HepG2
```{r}
base::row.names(hep.design) <- hep.design$ID
plus_files <- hep.design$bw_plus
minus_files <- hep.design$bw_minus

base::names(plus_files) <- base::rownames(hep.design)
base::names(minus_files) <- base::rownames(hep.design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)

hep.CTSSs <- CAGEfightR::quantifyCTSSs(plusStrand = bw_plus,
                       minusStrand = bw_minus,
                       design = hep.design)


hep.CTSSs <- hep.CTSSs %>% CAGEfightR::calcTPM() %>% CAGEfightR::calcPooled()
```

## [3.2.] Remove non standard chromosomes
```{r}
keep <- GenomeInfoDb::seqlevels(hct.CTSSs)[1:23]

a549.CTSSs <- GenomeInfoDb::keepSeqlevels(a549.CTSSs, keep, pruning.mode = "coarse")
hct.CTSSs <- GenomeInfoDb::keepSeqlevels(hct.CTSSs, keep, pruning.mode = "coarse")
hep.CTSSs <- GenomeInfoDb::keepSeqlevels(hep.CTSSs, keep, pruning.mode = "coarse")
```


## [3.3.] Pool replicates
```{r}
#Pool replicates
a549.p.CTSSs <- PRIME::poolReplicates(a549.CTSSs,
                               replicates=colData(a549.CTSSs)$Type)

a549.p.CTSSs <- CAGEfightR::calcTPM(a549.p.CTSSs)

dummy <- hct.CTSSs[,1]
BiocGenerics::colnames(dummy) <- "dummy"
for (i in base::seq_along(SummarizedExperiment::assays(dummy))) {
    SummarizedExperiment::assays(dummy)[[i]][] <- 0
}
SummarizedExperiment::colData(dummy)[] <- 0
hct.dummy.CTSSs <- BiocGenerics::cbind(hct.CTSSs, dummy)

hct.p.CTSSs <- PRIME::poolReplicates(hct.dummy.CTSSs,
                               replicates=colData(hct.dummy.CTSSs)$Type)
hct.p.CTSSs <- hct.p.CTSSs[, BiocGenerics::colnames(hct.p.CTSSs) != "0"]

hct.p.CTSSs <- CAGEfightR::calcTPM(hct.p.CTSSs)

hep.p.CTSSs <- PRIME::poolReplicates(hep.CTSSs,
                               replicates=colData(hep.CTSSs)$Type)

hep.p.CTSSs <- CAGEfightR::calcTPM(hep.p.CTSSs)
```

## [3.3.] Susample to lowest seq depth per cell line
```{r}
#Calc minimum seq depth
a549.min <- base::min(SummarizedExperiment::colData(a549.p.CTSSs)$totalTags)
hct.min <- base::min(SummarizedExperiment::colData(hct.p.CTSSs)$totalTags)
hep.min <- base::min(SummarizedExperiment::colData(hep.p.CTSSs)$totalTags)

a549.p.s.CTSSs <- PRIME::subsampleTarget(a549.p.CTSSs, target = a549.min)
hct.p.s.CTSSs <- PRIME::subsampleTarget(hct.p.CTSSs, target = hct.min)
hep.p.s.CTSSs <- PRIME::subsampleTarget(hep.p.CTSSs, target = hep.min)
```

## [3.4.] Create assay with singletons removed
```{r}
a549.p.s.CTSSs <- PRIME::rmSingletons(a549.p.s.CTSSs)
hct.p.s.CTSSs <- PRIME::rmSingletons(hct.p.s.CTSSs)
hep.p.s.CTSSs <- PRIME::rmSingletons(hep.p.s.CTSSs)
```

## [3.4.] Save CTSS objects
```{r}
base::saveRDS(
  object = a549.p.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.A549.CTSSs.rmSingletons.rds"
    )
  )

base::saveRDS(
  object = hct.p.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.HCT116.CTSSs.rmSingletons.rds"
    )
  )

base::saveRDS(
  object = hep.p.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.HepG2.CTSSs.rmSingletons.rds"
    )
  )
```
