---
title: "K562.CTSSs"
output: html_document
date: "2024-01-25"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
##Packages for plotting
base::library(epistack)
base::library(foreach)
base::library(BiocParallel)
base::library(reshape2)
base::library(limma)
base::library(Matrix)
base::library(corrplot)
base::library(assertthat)
```

## [0.1.] Load CAGEfighR extensions
```{r}
# base::setwd("/home/mhf817")
# cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
base::source(base::paste(cfr.extensions.dir, "utils.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "support.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "cumulative.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "decompose.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "enhancers.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "noise.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "coverage.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "complexity.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "subsample.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "dispersion.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "normalize.R", sep = sep))
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
```

## [0.4.] Register server cores / threads
```{r}
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 30
    )
  )
```

## [0.5.] Set working directory
```{r eval=FALSE}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/1.CTSSs/")
```

## [0.6.] Function for pooling replicates
```{r}
poolReplicates <- function(object, replicates, inputAssay="counts") {
    
    ## dimensions
    assertthat::assert_that(
        base::ncol(object) == base::length(replicates),
        base::length(base::unique(replicates)) > 1
      )


    base::message("Calculating pooled counts...")
    a <- base::do.call(
        what = "cbind",
        args = base::lapply(
          X = base::unique(replicates),
          FUN = function(b){
            Matrix::rowSums(
                SummarizedExperiment::assay(object,inputAssay)[,replicates==b,drop=FALSE]
              )
            }
          )
        )
    a <- methods::as(a, "sparseMatrix")
    base::colnames(a) <- base::unique(replicates)

    base::message("Create a pasted design matrix...")
    c <- base::do.call(
      what = "rbind",
      args = base::lapply(
        X = base::unique(replicates),
        FUN = function(b){ 
          base::apply(
              X = SummarizedExperiment::colData(object)[replicates==b,],
              MARGIN = 2 ,
              FUN = base::paste ,
              collapse = ","
            )
          }
        )
      )

    rse <- SummarizedExperiment::SummarizedExperiment(
        assays = S4Vectors::SimpleList(counts = a),
        rowRanges = SummarizedExperiment::rowRanges(object),
        colData = base::as.data.frame(c)
      )
    rse <- CAGEfightR::calcTotalTags(rse)
    return(rse)
  }
```

## [0.7.] Function for writing out TMM normalized bigwig tracks
```{r}
writeBw <- function(object, replicates, dir, inputAssay="TMM"){
  
  ## Check consistency of replicates
  assertthat::assert_that(
      base::length(SummarizedExperiment::rowRanges(object)) == base::nrow(SummarizedExperiment::assay(object,"TMM")),
      base::all(replicates %in% rownames(SummarizedExperiment::colData(object)))
    )
  base::stopifnot("Chosen directory doesn't exist!" = base::dir.exists(dir))
  
  ## Parallel export across all chosen replicates
  foreach::foreach(i=1:base::length(replicates)) %dopar% {
    
    ## Convert individual replicate to GRanges object
    gr <- SummarizedExperiment::rowRanges(object)
    S4Vectors::mcols(gr)["score"] <- SummarizedExperiment::assay(object,inputAssay)[,replicates[i]]
    
    ## Export replicate's plus and minus strands separately
    for (s in base::c("+","-")){
        grs <- gr[BiocGenerics::strand(gr) == s,]
        s <- base::ifelse(s == "+","plus","minus")
        fn <- base::paste0(dir,reps[i],".tmm.",s,".bw")
        
        rtracklayer::export(
            object = grs,
            con = fn,
            format = "bigWig"
          )
        base::message(base::paste(fn,"done.",sep = " "))
      }
    }
  }
```

# [1.] Process K562 datasets -------------------------

## [1.0.] Read in K562 design matrix
```{r}
design_ <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "K562.design.txt"
      ),
    sep = "\t",
    header = T
  )
```

## [1.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::filter(use == "yes") %>%
  tibble::column_to_rownames(var = "ID")
```

## [1.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [1.2.] Quantify CTSSs with no reference genome set
```{r}
k562.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )

k562.CTSSs <- k562.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()
```

## [1.3.] Visualize amount of tags across libraries
```{r}
SummarizedExperiment::colData(k562.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      ID = base::paste(Cell.line,Type,replicate,sep = "_"),
      ID = forcats::as_factor(x = ID)
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = totalTags/1000000,
        y = ID #base::paste(Type,replicate,sep="_")
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_y_discrete(
      limits = rev
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank()
    )
```

## [1.4.] Pool replicates
#### Evaluate with SummarizedExperiment::colData(k562.p.CTSSs), SummarizedExperiment::colData(k562.CTSSs)
```{r}
k562.p.CTSSs <- .GlobalEnv$poolReplicates(
  object = k562.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(k562.CTSSs)$Cell.line,
      SummarizedExperiment::colData(k562.CTSSs)$Type,
      sep="_"
    )
  )

k562.p.CTSSs <- CAGEfightR::calcTPM(k562.p.CTSSs)
```

## [1.5.1.] Subsample original to minimum common depth
```{r}
base::set.seed(1337)

#Only subsample here to second minimum (should probably exclude K562_ProCap_pcr_1 due to low sequencing depth)
k562.s.CTSSs <- .GlobalEnv$subsampleTarget(
    object = k562.CTSSs,
    target = 29267326
  )
k562.s.CTSSs <- CAGEfightR::subsetBySupport(
    object = k562.s.CTSSs,
    minSamples = 0
  )
k562.s.CTSSs <- CAGEfightR::calcTPM(k562.s.CTSSs)
```

## [1.5.2.] Subsample pooled to minimum common depth
```{r}
k562.p.s.CTSSs <- .GlobalEnv$subsampleTarget(
  object = k562.p.CTSSs,
  target = base::min(
      SummarizedExperiment::colData(k562.p.CTSSs)$totalTags
    )
  )
k562.p.s.CTSSs <- CAGEfightR::subsetBySupport(
    object = k562.p.s.CTSSs,
    minSamples = 0
  )
k562.p.s.CTSSs <- CAGEfightR::calcTPM(k562.p.s.CTSSs)
```

## [1.6.] Combine into one object
```{r}
SummarizedExperiment::colData(k562.CTSSs)$pooled <- "no"
SummarizedExperiment::colData(k562.p.CTSSs)$pooled <- "yes"

k562.combined.CTSSs <- BiocGenerics::cbind(k562.CTSSs,k562.p.CTSSs)
```

## [1.7.] Subsample combined object
#### Three sets, lowest CAGE rep and lowest pooled (excluding proCap and groCap)
```{r}
rep.min <- base::min(
  SummarizedExperiment::colData(
    k562.combined.CTSSs)$totalTags[
        SummarizedExperiment::colData(k562.combined.CTSSs)$pooled == "no" &
        !SummarizedExperiment::colData(k562.combined.CTSSs)$method %in% base::c("ProCap","GroCap")
      ]
  ) #29267326
pool.min <- base::min(
  SummarizedExperiment::colData(k562.combined.CTSSs)$totalTags[
      SummarizedExperiment::colData(k562.combined.CTSSs)$pooled=="yes" &
      !SummarizedExperiment::colData(k562.combined.CTSSs)$method %in% base::c("ProCap,ProCap","GroCap")
    ]
  ) #98669216

k562.combined.repSub.CTSSs<- .GlobalEnv$subsampleTarget(
    object = k562.combined.CTSSs,
    target = rep.min
  )
k562.combined.poolSub.CTSSs<- .GlobalEnv$subsampleTarget(
    object = k562.combined.CTSSs,
    target = pool.min
  )

k562.combined.repSub.CTSSs <- CAGEfightR::calcTPM(k562.combined.repSub.CTSSs)
k562.combined.poolSub.CTSSs <- CAGEfightR::calcTPM(k562.combined.poolSub.CTSSs)

```

# [1.8.1.] Save CTSS objects
```{r}
base::saveRDS(
  object = k562.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "subsampled.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.p.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.p.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.subsampled.k562.CTSSs.rds"
    )
  )

base::saveRDS(
  object = k562.combined.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.combined.repSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rds"
    )
  )
base::saveRDS(
  object = k562.combined.poolSub.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rds"
    )
  )
```

## [1.8.2.] Load saved rds files
```{r}
k562.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.k562.CTSSs.rds"
    )
  )
```

## [1.13.] TMM normalise data
```{r}
facs <- edgeR::calcNormFactors(
    object = k562.CTSSs,
    method = "TMM",
    refColumn = NULL
  )
tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
    object = k562.CTSSs,
    sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
    inputAssay = "counts",
    outputAssay = "TMM"
  )
```

## [1.14.] Write out the TMM normalized replicates as bigwig files
```{r}
.GlobalEnv$writeBw(
    object = tmm.CTSSs,
    replicates = base::rownames(SummarizedExperiment::colData(tmm.CTSSs)),
    dir = base::paste0(here::here("1.CTSSs/"),"bw_files_tmm/")
  )
```

## [1.15.1.] Save TMM normalized CTSSs
```{r}
base::saveRDS(
  object = tmm.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_tmm.CTSSs.rds"
    )
  )
```

## [1.15.2.] Load saved rds files
```{r}
fn <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("tmm.CTSSs")

tmm.CTSSs <- base::readRDS(file = fn)
```

#GM12878

```{r}
#Read in design for GM12878
design <- read.table("GM12878.design.txt",sep="\t",header=T)

#Filter and format for use
design <- design[design$use=="yes",]
row.names(design) <- paste(design$ID,design$Seq_run,sep="_")

#Format bw_files
plus_files <- design$bw_plus
minus_files <- design$bw_minus
names(plus_files) <- names(minus_files) <- row.names(design)

bw_plus <- BigWigFileList(plus_files)
bw_minus <- BigWigFileList(minus_files)

```




```{r}
#Quantify CTSSs with no reference genome set
register(MulticoreParam(workers=20))

GM12878.CTSSs <- quantifyCTSSs(plusStrand = bw_plus,
                       minusStrand = bw_minus,
                       design = design)


GM12878.CTSSs <- GM12878.CTSSs %>% calcTPM() %>% calcPooled()

```


```{r}
#Pool replicates
GM12878.p.CTSSs <- poolReplicates(GM12878.CTSSs,
                               replicates=paste(colData(GM12878.CTSSs)$Cell.line,colData(GM12878.CTSSs)$cell_count,colData(GM12878.CTSSs)$Type, sep="_"))

GM12878.p.CTSSs <- calcTPM(GM12878.p.CTSSs)


#Pool only based on reqsequencing
GM12878.pID.CTSSs <- poolReplicates(GM12878.CTSSs,colData(GM12878.CTSSs)$ID)

GM12878.pID.CTSSs <- calcTPM(GM12878.pID.CTSSs)

#Subsample both original and pooled to minimum common depth
set.seed(1337)


#Subsample reads
GM12878.s.CTSSs <- subsampleTarget(GM12878.CTSSs,target=min(colData(GM12878.CTSSs)$totalTags))
GM12878.s.CTSSs <- subsetBySupport(GM12878.s.CTSSs, minSamples=0)
GM12878.s.CTSSs <- calcTPM(GM12878.s.CTSSs)

GM12878.p.s.CTSSs <- subsampleTarget(GM12878.p.CTSSs,target=min(colData(GM12878.p.CTSSs)$totalTags))
GM12878.p.s.CTSSs <- subsetBySupport(GM12878.p.s.CTSSs, minSamples=0)
GM12878.p.s.CTSSs <- calcTPM(GM12878.p.s.CTSSs)


GM12878.pID.s.CTSSs <- subsampleTarget(GM12878.pID.CTSSs,target=min(colData(GM12878.pID.CTSSs)$totalTags))
GM12878.pID.s.CTSSs <- subsetBySupport(GM12878.pID.s.CTSSs, minSamples=0)
GM12878.pID.s.CTSSs <- calcTPM(GM12878.pID.s.CTSSs)


```



```{r}
#Create object that also contains fully pooled samples. Create two subsamples one to the min replicate depth and one to min poole depth and keep original

#Main indicates if part of main analysis. Removing Triptolide and low input
colData(GM12878.pID.CTSSs)$main <- "yes"
colData(GM12878.pID.CTSSs)$main[colData(GM12878.pID.CTSSs)$Type %in% c("Cap_align","ctrl","trip")| colData(GM12878.pID.CTSSs)$cell_count %in% c("50K","250K,250K")] <- "no"
colData(GM12878.pID.CTSSs)$pooled <- "no"

colData(GM12878.p.CTSSs)$main <- "yes"
colData(GM12878.p.CTSSs)$main[colData(GM12878.p.CTSSs)$Type %in% c("Cap_align","ctrl,ctrl,ctrl","trip,trip,trip,trip")| colData(GM12878.p.CTSSs)$cell_count %in% c("50K,50K,50K","250K,250K,250K,250K,250K,250K")] <- "no"
colData(GM12878.p.CTSSs)$pooled <- "yes"


GM12878.combined.CTSSs <- cbind(GM12878.p.CTSSs,GM12878.pID.CTSSs)

rep.min <- min(colData(GM12878.combined.CTSSs)$totalTags[colData(GM12878.combined.CTSSs)$main=="yes" & colData(GM12878.combined.CTSSs)$pooled=="no"]) #29027358
pool.min <- min(colData(GM12878.combined.CTSSs)$totalTags[colData(GM12878.combined.CTSSs)$main=="yes" & colData(GM12878.combined.CTSSs)$pooled=="yes"]) #106618677

GM12878.combined.repSub.CTSSs<- subsampleTarget(GM12878.combined.CTSSs, target = rep.min)
GM12878.combined.poolSub.CTSSs<- subsampleTarget(GM12878.combined.CTSSs, target = pool.min)

GM12878.combined.repSub.CTSSs <- calcTPM(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- calcTPM(GM12878.combined.poolSub.CTSSs)
```



```{r}
#Save CTSS objects
saveRDS(GM12878.CTSSs, "all.GM12878.CTSSs.rds")
saveRDS(GM12878.s.CTSSs, "subsampled.GM12878.CTSSs.rds")
saveRDS(GM12878.p.CTSSs, "pooled.GM12878.CTSSs.rds")
saveRDS(GM12878.p.s.CTSSs, "pooled.subsampled.GM12878.CTSSs.rds")

saveRDS(GM12878.pID.CTSSs, "IDpooled.GM12878.CTSSs.rds")
saveRDS(GM12878.pID.s.CTSSs, "IDpooled.subsampled.GM12878.CTSSs.rds")

saveRDS(GM12878.combined.CTSSs, "all.combined.GM12878.CTSSs.rds")
saveRDS(GM12878.combined.repSub.CTSSs, "repSub.combined.GM12878.CTSSs.rds")
saveRDS(GM12878.combined.poolSub.CTSSs, "poolSub.combined.GM12878.CTSSs.rds")

```




