---
title: "Remove singletons from CTSS objects"
output: html_document
date: "2024-04-24"
---


## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
##Packages for plotting
base::library(epistack)
base::library(foreach)
base::library(BiocParallel)
base::library(reshape2)
base::library(limma)
base::library(Matrix)
base::library(corrplot)
base::library(assertthat)
```

## [0.1.] Load CAGEfighR extensions
```{r}
# base::setwd("/home/mhf817")
# cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
base::source(base::paste(cfr.extensions.dir, "utils.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "support.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "cumulative.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "decompose.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "enhancers.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "noise.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "coverage.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "complexity.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "subsample.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "dispersion.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "normalize.R", sep = sep))
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
```

## [0.4.] Register server cores / threads
```{r}
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 10
    )
  )
```

## [0.5.] Set working directory
```{r eval=FALSE}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/1.CTSSs/")
```

## [0.6.] Function for pooling replicates
```{r}
poolReplicates <- function(object, replicates, inputAssay="counts") {
    
    ## dimensions
    assertthat::assert_that(
        base::ncol(object) == base::length(replicates),
        base::length(base::unique(replicates)) > 1
      )

    base::message("Calculating pooled counts...")
    a <- base::do.call(
        what = "cbind",
        args = base::lapply(
          X = base::unique(replicates),
          FUN = function(b){
            Matrix::rowSums(
                SummarizedExperiment::assay(object,inputAssay)[,replicates==b,drop=FALSE]
              )
            }
          )
        )
    a <- methods::as(a, "sparseMatrix")
    base::colnames(a) <- base::unique(replicates)

    base::message("Create a pasted design matrix...")
    c <- base::do.call(
      what = "rbind",
      args = base::lapply(
        X = base::unique(replicates),
        FUN = function(b){ 
          base::apply(
              X = SummarizedExperiment::colData(object)[replicates==b,],
              MARGIN = 2 ,
              FUN = base::paste ,
              collapse = ","
            )
          }
        )
      )

    rse <- SummarizedExperiment::SummarizedExperiment(
        assays = S4Vectors::SimpleList(counts = a),
        rowRanges = SummarizedExperiment::rowRanges(object),
        colData = base::as.data.frame(c)
      )
    rse <- CAGEfightR::calcTotalTags(rse)
    return(rse)
  }
```

## [0.7.] Alternative subsampling to deal with dropped rows
```{r}
subsampleTarget <- function(object, inputAssay = "counts", target) {

    assertthat::assert_that(
        methods::is(object, "SummarizedExperiment"),
        inputAssay %in% SummarizedExperiment::assayNames(object),
        base::is.numeric(target),
        target > 0
      )
    
    a <- SummarizedExperiment::assay(object,inputAssay)
    n <- base::ncol(a)
    nz <- base::lapply(
      X = 1:n,
      FUN = function(i){
          .GlobalEnv$nonzero(a[,i,drop=FALSE])[,1]
        }
      )
    s <- Matrix::colSums(a)
    d <- base::unlist(
      base::lapply(
        X = 1:n,
        FUN = function(i){
        if (s[i]<=target){
              a[nz[[i]],i]
            } else {
              stats::rbinom(base::length(nz[[i]]),a[nz[[i]],i], target/s[i])
            }
          }
        )
      )
    keep <- base::which(
        base::sapply(nz,length)>0
      )
    
    # Create a matrix to store the subsampled counts, ensuring all rows are retained
    new_matrix <- Matrix::sparseMatrix(
      i = base::unlist(nz),
      j = base::unlist(
        base::lapply(
          X = keep,
          FUN = function(i){
              base::rep(i, base::length(nz[[i]]))
            }
          )
        ),
      x = d, 
      dims = base::dim(a),  # Preserve the original dimensions
      dimnames = base::list(
          base::rownames(a),
          base::colnames(a)[keep]
        )
      )
    
    # Replace the original assay with the new matrix
    SummarizedExperiment::assay(object, inputAssay) <- new_matrix
    
    # Recalculate the total count of reads for each sample
    object <- CAGEfightR::calcTotalTags(object)
    
    return(object)
  }
```


# [1.] Process K562 datasets -------------------------

## [1.0.] Read in K562 design matrix
```{r}
design_ <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "K562.design.txt"
      ),
    sep = "\t",
    header = T
  )
```

## [1.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::filter(use == "yes") %>%
  tibble::column_to_rownames(var = "ID")
```

## [1.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [1.2.] Quantify CTSSs with no reference genome set
```{r}
k562.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )

k562.CTSSs <- k562.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()
```

## [1.3.] Visualize amount of tags across libraries
```{r}
SummarizedExperiment::colData(k562.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      ID = base::paste(Cell.line,Type,replicate,sep = "_"),
      ID = forcats::as_factor(x = ID)
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = totalTags/1000000,
        y = ID #base::paste(Type,replicate,sep="_")
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_y_discrete(
      limits = rev
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank()
    )
```

## [1.4.] Remove non-standard chromosomes
#### Important! chrY and chrM are also excluded
```{r}
keep <- GenomeInfoDb::seqlevels(k562.CTSSs)[1:23]
k562.CTSSs <- GenomeInfoDb::keepSeqlevels(k562.CTSSs, keep, pruning.mode = "coarse")
```

## [1.5.] Pool replicates
#### Evaluate with SummarizedExperiment::colData(k562.p.CTSSs), SummarizedExperiment::colData(k562.CTSSs)
```{r}
k562.p.CTSSs <- .GlobalEnv$poolReplicates(
  object = k562.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(k562.CTSSs)$Cell.line,
      SummarizedExperiment::colData(k562.CTSSs)$Type,
      sep="_"
    )
  )

k562.p.CTSSs <- CAGEfightR::calcTPM(k562.p.CTSSs)
```

## [1.6.] Combine into one object
```{r}
SummarizedExperiment::colData(k562.CTSSs)$pooled <- "no"
SummarizedExperiment::colData(k562.p.CTSSs)$pooled <- "yes"

k562.combined.CTSSs <- BiocGenerics::cbind(k562.CTSSs,k562.p.CTSSs)
```

```{r}
#Find minimum seq depths
k562.combined.CTSSs <- calcTotalTags(k562.combined.CTSSs)
```

```{r}
#Subsample both original and pooled to minimum common depth
set.seed(1337)

#subsample combined. Three sets, lowest CAGE rep and lowest pooled (excluding proCap and groCap)
rep.min <- min(colData(k562.combined.CTSSs)$totalTags[colData(k562.combined.CTSSs)$pooled=="no"& !colData(k562.combined.CTSSs)$method %in% c("ProCap","GroCap")]) #28619974
pool.min <- min(colData(k562.combined.CTSSs)$totalTags[colData(k562.combined.CTSSs)$pooled=="yes" & !colData(k562.combined.CTSSs)$method %in% c("ProCap,ProCap","GroCap")]) #96351121

k562.combined.repSub.CTSSs<- subsampleTarget(k562.combined.CTSSs, target = rep.min)
k562.combined.poolSub.CTSSs<- subsampleTarget(k562.combined.CTSSs, target = pool.min)


k562.combined.repSub.CTSSs <- calcTPM(k562.combined.repSub.CTSSs)
k562.combined.poolSub.CTSSs <- calcTPM(k562.combined.poolSub.CTSSs)


#Create new assay with counts where 1 counts become 0
rmSingletons <- function(rse, inputAssay = "counts") {
  rmS <- assay(rse, inputAssay)
  # Find the indices where values are equal to 1
  idx <- which(rmS == 1, arr.ind = TRUE)
  # Replace the values at those indices with 0
  rmS[idx] <- 0
  # Assign the modified assay back to the SummarizedExperiment object
  assay(rse, "counts.noSingletons") <- rmS
  rse
}

k562.combined.repSub.CTSSs <- rmSingletons(k562.combined.repSub.CTSSs)
k562.combined.poolSub.CTSSs <- rmSingletons(k562.combined.poolSub.CTSSs)
```






```{r}
#Save as new objects

saveRDS(k562.combined.CTSSs, "all.combined.k562.CTSSs.rds")
saveRDS(k562.combined.repSub.CTSSs, "repSub.combined.k562.CTSSs.rmSingletons.rds")
saveRDS(k562.combined.poolSub.CTSSs, "poolSub.combined.k562.CTSSs.rmSingletons.rds")


```



#########
#GM12878
#########


```{r}
#Read in design for GM12878
design <- read.table("GM12878.design.txt",sep="\t",header=T)

#Filter and format for use
design <- design[design$use=="yes",]
row.names(design) <- paste(design$ID,design$Seq_run,sep="_")

#Format bw_files
plus_files <- design$bw_plus
minus_files <- design$bw_minus
names(plus_files) <- names(minus_files) <- row.names(design)

bw_plus <- BigWigFileList(plus_files)
bw_minus <- BigWigFileList(minus_files)

```




```{r}
#Quantify CTSSs with no reference genome set
register(MulticoreParam(workers=20))

GM12878.CTSSs <- quantifyCTSSs(plusStrand = bw_plus,
                       minusStrand = bw_minus,
                       design = design)


GM12878.CTSSs <- GM12878.CTSSs %>% calcTPM() %>% calcPooled()

GM12878.CTSSs <- keepSeqlevels(GM12878.CTSSs, keep, pruning.mode = "coarse")


```





```{r}
#Pool only based on reqsequencing
GM12878.pID.CTSSs <- poolReplicates(GM12878.CTSSs,colData(GM12878.CTSSs)$ID)

GM12878.pID.CTSSs <- calcTPM(GM12878.pID.CTSSs)
#Pool replicates
GM12878.p.CTSSs <- poolReplicates(GM12878.CTSSs,
                               replicates=paste(colData(GM12878.CTSSs)$Cell.line,colData(GM12878.CTSSs)$cell_count,colData(GM12878.CTSSs)$Type, sep="_"))

GM12878.p.CTSSs <- calcTPM(GM12878.p.CTSSs)


```



```{r}
#Main indicates if part of main analysis. Removing Triptolide and low input
colData(GM12878.pID.CTSSs)$main <- "yes"
colData(GM12878.pID.CTSSs)$main[colData(GM12878.pID.CTSSs)$Type %in% c("Cap_align","ctrl","trip")| colData(GM12878.pID.CTSSs)$cell_count %in% c("50K","250K,250K")] <- "no"
colData(GM12878.pID.CTSSs)$pooled <- "no"
#Subsample
colData(GM12878.p.CTSSs)$main <- "yes"
colData(GM12878.p.CTSSs)$main[colData(GM12878.p.CTSSs)$Type %in% c("Cap_align","ctrl,ctrl,ctrl","trip,trip,trip,trip")| colData(GM12878.p.CTSSs)$cell_count %in% c("50K,50K,50K","250K,250K,250K,250K,250K,250K")] <- "no"
colData(GM12878.p.CTSSs)$pooled <- "yes"


GM12878.combined.CTSSs <- cbind(GM12878.p.CTSSs,GM12878.pID.CTSSs)
GM12878.combined.CTSSs <- calcTotalTags(GM12878.combined.CTSSs)

rep.min <- min(colData(GM12878.combined.CTSSs)$totalTags[colData(GM12878.combined.CTSSs)$main=="yes" & colData(GM12878.combined.CTSSs)$pooled=="no"]) #28453942
pool.min <- min(colData(GM12878.combined.CTSSs)$totalTags[colData(GM12878.combined.CTSSs)$main=="yes" & colData(GM12878.combined.CTSSs)$pooled=="yes"]) #104774632

GM12878.combined.repSub.CTSSs<- subsampleTarget(GM12878.combined.CTSSs, target = rep.min)
GM12878.combined.poolSub.CTSSs<- subsampleTarget(GM12878.combined.CTSSs, target = pool.min)

GM12878.combined.repSub.CTSSs <- calcTPM(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- calcTPM(GM12878.combined.poolSub.CTSSs)
```



```{r}
#remove singletons

GM12878.combined.repSub.CTSSs <- rmSingletons(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- rmSingletons(GM12878.combined.poolSub.CTSSs)
```

```{r}
#Save as new objects
saveRDS(GM12878.combined.CTSSs, "all.combined.GM12878.CTSSs.rds")
saveRDS(GM12878.combined.repSub.CTSSs, "repSub.combined.GM12878.CTSSs.rmSingletons.rds")
saveRDS(GM12878.combined.poolSub.CTSSs, "poolSub.combined.GM12878.CTSSs.rmSingletons.rds")


```


