---
title: "CTSS level complexity"
output: html_document
date: "2024-01-26"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
##Packages for plotting
base::library(epistack)
base::library(BiocParallel)
base::library(reshape2)
base::library(limma)
base::library(Matrix)
base::library(corrplot)
```

## [0.1.] Load CAGEfighR extensions
```{r}
# base::setwd("/home/mhf817")
# cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
base::source(base::paste(cfr.extensions.dir, "utils.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "support.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "cumulative.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "decompose.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "enhancers.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "noise.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "coverage.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "complexity.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "subsample.R", sep = sep))
base::source(base::paste(cfr.extensions.dir, "dispersion.R", sep = sep))
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
                     
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )

gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )

gm.full.ada <- base::c(
    "#F5CE00","#F5CE00","#F5CE00","#FF8D00",
    "#FF4425","#FF4425","#FF4425","#DD4425",
    "#92B9C9","#92B9C9","#92B9C9","#00439B",
    "#29A957","#29A957","#29A957","#007A30",
    "#878784","#878784","#878784","#3D4138",
    "#AC97C2","#AC97C2","#AC97C2","#7D62D0"
  )
```

## [0.4.] Set working directory
```{r eval=FALSE}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/1.CTSSs/")
```

# [1.] K562 datasets -------------------------

## [0.5.] Read in K562 datasets
#### see '1.1CTSSs.processing.Rmd' for read objects
```{r}
k562.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "all.k562.CTSSs.rds"
    )
  )
k562.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "subsampled.k562.CTSSs.rds"
    )
  )
k562.p.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.k562.CTSSs.rds"
    )
  )
k562.p.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.subsampled.k562.CTSSs.rds"
    )
  )
```

## [1.0.] Compare all CTSS level correlation
```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = k562.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename =base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_k562.CTSSs.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 30,
      height = 30
    )
  }
```

```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = k562.s.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename =base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_subsampled.k562.CTSSs.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 30,
      height = 30
    )
  }
```

```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = k562.p.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_pooled.k562.CTSSs.corr",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = k562.p.s.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_pooled.subsampled.k562.CTSSs.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

## [1.1.] Compare CTSS>0TPM correlations
```{r}
m1 <- as.matrix(assay(k562.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/k562.CTSSs.rm0.corr.pdf",width = 30, height = 30)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

m1 <- as.matrix(assay(k562.s.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/subsampled.k562.CTSSs.rm0.corr.pdf",width = 30, height = 30)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

m1 <- as.matrix(assay(k562.p.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/pooled.k562.CTSSs.rm0.corr.pdf",width = 15, height = 15)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

m1 <- as.matrix(assay(k562.p.s.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/pooled.subsampled.k562.CTSSs.rm0.corr.pdf",width = 15, height = 15)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()
```

## [1.2.] Calculate CTSS level complexity
```{r}
base::set.seed(1337)

comp.k562.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = k562.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.k562.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = k562.s.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.k562.p.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = k562.p.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.k562.p.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = k562.p.s.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
```

## [1.3.] Save & load RData container
```{r}
base::save(
  comp.k562.CTSSs,
  comp.k562.s.CTSSs,
  comp.k562.p.CTSSs,
  comp.k562.p.s.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_K562.CTSS.complexity.RData"
    )
  )
```

```{r}
f <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>% 
  stringr::str_subset("complexity") %>%
  stringr::str_subset("K562")

base::load(file = f)
```

## [1.3.1.] #### Safety copy of 'comp.k562.combined.CTSSs' object
```{r}
base::saveRDS(
  object = comp.k562.combined.CTSSs,
  file =  base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_comp.k562.combined.CTSSs.rds"
    )
  )
```

```{r}
comp.k562.combined.CTSSs <- base::readRDS(
  file = base::paste0(
    here::here("1.CTSSs/"),
      "20240412_comp.k562.combined.CTSSs.rds"
    )
  )
```

## [1.4.] Plot complexities
```{r}
comp.k562.CTSSs$method <- stringr::str_sub(
    string = comp.k562.CTSSs$sample, start = 1, end = -2
  )
ggplot2::ggplot(
    data = comp.k562.CTSSs,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = method,
        group = sample
      )
    ) + 
  geom_line() +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  theme_bw() +
  scale_color_manual(
      values = cols3
    ) +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.5.] Save complexity plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("results/20240324_combined_analysis/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_comp.comb.CTSSs_remap",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.6.] Plot complexities for the other objects
```{r}
comp.k562.CTSSs$method <- str_sub(comp.k562.CTSSs$sample,1,-2)
pdf("./plots/k562.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.k562.CTSSs,aes(totalTags,numberCTSSs,color=method,group=sample)) + 
  geom_line() +
  theme_bw() + 
  scale_color_manual(values=custom.col)
dev.off()

comp.k562.CTSSs$method <- str_sub(comp.k562.CTSSs$sample,1,-2)
comp.k562.s.CTSSs$method <- str_sub(comp.k562.s.CTSSs$sample,1,-2)
pdf("./plots/subsampled.k562.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.k562.s.CTSSs,aes(totalTags,numberCTSSs,color=method,group=sample)) + 
  geom_line() +
  theme_bw() + 
  scale_color_manual(values=custom.col)
dev.off()

comp.k562.p.CTSSs$method <- str_sub(comp.k562.p.CTSSs$sample,1,-2)
pdf("./plots/pooled.k562.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.k562.p.CTSSs,aes(totalTags,numberCTSSs,color=method,group=sample)) + 
  geom_line() +
  theme_bw() + 
  scale_color_manual(values=custom.col)
dev.off()

comp.k562.p.s.CTSSs$method <- str_sub(comp.k562.p.s.CTSSs$sample,1,-2)
pdf("./plots/pooled.subsampled.k562.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.k562.p.s.CTSSs,aes(totalTags,numberCTSSs,color=method,group=sample)) + 
  geom_line() +
  theme_bw() + 
  scale_color_manual(values=custom.col)
dev.off()
```

## [1.7.] Data wrangling for all K562 samples & pools
#### Name change of "K562_10M_N", remove "K562_Cap_align" sample
```{r}
comp.k562.combined.CTSSs$sample[comp.k562.combined.CTSSs$sample=="K562_N_10M"] <- "K562_10M_N"

comp.k562.combined.CTSSs <- comp.k562.combined.CTSSs[comp.k562.combined.CTSSs$sample!="K562_Cap_align",]

comp.k562.combined.CTSSs$sample <- base::factor(
    comp.k562.combined.CTSSs$sample,
    levels = base::c(
        "K562_C1","K562_C2","K562_C3","K562_C",
        "K562_N4","K562_N5","K562_N6","K562_N",
        "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
        "K562_ProCap_pcr_1","K562_ProCap_pcr_2","K562_pcr_align",
        "K562_groCap_1"
      ),
      ordered = TRUE
  )

base::unique(comp.k562.combined.CTSSs$sample)
```

## [1.7.1.] Identify max values for coordinated plotting
```{r}
tt1 <- comp.k562.combined.CTSSs %>%
  dplyr::pull(totalTags) %>%
  base::max(na.rm = TRUE)
tt2 <- comp.GM12878.combined.CTSSs %>%
  dplyr::pull(totalTags) %>%
  base::max(na.rm = TRUE)
nc1 <- comp.k562.combined.CTSSs %>%
  dplyr::pull(numberCTSSs) %>%
  base::max(na.rm = TRUE)
nc2 <- comp.GM12878.combined.CTSSs %>%
  dplyr::pull(numberCTSSs) %>%
  base::max(na.rm = TRUE)
xmax <- base::max(tt1,tt2)
ymax <- base::max(nc1,nc2)
```

## [1.8.] Plot all complexity curves
#### /combined_plots/combined.k562.CTSSs.comp.pdf
```{r}
ggplot2::ggplot(
    data = comp.k562.combined.CTSSs,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color=sample
      )
    ) + 
  geom_line() +
  geom_vline(
      xintercept = base::c(29267326,98669216),
      linetype = 2
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
    ) +
  scale_color_manual(
      values = k.full.pal
    ) +
  theme_bw()  +
  theme(
      aspect.ratio = 1,
    )
```

#### /combined_plots/combined.k562.CTSSs.comp.scaled.pdf
```{r}
ggplot2::ggplot(
    data = comp.k562.combined.CTSSs,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  theme_bw() +
  scale_y_log10() +
  # scale_x_log10() +
  scale_color_manual(
      values = k.full.pal
    ) +
  geom_vline(
      xintercept = base::c(29267326,98669216),
      linetype = 2
    )
```

## [1.9.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexityCtss",
          "_K562",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 6
    )
  }
```

## [1.10.] Plot subsampled number of CTSSs per library
## Add barplots for subsampled levels
```{r}
p.max <- comp.k562.combined.CTSSs %>%
  dplyr::filter(target < 98669216) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
r.max <- comp.k562.combined.CTSSs %>%
  dplyr::filter(target < 29267326) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
```

```{r}
ggplot2::ggplot(
    data = r.max, # p.max,
    mapping = aes(
        x = compexity,
        y = forcats::fct_rev(sample),
        fill = sample
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = sample
      ),
      x = 0,
      hjust = -0.1,
      colour = "white",
      size = 3,
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  # scale_x_log10() +
  guides(fill = "none") +
  labs(
      x = "numberCTSSs",
      y = "Samples"
    ) +
  theme_bw() +
  theme(
    aspect.ratio = 2,
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

## [1.11.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexityCtss",
          "_K562",
          "_barplots",
          "_replicate",
          # "_pool",
          "_nonLog10",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 3,
      height = 6
    )
  }
```


# [2.] GM12878 -------------------------

## [0.6.] Read in GM12878 datasets
#### see '1.1CTSSs.processing.Rmd' for read objects
```{r}
GM12878.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "all.GM12878.CTSSs.rds"
    )
  )
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "subsampled.GM12878.CTSSs.rds"
    )
  )
GM12878.p.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.GM12878.CTSSs.rds"
    )
  )
GM12878.p.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.subsampled.GM12878.CTSSs.rds"
    )
  )
GM12878.pID.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "IDpooled.GM12878.CTSSs.rds"
    )
  )
GM12878.pID.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "IDpooled.subsampled.GM12878.CTSSs.rds"
    )
  )
GM12878.combined.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )
```

## [2.0.] Compare all CTSS level correlation
```{r}
pdf("./plots/GM12878.CTSSs.corr.pdf",width = 30, height = 30)
corrplot(cor(
  as.matrix(assay(GM12878.CTSSs,"TPM"))),
  method="number")
dev.off()

pdf("./plots/subsampled.GM12878.CTSSs.corr.pdf",width = 30, height = 30)
corrplot(cor(
  as.matrix(assay(GM12878.s.CTSSs,"TPM"))),
  method="number")
dev.off()

pdf("./plots/pooled.GM12878.CTSSs.corr.pdf",width = 15, height = 15)
corrplot(cor(
  as.matrix(assay(GM12878.p.CTSSs,"TPM"))),
  method="number")
dev.off()

pdf("./plots/pooled.subsampled.GM12878.CTSSs.corr.pdf",width = 15, height = 15)
corrplot(cor(
  as.matrix(assay(GM12878.p.s.CTSSs,"TPM"))),
  method="number")
dev.off()

pdf("./plots/IDpooled.GM12878.CTSSs.corr.pdf",width = 30, height = 30)
corrplot(cor(
  as.matrix(assay(GM12878.pID.CTSSs,"TPM"))),
  method="number")
dev.off()

pdf("./plots/IDpooled.subsampled.GM12878.CTSSs.corr.pdf",width = 30, height = 30)
corrplot(cor(
  as.matrix(assay(GM12878.pID.s.CTSSs,"TPM"))),
  method="number")
dev.off()

```

## [2.1.] Compare CTSS>0TPM correlations
```{r}
#m1 <- as.matrix(assay(GM12878.CTSSs,"TPM"))
#m1[m1==0] <- NA

#pdf("./plots/GM12878.CTSSs.rm0.corr.pdf",width = 30, height = 30)
#corrplot(cor(m1,use="complete.obs"),
#  method="number")
#dev.off()

m1 <- as.matrix(assay(GM12878.s.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/subsampled.GM12878.CTSSs.rm0.corr.pdf",width = 30, height = 30)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

m1 <- as.matrix(assay(GM12878.p.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/pooled.GM12878.CTSSs.rm0.corr.pdf",width = 15, height = 15)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

m1 <- as.matrix(assay(GM12878.p.s.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/pooled.subsampled.GM12878.CTSSs.rm0.corr.pdf",width = 15, height = 15)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

#m1 <- as.matrix(assay(GM12878.pID.CTSSs,"TPM"))
#m1[m1==0] <- NA

#pdf("./plots/IDpooled.GM12878.CTSSs.rm0.corr.pdf",width = 30, height = 30)
#corrplot(cor(m1,use="complete.obs"),
#  method="number")
#dev.off()

m1 <- as.matrix(assay(GM12878.pID.s.CTSSs,"TPM"))
m1[m1==0] <- NA

pdf("./plots/IDpooled.subsampled.GM12878.CTSSs.rm0.corr.pdf",width = 30, height = 30)
corrplot(cor(m1,use="complete.obs"),
  method="number")
dev.off()

```

## [2.2.] Calculate CTSS level complexity
```{r}
base::set.seed(1337)

comp.GM12878.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.s.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.p.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.p.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.p.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.p.s.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.pID.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.pID.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.pID.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.pID.s.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
comp.GM12878.combined.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.combined.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
```

## [2.3.] Save & load RData container
```{r}
#save for potential later use
base::save(
    comp.GM12878.CTSSs,
    comp.GM12878.s.CTSSs,
    comp.GM12878.p.CTSSs,
    comp.GM12878.pID.CTSSs,
    comp.GM12878.pID.s.CTSSs,
    comp.GM12878.combined.CTSSs,
    file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.CTSS.complexity.RData"
    )
  )
```

```{r}
f <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>% 
  stringr::str_subset("complexity") %>%
  stringr::str_subset("GM12878")

base::load(file = f)
```

## [1.3.1.] #### Safety copy of 'comp.GM12878.combined.CTSSs' object
```{r}
base::saveRDS(
  object = comp.GM12878.combined.CTSSs,
  file =  base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_comp.GM12878.combined.CTSSs.rds"
    )
  )
```

```{r}
comp.GM12878.combined.CTSSs <- base::readRDS(
  file = base::paste0(
    here::here("1.CTSSs/"),
      "20240415_comp.GM12878.combined.CTSSs.rds"
    )
  )
```

## [1.4.] Plot complexities
```{r}
ggplot2::ggplot(
    data = comp.GM12878.CTSSs,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = method,
        group = sample
      )
    ) + 
  geom_line() +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  theme_bw() +
  scale_color_manual(
      values = cols3
    ) +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

#### "plots/GM12878.CTSSs.comp.pdf"
```{r}
ggplot2::ggplot(
  data = comp.GM12878.CTSSs,
  mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  theme_bw()
```

#### "plots/GM12878.CTSSs.compSplit.pdf"
```{r}
comp.GM12878.CTSSs$method <- stringr::str_extract(comp.GM12878.CTSSs$sample, "_(.*?)_")

comp.GM12878.CTSSs$method <- gsub("_", "", comp.GM12878.CTSSs$method )

ggplot2::ggplot(
    data = comp.GM12878.CTSSs,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  theme_bw() + 
  facet_wrap(
      facets = ~method
    ) +
  theme(
      legend.position = "none",
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [2.5.] Plot complexities for the other objects
```{r}
pdf("./plots/subsampled.GM12878.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.GM12878.s.CTSSs,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() 
dev.off()

pdf("./plots/pooled.GM12878.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.GM12878.p.CTSSs,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() 
dev.off()

pdf("./plots/pooled.subsampled.GM12878.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.GM12878.p.s.CTSSs,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() 
dev.off()

pdf("./plots/pooledID.GM12878.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.GM12878.pID.CTSSs,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() 
dev.off()

pdf("./plots/pooledID.subsampled.GM12878.CTSSs.comp.pdf",width = 15, height = 10)
ggplot(comp.GM12878.pID.s.CTSSs,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() 
dev.off()


```

## [2.6.] Prepare plotting 
#### 'comp.GM12878.combined.CTSSs' object is loaded in [2.3.]
#### replaced: 'main <- comp.GM12878.combined.CTSSs[comp.GM12878.combined.CTSSs$sample %in% base::colnames(GM12878.combined.CTSSs)[GM12878.combined.CTSSs$main=="yes"],]'
```{r}
levs <- base::c(
    "GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3","GM12878_500K_C",
    "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3", "GM12878_500K_N",
    "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3","GM12878_1M_C",
    "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3", "GM12878_1M_N",
    "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3", "GM12878_5M_N",
    "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3", "GM12878_10M_N"
  )

types <- GM12878.combined.CTSSs %>% 
  SummarizedExperiment::colData() %>%
  base::as.data.frame() %>% 
  tibble::rownames_to_column(var = "type") %>%
  dplyr::filter(main == "yes") %>%
  dplyr::pull(type)
main <- comp.GM12878.combined.CTSSs %>%
  dplyr::filter(sample %in% types)

main$sample <- base::factor(
    main$sample,
    levels = levs,
    ordered = TRUE
  )
```

## [2.7.] Plot complexities including pooled samples
#### "/combined_plots/combined.GM12878.CTSSs.comp.pdf"
#### "/combined_plots/combined.GM12878.CTSSs.comp.scaled.pdf"
```{r}
ggplot2::ggplot(
    data = main,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  geom_vline(
      xintercept = base::c(29027358,106618677),
      linetype = 2
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
    ) +
  scale_color_manual(
      values = gm.full.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```
#### log10 scaled
```{r}
ggplot2::ggplot(
    data = main,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  theme_bw() +
  scale_y_log10() +
  scale_color_manual(
      values = gm.full.pal
    ) +
  geom_vline(
      xintercept = base::c(29027358,106618677),
      linetype = 2
    )
```

#### Only pooled samples
#### previously: Only 500k and 5M samples, 'gm.full.ada' color palette used
```{r}
main %>%
  dplyr::filter(
      stringr::str_detect(sample,"_N$|_C$")
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  geom_vline(
      xintercept = base::c(29027358,106618677),
      linetype = 2
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
    ) +
  scale_color_manual(
      values = gm.full.pal[base::seq(1,24,4)]
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.8.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexityCtss",
          "_GM12878",
          "_Pooled",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 6, #10
      height = 4
    )
  }
```

## [2.9.] Plot subsampled number of CTSSs per library
## Add barplots for subsampled levels
```{r}
p.max <- main %>%
  dplyr::filter(target < 106618677L) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
r.max <- main %>%
  dplyr::filter(target < 29027358) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
```

```{r}
ggplot2::ggplot(
    data = p.max, # r.max,
    mapping = aes(
        x = compexity,
        y = forcats::fct_rev(sample),
        fill = sample
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = sample
      ),
      x = 0,
      hjust = 0,
      colour = "white",
      size = 3,
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  # scale_x_log10() +
  guides(fill = "none") +
  labs(
      x = "numberCTSSs",
      y = "Samples"
    ) +
  theme_bw() +
  theme(
    aspect.ratio = 2,
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

## [2.10.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexityCtss",
          "_GM12878",
          "_barplots",
          # "_replicate",
          "_pool",
          "_nonLog10",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 3,
      height = 6
    )
  }
```


```{r}
p1 <- ggplot(main,aes(totalTags,numberCTSSs,color=sample)) + 
  geom_line() +
  theme_bw() + scale_y_log10() + scale_color_manual(values = gm.full.pal) + geom_vline(xintercept=c(29027358,106618677),linetype=2)+ guides(color="none") 

p2 <- ggplot(r.max,aes(fct_rev(sample),rep.compexity,fill=sample)) + geom_col() + theme_bw() + scale_fill_manual(values=gm.full.pal) + coord_flip() + scale_y_log10()+ guides(fill="none") + labs(y="numberCTSSs", x="Samples")

p3 <- ggplot(p.max,aes(fct_rev(sample),pooled.compexity,fill=sample)) + geom_col() + theme_bw() + scale_fill_manual(values=gm.full.pal) + coord_flip() + scale_y_log10() + labs(y="numberCTSSs", x="Samples")
```


```{r}
# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  p3 + theme(legend.box.margin = margin(0, 0, 0, 12))
)

#Save legend plot as separate object to work with
pdf("./combined_plots/GM12878.colorLegend.pdf")
grid.arrange(legend)
dev.off()

pdf("./combined_plots/GM12878.mainComplexity.pdf",width=30,height=15)
grid.arrange(p1,p2,p3+guides(fill="none"),
  widths = c(2, 1, 1),
  layout_matrix = rbind(c(1, 2, 3))
)
dev.off()
```

