---
title: "CTSS level complexity"
output: html_document
date: "2024-01-26"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(BiocParallel)
```

## [0.1.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
```

# [1.] K562 datasets -------------------------

## [1.0.] Read in K562 datasets
#### see '1.1CTSSs.processing.Rmd' for CTSSs objects
```{r}
K562.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.k562.CTSSs.rds"
    )
  )
# K562.s.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "repSub.combined.k562.CTSSs.rmSingletons.rds"
#     )
#   )
# K562.p.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.k562.CTSSs.rmSingletons.rds"
#     )
#   )
```

## [1.1.] Compare all CTSS level correlation
```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = K562.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562.CTSSs.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 8
    )
  }
```

```{r}
# corrplot::corrplot(
#     stats::cor(
#       x = base::as.matrix(
#         x = SummarizedExperiment::assay(
#             x = K562.s.CTSSs,
#             i = "TPM"
#           )
#         )
#       ),
#     method = "circle", # "number"
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename =base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_subsampled.K562.CTSSs.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 30,
#       height = 30
#     )
#   }
```

```{r}
# corrplot::corrplot(
#     stats::cor(
#       x = base::as.matrix(
#         x = SummarizedExperiment::assay(
#             x = K562.p.CTSSs,
#             i = "TPM"
#           )
#         )
#       ),
#     method = "circle", # "number"
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_pooled.K562.CTSSs.corr",
#           ".", ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 15,
#       height = 15
#     )
#   }
```

## [1.2.] Compare CTSS>0TPM correlations
```{r}
m1 <- base::as.matrix(
    x = SummarizedExperiment::assay(x = K562.CTSSs, i = "TPM")
  )
m1[m1 == 0] <- NA
corrplot::corrplot(
  corr = stats::cor(m1, use = "complete.obs"),
    method = "circle",
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562.CTSSs.rm0.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 8
    )
  }
```

```{r}
# m1 <- base::as.matrix(
#     x = SummarizedExperiment::assay(x = K562.s.CTSSs, i = "TPM")
#   )
# m1[m1 == 0] <- NA
# corrplot::corrplot(
#   corr = stats::cor(m1, use = "complete.obs"),
#     method = "circle",
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_subsampled.K562.CTSSs.rm0.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

```{r}
# m1 <- base::as.matrix(
#     x = SummarizedExperiment::assay(x = K562.p.CTSSs, i = "TPM")
#   )
# m1[m1 == 0] <- NA
# corrplot::corrplot(
#   corr = stats::cor(m1, use = "complete.obs"),
#     method = "circle",
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_pooled.K562.CTSSs.rm0.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

## [1.3.] Calculate CTSS level complexity
```{r}
base::set.seed(1337)

comp.K562.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = K562.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
# comp.K562.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
#     object = K562.s.CTSSs,
#     step = 2e+6,
#     minCTSSsupport = 1,
#     CTSSunexpressed = 0
#   )
# comp.K562.p.CTSSs <- .GlobalEnv$calcCTSSComplexity(
#     object = K562.p.CTSSs,
#     step = 2e+6,
#     minCTSSsupport = 1,
#     CTSSunexpressed = 0
#   )
```

## [1.4.] Save & load objects
```{r}
base::saveRDS(
  object = comp.K562.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_comp.K562.CTSSs",
      ".rds"
    )
  )
```

```{r}
comp.K562.CTSSs <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("comp.K562.CTSSs.rds")
  )
```

## [1.5.] Data wrangling for all K562 samples & pools
#### Name change of "K562_10M_N", remove "K562_Cap_align" sample
```{r}
 levs <- base::c(
    "K562_C", "K562_C1", "K562_C2", "K562_C3",
    "K562_N", "K562_N4", "K562_N5", "K562_N6",
    "K562_10M_N", "K562_10M_N1", "K562_10M_N2", "K562_10M_N3"
  )
alts <- base::c(
    "K562_ProCap_pcr_1", "K562_ProCap_pcr_2", "K562_pcr_align",
    "K562_groCap_1", "K562_Cap_align"
  )

comp.K562.CTSSs.plot <- comp.K562.CTSSs %>%
  dplyr::mutate(
    sample = dplyr::case_when(
        sample == "K562_N_10M" ~ "K562_10M_N",
        TRUE ~ sample
      )
    ) %>% 
  dplyr::filter(sample %in% levs) %>% 
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = levs,
        ordered = TRUE
      ),
    method = stringr::str_replace(
        sample,"^(.*?)(_)?(\\d)?$","\\1"
      )
    )
```

## [1.6.] Plot complexities
```{r}
ggplot2::ggplot(
    data = comp.K562.CTSSs.plot,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = method,
        group = sample
      )
    ) + 
  geom_line() +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  theme_bw() +
  scale_color_manual(
      values = cols3
    ) +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.7.] Plot complexities for the other objects
```{r}
# comp.K562.s.CTSSs$method <- stringr::str_sub(
#       string = comp.K562.s.CTSSs$sample, start = 1, end = -2
#     )
# ggplot2::ggplot(
#     data = comp.K562.s.CTSSs,
#     mapping = aes(
#         x = totalTags,
#         y = numberCTSSs,
#         color = method,
#         group = sample
#       )
#     ) + 
#   geom_line() +
#   geom_abline(
#       intercept = 0,
#       slope = 1,
#       linetype = "dashed",
#       color = "black",
#       size = 0.5
#     ) +
#   theme_bw() +
#   scale_color_manual(
#       values = cols3
#     ) +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.position = base::c(0.95,0.05),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

```{r}
# comp.K562.p.CTSSs$method <- stringr::str_sub(
#     string = comp.K562.p.CTSSs$sample, start = 1, end = -2
#   )
# ggplot2::ggplot(
#     data = comp.K562.p.CTSSs,
#     mapping = aes(
#         x = totalTags,
#         y = numberCTSSs,
#         color = method,
#         group = sample
#       )
#     ) + 
#   geom_line() +
#   geom_abline(
#       intercept = 0,
#       slope = 1,
#       linetype = "dashed",
#       color = "black",
#       size = 0.5
#     ) +
#   theme_bw() +
#   scale_color_manual(
#       values = cols3
#     ) +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.position = base::c(0.95,0.05),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

## [1.8.] Identify max values for coordinated plotting
#### See [2.5.] below for "comp.GM12878.CTSSs.plot" object
```{r}
tt1 <- comp.K562.CTSSs.plot %>% # 284000983
  dplyr::pull(totalTags) %>%
  base::max(na.rm = TRUE)
tt2 <- comp.GM12878.CTSSs.plot %>% # 399999736
  dplyr::pull(totalTags) %>%
  base::max(na.rm = TRUE)

nc1 <- comp.K562.CTSSs.plot %>% # 29717697
  dplyr::pull(numberCTSSs) %>%
  base::max(na.rm = TRUE)
nc2 <- comp.GM12878.CTSSs.plot %>% # 49703283
  dplyr::pull(numberCTSSs) %>%
  base::max(na.rm = TRUE)

xmax <- base::max(tt1,tt2) # 399999736
ymax <- base::max(nc1,nc2) # 49703283
```

## [1.9.] Test plot all complexity curves
#### If required change scaling with 'scale_y_log10()'
```{r}
ggplot2::ggplot(
    data = comp.K562.CTSSs.plot,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = method,
        group = sample
      )
    ) + 
  geom_line() +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  geom_vline(
      xintercept = base::c(28623910,96353854),
      linetype = 2
    ) +
    scale_x_continuous(
      limits = base::c(0,xmax),
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.10.] Plot complexity curves for replicates and pools for printing
#### If required change scaling with 'scale_y_log10()'
```{r}
ggplot2::ggplot(
    data = comp.K562.CTSSs.plot,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line(
      size = 1
    ) +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  geom_vline(
      xintercept = base::c(28623910,96353854),
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
      name = "Total number of tags"
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
      name = "Number of CTSSs"
    ) +
  scale_color_manual(
      values = k.full.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.11.] Save complexity plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_complexity.K562.CTSSs",
          "_all",
          # "_replicates",# "_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.12.] Plot subsampled number of CTSSs per library
## Add barplots for subsampled levels
```{r}
p.max <- comp.K562.CTSSs.plot %>%
  dplyr::filter(target < 96353854) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
r.max <- comp.K562.CTSSs.plot %>%
  dplyr::filter(target < 28623910) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
```

```{r}
ggplot2::ggplot(
    data = p.max, # r.max,
    mapping = aes(
        x = compexity,
        y = forcats::fct_rev(sample),
        fill = sample
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = sample
      ),
      x = 0,
      hjust = -0.1,
      colour = "white",
      size = 3,
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  # scale_x_log10() +
  guides(fill = "none") +
  labs(
      x = "Number of CTSSs"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
```

## [1.13.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexity.K562.CTSSs",
          "_barplots",
          "_all", # "_replicate", # "_pool",
          "_poolSub", # "_repSub",
          # "_nonLog10",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 3,
      height = 5
    )
  }
```

# [2.] GM12878 -------------------------

## [2.0.] Read in GM12878 datasets
#### see '1.1CTSSs.processing.Rmd' for read objects
```{r}
GM12878.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )
# GM12878.s.CTSSs <- base::readRDS(
#     file = base::paste0(
#       here::here("1.CTSSs/"),
#       "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
# GM12878.p.CTSSs <- base::readRDS(
#     file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
```

## [2.1.] Compare all CTSS level correlation
```{r}
corrplot::corrplot(
    stats::cor(
      x = base::as.matrix(
        x = SummarizedExperiment::assay(
            x = GM12878.CTSSs,
            i = "TPM"
          )
        )
      ),
    method = "circle", # "number"
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878.CTSSs.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 12,
      height = 12
    )
  }
```

```{r}
# corrplot::corrplot(
#     stats::cor(
#       x = base::as.matrix(
#         x = SummarizedExperiment::assay(
#             x = GM12878.s.CTSSs,
#             i = "TPM"
#           )
#         )
#       ),
#     method = "circle", # "number"
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename =base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_subsampled.GM12878.CTSSs.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

```{r}
# corrplot::corrplot(
#     stats::cor(
#       x = base::as.matrix(
#         x = SummarizedExperiment::assay(
#             x = GM12878.p.CTSSs,
#             i = "TPM"
#           )
#         )
#       ),
#     method = "circle", # "number"
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename =base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_pooled.GM12878.CTSSs.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

## [2.2.] Compare CTSS>0TPM correlations
```{r}
m1 <- base::as.matrix(
    x = SummarizedExperiment::assay(x = GM12878.CTSSs, i = "TPM")
  )
m1[m1 == 0] <- NA
corrplot::corrplot(
  corr = stats::cor(m1, use = "complete.obs"),
    method = "circle",
    order = 'hclust',
    addrect = 2,
    col = base::rev(corrplot::COL2("RdBu",200)),
    tl.col = 'black'
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878.CTSSs.rm0.corr",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 8
    )
  }
```

```{r}
# m1 <- base::as.matrix(
#     x = SummarizedExperiment::assay(x = GM12878.s.CTSSs, i = "TPM")
#   )
# m1[m1 == 0] <- NA
# corrplot::corrplot(
#   corr = stats::cor(m1, use = "complete.obs"),
#     method = "circle",
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_subsampled.GM12878.CTSSs.rm0.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

```{r}
# m1 <- base::as.matrix(
#     x = SummarizedExperiment::assay(x = GM12878.p.CTSSs, i = "TPM")
#   )
# m1[m1 == 0] <- NA
# corrplot::corrplot(
#   corr = stats::cor(m1, use = "complete.obs"),
#     method = "circle",
#     order = 'hclust',
#     addrect = 2,
#     col = base::rev(corrplot::COL2("RdBu",200)),
#     tl.col = 'black'
#   )
```

```{r}
# 
# for(ext in base::c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("1.CTSSs/plots/"),
#           base::format(base::Sys.time(), "%Y%m%d"),
#           "_pooled.GM12878.CTSSs.rm0.corr",
#           ".",ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 8,
#       height = 8
#     )
#   }
```

## [2.3.] Calculate CTSS level complexity
```{r}
base::set.seed(1337)

comp.GM12878.CTSSs <- .GlobalEnv$calcCTSSComplexity(
    object = GM12878.CTSSs,
    step = 2e+6,
    minCTSSsupport = 1,
    CTSSunexpressed = 0
  )
# comp.GM12878.s.CTSSs <- .GlobalEnv$calcCTSSComplexity(
#     object = GM12878.s.CTSSs,
#     step = 2e+6,
#     minCTSSsupport = 1,
#     CTSSunexpressed = 0
#   )
# comp.GM12878.p.CTSSs <- .GlobalEnv$calcCTSSComplexity(
#     object = GM12878.p.CTSSs,
#     step = 2e+6,
#     minCTSSsupport = 1,
#     CTSSunexpressed = 0
#   )
```

## [2.4.] Save & load objects
```{r}
base::saveRDS(
  object = comp.GM12878.CTSSs,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_comp.GM12878.CTSSs",
      ".rds"
    )
  )
```

```{r}
comp.GM12878.CTSSs <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("comp.GM12878.CTSSs.rds")
  )
```

## [2.5.] Data wrangling for all K562 samples & pools
#### Name change of "K562_10M_N", remove "K562_Cap_align" sample
```{r}
 levs <- base::c(
    "GM12878_500K_C", "GM12878_500K_C1", "GM12878_500K_C2", "GM12878_500K_C3",
    "GM12878_500K_N", "GM12878_500K_N1", "GM12878_500K_N2", "GM12878_500K_N3",
    "GM12878_1M_C", "GM12878_1M_C1", "GM12878_1M_C2", "GM12878_1M_C3",
    "GM12878_1M_N", "GM12878_1M_N1", "GM12878_1M_N2", "GM12878_1M_N3",
    "GM12878_5M_N", "GM12878_5M_N1", "GM12878_5M_N2", "GM12878_5M_N3", 
    "GM12878_10M_N", "GM12878_10M_N1", "GM12878_10M_N2", "GM12878_10M_N3"
  )
alts <- base::c(
    "GM12878_50K_C", "GM12878_50K_C1", "GM12878_50K_C2", "GM12878_50K_C3",
    "GM12878_250K_C", "GM12878_250K_C1", "GM12878_250K_C2", "GM12878_250K_C3",
    "GM12878_1M_trip", "GM12878_triptolide_C1", "GM12878_triptolide_C2", "GM12878_triptolide_C3", "GM12878_triptolide_C4",
    "GM12878_1M_ctrl", "GM12878_ctrl_C1", "GM12878_ctrl_C2", "GM12878_ctrl_C3",
    "GM12878_10M_Cap_align", "GM12878_groCap_1"
  )

comp.GM12878.CTSSs.plot <- comp.GM12878.CTSSs %>%
  dplyr::filter(
      sample %in% levs
    ) %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = levs,
        ordered = TRUE
      ),
    method = stringr::str_replace(
        sample,"^(.*?)(_)?(\\d)?$","\\1"
      )
    )
```

## [2.6.] Test plotting of complexity curves for replicates and pools
```{r}
ggplot2::ggplot(
    data = comp.GM12878.CTSSs.plot,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = method,
        group = sample
      )
    ) + 
  geom_line(
      size = 1
    ) +
  geom_vline(
      xintercept = base::c(28462994,104779864),
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [2.7.] Plot complexities for the other objects
#### "./plots/subsampled.GM12878.CTSSs.comp.pdf"
```{r}
# ggplot2::ggplot(
#     data = comp.GM12878.s.CTSSs,
#     mapping = aes(
#         x = totalTags,
#         y = numberCTSSs,
#         color = method,
#         group = sample
#       )
#     ) + 
#   geom_line() +
#   geom_abline(
#       intercept = 0,
#       slope = 1,
#       linetype = "dashed",
#       color = "black",
#       size = 0.5
#     ) +
#   theme_bw() +
#   scale_color_manual(
#       values = cols3
#     ) +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.position = base::c(0.95,0.05),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

#### "./plots/pooled.GM12878.CTSSs.comp.pdf"
```{r}
# ggplot2::ggplot(
#     data = comp.GM12878.p.CTSSs,
#     mapping = aes(
#         x = totalTags,
#         y = numberCTSSs,
#         color = method,
#         group = sample
#       )
#     ) + 
#   geom_line() +
#   geom_abline(
#       intercept = 0,
#       slope = 1,
#       linetype = "dashed",
#       color = "black",
#       size = 0.5
#     ) +
#   theme_bw() +
#   scale_color_manual(
#       values = cols3
#     ) +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.position = base::c(0.95,0.05),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

## [2.8.] Plot complexities including pooled samples
#### see [1.8.] for xmax, ymax
#### If required, 
```{r}
ggplot2::ggplot(
    data = comp.GM12878.CTSSs.plot,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line(
      size = 1
    ) +
  geom_vline(
      xintercept = base::c(28462994,104779864),
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  geom_abline(
      intercept = 0,
      slope = 1,
      linetype = "dashed",
      color = "black",
      size = 0.5
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
      name = "Total number of tags"
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
      name = "Number of CTSSs"
    ) +
  scale_color_manual(
      values = gm.full.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none"
    )
```

#### Only pooled samples
#### previously: Only 500k and 5M samples, 'gm.full.ada' color palette used
```{r}
main %>%
  dplyr::filter(
      stringr::str_detect(sample,"_N$|_C$")
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  geom_vline(
      xintercept = base::c(29027358,106618677),
      linetype = 2
    ) +
  scale_x_continuous(
      limits = base::c(0,xmax),
    ) +
  scale_y_continuous(
      limits = base::c(0,ymax),
    ) +
  scale_color_manual(
      values = gm.full.pal[base::seq(1,24,4)]
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.9.] Print data
```{r}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          format(Sys.time(), "%Y%m%d"),
          "_comp.GM12878.CTSSs",
          "_all",
          # "_replicates",# "_pools",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.10.] Plot subsampled number of CTSSs per library
## Add barplots for subsampled levels
```{r}
p.max <- main %>%
  dplyr::filter(target < 104779864) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
r.max <- main %>%
  dplyr::filter(target < 28462994) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
```

```{r}
p.max_pools <- main %>%
  dplyr::filter(
      stringr::str_detect(sample,"_N$|_C$")
    ) %>%
  dplyr::filter(target < 106618677L) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
r.max_pools <- main %>%
  dplyr::filter(
      stringr::str_detect(sample,"_N$|_C$")
    ) %>%
  dplyr::filter(target < 29027358) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarize(
      compexity = base::max(numberCTSSs)
    )
```

```{r}
ggplot2::ggplot(
    data = r.max_pools, # p.max_pools, # p.max, # r.max,
    mapping = aes(
        x = compexity,
        y = forcats::fct_rev(sample),
        fill = sample
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = sample
      ),
      x = 0,
      hjust = 0,
      colour = "white",
      size = 3,
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  # scale_x_log10() +
  guides(fill = "none") +
  labs(
      x = "numberCTSSs",
      y = "Samples"
    ) +
  theme_bw() +
  theme(
    aspect.ratio = 2,
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

## [2.10.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          format(Sys.time(), "%Y%m%d"),
          "_complexityCtss",
          "_GM12878",
          "_barplots",
          "_replicate",
          # "_pool",
          "_OnlyPool",
          "_nonLog10",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 3,
      height = 6
    )
  }
```

## [2.11.] Combined plot
```{r}
p1 <- ggplot2::ggplot(
    data = main,
    mapping = aes(
        x = totalTags,
        y = numberCTSSs,
        color = sample
      )
    ) + 
  geom_line() +
  theme_bw() +
  scale_y_log10() +
  scale_color_manual(
      values = gm.full.pal
    ) +
  geom_vline(
      xintercept = base::c(29027358,106618677),
      linetype = 2
    ) +
  guides(color="none") 

p2 <- ggplot2::ggplot(
  data = r.max,
  mapping = aes(
    x = forcats::fct_rev(sample),
    y = rep.compexity,
    fill = sample
    )
  ) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  coord_flip() +
  scale_y_log10() +
  guides(fill = "none") +
  labs(y = "numberCTSSs", x = "Samples")

p3 <- ggplot2::ggplot(
  data = p.max,
  mapping = aes(
      x = forcats::fct_rev(sample),
      y = pooled.compexity,
      fill = sample
    )
  ) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(
      values = gm.full.pal
    ) +
  coord_flip() +
  scale_y_log10() +
  labs(y = "numberCTSSs", x = "Samples")
```

## "./combined_plots/GM12878.mainComplexity.pdf"
# Extract the legend after adding some space to the left
```{r}
p3_leg <- p3 +
  theme(
      legend.box.margin = margin(0, 0, 0, 12)
    )
p3_noleg <- p3 +
  guides(
      fill = "none"
    )
legend <- get_legend(p3_leg)

gridExtra::grid.arrange(
  p1,p2,p3_noleg,
  widths = base::c(2, 1, 1),
  layout_matrix = base::rbind(
      base::c(1, 2, 3)
    )
  )
```
