---
title: "CTSS noise"
output: html_document
date: "2024-01-31"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(PRIME)
```

## [0.1.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.2.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
cols4 <- base::c(
    "#FE8D02","#FF4425","#00439B",
    "#008F72","#878784","#7D62CF"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
```

## [0.3.] dhsPlot() helper function
```{r}
dhsPlots <- function(data,pal){
  cur <- ggplot2::ggplot(
      data = data,
      mapping = ggplot2::aes(
          x = distance,
          y = value,
          color = method,
          group = variable
        )
      ) +
    ggplot2::geom_line() +
    ggplot2::geom_vline(
        xintercept = 200,
        alpha = 0.5,
        linetype = "dashed"
      ) +
    ggplot2::scale_color_manual(
        values = pal
      ) +
    ggplot2::labs(
        y = "CTSS to DHS cumulative fraction",
        x = "Distance to DHSs center",
        color = "sample"
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(
        legend.position = "none",
        aspect.ratio = 1.15
      )
  
  dis200 <- data %>% 
      dplyr::filter(
          distance == 200
        ) %>%
    ggplot2::ggplot(
        data = .,
        mapping = ggplot2::aes(
            x = value,
            y = variable,
            fill = method
          )
        ) +
      ggplot2::geom_col() +
      ggplot2::geom_text(
          mapping = ggplot2::aes(
              label = variable
            ),
          hjust = 1.1,
          color = "white",
          fontface = "bold"
        ) +
      ggplot2::scale_fill_manual(
          values = pal
        ) +
      ggplot2::labs(
          x = "Cumulative CTSS fraction at 200bp",
          y = "sample",
          fill = "sample"
        ) +
      ggplot2::theme_bw() +
      ggplot2::theme(
          aspect.ratio = 1,
          legend.position = "none",
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank()
        )
  
  return(
    gridExtra::grid.arrange(
        cur, dis200,
        ncol = 2
      )
    )
  }
```

# [1.] Cumulative fraction of K562 samples towards ENCODE DHS sites specific for K562 -------------------------

## [1.0.] Read in pre made CTSS objects
```{r}
# K562.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "all.combined.k562.CTSSs.rds"
#     )
#   )
K562.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
# K562.p.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.k562.CTSSs.rmSingletons.rds"
#     )
#   )
```

## [1.1.] Read in GM12878-specific DHSs & prepare as GRanges object
```{r}
K562.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E123-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

K562.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(K562.dhs) + (
        (BiocGenerics::end(K562.dhs) - BiocGenerics::start(K562.dhs))/2
      )
  )
```

## [1.2.] Ensure using only standard chromosomes
```{r}
K562.CTSSs.stand <- GenomeInfoDb::keepStandardChromosomes(
    x = K562.s.CTSSs,
    pruning.mode = "coarse"
  )
K562.dhs <- GenomeInfoDb::keepStandardChromosomes(
    x = K562.dhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(K562.CTSSs.stand) == GenomeInfoDb::seqlevels(K562.s.CTSSs)),
    base::all(GenomeInfoDb::seqlevels(K562.s.CTSSs) %in% GenomeInfoDb::seqlevels(K562.dhs))
  )
```

## [1.3.] Calculate cumulative fractions as a function of the distance to DHSs
```{r}
K562.cumFrac <- PRIME::cumulativeFractionsAroundLoci(
    loci = K562.dhs,
    ctss = K562.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = K562.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_K562.E123dhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
K562.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("K562.E123dhs.cumFrac")
  )
```

## [1.4.] Data wrangling
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [1.5.] Preparing data for plotting
```{r}
vrs <- base::c(
   "K562_C","K562_C1","K562_C2","K562_C3",
   "K562_N","K562_N1","K562_N2","K562_N3",
   "K562_10M_N","K562_10M_N1","K562_10M_N2","K562_10M_N3"
  )

ctss.plot <- ctss.cumfrac %>%
  dplyr::filter(variable %in% vrs) %>%
  dplyr::mutate(
      variable = base::factor(
        x = variable,
        levels = base::rev(vrs),
        ordered = TRUE
      )
    )# %>%
  # dplyr::filter( ## Filter for replicates or pools only
  #     stringr::str_detect(
  #       variable, "^.*_(C|N)\\d$" # "^.*_(C|N)$" = pools
  #     )
  #   )
```

#### Evaluate & test plotting
```{r}
dplyr::distinct(ctss.plot,variable)
dplyr::distinct(ctss.plot,method)
.GlobalEnv$dhsPlots(ctss.plot,cols3)
```

## [1.6.] Plotting
```{r}
ggplot2::ggplot(
    data = ctss.plot,
    mapping = aes(
        x = distance,
        y = value,
        color = variable
      )
    ) +
  geom_line() +
  geom_vline(
      xintercept = 200,
      alpha = 0.5,
      linetype = "dashed"
    ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "Distance to DHSs center"
    ) +
  scale_color_manual(
      values = gm.full.pal
    ) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
    # legend.title = element_blank(),
    # legend.justification = base::c(1,0),
    # legend.position = base::c(0.95,0.05),
    # legend.background = element_rect(
    #     fill = "grey"
    #   )
    )
```

## [1.7.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562",
          "_cumFracDhs",
          "_all",
          # "_replicates", # "_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.8.] Plotting
```{r}
ctss.plot %>% 
  dplyr::filter(
      distance == 200
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          x = value,
          y = variable,
          fill = variable
        )
      ) +
    geom_col() +
    geom_text(
        mapping = aes(
            label = variable
          ),
        x = 0,
        hjust = -1.1,
        color = "white",
        fontface = "bold"
      ) +
    scale_fill_manual(
        values = gm.full.pal # cols4
      ) +
    labs(
        x = "Cumulative CTSS fraction at 200bp",
        y = "sample",
        fill = "sample"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
      )
```

## [1.9.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562",
          "_cumFracDhs",
          "_all",
          # "_replicates", # "_pools",
          "_dist200",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.10.] Data wrangling for cumulative fractions of unique tags
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [1.11.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [1.12.] Data wrangling for cumulative fractions of unique tags
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [1.13.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

# [2.] Cumulative fraction of K562 samples towards all ENCODE DHS sites -------------------------

## [2.0.] Read in rDHS bed file form encode 
#### File copied from "/projects/ralab/people/pcr980/DeepCompare/Raw_data"
```{r}
rdhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "rDHS_ENCFF619EJE.bed"
        ),
      col_names = base::c("chr","start","end","site"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        site = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

rdhs$thick <- IRanges::IRanges(
  start = BiocGenerics::start(rdhs) + (
      (BiocGenerics::end(rdhs) - BiocGenerics::start(rdhs))/2
    )
  )
```

## [2.1.] Ensure using only standard chromosomes
```{r}
rdhs <- GenomeInfoDb::keepStandardChromosomes(
    x = rdhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(K562.s.CTSSs) %in% GenomeInfoDb::seqlevels(rdhs))
  )
```

## [2.2.] Calculate cumulative fractions as a function of the distance to DHSs
#### see [1.2.] for 'K562.CTSSs.stand' object
```{r}
K562.rDHS.cumFrac <- PRIME::cumulativeFractionsAroundLoci(
    loci = rdhs,
    ctss = K562.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = K562.rDHS.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_K562.rdhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
K562.rDHS.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("K562.rdhs.cumFrac")
  )
```

## [2.3.] Data wrangling
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.rDHS.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [2.4.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [2.5.] Data wrangling
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.rDHS.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [2.6.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [2.7.] Data wrangling
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.rDHS.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [2.8.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

# [3.] Cumulative fraction of GM12878 samples towards ENCODE DHS sites specific for GM12878 -------------------------

## [3.0.] Read in pre-made CTSS objects
#### see "1.1.CTSSs.processing.Rmd"
```{r}
# GM12878.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "all.combined.GM12878.CTSSs.rds"
#     )
#   )
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
# GM12878.p.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
```

## [3.1.] Read in GM12878-specific DHSs & prepare as GRanges object
```{r}
GM12878.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E116-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(GM12878.dhs) + (
        (BiocGenerics::end(GM12878.dhs) - BiocGenerics::start(GM12878.dhs))/2
      )
  )
```

## [3.2.] Ensure using only standard chromosomes
```{r}
GM12878.CTSSs.stand <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.s.CTSSs,
    pruning.mode = "coarse"
  )
GM12878.dhs <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.dhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(GM12878.CTSSs.stand) == GenomeInfoDb::seqlevels(GM12878.s.CTSSs)),
    base::all(GenomeInfoDb::seqlevels(GM12878.s.CTSSs) %in% GenomeInfoDb::seqlevels(GM12878.dhs))
  )
```

## [3.3.] Calculate cumulative fractions as a function of the distance to DHSs
#### 'GM12878.p.s.CTSSs' object used for pool-plots
#### 'GM12878.pID.s.CTSSs' object used for replicate-plots
```{r}
GM12878.cumFrac <- PRIME::cumulativeFractionsAroundLoci(
    loci = GM12878.dhs,
    ctss = GM12878.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = GM12878.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.E116dhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
GM12878.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.E116dhs.cumFrac")
  )
```

## [3.4.] Data wrangling
#### Previous regex to extract 'C_N' column: 'regex = "^(.*)_(.*?)_(C|N)\\d$"' or 'regex = "^(.*)_(.*?)_(C|N)$"'
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
      method = stringr::str_replace(
          variable, "^(.*)_(.*?)_(.*)$","\\2"
        ),
      C_N = stringr::str_replace(
          variable, regex,"\\3"
        ),
      type = base::paste0(C_N,"_",method)
    )
```

#### Evaluate & test plotting
```{r}
dplyr::distinct(ctss.cumfrac,variable)
dplyr::distinct(ctss.cumfrac,method)
dplyr::distinct(ctss.cumfrac,type)
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [3.5.] Preparing data for plotting
```{r}
tps <- base::c(
    "C_500K","N_500K","C_1M",
    "N_1M","N_5M","N_10M"
  )
vrs <- base::c(
    "GM12878_500K_C","GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3",
    "GM12878_500K_N","GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3",
    "GM12878_1M_C","GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3",
    "GM12878_1M_N","GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3",
    "GM12878_5M_N","GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3",
    "GM12878_10M_N","GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3"
  )

ctss.plot <- ctss.cumfrac %>%
  dplyr::filter(type %in% tps) %>%
  dplyr::mutate(
      type = base::factor(
          x = type,
          levels = base::rev(tps),
          ordered = TRUE
        ),
      variable = base::factor(
        x = variable,
        levels = base::rev(vrs),
        ordered = TRUE
      )
    ) # %>%
  # dplyr::filter( ## Filter for replicates or pools only
  #     stringr::str_detect(
  #       variable, "^.*_(C|N)$" # "^.*_(C|N)\\d$"
  #     )
  #   )
```

## [3.6.] Plotting
```{r}
ggplot2::ggplot(
      data = ctss.plot,
      mapping = aes(
          x = distance,
          y = value,
          # group = variable, #type,
          color = variable #type
        )
      ) +
    geom_vline(
        xintercept = 200L,
        linetype = "dashed"
      ) +
    geom_line(
        size = 1
      ) +
    # facet_wrap(
    #     facets = ~ method #+ C_N
    #   ) +
    # guides(
    #     color = "none"
    #   ) +
    labs(
        y = "CTSS to DHS cumulative fraction",
        x = "Distance to DHSs center"
      ) +
    scale_color_manual(
        values = gm.full.pal #base::rev(cols4)
      ) +
    theme_bw() +
    theme(
      aspect.ratio = 1,
      legend.position = "none"
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.position = base::c(0.95,0.05),
      # legend.background = element_rect(
      #     fill = "grey"
      #   )
      )
```

## [3.7.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          "_all",
          # "_replicates", #"_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [3.8.] Plotting
```{r}
ctss.plot %>%
  dplyr::filter(
      distance == 200L
    ) %>%
ggplot2::ggplot(
    data = ., #ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = variable, #type,
        x = value,
        fill = variable #type
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = variable
      ),
      x = 0,
      hjust = -1.1,
      color = "white",
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = base::rev(gm.full.pal) # base::rev(cols4)
    ) +
  labs(
      x = "Cumulative CTSS fraction at 200bp",
      y = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.text.y = element_blank()
    )
```

## [3.9.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          "_all",
          # "_replicates", #"_pools",
          "_dist200",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [3.10.] Data wrangling for cumulative fraction of unique tags
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

#### Evaluate
```{r}
dplyr::distinct(ctss.cumfrac, method)
```

## [3.11.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [3.12.] Data wrangling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [3.13.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

# [4.] Cumulative fraction of GM12878 samples towards all ENCODE DHS sites -------------------------

## [4.0.] Ensure using only standard chromosomes
#### see [2.0.] above for generating 'rdhs' object
```{r}
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(GM12878.s.CTSSs) %in% GenomeInfoDb::seqlevels(rdhs))
  )
```

## [4.1.] Calculate cumulative fractions as a function of the distance to DHSs
#### 'GM12878.p.s.CTSSs' object used for pool-plots
#### 'GM12878.pID.s.CTSSs' object used for replicate-plots
```{r}
GM12878.rDHS.cumFrac <- PRIME::cumulativeFractionsAroundLoci(
    loci = rdhs,
    ctss = GM12878.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = gm12878.rDHS.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.rdhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
GM12878.rDHS.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.rdhs.cumFrac")
  )
```

## [4.2.] Data wrangling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [4.3.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [4.4.] Data wragnling for cumulative fraction of unique tags
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [4.5.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```

## [4.6.] Data wragnling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [4.7.] Plotting
```{r}
.GlobalEnv$dhsPlots(data = ctss.cumfrac, pal = cols3)
```
