---
title: "1.3Noise"
output: html_document
date: "2024-01-31"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(BiocParallel)
```

## [0.1.] Load CAGEfighR extensions
#### previous: base::setwd("/home/mhf817"), cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
cols4 <- base::c(
    "#FE8D02","#FF4425","#00439B",
    "#008F72","#878784","#7D62CF"
  )
```

## [0.4.] cumulativeFraction function
```{r}
cumulativeFractions <- function(loci, ctss, max_dist = 1000, inputAssay = "counts") {
    
    g <- GenomicRanges::GRanges(
        seqnames = GenomeInfoDb::seqnames(loci),
        ranges = loci$thick,
        strand = "*"
      )
    t <- methods::as(SummarizedExperiment::rowRanges(ctss), "GRanges")
    
    base::message("finding nearest loci...")
    g_nearest <- IRanges::nearest(t, g)
    
    base::message("calculating distances...")
    d <- base::rep(Inf, base::length(t))
    non.na <- base::which(!base::is.na(g_nearest))
    d[non.na] <- IRanges::distance(t[non.na], g[g_nearest[non.na]])
    
    base::message("calculating totalTags and unique Tags...")
    ctss <- base::suppressWarnings(
      CAGEfightR::calcTotalTags(
          object = ctss,
          inputAssay = "counts"
        )
      )
    SummarizedExperiment::colData(ctss)$uniqTags <- Matrix::colSums(
        SummarizedExperiment::assay(ctss,inputAssay="counts") > 0
      )
    
    base::message("calculating cumulative fractions...")
    focus <- base::which(d <= max_dist)
    d <- d[focus]
    ctss <- ctss[focus]
    
    o <- base::order(d)
    d <- d[o]
    
    unique_loc_frac <- base::sapply(
        X = base::colnames(ctss),
        FUN = function(n) {
          base::cat(".")
          a <- base::as.numeric(
              SummarizedExperiment::assay(ctss, inputAssay)[o, n]
            )
          a[a>0] <- 1
          nz <- base::which(a > 0)
          da <- d[nz]
          a_sub <- a[nz]
          return(
            base::sapply(
              X = base::seq(0, max_dist, by = 1),
              FUN = function(i){base::sum(a_sub[da == i])}
            )
          )
        }
      )
    base::cat("\n")
    
    fracs <- base::sapply(
      X = base::colnames(ctss),
      FUN = function(n) {
        base::cat(".")
        a <- base::as.numeric(
            SummarizedExperiment::assay(ctss, inputAssay)[o, n]
          )
        nz <- base::which(a > 0)
        da <- d[nz]
        a_sub <- a[nz]
        return(
          base::sapply(
              X = base::seq(0, max_dist, by = 1),
              FUN = function(i){base::sum(a_sub[da == i])}
            )
          )
        }
      )
    base::cat("\n")
    
    t <- methods::as(SummarizedExperiment::rowRanges(ctss), "GRanges")
    
    loc_fracs <- base::sapply(
        X = base::colnames(ctss),
        FUN = function(n) {
          base::cat(".")
          a <- base::as.numeric(
              SummarizedExperiment::assay(ctss, inputAssay)[o, n]
            )
        nz <- t[base::which(a > 0)]
        dist <- IRanges::distanceToNearest(g,nz)
        dist <- base::as.data.frame(dist)[,3]
        return(
            base::sapply(
              X = base::seq(0, max_dist, by = 1),
              FUN = function(i){base::sum(dist==i)}
            )
          )
        }
      )
    
    unique_loc_frac <- base::sapply(
        X = base::colnames(unique_loc_frac),
        FUN = function(n){
          base::cumsum(unique_loc_frac[, n] / ctss$uniqTags[n])
        }
      )
    fracs <- base::sapply(
      X = base::colnames(fracs),
      FUN = function(n){
          base::cumsum(fracs[, n] / ctss$totalTags[n])
        }
      )
    loc_fracs <- base::sapply(
      X = base::colnames(loc_fracs),
      FUN = function(n){
          base::cumsum(loc_fracs[, n] / base::length(g))
        }
      )
    
    ucumfrac <- base::cbind(
      base::data.frame(
          distance = base::seq(0, max_dist, by = 1),
          unique_loc_frac
        )
      )
    cumfrac <- base::cbind(
      base::data.frame(
          distance = base::seq(0, max_dist, by = 1),
          fracs
        )
      )
    locifrac <- base::cbind(
      base::data.frame(
          distance = base::seq(0, max_dist, by = 1),
          loc_fracs
        )
      )
    
    fraclist <- base::list(cumfrac,ucumfrac,locifrac)
    return(fraclist)
  }
```

# [1.] K562 data -------------------------

## [1.0.] Read in pre made CTSS objects
```{r}
# K562.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "all.combined.k562.CTSSs.rds"
#     )
#   )
K562.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
# K562.p.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.k562.CTSSs.rmSingletons.rds"
#     )
#   )
```

## To be discarded
```{r}
# k562.CTSSs <- readRDS("all.k562.CTSSs.rds")
# k562.s.CTSSs <- readRDS("subsampled.k562.CTSSs.rds")
# k562.p.CTSSs <- readRDS("pooled.k562.CTSSs.rds")
# k562.p.s.CTSSs <- readRDS("pooled.subsampled.k562.CTSSs.rds")
```

## [1.1.] Read in GM12878-specific DHSs & prepare as GRanges object
```{r}
K562.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E123-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

K562.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(K562.dhs) + (
        (BiocGenerics::end(K562.dhs) - BiocGenerics::start(K562.dhs))/2
      )
  )
```

## To be discarded
```{r}
# k562.dhs <- read.table("../0.External_resources/E123-DNase.macs2.hg38.narrowPeak")
# colnames(k562.dhs) <- c("chr","start","end")
# k562.dhs <- GRanges(k562.dhs[,1:3])
# k562.dhs$thick <- IRanges(start=start(k562.dhs)+((end(k562.dhs)-start(k562.dhs))/2))
```

## [1.2.] Ensure using only standard chromosomes
```{r}
K562.CTSSs.stand <- GenomeInfoDb::keepStandardChromosomes(
    x = K562.s.CTSSs,
    pruning.mode = "coarse"
  )
K562.dhs <- GenomeInfoDb::keepStandardChromosomes(
    x = K562.dhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(K562.CTSSs.stand) == GenomeInfoDb::seqlevels(K562.s.CTSSs)),
    base::all(GenomeInfoDb::seqlevels(K562.s.CTSSs) %in% GenomeInfoDb::seqlevels(K562.dhs))
  )
```

## [1.3.] Calculate cumulative fractions as a function of the distance to DHSs
```{r}
K562.cumFrac <- .GlobalEnv$cumulativeFractions(
    loci = K562.dhs,
    ctss = K562.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = K562.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_K562.E123dhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
K562.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("K562.E123dhs.cumFrac")
  )
```

## To be discarded
```{r}
# k562.CTSSs.stand <- keepStandardChromosomes(k562.s.CTSSs,pruning.mode = "coarse")
# k562.dhs<- keepStandardChromosomes(k562.dhs,pruning.mode = "coarse")
# cumFrac <- cumulativeFractions(k562.dhs,k562.CTSSs.stand) 
# 
# ctss.cumfrac <- reshape2::melt(K562.cumFrac[[1]], id.var="distance")
# ctss.cumfrac$method <- stringr::str_sub(ctss.cumfrac$variable,1,-2)
```

## [1.4.] Data wrangling
```{r}
regex = "^(K562_)(.*?)(_)?(\\d)?$"

ctss.cumfrac <- K562.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("K562_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = dplyr::case_when(
        variable == "K562_N_10M" ~ "K562_10M_N",
        variable == "K562_N4" ~ "K562_N1",
        variable == "K562_N5" ~ "K562_N2",
        variable == "K562_N6" ~ "K562_N3",
        TRUE ~ variable
      ),
    method = stringr::str_replace(
        variable, regex, "\\2"
      )
    )
```

## [1.5.] Preparing data for plotting
```{r}
vrs <- base::c(
   "K562_C","K562_C1","K562_C2","K562_C3",
   "K562_N","K562_N1","K562_N2","K562_N3",
   "K562_10M_N","K562_10M_N1","K562_10M_N2","K562_10M_N3"
  )

ctss.plot <- ctss.cumfrac %>%
  dplyr::filter(variable %in% vrs) %>%
  dplyr::mutate(
      variable = base::factor(
        x = variable,
        levels = base::rev(vrs),
        ordered = TRUE
      )
    ) %>%
  dplyr::filter( ## Filter for replicates or pools only
      stringr::str_detect(
        variable, "^.*_(C|N)\\d$" # "^.*_(C|N)$" = pools
      )
    )
```

#### Evaluate
```{r}
dplyr::distinct(ctss.plot,variable)
dplyr::distinct(ctss.plot,method)
```

## [1.6.] Plotting
```{r}
ggplot2::ggplot(
    data = ctss.plot,
    mapping = aes(
        x = distance,
        y = value,
        color = method,
        group = variable
      )
    ) +
  geom_line() +
  geom_vline(
      xintercept = 200,
      alpha = 0.5,
      linetype = "dashed"
    ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "Distance to DHSs center"
    ) +
  scale_color_manual(
      values = cols4
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.position = base::c(0.95,0.05),
      legend.background = element_rect(
        fill = "grey"
      )
    )
```

## [1.7.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562",
          "_cumFracDhs",
          "_replicates",
          # "_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.8.] Plotting
```{r}
ctss.plot %>% 
  dplyr::filter(
      distance == 200
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          x = value,
          y = variable,
          fill = method
        )
      ) +
    geom_col() +
    geom_text(
        mapping = aes(
            label = variable
          ),
        x = 0,
        hjust = -1.1,
        color = "white",
        fontface = "bold"
      ) +
    scale_fill_manual(
        values = cols4
      ) +
    labs(
        x = "Cumulative CTSS fraction at 200bp",
        y = "sample",
        fill = "sample"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
      )
```

## [1.9.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_K562",
          "_cumFracDhs",
          "_replicates",
          # "_pools",
          "_dist200",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

```{r}
uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
```

```{r}
ggplot2::ggplot(
    data = uctss.cumfrac,
    mapping = aes(
        x = distance,
        y = value,
        color = method,
        group = variable
      )
    ) +
  geom_line() +
  theme_bw() +
  labs(
      y = "unique CTSS to DHS cumulative fraction",
      x = "distance to DHSs center",
      color = "sample"
    ) +
  scale_color_manual(
      values = custom.col
    ) +
  geom_vline(
      xintercept = 200,
      alpha = 0.5,
      linetype = "dashed"
    ) +
  theme(
      legend.position = "none"
    )
```


```{r}
ggplot2::ggplot(
    data = uctss.cumfrac[uctss.cumfrac$distance==200,],
    mapping = aes(
        x = variable,
        y = value,
        fill = method
      )
    ) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(
      values = custom.col
    ) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
```

```{r}
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
pdf("./plots/cumfrac.k562.subsampled.DHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

## [1.4.] Data wrangling & plotting
```{r}
#Remove non standard chromsosomes for fair comparison
k562.CTSSs.stand <- keepStandardChromosomes(k562.CTSSs,pruning.mode = "coarse")
k562.dhs<- keepStandardChromosomes(k562.dhs,pruning.mode = "coarse")
cumFrac <- cumulativeFractions(k562.dhs,k562.CTSSs.stand) 

ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
```

```{r}
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")
p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
```

```{r}
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
```

```{r}
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
pdf("./plots/cumfrac.k562.DHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

```{r}
k562.rdhs.cumFrac <- cumulativeFractions(rdhs,k562.s.CTSSs.stand) 


ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")


p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")


uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")

p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")


loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")


p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")



pdf("./plots/cumfrac.k562.subsampled.rDHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

```{r}
k562.rdhs.cumFrac <- cumulativeFractions(rdhs,k562.CTSSs.stand) 


ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")


p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")


uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")

p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")


loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")


p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")



pdf("./plots/cumfrac.k562.rDHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

# [2.] GM12878 data -------------------------

## [2.0.] Read in pre-made CTSS objects
#### see "1.1.CTSSs.processing.Rmd"
```{r}
# GM12878.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "all.combined.GM12878.CTSSs.rds"
#     )
#   )
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
# GM12878.p.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
```

## [2.1.] Read in GM12878-specific DHSs & prepare as GRanges object
```{r}
GM12878.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E116-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(GM12878.dhs) + (
        (BiocGenerics::end(GM12878.dhs) - BiocGenerics::start(GM12878.dhs))/2
      )
  )
```

## [2.2.] Ensure using only standard chromosomes
```{r}
GM12878.CTSSs.stand <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.s.CTSSs,
    pruning.mode = "coarse"
  )
GM12878.dhs <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.dhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(GM12878.CTSSs.stand) == GenomeInfoDb::seqlevels(GM12878.s.CTSSs)),
    base::all(GenomeInfoDb::seqlevels(GM12878.s.CTSSs) %in% GenomeInfoDb::seqlevels(GM12878.dhs))
  )
```

## [2.3.] Calculate cumulative fractions as a function of the distance to DHSs
#### 'GM12878.p.s.CTSSs' object used for pool-plots
#### 'GM12878.pID.s.CTSSs' object used for replicate-plots
```{r}
GM12878.cumFrac <- .GlobalEnv$cumulativeFractions(
    loci = GM12878.dhs,
    ctss = GM12878.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = GM12878.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.E116dhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
GM12878.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.E116dhs.cumFrac")
  )
```

## [2.4.] Data wrangling
#### Previous regex to extract 'C_N' column: 'regex = "^(.*)_(.*?)_(C|N)\\d$"' or 'regex = "^(.*)_(.*?)_(C|N)$"'
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM128787.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

#### Evaluate
```{r}
dplyr::distinct(ctss.cumfrac,variable)
dplyr::distinct(ctss.cumfrac,method)
dplyr::distinct(ctss.cumfrac,type)
```

## [2.5.] Preparing data for plotting
```{r}
tps <- base::c(
    "C_500K","N_500K","C_1M",
    "N_1M","N_5M","N_10M"
  )
vrs <- base::c(
    "GM12878_500K_C","GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3",
    "GM12878_500K_N","GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3",
    "GM12878_1M_C","GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3",
    "GM12878_1M_N","GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3",
    "GM12878_5M_N","GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3",
    "GM12878_10M_N","GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3"
  )

ctss.plot <- ctss.cumfrac %>%
  dplyr::filter(type %in% tps) %>%
  dplyr::mutate(
      type = base::factor(
          x = type,
          levels = base::rev(tps),
          ordered = TRUE
        ),
      variable = base::factor(
        x = variable,
        levels = base::rev(vrs),
        ordered = TRUE
      )
    ) %>%
  dplyr::filter( ## Filter for replicates or pools only
      stringr::str_detect(
        variable, "^.*_(C|N)$" # "^.*_(C|N)\\d$"
      )
    )
```

## [2.6.] Plotting
```{r}
ggplot2::ggplot(
      data = ctss.plot,
      mapping = aes(
          x = distance,
          y = value,
          group = variable, #type,
          color = type
        )
      ) +
    geom_vline(
        xintercept = 200L,
        linetype = "dashed"
      ) +
    geom_line(
        size = 1
      ) +
    # facet_wrap(
    #     facets = ~ method #+ C_N
    #   ) +
    # guides(
    #     color = "none"
    #   ) +
    labs(
        y = "CTSS to DHS cumulative fraction",
        x = "Distance to DHSs center"
      ) +
    scale_color_manual(
        values = base::rev(cols4)
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        legend.title = element_blank(),
        legend.justification = base::c(1,0),
        legend.position = base::c(0.95,0.05),
        legend.background = element_rect(
          fill = "grey"
        )
      )
```

## [2.7.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          # "_replicates",
          "_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.8.] Plotting
```{r}
ctss.plot %>%
  dplyr::filter(
      distance == 200L
    ) %>%
ggplot2::ggplot(
    data = ., #ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = variable, #type,
        x = value,
        fill = type
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = type
      ),
      x = 0,
      hjust = -1.1,
      color = "white",
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = base::rev(cols4)
    ) +
  labs(
      x = "Cumulative CTSS fraction at 200bp",
      y = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.text.y = element_blank()
    )
```

## [2.9.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          # "_replicates",
          "_pools",
          "_dist200",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.10.] Data wrangling for cumulative fraction of unique tags
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

#### Evaluate
```{r}
dplyr::distinct(ctss.cumfrac, method)
```

## [2.11.]
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac,
    mapping = aes(
        x = distance,
        y = value,
        group = variable,
        color = method
      )
    ) +
  geom_line() +
  # facet_wrap(
  #     facets = ~ method + C_N
  #   ) +
  # guides(
  #     color="none"
  #   ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "distance to DHSs center"
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.12.]
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = variable,
        x = value,
        fill = method
      )
    ) +
  geom_col() +
  scale_fill_manual(
      values = cols3
    ) +
  labs(
      x = "Cumulative unique CTSS fraction at 200bp",
      y = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.13.] Data wrangling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [2.14.]
```{r}
ggplot2::ggplot(
  data = ctss.cumfrac,
  mapping = aes(
      x = distance,
      y = value,
      group = variable,
      color = method
    )
  ) +
  geom_line() +
  facet_wrap(
      facets = ~ method + C_N
    ) +
  guides(
      color = "none"
    ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "distance to DHSs center"
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.15.]
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = variable,
        x = value,
        fill = method
      )
    ) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(
      values = cols3
    ) +
  labs(
      x = "Cumulative DHS fraction at 200bp",
      y = "sample",
      fill = "sample"
    )
```

# [3.] Cumulative fraction towards all ENCODE DHS sites -------------------------

## [3.0.] Read in rDHS bed file form encode 
#### File copied from "/projects/ralab/people/pcr980/DeepCompare/Raw_data"
```{r}
rdhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "rDHS_ENCFF619EJE.bed"
        ),
      col_names = base::c("chr","start","end","site"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        site = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

rdhs$thick <- IRanges::IRanges(
  start = BiocGenerics::start(rdhs) + (
      (BiocGenerics::end(rdhs) - BiocGenerics::start(rdhs))/2
    )
  )
```

## [3.1.] Ensure using only standard chromosomes
```{r}
rdhs <- GenomeInfoDb::keepStandardChromosomes(
    x = rdhs,
    pruning.mode = "coarse"
  )
assertthat::assert_that(
    base::all(GenomeInfoDb::seqlevels(GM12878.s.CTSSs) %in% GenomeInfoDb::seqlevels(rdhs))
  )
```

## [3.2.] Calculate cumulative fractions as a function of the distance to DHSs
#### 'GM12878.p.s.CTSSs' object used for pool-plots
#### 'GM12878.pID.s.CTSSs' object used for replicate-plots
```{r}
GM12878.rDHS.cumFrac <- .GlobalEnv$cumulativeFractions(
    loci = rdhs,
    ctss = GM12878.CTSSs.stand
  )
```

```{r}
base::saveRDS(
  object = gm12878.rDHS.cumFrac,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.rdhs.cumFrac",
      ".rds"
    )
  )
```

```{r}
GM12878.rDHS.cumFrac <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.rdhs.cumFrac")
  )
```

## [3.3.] Data wrangling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [3.4.] Plotting
```{r}
ggplot2::ggplot(
  data = ctss.cumfrac,
  mapping = aes(
      x = distance,
      y = value,
      group = variable,
      color = method
    )
  ) +
  geom_line() +
  # facet_wrap(
  #     facets = ~ method + C_N
  #   ) +
  # guides(
  #     color = "none"
  #   ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "distance to DHSs center"
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.5.] Plotting
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        x = value,
        y = variable,
        fill = method
      )
    ) +
  geom_col() +
  scale_fill_manual(
      values = cols3
    ) +
  labs(
      x = "Cumulative CTSS fraction at 200bp",
      y = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.6.] Data wragnling for cumulative fraction of unique tags
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(2) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [3.7.] Plotting
```{r}
ggplot2::ggplot(
  data = ctss.cumfrac,
  mapping = aes(
      x = distance,
      y = value,
      group = variable,
      color = method
    )
  ) +
  geom_line() +
  # facet_wrap(
  #     facets = ~ method + C_N
  #   ) +
  # guides(
  #     color = "none"
  #   ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "distance to DHSs center"
    ) +
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.8.] Plotting
```{r}
ggplot2::ggplot(
  data = ctss.cumfrac[ctss.cumfrac$distance==200,],
  mapping = aes(
      x = value,
      y = variable,
      fill = method
    )
  ) +
  geom_col() +
  scale_fill_manual(
      values = cols3
    ) +
  labs(
      x = "Cumulative unique CTSS fraction at 200bp",
      y = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.9.] Data wragnling for cumulative fraction
```{r}
regex = "^(.*)_(.*?)_(C|N|trip|ctrl|Cap_align)(\\d?)$"

ctss.cumfrac <- GM12878.rDHS.cumFrac %>%
  purrr::pluck(3) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )
```

## [3.10.] Plotting
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac,
    mapping = aes(
        x = distance,
        y = value,
        group = variable,
        color = method
      )
    ) +
  geom_line() +
  # facet_wrap(
  #     facets = . ~ method + C_N
  #   ) +
  # guides(
  #     color = "none"
  #   ) +
  labs(
      y = "CTSS to DHS cumulative fraction",
      x = "distance to DHSs center"
    ) + 
  scale_color_manual(
      values = cols3
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.11.] Plotting
```{r}
ggplot2::ggplot(
    data = ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = variable,
        x = value,
        fill = method
      )
    ) +
    geom_col() +
    scale_fill_manual(
        values = cols3
      ) +
    labs(
        x = "Cumulative DHS fraction at 200bp",
        y = "sample",
        fill = "sample"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1
      )
```
