---
title: "1.3Noise"
output: html_document
date: "2024-01-31"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
##Packages for plotting
base::library(epistack)
base::library(BiocParallel)
base::library(reshape2)
base::library(limma)
base::library(Matrix)
base::library(corrplot)
```

## [0.1.] Load CAGEfighR extensions
#### previous: base::setwd("/home/mhf817"), cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
```{r message=FALSE}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
cols4 <- base::c(
    "#FE8D02","#FF4425","#00439B",
    "#008F72","#878784","#7D62CF"
  )
```

## [0.4.] Set working directory
```{r eval=FALSE}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/1.CTSSs/")
```

## [0.5.] cumulativeFraction function
```{r}
cumulativeFractions <- function(loci, ctss, max_dist = 1000, inputAssay = "counts") {
  
  g <- GRanges(seqnames = seqnames(loci), loci$thick, strand = "*")
  t <- as(rowRanges(ctss), "GRanges")
  
  message("finding nearest loci...")
  g_nearest <- nearest(t, g)
  
  message("calculating distances...")
  d <- rep(Inf, length(t))
  non.na <- which(!is.na(g_nearest))
  d[non.na] <- distance(t[non.na], g[g_nearest[non.na]])
  
  
  message("calculating totalTags and unique Tags...")
  ctss <- suppressWarnings(calcTotalTags(ctss, inputAssay = "counts"))
  colData(ctss)$uniqTags <- colSums(assay(ctss,inputAssay="counts")>0)
  
  message("calculating cumulative fractions...")
  focus <- which(d <= max_dist)
  d <- d[focus]
  ctss <- ctss[focus]
  
  o <- order(d)
  d <- d[o]
  
  unique_loc_frac <- sapply(colnames(ctss), function(n) {
    cat(".")
    a <- as.numeric(assay(ctss, inputAssay)[o, n])
    a[a>0] <- 1
    nz <- which(a > 0)
    da <- d[nz]
    a_sub <- a[nz]
    sapply(seq(0, max_dist, by = 1), function(i) sum(a_sub[da == i]))
  })
  cat("\n")
  
  
  fracs <- sapply(colnames(ctss), function(n) {
    cat(".")
    a <- as.numeric(assay(ctss, inputAssay)[o, n])
    nz <- which(a > 0)
    da <- d[nz]
    a_sub <- a[nz]
    sapply(seq(0, max_dist, by = 1), function(i) sum(a_sub[da == i]))
  })
  cat("\n")
  
  t <- as(rowRanges(ctss), "GRanges")
  
  loc_fracs <- sapply(colnames(ctss), function(n) {
    cat(".")
    a <- as.numeric(assay(ctss, inputAssay)[o, n])
    nz <- t[which(a > 0)]
    dist <- distanceToNearest(g,nz)
    dist <- as.data.frame(dist)[,3]
    sapply(seq(0, max_dist, by = 1), function(i) sum(dist==i))
  })
  
  unique_loc_frac <- sapply(colnames(unique_loc_frac), function(n) cumsum(unique_loc_frac[, n] / ctss$uniqTags[n]))
  
  fracs <- sapply(colnames(fracs), function(n) cumsum(fracs[, n] / ctss$totalTags[n]))
  
  loc_fracs <- sapply(colnames(loc_fracs), function(n) cumsum(loc_fracs[, n] / length(g)))
  
  ucumfrac <- cbind(data.frame(distance = seq(0, max_dist, by = 1), unique_loc_frac))
  
  cumfrac <- cbind(data.frame(distance = seq(0, max_dist, by = 1), fracs))
  
  locifrac <- cbind(data.frame(distance = seq(0, max_dist, by = 1), loc_fracs))
  
  fraclist <- list(cumfrac,ucumfrac,locifrac)
}
```

# [1.] K562 data -------------------------

## [1.1.] Read in pre made CTSS objects
```{r}

k562.CTSSs <- readRDS("all.k562.CTSSs.rds")
k562.s.CTSSs <- readRDS("subsampled.k562.CTSSs.rds")
k562.p.CTSSs <- readRDS("pooled.k562.CTSSs.rds")
k562.p.s.CTSSs <- readRDS("pooled.subsampled.k562.CTSSs.rds")
```

## [1.2.] Read in DHS bed file form encode 
```{r}
k562.dhs <- read.table("../0.External_resources/E123-DNase.macs2.hg38.narrowPeak")
colnames(k562.dhs) <- c("chr","start","end")
k562.dhs <- GRanges(k562.dhs[,1:3])
k562.dhs$thick <- IRanges(start=start(k562.dhs)+((end(k562.dhs)-start(k562.dhs))/2))
```

## [1.3.] Data wrangling & plotting
```{r}
#Remove non standard chromsosomes for fair comparison
k562.CTSSs.stand <- keepStandardChromosomes(k562.s.CTSSs,pruning.mode = "coarse")
k562.dhs<- keepStandardChromosomes(k562.dhs,pruning.mode = "coarse")
cumFrac <- cumulativeFractions(k562.dhs,k562.CTSSs.stand) 

ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
```

```{r}
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")


p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
```

```{r}
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")

p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
```

```{r}
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
pdf("./plots/cumfrac.k562.subsampled.DHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

## [1.4.] Data wrangling & plotting
```{r}
#Remove non standard chromsosomes for fair comparison
k562.CTSSs.stand <- keepStandardChromosomes(k562.CTSSs,pruning.mode = "coarse")
k562.dhs<- keepStandardChromosomes(k562.dhs,pruning.mode = "coarse")
cumFrac <- cumulativeFractions(k562.dhs,k562.CTSSs.stand) 

ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
```

```{r}
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")
p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
```

```{r}
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
```

```{r}
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")
p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
pdf("./plots/cumfrac.k562.DHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

```{r}
k562.rdhs.cumFrac <- cumulativeFractions(rdhs,k562.s.CTSSs.stand) 


ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")


p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")


uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")

p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")


loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")


p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")



pdf("./plots/cumfrac.k562.subsampled.rDHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

```{r}
k562.rdhs.cumFrac <- cumulativeFractions(rdhs,k562.CTSSs.stand) 


ctss.cumfrac <- melt(cumFrac[[1]], id.var="distance")
ctss.cumfrac$method <- str_sub(ctss.cumfrac$variable,1,-2)
p1 <- ggplot(ctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() +  labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample") + scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed") + theme(legend.position="none")


p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")


uctss.cumfrac <- melt(cumFrac[[2]], id.var="distance")
uctss.cumfrac$method <- str_sub(uctss.cumfrac$variable,1,-2)
p2 <- ggplot(uctss.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw() + labs(y="unique CTSS to DHS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")

p2.2 <- ggplot(uctss.cumfrac[uctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")


loci.cumfrac <- melt(cumFrac[[3]], id.var="distance")
loci.cumfrac$method <- str_sub(loci.cumfrac$variable,1,-2)
p3 <- ggplot(loci.cumfrac,aes(distance,value, color=method, group=variable)) + geom_line() + theme_bw()  + labs(y="DHS to CTSS cumulative fraction", x= "distance to DHSs center", color="sample")+ scale_color_manual(values=custom.col) + geom_vline(xintercept=200,alpha=0.5,linetype="dashed")+ theme(legend.position="none")


p3.2 <- ggplot(loci.cumfrac[loci.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col) + labs(y="Cumulative unique DHS fraction at 200bp", x= "sample", fill="sample")



pdf("./plots/cumfrac.k562.rDHS.pdf", height=20, width = 20)
grid.arrange(p1,p1.2,p2,p2.2,p3,p3.2 ,ncol=2)
dev.off()

```

# [2.] GM12878 data -------------------------

## [2.0.] Read in pre-made CTSS objects
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.GM12878.CTSSs.rds"
    )
  )
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "subsampled.GM12878.CTSSs.rds"
    )
  )
GM12878.p.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.GM12878.CTSSs.rds"
    )
  )
GM12878.p.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "pooled.subsampled.GM12878.CTSSs.rds"
    )
  )
GM12878.pID.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "IDpooled.GM12878.CTSSs.rds"
    )
  )
GM12878.pID.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "IDpooled.subsampled.GM12878.CTSSs.rds"
    )
  )
```

## [2.1.] Read in DHSs & prepare as GRanges object
```{r}
GM12878.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E116-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(GM12878.dhs) + (
        (BiocGenerics::end(GM12878.dhs) - BiocGenerics::start(GM12878.dhs))/2
      )
  )
```

## [2.2.] Calculate cumulative fractions as a function of the distance to DHSs
#### 'gm12878.cumFrac <- cumulativeFractions(GM12878.dhs,GM12878.CTSSs.stand)'
```{r}
GM12878.CTSSs.stand <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.p.s.CTSSs, # GM12878.pID.s.CTSSs,
    pruning.mode = "coarse"
  )
GM12878.dhs <- GenomeInfoDb::keepStandardChromosomes(
    x = GM12878.dhs,
    pruning.mode = "coarse"
  )
gm12878.cumFrac <- .GlobalEnv$cumulativeFractions(
    loci = GM12878.dhs,
    ctss = GM12878.CTSSs.stand
  )

regex = "^(.*)_(.*?)_(C|N)$"# "^(.*)_(.*?)_(C|N)\\d$"

ctss.cumfrac <- gm12878.cumFrac %>%
  purrr::pluck(1) %>%
  tidyr::pivot_longer(
      cols = dplyr::contains("GM12878_"),
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    method = stringr::str_replace(
        variable, "^(.*)_(.*?)_(.*)$","\\2"
      ),
    C_N = stringr::str_replace(
        variable, regex,"\\3"
      ),
    type = base::paste0(C_N,"_",method)
  )

# ctss.cumfrac <- melt(gm12878.cumFrac[[1]], id.var="distance")
# ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
# ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)
# ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )
```

#### Evaluate
```{r}
dplyr::distinct(ctss.cumfrac,variable)
dplyr::distinct(ctss.cumfrac,method)
dplyr::distinct(ctss.cumfrac,type)
```

## [2.3.] Data wrangling - preparing plotting
```{r}
tps <- base::c(
    "C_500K","N_500K","C_1M",
    "N_1M","N_5M","N_10M"
  )
vrs <- base::c(
    "GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3",
    "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3",
    "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3",
    "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3",
    "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3",
    "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3"
  )

ctss.plot <- ctss.cumfrac %>%
  dplyr::filter(type %in% tps) %>%
  dplyr::mutate(
      type = base::factor(
          x = type,
          levels = base::rev(tps),
          ordered = TRUE
        ),
      variable = base::factor(
        x = variable,
        levels = base::rev(vrs),
        ordered = TRUE
      )
    )
```

## [2.3.] Plotting
#### p1
```{r}
  ggplot2::ggplot(
    data = ctss.plot,
    mapping = aes(
        x = distance,
        y = value,
        group = type, #variable,
        color = type
      )
    ) +
    geom_vline(
        xintercept = 200L,
        linetype = "dashed"
      ) +
    geom_line() +
    # facet_wrap(
    #     facets = ~ method #+ C_N
    #   ) +
    # guides(
    #     color = "none"
    #   ) +
    labs(
        y = "CTSS to DHS cumulative fraction",
        x = "Distance to DHSs center"
      ) +
    scale_color_manual(
        values = cols4
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        legend.title = element_blank(),
        legend.justification = base::c(1,0),
        legend.position = base::c(0.95,0.05),
        legend.background = element_rect(
          fill = "grey"
        )
      )
```

## [2.4.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          # "_replicates",
          "_pools",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.5.] Plotting
#### p1.2
```{r}
ctss.plot %>%
  dplyr::filter(
      distance == 200L
    ) %>%
ggplot2::ggplot(
    data = ., #ctss.cumfrac[ctss.cumfrac$distance==200,],
    mapping = aes(
        y = type, #variable,
        x = value,
        fill = type
      )
    ) +
  geom_col() +
  geom_text(
      mapping = aes(
        label = type
      ),
      x = 0,
      hjust = -1.1,
      color = "white",
      fontface = "bold"
    ) +
  scale_fill_manual(
      values = base::rev(cols4)
    ) +
  labs(
      y = "Cumulative CTSS fraction at 200bp",
      x = "sample",
      fill = "sample"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.text.y = element_blank()
    )
```

## [2.6.] Save plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878",
          "_cumFracDhs",
          "_replicates",
          # "_pools",
          "_dist200",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```


```{r}
ctss.cumfrac <- melt(gm12878.cumFrac[2], id.var="distance")

ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)

ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )
```

```{r}
p2 <-  ggplot(ctss.cumfrac,aes(distance,value, group=variable, color=method)) + geom_line() + theme_bw() + facet_wrap(~method+C_N) + guides(color="none") + labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center")+ 
scale_color_manual(values=custom.col)

p2.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")
```

```{r}
ctss.cumfrac <- melt(gm12878.cumFrac[3], id.var="distance")


ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)

ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )
```

```{r}
p3 <-  ggplot(ctss.cumfrac,aes(distance,value, group=variable, color=method)) + geom_line() + theme_bw() + facet_wrap(~method+C_N) + guides(color="none") + labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center")+ 
scale_color_manual(values=custom.col)

p3.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative DHS fraction at 200bp", x= "sample", fill="sample")


pdf("./plots/cumfrac.GM12878.pID.subsampled.DHS.pdf", height=30, width = 20)
grid.arrange(arrangeGrob(p1,p2,p3, ncol=1, nrow=3),
         arrangeGrob(p1.2,p2.2,p3.2, ncol=1, nrow=3), widths=c(2,1))
dev.off()
```





```{r}
#Read in rDHS bed file form encode 
rdhs <- read.table("/maps/projects/ralab/people/pcr980/DeepCompare/Raw_data/rDHS_ENCFF619EJE.bed")
colnames(rdhs) <- c("chr","start","end")
rdhs <- GRanges(rdhs[,1:3])
rdhs$thick <- IRanges(start=start(rdhs)+((end(rdhs)-start(rdhs))/2))
```


```{r}
rdhs <- keepStandardChromosomes(rdhs, pruning.mode="coarse")
gm12878.rDHS.cumFrac <- cumulativeFractions(rdhs,GM12878.CTSSs.stand)
```
```{r}
ctss.cumfrac <- melt(gm12878.rDHS.cumFrac[[1]], id.var="distance")

ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)

ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )

p1 <- ggplot(ctss.cumfrac,aes(distance,value, group=variable, color=method)) + geom_line() + theme_bw() + facet_wrap(~method+C_N) + guides(color="none") + labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center")+ scale_color_manual(values=custom.col)

p1.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative CTSS fraction at 200bp", x= "sample", fill="sample")


ctss.cumfrac <- melt(gm12878.rDHS.cumFrac[2], id.var="distance")

ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)

ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )

p2 <-  ggplot(ctss.cumfrac,aes(distance,value, group=variable, color=method)) + geom_line() + theme_bw() + facet_wrap(~method+C_N) + guides(color="none") + labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center")+ 
scale_color_manual(values=custom.col)

p2.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative unique CTSS fraction at 200bp", x= "sample", fill="sample")


ctss.cumfrac <- melt(gm12878.rDHS.cumFrac[3], id.var="distance")


ctss.cumfrac$method <- str_extract(ctss.cumfrac$variable, "_(.*?)_")
ctss.cumfrac$C_N <- str_sub(ctss.cumfrac$variable,-2,-2)

ctss.cumfrac$method <- gsub("_", "", ctss.cumfrac$method )

p3 <-  ggplot(ctss.cumfrac,aes(distance,value, group=variable, color=method)) + geom_line() + theme_bw() + facet_wrap(~method+C_N) + guides(color="none") + labs(y="CTSS to DHS cumulative fraction", x= "distance to DHSs center")+ 
scale_color_manual(values=custom.col)

p3.2 <- ggplot(ctss.cumfrac[ctss.cumfrac$distance==200,], aes(variable, value,fill=method)) + geom_col() + theme_bw() + scale_fill_manual(values=custom.col)+ labs(y="Cumulative DHS fraction at 200bp", x= "sample", fill="sample")


pdf("./plots/cumfrac.GM12878.pID.subsampled.rDHS.pdf", height=30, width = 20)
grid.arrange(arrangeGrob(p1,p2,p3, ncol=1, nrow=3),
         arrangeGrob(p1.2,p2.2,p3.2, ncol=1, nrow=3), widths=c(2,1))
dev.off()
```










