---
title: "1.4.2.CTSSs.annFracs.Rmd"
output: html_document
date: "2024-02-29"
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(BiocParallel)
```

## [0.1.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
custom.col.ord <- base::c(
    "#FFDB6D","#F4EDCA","#52854C",
    "#4E84C4","#293352","#C3D7A4",
    "#D16103","grey","#C4961A"
  )
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```

## [0.4.] Add additional functions
```{r}
calcAnnoCTSS <- function(data, txModels, uniqueCTSS = FALSE) {
    
    annos <- CAGEfightR::assignTxType(
        object = data,
        txModels = txModels
      )
    
    #Identify all txTypes
    types <- base::unique(SummarizedExperiment::rowData(annos)$txType)
    
    # Initialize list to store total tags for each txType
    totalTags_list <- base::vector("list", length = base::length(types))
    
    # Loop through each txType
    for (i in base::seq_along(types)) {
      
        # Subset data based on current txType
        current_subset <- S4Vectors::subset(annos, txType %in% types[i])
        if(uniqeCTSS){
          
            # Calculate totalTags for current subset
            current_subset <- Matrix::colSums(
                SummarizedExperiment::assay(current_subset) > 0
              ) 
            totalTags_list[[i]] <- current_subset
            
          } else {
            
            # Calculate totalTags for current subset
            current_subset <- SummarizedExperiment::colData(
                CAGEfightR::calcTotalTags(current_subset)
              )
            totalTags_list[[i]] <- current_subset$totalTags
            
          }
      }
    
    # Combine totalTags into a data frame
    ann_counts <- base::as.data.frame(
        base::do.call(base::cbind, totalTags_list)
      )
    base::colnames(ann_counts) <- types
    return(ann_counts)
  }
```

## To be discarded
```{r}
# calc_anno_CTSS <- function(data, txModels) {
#     
#     data <- CAGEfightR::assignTxType(
#         object = data,
#         txModels = txModels
#       )
#     
#     #Identify all txTypes
#     types <- base::unique(SummarizedExperiment::rowData(data)$txType)
#     
#     # Initialize list to store total tags for each txType
#     totalTags_list <- base::vector("list", length = base::length(types))
#     
#     # Loop through each txType
#     for (i in base::seq_along(types)) {
#       
#         # Subset data based on current txType
#         current_subset <- IRanges::subset(data, txType %in% types[i])
#             
#         # Calculate totalTags for current subset
#         current_subset <- SummarizedExperiment::colData(
#             CAGEfightR::calcTotalTags(current_subset)
#           )
#         totalTags_list[[i]] <- current_subset$totalTags
#       }
#     
#     # Combine totalTags into a data frame
#     ann_counts <- base::as.data.frame(
#         base::do.call(base::cbind, totalTags_list)
#       )
#     base::colnames(ann_counts) <- types
#     return(ann_counts)
#   }
```

## To be discarded
```{r}
# calc_anno_uniqCTSS <- function(data, txModels) {
#   
#   data <- CAGEfightR::assignTxType(
#         object = data,
#         txModels = txModels
#       )
#     
#     #Identify all txTypes
#     types <- base::unique(SummarizedExperiment::rowData(data)$txType)
#     
#     # Initialize list to store total tags for each txType
#     totalTags_list <- base::vector("list", length = base::length(types))
#     
#     # Loop through each txType
#     for (i in seq_along(Types)) {
#       
#         # Subset data based on current txType
#         current_subset <- IRanges::subset(data, txType %in% types[i])
#         
#         # Calculate totalTags for current subset
#         current_subset <- Matrix::colSums(
#             SummarizedExperiment::assay(current_subset) > 0
#           ) 
#         totalTags_list[[i]] <- current_subset
#       }
#     
#     # Combine totalTags into a data frame
#     ann_counts <- base::as.data.frame(
#         base::do.call(base::cbind, totalTags_list)
#       )
#     base::colnames(ann_counts) <- types
#     return(ann_counts)
#   }
```

# [1.] K562 analysis -------------------------

## [1.0.] Read in datasets for K562 & GM12878
#### see [1.10.] in "1.1.CTSSs.processing.Rmd"
```{r}
K562.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
# K562.p.CTSSs <- base::readRDS(
#     file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.k562.CTSSs.rmSingletons.rds"
#     )
#   )
```

#### Factor levels for genomic features
```{r}
feats <- base::c(
    "intergenic","antisense","proximal",
    "intron","exon","promoter",
    "fiveUTR","CDS","threeUTR"
  )
```

## [1.0.] Quantify CTSS level annotations
```{r}
k562.ann.counts <- .GlobalEnv$calc_anno_CTSS(
    data = k562.s.CTSSs,
    txModels = txdb
  )
k562.ann.fracs <- k562.ann.counts/SummarizedExperiment::colData(k562.s.CTSSs)$totalTags
```

## Data wrangling - prepare plotting
```{r}
k562.ann.fracs.m <- k562.ann.fracs %>%
  tibble::rownames_to_column("sample") %>%
  dplyr::mutate(
      sample = dplyr::case_when(
        sample == "K562_N_10M" ~ "K562_10M_N",
        TRUE ~ sample
      ),
    ) %>%
  dplyr::filter(
      sample != "K562_Cap_align"
    ) %>% 
  tidyr::pivot_longer(
      cols = !sample,
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    variable = base::factor(
      x = variable,
      levels = feats,
      ordered = TRUE
    )
  )
```

#### Previous code chunk
```{r}
# k562.ann.counts <- calc_anno_CTSS(k562.s.CTSSs,txModels = txdb)
# k562.ann.fracs <- k562.ann.counts/colData(k562.s.CTSSs)$totalTags
# 
# k562.ann.uniqcounts <- calc_anno_uniqCTSS(k562.s.CTSSs,txModels = txdb)
# k562.ann.uniqfracs <- k562.ann.uniqcounts/rowSums(k562.ann.uniqcounts)
# 
# k562.ann.counts.p <- calc_anno_CTSS(k562.s.CTSSs,txModels = txdb)
# k562.ann.fracs.p <- k562.ann.counts/colData(k562.s.CTSSs)$totalTags
# 
# k562.ann.uniqcounts.p <- calc_anno_uniqCTSS(k562.p.CTSSs,txModels = txdb)
# k562.ann.uniqfracs.p <- k562.ann.uniqcounts.p/rowSums(k562.ann.uniqcounts.p)
```

## [1.1.] Save & laod K562 data objects
```{r}
base::save(
    k562.ann.counts,
    k562.ann.fracs,
    k562.ann.uniqcounts,
    k562.ann.uniqfracs,
    k562.ann.counts,
    k562.ann.fracs.p,
    k562.ann.uniqcounts.p,
    k562.ann.uniqfracs.p,
    file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_annFracs.k562.RData"
    )
  )
```

```{r}
base::load(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "annFracs.k562.RData"
    )
  )
```

## [1.2.] Plotting
```{r}
ggplot2::ggplot(
    data = k562.ann.fracs.m,
    mapping = aes(
        x = value,
        y = sample,
        fill = variable
      )
    ) +
  geom_col() +
  scale_fill_manual(
      values = custom.col
    ) +
  theme_bw() +
  theme(
      axis.title = element_blank()
    )
```

```{r}
ggplot2::ggplot(
    data = k562.ann.fracs,
    mapping = aes(
        x = intergenic,
        y = intron,
        color = sample
      )
    ) +
  geom_point(
      size = 2
    ) +
  theme_bw() +
  geom_text(
      label = rownames(k562.ann.fracs), 
      check_overlap = F,
      size = 3,
      nudge_y = 0.003
    ) + 
  guides(color="none") +
  scale_color_manual(
      values = k.full.pal
    )
```

```{r}
ggplot2::ggplot(
  data = k562.ann.fracs,
  mapping = aes(
        x = exon,
        y = intron,
        color = sample
      )
    ) +
  geom_point(
      size = 2
    ) +
  theme_bw() +
  geom_text(
      label = rownames(k562.ann.fracs),
      check_overlap = F,
      size = 3,
      nudge_y = 0.003
    ) + 
  guides(color="none") +
  scale_color_manual(
      values = k.full.pal
    )

pdf("./combined_plots/annFracs.k562.repSub.pdf", height=20, width = 10)
grid.arrange(p1,p2,p3)
dev.off()
```


```{r}
k562.ann.fracs.p$sample <- row.names(k562.ann.fracs.p)

k562.ann.fracs.p$sample[k562.ann.fracs.p$sample=="K562_N_10M"] <- "K562_10M_N"
k562.ann.fracs.p<- k562.ann.fracs.p[k562.ann.fracs.p$sample!="K562_Cap_align",]

k562.ann.fracs.m <- melt(k562.ann.fracs)

k562.ann.fracs.p.m <- melt(k562.ann.fracs.p)
```

## Plotting
```{r}
p1 <- ggplot(k562.ann.fracs.p.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(k562.ann.fracs.p, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.fracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)

p3 <- ggplot(k562.ann.fracs.p, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.fracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)

pdf("./combined_plots/annFracs.k562.poolSub.pdf", height=20, width = 10)
grid.arrange(p1,p2,p3)
dev.off()
```

## Unique CTSS fraction per annotation
```{r}
k562.ann.uniqfracs$sample <- row.names(k562.ann.uniqfracs)

k562.ann.uniqfracs$sample[k562.ann.uniqfracs$sample=="K562_N_10M"] <- "K562_10M_N"
k562.ann.uniqfracs<- k562.ann.uniqfracs[k562.ann.uniqfracs$sample!="K562_Cap_align",]

k562.ann.uniqfracs.m <- melt(k562.ann.uniqfracs)
```

## Plotting
```{r}
p1 <- ggplot(k562.ann.uniqfracs.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(k562.ann.uniqfracs, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.uniqfracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)

p3 <- ggplot(k562.ann.uniqfracs, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.uniqfracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)


pdf("./combined_plots/annuniqFracs.k562.repSub.pdf", height=20, width = 10)
grid.arrange(p1,p2,p3)
dev.off()
```

```{r}
k562.ann.uniqfracs.p$sample <- row.names(k562.ann.uniqfracs.p)

k562.ann.uniqfracs.p$sample[k562.ann.uniqfracs.p$sample=="K562_N_10M"] <- "K562_10M_N"
k562.ann.uniqfracs.p<- k562.ann.uniqfracs.p[k562.ann.uniqfracs.p$sample!="K562_Cap_align",]

k562.ann.uniqfracs.p.m <- melt(k562.ann.uniqfracs.p)
```

## Plotting
```{r}
p1 <- ggplot(k562.ann.uniqfracs.p.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(k562.ann.uniqfracs.p, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.uniqfracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)

p3 <- ggplot(k562.ann.uniqfracs.p, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(k562.ann.uniqfracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=k.full.pal)

pdf("./combined_plots/annuniqFracs.k562.poolSub.pdf", height=20, width = 10)
grid.arrange(p1,p2,p3)
dev.off()

```


# [2.] GM12878 analysis -------------------------

## [2.0.] Read in datasets for GM12878
#### see [1.10.] in "1.1.CTSSs.processing.Rmd"
```{r}
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
GM12878.p.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```

## [2.1.] Quantify CTSS level annotations
```{r}
subset <- SummarizedExperiment::colData(GM12878.s.CTSSs) %>%
  base::as.data.frame() %>%
  tibble::rownames_to_column(var = "sample") %>%
  dplyr::filter(main == "yes")
samples <- dplyr::pull(subset,sample)

GM12878.ann.counts <- .GlobalEnv$calcAnnoCTSS(
    data = GM12878.s.CTSSs[,samples],
    txModels = txdb,
    uniqueCTSS = FALSE # TRUE
  )

tags <- dplyr::pull(subset,totalTags)
GM12878.ann.fracs <- GM12878.ann.counts/tags
```

```{r}
subset <- SummarizedExperiment::colData(GM12878.p.CTSSs) %>%
  base::as.data.frame() %>%
  tibble::rownames_to_column(var = "sample") %>%
  dplyr::filter(main == "yes")
samples <- dplyr::pull(subset,sample)

GM12878.ann.counts <- .GlobalEnv$calcAnnoCTSS(
    data = GM12878.p.CTSSs[,samples],
    txModels = txdb,
    uniqueCTSS = FALSE # TRUE
  )

tags <- dplyr::pull(subset,totalTags)
GM12878.ann.fracs <- GM12878.ann.counts/tags
```

## [2.2.] Factor levels for genomic features & cell amounts
```{r}
feats <- base::c(
    "intergenic","proximal","promoter",    
    "fiveUTR","CDS","exon",
    "intron","threeUTR","antisense"
  )
amts <- base::c("500K","1M","5M","10M")
reps <- base::c("pool","1","2","3")

# sams <- base::c(
#     "GM12878_500K_C","GM12878_500K_N",
#     "GM12878_1M_C","GM12878_1M_N",
#     "GM12878_5M_N","GM12878_10M_N"
#   )
```

## [2.3.] Data wrangling - prepare plotting of fractions
#### Previous code: gm12878.ann.fracs$sample <- row.names(gm12878.ann.fracs); gm12878.ann.fracs.m <- melt(gm12878.ann.fracs)
```{r}
GM12878.ann.fracs.m <- GM12878.ann.fracs %>%
  tibble::rownames_to_column("sample") %>%
  tidyr::pivot_longer(
      cols = !sample,
      names_to = "variable",
      values_to = "value"
    ) %>%
  tidyr::separate_wider_delim(
      cols = sample,
      delim = "_",
      names = base::c("cell","amount","type"),
      cols_remove = FALSE
    ) %>%
  tidyr::separate(
      col = type,
      sep = 1,
      into = base::c("source","rep"),
      convert = TRUE ## Creates NAs
    ) %>%
  dplyr::mutate(
    rep = dplyr::case_when(
        base::is.na(rep) ~ "pool",
        TRUE ~ base::as.character(rep)
      ),
    rep = base::factor(
        x = rep,
        levels = base::rev(reps),
        ordered = TRUE
      ),
    amount = base::factor(
        x = amount,
        levels = amts,
        ordered = TRUE
      ),
    source = base::factor(
        x = source,
        levels = base::c("C","N"),
        ordered = TRUE
      ),
    variable = base::factor(
        x = variable,
        levels = base::rev(feats),
        ordered = TRUE
      )
    )
```

## [2.4.] Save & load data
```{r}
base::saveRDS(
  object = GM12878.ann.counts,
  file = base::paste0(
      here::here("1.CTSSs/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.ann.counts",
      ".rds"
    )
  )
```

```{r}
GM12878.ann.counts <- base::readRDS(
  file = base::list.files(
        path = here::here("1.CTSSs"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.ann.counts")
  )
```

## [2.5.] Plotting of fractions
```{r}
ggplot2::ggplot(
    data = GM12878.ann.fracs.m,
    mapping = aes(
        x = value,
        y = rep,
        fill = variable
      )
    ) +
  geom_col() +
  facet_wrap(
      facets = source ~ amount,
      ncol = 1,
      strip.position = "right",
      axis.labels = "margins",
      axes = "margins"
    ) +
  theme_bw() +
  scale_fill_manual(
      values = base::rev(custom.col.ord) #base::rev(custom.col)
    ) +
  theme(
      axis.title = element_blank(),
      panel.spacing = grid::unit(0.1,"lines"),
      aspect.ratio = 1/6,
      # legend.position = "none"
      legend.title = element_blank(),
      legend.background = element_blank()
    )
```

## [2.6.] Save complexity plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878.ann.fracs.m",
          "_allReplicatesPool",
          "_reordered",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 7,
      height = 5
    )
  }
```

## [2.7.] Plotting of fractions for pooled samples only
```{r}
GM12878.ann.fracs.m %>% 
  dplyr::filter(rep == "pool") %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(sams),
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          x = value,
          y = sample,
          fill = variable
        )
      ) +
    geom_col() +
    scale_fill_manual(
        values = base::rev(custom.col.ord) #base::rev(custom.col)
      ) +
    theme_bw() +
    theme(
        axis.title = element_blank(),
        panel.spacing = grid::unit(0.1,"lines"),
        aspect.ratio = 1/6,
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.position = "bottom"
      )
```

## [2.8.] Save complexity plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("1.CTSSs/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878.ann.fracs.m",
          "_pool",
          "_reordered",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [2.9.] Quantify CTSS level annotations
```{r}
GM12878.ann.uniqcounts <- .GlobalEnv$calcAnnoCTSS(
    data = GM12878.s.CTSSs[,samples],
    txModels = txdb,
    uniqueCTSS = TRUE
  )

GM12878.ann.uniqfracs <- GM12878.ann.uniqcounts/base::rowSums(GM12878.ann.uniqcounts)

gm12878.ann.uniqfracs$sample <- row.names(gm12878.ann.uniqfracs)
gm12878.ann.uniqfracs.m <- melt(gm12878.ann.uniqfracs)


GM12878.ann.uniqfracs.m <- GM12878.ann.uniqfracs %>%
  tibble::rownames_to_column("sample") %>%
  tidyr::pivot_longer(
      cols = !sample,
      names_to = "variable",
      values_to = "value"
    ) %>%
  tidyr::separate_wider_delim(
      cols = sample,
      delim = "_",
      names = base::c("cell","amount","type"),
      cols_remove = FALSE
    ) %>%
  tidyr::separate(
      col = type,
      sep = 1,
      into = base::c("source","rep"),
      convert = TRUE ## Creates NAs
    ) %>%
  dplyr::mutate(
    rep = dplyr::case_when(
        base::is.na(rep) ~ "pool",
        TRUE ~ base::as.character(rep)
      ),
    rep = base::factor(
        x = rep,
        levels = base::rev(reps),
        ordered = TRUE
      ),
    amount = base::factor(
        x = amount,
        levels = amts,
        ordered = TRUE
      ),
    source = base::factor(
        x = source,
        levels = base::c("C","N"),
        ordered = TRUE
      ),
    variable = base::factor(
        x = variable,
        levels = base::rev(feats),
        ordered = TRUE
      )
    )
```


```{r}
gm12878.ann.counts.p <- .GlobalEnv$calc_anno_CTSS(
    GM12878.p.CTSSs[,colData(GM12878.p.CTSSs)$main=="yes"],
    txModels = txdb
  )
gm12878.ann.fracs.p <- gm12878.ann.counts.p/colData(GM12878.p.CTSSs)$totalTags[colData(GM12878.p.CTSSs)$main=="yes"]
gm12878.ann.fracs.p$sample <- row.names(gm12878.ann.fracs.p)
gm12878.ann.fracs.p.m <- melt(gm12878.ann.fracs.p)

gm12878.ann.uniqcounts.p <- .GlobalEnv$calc_anno_uniqCTSS(
    GM12878.p.CTSSs[,colData(GM12878.p.CTSSs)$main=="yes"],
    txModels = txdb
  )
gm12878.ann.uniqfracs.p <- gm12878.ann.uniqcounts.p/rowSums(gm12878.ann.uniqcounts.p)
gm12878.ann.uniqfracs.p$sample <- row.names(gm12878.ann.uniqfracs.p)
gm12878.ann.uniqfracs.p.m <- melt(gm12878.ann.uniqfracs.p)
```

#### /combined_plots/annFracs.GM12878.repSub.pdf
#### See chunk in [1.5.] for previous plot part "p1"
```{r}
p2 <- ggplot(gm12878.ann.fracs, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.fracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)

p3 <- ggplot(gm12878.ann.fracs, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.fracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)
```

#### /combined_plots/annFracs.GM12878.poolSub.pdf
```{r}
p1 <- ggplot(gm12878.ann.fracs.p.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(gm12878.ann.fracs.p, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.fracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)

p3 <- ggplot(gm12878.ann.fracs.p, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.fracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)
```

#### /combined_plots/annuniqFracs.GM12878.repSub.pdf
```{r}
p1 <- ggplot(gm12878.ann.uniqfracs.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(gm12878.ann.uniqfracs, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.uniqfracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)

p3 <- ggplot(gm12878.ann.uniqfracs, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.uniqfracs), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)
```

#### /combined_plots/annuniqFracs.GM12878.poolSub.pdf
```{r}
p1 <- ggplot(gm12878.ann.uniqfracs.p.m, aes(sample, value, fill=variable)) + geom_col() + coord_flip() +
  theme_bw() + scale_fill_manual(values=custom.col)

p2 <- ggplot(gm12878.ann.uniqfracs.p, aes(intergenic, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.uniqfracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)

p3 <- ggplot(gm12878.ann.uniqfracs.p, aes(exon, intron, color=sample)) +
  geom_point(size=2) + theme_bw() + geom_text( label=rownames(gm12878.ann.uniqfracs.p), 
    check_overlap = F, size=3,nudge_y = 0.003) + 
  guides(color="none") + scale_color_manual(values=gm.full.pal)
```

