---
title: "3.1. GM12878 divergent loci calling"
output: html_document
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")

base::library("BiocParallel")
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 30
    )
  )
```

## [0.2.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D", "#C4961A", "#F4EDCA",
    "#D16103", "#C3D7A4", "#52854C",
    "#4E84C4", "#293352","grey"
  )
```

## This is part of PRIME, no?

## [0.5.] Helper function mergeLoci() & divergentLociTCsSummit()
```{r}

```

```{r}

```
# [1.] Calling of divergent locis -------------------------

#### Annotation of txType and overlaps with atac-seq regions

## [1.0.] Load CTSSs
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```

## [1.1.] Run divergentLociTCsSummit() across chosen samples
#### "GM12878.DLs <- lapply(colnames(GM12878.CTSSs),function(x) divergentLociTCsSummit(GM12878.CTSSs[,x],callingAssay="counts"))"
#### Evaluate with 's = "GM12878_500K_C"', "ctss = CTSSs.by.chr[["chrX"]]"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

## Check that all selected samples are present in GM12878.CTSSs
assertthat::are_equal(
    main_samples,
    base::colnames(GM12878.CTSSs)[base::colnames(GM12878.CTSSs) %in% main_samples]
  )
```

```{r}
GM12878.DLs.rmS <- base::lapply(
  X = main_samples[1],
  FUN = function(s) {
    base::message("Currently processed sample:\t",s)
    ## Split sample into chromosomes
    CTSSs.by.chr <- base::split(
        x = GM12878.CTSSs[,s],
        f = GenomeInfoDb::seqnames(GM12878.CTSSs[,s])
      )
    ## Run divergentLociTCsSummit() across chromosomes
    DLs.by.chr <- base::lapply(
      X = base::names(CTSSs.by.chr),
      FUN = function(chr) {
        base::message("\tCurrently processed chromosome:\t",chr)
        .GlobalEnv$divergentLociTCsSummit(
            ctss = CTSSs.by.chr[[chr]],
            callingAssay = "counts.noSingletons"
          )
        }
      )
    return(
      ## Merge information for all chromosomes of one sample
      base::suppressWarnings(
        base::unlist(
            methods::as(DLs.by.chr,"GRangesList")
          )
        )
      )
    }
  )
```

## [1.2.] Save output
#### "saveRDS(GM12878.DLs, "poolSub.GM12878.summitDLs.rds")"
```{r}
# base::saveRDS(
#     object = GM12878.DLs.rmS,
#     file = base::paste0(
#       here::here("3.DivergentLoci/"),
#       "poolSub.GM12878.summitDLs.rmSingletons.rds"
#     )
#   )
```
