---
title: "3.1.5.GM12878.summitDLs.Rmd"
output: html_document
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")

base::library("BiocParallel")
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 30
    )
  )
```

## [0.2.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D", "#C4961A", "#F4EDCA",
    "#D16103", "#C3D7A4", "#52854C",
    "#4E84C4", "#293352","grey"
  )
```

## [0.5.] Helper function mergeLoci() & divergentLociTCsSummit()
```{r}
mergeLoci <- function(m, p) {
    m.i <- 1
    p.i <- 1
    while ((p[p.i,start] < m[m.i,start]) || ((p[p.i,start] - m[m.i,start]) > max_gap)) {
        if ((p[p.i,score] < m[m.i,score]) && (p.i < base::nrow(p)) && ((p[p.i+1,start] - m[m.i,start]) < max_gap)) {
            p.i <- p.i + 1
          } else {
            m.i <- m.i + 1
          }
      }
    base::floor(m[m.i,start] + ((p[p.i,start] - m[m.i,start]) / 2))
  }
```

```{r}
divergentLociTCsSummit<- function(ctss, max_gap = 400, win_size = 200, inputAssay = "counts", callingAssay = "counts") {
    
    base::message(base::paste("Running DL calling for",base::colnames(ctss)))
    
    ## Pool & subset
    ctss.o <- ctss
    ctss <- base::suppressWarnings(
        CAGEfightR::calcPooled(ctss, inputAssay = callingAssay)
      )
    ctss <- base::subset(ctss, score > 0)
    
    #Call TCs for each sample
    object <- CAGEfightR::clusterUnidirectionally(ctss)
    object <- CAGEfightR::quantifyClusters(ctss,object)
    object <- base::suppressWarnings(
        CAGEfightR::calcPooled(object, inputAssay = inputAssay)
      )
    object <- SummarizedExperiment::rowRanges(object) 
    
    base::message("Removing overlapping TCs by strand...")
    object <- CAGEfightR::swapRanges(object) #Summit focused
    TCsByStrand <- CAGEfightR:::splitByStrand(object)
    
    ## Find overlapping and book-ended summits between strands
    olaps <- IRanges::findOverlaps(
        TCsByStrand$'-',
        TCsByStrand$'+',
        maxgap = 0,
        type = "any",
        select = "all",
        ignore.strand = TRUE
      )
    m_score <-  S4Vectors::mcols(TCsByStrand$'-')$score
    p_score <-  S4Vectors::mcols(TCsByStrand$'+')$score
    
    m_rem <- S4Vectors::queryHits(olaps)[
        base::which(m_score[S4Vectors::queryHits(olaps)] <= p_score[S4Vectors::subjectHits(olaps)])
      ]
    p_rem <- S4Vectors::subjectHits(olaps)[
        base::which(p_score[S4Vectors::subjectHits(olaps)] < m_score[S4Vectors::queryHits(olaps)])
      ]
    
    ## remove overlapping TCs
    if (base::length(m_rem) > 0) {
        TCsByStrand$'-' <- TCsByStrand$'-'[-m_rem]
      }
    if (base::length(p_rem) > 0) {
        TCsByStrand$'+' <- TCsByStrand$'+'[-p_rem]
      }
    
    base::message("Finding divergent TC pairs...")
    ## Find divergent TC pairs
    m_pad <- IRanges::flank(
        TCsByStrand$'-',
        width = max_gap,
        start = TRUE,
        both = FALSE
      )
    pairs <- IRanges::findOverlaps(
        m_pad,
        TCsByStrand$'+',
        maxgap = -1,
        type = "any",
        select = "all",
        ignore.strand = TRUE
      )
    
    ## Find connected components of TC pair graphs
    edge_list <- base::cbind(
        base::names(TCsByStrand$'-')[S4Vectors::queryHits(pairs)],
        base::names(TCsByStrand$'+')[S4Vectors::subjectHits(pairs)]
      )
    g <- igraph::graph_from_edgelist(edge_list, directed = FALSE)
    con <- igraph::components(g)
    
    ## Keep only relevant TCs
    object <- object[base::names(con$membership)]
    
    base::message("Merging into divergent loci...")
    covByStrand <- CAGEfightR:::splitPooled(
        methods::as(SummarizedExperiment::rowRanges(ctss),"GRanges")
      )
    
    dt <- data.table::data.table(
        strand = base::as.character(BiocGenerics::strand(object)),
        start = IRanges::start(object),
        score = object$score,
        locus = con$membership,
        stringsAsFactors = FALSE
      )
    
    dt <- base::split(dt, dt$strand)
    data.table::setorder(dt$'-', -score, start) ## order by score, then by position (prioritize rightmost)
    data.table::setorder(dt$'+', -score, -start) ## order by score, then by position (prioritize leftmost)
    
    loci.m <- base::split(dt$'-', dt$'-'$'locus')
    loci.p <- base::split(dt$'+', dt$'+'$'locus')
    
   mergeLoci <- function(m, p) {
      m.i <- 1
      p.i <- 1
      while ((p[p.i,start] < m[m.i,start]) || ((p[p.i,start] - m[m.i,start]) > max_gap)) {
        if ((p[p.i,score] < m[m.i,score]) && (p.i < nrow(p)) && ((p[p.i+1,start] - m[m.i,start]) < max_gap)) {
          p.i <- p.i + 1
        } else {
          m.i <- m.i + 1
        }
      }
      base::floor(m[m.i,start] + ((p[p.i,start] - m[m.i,start]) / 2))
    }
    
    mid <- BiocParallel::bplapply(
        X = 1:length(loci.m),
        FUN = function(i) mergeLoci(loci.m[[i]], loci.p[[i]])
      )
    mid <- base::unlist(mid)
    
    ## Extract seqnames for loci
    div_chr <- con$membership[S4Vectors::match(1:con$no, con$membership)]
    div_chr <- base::as.character(
      base::sapply(
          X = base::names(div_chr),
          FUN = function(n) {
            base::strsplit(n,":")[[1]][1]
          }
        )
      )
    #Revert to counts
    ctss <- ctss.o
    ctss <- CAGEfightR::calcPooled(ctss, inputAssay = inputAssay)
    ctss <- base::subset(ctss, score > 0)
    
    gr <- GenomicRanges::GRanges(
      seqnames = div_chr,
      IRanges::IRanges(
          start = mid,
          end = mid
        )
      )
    GenomeInfoDb::seqlevels(ctss, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(gr)
    GenomeInfoDb::seqinfo(ctss) <- GenomeInfoDb::seqinfo(gr)
    
    base::cat("\r")
    base::message("Calculating directionality...")
    
    win_1 <- IRanges::flank(
        x = gr,
        width = win_size,
        start = TRUE,
        both = FALSE
      )
    win_2 <- IRanges::flank(
        x = gr,
        width = win_size,
        start = FALSE,
        both = FALSE
      )
    
    covByStrand <- CAGEfightR:::splitPooled(methods::as(rowRanges(ctss),"GRanges"))
    
    ## Quantify strand-wise in flanking windows around midpoint
    M1 <- base::unlist(IRanges::viewSums(IRanges::Views(covByStrand$`-`, win_1)))
    P2 <- base::unlist(IRanges::viewSums(IRanges::Views(covByStrand$`+`, win_2)))
    M2 <- base::unlist(IRanges::viewSums(IRanges::Views(covByStrand$`-`, win_2)))
    P1 <- base::unlist(IRanges::viewSums(IRanges::Views(covByStrand$`+`, win_1)))
    
    ## Calculate directionality
    pooled_directionality <- (P2-M1) / (P2+M1)
    
    ## Test if divergent
    divergent <- (M1>P1) & (P2>M2)
    
    base::message("Calculating coverage across samples...")
    
    ## Quantify strand-wise in flanking windows around midpoint
    BiocGenerics::strand(win_1) <- "-"
    BiocGenerics::strand(win_2) <- "+"
    mat_2_plus <- base::suppressMessages(
      SummarizedExperiment::assay(
          CAGEfightR::quantifyClusters(
              object = ctss,
              clusters = win_2,
              inputAssay = inputAssay
            ),
          i = inputAssay
        ) > 0
      )
    mat_1_minus <- base::suppressMessages(
      SummarizedExperiment::assay(
          CAGEfightR::quantifyClusters(
              object = ctss,
              clusters = win_1,
              inputAssay = inputAssay
            ),
          i = inputAssay
        ) > 0
      )
    
    ## Quntify number of bidirectional cases (both strands expressed)
    bidirectional <- base::rowSums(mat_1_minus & mat_2_plus)
    
    base::message("Preparing output...")
    
    ## Build GRanges object
    IRanges::start(gr) <- IRanges::start(gr) - win_size
    IRanges::end(gr) <- IRanges::end(gr) + win_size
    gr$score <- M1 + P2
    gr$thick <- IRanges::IRanges(start = mid, width = 1)
    
    S4Vectors::mcols(gr)[, "directionality"] <- pooled_directionality
    S4Vectors::mcols(gr)[, "bidirectionality"] <- bidirectional
    S4Vectors::mcols(gr)[, "divergent"] <- divergent
    
    ids <- base::paste0(
        GenomeInfoDb::seqnames(gr), ":", IRanges::start(gr), "-", IRanges::end(gr)
      )
    base::names(gr) <- ids
    base::names(gr$thick) <- ids
    
    gr$sample <- base::colnames(ctss)
    
    return(gr)
  }
```
# [1.] Calling of divergent locis -------------------------

#### Annotation of txType and overlaps with atac-seq regions

## [1.0.] Load CTSSs
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```

## [1.1.] Run divergentLociTCsSummit() across chosen samples
#### "GM12878.DLs <- lapply(colnames(GM12878.CTSSs),function(x) divergentLociTCsSummit(GM12878.CTSSs[,x],callingAssay="counts"))"
#### Evaluate with 's = "GM12878_500K_C"', "ctss = CTSSs.by.chr[["chrX"]]"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

## Check that all selected samples are present in GM12878.CTSSs
assertthat::are_equal(
    main_samples,
    base::colnames(GM12878.CTSSs)[base::colnames(GM12878.CTSSs) %in% main_samples]
  )
```

```{r}
GM12878.DLs.rmS <- base::lapply(
  X = main_samples[1],
  FUN = function(s) {
    base::message("Currently processed sample:\t",s)
    ## Split sample into chromosomes
    CTSSs.by.chr <- base::split(
        x = GM12878.CTSSs[,s],
        f = GenomeInfoDb::seqnames(GM12878.CTSSs[,s])
      )
    ## Run divergentLociTCsSummit() across chromosomes
    DLs.by.chr <- base::lapply(
      X = base::names(CTSSs.by.chr),
      FUN = function(chr) {
        base::message("\tCurrently processed chromosome:\t",chr)
        .GlobalEnv$divergentLociTCsSummit(
            ctss = CTSSs.by.chr[[chr]],
            callingAssay = "counts.noSingletons"
          )
        }
      )
    return(
      ## Merge information for all chromosomes of one sample
      base::suppressWarnings(
        base::unlist(
            methods::as(DLs.by.chr,"GRangesList")
          )
        )
      )
    }
  )
```

## [1.2.] Save output
#### "saveRDS(GM12878.DLs, "poolSub.GM12878.summitDLs.rds")"
```{r}
# base::saveRDS(
#     object = GM12878.DLs.rmS,
#     file = base::paste0(
#       here::here("3.DivergentLoci/"),
#       "poolSub.GM12878.summitDLs.rmSingletons.rds"
#     )
#   )
```
