---
title: "3.2.2.DL cumulative fractions vs expression"
output: html_document
date: "2024-03-01"
---

# [-1.] Markdown is an adaptation of the version from 2024-02-13
# title: '3.2.DL cumulative fractions vs expression'
# output: html_document
# date: "2024-02-13"

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")
base::library("Seurat")

# base::library("BiocParallel")
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 30
#     )
#   )
```

```{r}
# library(TxDb.Hsapiens.UCSC.hg38.knownGene)
# txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
# 
# #Packages for plotting
# library(CAGEfightR)
# library(epistack)
# library(BiocParallel)
# library(dplyr)
# library(reshape2)
# library(limma)
# library(ggplot2)
# library(stringr)
```

## [0.2.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.3.] Load annotations
```{r}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.4.] Add custom color palettes
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA", 
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```

```{r}
#Load in previously called DLs 
k562.DLs.CTSS <- readRDS("repSub.k562.DLs.CTSS.rds")
k562.DLs.TCs <- readRDS("repSub.k562.DLs.TCs.rds")

k562.DLs.CTSS.p <- readRDS("poolSub.k562.DLs.CTSS.rds")
k562.DLs.TCs.p <- readRDS("poolSub.k562.DLs.TCs.rds")

```

## [1.] K562 -------------------------

```{r}
k562.DLs.df <- as.data.frame(k562.DLs.CTSS, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/repSub.k562.DLs.CTSS.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```

```{r}
k562.DLs.df <- as.data.frame(k562.DLs.TCs, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed$sample <- factor(k562.DLs.summed$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=3,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub$sample <- factor(k562.DLs.summed.sub$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


k562.DLs.summed.sub.bidir = k562.DLs.df[abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir$sample <- factor(k562.DLs.summed.sub.bidir$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub.bidir.olap$sample <- factor(k562.DLs.summed.sub.bidir.olap$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


pdf("./plots.2/repSub.k562.DLs.TCs.cumulativeSums.pdf",width=20, height=15)
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage of all DLs") 

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_x") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal)+ labs(title="cumulative percentage of counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts of DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage of bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal)  + labs(title="cumulative counts bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage  bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```

```{r}


k562.DLs.df <- as.data.frame(k562.DLs.CTSS.p, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/poolSub.k562.DLs.CTSS.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```

```{r}
k562.DLs.df <- as.data.frame(k562.DLs.TCs.p, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed$sample <- factor(k562.DLs.summed$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=3,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub$sample <- factor(k562.DLs.summed.sub$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=3 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir$sample <- factor(k562.DLs.summed.sub.bidir$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=3 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub.bidir.olap$sample <- factor(k562.DLs.summed.sub.bidir.olap$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

pdf("./plots.2/poolSub.k562.DLs.TCs.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```

# [2.] GM12878 -------------------------

## [2.0.] Read in divergent loci from GM12878 CTSSs
#### see [1.2.] in '3.DivergentLoci/3.1.5.GM12878.summitDLs.R(md)' for "poolSub.GM12878.summitDLs.rmSingletons.rds"
```{r}
GM12878.DLs <- base::readRDS(
  file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.GM12878.summitDLs.rmSingletons.rds"
    )
  )
```

## [2.1.] Combine & subset data
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

GM12878.DLs <- base::unlist(methods::as(GM12878.DLs,"GRangesList"))
# base::unique(GM12878.DLs$sample)
GM12878.DLs <- GM12878.DLs[GM12878.DLs$sample %in% main_samples,] #1.875.808 vs 2.004.109
# base::unique(GM12878.DLs$sample)
```

## [2.2.] Load GM12878 DHSs and reformat them
```{r}
GM12878.dhs <- readr::read_delim(
    file = base::paste0(
        here::here("0.External_resources/"),
        "E116-DNase.macs2.hg38.narrowPeak"
      ),
    delim = "\t",
    col_names = base::c("chr","start","end","rank"),
    col_types = readr::cols(
        .default = readr::col_integer(),
        chr = readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = IRanges::start(GM12878.dhs) + ((IRanges::end(GM12878.dhs) - IRanges::start(GM12878.dhs))/2)
  )
GM12878.dhs <- GM12878.dhs %>%
  CAGEfightR::swapRanges() %>%
  IRanges::promoters(
      upstream = 200,
      downstream = 200
    )
```

## [2.3.] Identify DLs overlapping with DHSs and assign annotations
#### see [0.2.] for laoding of annotations
```{r}
GM12878.DLs$dhs.olap <- IRanges::countOverlaps(
    query = GM12878.DLs,
    subject = GM12878.dhs
  )
GM12878.DLs <- CAGEfightR::assignTxType(
    object = GM12878.DLs,
    txModels = txdb
  )
```

## To be discarded
```{r}
# GM12878.DLs.CTSS <- base::readRDS("repSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs <- base::readRDS("repSub.GM12878.DLs.TCs.rds")
# 
# GM12878.DLs.CTSS.p <- base::readRDS("poolSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs.p <- base::readRDS("poolSub.GM12878.DLs.TCs.rds")

# GM12878 <- readRDS("../1.CTSSs/poolSub.combined.GM12878.CTSSs.rds")
# main <- row.names(colData(GM12878))[colData(GM12878)$main=="yes"]
```

## [2.1.] Sum and accumulate 
#### To be discarded
```{r}
# GM12878.DLs.df <- GM12878.DLs %>% # GM12878.DLs.CTSS
#   base::as.data.frame(row.names = NULL)
# # GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
# 
# GM12878.DLs.summed <- GM12878.DLs.df %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub = GM12878.DLs.df %>%
#   dplyr::filter(score >= 2) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub.bidir = GM12878.DLs.df %>%
#   dplyr::filter(
#     score >= 2 & base::abs(directionality) <= 0.8) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df %>%
#   dplyr::filter(
#       score >= 2 & base::abs(directionality) <= 0.8 & dhs.olap >= 1
#     ) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#    dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
```

## [2.2.] CAGE-dependent expression differences of DHSs and non-DHSs
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    dhs = dplyr::case_when(
        dhs.olap >= 1 ~ "pos",
        TRUE ~ "neg"
      ),
    sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    # comb = dplyr::case_when(
    #     divergent == TRUE & dhs.olap >= 1 ~ "divergent DHS",
    #     divergent == FALSE & dhs.olap >= 1 ~ "non-divergent DHS",
    #     divergent == TRUE & dhs.olap == 0 ~ "divergent non-DHS",
    #     divergent == FALSE & dhs.olap == 0 ~ "non-divergent non-DHS"
    #   )
    ) %>%
  dplyr::group_by(sample, dhs) %>% # comb
  dplyr::arrange(score) %>%
  dplyr::mutate(
      rank = dplyr::row_number()/dplyr::n()
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = rank,
        y = base::log10(score),
        color = dhs # comb
      )
    ) +
  geom_line() +
  scale_colour_manual(
      # values = base::c("#226764","#71266E","#A7383E","#579757"),
      values = base::c("#500711", "#ED9BBD", "#B0041A", "#FFC95F")
    ) +
  facet_wrap(
      facets = . ~ sample, # divergent ~ sample
      nrow = 1
    ) +
  guides(
      color = guide_legend(position = "inside")
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position.inside = base::c(0.05,0.95),
      legend.justification = base::c(0,1),
      legend.title = element_blank(),
      legend.background = element_rect(
          fill = "grey75"
        ),
      legend.key.size = grid::unit(0.4, 'cm'),
      legend.key.height = grid::unit(0.4, 'cm'),
      legend.key.width = grid::unit(0.4, 'cm'),
      legend.text = element_text(size = 7),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [2.3.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_divergentDHS_expression",
          "_part2",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [2.4.] Divergency of DHSs and non-DHSs
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    dhs = dplyr::case_when(
        dhs.olap >= 1 ~ "pos",
        TRUE ~ "neg"
      ),
    sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, dhs) %>% # comb
  dplyr::arrange(directionality) %>%
  dplyr::mutate(
      rank = dplyr::row_number()/dplyr::n()
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = rank,
        y = directionality,
        color = dhs # comb
      )
    ) +
  geom_line() +
  scale_colour_manual(
      # values = base::c("#226764","#71266E","#A7383E","#579757"),
      values = base::c("#500711", "#ED9BBD", "#B0041A", "#FFC95F")
    ) +
  facet_wrap(
      facets = . ~ sample, # divergent ~ sample
      nrow = 1
    ) +
  guides(
      color = guide_legend(position = "inside")
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position.inside = base::c(0.05,0.95),
      legend.justification = base::c(0,1),
      legend.title = element_blank(),
      legend.background = element_rect(
          fill = "grey75"
        ),
      legend.key.size = grid::unit(0.4, 'cm'),
      legend.key.height = grid::unit(0.4, 'cm'),
      legend.key.width = grid::unit(0.4, 'cm'),
      legend.text = element_text(size = 7),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [2.5.] Divergent regulatory elements across genomic features
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>% 
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = main_samples,
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, divergent, txType) %>%
  dplyr::summarise(
      num = dplyr::n(),
      .groups = "drop"
    ) %>%
  dplyr::arrange(sample, txType, desc(divergent)) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = sample,
        y = base::log10(num),
        fill = divergent
      )
    ) +
  geom_bar(
      stat = "identity",
      position = position_identity() # "stack"
    ) +
  facet_wrap(
      facets = . ~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1, "lines"),
      axis.title.x = element_blank(),
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5
      )
    )
```

## [2.6.] Plot numbers of overlaps between divergent and open sites (i.e., DHSs)
```{r}
# cols <- grDevices::colorRampPalette(
#     base::rev(RColorBrewer::brewer.pal(11,"RdBu"))
#   )(100)

coef = 0.01 # 0.3
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#0099dc", "#95cbee", "#c9e2f6", "#e7f0fa")
    )(coef*100),
  grDevices::colorRampPalette(
      base::c("#FDE2E0","#FFBCB5","#FE9083","#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

GM12878.DLs %>%
  tibble::as_tibble() %>% 
  dplyr::mutate(
      dhs = dplyr::case_when(
          dhs.olap >= 1 ~ "+",
          TRUE ~ "-"
        ),
      divergent = dplyr::case_when(
          divergent == TRUE ~ "+",
          TRUE ~ "-"
        ),
      sample = base::factor(
        x = sample,
        levels = main_samples,
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, divergent, dhs) %>%
  dplyr::summarise(
      num = dplyr::n(),
      .groups = "drop"
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = divergent,
        y = dhs,
        fill = num # base::log10(num)
      )
    ) +
  geom_tile() +
  geom_text(
    mapping = aes(
          # color = base::ifelse(base::log10(num) > 5,"on","off"),
          color = base::ifelse(num > 400000,"on","off"),
          label = base::format(num, big.mark = ".")
        ),
      size = 3
    ) +
  facet_wrap(
      facets = . ~ sample,
      nrow = 1
    ) +
  scale_fill_gradientn(
      limits = base::c(0L,907592), # base::log10(907592) # 6L
      colors = .GlobalEnv$cols
    ) +
  scale_color_manual(
      values = base::c(on="white",off="black")
    ) +
  theme_bw() +
  theme(
      axis.text = element_text(
          size = 15
        ),
      aspect.ratio = 1,
      # axis.title = element_blank(),
      panel.spacing = grid::unit(0.1,"lines")
    ) +
  guides(
      fill = guide_colorbar(
          label.position = "right",
          title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5
            ),
          ticks = TRUE,
          nbin = 100,
          barheight = 5,
          label = TRUE,
          barwidth = 0.5
        ),
      color = "none"
    )
```

## [2.7.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_divergentDHS_numbers",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 3
    )
  }
```

## plots.2/repSub.GM12878.DLs.CTSS.cumulativeSums.pdf
```{r}
# ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() + theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")
# 
# ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +theme(legend.position="bottom") + 
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
```

```{r}
# GM12878.DLs.df <- as.data.frame(GM12878.DLs.TCs, row.names=NULL)
# GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
# GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# 
# GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
```

## plots.2/repSub.GM12878.DLs.TCs.cumulativeSums.pdf
```{r}
# ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")
# 
# ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
```

```{r}
# GM12878.DLs.df <- as.data.frame(GM12878.DLs.CTSS.p, row.names=NULL)
# GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
# GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# 
# GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
```

## plots.2/poolSub.GM12878.DLs.CTSS.cumulativeSums.pdf
```{r}
# ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")
# 
# ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
```

```{r}
# GM12878.DLs.df <- as.data.frame(GM12878.DLs.TCs.p, row.names=NULL)
# GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
# GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# 
# GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
# 
# 
# GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
#          mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))
```

## plots.2/poolSub.GM12878.DLs.TCs.cumulativeSums.pdf
```{r}
# ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")
# 
# ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
# 
# ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
#   geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
#   scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")
```

# [3.] Single nuclei ATACseq data integration -------------------------

## [3.0.] Define paths for single nuclei ATACseq files
#### Data mapped to "GRCh38_no_alt_analysis_set_GCA_000001405.15.fasta" genome version
#### Source of that fasta file: 'https://cf.10xgenomics.com/supp/cell-arc/refdata-cellranger-arc-GRCh38-2020-A-2.0.0.tar.gz', 'https://www.encodeproject.org/files/GRCh38_no_alt_analysis_set_GCA_000001405.15/@@download/GRCh38_no_alt_analysis_set_GCA_000001405.15.fasta.gz'
#### STAR index build with the same hg38 genome fasta file: "GRCh38_no_alt_analysis_set_GCA_000001405.15.fasta"
```{r}
count_matrix_dir <- base::paste0(
    "/maps/projects/ralab_nnfc-AUDIT/people/lpm537/project/",
    "E2G/processed/10x_multiome_240614/GM12878/cellranger_res/GM12878/outs",
    "/filtered_feature_bc_matrix/"
  )

fragments_path <- base::paste0(
    "/maps/projects/ralab_nnfc-AUDIT/people/lpm537/project/",
    "E2G/processed/10x_multiome_240614/GM12878/cellranger_res/GM12878/outs/",
    "atac_fragments.tsv.gz"
  )
metadata_path <- base::paste0(
    "/maps/projects/ralab_nnfc-AUDIT/people/lpm537/project/",
    "E2G/processed/10x_multiome_240614/GM12878/cellranger_res/GM12878/outs/",
    "per_barcode_metrics.csv"
  )
```

## [3.1.] Import count matrix as 'single nuclei atac seq' (snas) count data object
```{r}
snas_count_data <- Seurat::Read10X(
    data.dir = count_matrix_dir
  )
snas_count_data$Peaks[1:5,1:5]
```

## [3.2.] Import metadata
```{r}
pbmc_metadata <- utils::read.csv(
    file = metadata_path,
    header = TRUE,
    row.names = 1
  )
pbmc_metadata[1:5,]
```

## [3.3.] Create seurat/signac object
```{r warning=FALSE}
# # Create chromatinAssay object
# chrom_assay <- CreateChromatinAssay(
#   counts = pbmc_count_data$Peaks, ## Specify scATAC-seq
#   sep = c(":", "-"),
#   fragments = fragments_path,
#   min.cells = 10, # include peaks detected in at least 10 cells
#   min.features = 200 # include cells detecting more than 200 peaks
# )
# 
# # Create seurat object from chromatinAssay object and integrate metadata
# pbmc_atac <- CreateSeuratObject(
#   counts = chrom_assay,
#   assay = "peaks", # assay name
#   meta.data = pbmc_metadata
# )
# 
# # Remove peaks in non-standard chromosomes
# peaks.keep <- seqnames(granges(pbmc_atac)) %in% standardChromosomes(granges(pbmc_atac))
# pbmc_atac <- pbmc_atac[as.vector(peaks.keep), ]
# 
# # Add gene annotation information
# annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
# seqlevels(annotation) <- paste0('chr', seqlevels(annotation))
# Annotation(pbmc_atac) <- annotation
```

## [3.4.] Transform data
#### Count all fragments across all cells: Matrix::rowSums(snas_count_data$Peaks)
#### Absolute number of cells & fraction of cells as two alternatives
```{r}
intm_ <- Matrix::rowSums(snas_count_data$Peaks >= 1) / base::ncol(snas_count_data$Peaks)
intm <- tibble::as_tibble(intm_, rownames = "atac_site")

snas <- intm %>%
  dplyr::mutate(
      chr = stringr::str_replace(atac_site,"^(.*?):.*$","\\1"),
      start = stringr::str_replace(atac_site,"^.*?:(.*?)-.*$","\\1"),
      end = stringr::str_replace(atac_site,"^.*?:.*?-(.*?)$","\\1")
    ) %>%
  dplyr::relocate(chr,start,end) %>%
  dplyr::select(-atac_site) %>%
  GenomicRanges::GRanges()
```

## [3.5.] Read out overlap between CAGE TCs & single nuclei ATAC seq
```{r}
# hits <- GenomicRanges::countOverlaps(GM12878.DLs,snas)
# GM12878.DLs$snas.olap <- hits
# base::table(GM12878.DLs$snas.olap)

olaps <- GenomicRanges::findOverlaps( # 251.644
    query = GM12878.DLs, subject = snas
  )

snas_olap <- snas[S4Vectors::subjectHits(olaps),] %>% # 251.644
  tibble::as_tibble() %>%
  dplyr::mutate(
    snas_sites = base::paste0(
        seqnames,":",start,"-",end
      )
    ) %>%
  dplyr::select(
      snas_sites, nr_cells = "value"
    )

dls_olap <- GM12878.DLs[S4Vectors::queryHits(olaps),] %>% # 251.644
  tibble::as_tibble() %>%
   dplyr::mutate(
    DL_sites = base::paste0(
        seqnames,":",start,"-",end
      )
    ) %>%
  dplyr::select(
      DL_sites, score, directionality,
      bidirectionality, divergent, sample,
      dhs.olap, txType, snas.olap
    )

comb <- dplyr::bind_cols( # 251.644 
    snas_olap, dls_olap
  )
```

## [3.6.] Reintegrate overlaps into the 'GM12878.DLs' object
#### Evaluate with 'dplyr::filter(comb, snas.olap > 1)'
#### Example for a DL overlapping 2 ATAC sites & occuring in >1 sample
```{r}
intgr <- comb %>% # 249.989
  dplyr::group_by(DL_sites,sample) %>%
  dplyr::filter(
      nr_cells == base::max(nr_cells)
    ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
      DL_sites, snas_sites, nr_cells, sample
    ) # %>% distinct(DL_sites, snas_sites, nr_cells,sample)

DL_snas <- GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    DL_sites = base::paste0(
        seqnames,":",start,"-",end
      )
    ) %>%
  dplyr::select(
      DL_sites, score, directionality,
      bidirectionality, divergent, sample,
      dhs.olap, txType, snas.olap
    ) %>%
  dplyr::left_join(
      intgr,
      by = base::c("DL_sites","sample")
    ) %>%
  tidyr::replace_na(
      replace = base::list(
        snas_sites = "-",
        nr_cells = 0L
      )
    ) %>%
  dplyr::mutate(
    class = dplyr::case_when(
        divergent == TRUE & dhs.olap >= 1 ~ "divergent DHS",
        divergent == FALSE & dhs.olap >= 1 ~ "non-divergent DHS",
        divergent == TRUE & dhs.olap == 0 ~ "divergent non-DHS",
        divergent == FALSE & dhs.olap == 0 ~ "non-divergent non-DHS"
      ),
    sample = base::factor(
        x = sample,
        levels = main_samples,
        ordered = TRUE
      )
    )
```

## [3.7.] Plot overlaps
```{r}
coef = 0.25 # 0.3
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

upper_y <- base::log10(base::max(DL_snas$score))
upper_x <- base::max(DL_snas$nr_cells)

plot <- ggplot2::ggplot(
      data = DL_snas,
      mapping = aes(
        x = nr_cells,
        y = base::log10(score)
        # color = class
      )
    ) +
  # geom_point(
  #     alpha = 0.1
  #   ) +
  # geom_density2d(
  #     bins = 1000,
  #     contour_var = "density" # "count" # "ndensity"
  #   ) +
  geom_hex(
      mapping = aes(
        fill = stat(
            base::log10(count)
          )
        ),
      bins = 30
    ) +
  facet_grid(
      cols = vars(sample),
      rows = vars(class)
      # ncol = 6
    ) +
  scale_fill_gradientn(
      colours = cols
    ) +
  scale_x_continuous(
        limits = base::c(0,upper_x)
      ) +
  scale_y_continuous(
      limits = base::c(0,upper_y)
    ) +
  theme_bw() +
  theme(
      # legend.position = "none",
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines")
    ) +
  guides(
      fill = guide_colorbar(
          label.position = "right",
          # title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5,
              size = 10
            ),
          ticks = T,
          nbin = 100,
          barheight = 5,
          label = TRUE,
          barwidth = 0.5
        ),
      color = "none"
    )
base::print(plot)
```

## [3.8.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_divergentDHS_snATACseq_numbers",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 12,
      height = 7
    )
  }
```

## [3.9.] Plot only overlaps for 'GM12878_1M_N' sample
```{r}
upper_fill_value <- base::max(ggplot2::ggplot_build(plot)$data[[1]]$count)

plot_1MN <- DL_snas %>%
  dplyr::filter(sample == "GM12878_1M_N") %>% 
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          x = nr_cells,
          y = base::log10(score)
        )
      ) +
    geom_hex(
        mapping = aes(
          fill = stat(
              base::log10(count)
            )
          ),
        bins = 30
      ) +
    facet_wrap(
        facets = . ~ class,
        nrow = 1
      ) +
    scale_fill_gradientn(
        limits = base::c(0,base::log10(upper_fill_value)),
        colours = cols
      ) +
    scale_x_continuous(
        limits = base::c(0,upper_x)
      ) +
    scale_y_continuous(
        limits = base::c(0,upper_y)
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        panel.spacing = grid::unit(0.1,"lines")
      ) +
    guides(
        fill = guide_colorbar(
            label.position = "right",
            # title = NULL,
            title.position = "left",
            title.theme = element_text(
                angle = 90,
                hjust = 0.5,
                size = 10
              ),
            ticks = T,
            nbin = 100,
            barheight = 5,
            label = TRUE,
            barwidth = 0.5
          ),
        color = "none"
      )
base::print(plot_1MN)
```

## [3.10.] Print plot only overlaps for 'GM12878_1M_N' sample
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_divergentDHS_snATACseq_1Mnuclei",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [3.11.] Plot all panels separateley 
#### Evaluate with c = "non-divergent DHS" and s = "GM12878_1M_N"
```{r}
classes <- DL_snas %>% 
  dplyr::distinct(class) %>%
  dplyr::pull(class)

plots <- tibble::tibble(
      sample = base::rep(main_samples[1:6],base::length(classes)),
      class = base::sort(base::rep(classes,base::length(main_samples[1:6])))
    ) %>%
  dplyr::mutate(
    plot = purrr::pmap(
      .l = base::list(sample,class),
      .f = function(s,c){
        base::cat(s, "\t", c, "\n")
        plot <- DL_snas %>%
          dplyr::filter(
              sample == s & class == c
            ) %>%
          ggplot2::ggplot(
            data = .,
            mapping = aes(
              x = nr_cells,
              y = base::log10(score),
              fill = sample
            )
          ) +
        ggrastr::geom_point_rast(
            alpha = 0.01
          ) +
        geom_text(
            x = upper_x*0.95,
            y = upper_y*0.95,
            hjust = 1,
            vjust = 1,
            label = base::paste0(s," ", c)
          ) +
        scale_x_continuous(
            limits = base::c(0,upper_x)
          ) +
        scale_y_continuous(
            limits = base::c(0,upper_y)
          ) +
        theme_bw() +
        theme(
            aspect.ratio = 1,
            legend.position = "none"
          )
      return(
        ggExtra::ggMarginal(
            p = plot,
            groupFill = TRUE
          )
        )
      }
    )
  )
```

## [3.12.] Print all panel plots separately
```{r}
plots %>%
  dplyr::mutate(
    path = purrr::pmap_chr(
      .l = base::list(sample, class, plot),
      .f = function(s, c, p){
        base::print(p)
        fn <- base::paste0(
            here::here("3.DivergentLoci/plots.2/"),
            base::format(base::Sys.time(), "%Y%m%d"),
            "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
            "_divergentDHS_snATACseq",
            "_", s,
            "_", stringr::str_replace_all(c, " |-", ""),
            "_rastr"
          )
        for(ext in base::c("pdf", "png")){
          ggplot2::ggsave(
            filename = base::paste0(fn,".", ext),
            plot = p,
            device = ext,
            dpi = 300,
            width = 5,
            height = 5
          )
        }
        return(fn)
      }
    )
  )
```

```{r}
DL_snas %>%
  dplyr::group_by(sample, class) %>%
  dplyr::summarise(
      frac = base::sum(nr_cells > 0)/dplyr::n(),
      abso = base::sum(nr_cells > 0)
    ) %>%
  dplyr::filter(
      sample == "GM12878_1M_N"
    )
```

