---
title: "3.2.2.DL cumulative fractions vs expression"
output: html_document
date: "2024-03-01"
---

# [-1.] Markdown is an adaptation of the version from 2024-02-13
# title: '3.2.DL cumulative fractions vs expression'
# output: html_document
# date: "2024-02-13"

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")

# base::library("BiocParallel")
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 30
#     )
#   )
```

```{r}
# library(TxDb.Hsapiens.UCSC.hg38.knownGene)
# txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
# 
# #Packages for plotting
# library(CAGEfightR)
# library(epistack)
# library(BiocParallel)
# library(dplyr)
# library(reshape2)
# library(limma)
# library(ggplot2)
# library(stringr)
```

## [0.2.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.3.] Add custom color palettes
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA", 
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```



```{r}
#Load in previously called DLs 
k562.DLs.CTSS <- readRDS("repSub.k562.DLs.CTSS.rds")
k562.DLs.TCs <- readRDS("repSub.k562.DLs.TCs.rds")

k562.DLs.CTSS.p <- readRDS("poolSub.k562.DLs.CTSS.rds")
k562.DLs.TCs.p <- readRDS("poolSub.k562.DLs.TCs.rds")

```


## [1.] K562 -------------------------

```{r}
k562.DLs.df <- as.data.frame(k562.DLs.CTSS, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/repSub.k562.DLs.CTSS.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```




```{r}
k562.DLs.df <- as.data.frame(k562.DLs.TCs, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed$sample <- factor(k562.DLs.summed$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=3,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub$sample <- factor(k562.DLs.summed.sub$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


k562.DLs.summed.sub.bidir = k562.DLs.df[abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir$sample <- factor(k562.DLs.summed.sub.bidir$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub.bidir.olap$sample <- factor(k562.DLs.summed.sub.bidir.olap$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_10M_N",
                                            "K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align","K562_groCap_1"))


pdf("./plots.2/repSub.k562.DLs.TCs.cumulativeSums.pdf",width=20, height=15)
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage of all DLs") 

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_x") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal)+ labs(title="cumulative percentage of counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts of DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage of bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal)  + labs(title="cumulative counts bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative percentage  bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
    geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free") + theme_bw()+ scale_x_log10() + 
    scale_color_manual(values=k.full.pal) + labs(title="cumulative counts bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```





```{r}


k562.DLs.df <- as.data.frame(k562.DLs.CTSS.p, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=2 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/poolSub.k562.DLs.CTSS.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```




```{r}
k562.DLs.df <- as.data.frame(k562.DLs.TCs.p, row.names=NULL)
k562.DLs.df <- k562.DLs.df[k562.DLs.df$sample!="K562_Cap_align",] 
k562.DLs.df$sample[k562.DLs.df$sample=="K562_N_10M"] <- "K562_10M_N"
k562.DLs.summed = k562.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed$sample <- factor(k562.DLs.summed$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub = k562.DLs.df[k562.DLs.df$score>=3,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub$sample <- factor(k562.DLs.summed.sub$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub.bidir = k562.DLs.df[k562.DLs.df$score>=3 & abs(k562.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


k562.DLs.summed.sub.bidir$sample <- factor(k562.DLs.summed.sub.bidir$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

k562.DLs.summed.sub.bidir.olap = k562.DLs.df[k562.DLs.df$score>=3 & abs(k562.DLs.df$directionality)<=0.8& k562.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

k562.DLs.summed.sub.bidir.olap$sample <- factor(k562.DLs.summed.sub.bidir.olap$sample, levels=c("K562_C1","K562_C2","K562_C3","K562_C","K562_N4","K562_N5","K562_N6","K562_N",
                                            "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
                                            "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1","K562_pcr_align"))

pdf("./plots.2/poolSub.k562.DLs.TCs.cumulativeSums.pdf")
ggplot(k562.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(k562.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(k562.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(k562.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(k562.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(k562.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=k.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```

# [2.] GM12878 -------------------------

## [2.0.] Read in divergent loci from GM12878 CTSSs
#### see [1.2.] in '3.DivergentLoci/3.1.5.GM12878.summitDLs.R(md)' for "poolSub.GM12878.summitDLs.rmSingletons.rds"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )
GM12878.DLs <- base::readRDS(
  file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.GM12878.summitDLs.rmSingletons.rds"
    )
  )
```

## To be discarded
```{r}
# GM12878.DLs.CTSS <- base::readRDS("repSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs <- base::readRDS("repSub.GM12878.DLs.TCs.rds")
# 
# GM12878.DLs.CTSS.p <- base::readRDS("poolSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs.p <- base::readRDS("poolSub.GM12878.DLs.TCs.rds")

# GM12878 <- readRDS("../1.CTSSs/poolSub.combined.GM12878.CTSSs.rds")
# main <- row.names(colData(GM12878))[colData(GM12878)$main=="yes"]
```

## [2.1.] Sum and accumulate 
#### To be discarded
```{r}
# GM12878.DLs.df <- GM12878.DLs %>% # GM12878.DLs.CTSS
#   base::as.data.frame(row.names = NULL)
# # GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
# 
# GM12878.DLs.summed <- GM12878.DLs.df %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub = GM12878.DLs.df %>%
#   dplyr::filter(score >= 2) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub.bidir = GM12878.DLs.df %>%
#   dplyr::filter(
#     score >= 2 & base::abs(directionality) <= 0.8) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#   dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
# 
# GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df %>%
#   dplyr::filter(
#       score >= 2 & base::abs(directionality) <= 0.8 & dhs.olap >= 1
#     ) %>%
#   dplyr::group_by(sample,txType) %>%
#   dplyr::count(score) %>%
#    dplyr::mutate(
#       pct = n/base::sum(n),
#       cumsum = base::cumsum(pct),
#       cumsumCounts = base::cumsum(n)
#     )
```

## [2.2.] CAGE-dependent expression differences of DHSs and non-DHSs
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    dhs = dplyr::case_when(
        dhs.olap >= 1 ~ "pos",
        TRUE ~ "neg"
      ),
    sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    # comb = dplyr::case_when(
    #     divergent == TRUE & dhs.olap >= 1 ~ "divergent DHS",
    #     divergent == FALSE & dhs.olap >= 1 ~ "non-divergent DHS",
    #     divergent == TRUE & dhs.olap == 0 ~ "divergent non-DHS",
    #     divergent == FALSE & dhs.olap == 0 ~ "non-divergent non-DHS"
    #   )
    ) %>%
  dplyr::group_by(sample, dhs) %>% # comb
  dplyr::arrange(score) %>%
  dplyr::mutate(
      rank = dplyr::row_number()/dplyr::n()
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = rank,
        y = base::log10(score),
        color = dhs # comb
      )
    ) +
  geom_line() +
  scale_colour_manual(
      # values = base::c("#226764","#71266E","#A7383E","#579757"),
      values = base::c("#500711", "#ED9BBD", "#B0041A", "#FFC95F")
    ) +
  facet_wrap(
      facets = . ~ sample, # divergent ~ sample
      nrow = 1
    ) +
  guides(
      color = guide_legend(position = "inside")
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position.inside = base::c(0.05,0.95),
      legend.justification = base::c(0,1),
      legend.title = element_blank(),
      legend.background = element_rect(
          fill = "grey75"
        ),
      legend.key.size = grid::unit(0.4, 'cm'),
      legend.key.height = grid::unit(0.4, 'cm'),
      legend.key.width = grid::unit(0.4, 'cm'),
      legend.text = element_text(size = 7),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [2.3.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_divergentDHS_expression",
          "_part2",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [2.4.] Divergency of DHSs and non-DHSs
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    dhs = dplyr::case_when(
        dhs.olap >= 1 ~ "pos",
        TRUE ~ "neg"
      ),
    sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, dhs) %>% # comb
  dplyr::arrange(directionality) %>%
  dplyr::mutate(
      rank = dplyr::row_number()/dplyr::n()
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = rank,
        y = directionality,
        color = dhs # comb
      )
    ) +
  geom_line() +
  scale_colour_manual(
      # values = base::c("#226764","#71266E","#A7383E","#579757"),
      values = base::c("#500711", "#ED9BBD", "#B0041A", "#FFC95F")
    ) +
  facet_wrap(
      facets = . ~ sample, # divergent ~ sample
      nrow = 1
    ) +
  guides(
      color = guide_legend(position = "inside")
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position.inside = base::c(0.05,0.95),
      legend.justification = base::c(0,1),
      legend.title = element_blank(),
      legend.background = element_rect(
          fill = "grey75"
        ),
      legend.key.size = grid::unit(0.4, 'cm'),
      legend.key.height = grid::unit(0.4, 'cm'),
      legend.key.width = grid::unit(0.4, 'cm'),
      legend.text = element_text(size = 7),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [2.5.] Divergent regulatory elements across genomic features
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>% 
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = main_samples,
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, divergent, txType) %>%
  dplyr::summarise(
      num = dplyr::n(),
      .groups = "drop"
    ) %>%
  dplyr::arrange(sample, txType, desc(divergent)) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = sample,
        y = base::log10(num),
        fill = divergent
      )
    ) +
  geom_bar(
      stat = "identity",
      position = position_identity() # "stack"
    ) +
  facet_wrap(
      facets = . ~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1, "lines"),
      axis.title.x = element_blank(),
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5
      )
    )
```

## [2.6.]
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>% 
  dplyr::mutate(
      dhs = dplyr::case_when(
          dhs.olap >= 1 ~ TRUE,
          TRUE ~ FALSE
        ),
      sample = base::factor(
        x = sample,
        levels = main_samples,
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(sample, divergent, dhs) %>%
  dplyr::summarise(
      num = dplyr::n(),
      .groups = "drop"
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = divergent,
        y = dhs,
        fill = num
      )
    ) +
  geom_tile() +
  facet_wrap(
      facets = . ~ sample,
      nrow = 1
    ) +
  theme_bw()
```



```{r}
pdf("./plots.2/repSub.GM12878.DLs.CTSS.cumulativeSums.pdf")
ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() + theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +theme(legend.position="bottom") + 
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```




```{r}
GM12878.DLs.df <- as.data.frame(GM12878.DLs.TCs, row.names=NULL)
GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/repSub.GM12878.DLs.TCs.cumulativeSums.pdf")
ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```





```{r}


GM12878.DLs.df <- as.data.frame(GM12878.DLs.CTSS.p, row.names=NULL)
GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/poolSub.GM12878.DLs.CTSS.cumulativeSums.pdf")
ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```




```{r}
GM12878.DLs.df <- as.data.frame(GM12878.DLs.TCs.p, row.names=NULL)
GM12878.DLs.df <- GM12878.DLs.df[GM12878.DLs.df$sample%in%main,] 
GM12878.DLs.summed = GM12878.DLs.df %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))

GM12878.DLs.summed.sub = GM12878.DLs.df[GM12878.DLs.df$score>=2,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))



GM12878.DLs.summed.sub.bidir = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


GM12878.DLs.summed.sub.bidir.olap = GM12878.DLs.df[GM12878.DLs.df$score>=2 & abs(GM12878.DLs.df$directionality)<=0.8& GM12878.DLs.df$dhs.olap>=1,] %>% group_by(sample,txType) %>% count(score) %>%
         mutate(pct = n/sum(n)) %>% mutate(cumsum=cumsum(pct), cumsumCounts=cumsum(n))


pdf("./plots.2/poolSub.GM12878.DLs.TCs.cumulativeSums.pdf")
ggplot(GM12878.DLs.summed, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of all DLs")

ggplot(GM12878.DLs.summed, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of all DLs")


ggplot(GM12878.DLs.summed.sub, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts DLs")

ggplot(GM12878.DLs.summed.sub, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts DLs")


ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative percentage of >=3 counts & bidirScore<0.8  DLs")

ggplot(GM12878.DLs.summed.sub.bidir, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8  DLs")



ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsum,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType) + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500))+ labs(title="cumulative percentage of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

ggplot(GM12878.DLs.summed.sub.bidir.olap, aes(score,cumsumCounts,color=sample)) +
  geom_line(aes(group=sample)) + facet_wrap(~txType, scales = "free_y") + theme_bw() +theme(legend.position="bottom") + 
  scale_color_manual(values=gm.full.pal) + xlim(c(0,500)) + labs(title="cumulative counts of >=3 counts & bidirScore<0.8 & DHS overlap  DLs")

dev.off()
```


