---
title: "3.3. DL signatures"
output: html_document
date: "2024-03-01"
---

# [0.] Prerequisites -------------------------

## [0.0.] Install necessary packages
```{r}
#utils::install.packages("ggbreak")
```

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")
```

## [0.2.] Load annotations
```{r}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## To be discarded
```{r}
# base::library("BiocParallel")
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 30
#     )
#   )
# 
# #Packages for plotting
# library(CAGEfightR)
# library(epistack)
# library(BiocParallel)
# library(dplyr)
# library(reshape2)
# library(limma)
# library(ggplot2)
# library(stringr)
```

## [0.3.] Add custom color palettes
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA", 
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#00439B","#FF4425",
    "#009072","#878784","#7D62D0",
    "hotpink"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```

## [1.] GM12878 -------------------------

## [1.0.] Read in divergent loci from GM12878 CTSSs
#### see [1.2.] in '3.DivergentLoci/3.1.5.GM12878.summitDLs.Rmd' for "poolSub.GM12878.summitDLs.rmSingletons.rds"
```{r}
GM12878.DLs.CTSS <- base::readRDS(
  file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.GM12878.summitDLs.rmSingletons.rds"
    )
  )
```

## [2.1.] Subset samples for main_samples
## see [2.9.] in '1.CTSSs/1.1.CTSSs.processing.Rmd' for "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

GM12878.DLs <- base::unlist(methods::as(GM12878.DLs.CTSS,"GRangesList"))
# base::unique(GM12878.DLs$sample)
GM12878.DLs <- GM12878.DLs[GM12878.DLs$sample %in% main_samples,] #1.875.808 vs 2.004.10
# base::unique(GM12878.DLs$sample)
```

## [2.2.] Plot total amount of divergent loci
#### Resulting plots stored in 'plots.2/repSub.GM12878.nDLs.CTSS.pdf'
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>% # dplyr::group_by(sample) %>% dplyr::summarise(num = dplyr::n())
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          y = sample,
          fill = sample
        )
      ) +
    geom_bar() +
    # scale_x_log10() +
    # ggbreak::scale_x_break(
    #     breaks = base::c(200500,1500000)
    #   ) +
    # scale_x_continuous(
    #     limits = base::c(0,200500)
    #   ) +
    scale_fill_manual(
        limits = main_samples[1:6],
        values = gm.pooled.pal
      ) +
    labs(
        title = "total DLs detected"
      ) + 
    theme_bw() +
    theme(
        aspect.ratio = 1,
        axis.title.y = element_blank(),
        legend.position = "none"
      )
```

## [2.3.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_totalAmount",
          "_part2",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.4.] Plot subsetted divergent loci numbers
#### Replotting the number of DLs with a cutoff at score >= 3 is meaningless as the minimum score is 4
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(score)
```

#### Evaluate the range of all bidirectionality scores across samples
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>%
  dplyr::reframe(r = base::range(bidirectionality))
```

#### Plot nDLs across bidirectionality scores
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      sample = base::factor(
          x = sample,
          levels = base::rev(main_samples),
          ordered = TRUE
        ),
      absDir = base::abs(directionality)
    ) %>% 
  dplyr::group_by(sample) %>%
  dplyr::arrange(absDir) %>%
  dplyr::mutate(
      rank = dplyr::row_number()
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
      x = absDir,
      y = rank,
      color = sample
      )
    ) +
  geom_line() +
  geom_vline(
      xintercept = 0.8,
      linetype = "dashed"
    ) +
  scale_y_log10() +
  scale_color_manual(
      breaks = main_samples[1:6],
      values = gm.pooled.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.5.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_absDirScores",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.6.] Plot nDLs at for all sites with bidirectionality score of <= 0.8
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  as.data.frame(row.names = NULL) %>%
  dplyr::filter(
      base::abs(directionality) <= 0.8
    ) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>% # dplyr::group_by(sample) %>% dplyr::summarise(num = dplyr::n())
  ggplot(
    data = .,
    mapping = aes(
        y = sample,
        fill = sample
      )
    ) +
  geom_bar() +
  theme_bw() +
  # scale_x_log10() +
  scale_x_continuous(
      limits = base::c(0,200500) # 200500 # 1536846
    ) +
  scale_fill_manual(
      limits = main_samples[1:6],
      values = gm.pooled.pal
    ) +
  labs(
      title = "DLs detected directionality <= 0.8"
    ) +
  theme(
      aspect.ratio = 1,
      legend.position = "none"
    )
```

## [2.7.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_dirCutOff08",
          # "_part2",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.8.] Load GM12878 DHSs and reformat them
```{r}
GM12878.dhs <- readr::read_delim(
    file = base::paste0(
        here::here("0.External_resources/"),
        "E116-DNase.macs2.hg38.narrowPeak"
      ),
    delim = "\t",
    col_names = base::c("chr","start","end","rank"),
    col_types = readr::cols(
        .default = readr::col_integer(),
        chr = readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = IRanges::start(GM12878.dhs) + ((IRanges::end(GM12878.dhs) - IRanges::start(GM12878.dhs))/2)
  )
GM12878.dhs <- GM12878.dhs %>%
  CAGEfightR::swapRanges() %>%
  IRanges::promoters(
      upstream = 200,
      downstream = 200
    )
```

## [2.9.] Identify DLs overlapping with DHSs and assign annotations
#### see [0.2.] for laoding of annotations
```{r}
GM12878.DLs$dhs.olap <- IRanges::countOverlaps(
    query = GM12878.DLs,
    subject = GM12878.dhs
  )
GM12878.DLs <- CAGEfightR::assignTxType(
    object = GM12878.DLs,
    txModels = txdb
  )
```

## [2.10.] Plot nDLs at for all sites overlapping DHSs with bidirectionality score of <= 0.8
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::filter(
      base::abs(directionality) <= 0.8 &
      dhs.olap >= 1
    ) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>% # dplyr::group_by(sample) %>% dplyr::summarise(num = dplyr::n())
  ggplot(
    data = .,
    mapping = aes(
        y = sample,
        fill = sample
      )
    ) +
  geom_bar() +
  theme_bw() +
  scale_x_continuous(
      # trans = 'log10',
      limits = base::c(0L,200500) #1536846
    ) +
  scale_fill_manual(
    limits = main_samples[1:6],
    values = gm.pooled.pal
    ) +
  labs(
      title = "DLs overlapping DHSs with directionality <= 0.8"
    ) +
  theme(
      aspect.ratio = 1,
      legend.position = "none"
    )
```

## [2.11.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_DhsOverlap_dirCutOff08",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.12.] Plot nDLs with DHS overlap across bidirectionality scores
#### Upper limit: 1536846	from GM12878_10M_N
```{r}
GM12878.DLs %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      dhs.olap >= 1L
    ) %>% 
  dplyr::mutate(
      absDir = base::abs(directionality)
    ) %>% 
  dplyr::group_by(sample) %>%
  dplyr::arrange(absDir) %>%
  dplyr::mutate(
      sample = base::factor(
          x = sample,
          levels = base::rev(main_samples[1:6]),
          ordered = TRUE
        ),
      rank = dplyr::row_number()
    ) %>% # dplyr::summarise(ylim = base::max(rank))
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = absDir,
        y = rank,
        color = sample
      )
    ) +
  geom_line() +
  geom_vline(
      xintercept = 0.8,
      linetype = "dashed"
    ) +
  scale_y_continuous(
      limits = base::c(1L,1536846),
      trans = 'log10'
    ) +
  scale_color_manual(
      limits = main_samples[1:6],
      values = gm.pooled.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.13.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_DhsOverlap_dirCutOff08_absDirScores",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.14.] Directionality signature over all regions
#### "repSub.GM12878.bidirScore.CTSS.pdf"
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = directionality,
        y = sample,
        fill = sample
      )
    ) +
  ggridges::geom_density_ridges(
      scale = 3,
      alpha = 0.8
    ) +
  theme_bw() +
  labs(
      title = "bidir score over all DLs"
    ) +
  scale_fill_manual(
      breaks = main_samples[1:6],
      values = gm.pooled.pal
    ) +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.title.y = element_blank()
    )
```

## [2.15.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_dirScores",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [2.16.] Directionality signature across pooled samples separated by genomic features
#### title = "bidir score over all DLs", fill="annotation"
```{r}
 GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = directionality,
        y = sample,
        fill = sample
      )
    ) +
  # geom_histogram(bins = 25, position = "dodge") + 
  ggridges::geom_density_ridges(
      scale = 3,
      alpha = 0.8
    ) +
  facet_wrap(
      # scales = "free_y"
      facets = . ~ txType,
      nrow = 1
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines"),
      legend.position = "none"
    )
```


## [2.17.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_dirScores_genomicFeat",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```

## [2.18.] Directionality signature over all regions

```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  # dplyr::filter(
  #     dhs.olap >= 1
  #   ) %>%
  dplyr::mutate(
      dhs = dplyr::case_when(
          dhs.olap >= 1 ~ "pos",
          TRUE ~ "neg"
        ),
      sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = directionality,
        fill = dhs
      )
    ) +
  geom_density(
      alpha = 0.5
    ) +
  # geom_histogram(
  #     bins = 25,
  #     position = "dodge"
  #   ) + 
  facet_wrap(
      scales = "free_y",
      facets = . ~ sample
    ) +
  labs(title = "bidir score over all DLs", fill="DHS overlap") +
  scale_fill_manual(
      values = base::c("red","blue")
    ) +
  theme_bw()
```

```{r}
 GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::filter(
      sample != "GM12878_10M_N"
    ) %>% 
  dplyr::mutate(
      dhs = dplyr::case_when(
          dhs.olap >= 1 ~ "pos",
          TRUE ~ "neg"
        ),
      sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = directionality,
        y = sample,
        fill = dhs
      )
    ) +
  # geom_histogram(bins = 25, position = "dodge") + 
  ggridges::geom_density_ridges(
      scale = 1,
      alpha = 0.4
    ) +
  facet_wrap(
      # scales = "free_y"
      facets = . ~ txType,
      nrow = 1
    ) +
  scale_x_continuous(
      limits = base::c(-1,1)
    ) +
  scale_fill_manual(
      breaks = base::c("neg","pos"),
      values = base::c("red","blue")
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 3,
      panel.spacing = grid::unit(0.1,"lines"),
      legend.position = "none",
      axis.title.y = element_blank()
    )
```


## [2.19.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("3.DivergentLoci/plots.2/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.summitDLs.rmSingletons.main_samples",
          "_nDLs_DhsOverlap_dirScores_genomicFeat",
          "_v2",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```
