---
title: "3.3.2.DLs.signatures"
output: html_document
date: "2024-03-01"
---

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")

# base::library("BiocParallel")
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 30
#     )
#   )
```

```{r}
# library(TxDb.Hsapiens.UCSC.hg38.knownGene)
# txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
# 
# #Packages for plotting
# library(CAGEfightR)
# library(epistack)
# library(BiocParallel)
# library(dplyr)
# library(reshape2)
# library(limma)
# library(ggplot2)
# library(stringr)
```

## [0.2.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.3.] Add custom color palettes
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA", 
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```




```{r}
#Load in previously called DLs 
k562.DLs.CTSS <- readRDS("repSub.k562.DLs.CTSS.rds")
k562.DLs.TCs <- readRDS("repSub.k562.DLs.TCs.rds")

k562.DLs.CTSS <- k562.DLs.CTSS[k562.DLs.CTSS$sample!="K562_Cap_align",] 
k562.DLs.CTSS$sample[k562.DLs.CTSS$sample=="K562_N_10M"] <- "K562_10M_N"

k562.DLs.TCs <- k562.DLs.TCs[k562.DLs.TCs$sample!="K562_Cap_align",] 
k562.DLs.TCs$sample[k562.DLs.TCs$sample=="K562_N_10M"] <- "K562_10M_N"

k562.DLs.CTSS.p <- readRDS("poolSub.k562.DLs.CTSS.rds")
k562.DLs.TCs.p <- readRDS("poolSub.k562.DLs.TCs.rds")

k562.DLs.CTSS.p <- k562.DLs.CTSS.p[k562.DLs.CTSS.p$sample!="K562_Cap_align",] 
k562.DLs.CTSS.p$sample[k562.DLs.CTSS.p$sample=="K562_N_10M"] <- "K562_10M_N"

k562.DLs.TCs.p <- k562.DLs.TCs.p[k562.DLs.TCs.p$sample!="K562_Cap_align",] 
k562.DLs.TCs.p$sample[k562.DLs.TCs.p$sample=="K562_N_10M"] <- "K562_10M_N"
```


#K562
```{r}
#Simple plot of the numbers of DLs per samples


pdf("./plots.2/repSub.k562.nDLs.CTSS.pdf")
k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()


pdf("./plots.2/repSub.k562.nDLs.TCs.pdf")
k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()

```


```{r}
#Simple plot of the numbers of DLs per samples


pdf("./plots.2/poolSub.k562.nDLs.CTSS.pdf")
k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()

pdf("./plots.2/poolSub.k562.nDLs.TCs.pdf")
k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()

```






```{r}
#Directionality signature over different regions

pdf("./plots.2/repSub.k562.bidirScore.CTSS.pdf")
k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs") + scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)


k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()



pdf("./plots.2/repSub.k562.bidirScore.TCs.pdf", width = 20, height = 15)
k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)


k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```



```{r}
#Directionality signature over different regions

pdf("./plots.2/poolSub.k562.bidirScore.CTSS.pdf")
k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)


k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


pdf("./plots.2/poolSub.k562.bidirScore.TCs.pdf")
k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)


k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```

## [2.] GM12878 -------------------------

## [2.0.] Read in divergent loci from GM12878 CTSSs
#### see [1.2.] in '3.DivergentLoci/3.1.5.GM12878.summitDLs.R(md)' for "poolSub.GM12878.summitDLs.rmSingletons.rds"
```{r}
GM12878.DLs.CTSS <- base::readRDS(
  file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.GM12878.summitDLs.rmSingletons.rds"
    )
  )
```

## To be discarded
```{r}
# GM12878.DLs.CTSS <- base::readRDS("repSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs <- base::readRDS("repSub.GM12878.DLs.TCs.rds")
# 
# GM12878.DLs.CTSS.p <- base::readRDS("poolSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs.p <- base::readRDS("poolSub.GM12878.DLs.TCs.rds")
```

## [2.1.] Read in GM12878 CTSSs
## see [2.9.] in '1.CTSSs/1.1.CTSSs.processing.Rmd' for "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

GM12878.DLs <- base::unlist(methods::as(GM12878.DLs.CTSS,"GRangesList"))
# base::unique(GM12878.DLs$sample)
GM12878.DLs <- GM12878.DLs[GM12878.DLs$sample %in% main_samples,] #1.875.808 vs 2.004.10
# base::unique(GM12878.DLs$sample)
```

## To be discarded
```{r}
# GM12878.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
```

## To be discarded
```{r}
# main <- base::row.names(SummarizedExperiment::colData(GM12878.CTSSs))[SummarizedExperiment::colData(GM12878.CTSSs)$main == "yes"]
# 
# GM12878.DLs.CTSS <- GM12878.DLs.CTSS[GM12878.DLs.CTSS$sample%in%main,]
# GM12878.DLs.TCs <- GM12878.DLs.TCs[GM12878.DLs.TCs$sample%in%main,]
# 
# GM12878.DLs.CTSS.p <- GM12878.DLs.CTSS.p[GM12878.DLs.CTSS.p$sample%in%main,]
# GM12878.DLs.TCs.p <- GM12878.DLs.TCs.p[GM12878.DLs.TCs.p$sample%in%main,]
```


```{r}
# pdf("./plots.2/repSub.GM12878.nDLs.CTSS.pdf")
 GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        y = sample,
        fill = sample
      )
    ) +
  geom_bar() +
  scale_fill_manual(
    values = gm.full.pal) +
  labs(
      title = "total DLs detected"
    ) + 
  theme_bw() +
  theme(
      legend.position = "bottom"
    )
```


```{r}
GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()


pdf("./plots.2/repSub.GM12878.nDLs.TCs.pdf")
GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

```


```{r}
#Simple plot of the numbers of DLs per samples


pdf("./plots.2/poolSub.GM12878.nDLs.CTSS.pdf")
GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected") +theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

pdf("./plots.2/poolSub.GM12878.nDLs.TCs.pdf")
GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

```






```{r}
#Directionality signature over different regions

pdf("./plots.2/repSub.GM12878.bidirScore.CTSS.pdf")
GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()



pdf("./plots.2/repSub.GM12878.bidirScore.TCs.pdf")
GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```



```{r}
#Directionality signature over different regions

pdf("./plots.2/poolSub.GM12878.bidirScore.CTSS.pdf")
GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


pdf("./plots.2/poolSub.GM12878.bidirScore.TCs.pdf")
GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```

