---
title: "3.3.2.DLs.signatures"
output: html_document
date: "2024-03-01"
---

# [0.] Prerequisites -------------------------

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")
```

## [0.2.] Load annotations
```{r}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## To be discarded
```{r}
# base::library("BiocParallel")
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 30
#     )
#   )
# 
# #Packages for plotting
# library(CAGEfightR)
# library(epistack)
# library(BiocParallel)
# library(dplyr)
# library(reshape2)
# library(limma)
# library(ggplot2)
# library(stringr)
```

## [0.2.] Load CAGEfighR extensions
## To be discarded
```{r message=FALSE,results='hide'}
# sep <- .Platform$file.sep
# cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
# fs <- base::c(
#     "utils","support","cumulative",
#     "decompose","enhancers","noise",
#     "coverage","complexity","subsample",
#     "dispersion"
#   )
# 
# purrr::map(
#   .x = fs,
#   .f = function(f){
#     base::source(
#       base::paste0(
#         cfr.extensions.dir, sep, f, ".R"
#         )
#       )
#     }
#   )
```

## [0.3.] Add custom color palettes
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA", 
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#00439B","#FF4425",
    "#009072","#878784","#7D62D0",
    "hotpink"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```

## [1.] K562 -------------------------

## [1.0.] Read in divergent loci from K562 CTSSs
## To be discarded
```{r}
# k562.DLs.CTSS <- readRDS("repSub.k562.DLs.CTSS.rds")
# k562.DLs.TCs <- readRDS("repSub.k562.DLs.TCs.rds")
# 
# k562.DLs.CTSS <- k562.DLs.CTSS[k562.DLs.CTSS$sample!="K562_Cap_align",] 
# k562.DLs.CTSS$sample[k562.DLs.CTSS$sample=="K562_N_10M"] <- "K562_10M_N"
# 
# k562.DLs.TCs <- k562.DLs.TCs[k562.DLs.TCs$sample!="K562_Cap_align",] 
# k562.DLs.TCs$sample[k562.DLs.TCs$sample=="K562_N_10M"] <- "K562_10M_N"
# 
# k562.DLs.CTSS.p <- readRDS("poolSub.k562.DLs.CTSS.rds")
# k562.DLs.TCs.p <- readRDS("poolSub.k562.DLs.TCs.rds")
# 
# k562.DLs.CTSS.p <- k562.DLs.CTSS.p[k562.DLs.CTSS.p$sample!="K562_Cap_align",] 
# k562.DLs.CTSS.p$sample[k562.DLs.CTSS.p$sample=="K562_N_10M"] <- "K562_10M_N"
# 
# k562.DLs.TCs.p <- k562.DLs.TCs.p[k562.DLs.TCs.p$sample!="K562_Cap_align",] 
# k562.DLs.TCs.p$sample[k562.DLs.TCs.p$sample=="K562_N_10M"] <- "K562_10M_N"
```

## [2.1.] Plot of the numbers of DLs per samples
```{r}
pdf("./plots.2/repSub.k562.nDLs.CTSS.pdf")
k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()
```

```{r}
pdf("./plots.2/repSub.k562.nDLs.TCs.pdf")
k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()

```

## [2.2.] Plot of the numbers of DLs per samples
```{r}
pdf("./plots.2/poolSub.k562.nDLs.CTSS.pdf")
k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()
```

```{r}
pdf("./plots.2/poolSub.k562.nDLs.TCs.pdf")
k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="total DLs detected")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")
dev.off()
```

## [2.3.] Directionality signature over different regions
```{r}
pdf("./plots.2/repSub.k562.bidirScore.CTSS.pdf")
k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs") + scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)
dev.off()
```

```{r}
pdf("./plots.2/repSub.k562.bidirScore.TCs.pdf", width = 20, height = 15)
k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()
```

## [2.4.] Directionality signature over different regions
```{r}
pdf("./plots.2/poolSub.k562.bidirScore.CTSS.pdf")
k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)
dev.off()
```


```{r}
pdf("./plots.2/poolSub.k562.bidirScore.TCs.pdf")
k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=k.full.pal)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=k.full.pal)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

k562.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)
dev.off()
```

## [2.] GM12878 -------------------------

## [2.0.] Read in divergent loci from GM12878 CTSSs
#### see [1.2.] in '3.DivergentLoci/3.1.5.GM12878.summitDLs.R(md)' for "poolSub.GM12878.summitDLs.rmSingletons.rds"
```{r}
GM12878.DLs.CTSS <- base::readRDS(
  file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.GM12878.summitDLs.rmSingletons.rds"
    )
  )
```

## To be discarded
```{r}
# GM12878.DLs.CTSS <- base::readRDS("repSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs <- base::readRDS("repSub.GM12878.DLs.TCs.rds")
# 
# GM12878.DLs.CTSS.p <- base::readRDS("poolSub.GM12878.DLs.CTSS.rds")
# GM12878.DLs.TCs.p <- base::readRDS("poolSub.GM12878.DLs.TCs.rds")
```

## [2.1.] Subset samples for main_samples
## see [2.9.] in '1.CTSSs/1.1.CTSSs.processing.Rmd' for "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
```{r}
main_samples <- base::c(
    "GM12878_500K_C", "GM12878_1M_C", "GM12878_500K_N",
    "GM12878_1M_N", "GM12878_5M_N", "GM12878_10M_N",
    "GM12878_1M_trip", "GM12878_1M_ctrl", "GM12878_groCap_1"
  )

GM12878.DLs <- base::unlist(methods::as(GM12878.DLs.CTSS,"GRangesList"))
# base::unique(GM12878.DLs$sample)
GM12878.DLs <- GM12878.DLs[GM12878.DLs$sample %in% main_samples,] #1.875.808 vs 2.004.10
# base::unique(GM12878.DLs$sample)
```

## Read in GM12878 CTSSs
## To be discarded
```{r}
# GM12878.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("1.CTSSs/"),
#       "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
```

## To be discarded
```{r}
# main <- base::row.names(SummarizedExperiment::colData(GM12878.CTSSs))[SummarizedExperiment::colData(GM12878.CTSSs)$main == "yes"]
# 
# GM12878.DLs.CTSS <- GM12878.DLs.CTSS[GM12878.DLs.CTSS$sample%in%main,]
# GM12878.DLs.TCs <- GM12878.DLs.TCs[GM12878.DLs.TCs$sample%in%main,]
# 
# GM12878.DLs.CTSS.p <- GM12878.DLs.CTSS.p[GM12878.DLs.CTSS.p$sample%in%main,]
# GM12878.DLs.TCs.p <- GM12878.DLs.TCs.p[GM12878.DLs.TCs.p$sample%in%main,]
```

## [2.2.] Plot total amount of divergent loci
#### Resulting plots stored in 'plots.2/repSub.GM12878.nDLs.CTSS.pdf'
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          y = sample,
          fill = sample
        )
      ) +
    geom_bar() +
    scale_x_log10() +
    scale_fill_manual(
        limits = main_samples[1:6],
        values = gm.pooled.pal
      ) +
    labs(
        title = "total DLs detected"
      ) + 
    theme_bw() +
    theme(
        aspect.ratio = 1,
        axis.title.y = element_blank(),
        legend.position = "none"
      )
```

## [2.3.] Plot subsetted divergent loci numbers
#### Replotting the number of DLs with a cutoff at score >= 3 is meaningless as the minimum score is 4
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(score)
```

#### Evaluate the range of all bidirectionality scores across samples
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>%
  dplyr::reframe(r = base::range(bidirectionality))
```

#### Read out nDLs across bidirectionality scores for all samples
```{r}
all.nDLs <- tibble::tibble(
      sample = base::sort(
          base::rep(
            main_samples[1:6],
            base::length(base::seq(0,1,0.05))
          )
        ),
      biDirScore = base::rep(
        base::seq(0,1,0.05),
        base::length(main_samples[1:6])
      )
    ) %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = main_samples[1:6],
        ordered = TRUE
      )
    ) %>%
  dplyr::mutate(
    nDLs = purrr::pmap_dbl(
      .l = base::list(sample,biDirScore),
      .f = function(s,bds){
        return(
          GM12878.DLs %>% # GM12878.DLs.CTSS
            base::as.data.frame(row.names = NULL) %>%
            dplyr::filter(
                sample == s & base::abs(directionality) <= bds
              ) %>%
            base::nrow()
        )
      }
    )
  )
```

#### Plot nDLs across bidirectionality scores
```{r}
ggplot2::ggplot(
      data = all.nDLs,
      mapping = aes(
        x = biDirScore,
        y = log10(nDLs),
        color = sample
      )
    ) +
  geom_line() +
  geom_vline(
      xintercept = 0.8,
      linetype = "dashed"
    ) +
  scale_color_manual(
      values = gm.pooled.pal
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

#### Plot nDLs at for all sites with bidirectionality score of <= 0.8
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  as.data.frame(row.names = NULL) %>%
  dplyr::filter(
      base::abs(directionality) <= 0.8
    ) %>%
  dplyr::mutate(
      sample = base::factor(
        x = sample,
        levels = base::rev(main_samples),
        ordered = TRUE
      )
    ) %>% 
  ggplot(
    data = .,
    mapping = aes(
        y = sample,
        fill = sample
      )
    ) +
  geom_bar() +
  theme_bw() +
  scale_x_log10() +
  scale_fill_manual(
    limits = main_samples[1:6],
    values = gm.pooled.pal
    ) +
  labs(
      title = "DLs detected directionality <= 0.8"
    ) +
  theme(
      aspect.ratio = 1,
      legend.position = "none"
    )
```

## [2.4.] Load GM12878 DHSs and reformat them
```{r}
GM12878.dhs <- readr::read_delim(
    file = base::paste0(
        here::here("0.External_resources/"),
        "E116-DNase.macs2.hg38.narrowPeak"
      ),
    delim = "\t",
    col_names = base::c("chr","start","end","rank"),
    col_types = readr::cols(
        .default = readr::col_integer(),
        chr = readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

GM12878.dhs$thick <- IRanges::IRanges(
    start = IRanges::start(GM12878.dhs) + ((IRanges::end(GM12878.dhs) - IRanges::start(GM12878.dhs))/2)
  )
GM12878.dhs <- GM12878.dhs %>%
  CAGEfightR::swapRanges() %>%
  IRanges::promoters(
      upstream = 200,
      downstream = 200
    )
```

## [2.5.] Identify DLs overlapping with DHSs and assign annotations
#### see [0.2.] for laoding of annotations
```{r}
GM12878.DLs$dhs.olap <- IRanges::countOverlaps(
    query = GM12878.DLs,
    subject = GM12878.dhs
  )
GM12878.DLs <- CAGEfightR::assignTxType(
    object = GM12878.DLs,
    txModels = txdb
  )
```

## [2.6.] Plot nDLs at for all sites overlapping DHSs with bidirectionality score of <= 0.8
```{r}
GM12878.DLs %>% # GM12878.DLs.CTSS
  base::as.data.frame(row.names = NULL) %>%
  dplyr::filter(
      base::abs(directionality) <= 0.8 &
      dhs.olap >= 1
    ) %>%
  ggplot2::ggplot(
    mapping = aes(
        y = sample,
        fill = sample
      )
    ) +
  geom_bar() +
  theme_bw() +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  labs(
      title = "DLs overlapping DHSs with directionality <= 0.8"
    ) +
  theme(
      legend.position = "bottom"
    )
```


```{r}
pdf("./plots.2/repSub.GM12878.nDLs.TCs.pdf")
GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

```


```{r}
#Simple plot of the numbers of DLs per samples


pdf("./plots.2/poolSub.GM12878.nDLs.CTSS.pdf")
GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected") +theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

pdf("./plots.2/poolSub.GM12878.nDLs.TCs.pdf")
GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="total DLs detected")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8")+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3 & abs(directionality)<=0.8 & dhs.olap>=1) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=gm.full.pal) + labs(title="DLs detected score>=3 & directionality <=0.8 & DHS olverlap")+theme(legend.position="bottom") 
dev.off()

```






```{r}
#Directionality signature over different regions

pdf("./plots.2/repSub.GM12878.bidirScore.CTSS.pdf")
GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.CTSS %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()



pdf("./plots.2/repSub.GM12878.bidirScore.TCs.pdf")
GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.TCs %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```



```{r}
#Directionality signature over different regions

pdf("./plots.2/poolSub.GM12878.bidirScore.CTSS.pdf")
GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.CTSS.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


pdf("./plots.2/poolSub.GM12878.bidirScore.TCs.pdf")
GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() + labs(title="bidir score over all DLs")+ scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=sample)) + geom_histogram(bins=25) + facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3") + scale_fill_manual(values=gm.full.pal)+theme(legend.position="bottom") 


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=dhs.olap>=1)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="DHS overlap") + scale_fill_manual(values=custom.col[c(4,6)])



GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() + 
  labs(title="bidir score over all DLs", fill="annotation") + scale_fill_manual(values=custom.col)


GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with score >=3", fill="annotation") + scale_fill_manual(values=custom.col)



GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap", fill="annotation") + scale_fill_manual(values=custom.col)

GM12878.DLs.TCs.p %>% as.data.frame(row.names=NULL) %>% subset(dhs.olap>=1 & score>=3) %>%
  ggplot(aes(directionality,fill=txType)) + geom_histogram(bins=25, position="dodge") + 
  facet_wrap(~sample, scales="free_y") + theme_bw() +
  labs(title="bidir score over DLs with DHS overlap & score>=3", fill="annotation") + scale_fill_manual(values=custom.col)

dev.off()


```

