---
title: "CTSS processing including NET-CAGE"
date: "2026-02-25"
output: html_document
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
base::library(PRIME)
```

## [0.1.] Register server cores / threads
```{r}
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 10
    )
  )
```

# [2.] Process GM12878 datasets -------------------------

## [2.0.] Read in GM12878 design matrix
```{r}
design_ <- utils::read.table(
    file =  base::paste0(
      here::here("0.External_resources/NET_CAGE/"),
        "GM12878.design.txt"
      ),
    sep = "\t",
    header = TRUE
  )
```

## [2.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::filter(use == "yes") %>%
  dplyr::mutate(
      rn = base::paste(ID,Seq_run,sep="_")
    ) %>% 
  tibble::column_to_rownames(var = "rn")
```

## [2.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [2.3.] Quantify CTSSs with no reference genome set
```{r}
GM12878.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )

GM12878.CTSSs <- calcTotalTags(GM12878.CTSSs)
```

## [2.4.] Visualize amount of tags across libraries
```{r}
SummarizedExperiment::colData(GM12878.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      ID = base::paste(Cell.line,Type,replicate,sep = "_"),
      ID = forcats::as_factor(x = ID)
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = totalTags/1000000,
        y = ID #base::paste(Type,replicate,sep="_")
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_y_discrete(
      limits = rev
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 2,
      axis.title.y = element_blank()
    )
```

## [2.5.] Pool replicates
#### Evaluate with SummarizedExperiment::colData(k562.p.CTSSs), SummarizedExperiment::colData(k562.CTSSs)
```{r}
keep <- GenomeInfoDb::seqlevels(GM12878.CTSSs)[1:23]

GM12878.CTSSs <- GenomeInfoDb::keepSeqlevels(GM12878.CTSSs, keep, pruning.mode = "coarse")


GM12878.CTSSs <- GM12878.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()

GM12878.p.CTSSs <- PRIME::poolReplicates(
  object = GM12878.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(GM12878.CTSSs)$Cell.line,
      SummarizedExperiment::colData(GM12878.CTSSs)$cell_count,
      SummarizedExperiment::colData(GM12878.CTSSs)$Type,
      sep = "_"
    )
  )

GM12878.p.CTSSs <- CAGEfightR::calcTPM(GM12878.p.CTSSs)
```

## [2.6.] Pool only based on reqsequencing
```{r}
GM12878.pID.CTSSs <- PRIME::poolReplicates(
    object = GM12878.CTSSs,
    replicates = SummarizedExperiment::colData(GM12878.CTSSs)$ID
  )

GM12878.pID.CTSSs <- CAGEfightR::calcTPM(GM12878.pID.CTSSs)
```

## [2.7.] Create combined objects
# Create object that also contains fully pooled samples.
#### Main indicates if part of main analysis. Removing Triptolide and low input
```{r}
SummarizedExperiment::colData(GM12878.pID.CTSSs)$main <- "yes"
SummarizedExperiment::colData(GM12878.pID.CTSSs)$main[
    SummarizedExperiment::colData(GM12878.pID.CTSSs)$Type %in% base::c("Cap_align","ctrl","trip") |
    SummarizedExperiment::colData(GM12878.pID.CTSSs)$cell_count %in% base::c("50K","250K,250K")
  ] <- "no"
SummarizedExperiment::colData(GM12878.pID.CTSSs)$pooled <- "no"
```

```{r}
SummarizedExperiment::colData(GM12878.p.CTSSs)$main <- "yes"
SummarizedExperiment::colData(GM12878.p.CTSSs)$main[
    SummarizedExperiment::colData(GM12878.p.CTSSs)$Type %in% base::c("Cap_align","ctrl,ctrl,ctrl","trip,trip,trip,trip") |
    SummarizedExperiment::colData(GM12878.p.CTSSs)$cell_count %in% base::c("50K,50K,50K","250K,250K,250K,250K,250K,250K")
  ] <- "no"
SummarizedExperiment::colData(GM12878.p.CTSSs)$pooled <- "yes"
```

```{r}
GM12878.combined.CTSSs <- cbind(GM12878.p.CTSSs,GM12878.pID.CTSSs)
GM12878.combined.CTSSs <- calcTotalTags(GM12878.combined.CTSSs)
```

## [2.8.] Create two subsamples one to the min replicate depth and one to min pool depth and keep original
```{r}
rep.min <- base::min(
  SummarizedExperiment::colData(GM12878.combined.CTSSs)$totalTags[
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$main == "yes" &
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$pooled == "no"
    ]
  ) #29027358
pool.min <- base::min(
  SummarizedExperiment::colData(GM12878.combined.CTSSs)$totalTags[
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$main == "yes" &
      SummarizedExperiment::colData(GM12878.combined.CTSSs)$pooled == "yes"
    ]
  ) #106618677

GM12878.combined.repSub.CTSSs<- subsampleTarget(
    object = GM12878.combined.CTSSs,
    target = rep.min
  )
GM12878.combined.poolSub.CTSSs<- subsampleTarget(
      object = GM12878.combined.CTSSs,
      target = pool.min
    )

GM12878.combined.repSub.CTSSs <- CAGEfightR::calcTPM(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- CAGEfightR::calcTPM(GM12878.combined.poolSub.CTSSs)
```

```{r}
GM12878.combined.repSub.CTSSs <- rmSingletons(GM12878.combined.repSub.CTSSs)
GM12878.combined.poolSub.CTSSs <- rmSingletons(GM12878.combined.poolSub.CTSSs)
```

## [2.9.] Save CTSS objects
```{r}
base::saveRDS(
    object = GM12878.combined.CTSSs,
    file = "all.combined.GM12878.CTSSs.rds")

base::saveRDS(
  object = GM12878.combined.repSub.CTSSs,
  file ="repSub.combined.GM12878.CTSSs.rmSingletons.rds")

base::saveRDS(
  object = GM12878.combined.poolSub.CTSSs,
  file ="poolSub.combined.GM12878.CTSSs.rmSingletons.rds")
here::here()
```
