---
title: "CTSS annnotating including NET-CAGE"
date: "2026-02-25"
output: html_document
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(BiocParallel)
```

## [0.1.] Load CAGEfighR extensions
```{r message=FALSE,results='hide'}
sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
fs <- base::c(
    "utils","support","cumulative",
    "decompose","enhancers","noise",
    "coverage","complexity","subsample",
    "dispersion"
  )

purrr::map(
  .x = fs,
  .f = function(f){
    base::source(
      base::paste0(
        cfr.extensions.dir, sep, f, ".R"
        )
      )
    }
  )
```

## [0.2.] Load annotations
```{r message=FALSE}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.3.] Add custom color palette
```{r}
custom.col <- base::c(
    "#FFDB6D","#C4961A","#F4EDCA",
    "#D16103","#C3D7A4","#52854C",
    "#4E84C4","#293352","grey"
  )
custom.col.ord <- base::c(
    "#FFDB6D","#F4EDCA","#52854C",
    "#4E84C4","#293352","#C3D7A4",
    "#D16103","grey",   "#C4961A"
  )
cols3 <- base::c(
    "#ED4938","#ADC607","#F4C10D",
    "#333136","#161474","#7B3DBA",
    "#03CEC2","#2AB348","#EDA838",
    "#7297C7","#EB373C","#FFE45F"
  )
gm.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#DF613F","#DF5643","#DD4425",
    "#92B9C9","#76A5B6","#004352",
    "#29A957","#008E37","#007A30",
    "#4F5D5C","#444E4C","#3D4138",
    "#AC97C2","#977FB4","#7D62A3"
  )
gm.pooled.pal <- base::c(
    "#FF8D00","#FF4425","#00439B",
    "#009072","#878784","#7D62D0"
  )
gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0"
  )
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#4F5D5C"
  )
```

## [0.4.] Add additional functions
```{r}
calcAnnoCTSS <- function(data, txModels, uniqueCTSS = FALSE) {
    
    annos <- CAGEfightR::assignTxType(
        object = data,
        txModels = txModels
      )
    
    #Identify all txTypes
    types <- base::unique(SummarizedExperiment::rowData(annos)$txType)
    
    # Initialize list to store total tags for each txType
    totalTags_list <- base::vector("list", length = base::length(types))
    
    # Loop through each txType
    for (i in base::seq_along(types)) {
      
        # Subset data based on current txType
        current_subset <- S4Vectors::subset(annos, txType %in% types[i])
        if(uniqueCTSS){
          
            # Calculate totalTags for current subset
            current_subset <- Matrix::colSums(
                SummarizedExperiment::assay(current_subset) > 0
              ) 
            totalTags_list[[i]] <- current_subset
            
          } else {
            
            # Calculate totalTags for current subset
            current_subset <- SummarizedExperiment::colData(
                CAGEfightR::calcTotalTags(current_subset)
              )
            totalTags_list[[i]] <- current_subset$totalTags
            
          }
      }
    
    # Combine totalTags into a data frame
    ann_counts <- base::as.data.frame(
        base::do.call(base::cbind, totalTags_list)
      )
    base::colnames(ann_counts) <- types
    return(ann_counts)
  }
```

## [0.5.] Read in metainfo
```{r}
meta <- readr::read_delim(
    file = base::paste0(
        here::here("0.External_resources/"),
        "20260225_metainfo_GM12878libraries_v1.csv"
      ),
    delim = ";",
    col_names = TRUE,
    col_types = readr::cols(
      .default = readr::col_character()
    )
  )
```

# [2.] GM12878 analysis -------------------------

## [2.0.] Read in datasets for GM12878
#### see [1.10.] in "1.1.CTSSs.processing.Rmd"
```{r}
# GM12878.s.CTSSs <- base::readRDS(
#   file = base::paste0(
#       here::here("0.External_resources/NET_CAGE/"),
#       "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
#     )
#   )
GM12878.p.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("0.External_resources/NET_CAGE/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```

## [2.1.] Quantify CTSS level annotations
```{r}
GM12878.ann.counts <- .GlobalEnv$calcAnnoCTSS(
    data = GM12878.p.CTSSs,
    txModels = txdb,
    uniqueCTSS = FALSE # TRUE
  )

GM12878.ann.fracs <- GM12878.ann.counts/SummarizedExperiment::colData(GM12878.p.CTSSs)$totalTags
```

## [2.2.] Factor levels for genomic features & cell amounts
```{r}
feats <- base::c(
    "intergenic","proximal","promoter",    
    "fiveUTR","CDS","exon",
    "intron","threeUTR","antisense"
  )
amts <- base::c("50K","250K","500K","1M","5M","10M","20M")
reps <- base::c("pool","1","2","3","4")
scrs <- base::c("C","N","NET","trip","trip_ctrl","cap_align","groCap")

# sams <- base::c(
#     "500K_C","500K_N","1M_C",
#     "1M_N","5M_N","10M_N"
#   )
```

## [2.3.] Data wrangling - prepare plotting of fractions
```{r}
GM12878.ann.fracs.m <- GM12878.ann.fracs %>%
  tibble::rownames_to_column("sample") %>%
  tidyr::pivot_longer(
      cols = !sample,
      names_to = "variable",
      values_to = "value"
    ) %>%
  dplyr::left_join(
      x = ., y = meta, by = "sample"
    ) %>%
  # dplyr::mutate(
  #     cell = stringr::str_replace(sample, "^(.*?)_.*$", "\\1"),
  #     is_NET = stringr::str_detect(sample, "_NET_"),
  #     
  #     # NET samples
  #     rep = ifelse(is_NET, sub("^.*_(\\d+)$", "\\1", sample), NA),
  #     source = ifelse(is_NET, "NET", NA),
  #     amount = ifelse(is_NET, "20M", NA),
  #     
  #     # classic samples
  #     amount = ifelse(!is_NET, sub("^[^_]+_([^_]+)_.*$", "\\1", sample), amount),
  #     type   = ifelse(!is_NET, sub("^[^_]+_[^_]+_(.*)$", "\\1", sample), NA),
  #     source = ifelse(!is_NET, substr(type,1,1), source),
  #     rep    = ifelse(!is_NET, substr(type,2,nchar(type)), rep)
  #   ) %>%
  # dplyr::select(-type, -is_NET) %>%
  mutate(
    rep = base::factor(
        x = rep,
        levels = base::rev(reps),
        ordered = TRUE
      ),
    amount = base::factor(
        x = amount,
        levels = amts,
        ordered = TRUE
      ),
    source = base::factor(
        x = source,
        levels = scrs,
        ordered = TRUE
      ),
    variable = base::factor(
        x = variable,
        levels = base::rev(feats),
        ordered = TRUE
      )
    )
```

## [2.4.] Plotting of fractions
```{r}
ggplot2::ggplot(
    data = GM12878.ann.fracs.m,
    mapping = aes(
        x = value,
        y = rep,
        fill = variable
      )
    ) +
  geom_col() +
  facet_wrap(
      facets = source ~ amount,
      ncol = 1,
      strip.position = "right",
      axis.labels = "margins",
      axes = "margins"
    ) +
  theme_bw() +
  scale_fill_manual(
      values = base::rev(custom.col.ord) #base::rev(custom.col)
    ) +
  theme(
      axis.title = element_blank(),
      panel.spacing = grid::unit(0.1,"lines"),
      aspect.ratio = 1/6,
      # legend.position = "none"
      legend.title = element_blank(),
      legend.background = element_blank()
    )
```

## [2.5.] Print plot to file
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("0.External_resources/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_annFrac",
        "_AllSamples_AllReplicates",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 5,
    height = 8
  )
```
