---
title: "Plotting nascent enrichment in DHSs"
output: html_document
date: "2024-07-04"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(dplyr)
base::library(ggplot2)
base::library(tidyverse)
base::library(ggrastr)
```

## [0.2.] Add custom color palettes
```{r}
custom.col <- base::c(
  "#FFDB6D","#C4961A","#F4EDCA", 
  "#D16103","#C3D7A4","#52854C",
  "#4E84C4","#293352","grey"
  )


gm.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#DF613F","#DF5643","#DD4425",
  "#92B9C9","#76A5B6","#004352",
  "#29A957","#008E37","#007A30",
  "#4F5D5C","#444E4C","#3D4138",
  "#AC97C2","#977FB4","#7D62A3"
  )
                     
gm.pooled.pal <- base::c(
  "#FF8D00","#FF4425","#00439B",
  "#009072","#878784","#7D62D0"
  )

gm.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#92B9C9",
  "#76A5B6","#004352","#00439B",
  "#29A957","#008E37","#007A30",
  "#009072","#4F5D5C","#444E4C",
  "#3D4138","#878784","#AC97C2",
  "#977FB4","#7D62A3","#7D62D0"
  )

k.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#4F5D5C",
  "#92B9C9","#76A5B6","#004352",
  "#00439B","#29A957","#008E37",
  "#007A30"
  )
```

## [0.3.] Set working directory
```{r}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/2.Cellular_localization/")
```

# [1.] Plotting GM12878 pooled subsetted DHSs quantifications -------------------------

## [1.0.] Read DHS quantified regions
```{r}
GM12878.dhs.p <- base::readRDS(
    file = "poolSub.DHSs.quantified.GM12878.rds"
  )
## Rep subsetted used for gro-cap comparison
GM12878.dhs <- base::readRDS(
    file = "poolSub.DHSs.quantified.GM12878.rds"
  )
```

## [1.1.] Prepare plotting data frame with relevant metrics
```{r}
gm12878.expr.df.p <- base::as.data.frame(
    SummarizedExperiment::assay(GM12878.dhs.p, "TPM")
  )
gm12878.expr.df.p$txType <- SummarizedExperiment::rowData(GM12878.dhs.p)$txType

gm12878.expr.df.p$tripMean <- base::rowMeans(
    gm12878.expr.df.p[,base::c("GM12878_1M_trip","GM12878_1M_ctrl")]
  )

gm12878.expr.df.p$tripFC <- base::log2(
    base::c(gm12878.expr.df.p$GM12878_1M_trip+0.1)/(gm12878.expr.df.p$GM12878_1M_ctrl+0.1)
  )

gm12878.expr.df.p$groFC <- base::log2(
    base::c(gm12878.expr.df.p$GM12878_groCap_1+0.1)/(gm12878.expr.df.p$GM12878_1M_ctrl+0.1)
  )
```

## [1.1.] Plot C vs N with triptolide FC and gro-cap coloring, split between gene proximal and distal DHSs
```{r}
gm12878.expr.df.p$grp <- "proximal"
gm12878.expr.df.p$grp[gm12878.expr.df.p$txType %in% base::c("intron","intergenic")] <- "distal"
gm12878.expr.df.p$grp <- base::factor(gm12878.expr.df.p$grp, levels = base::c("proximal", "distal"))


gm12878.expr.df.p.m <- gm12878.expr.df.p %>%
  dplyr::select(
      "grp", "tripFC", "groFC",
      "GM12878_500K_C","GM12878_1M_N","GM12878_5M_N"
    ) %>%
  tidyr::pivot_longer(
      cols = base::c("GM12878_1M_N","GM12878_5M_N")
    )
```

## [1.2.] Prepare order of plotted points
```{r}
plot.df <- gm12878.expr.df.p.m %>% 
  dplyr::filter(
      GM12878_500K_C > 0 | value > 0
    ) %>%
  dplyr::mutate(
      tripGrp = dplyr::case_when(
          tripFC <= 0 ~ "low",
          tripFC > 0 ~ "high"
        ),
      tripGrp = base::factor(
        x = tripGrp,
        levels = base::c("low","high"),
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(tripGrp) %>% 
  dplyr::arrange(
      tripGrp, base::abs(tripFC)
    )
  # dplyr::filter(
  #   row_number() %in% base::c(1,dplyr::n())
  # )
```

## [1.3.] Plot data
```{r}
plot <- ggplot2::ggplot(
  data = plot.df,
     aes(
         x = GM12878_500K_C,
         y = value,
         fill = tripFC
       )
     ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.01
    ) +
  # geom_density_2d(
  #     color = "black"
  #   ) +
  scale_y_log10() +
  scale_x_log10() +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  geom_abline(
      intercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "500K cells CAGE [TPM]",
      y = "nuclei CAGE [TPM]"
    ) +
  facet_wrap(
      facets = tripGrp ~ grp + name,
      # facets =  ~ grp + name,
      nrow = 2
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(
          size = 8
        ),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

## [1.4.] Plot ratio vs. triptolide signal
```{r}
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  ggplot2::ggplot(
  data = .,
     aes(
         x = log2(ratio),
         y = tripFC,
         fill = tripFC
       )
     ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.05
    ) +
  geom_density_2d(
      color = "black",
      bins = 50,
      alpha = 0.5
    ) +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  geom_vline(
      xintercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "nuclei / 500K cells",
      y = "triptolide FC"
    ) +
  facet_grid(
    rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(
          size = 8
        ),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [1.4.] Print plots
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.GM12878.DHSs.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```

## [1.5.] Prepare groFC plotting
```{r}
plot.df <- gm12878.expr.df.p.m %>% 
  dplyr::filter(
      GM12878_500K_C > 0 | value > 0
    ) %>%
  dplyr::mutate(
      groGrp = dplyr::case_when(
          groFC <= 0 ~ "low",
          groFC > 0 ~ "high"
        ),
      groGrp = base::factor(
        x = groGrp,
        levels = base::c("low","high"),
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(groGrp) %>% 
  dplyr::arrange(
      groGrp, base::abs(groFC)
    )
```

## [1.6.] Plot groFC onto nuclei-vs.cell CAGE comparison
```{r}
plot <- ggplot2::ggplot(
    data = plot.df,
    mapping = aes(
        x = GM12878_500K_C,
        y = value,
        fill = groFC
      )
    ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.01
    ) +
  geom_abline(
      intercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "500K cells CAGE [TPM]",
      y = "nuclei CAGE [TPM]"
    ) +
  facet_wrap(
      facets = groGrp ~ grp + name,
      nrow = 2
    ) +
  scale_y_log10() +
  scale_x_log10() +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

```{r}
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  ggplot2::ggplot(
  data = .,
     aes(
         x = log2(ratio),
         y = groFC,
         fill = groFC
       )
     ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.05
    ) +
  geom_density_2d(
      color = "black",
      bins = 30,
      alpha = 0.5
    ) +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  geom_vline(
      xintercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "nuclei / 500K cells",
      y = "triptolide FC"
    ) +
  facet_grid(
    rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(
          size = 8
        ),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.GM12878.DHSs.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```






# [2.] Plotting GM12878 pooled subsetted DHSs quantifications

## [2.0.] Read DHS quantified regions
```{r}
GM12878.dhs <- base::readRDS("poolSub.DHSs.quantified.GM12878.rds")
```


## [2.1.] Prepare plotting data frame with relevant metrics
```{r}
gm12878.expr.df <- base::as.data.frame(assay(GM12878.dhs, "TPM"))
gm12878.expr.df$txType <- SummarizedExperiment::rowData(GM12878.dhs)$txType

gm12878.expr.df$tripMean <- base::rowMeans(
  gm12878.expr.df[,c("GM12878_1M_trip","GM12878_1M_ctrl")])

gm12878.expr.df$tripFC <- base::log2(
  c(gm12878.expr.df$GM12878_1M_trip+0.1)/(gm12878.expr.df$GM12878_1M_ctrl+0.1))
```


## [2.1.] Plot C vs N with triptolide FC and gro-cap coloring, split between gene proximal and distal DHSs
```{r}
gm12878.expr.df$grp <- "proximal"
gm12878.expr.df$grp[gm12878.expr.df$txType %in% c("intron","intergenic")] <- "distal"
gm12878.expr.df$grp <- base::factor(gm12878.expr.df$grp, levels = c("proximal", "distal"))

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="proximal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=tripFC)) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white")  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.proximal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="distal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=tripFC)) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white")  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.distal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }


ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="proximal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=log10(0.1+GM12878_groCap_1))) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white", midpoint = c(-0.25))  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.proximal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="distal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=log10(0.1+GM12878_groCap_1))) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white", midpoint = c(-0.25))  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.distal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```













