---
title: "Plotting nascent enrichment in DHSs"
output: html_document
date: "2024-07-04"
---

## [0.] Prerequsites -------------------------

## [-1.0.] Install necessary packages
```{r}
remotes::install_github("easystats/see")
utils::install.packages("gghalves")
```

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(tidyverse)
```

# [1.] Plotting GM12878 pooled subsetted DHSs quantifications -------------------------

## [1.0.] Read DHS quantified regions
#### Evaluate with 'SummarizedExperiment::colData(GM12878.dhs.p) %>% tibble::as_tibble(rownames = "type")'
```{r}
GM12878.dhs.p <- base::readRDS(
    file = "poolSub.DHSs.quantified.GM12878.rds"
  )
```

#### Rep subsetted used for gro-cap comparison
```{r eval=FALSE}
# GM12878.dhs <- base::readRDS(
#     file = "poolSub.DHSs.quantified.GM12878.rds"
#   )
```

## [1.1.] Transform RangedSummarizedExperiment into a dataframe
```{r}
gm12878.expr.df.p <- base::as.data.frame(
    SummarizedExperiment::assay(GM12878.dhs.p, "TPM")
  )
gm12878.expr.df.p$txType <- SummarizedExperiment::rowData(GM12878.dhs.p)$txType

gm12878.expr.df.p$tripMean <- base::rowMeans(
    gm12878.expr.df.p[,base::c("GM12878_1M_trip","GM12878_1M_ctrl")]
  )

gm12878.expr.df.p$tripFC <- base::log2(
    base::c(gm12878.expr.df.p$GM12878_1M_trip+0.1)/(gm12878.expr.df.p$GM12878_1M_ctrl+0.1)
  )

gm12878.expr.df.p$groFC <- base::log2(
    base::c(gm12878.expr.df.p$GM12878_groCap_1+0.1)/(gm12878.expr.df.p$GM12878_1M_ctrl+0.1)
  )

# SummarizedExperiment::rowRanges(GM12878.dhs.p)
gm12878.expr.df.p$loc <- base::paste0(
    GenomeInfoDb::seqnames(GM12878.dhs.p)@values, ":",
    BiocGenerics::start(GM12878.dhs.p), "-",
    BiocGenerics::end(GM12878.dhs.p)
  )
```

## [1.1.] Split between gene proximal and distal DHSs
```{r}
gm12878.expr.df.p$grp <- "proximal"
gm12878.expr.df.p$grp[
    gm12878.expr.df.p$txType %in% base::c("intron","intergenic")
  ] <- "distal"
gm12878.expr.df.p$grp <- base::factor(
    x = gm12878.expr.df.p$grp,
    levels = base::c("distal", "proximal"),
    ordered = TRUE
  )

gm12878.expr.df.p.m <- gm12878.expr.df.p %>%
  dplyr::select(
      "loc", "grp", "tripFC", "groFC",
      "GM12878_500K_C","GM12878_1M_N","GM12878_5M_N",
    ) %>%
  tidyr::pivot_longer(
      cols = base::c("GM12878_1M_N","GM12878_5M_N")
    )
```

## [1.2.] Prepare order of plotted points
```{r}
plot.df <- gm12878.expr.df.p.m %>% 
  dplyr::filter(
      GM12878_500K_C > 0 | value > 0
    ) %>%
  dplyr::mutate(
      tripGrp = dplyr::case_when(
          tripFC <= 0 ~ "low",
          tripFC > 0 ~ "high"
        ),
      tripGrp = base::factor(
        x = tripGrp,
        levels = base::c("low","high"),
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(tripGrp) %>% 
  dplyr::arrange(
      tripGrp, base::abs(tripFC)
    )
  # dplyr::filter(
  #   row_number() %in% base::c(1,dplyr::n())
  # )
```

## [1.3.] Plot data
```{r}
plot <- ggplot2::ggplot(
  data = plot.df,
     aes(
         x = GM12878_500K_C,
         y = value,
         fill = tripFC
       )
     ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.01
    ) +
  # geom_density_2d(
  #     color = "black"
  #   ) +
  scale_y_log10() +
  scale_x_log10() +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  geom_abline(
      intercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "500K cells CAGE [TPM]",
      y = "nuclei CAGE [TPM]"
    ) +
  facet_wrap(
      facets = tripGrp ~ grp + name,
      # facets =  ~ grp + name,
      nrow = 2
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(
          size = 8
        ),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

## [1.4.] Prepare annotations for triptolide plot
```{r}
annos <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  dplyr::group_by(name,grp) %>%
    dplyr::summarise(
        low_Left = base::sum(base::log2(ratio) < -1 & tripFC < -1, na.rm = TRUE),
        up_Left = base::sum(base::log2(ratio) < -1 & tripFC > 1, na.rm = TRUE),
        low_Right = base::sum(base::log2(ratio) > 1 & tripFC < -1, na.rm = TRUE),
        up_Right = base::sum(base::log2(ratio) > 1 & tripFC > 1, na.rm = TRUE),
        total = base::sum(tripFC < Inf, na.rm = TRUE),
        .groups = "drop"
      ) %>%
    tidyr::pivot_longer(
        cols = tidyr::contains("_"),
        names_to = "quadr",
        values_to = "count"
      ) %>%
    dplyr::mutate(
        xpos = base::rep(base::c(-Inf,-Inf,Inf,Inf),4),
        ypos = base::rep(base::c(-Inf, Inf,-Inf,Inf),4),
        hjustvar = base::rep(base::c(0,0,1,1),4),
        vjustvar = base::rep(base::c(0,1,0,1),4)
      )
```

## [1.5.] Plot ratio vs. triptolide signal
```{r}
coef = 0.1
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

plot <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
         # fill = tripFC,
         x = base::log2(ratio),
         y = tripFC
       )
     ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.05,
      fill = "white"
    ) +
  # geom_density_2d(
  #     color = "black",
  #     bins = 50,
  #     alpha = 0.5
  #   ) +
  geom_density2d(
      mapping = aes(
          color = ..level..
        ),
      # binwidth = 0.1,
      bins = 30,
      size = 0.5
    ) +
  geom_vline(
      xintercept = base::c(-1,1),
      linetype = "dashed"
    ) +
  geom_hline(
      yintercept = base::c(-1,1),
      linetype = "dashed"
    ) +
  geom_text(
      data = annos,
      mapping = aes(
        x = xpos,
        y = ypos,
        hjust = hjustvar,
        vjust = vjustvar,
        label = count
      )
    ) +
  scale_colour_gradientn(
      colours = cols
    ) +
  # scale_fill_gradient2(
  #     low = "blue",
  #     high = "darkred",
  #     mid = "white"
  #   )  +
  labs(
      x = "nuclei / 500K cells",
      y = "triptolide FC"
    ) +
  facet_grid(
      rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(
      #     size = 8
      #   ),
      # legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

## [1.6.] Print plots
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("2.Cellular_localization/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.GM12878.DHSs.main",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.7.] Prepare stats for plot
#### Run Wilcoxon rank sum test
```{r}
stats <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      ratGrp = dplyr::case_when(
        ratio < 0 ~ "neg",
        ratio >= 0 ~ "pos"
      )
    ) %>%
  dplyr::filter(
      ratio <= -1 | ratio >= 1
    ) %>%
  dplyr::group_by(grp, name) %>%
  tidyr::nest() %>%
  dplyr::mutate(
    wrst = purrr::map_dbl(
        .x = data,
        .f = function(df){
          return(
            stats::wilcox.test(
              formula = tripFC ~ ratGrp,
              data = df
            )$p.value
          )
        }
      )
    ) %>%
  dplyr::select(-data) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
      xpos = base::rep(1.5,4),
      ypos = base::rep(-6.5,4)
    )
```

## [1.8.] Plot data distributions
```{r}
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      ratGrp = dplyr::case_when(
        ratio < 0 ~ "neg",
        ratio >= 0 ~ "pos"
      )
    ) %>%
  dplyr::filter(
      ratio <= -1 | ratio >= 1
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = ratGrp,
        y = tripFC
      )
    ) +
  gghalves::geom_half_violin(
      nudge = 0.025
    ) +
  gghalves::geom_half_boxplot(
      side = "r",
      outlier.shape = NA,
      nudge = 0.025
    ) +
   geom_text(
      data = stats,
      mapping = aes(
        x = xpos,
        y = ypos,
        label = wrst
      )
    ) +
  ylim(-7.5,3.1) +
  # geom_violin() +
  # geom_boxplot() +
  # see::geom_violinhalf(
  #     position = position_nudge(
  #       x = 0.15,
  #       y = 0
  #     )
  #   ) +
  # geom_jitter(
  #     shape = 21,
  #     color = "black",
  #     alpha = 0.1,
  #     stroke = 0.05,
  #     height = 0,
  #     width = 0.1
  #   ) +
  facet_grid(
      rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [1.9.] Print plots
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("2.Cellular_localization/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.GM12878.DHSs.main",
          "_statistics",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

```{r}
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      intg = base::sqrt(ratio^2 + tripFC^2)
    ) %>%
  dplyr::filter(  
      ratio > 1 &
      tripFC < -1 &
      !base::is.infinite(ratio) &
      grp == "proximal" &
      name == "GM12878_5M_N"
    ) %>%
  dplyr::arrange(dplyr::desc(intg))


#chr1:120850732-120851131
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      intg = base::sqrt(ratio^2 + tripFC^2)
    ) %>%
  dplyr::filter(  
      ratio > 1 &
      tripFC < -1 &
      GM12878_500K_C > 1 &
      value > 1 & 
      !base::is.infinite(ratio) &
      grp == "proximal" &
      name == "GM12878_5M_N"
    ) %>%
  dplyr::arrange(dplyr::desc(intg))
```

## [1.10.] Prepare groFC plotting
```{r}
plot.df <- gm12878.expr.df.p.m %>% 
  dplyr::filter(
      GM12878_500K_C > 0 | value > 0
    ) %>%
  dplyr::mutate(
      groGrp = dplyr::case_when(
          groFC <= 0 ~ "low",
          groFC > 0 ~ "high"
        ),
      groGrp = base::factor(
        x = groGrp,
        levels = base::c("low","high"),
        ordered = TRUE
      )
    ) %>%
  dplyr::group_by(groGrp) %>% 
  dplyr::arrange(
      groGrp, base::abs(groFC)
    )
```

## [1.10.] Plot groFC onto nuclei-vs.cell CAGE comparison
```{r}
plot <- ggplot2::ggplot(
    data = plot.df,
    mapping = aes(
        x = GM12878_500K_C,
        y = value,
        fill = groFC
      )
    ) + 
  geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.01
    ) +
  geom_abline(
      intercept = 0,
      linetype = "dashed"
    ) +
  labs(
      x = "500K cells CAGE [TPM]",
      y = "nuclei CAGE [TPM]"
    ) +
  facet_wrap(
      facets = groGrp ~ grp + name,
      nrow = 2
    ) +
  scale_y_log10() +
  scale_x_log10() +
  scale_fill_gradient2(
      low = "blue",
      high = "darkred",
      mid = "white"
    )  +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

## [1.11.] Prepare annotations for groFC plot
```{r}
annos <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  dplyr::group_by(name,grp) %>%
    dplyr::summarise(
        low_Left = base::sum(base::log2(ratio) < -1 & groFC < -1, na.rm = TRUE),
        up_Left = base::sum(base::log2(ratio) < -1 & groFC > 1, na.rm = TRUE),
        low_Right = base::sum(base::log2(ratio) > 1 & groFC < -1, na.rm = TRUE),
        up_Right = base::sum(base::log2(ratio) > 1 & groFC > 1, na.rm = TRUE),
        total = base::sum(tripFC < Inf, na.rm = TRUE),
        .groups = "drop"
      ) %>%
    tidyr::pivot_longer(
        cols = tidyr::contains("_"),
        names_to = "quadr",
        values_to = "count"
      ) %>%
    dplyr::mutate(
        xpos = base::rep(base::c(-Inf,-Inf,Inf,Inf),4),
        ypos = base::rep(base::c(-Inf, Inf,-Inf,Inf),4),
        hjustvar = base::rep(base::c(0,0,1,1),4),
        vjustvar = base::rep(base::c(0,1,0,1),4)
      )
```

## [1.12.] Plot groFC onto nuclei-vs.cell CAGE comparison
```{r}
coef = 0.1
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

plot <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = value / GM12878_500K_C
    ) %>%
  ggplot2::ggplot(
     data = .,
     mapping = aes(
         x = base::log2(ratio),
         y = groFC
       )
     ) + 
    geom_point(
      shape = 21,
      color = "black",
      # alpha = 0.1,
      stroke = 0.05,
      fill = "white"
    ) +
  # geom_density_2d(
  #     color = "black",
  #     bins = 50,
  #     alpha = 0.5
  #   ) +
  geom_density2d(
      mapping = aes(
          color = ..level..
        ),
      # binwidth = 0.1,
      bins = 15,
      size = 0.5
    ) +
  geom_vline(
      xintercept = base::c(-1,1),
      linetype = "dashed"
    ) +
  geom_hline(
      yintercept = base::c(-1,1),
      linetype = "dashed"
    ) +
  geom_text(
      data = annos,
      mapping = aes(
        x = xpos,
        y = ypos,
        hjust = hjustvar,
        vjust = vjustvar,
        label = count
      )
    ) +
  scale_colour_gradientn(
      colours = cols
    ) +
  # scale_fill_gradient2(
  #     low = "blue",
  #     high = "darkred",
  #     mid = "white"
  #   )  +
  labs(
      x = "nuclei / 500K cells",
      y = "groFC"
    ) +
  facet_grid(
      rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(
      #     size = 8
      #   ),
      # legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Point",
    dpi = 300
  )
```

## [1.13.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("2.Cellular_localization/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.GM12878.DHSs.main",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [1.14.] Prepare stats for plot
#### Run Wilcoxon rank sum test
```{r}
stats <- gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      ratGrp = dplyr::case_when(
        ratio < 0 ~ "neg",
        ratio >= 0 ~ "pos"
      )
    ) %>%
  dplyr::filter(
      ratio <= -1 | ratio >= 1
    ) %>%
  dplyr::group_by(grp, name) %>%
  tidyr::nest() %>%
  dplyr::mutate(
    wrst = purrr::map_dbl(
        .x = data,
        .f = function(df){
          return(
            stats::wilcox.test(
              formula = groFC ~ ratGrp,
              data = df
            )$p.value
          )
        }
      )
    ) %>%
  dplyr::select(-data) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
      xpos = base::rep(1.5,4),
      ypos = base::rep(-6.5,4)
    )
```

## [1.15.] Plot data distributions
```{r}
gm12878.expr.df.p.m %>%
  dplyr::mutate(
      ratio = base::log2(value / GM12878_500K_C),
      ratGrp = dplyr::case_when(
        ratio < 0 ~ "neg",
        ratio >= 0 ~ "pos"
      )
    ) %>%
  dplyr::filter(
      ratio <= -1 | ratio >= 1
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = ratGrp,
        y = groFC
      )
    ) +
  gghalves::geom_half_violin(
      nudge = 0.025
    ) +
  gghalves::geom_half_boxplot(
      side = "r",
      outlier.shape = NA,
      nudge = 0.025
    ) +
   geom_text(
      data = stats,
      mapping = aes(
        x = xpos,
        y = ypos,
        label = wrst
      )
    ) +
  ylim(-7.5,3.1) +
  # geom_violin() +
  # geom_boxplot() +
  # see::geom_violinhalf(
  #     position = position_nudge(
  #       x = 0.15,
  #       y = 0
  #     )
  #   ) +
  # geom_jitter(
  #     shape = 21,
  #     color = "black",
  #     alpha = 0.1,
  #     stroke = 0.05,
  #     height = 0,
  #     width = 0.1
  #   ) +
  facet_grid(
      rows = name ~ grp
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [1.16.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("2.Cellular_localization/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.GM12878.DHSs.main",
          "_statistics",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

# [2.] Plotting GM12878 pooled subsetted DHSs quantifications -------------------------

## [2.0.] Read DHS quantified regions
```{r}
GM12878.dhs <- base::readRDS("poolSub.DHSs.quantified.GM12878.rds")
```

## [2.1.] Prepare plotting data frame with relevant metrics
```{r}
gm12878.expr.df <- base::as.data.frame(assay(GM12878.dhs, "TPM"))
gm12878.expr.df$txType <- SummarizedExperiment::rowData(GM12878.dhs)$txType

gm12878.expr.df$tripMean <- base::rowMeans(
  gm12878.expr.df[,c("GM12878_1M_trip","GM12878_1M_ctrl")])

gm12878.expr.df$tripFC <- base::log2(
  c(gm12878.expr.df$GM12878_1M_trip+0.1)/(gm12878.expr.df$GM12878_1M_ctrl+0.1))
```


## [2.1.] Plot C vs N with triptolide FC and gro-cap coloring, split between gene proximal and distal DHSs
```{r}
gm12878.expr.df$grp <- "proximal"
gm12878.expr.df$grp[gm12878.expr.df$txType %in% c("intron","intergenic")] <- "distal"
gm12878.expr.df$grp <- base::factor(gm12878.expr.df$grp, levels = c("proximal", "distal"))

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="proximal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=tripFC)) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white")  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.proximal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="distal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=tripFC)) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white")  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_tripFC.distal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }


ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="proximal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=log10(0.1+GM12878_groCap_1))) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white", midpoint = c(-0.25))  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.proximal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }

ggplot2::ggplot(gm12878.expr.df[(gm12878.expr.df$GM12878_1M_C>0 | gm12878.expr.df$GM12878_5M_N>0) & gm12878.expr.df$grp=="distal",],
       aes(GM12878_1M_C,GM12878_5M_N,color=log10(0.1+GM12878_groCap_1))) + 
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_colour_gradient2(low = "steelblue", high = "orange",mid="white", midpoint = c(-0.25))  +
  geom_abline(intercept = 0) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
      fill = "grey95"
        ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="1M cells CAGE [TPM]", y="5M nuclei CAGE [TPM]")

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_groCap.distal.GM12878.DHSs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```













