---
title: "20240417_workflow_fingerprints"
output: html_document
date: "2024-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [0.0.] Load necessary packages
```{r}
base::library(tidyverse)
```

## [0.1.] Read in metainfo
```{r}
meta <- readr::read_delim(
    file = base::paste0(
        here::here("1.CTSSs/"),
        "GM12878.design.txt"
      ),
    delim = "\t",
    col_names = TRUE,
    col_types = readr::cols(
      .default = readr::col_character()
    )
  )
```

## [1.0.] Read in data
```{r}
cnts_ <- readr::read_delim(
    file = base::paste0(
        here::here("6.BamFingerprints/"),
        "20240417_counts_fingerprints.txt"
      ),
    skip = 1,
    delim = "\t",
    col_names = TRUE,
    col_types = readr::cols(
      .default = readr::col_integer()
    )
  )
```

## [1.1.] Data wrangling
```{r}
cnts <- cnts_ %>%
  tidyr::pivot_longer(
      cols = dplyr::everything(),
      names_to = "sample",
      values_to = "count"
    ) %>%
  dplyr::mutate(
    sample = stringr::str_replace(
      sample, "^.*\\/bam_files\\/(.*)_STAR.*$","\\1"
    )
  )
```

## [1.2.] Calculate relative ranks & cumsums
#### All samples have 500435 entries, evaluate with 'dplyr::summarise(sms, len = dplyr::n())'
#### evaluate with 'base::max(fsms$relagsum)'
```{r}
sms <- cnts %>%
  dplyr::group_by(sample) %>%
  dplyr::arrange(count) %>%
  dplyr::mutate(
    agsum = base::cumsum(count),
    rank = dplyr::row_number(),
    relrank = rank/500435
  )

ref <- sms %>%
  dplyr::summarise(
      fcnt = base::max(agsum)
    )

fsms <- sms %>%
  tidyr::nest() %>% 
  dplyr::mutate(
    newdata = purrr::pmap(
      .l = base::list(sample,data),
      .f = function(s,df){
        c <- ref %>%
          dplyr::filter(sample == s) %>%
          dplyr::pull(fcnt)
        return(
          df %>%
            dplyr::mutate(
              relagsum = agsum/c
            )
          )
      }
    )
  ) %>%
  dplyr::select(-data) %>%
  tidyr::unnest(cols = "newdata")
```

## [1.3.] Test plotting of K562 samples
```{r}
fsms %>%
  dplyr::filter(
      stringr::str_detect(sample,"K562")
    ) %>%
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          x = relrank,
          y = relagsum,
          color = sample
        )
      ) +
    geom_line()
```

## [1.4.] Evaluate samples
```{r}
fsms %>% 
  dplyr::distinct(sample) %>%
  dplyr::filter(
      # stringr::str_detect(sample, "250k|50k|K562", negate = TRUE)
    stringr::str_detect(sample, "500k", negate = TRUE)
    ) %>%
  dplyr::pull(sample)
```

## [1.5.] Subset samples & separate sample name information
#### Evaluate with 'dplyr::distinct(fsmd,sample)', 'dplyr::distinct(fsmd,repseq)'
```{r}
samples <- meta %>%
  dplyr::filter(
      use == "yes" &
      stringr::str_detect(ID,"groCap",negate =TRUE)
    ) %>%
  dplyr::mutate(
    sam = stringr::str_replace(
        bw_plus,"^.*\\/bw_files\\/(.*)_STAR.*","\\1"
      )
    ) %>%
  dplyr::pull(sam)
  
fsmd <- fsms %>%
  dplyr::filter(sample %in% samples) %>% 
  dplyr::mutate(
    seq = stringr::str_extract(
        sample,"reseq"
      ),
    sample = stringr::str_replace_all(
        sample,"GM12878_|CAGE_reseq_|CAGE_",""
      )
    ) %>%
  tidyr::separate(
      col = sample,
      sep = "_",
      into = base::c("source","amount","rep"),
      remove = FALSE
    ) %>%
  dplyr::mutate(
    repseq = dplyr::case_when(
      seq == "reseq" ~ base::paste0(rep,"_",seq),
      TRUE ~ rep
    )
  )
```

## [1.6.] Convert arguments into factors
```{r}
amts <- base::c("500k","1mio","5mio","10mio")
curlst <- base::c(
    "C_500k_1", "C_500k_2", "C_500k_3",
    "C_1mio_1", "C_1mio_2", "C_1mio_3",
    "N_500k_1", "N_500k_2", "N_500k_3",
    "N_1mio_1", "N_1mio_2", "N_1mio_3",
    "N_5mio_1", "N_5mio_2", "N_5mio_3",
    "N_10mio_1", "N_10mio_2","N_10mio_3"
  )
```

```{r}
fsmp <- fsmd %>%
  dplyr::mutate(
    amount = base::factor(
        x = amount,
        levels = amts,
        ordered = TRUE
      ),
    source = base::c(
      x = source,
      levels = base::c("C","N"),
      ordered = TRUE
    )
  )
```


```{r}
  ggplot2::ggplot(
        data = fsmp,
        mapping = aes(
          x = relrank,
          y = relagsum,
          colour = rep
        )
      ) +
    geom_line() +
    facet_wrap(
        facets = amount ~ source + seq,
        nrow = 1
      ) +
    theme_bw() +
    theme(
        panel.spacing = grid::unit(0.1,"lines"),
        aspect.ratio = 1
      )
```

