---
title: "6.1. Fingerprint plots"
output: html_document
date: "2024-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [0.] Prerequisites -------------------------

## [0.0.] Load necessary packages
```{r message=FALSE}
base::library(PRIME)
base::library(tidyverse)
base::library(foreach)
base::library(doParallel)
doParallel::registerDoParallel(
    cores = 10
  )
```


```{r message=FALSE}
# base::library(BiocParallel)
# BiocParallel::register(
#   BPPARAM = BiocParallel::MulticoreParam(
#       workers = 15
#     )
#   )
```

## [0.2.] Read in data
```{r}
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )
```

## [0.3.] Print out bigWig files for "counts.noSingletons" assay
```{r}
subset <- SummarizedExperiment::colData(GM12878.s.CTSSs) %>%
  base::as.data.frame() %>%
  tibble::rownames_to_column(var = "sample") %>%
  dplyr::filter(main == "yes")
samples <- dplyr::pull(subset,sample)

PRIME::writeBw(
    object = GM12878.s.CTSSs,
    replicates = samples,
    dir = here::here("6.BamFingerprints/20260326_bw_files/"),
    splitByStrand = FALSE,
    inputAssay = "counts.noSingletons"
  )
```

## [0.4.] Read in metainfo
```{r}
meta <- readr::read_delim(
    file = base::paste0(
        here::here("1.CTSSs/"),
        "GM12878.design.txt"
      ),
    delim = "\t",
    col_names = TRUE,
    col_types = readr::cols(
      .default = readr::col_character()
    )
  )
```

# [1.] deepTools plotFingerprint -------------------------

## [1.0.] Read in data
```{r}
cnts_ <- readr::read_delim(
    file = base::paste0(
        here::here("6.BamFingerprints/"),
        "20240417_counts_fingerprints.txt"
      ),
    skip = 1,
    delim = "\t",
    col_names = TRUE,
    col_types = readr::cols(
      .default = readr::col_integer()
    )
  )
```

## [1.1.] Data wrangling
```{r}
cnts <- cnts_ %>%
  tidyr::pivot_longer(
      cols = dplyr::everything(),
      names_to = "sample",
      values_to = "count"
    ) %>%
  dplyr::mutate(
    sample = stringr::str_replace(
      sample, "^.*\\/bam_files\\/(.*)_STAR.*$","\\1"
    )
  )
```

## [1.2.] Calculate relative ranks & cumsums
#### All samples have 500435 entries, evaluate with 'dplyr::summarise(sms, len = dplyr::n())'
#### evaluate with 'base::max(fsms$relagsum)'
```{r}
sms <- cnts %>%
  dplyr::group_by(sample) %>%
  dplyr::arrange(count) %>%
  dplyr::mutate(
    agsum = base::cumsum(count),
    rank = dplyr::row_number(),
    relrank = rank/500435
  )

ref <- sms %>%
  dplyr::summarise(
      fcnt = base::max(agsum)
    )

fsms <- sms %>%
  tidyr::nest() %>% 
  dplyr::mutate(
    newdata = purrr::pmap(
      .l = base::list(sample,data),
      .f = function(s,df){
        c <- ref %>%
          dplyr::filter(sample == s) %>%
          dplyr::pull(fcnt)
        return(
          df %>%
            dplyr::mutate(
              relagsum = agsum/c
            )
          )
      }
    )
  ) %>%
  dplyr::select(-data) %>%
  tidyr::unnest(cols = "newdata")
```

## [1.3.] Test plotting of K562 samples
```{r}
fsms %>%
  dplyr::filter(
      stringr::str_detect(sample,"K562")
    ) %>%
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          x = relrank,
          y = relagsum,
          color = sample
        )
      ) +
    geom_line()
```

## [1.4.] Evaluate samples
```{r}
fsms %>% 
  dplyr::distinct(sample) %>%
  dplyr::filter(
      # stringr::str_detect(sample, "250k|50k|K562", negate = TRUE)
    stringr::str_detect(sample, "500k", negate = TRUE)
    ) %>%
  dplyr::pull(sample)
```

## [1.5.] Subset samples & separate sample name information
#### Evaluate with 'dplyr::distinct(fsmd,sample)', 'dplyr::distinct(fsmd,repseq)'
```{r}
samples <- meta %>%
  dplyr::filter(
      use == "yes" &
      stringr::str_detect(ID,"groCap",negate =TRUE)
    ) %>%
  dplyr::mutate(
    sam = stringr::str_replace(
        bw_plus,"^.*\\/bw_files\\/(.*)_STAR.*","\\1"
      )
    ) %>%
  dplyr::pull(sam)
  
fsmd <- fsms %>%
  dplyr::filter(sample %in% samples) %>% 
  dplyr::mutate(
    seq = stringr::str_extract(
        sample,"reseq"
      ),
    sample = stringr::str_replace_all(
        sample,"GM12878_|CAGE_reseq_|CAGE_",""
      )
    ) %>%
  tidyr::separate(
      col = sample,
      sep = "_",
      into = base::c("source","amount","rep"),
      remove = FALSE
    ) %>%
  dplyr::mutate(
    repseq = dplyr::case_when(
        seq == "reseq" ~ base::paste0(rep,"_",seq),
        TRUE ~ rep
      ),
    srcamt = base::paste0(source,"_",amount)
  )
```

## [1.6.] Convert arguments into factors
```{r}
amts <- base::c("500k","1mio","5mio","10mio")
curlst <- base::c(
    "C_500k_1", "C_500k_2", "C_500k_3",
    "C_1mio_1", "C_1mio_2", "C_1mio_3",
    "N_500k_1", "N_500k_2", "N_500k_3",
    "N_1mio_1", "N_1mio_2", "N_1mio_3",
    "N_5mio_1", "N_5mio_2", "N_5mio_3",
    "N_10mio_1", "N_10mio_2","N_10mio_3"
  )
srcamts <- base::c(
    "C_500k","C_1mio","N_500k",
    "N_1mio","N_5mio","N_10mio"
  )
```

```{r}
fsmp <- fsmd %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(
      sample %in% curlst
    ) %>% #distinct(source)
  dplyr::mutate(
    srcamt = base::factor(
        x = srcamt,
        levels = srcamts,
        ordered = TRUE
      )
    )
```

## [1.7.] Plot data
```{r}
ggplot2::ggplot(
      data = fsmp,
      mapping = aes(
        x = relrank,
        y = relagsum,
        colour = repseq
      )
    ) +
  geom_line() +
  facet_wrap(
      facets = srcamt ~ .,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      panel.spacing = grid::unit(0.1,"lines"),
      aspect.ratio = 1
    )
```

# [2.] deepTools plotFingerprint - QC data -------------------------

## [2.0.] Read in QC data
```{r}
qcs <- readr::read_delim(
      file = base::paste0(
          here::here("6.BamFingerprints/"),
          "20240417_qc_fingerprints.txt"
        ),
      delim = "\t",
      col_names = TRUE,
      col_types = readr::cols(
        .default = readr::col_double(),
        Sample = readr::col_character()
      )
    ) %>%
  dplyr::rename(sample = "Sample")
```

## [2.1.] Read in QC data
```{r}
qcp <- qcs %>%
  dplyr::mutate(
      sample = stringr::str_replace(
        sample, "^(.*)\\/bam_files\\/(.*)_STAR.*$","\\2"
      )
    ) %>%
  dplyr::mutate(
    seq = stringr::str_extract(
        sample,"reseq"
      ),
    sample = stringr::str_replace_all(
        sample,"GM12878_|CAGE_reseq_|CAGE_",""
      )
    ) %>%
  tidyr::separate(
      col = sample,
      sep = "_",
      into = base::c("source","amount","rep"),
      remove = FALSE
    ) %>%
  dplyr::mutate(
      repseq = dplyr::case_when(
          seq == "reseq" ~ base::paste0(rep,"_",seq),
          TRUE ~ rep
        ),
      srcamt = base::paste0(source,"_",amount)
    ) %>%
  dplyr::filter(
      sample %in% curlst
    ) %>%
  dplyr::mutate(
    srcamt = base::factor(
        x = srcamt,
        levels = srcamts,
        ordered = TRUE
      )
    )
```

## [2.2.] PLot data
```{r}
qcp_sub <- qcp %>% 
 dplyr::filter(
      stringr::str_detect(repseq,"reseq")
    )
plot <- fsmp %>%
   dplyr::filter(
      stringr::str_detect(repseq,"reseq")
    ) %>% 
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          x = relrank,
          y = relagsum,
          colour = repseq
        )
      ) +
    geom_line() +
    geom_vline(
        data = qcp_sub,
        mapping = aes(
            xintercept = `X-intercept`
          ),
        linetype = "dashed",
        size = 0.1
      ) +
    geom_vline(
        data = qcp_sub,
        mapping = aes(
            xintercept = `Elbow Point`
          ),
        linetype = "dashed",
        color = "blue",
        size = 0.1
      ) +
    facet_wrap(
        facets = srcamt ~ .,
        nrow = 1
      ) +
    scale_colour_manual(
        limits = base::c("1_reseq","2_reseq","3_reseq"),
        breaks = base::c("1_reseq","2_reseq","3_reseq"),
        values = base::c("#20639B","#3CAEA3","#ED553B"),
      ) +
    theme_bw() +
    theme(
        panel.spacing = grid::unit(0.1,"lines"),
        aspect.ratio = 1,
        legend.position = "none"
      )
ggrastr::rasterise(
    input = plot,
    layers = "Line",
    dpi = 300
  )
```

## [2.3.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("6.BamFingerprints/"),
          format(Sys.time(), "%Y%m%d"),
          "_fingerprint",
          "_AllReseqSamples",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 3
    )
  }
```

## [2.4.] Plot qc data
```{r}
qcp %>%
  dplyr::filter(
      stringr::str_detect(repseq,"reseq")
    ) %>% 
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          y = 1-`Elbow Point`,
          x = repseq
        )
      ) +
    geom_col() +
    facet_wrap(
        facets = srcamt ~ .,
        nrow = 1
      ) +
    scale_y_continuous(
        name = "Enriched genome"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        panel.spacing = grid::unit(0.1,"lines"),
        axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
        axis.title.x = element_blank()
      )
```

## [2.5.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("6.BamFingerprints/"),
          format(Sys.time(), "%Y%m%d"),
          "_fingerprint",
          "_AllReseqSamples",
          "_ElbowEnrichment",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 3
    )
  }
```

## [2.6.] Print data
```{r}
qcp %>%
  dplyr::filter(
      stringr::str_detect(repseq,"reseq")
    ) %>%
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          y = 1-`X-intercept`,
          x = repseq
        )
      ) +
    geom_col() +
    facet_wrap(
        facets = srcamt ~ .,
        nrow = 1
      ) +
    scale_y_continuous(
        name = "Covered genome"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        panel.spacing = grid::unit(0.1,"lines"),
        axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
        axis.title.x = element_blank()
      )
```

## [2.7.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("6.BamFingerprints/"),
          format(Sys.time(), "%Y%m%d"),
          "_fingerprint",
          "_AllReseqSamples",
          "_xInterceptEnrichment",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 3
    )
  }
```

## [2.8.] Plot data
```{r}
qcp %>%
  dplyr::filter(
      stringr::str_detect(repseq,"reseq")
    ) %>%
  dplyr::mutate(
      coeff = (1-`Elbow Point`)/(1-`X-intercept`)
    ) %>%
   ggplot2::ggplot(
        data = .,
        mapping = aes(
          y = coeff,
          x = repseq
        )
      ) +
    geom_col() +
    facet_wrap(
        facets = srcamt ~ .,
        nrow = 1
      ) +
    scale_y_continuous(
        name = "Enriched genome / Covered genome"
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        panel.spacing = grid::unit(0.1,"lines"),
        axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
        axis.title.x = element_blank()
      )
```

## [2.7.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("6.BamFingerprints/"),
          format(Sys.time(), "%Y%m%d"),
          "_fingerprint",
          "_AllReseqSamples",
          "_EnrichedOverCovered",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [3.] Custom fingerprints -------------------------

## [3.0.] Function for generating fingerprint data
```{r}
fingerPrint <- function(object,replicates,inputAssay = "counts",n = 100000){
      all <- base::data.frame()
      for(i in 1:base::length(replicates)){
          sam <- replicates[i]
          col <- base::sort(
              SummarizedExperiment::assay(object,)[,sam]
            )
          
          ## Calculate cumSum as a function of CTSSs rank
          csum <- base::cumsum(col)
          totalLength <- base::length(csum)
          totalCTSSs <- csum[totalLength]
          csumRel <- csum / totalCTSSs
          rank <- base::seq(1L,totalLength,1L) / totalLength
          
          ## Subset to memory-efficient 100k values
          subset <- base::c(
              base::seq(1,totalLength,base::floor(totalLength/n)),
              totalLength
            )
          
          df <- tibble::tibble(
              sample = sam,
              raw = col[subset],
              csum = csum[subset],
              csumRel = csumRel[subset],
              rank = rank[subset]
            )
          ## Combine
          all <- dplyr::bind_rows(all,df)
        }
    return(all)
  }
```

## [3.1.] Apply fingerPrint() to subsetted data
#### columns subsetted via samples and rows via emptiness # 19.987.456
```{r}
subset <- SummarizedExperiment::colData(GM12878.s.CTSSs) %>%
  base::as.data.frame() %>%
  tibble::rownames_to_column(var = "sample") %>%
  dplyr::filter(main == "yes")
samples <- dplyr::pull(subset,sample)

keep <- Matrix::rowSums(
    SummarizedExperiment::assay(GM12878.s.CTSSs,"counts.noSingletons")[,samples]
  ) > 0
GM12878.s.CTSSs.mod <- GM12878.s.CTSSs[keep,samples]
# SummarizedExperiment::assay(GM12878.s.CTSSs.mod,"counts.noSingletons")

GM12878.finger <- PRIME::fingerPrint(
    object = GM12878.s.CTSSs.mod,
    replicates = samples,
    inputAssay = "counts.noSingletons"
  )
# dplyr::distinct(GM12878.finger,sample)
```

## [3.2.] Save & load data
```{r}
base::saveRDS(
  object = GM12878.finger,
  file = base::paste0(
      here::here("6.BamFingerprints/"),
      base::format(base::Sys.time(), "%Y%m%d"),
      "_GM12878.finger",
      "_downsampled",
      ".rds"
    )
  )
```

```{r}
GM12878.finger <- base::readRDS(
  file = base::list.files(
        path = here::here("6.BamFingerprints"),
        full.names = TRUE
      ) %>% 
    stringr::str_subset("GM12878.finger_down")
  )
```

## [3.3.] Data wrangling to prepare plotting
```{r}
amts <- base::c("500K","1M","5M","10M")
reps <- base::c("pool","1","2","3")

GM12878.finger.plot <- GM12878.finger %>%
  tidyr::separate_wider_delim(
      cols = sample,
      delim = "_",
      names = base::c("cell","amount","type"),
      cols_remove = FALSE
    ) %>%
  tidyr::separate(
      col = type,
      sep = 1,
      into = base::c("source","rep"),
      convert = TRUE ## Creates NAs
    ) %>%
  dplyr::mutate(
    rep = dplyr::case_when(
        base::is.na(rep) ~ "pool",
        TRUE ~ base::as.character(rep)
      ),
    rep = base::factor(
        x = rep,
        levels = base::rev(reps),
        ordered = TRUE
      ),
    amount = base::factor(
        x = amount,
        levels = amts,
        ordered = TRUE
      ),
    source = base::factor(
        x = source,
        levels = base::c("C","N"),
        ordered = TRUE
      )
    ) %>%
  dplyr::filter(rank >= 0.7)
```

```{r}
intercepts <- GM12878.finger.plot %>%
  dplyr::filter(csumRel > 0) %>%
  dplyr::group_by(sample) %>%
  dplyr::filter(
      csumRel == base::min(csumRel)
    )
# base::min(intercepts$rank)
```

#### Evaluate
```{r}
GM12878.finger.plot %>%
  dplyr::group_by(sample) %>%
  dplyr::filter(
      csumRel == base::max(csumRel)
    )
```

## [3.4.] Plot fingerPrints
```{r}
 ggplot2::ggplot(
      data = GM12878.finger.plot,
      mapping = aes(
        x = rank,
        y = csumRel,
        color = rep
      )
    ) +
  geom_line(linewidth = 1L) +
  geom_vline(
      data = intercepts,
      mapping = aes(
          xintercept = rank,
          color = rep
        ),
      linetype = "dashed"
    ) +
  ggrepel::geom_text_repel(
      data = intercepts,
        mapping = aes(
          label = base::round(rank,digits = 3)
        ),
      angle = 90,
      ylim = base::c(-0.8,0.0),
      # direction = "x",
      min.segment.length = 0,
      segment.size = 0.2,
      max.overlaps = Inf,
      box.padding = 1,
      max.time = 10,
      force_pull = 0,
      force = 0.5,
      vjust = 0.5
    ) +
  facet_wrap(
      facets = source ~ amount,
      nrow = 1
    ) +
  scale_x_continuous(
      limits = base::c(0.7,1),
      breaks = base::c(0.7,0.8,0.9,1),
      position = "top"
    ) +
  coord_cartesian(clip = "off") +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1,"lines"),
      plot.margin = grid::unit(base::c(0.1, 0.1, 2, 0.1), "cm")
    )
```

## [3.5.] Save complexity plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("6.BamFingerprints/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_GM12878.fingerprint",
          "_all",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```
