---
title: "20240411_workflow_trackVisualization"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Packages -------------------------

## [0.0.] Install necessary packages
#### Maybe required: "base::detach(name = tidyverse, unload = TRUE)" & "utils::remove.packages("tidyverse")"
```{r eval=FALSE}
BiocManager::install("AnnotationHub")
BiocManager::install("remotes")
BiocManager::install("Bioconductor/RBGL")
BiocManager::install("ggbio")
devtools::install_github("thomasp85/patchwork")
utils::install.packages("ggh4x")
utils::install.packages("S7")
utils::install.packages(pkgs = "tidyverse", lib = "/maps/projects/ralab/people/pbr316/bin/Rpackages")
```

```{r}
install.packages("Cairo") #, "cli", "digest", "ff", "FNN", "glue", "knitr", "matrixStats", "polyclip", "Rcpp", "RCurl", "RJSONIO", "rlang", "tinytex", "xfun", "XML", "yaml"))
```


## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library(S7)
base::library("tidyverse", lib.loc = "/maps/projects/ralab/people/pbr316/bin/Rpackages")
Biobase::package.version("ggplot2")
# base::library("AnnotationHub")
# base::library("parallel")
# base::library("foreach")
# base::library("doParallel")
# doParallel::registerDoParallel(
#     cores = 15
#   )

# utils::installed.packages() %>%
#   tibble::as_tibble() %>% 
#   dplyr::filter(
#     stringr::str_detect(Package,"TxDb.Hsapiens")
#   )
base::library("TxDb.Hsapiens.UCSC.hg38.knownGene")

# sep <- .Platform$file.sep
# cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
# base::source(base::paste(cfr.extensions.dir, "normalize.R", sep = sep))
```

## [0.2.] Helper function readRange()
#### evaluate with 'f = fs[1]', 'files = fs', 'range = ran'
```{r}
readRange <- function(files, range){
    ## Prepare range for rtracklayer
    ran_sel <- rtracklayer::BigWigSelection(
        ranges = range
      )
    ## Reading in scores in range from files serially
    raw <- purrr::map(
      .x = files,
      .f = function(f){
          s <- base::basename(f) %>%
            stringr::str_replace(.,"^(.*?).bw$","\\1")
          return(
            base::suppressWarnings(
              rtracklayer::import(
                  con = f,
                  as = "GRanges",
                  selection = ran_sel
                ) %>%
              tibble::as_tibble() %>%
              dplyr::mutate(
                sample = s,
                strand = stringr::str_replace(
                    s, "^(.*)\\.(.*?)$", "\\2"
                  ),
                score = dplyr::case_when(
                    strand == "minus" ~ (-1)*score,
                    TRUE ~ score
                  )
                )
              )
            )
          }
        ) %>%
      dplyr::bind_rows()
    
    ## Fill missing scores across samples with zeros
    out <- .GlobalEnv$fillGaps(raw)
    
    return(out)
  }
```

## [0.3.] Helper function fillGaps()
#### Evaluate with 'data = raw'
```{r}
fillGaps <- function(data){
    columns <- data %>% 
      dplyr::distinct(sample) %>% 
      dplyr::pull(sample)
  
    df_filled <- data %>% 
      tidyr::pivot_wider(
          names_from = "sample",
          values_from = "score"
        ) %>%
      dplyr::mutate(
        dplyr::across(
          .cols = tidyselect::all_of(columns),
          .fns = function(x){
              tidyr::replace_na(x, 0)
            }
          )
        ) %>%
      tidyr::pivot_longer(
          cols = tidyselect::all_of(columns),
          names_to = "sample",
          values_to = "score"
        ) %>%
      dplyr::mutate(
        sample = stringr::str_replace(
            sample, ".minus|.plus", ""
          )
        )
    return(df_filled)
  }
```

## [0.4.] Helper function plotTracks()
#### Evaluate with 'data = cage_sel'
```{r}
plotTracks <- function(data, range){
    start_pos <- BiocGenerics::start(range)
    end_pos <- BiocGenerics::end(range)
    
    plot <- ggplot2::ggplot(
        data = data,
        mapping = aes(
          x = start,
          y = score,
          colour = strand,
          linetype = strand
        )
      ) +
    ggplot2::geom_bar(
        stat = "identity",
        size = 0.5
      ) +
    ggplot2::scale_x_continuous(
        limits = base::c(start_pos,end_pos),  
        position = "top"
      ) +
    ggplot2::scale_color_manual(
        limits = base::c("minus","plus"),
        values = base::c("#FF6400","#632770"),
        guide = "none"
      ) +
    ggplot2::scale_linetype_manual(
        limits = base::c("minus","plus"),
        breaks = base::c("minus","plus"),
        values = base::c("solid","solid"),
        guide = "none"
      ) +
    ggplot2::facet_wrap(
        facets = . ~ sample,
        ncol = 1,
        strip.position = "right"
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(
        # legend.position = "none",
        axis.title.x = element_blank(),
        panel.spacing = grid::unit(0.05,"lines"),
        panel.grid.minor = element_blank(),
        plot.margin = grid::unit(base::c(0,0,0,0),"cm")
      )
    return(plot)
  }
```

## [0.4.] Helper function writeBw()
```{r}
writeBw <- function(object, dir, replicates = "all", inputAssay = "counts", splitByStrand = TRUE){
  
  if(replicates == "all"){
      replicates <- base::rownames(SummarizedExperiment::colData(object))
    }
  
  ## Check consistency of replicates
  assertthat::assert_that(
      base::length(SummarizedExperiment::rowRanges(object)) == base::nrow(SummarizedExperiment::assay(object,inputAssay)),
      base::all(replicates %in% base::rownames(SummarizedExperiment::colData(object))),
      inputAssay %in% base::names(SummarizedExperiment::assays(object))
    )
  base::stopifnot("Chosen directory doesn't exist!" = base::dir.exists(dir))
  
  gr <- GenomicRanges::GRanges(SummarizedExperiment::rowRanges(object))
  
  ## Parallel export across all chosen replicates
  foreach::foreach(i=1:base::length(replicates)) %dopar% {
    
    ## Print replicate when exiting
    base::on.exit(base::cat(base::paste0("Exit due to ",replicates[i],".\n")), add = TRUE)
    
    ## Convert individual replicate to GRanges object
    S4Vectors::mcols(gr)["score"] <- SummarizedExperiment::assay(object,inputAssay)[,replicates[i]]
    
    ## Export replicate's plus and minus strands separately
    if(splitByStrand){
      for (s in base::c("+","-")){
          grs <- gr[BiocGenerics::strand(gr) == s,]
          s <- base::ifelse(s == "+","plus","minus")
          
          fn <- base::paste0(dir,replicates[i],".",inputAssay,".",s,".bw")
          rtracklayer::export(
              object = grs,
              con = fn,
              format = "bigWig"
            )
          base::message(base::paste(fn,"done.",sep = " "))
        }
      } else {
        
        # Aggregate counts from both strands
        grred <- IRanges::reduce(gr, ignore.strand = TRUE, with.revmap = TRUE, min.gapwidth = 0)
        S4Vectors::mcols(grred) <- S4Vectors::aggregate(
            gr, S4Vectors::mcols(grred)$revmap, score = base::sum(score), drop = FALSE
          )
        
        # Keep only score metadata column
        grred@elementMetadata <- grred@elementMetadata[,"score",drop = FALSE]
    
        fn <- base::paste0(dir,replicates[i],".",inputAssay,".bw")
        rtracklayer::export(
              object = grred,
              con = fn,
              format = "bigWig"
            )
        base::message(base::paste(fn,"done.",sep = " "))
      }
    }
 }
```

## [1.] Prerequisites --------------------

## [1.0.] Color palette
```{r}
cols <- base::c(
    "#659DBD", "#FFC425", "#AFD275",
    "#9A1750", "#FC4445", "#55BCC9",
    "#242582", "#00B159", "#8860D0",
    "#41B3A3", "#C38D9E", "#8EE4AF",
    "#1F2833", "#D79922", "#7E685A",
    "#8D8741", "#E27D60", "#E8A87C",
    "#9B786F", "#DAAD86", "#EDF5E1",
    "#64485C"
  )
```

## [1.1.] Read in annotations - convert from gff to gr
#### unique entries: 'hg38_gr[which(!is.na(elementMetadata(dm6_gr)[,"Name"])),]'
#### 'hg38_tbl <- tibble::as_tibble(hg38_gr)', 'dplyr::distinct(hg38_tbl,type)'
#### 'base::unique(S4Vectors::elementMetadata(hg38_gr)[,"type"])'
#### previous '/isdata/alab/genome/annotations/hg38/gencode.v31.annotation.gtf'
#### see 'https://www.gencodegenes.org/human/' for updates
```{r}
# hg38_gff <- rtracklayer::readGFF(
#     filepath = base::paste0(
#       here::here("0.External_resources/"),
#       "gencode.v39.annotation.gtf"
#     )
#   )
# dplyr::distinct(hg38_gff,type)
# hg38_gr <- GenomicRanges::makeGRangesFromDataFrame(
#     df = hg38_gff,
#     keep.extra.columns = TRUE
#   )
# 
# types <- base::c(
#     "transcription_start_site", "exon_junction", "exon",
#     "gene", "pseudogene", "ncRNA",
#     "intron", "mRNA", "protein",
#     "CDS", "five_prime_UTR", "three_prime_UTR",
#     "polyA_site", "snoRNA", "pre_miRNA",
#     "miRNA", "silencer", "enhancer",
#     "snRNA", "tRNA", "rRNA",
#     "UTR"
#   )
# hg38_gr_sub <- hg38_gr[
#     S4Vectors::elementMetadata(hg38_gr)[,"type"] %in% types,
#   ]
```

## [1.2.] Build lookup table for colors
#### 'col2rgb(cols3[1])[1:3,] %>% unname() %>% paste(collapse = ",")'
```{r}
# annos <- base::unique(
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"type"]
#   )
# 
# lkup <- tibble::tibble(
#     txType = annos,
#     hex = cols[1:base::length(annos)]
#   )
```

## [1.3.] Identify K562 and GM12878 bigwig files to be read
```{r}
fs <- base::list.files(
      path = here::here("trackHubFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("PoolSub|repSub", negate = TRUE) %>% 
  stringr::str_subset("_500K_|_5M_") # GM12878-specific
  # stringr::str_subset("K562") # K562-specific
```

## [1.4.] Identify FANTOM5 immune tracks (Fig6)
```{r}
facets_sel <- base::c(
    "mast cell", "dendritic cell", "monocyte",
    "T cell", "reticulocyte"
  )

fs <- base::list.files(
      path = here::here("0.External_resources/FANTOM5/pooled_bw"),
      full.names = TRUE
    ) %>%
  stringr::str_subset(base::paste0(facets_sel,collapse = "|"))
```

## [1.5.] Load HepG2-, HCT116-, A549-CTSS objects & adapt names (Fig5)
#### Evaluate with 'rsen = "A549"', 'SummarizedExperiment::colData(CTSSs.objects[[rsen]])'
```{r}
CTSSs.objects_ <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("\\/poolSub") %>%
  stringr::str_subset("HepG2|HCT116|A549")
  # stringr::str_subset("HepG2|HCT116|A549|combined.GM12878|combined.k562")

CTSSs.objects <- purrr::map(
    .x = CTSSs.objects_,
    .f = base::readRDS
  )
base::names(CTSSs.objects) <- stringr::str_replace(
    CTSSs.objects_, ".*/poolSub(\\.combined)?\\.(.*?)\\.CTSSs.*", "\\2"
  )
CTSSs.objects <- purrr::map(
    .x = base::names(CTSSs.objects),
    .f = function(rsen){
      base::rownames(SummarizedExperiment::colData(CTSSs.objects[[rsen]])) <- base::paste0(
          rsen, "_",
          base::rownames(SummarizedExperiment::colData(CTSSs.objects[[rsen]]))
        )
      return(CTSSs.objects[[rsen]])
    }
  )
base::names(CTSSs.objects) <- stringr::str_replace(
    CTSSs.objects_, ".*/poolSub(\\.combined)?\\.(.*?)\\.CTSSs.*", "\\2"
  )
```

## [1.6.] Add the corresponding GM12878 and K562 poolSub data to CTSSs.objects of [1.5.] (Fig5)
#### Evaluate with SummarizedExperiment::colData(K562.CTSSs.sub), SummarizedExperiment::assay(K562.CTSSs.sub, i = "counts.noSingletons")
```{r}
K562.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = K562.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
      stringr::str_detect(type, "K562_C$|K562_N$")
    ) %>%
  dplyr::pull(ID)

# K562.CTSSs.sub <- S4Vectors::subset(
#     x = K562.CTSSs,
#     select = SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
#   )

tpm <- SummarizedExperiment::assay(K562.CTSSs,"TPM")[,SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids]
K562.CTSSs.sub <- SummarizedExperiment::SummarizedExperiment(
    assays = base::list(
        TPM = SummarizedExperiment::assay(K562.CTSSs,"TPM")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts = SummarizedExperiment::assay(K562.CTSSs,"counts")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts.noSingletons = SummarizedExperiment::assay(K562.CTSSs,"counts.noSingletons")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ] # 8810502, 2
      ),
    colData = SummarizedExperiment::colData(K562.CTSSs)[SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids,],
    rowRanges = SummarizedExperiment::rowRanges(K562.CTSSs)[(Matrix::rowSums(tpm) != 0),]
  )

CTSSs.objects[["K562"]] <- K562.CTSSs.sub
base::remove(K562.CTSSs,K562.CTSSs.sub)
```

#### Evaluations
```{r}
SummarizedExperiment::assay(K562.CTSSs,"TPM")[1:13,]
SummarizedExperiment::assay(K562.CTSSs.sub,"TPM")[1:13,]

assertthat::are_equal(
    SummarizedExperiment::rowRanges(K562.CTSSs)[1030345:1030545],
    base::sort(SummarizedExperiment::rowRanges(K562.CTSSs))[1030345:1030545]
  )
setdiff(
    SummarizedExperiment::rowRanges(K562.CTSSs),
    base::sort(SummarizedExperiment::rowRanges(K562.CTSSs))
  )
tpm <- SummarizedExperiment::assay(K562.CTSSs.sub,"TPM")
tpm[(Matrix::rowSums(tpm) != 0),] # 8,810,502
SummarizedExperiment::rowRanges(K562.CTSSs)[(Matrix::rowSums(tpm) != 0),] # 8,810,502
```

```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
     stringr::str_detect(type, "GM12878_1M_C$|GM12878_1M_N$")
    ) %>%
  dplyr::pull(ID)

# GM12878.CTSSs.sub <- S4Vectors::subset(
#     x = GM12878.CTSSs,
#     select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
#   )

tpm <- SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[,SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids]
GM12878.CTSSs.sub <- SummarizedExperiment::SummarizedExperiment(
    assays = base::list(
        TPM = SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts = SummarizedExperiment::assay(GM12878.CTSSs,"counts")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts.noSingletons = SummarizedExperiment::assay(GM12878.CTSSs,"counts.noSingletons")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ] # 8810502, 2
      ),
    colData = SummarizedExperiment::colData(GM12878.CTSSs)[SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids,],
    rowRanges = SummarizedExperiment::rowRanges(GM12878.CTSSs)[(Matrix::rowSums(tpm) != 0),]
  )

CTSSs.objects[["GM12878"]] <- GM12878.CTSSs.sub
# SummarizedExperiment::colData(GM12878.CTSSs.sub)
base::remove(GM12878.CTSSs,GM12878.CTSSs.sub)
```

#### Evaluations
```{r}
SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[1:13,]
SummarizedExperiment::assay(GM12878.CTSSs.sub,"TPM")[13,]
assertthat::are_equal(
    SummarizedExperiment::rowRanges(GM12878.CTSSs)[1030345:1030545],
    base::sort(SummarizedExperiment::rowRanges(GM12878.CTSSs))[1030345:1030545]
  )
setdiff(
    SummarizedExperiment::rowRanges(GM12878.CTSSs),
    base::sort(SummarizedExperiment::rowRanges(GM12878.CTSSs))
  )
```

## [1.7.] Load prepared 'all CTSS' - objects for GM12878 & K562
#### Evaluate with 'SummarizedExperiment::colData(GM12878.CTSSs.sub)' & 'SummarizedExperiment::assay(GM12878.CTSSs.sub, "counts")'
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
      (pooled == "yes" & stringr::str_detect(type, "trip|ctrl")) |
      (stringr::str_detect(type, "groCap")) |
      (pooled == "yes" & stringr::str_detect(ID, "500K|5M"))
    ) %>%
  dplyr::pull(ID)

GM12878.CTSSs.sub <- S4Vectors::subset(
    x = GM12878.CTSSs,
    select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
  )
base::remove(GM12878.CTSSs)
```

#### Evaluate with 'SummarizedExperiment::colData(K562.CTSSs.sub)' & 'SummarizedExperiment::assay(K562.CTSSs.sub, "counts")'
```{r}
K562.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.k562.CTSSs.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = K562.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      # pooled == "yes" &
      stringr::str_detect(ID, "Cap", negate = TRUE)
    ) %>%
  dplyr::pull(ID)

K562.CTSSs.sub <- S4Vectors::subset(
    x = K562.CTSSs,
    select = SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
  )
```

## [1.8.] Subset for other samples of interest, e.g., 5M, Triptolide or GROcap
```{r}
# ##GM12878 5M and 500k samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       pooled == "yes" &
#       stringr::str_detect(ID, "500K|5M")
#     ) %>%
#   dplyr::pull(ID)
# 
# ##Triptolide samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       stringr::str_detect(Pool, "trip")
#     ) %>%
#   dplyr::pull(ID)
# 
# ##GRO-cap samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       stringr::str_detect(type, "groCap")
#     ) %>%
#   dplyr::pull(ID)
```

## [1.9.] Write out bigWig files
```{r}
samples = base::rownames(
    SummarizedExperiment::colData(GM12878.CTSSs.sub)
  )

.GlobalEnv$writeBw(
    object = GM12878.CTSSs,
    replicates = samples,
    dir = base::paste0(
        here::here("trackHubFiles/"),
        "20240718_bigWig/"
      ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

```{r}
samples = base::rownames(
    SummarizedExperiment::colData(K562.CTSSs.sub)
  )

.GlobalEnv$writeBw(
    object = K562.CTSSs,
    replicates = samples,
    dir = base::paste0(
        here::here("trackHubFiles/"),
        "20240718_bigWig/"
      ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

#### Write out HepG2-, HCT116-, A549-CTSS samples
```{r}
purrr::map(
    .x = CTSSs.objects,
    .f = .GlobalEnv$writeBw,
    dir = base::paste0(
          here::here("trackHubFiles/"),
          "20240718_bigWig/"
        ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

## [1.10.] TMM normalise data
```{r}
# base::debugonce(edgeR::calcNormFactors)
# facs <- edgeR::calcNormFactors(
#     object = GM12878.CTSSs.sub,
#     method = "TMM",
#     refColumn = NULL
#   )
# GM12878.tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
#     object = GM12878.CTSSs.sub,
#     sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#     inputAssay = "counts",
#     outputAssay = "TMM"
#   )
```

#### Use TMM values as "counts" and calculate TPMs on top of this
```{r}
# GM12878.tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
#       object = GM12878.CTSSs.sub,
#       sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#       inputAssay = "counts",
#       outputAssay = "counts"
#     ) %>%
#   CAGEfightR::calcTPM()
```

#### Deconstructed edgeR::calcNormFactors() function
#### "GM12878_5M_N2" is the refColumn
```{r}
# object <- edgeR::SE2DGEList(GM12878.CTSSs.sub)
# x <- base::as.matrix(object)
# nsamples <- base::ncol(x)
# lib.size <- Matrix::colSums(x)
# allzero <- .Internal(rowSums(x > 0, base::nrow(x), nsamples, na.rm = FALSE)) == 0L
# x <- x[!allzero, , drop = FALSE]
# f75 <- base::suppressWarnings(
#    edgeR:::.calcFactorQuantile(
#        data = x,
#        lib.size = lib.size,
#        p = 0.75
#      )
#    )
# if (stats::median(f75) < 1e-20) {
#     refColumn <- BiocGenerics::which.max(base::colSums(base::sqrt(x)))
#   } else {
#     refColumn <- BiocGenerics::which.min(base::abs(f75 - base::mean(f75)))
#   }
```

## [1.11.] Load tmm.CTSSs object
#### see [1.15.2.] in "1.1.CTSSs.processing.Rmd" for more info
```{r}
# fn <- base::list.files(
#       path = here::here("1.CTSSs"),
#       full.names = TRUE
#     ) %>%
#   stringr::str_subset("tmm.CTSSs")
# 
# tmm.CTSSs <- base::readRDS(file = fn)
```

## [1.12.] Read in CRISPR-gRNA locations
```{r}
dir = "/projects/ralab/people/pbr316/20221208_scCAGE/resources/"
crispr <- readr::read_delim(
    file = base::paste0(
        dir,
        "EPCrisprBenchmark_ensemble_data_GRCh38.tsv.gz"
      ),
    delim = "\t",
    col_names = TRUE
  )

myc_locus <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::summarise(
      chr = dplyr::first(chrom),
      start = base::min(chromStart),
      end = base::max(chromEnd)
    ) %>%
  GenomicRanges::GRanges()

crispr_data <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::mutate(
      y = dplyr::row_number(),
      y_alt = 1,
      summit = chromStart + base::round((chromEnd-chromStart+1)/2)
    )
```

# [2.] Prepare coordinate objects -------------------------

## [2.0.] Define coordinates
#### suggested region "chr8:144,407,529-144,416,261", "chr10:110,291,360-110,366,023", 
#### wider area: "chrX:48,670,569-48,746,869" with enh1: "chrX:48,676,170-48,676,946" and enh2: "chrX:48,737,065-48,737,528"
#### enhancers - "chr9:41358750-41358940", "chr9:41,356,277-41,359,588", "chr8:134563800-134564100", "chr8:134,563,186-134,566,375"
#### MYC-enhancer: "chr8:127,884,215-127,915,852"
#### MYC-locus & PVT1-promoter: "chr8:127,704,300-127,810,400"
#### HDAC6 promoter: "chrX:48,801,000-48,803,000"
#### RUNX1 promoter: "chr21:35,048,764-35,049,617"
#### MEF2A promoter: "chr15:99,565,263-99,566,083"
#### GATA1 promoters: "chrX:48,786,301-48,788,293" & enhancer: "chrX:48,787,857-48,788,321"
#### KLF2 promoter: "chr19:16,324,467-16,325,106"
#### BCL2 promoter/PROMPTs: "chr18:63,317,868-63,320,926"
```{r}
# l_ <- "chr4:159,101,737-159,104,305"
l_ <- "chr4:144,080,727-144,158,976"
```

#### see [4.10.] in 10.3.Cell_line_vs_trait_enrichment.Rmd for 'chr22:37109513:37109512'

### MCHC trait (Fig5)
#### Whole ABCA7 gene including "chr19:1050876-1050875": "chr19:1039521-1065670"
#### Whole TMPRSS6 gene including "chr22:37109513:37109512": "chr22:37,065,000-37,110,700"
```{r}
l_ <- "chr19:1,039,521-1,065,670" # ABCA7
l_ <- "chr19:1,050,492-1,051,269" # ABCA7 Primeloci

l_ <- "chr22:37,065,000-37,110,700" # TMPRSS6
l_ <- "chr22:37,109,199-37,109,898" # TMPRSS6 Primeloci
```

### ApoB trait (Fig5)
#### Whole ApoB gene locus as control: "chr2:20,999,429-21,046,073"
#### Whole ABO gene including "chr9:133274296:133274295" and "chr9:133263363:133263374": "chr9:133,250,000-133,276,238"
#### Whole LPL gene including: "chr8:19961929-19961928": "chr8:19,938,800-19,968,800"
#### Whole LIPC gene including: "chr15:58430392:58430391": "chr15:58,400,000-58,570,000"

#### Whole APOC1 gene including "chr19:44913485:44913484": "chr19:44,914,000-44,920,000" & with alternatively entire promoter proximal region "chr19:44,913,101-44,914,607"
#### APOE - APOC1: "chr19:44,908,532-44,915,200"
```{r}
l_ <- "chr2:20,999,429-21,046,073"

l_ <- "chr9:133,250,000-133,276,238" # ABO
l_ <- "chr9:133,273,899-133,274,680" # ABO PRIMElocus

l_ <- "chr8:19,938,800-19,968,800" # LPL
l_ <- "chr8:19,961,537-19,962,318" # LPL Primelocus

l_ <- "chr15:58,400,000-58,570,000" # LIPCs
l_ <- "chr15:58429996-58435030" # LIPCs PRIMEloci

l_ <- "chr19:44,908,532-44,915,200" # APOC1
```

### c-KIT & KLF6 (Fig6)
```{r}
l_ <- "chr4:54,360,000-55,036,000" # c-KIT
l_ <- "chr10:3,700,000-4,146,000" # KLF6
```

## [2.1.] Get genomic region from coordinates
```{r}
l <- stringr::str_replace_all(l_,",","")
chr = stringr::str_replace(l,"^(chr.*):.*","\\1")
s = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\1") %>%
  base::as.integer()
e = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\2") %>%
  base::as.integer()
out <- base::paste0("_",chr,"_",s,"_",e)
```

```{r}
extend = 0
ran <- GenomicRanges::GRanges(
    seqnames = chr,
    ranges = IRanges::IRanges(start = s, end = e)
  )
BiocGenerics::start(ran) <- BiocGenerics::start(ran) - extend
BiocGenerics::end(ran) <- BiocGenerics::end(ran) + extend
GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
```

## [2.2.] Get genomic region from gene names
#### examples are SLC39A4, YLPM1, SMNDC1
```{r}
# omr <- base::c(
#     "AREL1","FCF1","YLPM1","XR_944033.2",
#     "PROX2"
#   )
# ran_ <- hg38_gr_sub[
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"gene_name"] %in% omr,
#   ]
# base::unique(ran_$gene_name)
# ran <- base::range(ran_,ignore.strand = TRUE)
# 
# GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
# if(!stringr::str_detect(GenomeInfoDb::seqlevels(ran)[1],"chr")){
#   GenomeInfoDb::seqlevels(ran) <- base::gsub(
#       pattern = '^(.*)$',
#       replacement = 'chr\\1',
#       x = GenomeInfoDb::seqlevels(ran)
#     )
#   }
# GenomeInfoDb::seqlevels(ran)
```

## [2.3.] Get genomic regions from GRanges input
```{r}
# gr <- crispr %>%
#   dplyr::filter(
#       measuredGeneSymbol == "MYC" &
#       Significant == TRUE
#     ) %>%
#   dplyr::mutate(
#       summit = chromStart + base::round((chromEnd - chromStart +1)/2)
#     ) %>% 
#   GenomicRanges::GRanges()
# extend = 1000
# BiocGenerics::start(gr) <- BiocGenerics::start(gr) - extend
# BiocGenerics::end(gr) <- BiocGenerics::end(gr) + extend
# GenomeInfoDb::seqlevelsStyle(gr) <- "UCSC"
# myce38 <- IRanges::reduce(gr, ignore.strand = TRUE, with.revmap = TRUE)
# S4Vectors::mcols(myce38)$smts <- S4Vectors::aggregate(gr, S4Vectors::mcols(myce38)$revmap, smts = summit)$smts
# i = 1L
# ran <- myce38[i]
```

## [2.4.] Extract coordinates
```{r}
chrom <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)
```

## [3.] CAGE_preparation_BigWig --------------------

## [3.1.] Read in tpm scores in selected ranges
#### see [2.0.] in '20200721_workflow_DrosophilaCAGE.Rmd'
```{r}
cage_sel_ <- .GlobalEnv$readRange(fs,ran)
```

## [3.2.] Summarise data
```{r}
cage_sel <- cage_sel_ %>%
  dplyr::mutate(
    replicate = stringr::str_replace(
        sample, ".*_(.*?)\\.TPM$", "\\1"
      ),
    type = stringr::str_replace(
        sample, "^(.*?)_(.*)_(.*)\\.TPM$", "\\2"
      ),
    )
```

## [3.3.] Adjust source order
#### dplyr::distinct(cage_sel,source)
```{r}
levs <- base::c(
    "C", "C1", "C2", "C3",
    "N", "N1", "N2", "N3"
  )
cage_sel <- cage_sel %>%
  dplyr::mutate(
      replicate = base::factor(
          x = replicate,
          levels = levs,
          ordered = TRUE
        )
    ) %>%
  dplyr::arrange(
      type,replicate,seqnames,start,end
    )
```

## [4.] CAGE_preparation_CTSSs --------------------

## [4.0.] Subset & convert scores from RangedSummarizedExperiment to tibble
#### Evaluate samples "base::rownames(SummarizedExperiment::colData(GM12878.CTSSs.sub))"
```{r}
# ovl <- IRanges::findOverlaps(
#     query = ran,
#     subject = GM12878.CTSSs.sub # K562.CTSSs.sub
#   )
# gr <- SummarizedExperiment::rowRanges(GM12878.CTSSs.sub)[ovl@to,] # K562.CTSSs.sub
# 
# crds <- gr %>% 
#   base::as.data.frame() %>%
#   tibble::as_tibble() %>%
#   dplyr::mutate(
#       start = pos,
#       end = pos
#     ) %>%
#   dplyr::select(
#       seqnames,start,end,strand
#     )
# 
# scrs <- SummarizedExperiment::assay(GM12878.CTSSs.sub,"TPM")[ovl@to,] %>% # K562.CTSSs.sub
#   base::as.matrix() %>%
#   base::as.data.frame()
# 
# cage_sel_ <- dplyr::bind_cols(crds,scrs)
```

#### Evaluate with 'c = "HepG2"'
#### Alternative for combining data: base::do.call(what = "rbind",args = base::lapply(X = base::names(ovl),FUN = function(c){}))'
```{r}
# ovl <- purrr::map(
#   .x = base::names(CTSSs.objects),
#   .f = function(c){
#       ovl <- IRanges::findOverlaps(
#           query = ran,
#           subject = CTSSs.objects[[c]]
#         )
#       gr <- SummarizedExperiment::rowRanges(CTSSs.objects[[c]])[ovl@to,]
#       return(methods::as(gr,"GRanges"))
#     }
#   )
# base::names(ovl) <- base::names(CTSSs.objects)
# 
# crds <- purrr::map_dfr(
#   .x = base::names(ovl),
#   .f = function(c){
#     ovl[[c]] %>% 
#       base::as.data.frame() %>%
#         tibble::as_tibble() %>%
#         dplyr::mutate(
#             celltype = c
#           ) %>%
#         dplyr::select(
#             celltype,seqnames,start,end,strand
#           )
#         }
#       )
```

```{r}
ovl <- purrr::map(
  .x = base::names(CTSSs.objects),
  .f = function(c){
      return(
        IRanges::findOverlaps(
          query = ran,
          subject = CTSSs.objects[[c]]
        )
      )
    }
  )
base::names(ovl) <- base::names(CTSSs.objects)

crds <- purrr::map_dfr(
  .x = base::names(ovl),
  .f = function(c){
    gr <- SummarizedExperiment::rowRanges(CTSSs.objects[[c]])[ovl[[c]]@to,]
    methods::as(gr,"GRanges") %>%
      base::as.data.frame() %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = c,
            rowID = ovl[[c]]@to
          ) %>%
        dplyr::select(
            type,seqnames,start,
            end,strand,rowID
          )
        }
      )
```

#### Evaluate with 'c = "HCT116"', 'c = "A549"','c = "K562"', 'c = "HepG2"'
```{r}
scrs <-  purrr::map_dfr(
  .x = base::names(ovl),
  .f = function(c){
    
    mat <- CTSSs.objects[[c]]@assays@data$TPM %>%
      base::as.matrix()
    
    base::colnames(mat) <- stringr::str_extract(base::colnames(mat), "C|N")
    return(
      mat %>% 
        tibble::as_tibble() %>%
        dplyr::filter(
            dplyr::row_number() %in% ovl[[c]]@to
          ) %>%
        dplyr::mutate(
            type = c,
            rowID = ovl[[c]]@to
          )
      )
    }
  )

cage_sel_ <- dplyr::left_join(
      crds, scrs, by = base::c("type","rowID")
    ) %>%
  dplyr::mutate(
      dplyr::across(
        C:N, ~ base::replace(., base::is.na(.), 0)
      )
    ) %>%
  dplyr::select(-rowID)
```

## [4.1.] Summarise data
#### Evaluate with 'dplyr::arrange(cage_sel,dplyr::desc(score))'
#### Use 'TRUE ~ (20)*score' for GATA1 enhancer example
```{r}
# regex_f = "(?<=C|N)\\d$" # ".*(.)$"
# regex_s = "^(.*?)_(.*?)(\\d)?$" # "^(.*?)_(.*)(_)?(.)"
# 
# cage_sel <- cage_sel_ %>%
#   tidyr::pivot_longer(
#       cols = tidyselect::contains("_"),
#       names_to = "sample",
#       values_to = "score"
#     ) %>%
#   dplyr::mutate(
#     replicate = stringr::str_extract(
#         sample, regex_f
#       ),
#     replicate = dplyr::case_when(
#         base::is.na(replicate) ~ "pool",
#         TRUE ~ replicate
#       ),
#     type = stringr::str_replace(
#           sample, regex_s, "\\2"
#         ) %>%
#       stringr::str_replace(
#         .,"^(.*?)[_]$","\\1"
#       )
#     ) %>%
#   dplyr::mutate(
#     score = dplyr::case_when(
#         strand == "-" ~ (-1)*score,
#         TRUE ~ score
#       ),
#     strand = dplyr::case_when(
#         strand == "+" ~ "plus",
#         strand == "-" ~ "minus"
#       )
#     )
```

```{r}
cage_sel <- cage_sel_ %>%
  tidyr::pivot_longer(
      cols = base::c("C","N"),
      names_to = "sample",
      values_to = "score"
    ) %>%
  dplyr::mutate(
    score = dplyr::case_when(
        strand == "-" ~ (-1)*score,
        TRUE ~ score
      ),
    strand = dplyr::case_when(
        strand == "+" ~ "plus",
        strand == "-" ~ "minus"
      )
    )
```

## [4.2.] Add mock lines for missing cell types
#### Evaluate with "ins <- base::names(CTSSs.objects)"
```{r}
ins <- cage_sel %>% 
  dplyr::distinct(type) %>%
  dplyr::pull(type)

ous <- base::names(CTSSs.objects)[! base::names(CTSSs.objects) %in% ins]

insert <- tibble::tibble(
    type = base::sort(base::rep(ous,2)),
    seqnames = base::rep(chrom,2*base::length(ous)),
    start = base::rep(start_pos,2*base::length(ous)),
    end = base::rep(end_pos,2*base::length(ous)),
    strand = base::rep("plus",2*base::length(ous)),
    sample = base::rep(base::c("C","N"),base::length(ous)),
    score = base::rep(base::as.double(0),2*base::length(ous))
  )

cage_sel <- dplyr::bind_rows(cage_sel,insert)
```

# [5.] Plot data -------------------------

## [5.0.] Plot genome tracks
```{r}
# plot <- .GlobalEnv$plotTracks(cage_sel, ran) +
#   ggplot2::facet_wrap(
#       facets = . ~ type + replicate,
#       ncol = 1,
#       strip.position = "right",
#       scales = "free_y"
#     ) +
#   ggh4x::facetted_pos_scales(
#     y = base::list(
#       type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ scale_y_continuous(
#           limits = base::c(-3,3)
#         ),
#       type %in% base::c("500K_N","5M_N","500K_C") ~ scale_y_continuous(
#           limits = base::c(0,300)
#         )
#       )
#     )
# ggrastr::rasterise(
#     input = plot,
#     layers = "Bar",
#     dpi = 300
#   )
```

```{r}
# plot <- cage_sel %>%
#   dplyr::filter(
#       sample != "GM12878_10M_Cap_align"
#     ) %>% 
#   dplyr::mutate(
#     type = base::factor(
#         x = type,
#         levels = base::c("500K_C","500K_N","5M_N","1M_trip","1M_ctrl","groCap","10M_Cap_align"),
#         ordered = TRUE
#       ),
#     score = dplyr::case_when(
#         strand == "minus" & type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ score,
#         strand == "minus" & type %in% base::c("500K_N","5M_N","500K_C") ~ score*200,
#         TRUE ~ score
#       )
#     ) %>%
#   .GlobalEnv$plotTracks(., ran) +
#     ggplot2::facet_wrap(
#         facets = . ~ type,
#         ncol = 1,
#         strip.position = "right",
#         scales = "free_y"
#       ) +
#     ggh4x::facetted_pos_scales(
#       y = base::list(
#         type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ scale_y_continuous(
#             limits = base::c(-3,3)
#           ),
#         type %in% base::c("500K_N","5M_N","500K_C") ~ scale_y_continuous(
#             limits = base::c(-300,300),
#             breaks = base::c(-200,0,200),
#             labels = base::c(-1,0,200)
#           )
#         )
#       )
# ggrastr::rasterise(
#     input = plot,
#     layers = "Bar",
#     dpi = 300
#   )
```


## [5.1.] Print data
```{r eval=FALSE}
# for(ext in c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("trackHubFiles/"),
#           format(Sys.time(), "%Y%m%d"),
#           out,
#           # "_enh2",
#           "_500k5M",
#           "_TmmTpm",
#           ".", ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 4,
#       height = 8
#     )
#   }
```

## [5.2.] E.g., APOC1 example (ABO)
#### see [4.10.] in "10.3.Cell_line_vs_trait_enrichment.Rmd"
```{r}
sites_ <- base::readRDS(
  file = base::list.files(
        path = here::here("10.External_validation"),
        full.names = TRUE
      ) %>%
    stringr::str_subset("ApoB.rds")
    # stringr::str_subset("MCHC.rds")
  )
```

## E.g., Filter out ApoB-related genetic variants in the vicinity of APOC1 gene (ABO)
#### see also [2.3.] for 'start_pos' and 'end_pos' objects
```{r warning=FALSE}
chr <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

sites <- sites_ %>% 
  tidyr::pivot_wider( # 218
      names_from = "cell.line",
      values_from = "score"
    ) %>%
  # dplyr::filter(
  #     stringr::str_detect(variant, "chr19:44")
  #   ) %>%
  tidyr::separate(
      col = variant,
      sep = ":",
      into = base::c("chrom","start","end","gv"),
      convert = TRUE
    ) %>%
  dplyr::filter(
      chrom == chr & start > start_pos & end < end_pos,
    ) %>%
  dplyr::mutate(
      end = start
    ) %>% 
  # dplyr::select(chrom, start, gv) %>% 
  GenomicRanges::makeGRangesFromDataFrame()
  # dplyr::mutate(
  #     start = base::as.double(start) - extend,
  #     end = base::as.double(end) + extend
  #   ) %>% 
  # dplyr::select(start, end)
```

#### see [1.0.-1.4.] in "10.3.Cell_line_vs_trait_enrichment.Rmd" for 'unfiltered.PRIMEloci'-object
#### Alternative: Use IRanges::reduce() on olap from unfiltered.PRIMEloci and filter for the highest PRIMEloci score
#### evaluate with 'n = "HepG2_N"' # 61,589

#### Pooled PRIMEloci data
```{r}
sites_scores_ <- purrr::map_dfr(
  .x = base::names(pooled.PRIMEloci),
  .f = function(n){
    pd <- pooled.PRIMEloci[[n]]
    olap <- IRanges::findOverlaps(
        query = sites,
        subject = pd,
        minoverlap = 1
      )
    return(
      pd[S4Vectors::subjectHits(olap),] %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = stringr::str_replace(n,"(.*?)_.*","\\1"),
            sample = stringr::str_replace(n,"(.*?)_(.*)","\\2")
          )
      )
    }
  )
```

#### Unfiltered PRIMEloci data
```{r}
sites_scores_ <- purrr::map_dfr(
  .x = base::names(unfiltered.PRIMEloci),
  .f = function(n){
    pd <- unfiltered.PRIMEloci[[n]]
    olap <- IRanges::findOverlaps(
        query = sites,
        subject = pd,
        minoverlap = 1
      )
    
    gr <- pd[S4Vectors::subjectHits(olap),]
    out <- IRanges::reduce(x = gr, with.revmap = TRUE)
    S4Vectors::mcols(out)$score <- base::do.call(
          what = rbind,
          args = base::lapply(
            X = S4Vectors::mcols(out)$revmap,
            FUN = function(i) {
              base::max(S4Vectors::mcols(gr)$score[i])
            }
          )
        )
    
    return(
      out %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = stringr::str_replace(n,"(.*?)_.*","\\1"),
            sample = stringr::str_replace(n,"(.*?)_(.*)","\\2")
          )
      )
    }
  )
```

#### Cut sites to plotted range
```{r}
sites_scores <- sites_scores_ %>%
  dplyr::mutate(
    start = dplyr::case_when(
        start <= start_pos ~ start_pos,
        .default = start
      ),
    end = dplyr::case_when(
        end >= end_pos ~ end_pos,
        .default = end
      )
    )
```


#### Plot data
```{r}
# base::range(cage_sel$score)
minY = base::range(cage_sel$score)[1] # -2.1 # -0.5 # -0.15
maxY = base::range(cage_sel$score)[2] # 2.1 # 0.5 # 0.15
coef = 0.75
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

plot_ <- .GlobalEnv$plotTracks(cage_sel, ran) +
  geom_rect(
      inherit.aes = FALSE,
      data = sites_scores,
      mapping = aes(
          xmin = start,
          xmax = end,
          fill = score
        ),
      ymin = minY,
      ymax = maxY,
      # fill = "grey",
      alpha = 0.5
    ) +
  facet_wrap(
      facets = . ~ type + sample,
      ncol = 1,
      strip.position = "right",
      scales = "fixed" #"fixed" "free_y"
    ) +
  scale_fill_gradientn(
      limits = base::c(0,1),
      colours = cols
    ) +
  ylim(minY,maxY)  +
  guides(
      fill = guide_colourbar(
        label.position = "right",
        title = NULL,
        title.position = "left",
        title.theme = element_text(
            angle = 90,
            hjust = 0.5
          ),
        ticks=T,
        nbin=100,
        barheight=10,
        label=T,
        barwidth=0.5
      )
    ) # +
  # theme(
  #     aspect.ratio = 0.8
  #   )
  
plot <- ggrastr::rasterise(
    input = plot_,
    layers = "Bar",
    dpi = 300
  )
plot
```

## Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("trackHubFiles/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          # "_ABOprimelocus",
          # "_LPLprimelocus",
          # "_LIPCprimeloci",
          "_ABCA7primelocus",
          # "_TMPRSS6primelocus",
          "_unfilteredPRIMEloci",
          out,
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [5.2.] Plot c-KIT & KLF6 (Fig6)
```{r}
plot <- cage_sel_ %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = facets_sel,
        ordered = TRUE
      )
    # score = dplyr::case_when(
    #     strand == "minus" & type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ score,
    #     strand == "minus" & type %in% base::c("500K_N","5M_N","500K_C") ~ score*200,
    #     TRUE ~ score
    #   )
    ) %>%
  .GlobalEnv$plotTracks(., ran) +
    ggplot2::facet_wrap(
        facets = . ~ sample,
        ncol = 1,
        strip.position = "right",
        scales = "free_y"
      ) +
    scale_y_continuous(
        limits = base::c(-50,50),
        breaks = base::seq(-50,50,10)
      )
    # ggh4x::facetted_pos_scales(
    #   y = base::list(
    #     type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ scale_y_continuous(
    #         limits = base::c(-3,3)
    #       ),
    #     type %in% base::c("500K_N","5M_N","500K_C") ~ scale_y_continuous(
    #         limits = base::c(-300,300),
    #         breaks = base::c(-200,0,200),
    #         labels = base::c(-1,0,200)
    #       )
    #     )
    #   )
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

```{r}
start_pos <- BiocGenerics::start(range)
end_pos <- BiocGenerics::end(range)


test <- cage_sel_ %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = facets_sel,
        ordered = TRUE
      ),
    start = base::as.double(start)
    )
  ggplot(
        data = test,
        mapping = aes(
          x = start,
          y = score
          # colour = strand,
          # linetype = strand
        )
      ) +
    geom_bar()
        stat = "identity"
        # size = 0.5
      ) +
    ggplot2::scale_x_continuous(
        limits = base::c(start_pos,end_pos),  
        position = "top"
      ) +
    ggplot2::scale_color_manual(
        limits = base::c("minus","plus"),
        values = base::c("#FF6400","#632770"),
        guide = "none"
      ) +
    ggplot2::scale_linetype_manual(
        limits = base::c("minus","plus"),
        breaks = base::c("minus","plus"),
        values = base::c("solid","solid"),
        guide = "none"
      ) +
    ggplot2::facet_wrap(
        facets = . ~ sample,
        ncol = 1,
        strip.position = "right"
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(
        # legend.position = "none",
        axis.title.x = element_blank(),
        panel.spacing = grid::unit(0.05,"lines"),
        panel.grid.minor = element_blank(),
        plot.margin = grid::unit(base::c(0,0,0,0),"cm")
      )
```


## [6.] Genomic_landscape -------------------------

## [6.0.] Plot genomic environment
#### evaluate with 'columns(txdb)'
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

plot_land <- ggbio::autoplot(
      object = txdb,
      which = ran,
      truncate.gaps = FALSE,
      truncate.fun = NA,
      stat = "identity",
      label = FALSE
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  theme_bw() +
  theme(
      plot.margin = grid::unit(base::c(0.05,0,0,0), "lines")
    )
base::print(plot_land)
```

## [6.1.] Align genomic landscape and score plot
```{r}
g <- ggplot2::ggplotGrob(plot)
l <- ggplot2::ggplotGrob(plot_land@ggplot)

l_add <- gtable::gtable_add_cols(
    x = l,
    widths = grid::unit(1, "lines"),
    pos = -1
  )

l_add$widths <- g$widths

gridExtra::grid.arrange(
    g,l_add,
    heights = grid::unit(base::c(0.8, 0.2),"npc")
  )
```

## [6.2.] Print data
```{r}
# outfile <- base::paste0(
#     here::here("trackHubFiles/plots/"),
#     base::format(base::Sys.time(), "%Y%m%d"),
#     "_proximalTripGro",
#     out,
#     "_RAPGEF2",
#     "_GM12878",
#     "_poolsOnly",
#     ".pdf"
#   )
# 
# grDevices::pdf(
#     file = outfile,
#     width = 8,
#     height = 4
#   )
# base::print(
#     gridExtra::grid.arrange(
#       g,l_add,
#       heights = grid::unit(base::c(0.8, 0.2),"npc")
#     )
#   )
# grDevices::dev.off()
```

```{r}
outfile <- base::paste0(
    here::here("trackHubFiles/plots/"),
    base::format(base::Sys.time(), "%Y%m%d"),
    "_APOBgene",
    # "_ABCA7gene", # "_ABCA7enhancer",
    # "_TMPRSS6gene", # "_TMPRSS6promoter",
    # "_ABOgene", # "_ABOprimelocus", # "_ABOgene", # "_ABOenhancer_2",
    # "_LPLgene", # "_LPLenhancer",
    # "_LIPCgene",
    # "_APOC1intergenic",
    "_unfilteredPRIMEloci",
    # "_pooledPRIMEloci",
    # "_rescaled",
    out,
    ".pdf"
  )

grDevices::pdf(
    file = outfile,
    width = 8, # 4 # 8
    height = 4 # 6 # 4
  )
base::print(
    gridExtra::grid.arrange(
      g,l_add,
      heights = grid::unit(base::c(0.8, 0.2),"npc")
    )
  )
grDevices::dev.off()
```

```{r}
library(tidyverse)
ggplot(data = mtcars, aes(x = wt, y = mpg)) + geom_point()
```

