---
title: "20240411_workflow_trackVisualization"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [0.0.] Install necessary packages
```{r}
BiocManager::install("AnnotationHub")
BiocManager::install("remotes")
BiocManager::install("Bioconductor/RBGL")
BiocManager::install("ggbio")
```

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")
base::library("magrittr")
base::library("ggpubr")
base::library("here")
base::library("ggrastr")

base::library("AnnotationHub")

utils::installed.packages() %>%
  tibble::as_tibble() %>% 
  dplyr::filter(
    stringr::str_detect(Package,"TxDb.Hsapiens")
  )
base::library("TxDb.Hsapiens.UCSC.hg38.knownGene")

sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
base::source(base::paste(cfr.extensions.dir, "normalize.R", sep = sep))
```

## [1.] Prerequisites --------------------

## [1.0.] Read in annotations - convert from gff to gr
#### unique entries: 'hg38_gr[which(!is.na(elementMetadata(dm6_gr)[,"Name"])),]'
#### 'hg38_tbl <- tibble::as_tibble(hg38_gr)', 'dplyr::distinct(hg38_tbl,type)'
#### 'base::unique(S4Vectors::elementMetadata(hg38_gr)[,"type"])'
#### previous '/isdata/alab/genome/annotations/hg38/gencode.v31.annotation.gtf'
#### see 'https://www.gencodegenes.org/human/' for updates
```{r}
hg38_gff <- rtracklayer::readGFF(
    filepath = base::paste0(
      here::here("0.External_resources/"),
      "gencode.v39.annotation.gtf"
    )
  )
dplyr::distinct(hg38_gff,type)
hg38_gr <- GenomicRanges::makeGRangesFromDataFrame(
    df = hg38_gff,
    keep.extra.columns = TRUE
  )

types <- base::c(
    "transcription_start_site", "exon_junction", "exon",
    "gene", "pseudogene", "ncRNA",
    "intron", "mRNA", "protein",
    "CDS", "five_prime_UTR", "three_prime_UTR",
    "polyA_site", "snoRNA", "pre_miRNA",
    "miRNA", "silencer", "enhancer",
    "snRNA", "tRNA", "rRNA",
    "UTR"
  )
hg38_gr_sub <- hg38_gr[
    S4Vectors::elementMetadata(hg38_gr)[,"type"] %in% types,
  ]
```

## [1.1.] Color palette
```{r}
cols <- base::c(
    "#659DBD", "#FFC425", "#AFD275",
    "#9A1750", "#FC4445", "#55BCC9",
    "#242582", "#00B159", "#8860D0",
    "#41B3A3", "#C38D9E", "#8EE4AF",
    "#1F2833", "#D79922", "#7E685A",
    "#8D8741", "#E27D60", "#E8A87C",
    "#9B786F", "#DAAD86", "#EDF5E1",
    "#64485C"
  )
```

## [1.2.] Build lookup table for colors
#### 'col2rgb(cols3[1])[1:3,] %>% unname() %>% paste(collapse = ",")'
```{r}
annos <- base::unique(
    S4Vectors::elementMetadata(hg38_gr_sub)[,"type"]
  )

lkup <- tibble::tibble(
    txType = annos,
    hex = cols[1:base::length(annos)]
  )
```

## [1.4.] Load metainfo object
#### __not_in_use__
```{r eval=FALSE}
# metainfo <- SummarizedExperiment::colData(ctss) %>%
#   tibble::as_tibble(
#     rownames = "id"
#   )
```

## [1.5.] Identify files
```{r}
fs <- base::list.files(
      path = here::here("trackHubFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("PoolSub|repSub", negate = TRUE) %>% 
  stringr::str_subset("_500K_|_5M_")
```

## [1.6.] Load tmm.CTSSs object
#### see [1.15.2.] in "1.1.CTSSs.processing.Rmd" for more info
```{r}
fn <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("tmm.CTSSs")

tmm.CTSSs <- base::readRDS(file = fn)
```

## [1.7.] Load  object
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )
```

```{r}
ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      pooled == "no" &
      stringr::str_detect(ID, "500K|5M")
    ) %>%
  dplyr::pull(ID)

GM12878.CTSSs.sub <- S4Vectors::subset(
    x = GM12878.CTSSs,
    select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
  )
SummarizedExperiment::colData(GM12878.tmm.CTSSs.sub)
SummarizedExperiment::assay(GM12878.tmm.CTSSs.sub, "TMM")
```


## [1.8.] TMM normalise data
```{r}
facs <- edgeR::calcNormFactors(
    object = GM12878.tmm.CTSSs.sub,
    method = "TMM",
    refColumn = NULL
  )
GM12878.tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
    object = GM12878.CTSSs,
    sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
    inputAssay = "counts",
    outputAssay = "TMM"
  )
```

## [2.] CAGE_preparation_BigWig --------------------

## [2.0.] Helper function reader()
#### evaluate with 'f = fs[1]'
```{r}
readcage <- function(files, range){
  return(
    purrr::map(
      .x = files,
      .f = function(f){
          s <- base::basename(f) %>%
            stringr::str_replace(.,"^(.*?).bw$","\\1")
          return(
            base::suppressWarnings(
              rtracklayer::import(
                  con = f,
                  as = "GRanges",
                  selection = range
                ) %>%
              tibble::as_tibble() %>%
              dplyr::mutate(
                  sample = s
                )
              )
            )
          }
        ) %>%
      dplyr::bind_rows()
    )
  }
```

## [2.1.] Get genomic region from coordinates
#### suggested region "chr8:144,407,529-144,416,261", "chr10:110,291,360-110,366,023", "chrX:48,670,569-48,746,869"
#### enhancers: "chrX:48,676,170-48,676,946", "chrX:48,737,065-48,737,528", "chr9:41358750-41358940", "chr9:41,356,277-41,359,588", "chr8:134563800-134564100", "chr8:134,563,186-134,566,375"
```{r}
l_ <-  "chrX:48,670,569-48,746,869"
l <- stringr::str_replace_all(l_,",","")
chr = stringr::str_replace(l,"^(chr.*):.*","\\1")
s = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\1") %>%
  base::as.integer()
e = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\2") %>%
  base::as.integer()
out <- base::paste0("_",chr,"_",s,"_",e)
```

```{r}
extend = 0
ran <- GenomicRanges::GRanges(
    seqnames = chr,
    ranges = IRanges::IRanges(start = s, end = e)
  )
BiocGenerics::start(ran) <- BiocGenerics::start(ran) - extend
BiocGenerics::end(ran) <- BiocGenerics::end(ran) + extend
GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
```

## [2.2.] Get genomic region from gene names
#### examples are SLC39A4, YLPM1, SMNDC1
```{r}
# omr <- base::c(
#     "AREL1","FCF1","YLPM1","XR_944033.2",
#     "PROX2"
#   )
# ran_ <- hg38_gr_sub[
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"gene_name"] %in% omr,
#   ]
# base::unique(ran_$gene_name)
# ran <- base::range(ran_,ignore.strand = TRUE)
# 
# GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
# if(!stringr::str_detect(GenomeInfoDb::seqlevels(ran)[1],"chr")){
#   GenomeInfoDb::seqlevels(ran) <- base::gsub(
#       pattern = '^(.*)$',
#       replacement = 'chr\\1',
#       x = GenomeInfoDb::seqlevels(ran)
#     )
#   }
# GenomeInfoDb::seqlevels(ran)
```

## [2.3.] Extract coordinates
```{r}
chrom <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)
```

## [2.4.] Read in tpm scores in selected ranges
#### see [2.0.] in '20200721_workflow_DrosophilaCAGE.Rmd'
```{r}
ran_sel <- rtracklayer::BigWigSelection(
    ranges = ran
  )

cage_sel_ <- .GlobalEnv$readcage(fs,ran_sel)
```


```{r}
ovl <- IRanges::findOverlaps(
    query = ran,
    subject = GM12878.tmm.CTSSs #tmm.CTSSs
  )
gr <- SummarizedExperiment::rowRanges(GM12878.tmm.CTSSs)[ovl@to,] #tmm.CTSSs

crds <- gr %>% 
  base::as.data.frame() %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      start = pos,
      end = pos
    ) %>%
  dplyr::select(
      seqnames,start,end,strand
    )

scrs <- SummarizedExperiment::assay(GM12878.tmm.CTSSs,"TMM")[ovl@to,] %>% #tmm.CTSSs
  base::as.matrix() %>% 
  base::as.data.frame()

cage_sel_ <- dplyr::bind_cols(crds,scrs)
```

```{r}
cage_sel <- cage_sel_ %>%
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "sample",
      values_to = "tmm"
    ) %>%
  dplyr::mutate(
    replicate = stringr::str_replace(
        sample, ".*(.)$", "\\1"
      ),
    type = stringr::str_replace(
          sample, "^(.*?)_(.*)(_)?(.)", "\\2"
        ) %>%
      stringr::str_replace(
        .,"^(.*?)[_]$","\\1"
      )
    ) %>%
  dplyr::mutate(
    tmm = dplyr::case_when(
        strand == "-" ~ (-1)*tmm,
        TRUE ~ tmm
      )
    )
```

#### subset for "500K_C", "500K_N" and "5M_N" repliactes only
```{r}
cage_sel_reps <- cage_sel %>%
  dplyr::filter(
      stringr::str_detect(replicate, "\\d") &
      stringr::str_detect(type, "500K|5M")
    ) %>%
  dplyr::mutate(
      strand = dplyr::case_when(
        strand == "+" ~ "plus",
        strand == "-" ~ "minus"
      )
    )
```

## [2.5.] Summarise data
```{r}
cage_sel <- cage_sel_ %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "score"
    ) %>%
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::contains("_"),
        .fns = function(x) tidyr::replace_na(x, 0)
      )
    ) %>%
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "sample",
      values_to = "tpm"
    ) %>%
  dplyr::mutate(
    strand = stringr::str_replace(
        sample, "(.*)\\.TPM\\.(.*)", "\\2"
      ),
    replicate = stringr::str_replace(
        sample, ".*_(.*)\\.TPM\\.(.*)", "\\1"
      ),
    type = stringr::str_replace(
        sample, "^(.*?)_(.*)_(.*)", "\\2"
      ),
    ) %>%
  dplyr::mutate(
    tpm = dplyr::case_when(
        strand == "minus" ~ (-1)*tpm,
        TRUE ~ tpm
      )
    )
```

## [2.6.] Adjust source order
#### dplyr::distinct(cage_sel,source)
```{r}
levs <- base::c(
    "C", "C1", "C2", "C3",
    "N", "N1", "N2", "N3"
  )
cage_sel <- cage_sel %>%
  dplyr::mutate(
      replicate = base::factor(
          x = replicate,
          levels = levs,
          ordered = TRUE
        )
    ) %>%
  dplyr::arrange(
      type,replicate,seqnames,start,end
    )
```

## [3.] Plot data -------------------------

## [3.0.] Plot data
```{r}
plot <- ggplot2::ggplot(
      data = cage_sel_reps, #cage_sel,
      mapping = aes(
        x = start,
        y = tmm, #tpm,
        colour = strand,
        linetype = strand
      )
    ) +
  geom_bar(
      stat = "identity",
      size = 0.5
    ) +
  scale_color_manual(
      limits = base::c("minus","plus"),
      values = base::c("#FF6400","#632770")
    ) +
  scale_linetype_manual(
      limits = base::c("minus","plus"),
      breaks = base::c("minus","plus"),
      values = base::c("solid","solid")
    ) +
  facet_wrap(
      facets = . ~ type + replicate,
      ncol = 1,
      strip.position = "right"
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  # scale_y_continuous(
  #     name = "Average TPM",
  #     limits = #yRange #base::c((-1)*y,y)
  #   ) +
  # coord_cartesian(
  #     ylim = base::c((-1)*y,y)
  #   ) +
  theme_bw() +
  theme(
      aspect.ratio = 1/4,
      legend.position = "none",
      axis.title.x = element_blank(),
      panel.spacing = grid::unit(0.05,"lines"),
      panel.grid.minor = element_blank()
    )
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

## [3.2.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("trackHubFiles/"),
          format(Sys.time(), "%Y%m%d"),
          out,
          "_enhChr8Wider",
          "_500k5M",
          "_tpm",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 4,
      height = 8
    )
  }
```

## [4.] Genomic_landscape -------------------------

## [4.0.] Plot genomic environment
#### evaluate with 'columns(txdb)'
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

# base::debugonce(ggbio::autoplot)
plot_land <- ggbio::autoplot(
      txdb,
      which = ran,
      truncate.gaps = FALSE,
      truncate.fun = NA,
      coord = "genome",
      stat = "identity",
      label = FALSE
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1/4
    )
base::print(plot_land)
```

## [4.1.] Print data
```{r}
outfile <- base::paste0(
    here::here("trackHubFiles/"),
    format(Sys.time(), "%Y%m%d"),
    out,
    "_enh2",
    "_500k5M",
    "_landscape",
    ".pdf"
  )

grDevices::pdf(
    file = outfile,
    width = 8,
    height = 3
  )
base::print(
  ggbio::autoplot(
      txdb,
      which = ran,
      truncate.gaps = FALSE,
      truncate.fun = NA,
      coord = "genome",
      stat = "identity",
      label = FALSE
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  theme_bw() +
  theme(
        aspect.ratio = 1/4
      )
    )
grDevices::dev.off()
```

