---
title: "20240411_workflow_trackVisualization"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Packages -------------------------

## [0.0.] Install necessary packages
```{r}
BiocManager::install("AnnotationHub")
BiocManager::install("remotes")
BiocManager::install("Bioconductor/RBGL")
BiocManager::install("ggbio")
devtools::install_github("thomasp85/patchwork")
```

## [0.1.] Load necessary packages
```{r message=FALSE, warning=FALSE}
base::library("tidyverse")
base::library("magrittr")
base::library("ggpubr")
base::library("here")
base::library("ggrastr")
base::library("patchwork")

base::library("AnnotationHub")

utils::installed.packages() %>%
  tibble::as_tibble() %>% 
  dplyr::filter(
    stringr::str_detect(Package,"TxDb.Hsapiens")
  )
base::library("TxDb.Hsapiens.UCSC.hg38.knownGene")

sep <- .Platform$file.sep
cfr.extensions.dir <- "/projects/ralab/people/pbr316/scripts/CAGEfightR_extensions"
base::source(base::paste(cfr.extensions.dir, "normalize.R", sep = sep))
```

## [0.2.] Helper function readRange()
#### evaluate with 'f = fs[1]', 'files = fs', 'range = ran'
```{r}
readRange <- function(files, range){
    ## Prepare range for rtracklayer
    ran_sel <- rtracklayer::BigWigSelection(
        ranges = range
      )
    ## Reading in scores in range from files serially
    raw <- purrr::map(
      .x = files,
      .f = function(f){
          s <- base::basename(f) %>%
            stringr::str_replace(.,"^(.*?).bw$","\\1")
          return(
            base::suppressWarnings(
              rtracklayer::import(
                  con = f,
                  as = "GRanges",
                  selection = ran_sel
                ) %>%
              tibble::as_tibble() %>%
              dplyr::mutate(
                sample = s,
                strand = stringr::str_replace(
                    s, "^(.*)\\.(.*?)$", "\\2"
                  ),
                score = dplyr::case_when(
                    strand == "minus" ~ (-1)*score,
                    TRUE ~ score
                  )
                )
              )
            )
          }
        ) %>%
      dplyr::bind_rows()
    
    ## Fill missing scores across samples with zeros
    out <- .GlobalEnv$fillGaps(raw)
    
    return(out)
  }
```

## [0.3.] Helper function fillGaps()
#### Evaluate with 'data = raw'
```{r}
fillGaps <- function(data){
    columns <- data %>% 
      dplyr::distinct(sample) %>% 
      dplyr::pull(sample)
  
    df_filled <- data %>% 
      tidyr::pivot_wider(
          names_from = "sample",
          values_from = "score"
        ) %>%
      dplyr::mutate(
        dplyr::across(
          .cols = columns,
          .fns = function(x){
              tidyr::replace_na(x, 0)
            }
          )
        ) %>%
      tidyr::pivot_longer(
          cols = tidyselect::all_of(columns),
          names_to = "sample",
          values_to = "score"
        ) %>%
      dplyr::mutate(
        sample = stringr::str_replace(
            sample, ".minus|.plus", ""
          )
        )
    return(df_filled)
  }
```

## [0.4.] Helper function plotTracks()
#### Evaluate with 'data = cage_sel'
```{r}
plotTracks <- function(data, range){
    start_pos <- BiocGenerics::start(range)
    end_pos <- BiocGenerics::end(range)
    
    plot <- ggplot2::ggplot(
        data = data,
        mapping = aes(
          x = start,
          y = score,
          colour = strand,
          linetype = strand
        )
      ) +
    ggplot2::geom_bar(
        stat = "identity",
        size = 0.5
      ) +
    ggplot2::scale_x_continuous(
        limits = base::c(start_pos,end_pos),  
        position = "top"
      ) +
    ggplot2::scale_color_manual(
        limits = base::c("minus","plus"),
        values = base::c("#FF6400","#632770")
      ) +
    ggplot2::scale_linetype_manual(
        limits = base::c("minus","plus"),
        breaks = base::c("minus","plus"),
        values = base::c("solid","solid")
      ) +
    ggplot2::facet_wrap(
        facets = . ~ sample,
        ncol = 1,
        strip.position = "right"
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.spacing = grid::unit(0.05,"lines"),
        panel.grid.minor = element_blank(),
        plot.margin = grid::unit(base::c(0,0,0,0),"cm")
      )
    return(plot)
  }
```

## [1.] Prerequisites --------------------

## [1.0.] Read in annotations - convert from gff to gr
#### unique entries: 'hg38_gr[which(!is.na(elementMetadata(dm6_gr)[,"Name"])),]'
#### 'hg38_tbl <- tibble::as_tibble(hg38_gr)', 'dplyr::distinct(hg38_tbl,type)'
#### 'base::unique(S4Vectors::elementMetadata(hg38_gr)[,"type"])'
#### previous '/isdata/alab/genome/annotations/hg38/gencode.v31.annotation.gtf'
#### see 'https://www.gencodegenes.org/human/' for updates
```{r}
# hg38_gff <- rtracklayer::readGFF(
#     filepath = base::paste0(
#       here::here("0.External_resources/"),
#       "gencode.v39.annotation.gtf"
#     )
#   )
# dplyr::distinct(hg38_gff,type)
# hg38_gr <- GenomicRanges::makeGRangesFromDataFrame(
#     df = hg38_gff,
#     keep.extra.columns = TRUE
#   )
# 
# types <- base::c(
#     "transcription_start_site", "exon_junction", "exon",
#     "gene", "pseudogene", "ncRNA",
#     "intron", "mRNA", "protein",
#     "CDS", "five_prime_UTR", "three_prime_UTR",
#     "polyA_site", "snoRNA", "pre_miRNA",
#     "miRNA", "silencer", "enhancer",
#     "snRNA", "tRNA", "rRNA",
#     "UTR"
#   )
# hg38_gr_sub <- hg38_gr[
#     S4Vectors::elementMetadata(hg38_gr)[,"type"] %in% types,
#   ]
```

## [1.1.] Color palette
```{r}
cols <- base::c(
    "#659DBD", "#FFC425", "#AFD275",
    "#9A1750", "#FC4445", "#55BCC9",
    "#242582", "#00B159", "#8860D0",
    "#41B3A3", "#C38D9E", "#8EE4AF",
    "#1F2833", "#D79922", "#7E685A",
    "#8D8741", "#E27D60", "#E8A87C",
    "#9B786F", "#DAAD86", "#EDF5E1",
    "#64485C"
  )
```

## [1.2.] Build lookup table for colors
#### 'col2rgb(cols3[1])[1:3,] %>% unname() %>% paste(collapse = ",")'
```{r}
# annos <- base::unique(
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"type"]
#   )
# 
# lkup <- tibble::tibble(
#     txType = annos,
#     hex = cols[1:base::length(annos)]
#   )
```

## [1.4.] Load metainfo object
#### __not_in_use__
```{r eval=FALSE}
# metainfo <- SummarizedExperiment::colData(ctss) %>%
#   tibble::as_tibble(
#     rownames = "id"
#   )
```

## [1.5.] Identify files
```{r}
fs <- base::list.files(
      path = here::here("trackHubFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("PoolSub|repSub", negate = TRUE) %>% 
  stringr::str_subset("_500K_|_5M_")
```

## [1.6.] Load tmm.CTSSs object
#### see [1.15.2.] in "1.1.CTSSs.processing.Rmd" for more info
```{r}
# fn <- base::list.files(
#       path = here::here("1.CTSSs"),
#       full.names = TRUE
#     ) %>%
#   stringr::str_subset("tmm.CTSSs")
# 
# tmm.CTSSs <- base::readRDS(file = fn)
```

## [1.7.] Load  object
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )
```

## [1.8.] Subset for 500k nuclei and cell samples as well as 5mio nuclei
```{r}
ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      # pooled == "no" &
      stringr::str_detect(ID, "500K|5M")
    ) %>%
  dplyr::pull(ID)

GM12878.CTSSs.sub <- S4Vectors::subset(
    x = GM12878.CTSSs,
    select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
  )
SummarizedExperiment::colData(GM12878.CTSSs.sub)
SummarizedExperiment::assay(GM12878.CTSSs.sub, "counts")
```

## [1.9.] TMM normalise data
```{r}
# base::debugonce(edgeR::calcNormFactors)
# facs <- edgeR::calcNormFactors(
#     object = GM12878.CTSSs.sub,
#     method = "TMM",
#     refColumn = NULL
#   )
# GM12878.tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
#     object = GM12878.CTSSs.sub,
#     sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#     inputAssay = "counts",
#     outputAssay = "TMM"
#   )
```

#### Use TMM values as "counts" and calculate TPMs on top of this
```{r}
# GM12878.tmm.CTSSs <- .GlobalEnv$TPMnormalizeBySizeFactors(
#       object = GM12878.CTSSs.sub,
#       sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#       inputAssay = "counts",
#       outputAssay = "counts"
#     ) %>%
#   CAGEfightR::calcTPM()
```

#### Deconstructed edgeR::calcNormFactors() function
#### "GM12878_5M_N2" is the refColumn
```{r}
# object <- edgeR::SE2DGEList(GM12878.CTSSs.sub)
# x <- base::as.matrix(object)
# nsamples <- base::ncol(x)
# lib.size <- Matrix::colSums(x)
# allzero <- .Internal(rowSums(x > 0, base::nrow(x), nsamples, na.rm = FALSE)) == 0L
# x <- x[!allzero, , drop = FALSE]
# f75 <- base::suppressWarnings(
#    edgeR:::.calcFactorQuantile(
#        data = x,
#        lib.size = lib.size,
#        p = 0.75
#      )
#    )
# if (stats::median(f75) < 1e-20) {
#     refColumn <- BiocGenerics::which.max(base::colSums(base::sqrt(x)))
#   } else {
#     refColumn <- BiocGenerics::which.min(base::abs(f75 - base::mean(f75)))
#   }
```

## [1.10.] Read in CRISPR-gRNA locations
```{r}
dir = "/projects/ralab/people/pbr316/20221208_scCAGE/resources/"
crispr <- readr::read_delim(
    file = base::paste0(
        dir,
        "EPCrisprBenchmark_ensemble_data_GRCh38.tsv.gz"
      ),
    delim = "\t",
    col_names = TRUE
  )

myc_locus <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::summarise(
      chr = dplyr::first(chrom),
      start = base::min(chromStart),
      end = base::max(chromEnd)
    ) %>%
  GenomicRanges::GRanges()

crispr_data <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::mutate(
      y = dplyr::row_number(),
      y_alt = 1,
      summit = chromStart + base::round((chromEnd-chromStart+1)/2)
    )
```

# [2.] Prepare coordinate objects -------------------------

## [2.0.] Get genomic region from coordinates
#### suggested region "chr8:144,407,529-144,416,261", "chr10:110,291,360-110,366,023", 
#### wider area: "chrX:48,670,569-48,746,869" with enh1: "chrX:48,676,170-48,676,946" and enh2: "chrX:48,737,065-48,737,528"
#### enhancers - "chr9:41358750-41358940", "chr9:41,356,277-41,359,588", "chr8:134563800-134564100", "chr8:134,563,186-134,566,375"
```{r}
l_ <-  "chr10:110,291,360-110,366,023"
l <- stringr::str_replace_all(l_,",","")
chr = stringr::str_replace(l,"^(chr.*):.*","\\1")
s = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\1") %>%
  base::as.integer()
e = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\2") %>%
  base::as.integer()
out <- base::paste0("_",chr,"_",s,"_",e)
```

```{r}
extend = 0
ran <- GenomicRanges::GRanges(
    seqnames = chr,
    ranges = IRanges::IRanges(start = s, end = e)
  )
BiocGenerics::start(ran) <- BiocGenerics::start(ran) - extend
BiocGenerics::end(ran) <- BiocGenerics::end(ran) + extend
GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
```

## [2.1.] Get genomic region from gene names
#### examples are SLC39A4, YLPM1, SMNDC1
```{r}
# omr <- base::c(
#     "AREL1","FCF1","YLPM1","XR_944033.2",
#     "PROX2"
#   )
# ran_ <- hg38_gr_sub[
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"gene_name"] %in% omr,
#   ]
# base::unique(ran_$gene_name)
# ran <- base::range(ran_,ignore.strand = TRUE)
# 
# GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
# if(!stringr::str_detect(GenomeInfoDb::seqlevels(ran)[1],"chr")){
#   GenomeInfoDb::seqlevels(ran) <- base::gsub(
#       pattern = '^(.*)$',
#       replacement = 'chr\\1',
#       x = GenomeInfoDb::seqlevels(ran)
#     )
#   }
# GenomeInfoDb::seqlevels(ran)
```

## [2.2.] Get genomic regions from GRanges input
```{r}
gr <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC" &
      Significant == TRUE
    ) %>%
  dplyr::mutate(
      summit = chromStart + base::round((chromEnd - chromStart +1)/2)
    ) %>% 
  GenomicRanges::GRanges()
extend = 1000
BiocGenerics::start(gr) <- BiocGenerics::start(gr) - extend
BiocGenerics::end(gr) <- BiocGenerics::end(gr) + extend
GenomeInfoDb::seqlevelsStyle(gr) <- "UCSC"
myce38 <- IRanges::reduce(gr, ignore.strand = TRUE, with.revmap = TRUE)
S4Vectors::mcols(myce38)$smts <- S4Vectors::aggregate(gr, S4Vectors::mcols(myce38)$revmap, smts = summit)$smts
i = 1L
ran <- myce38[i]
```

## [2.2.] Extract coordinates
```{r}
chrom <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)
```

## [3.] CAGE_preparation_BigWig --------------------

## [3.1.] Read in tpm scores in selected ranges
#### see [2.0.] in '20200721_workflow_DrosophilaCAGE.Rmd'
```{r}
cage_sel_ <- .GlobalEnv$readRange(fs,ran)
```

## [3.2.] Summarise data
```{r}
cage_sel <- cage_sel_ %>%
  dplyr::mutate(
    replicate = stringr::str_replace(
        sample, ".*_(.*?)\\.TPM$", "\\1"
      ),
    type = stringr::str_replace(
        sample, "^(.*?)_(.*)_(.*)\\.TPM$", "\\2"
      ),
    )
```

## [3.3.] Adjust source order
#### dplyr::distinct(cage_sel,source)
```{r}
levs <- base::c(
    "C", "C1", "C2", "C3",
    "N", "N1", "N2", "N3"
  )
cage_sel <- cage_sel %>%
  dplyr::mutate(
      replicate = base::factor(
          x = replicate,
          levels = levs,
          ordered = TRUE
        )
    ) %>%
  dplyr::arrange(
      type,replicate,seqnames,start,end
    )
```

## [4.] CAGE_preparation_CTSSs --------------------

## [4.0.] Subset & convert scores from RangedSummarizedExperiment to tibble
```{r}
ovl <- IRanges::findOverlaps(
    query = ran,
    subject = GM12878.CTSSs.sub # GM12878.tmm.CTSSs
  )
gr <- SummarizedExperiment::rowRanges(GM12878.CTSSs.sub)[ovl@to,] # GM12878.tmm.CTSSs

crds <- gr %>% 
  base::as.data.frame() %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      start = pos,
      end = pos
    ) %>%
  dplyr::select(
      seqnames,start,end,strand
    )

# scrs <- SummarizedExperiment::assay(GM12878.tmm.CTSSs,"TMM")[ovl@to,] %>%
#   base::as.matrix() %>% 
#   base::as.data.frame()

scrs <- SummarizedExperiment::assay(GM12878.CTSSs.sub,"TPM")[ovl@to,] %>% # GM12878.tmm.CTSSs
  base::as.matrix() %>%
  base::as.data.frame()

cage_sel_ <- dplyr::bind_cols(crds,scrs)
```

## [4.1.] Summarise data
#### Evaluate with 'dplyr::arrange(cage_sel,dplyr::desc(score))'
```{r}
cage_sel <- cage_sel_ %>%
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "sample",
      values_to = "score"
    ) %>%
  dplyr::mutate(
    replicate = stringr::str_replace(
        sample, ".*(.)$", "\\1"
      ),
    type = stringr::str_replace(
          sample, "^(.*?)_(.*)(_)?(.)", "\\2"
        ) %>%
      stringr::str_replace(
        .,"^(.*?)[_]$","\\1"
      )
    ) %>%
  dplyr::mutate(
    score = dplyr::case_when(
        strand == "-" ~ (-1)*score,
        TRUE ~ score
      ),
    strand = dplyr::case_when(
        strand == "+" ~ "plus",
        strand == "-" ~ "minus"
      )
    )
```

# [5.] Plot data -------------------------

## [5.0.] Plot genome tracks
```{r}
plot <- .GlobalEnv$plotTracks(cage_sel, ran) +
  ggplot2::facet_wrap(
      facets = . ~ type + replicate,
      ncol = 1,
      strip.position = "right"
    )
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

## [5.1.] Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("trackHubFiles/"),
          format(Sys.time(), "%Y%m%d"),
          out,
          # "_enh2",
          "_500k5M",
          "_TmmTpm",
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 4,
      height = 8
    )
  }
```

## [6.] Genomic_landscape -------------------------

## [6.0.] Plot genomic environment
#### evaluate with 'columns(txdb)'
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

plot_land <- ggbio::autoplot(
      object = txdb,
      which = ran,
      truncate.gaps = FALSE,
      truncate.fun = NA,
      stat = "identity",
      label = FALSE
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  theme_bw() +
  theme(
      plot.margin = grid::unit(base::c(0.05,0,0,0), "lines")
    )
base::print(plot_land)
```

## [6.1.] Align genomic landscape and score plot
```{r}
g <- ggplot2::ggplotGrob(plot)
l <- ggplot2::ggplotGrob(plot_land@ggplot)

l_add <- gtable::gtable_add_cols(
    x = l,
    widths = grid::unit(1, "lines"),
    pos = -1
  )

l_add$widths <- g$widths

gridExtra::grid.arrange(
    g,l_add,
    heights = grid::unit(base::c(0.8, 0.2),"npc")
  )
```

## [6.1.] Print data
```{r}
outfile <- base::paste0(
    here::here("trackHubFiles/"),
    format(Sys.time(), "%Y%m%d"),
    out,
    "_500k5M",
    ".pdf"
  )

grDevices::pdf(
    file = outfile,
    width = 8,
    height = 4
  )
base::print(
    gridExtra::grid.arrange(
      g,l_add,
      heights = grid::unit(base::c(0.8, 0.2),"npc")
    )
  )
grDevices::dev.off()
```
