---
title: "workflow_trackVisualization"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Packages -------------------------

## [0.0.] Install necessary packages
#### Maybe required: "base::detach(name = tidyverse, unload = TRUE)", "utils::remove.packages("tidyverse")", "utils::remove.packages("ggplot2")"
#### Lib locations: "/maps/projects/scarball/apps/Modules/software/R-packages/4.2.2", "/maps/projects/ralab/people/pbr316/bin/Rpackages"
```{r eval=FALSE}
# # tidyverse
# utils::install.packages(
#     pkgs = "tidyverse",
#     lib = "/maps/projects/ralab/people/pbr316/bin/Rpackages"
#   )
# 
# ## Ensure ggplot2 version 3.5.2 is installed
# utils::remove.packages("ggplot2")
# utils::install.packages(
#     "https://cran.r-project.org/src/contrib/Archive/ggplot2/ggplot2_3.5.2.tar.gz",
#     type = "source",
#     repos = NULL
#   )
# utils::packageVersion("ggplot2")
# 
# BiocManager::install("AnnotationHub")
# BiocManager::install("remotes")
# BiocManager::install("Bioconductor/RBGL")
# BiocManager::install("ggbio")
# devtools::install_github("thomasp85/patchwork")
# utils::install.packages("ggh4x")
# 
# utils::install.packages(
#     "xfun",
#     dependencies = TRUE,
#     INSTALL_opts = '--no-lock'
#   )
# utils::install.packages("S7")
```

## [0.1.] Load necessary packages
#### Evaluate with sessioninfo::session_info()
```{r message=FALSE, warning=FALSE}
base::library(tidyverse)
base::library(PRIME)
utils::packageVersion("ggplot2") # 4.0.1 # 4.0.3 (20260429)

# base::library(tidyverse, lib.loc = "/maps/projects/ralab/people/pbr316/bin/Rpackages")
# base::library("AnnotationHub")
# base::library("parallel")
# base::library("foreach")
# base::library("doParallel")
# doParallel::registerDoParallel(
#     cores = 15
#   )

# utils::installed.packages() %>%
#   tibble::as_tibble() %>% 
#   dplyr::filter(
#     stringr::str_detect(Package,"TxDb.Hsapiens")
#   )
base::library("TxDb.Hsapiens.UCSC.hg38.knownGene")
base::library("BSgenome.Hsapiens.UCSC.hg38")

```


## [1.] Read in data --------------------

## [1.0.] Color palette
```{r}
cols <- base::c(
    "#659DBD", "#FFC425", "#AFD275",
    "#9A1750", "#FC4445", "#55BCC9",
    "#242582", "#00B159", "#8860D0",
    "#41B3A3", "#C38D9E", "#8EE4AF",
    "#1F2833", "#D79922", "#7E685A",
    "#8D8741", "#E27D60", "#E8A87C",
    "#9B786F", "#DAAD86", "#EDF5E1",
    "#64485C"
  )
```

## [1.1.] Read in annotations - convert from gff to gr
#### unique entries: 'hg38_gr[which(!is.na(elementMetadata(dm6_gr)[,"Name"])),]'
#### 'hg38_tbl <- tibble::as_tibble(hg38_gr)', 'dplyr::distinct(hg38_tbl,type)'
#### 'base::unique(S4Vectors::elementMetadata(hg38_gr)[,"type"])'
#### previous '/isdata/alab/genome/annotations/hg38/gencode.v31.annotation.gtf'
#### see 'https://www.gencodegenes.org/human/' for updates
```{r}
# hg38_gff <- rtracklayer::readGFF(
#     filepath = base::paste0(
#       here::here("0.External_resources/"),
#       "gencode.v39.annotation.gtf"
#     )
#   )
# dplyr::distinct(hg38_gff,type)
# hg38_gr <- GenomicRanges::makeGRangesFromDataFrame(
#     df = hg38_gff,
#     keep.extra.columns = TRUE
#   )
# 
# types <- base::c(
#     "transcription_start_site", "exon_junction", "exon",
#     "gene", "pseudogene", "ncRNA",
#     "intron", "mRNA", "protein",
#     "CDS", "five_prime_UTR", "three_prime_UTR",
#     "polyA_site", "snoRNA", "pre_miRNA",
#     "miRNA", "silencer", "enhancer",
#     "snRNA", "tRNA", "rRNA",
#     "UTR"
#   )
# hg38_gr_sub <- hg38_gr[
#     S4Vectors::elementMetadata(hg38_gr)[,"type"] %in% types,
#   ]
```

## [1.2.] Build lookup table for colors associated with annotations
#### 'col2rgb(cols3[1])[1:3,] %>% unname() %>% paste(collapse = ",")'
```{r}
# annos <- base::unique(
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"type"]
#   )
# 
# lkup <- tibble::tibble(
#     txType = annos,
#     hex = cols[1:base::length(annos)]
#   )
```

## ---------------

## [1.3.] Read in bigwig files - K562 and GM12878
```{r}
fs <- base::list.files(
      path = here::here("trackHubFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("PoolSub|repSub", negate = TRUE) %>% 
  stringr::str_subset("_500K_|_5M_") # GM12878-specific
  # stringr::str_subset("K562") # K562-specific
```

## [1.4.] Read in bigwig files - FANTOM5 immune tracks (Fig6)
#### SMAD3 (Fig6)
```{r}
facets_sel <- base::c(
    "mast cell", "dendritic cell", "monocyte",
    "T cell", "reticulocyte"
  )

fs <- base::list.files(
      path = here::here("0.External_resources/FANTOM5/pooled_bw"),
      full.names = TRUE
    ) %>%
  stringr::str_subset(base::paste0(facets_sel,collapse = "|"))
```

#### ACOT11, NCOR2, TRDMT1, CD82 (Fig6)
```{r}
# facets_sel <- base::c(
#     "B_lymphoblastoid", "Hep-2", "astrocyte",
#     "chronic_myelogenous_leukemia", "epitheloid_carcinoma", "kidney epithelial cell",
#     "mast cell", "mesenchymal cell", "monocyte",
#     "natural killer cell", "neuroblastoma", "skin fibroblast",
#     "small_cell_lung_carcinoma", "thymus", "vascular associated smooth muscle cell"
#   )
# facets_sel <- base::c(
#     "mast cell", "dendritic cell", "monocyte",
#     "T cell", "reticulocyte", "cervical_cancer",
#     "placental epithelial cell", "uterus"
#   )
# facets_sel <- base::c(
#     "mast cell", "dendritic cell", "monocyte",
#     "T cell", "reticulocyte", "hepatocyte",
#     "large intestine", "small intestine", "preadipocyte",
#     "liposarcoma", "fat cell", "hepatocellular_carcinoma",
#     "Hep-2", "liver", "hepatic stellate cell",
#     "endothelial cell of hepatic sinusoid"
#   )
# facets_sel <- base::c(
#     "mast cell", "dendritic cell", "monocyte",
#     "T cell", "reticulocyte", "natural killer cell"
#   )
# facets_sel <- base::c(
#     "mast cell", "dendritic cell", "monocyte",
#     "T cell", "reticulocyte", "acute_myeloid_leukemia_\\(FAB_M6\\)",
#     "neutrophil"
#   )

fs <- base::list.files( 
      path = here::here("0.External_resources/FANTOM5/pooled_bw"),
      full.names = TRUE
    ) %>%
  stringr::str_subset(base::paste0(facets_sel,collapse = "|"))
```

#### FANTOM5 kidney-related tracks (Fig6)
```{r}
facets_sel_ <- base::c(
    "spinal cord", "internal male genitalia", "throat",
    "tongue", "testis", "thyroid gland",
    "blood vessel", "spleen", "thymus",
    "meninx", "eye", "kidney",
    "uterus", "lung", "large.intestine",
    "small.intestine"
  )

facets_sel <- facets_sel_ %>%
  base::paste0("/",., "\\.") %>%
  base::paste0(collapse = "|")

fs <- base::list.files(
      path = here::here("0.External_resources/FANTOM5/pooled_bw"),
      full.names = TRUE
    ) %>%
  stringr::str_subset(facets_sel)
  # stringr::str_subset("TPM1")
```

## ---------------

## [1.5.] Read in CTSSs - Load HepG2-, HCT116-, A549-CTSS objects & adapt names (Fig5)
#### Evaluate with 'rsen = "A549"', 'SummarizedExperiment::colData(CTSSs.objects[[rsen]])'
```{r}
CTSSs.objects_ <- base::list.files(
      path = here::here("1.CTSSs"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("\\/poolSub") %>%
  stringr::str_subset("HepG2|HCT116|A549")
  # stringr::str_subset("HepG2|HCT116|A549|combined.GM12878|combined.k562")

CTSSs.objects <- purrr::map(
    .x = CTSSs.objects_,
    .f = base::readRDS
  )
base::names(CTSSs.objects) <- stringr::str_replace(
    CTSSs.objects_, ".*/poolSub(\\.combined)?\\.(.*?)\\.CTSSs.*", "\\2"
  )
CTSSs.objects <- purrr::map(
    .x = base::names(CTSSs.objects),
    .f = function(rsen){
      base::rownames(SummarizedExperiment::colData(CTSSs.objects[[rsen]])) <- base::paste0(
          rsen, "_",
          base::rownames(SummarizedExperiment::colData(CTSSs.objects[[rsen]]))
        )
      return(CTSSs.objects[[rsen]])
    }
  )
base::names(CTSSs.objects) <- stringr::str_replace(
    CTSSs.objects_, ".*/poolSub(\\.combined)?\\.(.*?)\\.CTSSs.*", "\\2"
  )
```

## [1.6.] Read in CTSSs - Add the corresponding K562 poolSub data to CTSSs.objects of [1.5.] (Fig5)
#### Evaluate with SummarizedExperiment::colData(K562.CTSSs.sub), SummarizedExperiment::assay(K562.CTSSs.sub, i = "counts.noSingletons")
```{r}
K562.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = K562.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
      # stringr::str_detect(type, "K562_C$|K562_N$")
    stringr::str_detect(type, "K562_C$|K562_N$|K562_N_10M$") # (20260408)
    ) %>%
  dplyr::pull(ID)

# K562.CTSSs.sub <- S4Vectors::subset(
#     x = K562.CTSSs,
#     select = SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
#   )

tpm <- SummarizedExperiment::assay(K562.CTSSs,"TPM")[,SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids]
K562.CTSSs.sub <- SummarizedExperiment::SummarizedExperiment(
    assays = base::list(
        TPM = SummarizedExperiment::assay(K562.CTSSs,"TPM")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts = SummarizedExperiment::assay(K562.CTSSs,"counts")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts.noSingletons = SummarizedExperiment::assay(K562.CTSSs,"counts.noSingletons")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
          ] # 8810502, 2
      ),
    colData = SummarizedExperiment::colData(K562.CTSSs)[SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids,],
    rowRanges = SummarizedExperiment::rowRanges(K562.CTSSs)[(Matrix::rowSums(tpm) != 0),]
  )

CTSSs.objects[["K562"]] <- K562.CTSSs.sub
base::remove(K562.CTSSs,K562.CTSSs.sub)
```

#### Evaluations
```{r}
SummarizedExperiment::assay(K562.CTSSs,"TPM")[1:13,]
SummarizedExperiment::assay(K562.CTSSs.sub,"TPM")[1:13,]
SummarizedExperiment::assay(CTSSs.objects[["K562"]],"TPM")[1:13,]

assertthat::are_equal(
    SummarizedExperiment::rowRanges(K562.CTSSs)[1030345:1030545],
    base::sort(SummarizedExperiment::rowRanges(K562.CTSSs))[1030345:1030545]
  )
setdiff(
    SummarizedExperiment::rowRanges(K562.CTSSs),
    base::sort(SummarizedExperiment::rowRanges(K562.CTSSs))
  )
tpm <- SummarizedExperiment::assay(K562.CTSSs.sub,"TPM")
tpm[(Matrix::rowSums(tpm) != 0),] # 8,810,502
SummarizedExperiment::rowRanges(K562.CTSSs)[(Matrix::rowSums(tpm) != 0),] # 8,810,502
```

## [1.7.] Read in CTSSs - Add the corresponding GM12878 poolSub data to CTSSs.objects of [1.5.] (Fig5)
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
     stringr::str_detect(type, "GM12878_1M_C$|GM12878_1M_N$")
    ) %>%
  dplyr::pull(ID)

# GM12878.CTSSs.sub <- S4Vectors::subset(
#     x = GM12878.CTSSs,
#     select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
#   )

tpm <- SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[,SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids]
GM12878.CTSSs.sub <- SummarizedExperiment::SummarizedExperiment(
    assays = base::list(
        TPM = SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts = SummarizedExperiment::assay(GM12878.CTSSs,"counts")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ], # 8810502, 2
        counts.noSingletons = SummarizedExperiment::assay(GM12878.CTSSs,"counts.noSingletons")[
            (Matrix::rowSums(tpm) != 0),
            SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
          ] # 8810502, 2
      ),
    colData = SummarizedExperiment::colData(GM12878.CTSSs)[SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids,],
    rowRanges = SummarizedExperiment::rowRanges(GM12878.CTSSs)[(Matrix::rowSums(tpm) != 0),]
  )

CTSSs.objects[["GM12878"]] <- GM12878.CTSSs.sub
# SummarizedExperiment::colData(GM12878.CTSSs.sub)
base::remove(GM12878.CTSSs,GM12878.CTSSs.sub)
```

#### Evaluations
```{r}
base::names(CTSSs.objects)
SummarizedExperiment::assay(GM12878.CTSSs,"TPM")[1:13,]
SummarizedExperiment::assay(GM12878.CTSSs.sub,"TPM")[13,]
SummarizedExperiment::assay(CTSSs.objects[["GM12878"]],"TPM")[1:13,]

assertthat::are_equal(
    SummarizedExperiment::rowRanges(GM12878.CTSSs)[1030345:1030545],
    base::sort(SummarizedExperiment::rowRanges(GM12878.CTSSs))[1030345:1030545]
  )
setdiff(
    SummarizedExperiment::rowRanges(GM12878.CTSSs),
    base::sort(SummarizedExperiment::rowRanges(GM12878.CTSSs))
  )
```

## ---------------

## [1.8.] Read in CTSSs - Load prepared 'all CTSS' - objects for GM12878 & K562 (Fig3)
#### Evaluate with 'SummarizedExperiment::colData(GM12878.CTSSs.sub)' & 'SummarizedExperiment::assay(GM12878.CTSSs.sub, "counts")'
```{r}
GM12878.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.GM12878.CTSSs.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
  tibble::as_tibble(rownames = "type") %>%
  dplyr::filter(
      (pooled == "yes" & stringr::str_detect(type, "trip|ctrl")) |
      (stringr::str_detect(type, "groCap")) |
      (pooled == "yes" & stringr::str_detect(ID, "500K|5M"))
    ) %>%
  dplyr::pull(ID)

GM12878.CTSSs.sub <- S4Vectors::subset(
    x = GM12878.CTSSs,
    select = SummarizedExperiment::colData(GM12878.CTSSs)$ID %in% ids
  )
base::remove(GM12878.CTSSs)
```

#### Identify and subset for specific samples, e.g., 5M, Triptolide or GROcap
```{r}
# ##GM12878 5M and 500k samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       pooled == "yes" &
#       stringr::str_detect(ID, "500K|5M")
#     ) %>%
#   dplyr::pull(ID)
# 
# ##Triptolide samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       stringr::str_detect(Pool, "trip")
#     ) %>%
#   dplyr::pull(ID)
# 
# ##GRO-cap samples
# ids <- SummarizedExperiment::colData(x = GM12878.CTSSs) %>%
#   tibble::as_tibble(rownames = "type") %>%
#   dplyr::filter(
#       stringr::str_detect(type, "groCap")
#     ) %>%
#   dplyr::pull(ID)
```

#### Evaluate with 'SummarizedExperiment::colData(K562.CTSSs.sub)' & 'SummarizedExperiment::assay(K562.CTSSs.sub, "counts")'
#### Different from "poolSub.combined.GM12878.CTSSs.rmSingletons.rds" in [1.6.]
```{r}
K562.CTSSs <- base::readRDS(
  file = base::paste0(
      here::here("1.CTSSs/"),
      "all.combined.k562.CTSSs.rds"
    )
  )

ids <- SummarizedExperiment::colData(x = K562.CTSSs) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      pooled == "yes" & # Only pooled replicates for figure plots
      stringr::str_detect(ID, "Cap", negate = TRUE)
    ) %>%
  dplyr::pull(ID)

K562.CTSSs.sub <- S4Vectors::subset(
    x = K562.CTSSs,
    select = SummarizedExperiment::colData(K562.CTSSs)$ID %in% ids
  )
```

## ---------------

## [1.9.] Write out bigWig files
```{r}
samples = base::rownames(
    SummarizedExperiment::colData(GM12878.CTSSs.sub)
  )

PRIME::writeBw(
    object = GM12878.CTSSs,
    replicates = samples,
    dir = base::paste0(
        here::here("trackHubFiles/"),
        "20240718_bigWig/"
      ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

```{r}
samples = base::rownames(
    SummarizedExperiment::colData(K562.CTSSs.sub)
  )

PRIME::writeBw(
    object = K562.CTSSs,
    replicates = samples,
    dir = base::paste0(
        here::here("trackHubFiles/"),
        "20240718_bigWig/"
      ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

#### Write out HepG2-, HCT116-, A549-CTSS samples
```{r}
purrr::map(
    .x = CTSSs.objects,
    .f = PRIME::writeBw,
    dir = base::paste0(
          here::here("trackHubFiles/"),
          "20240718_bigWig/"
        ),
    inputAssay = "TPM",
    splitByStrand = TRUE
  )
```

## ---------------

## [1.10.] TMM normalise data
```{r}
# base::debugonce(edgeR::calcNormFactors)
# facs <- edgeR::calcNormFactors(
#     object = GM12878.CTSSs.sub,
#     method = "TMM",
#     refColumn = NULL
#   )
# GM12878.tmm.CTSSs <- PRIME::TPMnormalizeBySizeFactors(
#     object = GM12878.CTSSs.sub,
#     sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#     inputAssay = "counts",
#     outputAssay = "TMM"
#   )
```

#### Use TMM values as "counts" and calculate TPMs on top of this
```{r}
# GM12878.tmm.CTSSs <- PRIME::TPMnormalizeBySizeFactors(
#       object = GM12878.CTSSs.sub,
#       sizeFactors = facs$samples$lib.size * facs$samples$norm.factors,
#       inputAssay = "counts",
#       outputAssay = "counts"
#     ) %>%
#   CAGEfightR::calcTPM()
```

#### Deconstructed edgeR::calcNormFactors() function
#### "GM12878_5M_N2" is the refColumn
```{r}
# object <- edgeR::SE2DGEList(GM12878.CTSSs.sub)
# x <- base::as.matrix(object)
# nsamples <- base::ncol(x)
# lib.size <- Matrix::colSums(x)
# allzero <- .Internal(rowSums(x > 0, base::nrow(x), nsamples, na.rm = FALSE)) == 0L
# x <- x[!allzero, , drop = FALSE]
# f75 <- base::suppressWarnings(
#    edgeR:::.calcFactorQuantile(
#        data = x,
#        lib.size = lib.size,
#        p = 0.75
#      )
#    )
# if (stats::median(f75) < 1e-20) {
#     refColumn <- BiocGenerics::which.max(base::colSums(base::sqrt(x)))
#   } else {
#     refColumn <- BiocGenerics::which.min(base::abs(f75 - base::mean(f75)))
#   }
```

#### Load tmm.CTSSs object
#### see [1.15.2.] in "1.1.CTSSs.processing.Rmd" for more info
```{r}
# fn <- base::list.files(
#       path = here::here("1.CTSSs"),
#       full.names = TRUE
#     ) %>%
#   stringr::str_subset("tmm.CTSSs")
# 
# tmm.CTSSs <- base::readRDS(file = fn)
```

## ---------------

## [1.12.] Read in CRISPR-gRNA locations
```{r}
dir = "/projects/ralab/people/pbr316/20221208_scCAGE/resources/"
crispr <- readr::read_delim(
    file = base::paste0(
        dir,
        "EPCrisprBenchmark_ensemble_data_GRCh38.tsv.gz"
      ),
    delim = "\t",
    col_names = TRUE
  )

myc_locus <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::summarise(
      chr = dplyr::first(chrom),
      start = base::min(chromStart),
      end = base::max(chromEnd)
    ) %>%
  GenomicRanges::GRanges()

crispr_data <- crispr %>%
  dplyr::filter(
      measuredGeneSymbol == "MYC"
    ) %>%
  dplyr::mutate(
      y = dplyr::row_number(),
      y_alt = 1,
      summit = chromStart + base::round((chromEnd-chromStart+1)/2)
    )
```

## [1.13.] Read in MYC enhancer coordinates (20260430)
```{r}
mycs <- rtracklayer::import.bed(
    con = base::paste0(
        here::here("0.External_resources/"),
        "20260501_MYC_superenhancers_hg38.bed"
      )
    ) %>%
  tibble::as_tibble() %>%
  dplyr::select(seqnames, start, end)
```

## [1.14.] Read in K562 enhancer coordinates (20260505)
```{r}
enhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "20260505_enhancers_K562_v1.csv"
        ),
      delim = ";",
      col_names = TRUE,
      col_types = readr::cols(
        .default = readr::col_integer(),
        sample = readr::col_character(),
        seqname = readr::col_character(),
        type = readr::col_character(),
        reference = readr::col_character(),
        comment = readr::col_character()
      )
    ) %>%
  dplyr::mutate(
      loc = base::paste0(seqname, ":", start, "-", end)
    )
```

# [2.] Prepare coordinate objects -------------------------

## [2.0.] Define coordinates

#### Fig1 examples
#### suggested region "chr8:144,407,529-144,416,261", "chr10:110,291,360-110,366,023", 
#### wider area: "chrX:48,670,569-48,746,869" with enh1: "chrX:48,676,170-48,676,946" and enh2: "chrX:48,737,065-48,737,528"
#### enhancers - "chr9:41358750-41358940", "chr9:41,356,277-41,359,588", "chr8:134563800-134564100", "chr8:134,563,186-134,566,375"
#### MYC-enhancer: "chr8:127,884,215-127,915,852"
#### MYC-locus & PVT1-promoter: "chr8:127,704,300-127,810,400"
#### HDAC6 promoter: "chrX:48,801,000-48,803,000"
#### RUNX1 promoter: "chr21:35,048,764-35,049,617"
#### MEF2A promoter: "chr15:99,565,263-99,566,083"
#### GATA1 promoters: "chrX:48,786,301-48,788,293" & enhancer: "chrX:48,787,857-48,788,321"
#### KLF2 promoter: "chr19:16,324,467-16,325,106"
#### BCL2 promoter/PROMPTs: "chr18:63,317,868-63,320,926"
#### NAA1 / LIN01088 intron: "chr4:144,080,727-144,158,976"
```{r}
l_ <- "chrX:48,786,301-48,788,293" # GATA1 promoter
l_ <- "chrX:48,787,857-48,788,321" # GATA1 enhancer

l_ <- "chrX:12950917-12951208" # TMSB4X enhancer

l_ <- "chr20:47360724-47361135" # ZMYND8 enhancer

l_ <- "chr7:12,684,956-12,724,970" # ARL4A genewider
l_ <- "chr7:12722785-12723156" # ARL4A enhancer
```

#### see [4.10.] in 10.3.Cell_line_vs_trait_enrichment.Rmd for 'chr22:37109513:37109512'

#### RAPGEF2 promoter: "chr4:159,101,737-159,104,305" (Fig3)
```{r}
l_ <- "chr4:159,101,737-159,104,305" # RAPGEF2 promoter combined

l_ <- "chr4:159,101,737-159,102,970" # RAPGEF2 PROMPT wider
l_ <- "chr4:159,102,750-159,102,970" # RAPGEF2 PROMPT

l_ <- "chr4:159,102,971-159,104,305" # RAPGEF2 promoter wider
l_ <- "chr4:159,102,971-159,103,250" # RAPGEF2 promoter
```

### MCHC trait (Fig5)
#### Whole ABCA7 gene including "chr19:1050876-1050875": "chr19:1039521-1065670"
#### Whole TMPRSS6 gene including "chr22:37109513:37109512": "chr22:37,065,000-37,110,700"
```{r}
l_ <- "chr19:1,039,521-1,065,670" # ABCA7
l_ <- "chr19:1,050,492-1,051,269" # ABCA7 Primeloci

l_ <- "chr22:37,065,000-37,110,700" # TMPRSS6
l_ <- "chr22:37,109,199-37,109,898" # TMPRSS6 Primeloci
```

### ApoB trait (Fig5)
#### Whole ApoB gene locus as control: "chr2:20,999,429-21,046,073"
#### Whole ABO gene including "chr9:133274296:133274295" and "chr9:133263363:133263374": "chr9:133,250,000-133,276,238"
#### Whole LPL gene including: "chr8:19961929-19961928": "chr8:19,938,800-19,968,800"
#### Whole LIPC gene including: "chr15:58430392:58430391": "chr15:58,400,000-58,570,000"

#### Whole APOC1 gene including "chr19:44913485:44913484": "chr19:44,914,000-44,920,000" & with alternatively entire promoter proximal region "chr19:44,913,101-44,914,607"
#### APOE - APOC1: "chr19:44,908,532-44,915,200"
```{r}
l_ <- "chr2:20,999,429-21,046,073"

l_ <- "chr9:133,250,000-133,276,238" # ABO
l_ <- "chr9:133,273,899-133,274,680" # ABO PRIMElocus

l_ <- "chr8:19,938,800-19,968,800" # LPL
l_ <- "chr8:19,961,537-19,962,318" # LPL Primelocus

l_ <- "chr15:58,400,000-58,570,000" # LIPCs
l_ <- "chr15:58429996-58435030" # LIPCs PRIMEloci

l_ <- "chr19:44,908,532-44,915,200" # APOC1
```

### c-KIT & KLF6 (Fig6)
```{r}
l_ <- "chr4:54,360,000-55,036,000" # c-KIT gene wider
l_ <- "chr4:54,538,000-54,760,000" # c-KIT gene
l_ <- "chr4:54,635,119-54,635,915" # c-KIT enhancer
# l_ <- "chr4:54,708,916-54,709,316" # (20260228)
l_ <- "chr4:54542425-54543225" # c-KIT enhancer (20260303, deficiency anemia)

# l_ <- "chr10:3,700,000-4,146,000" # KLF6 # wider area
l_ <- "chr10:3,715,630-3,960,899" # KLF6
l_ <- "chr10:3,942,689-3,943,182" # KLF6 enhancer
#l_ <- "chr10:3,740,001-3,740,401" # (20260228)
```

#### Examples immune (Fig6)
```{r}
# l_ <- "chr9:127,737,317-127,779,666" # SH2D3C
# l_ <- "chr9:127776431-127776831"
# l_ <- "chr21:37,361,584-37,657,699" # DYRK1A
# l_ <- "chr21:37563548-37563948"

# l_ <- "chr1:6,219,522-6,361,546" # GPR153 (ACOT7)
# l_ <- "chr1:6301631-6302031"

# l_ <-  "chr3:100,679,296-100,760,074" # TFG
# l_ <- "chr3:100,702,114-100,702,514" # TFG enhancer
# l_ <- "chr3:100,696,090-100,710,895" # TFG enhancer wider

l_ <- "chr17:34,347,028-34,420,319" # CCL1
l_ <- "chr17:34412487-34412887" # CCL1 enhancer
l_ <- "chr17:34412487-34413427" # CCL1 enhancers

l_ <- "chr11:75,075,603-75,209,784" # OR2AT4 / SLCO2B1 gene wider
l_ <- "chr11:75,100,000-75,109,986" # OR2AT4 gene
l_ <- "chr11:75103056-75104456" # OR2AT4 enhancer

l_ <- "chr14:24,497,494-24,526,450" # CMA1 gene
l_ <- "chr14:24,423,851-24,585,580" # CMA1 gene wider (20260314)
l_ <- "chr14:24,423,851-24,550,000" # CMA1 gene (20260326)
l_ <- "chr14:24517772-24518172" # CMA1 enhancer

l_ <- "chr15:67,039,999-67,195,802" # SMAD3 gene
l_ <- "chr15:67120991-67121491" # SMAD3 intronic enhancer

l_ <- "chr10:8,052,466-8,075,704" # GATA3 gene
l_ <- "chr10:8073007-8073407" # GATA3 intronic enhancer
```

#### Examples kidney-related (Fig6)
```{r}
l_ <- "chr10:117,166,998-118,047,656" # SLC18A2 / EMX2 gene
l_ <- "chr10:117720703-117721303" # SLC18A2 enhancer

l_ <- "chr2:113,171,623-113,282,700" # PAX8 gene
l_ <- "chr2:113213742-113214142" # PAX8 enhancer

l_ <- "chr2:54,448,250-54,680,036" # Cystatin C / SPTBN1 gene
l_ <- "chr2:54559889-54560289" # Cystatin C / SPTBN1 enhancer
```

```{r}
l_ <- "chr12:124,284,664-124,887,451" # NCOR2 gene
l_ <- "chr12:124626205-124626605" # NCOR2 enhancer

l_ <- "chr1:54,220,981-54,629,250" # ACOT11 gene
l_ <- "chr1:54488214-54488614" # ACOT11 enhancer

l_ <- "chr11:44,559,802-44,621,474" # CD82 gene
l_ <- "chr11:44570516-44570916" # CD82 enhancer

l_ <- "chr10:17,133,279-17,243,033" # TRDMT1 gene
l_ <- "chr10:17212828-17213228" # TRDMT1 enhancer

l_ <- "chr17:39,751,261-39,885,067" # IKFZ3 gene
l_ <- "chr17:39813772-39814172" # IKFZ3 enhancer
```

#### MYC example (Fig4)
```{r}
l_ <- "chr8:127,212,430-127,229,803" # PC91
l_ <- "chr8:127,218,611-127,220,250" # PC91_subset
l_ <- "chr8:127,264,015-127,266,715" # HT292
l_ <- "chr8:127,291,864-127,302,271" # PC92
l_ <- "chr8:127,595,945-127,597,132" # PC93
l_ <- "chr8:127,612,703-127,620,929" # PC94 # -0.06045953  0.26724026
l_ <- "chr8:127,618,500-127,618,700" # PC94_subset
l_ <- "chr8:127,618,483-127,618,802" # PC94_subset_2
l_ <- "chr8:127,667,949-127,670,942" # PC95 # -0.2159349  0.1049872
l_ <- "chr8:127,668,650-127,669,350" # PC95_subset
l_ <- "chr8:127,794,101-127,796,071" # PVT1
l_ <- "chr8:127,808,598-127,811,423" # A3751
l_ <- "chr8:127,850,678-127,855,822" # A3752 # -0.09139276  0.31496174
l_ <- "chr8:127,851,517-127,851,856" # A3752_subset1
l_ <- "chr8:127,854,256-127,854,835" # A3752_subset2
l_ <- "chr8:127,890,376-127,890,404" # A3753
l_ <- "chr8:127,906,653-127,909,763" # A3754
l_ <- "chr8:128,076,404-128,077,909" # A3755
l_ <- "chr8:128,175,430-128,177,526" # MDA1 # -0.11624159  0.08464334
l_ <- "chr8:128,176,675-128,177,208" # MDA1_subset
l_ <- "chr8:128,182,061-128,184,664" # MDA2
l_ <- "chr8:128,324,390-128,324,940" # A3756
l_ <- "chr8:128,551,556-128,560,975" # A5493 # -0.9652244  0.2906040
l_ <- "chr8:128,555,167-128,555,466" # A5493_subset1
l_ <- "chr8:128,556,690-128,557,011" # A5493_subset2
l_ <- "chr8:128,653,951-128,654,391" # HepG21
l_ <- "chr8:129,707,012-129,713,257" # HCT1161 # -0.1571948  0.6529629
l_ <- "chr8:129,707,566-129,708,269" # HCT1161_subset

l_ <- "chr8:127,892,754-127,907,754" # e1
l_ <- "chr8:127,952,754-127,972,754" # e2
l_ <- "chr8:128,042,754-128,052,754" # e3-e4
l_ <- "chr8:129,577,754-129,592,754" # e5
l_ <- "chr8:129,682,754-129,697,754" # e6-e7 #  -2.1276451  0.3113627
l_ <- "chr8:129,690,869-129,691,415" # e6
l_ <- "chr8:129,692,149-129,693,315" # e7

l_ <- "chr8:127,727,331-129,697,754" # MYC gene wider 1
l_ <- "chr8:127,173,354-129,805,804" # MYC gene wider 2
l_ <- "chr8:127,670,067-129,805,029" # MYC PVT1 GSDMC genes
```

#### ProbeSet K562_enhancers
#### see [1.4.] in '10.1.K562_external_validation_copy.Rmd'
```{r}
l_ <- "chr12:34335869-34336269" # intergenic (LINC02963)
l_ <- "chr4:702557-702957" # intergenic (PCGF3)
l_ <- "chr9:135689838-135690238" # intergenic (SOLH1)
# l_ <- "chr9:135689838-135690260" # intergenic (SOLH1)
l_ <- "chr10:77836260-77836660" # intronic (DLG5)
# l_ <- "chr10:77836250-77836660" # intronic (DLG5)
l_ <- "chr9:136511891-136512291" # intronic (NOTCH1) ?
l_ <- "chr4:128471633-128472033" # intronic (LINC02615)
l_ <- "chr11:65493822-65494222" # intergenic (MALAT1)

l_ <- "chr1:36386416-36386816" # exon (LOC124904012)
l_ <- "chr7:76166443-76166843" # intergenic (LOC124901674)
# l_ <- "chr7:76166443-76166860" # intergenic (LOC124901674)
l_ <- "chr20:37379147-37379547" # intronic (SRC)
l_ <- "chr5:119321597-119321997" # intronic (TNFAIP8)
l_ <- "chr15:70099475-70099875" # intergenic (TLE3)
# l_ <- "chr15:70099430-70099930" # intergenic (TLE3)

l_ <- "chr2:142930954-142931354" # intronic (KYNU)
l_ <- "chr11:9364346-9364746" # intergenic (IPO7)
l_ <- "chr5:141849550-141849950" # intergenic (PCDH1)
l_ <- "chr6:53359250-53359650" # intergenic (ELOVL5)
# l_ <- "chr6:53359250-53359665" # intergenic (ELOVL5)
```

#### ProbeSet K562_nonenhancers
#### see [1.2.2.] in '10.3.Cell_line_vs_trait_enrichment_copy.Rmd'
```{r}
l_ <- "chr10:72319524-72320183" # intergenic (DNAJB12)
l_ <- "chr6:142926574-142927393" # intergenic (HIVEP2) ?
l_ <- "chr6:44001262-44001770" # intronic (SCIRT) ?
l_ <- "chr2:84927005-84927424" # intergenic (TMSB10) ?
l_ <- "chr2:231680936-231681515" # intergenic (PTMA)
l_ <- "chr18:8607550-8608249" # intergenic (RAB12) # excluded
l_ <- "chr1:2578073-2578772" # intronic (LOC100996583)
l_ <- "chr16:509187-509745" # intronic (RAB11FIP3)
```

## [2.1.] Get genomic region from coordinates
```{r}
l <- stringr::str_replace_all(l_,",","")
chr = stringr::str_replace(l,"^(chr.*):.*","\\1")
s = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\1") %>%
  base::as.integer()
e = stringr::str_replace(l,"^chr.*:(.*)-(.*)","\\2") %>%
  base::as.integer()
out <- base::paste0("_",chr,"_",s,"_",e)
```

```{r}
extend = 0 # 200 # 500
ran <- GenomicRanges::GRanges(
    seqnames = chr,
    ranges = IRanges::IRanges(start = s, end = e)
  )
BiocGenerics::start(ran) <- BiocGenerics::start(ran) - extend
BiocGenerics::end(ran) <- BiocGenerics::end(ran) + extend
GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
```

## [2.2.] Extract coordinates
```{r}
chrom <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)
```

## [2.3.] Get genomic region from gene names
#### examples are SLC39A4, YLPM1, SMNDC1
```{r}
# omr <- base::c(
#     "AREL1","FCF1","YLPM1","XR_944033.2",
#     "PROX2"
#   )
# ran_ <- hg38_gr_sub[
#     S4Vectors::elementMetadata(hg38_gr_sub)[,"gene_name"] %in% omr,
#   ]
# base::unique(ran_$gene_name)
# ran <- base::range(ran_,ignore.strand = TRUE)
# 
# GenomeInfoDb::seqlevelsStyle(ran) <- "UCSC"
# if(!stringr::str_detect(GenomeInfoDb::seqlevels(ran)[1],"chr")){
#   GenomeInfoDb::seqlevels(ran) <- base::gsub(
#       pattern = '^(.*)$',
#       replacement = 'chr\\1',
#       x = GenomeInfoDb::seqlevels(ran)
#     )
#   }
# GenomeInfoDb::seqlevels(ran)
```

## [2.4.] Get genomic regions from GRanges input
```{r}
# gr <- crispr %>%
#   dplyr::filter(
#       measuredGeneSymbol == "MYC" &
#       Significant == TRUE
#     ) %>%
#   dplyr::mutate(
#       summit = chromStart + base::round((chromEnd - chromStart +1)/2)
#     ) %>% 
#   GenomicRanges::GRanges()
# extend = 1000
# BiocGenerics::start(gr) <- BiocGenerics::start(gr) - extend
# BiocGenerics::end(gr) <- BiocGenerics::end(gr) + extend
# GenomeInfoDb::seqlevelsStyle(gr) <- "UCSC"
# myce38 <- IRanges::reduce(gr, ignore.strand = TRUE, with.revmap = TRUE)
# S4Vectors::mcols(myce38)$smts <- S4Vectors::aggregate(gr, S4Vectors::mcols(myce38)$revmap, smts = summit)$smts
# i = 1L
# ran <- myce38[i]
```

## [3.] CAGE preparation based on bigwigs --------------------

## [3.1.] Read in tpm scores in selected ranges
#### see [2.0.] in '20200721_workflow_DrosophilaCAGE.Rmd'
#### Evaluate with 'dplyr::distinct(cage_sel_,sample)'
```{r}
cage_sel_ <- PRIME::readRange(fs,ran)
```

## [3.2.] Summarise data
```{r}
cage_sel <- cage_sel_ %>%
  dplyr::mutate(
    replicate = stringr::str_replace(
        sample, ".*_(.*?)\\.TPM$", "\\1"
      ),
    type = stringr::str_replace(
        sample, "^(.*?)_(.*)_(.*)\\.TPM$", "\\2"
      ),
    )
```

## [3.3.] Adjust source order
#### dplyr::distinct(cage_sel,source)
```{r}
levs <- base::c(
    "C", "C1", "C2", "C3",
    "N", "N1", "N2", "N3"
  )
cage_sel <- cage_sel %>%
  dplyr::mutate(
      replicate = base::factor(
          x = replicate,
          levels = levs,
          ordered = TRUE
        )
    ) %>%
  dplyr::arrange(
      type,replicate,seqnames,start,end
    )
```

## [4.] CAGE preparation based on CTSSs --------------------

## [4.0.1.] Subset & convert scores from RangedSummarizedExperiment to tibble
#### Based on objects read in in [1.8.]
#### Evaluate samples "base::rownames(SummarizedExperiment::colData(GM12878.CTSSs.sub))"
```{r}
# object <- GM12878.CTSSs.sub # GM12878.CTSSs.sub # K562.CTSSs.sub
# ovl <- IRanges::findOverlaps(
#     query = ran,
#     subject = object
#   )
# gr <- SummarizedExperiment::rowRanges(object)[ovl@to,]
# 
# crds <- gr %>%
#   base::as.data.frame() %>%
#   tibble::as_tibble() %>%
#   dplyr::mutate(
#       start = pos,
#       end = pos
#     ) %>%
#   dplyr::select(
#       seqnames,start,end,strand
#     )
# 
# scrs <- SummarizedExperiment::assay(object,"TPM")[ovl@to,] %>%
#   base::as.matrix() %>%
#   base::as.data.frame()
# 
# cage_sel_ <- dplyr::bind_cols(crds,scrs)
```

## [4.0.2.] Subset & convert scores from RangedSummarizedExperiment to tibble
#### Based on objects read in in [1.5.] - [1.7.]
#### Evaluate with 'c = "HepG2"'
#### Alternative for combining data: base::do.call(what = "rbind",args = base::lapply(X = base::names(ovl),FUN = function(c){}))'
```{r}
# ovl <- purrr::map(
#   .x = base::names(CTSSs.objects),
#   .f = function(c){
#       ovl <- IRanges::findOverlaps(
#           query = ran,
#           subject = CTSSs.objects[[c]]
#         )
#       gr <- SummarizedExperiment::rowRanges(CTSSs.objects[[c]])[ovl@to,]
#       return(methods::as(gr,"GRanges"))
#     }
#   )
# base::names(ovl) <- base::names(CTSSs.objects)
# 
# crds <- purrr::map_dfr(
#   .x = base::names(ovl),
#   .f = function(c){
#     ovl[[c]] %>% 
#       base::as.data.frame() %>%
#         tibble::as_tibble() %>%
#         dplyr::mutate(
#             celltype = c
#           ) %>%
#         dplyr::select(
#             celltype,seqnames,start,end,strand
#           )
#         }
#       )
```

```{r}
ovl <- purrr::map(
  .x = base::names(CTSSs.objects),
  .f = function(c){
      return(
        IRanges::findOverlaps(
          query = ran,
          subject = CTSSs.objects[[c]]
        )
      )
    }
  )
base::names(ovl) <- base::names(CTSSs.objects)

crds <- purrr::map_dfr(
  .x = base::names(ovl),
  .f = function(c){
    gr <- SummarizedExperiment::rowRanges(CTSSs.objects[[c]])[ovl[[c]]@to,]
    methods::as(gr,"GRanges") %>%
      base::as.data.frame() %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = c,
            rowID = ovl[[c]]@to
          ) %>%
        dplyr::select(
            type,seqnames,start,
            end,strand,rowID
          )
        }
      )
```

#### Evaluate with 'c = "HCT116"', 'c = "A549"','c = "K562"', 'c = "HepG2"'
```{r}
scrs <-  purrr::map_dfr(
  .x = base::names(ovl),
  .f = function(c){
    
    mat <- CTSSs.objects[[c]]@assays@data$TPM %>%
      base::as.matrix()
    
    # base::colnames(mat) <- stringr::str_extract(base::colnames(mat), "C|N")
    # base::colnames(mat) <- stringr::str_replace(base::colnames(mat), "^(.*?)_(.*?)$", "\\2")
    base::colnames(mat) <- stringr::str_match(base::colnames(mat), "N_10M|C|N") %>%
      base::as.character()
      
    return(
      mat %>%
        tibble::as_tibble() %>%
        dplyr::filter(
            dplyr::row_number() %in% ovl[[c]]@to
          ) %>%
        dplyr::mutate(
          type = c,
          rowID = ovl[[c]]@to
        )
      )
    }
  )

cage_sel_ <- dplyr::left_join(
      crds, scrs, by = base::c("type","rowID")
    ) %>%
  # dplyr::mutate(
  #     dplyr::across(
  #       C:N, ~ base::replace(., base::is.na(.), 0)
  #     )
  #   ) %>%
  dplyr::select(-rowID)
```

## [4.1.1.] Summarise data
#### Used with e.g., GM12878.CTSSs.sub, K562.CTSSs.sub, see [1.8.] & [4.0.1]
#### Evaluate with 'dplyr::arrange(cage_sel,dplyr::desc(score))'
#### Use 'TRUE ~ (20)*score' for GATA1 enhancer example
```{r}
# regex_f = "(?<=C|N)\\d$" # ".*(.)$"
# regex_s = "^(.*?)_(.*?)(\\d)?$" # "^(.*?)_(.*)(_)?(.)"
# 
# cage_sel <- cage_sel_ %>%
#   tidyr::pivot_longer(
#       cols = tidyselect::contains("_"),
#       names_to = "sample",
#       values_to = "score"
#     ) %>%
#   dplyr::mutate(
#     replicate = stringr::str_extract(
#         sample, regex_f
#       ),
#     replicate = dplyr::case_when(
#         base::is.na(replicate) ~ "pool",
#         TRUE ~ replicate
#       ),
#     type = stringr::str_replace(
#           sample, regex_s, "\\2"
#         ) %>%
#       stringr::str_replace(
#         .,"^(.*?)[_]$","\\1"
#       )
#     ) %>%
#   dplyr::mutate(
#     score = dplyr::case_when(
#         strand == "-" ~ (-1)*score,
#         TRUE ~ score
#       ),
#     strand = dplyr::case_when(
#         strand == "+" ~ "plus",
#         strand == "-" ~ "minus"
#       )
#     )
```

## [4.1.2.] Summarise data
#### Used with e.g., GM12878.CTSSs.sub, K562.CTSSs.sub, see [1.5.]-[1.7.] & [4.0.2]
#### Evaluate with 'dplyr::distinct(cage_sel,sample)'
```{r}
cage_sel <- cage_sel_ %>%
  tidyr::pivot_longer(
      cols = base::c("C","N", "N_10M"),
      names_to = "sample",
      values_to = "score",
      values_drop_na = TRUE
    ) %>%
  dplyr::mutate(
    score = dplyr::case_when(
        strand == "-" ~ (-1)*score,
        TRUE ~ score
      ),
    strand = dplyr::case_when(
        strand == "+" ~ "plus",
        strand == "-" ~ "minus"
      )
    )
```

## [4.2.] Add mock lines for missing cell types
#### Use with 'CTSSs.objects'
#### Evaluate with "ins <- base::names(CTSSs.objects)"
```{r}
ins <- cage_sel %>%
  dplyr::distinct(type) %>%
  dplyr::pull(type)

ous <- base::names(CTSSs.objects)[! base::names(CTSSs.objects) %in% ins]

insert <- tibble::tibble(
    type = base::sort(base::rep(ous,2)),
    seqnames = base::rep(chrom,2*base::length(ous)),
    start = base::rep(start_pos,2*base::length(ous)),
    end = base::rep(end_pos,2*base::length(ous)),
    strand = base::rep("plus",2*base::length(ous)),
    sample = base::rep(base::c("C","N"),base::length(ous)),
    score = base::rep(base::as.double(0),2*base::length(ous))
  )

cage_sel <- dplyr::bind_rows(cage_sel,insert)
```

#### Evaluate with 'ins <- base::c("dendritic cell", "mast cell", "monocyte", "reticulocyte")'
#### 'dplyr::distinct(cage_sel,sample)'
#### For individual samples 'insert <- tibble::tibble(sample = "reticulocyte",seqnames = chrom,start = start_pos,end = end_pos,strand = "plus",score = base::as.double(0))'
```{r}
# cage_sel <- cage_sel_
# ins <- cage_sel %>%
#   dplyr::distinct(sample) %>%
#   dplyr::pull(sample)
# 
# ous <- facets_sel[!facets_sel %in% ins]
# # ous <- facets_sel_[!facets_sel_ %in% ins]
# 
# insert <- tibble::tibble(
#     sample = ous,
#     seqnames = chrom,
#     start = start_pos,
#     end = end_pos,
#     strand = "plus",
#     score = base::as.double(0)
#   )
# 
# cage_sel <- dplyr::bind_rows(cage_sel,insert)
```

# [5.] Plot data -------------------------

## [5.0.] Plot genome tracks
```{r}
plot <- PRIME::plotTracks(cage_sel, ran) +
  ggplot2::facet_wrap(
      facets = . ~ type + replicate,
      ncol = 1,
      # scales = "free_y",
      strip.position = "right"
    ) +
  coord_cartesian(
      ylim = base::c(-0.1,0.1) # -250,250 # -3,3
    ) # +
  # ggh4x::facetted_pos_scales(
  #   y = base::list(
  #     type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ scale_y_continuous(
  #         limits = base::c(-3,3)
  #       ),
  #     type %in% base::c("500K_N","5M_N","500K_C") ~ scale_y_continuous(
  #         limits = base::c(0,300)
  #       )
  #     )
  #   )
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

#### Plot GATA1 enhancer: "chrX:48,787,857-48,788,321"
```{r}
# plot <- cage_sel %>%
#   dplyr::mutate(
#     type = base::factor(
#         x = type,
#         levels = base::c("C","N","N_10M"),
#         ordered = TRUE
#       ),
#     score = dplyr::case_when(
#         strand == "plus" ~ score*10,
#         TRUE ~ score
#       )
#     ) %>%
#   PRIME::plotTracks(., ran) +
#     ggplot2::facet_wrap(
#         facets = . ~ type,
#         ncol = 1,
#         strip.position = "right"
#         # scales = "free_y"
#       ) +
#     ggh4x::facetted_pos_scales(
#       y = base::list(
#         type %in% base::c("C","N","N_10M") ~ scale_y_continuous(
#             limits = base::c(-7.5,7.5),
#             breaks = base::c(-5,0,5),
#             labels = base::c(-5,0,0.5)
#           )
#         )
#       )
# ggrastr::rasterise(
#     input = plot,
#     layers = "Bar",
#     dpi = 300
#   )
```

## Print data
```{r eval=FALSE}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("trackHubFiles/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        # "_GATA1", "_RAPGEF2"
        # "_TMSB4X", "_ZMYND8", "_ARL4A",
        "_RAPGEF2",
        # "_K562CTSSPool", "_Gm12878CTSSPool",
        # "_enhancer", "_gene", "_promoter", _genewider, "_PROMPT"
        "_PROMPT_v2",
        out,
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    device = "pdf",
    dpi = 300,
    width = 2, # 2.5 # 8
    height = 5
  )
```

#### Plot MYC example (20260408)
```{r}
# plot <- PRIME::plotTracks(cage_sel, ran) +
#   ggplot2::facet_wrap(
#       facets = . ~ type,
#       ncol = 1,
#       # scales = "free_y",
#       strip.position = "right"
#     ) # +
#   # ggh4x::facetted_pos_scales(
#   #   y = base::list(
#   #     type %in% base::c("C","N","N_10M") ~ scale_y_continuous(
#   #         limits = base::c(-5,5)
#   #       )
#   #     )
#   #   )
# ggrastr::rasterise(
#     input = plot,
#     layers = "Bar",
#     dpi = 300
#   )
```

#### Plot MYC example (20260501)
```{r}
cage_sel <- cage_sel %>%
  dplyr::filter(!(type == "HCT116" & sample == "N"))

plot <- PRIME::plotTracks(cage_sel, ran) +
  ggplot2::facet_wrap(
      facets = . ~ type + sample,
      ncol = 1,
      # scales = "free_y",
      strip.position = "right"
    ) +
  coord_cartesian(
      ylim = base::c(-0.25,0.25)
    ) +
  scale_y_continuous(
      labels = base::c(-0.2,0,0.2),
      breaks = base::c(-0.2,0,0.2)
    )
  # theme(
  #     aspect.ratio = 1
  #   ) # +
  # geom_rect(
  #     inherit.aes = FALSE,
  #     data = tibble::tibble(
  #         start = base::c(129690933, 129692279),
  #         end = base::c(129691317, 129693646)
  #       ),
  #     mapping = aes(
  #         xmin = start,
  #         xmax = end
  #       ),
  #     ymin = -2,
  #     ymax = 0.5,
  #     fill = "grey",
  #     alpha = 0.5
  #   )
  # geom_rect(
  #     inherit.aes = FALSE,
  #     data = mycs,
  #     mapping = aes(
  #         xmin = start,
  #         xmax = end
  #       ),
  #     ymin = -5,
  #     ymax = 5,
  #     fill = "grey",
  #     alpha = 0.5
  #   )
  # ggh4x::facetted_pos_scales(
  #   y = base::list(
  #     type %in% base::c("C","N","N_10M") ~ scale_y_continuous(
  #         limits = base::c(-5,5)
  #       )
  #     )
  #   )
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

#### Plot ProbeSet enhancers (20260505)
```{r}
# cage_sel <- cage_sel %>%
#   dplyr::filter(!(type == "HCT116" & sample == "N"))
# 
# probes <- tibble::tibble(
#     start = base::c(509242, 509408),
#     end = base::c(509391, 509557),
#     sample = base::c("N_10M","N_10M"),
#     type = base::c("K562","K562")
#   )
```

```{r}
probes <- enhs %>%
  dplyr::select(-comment) %>% 
  drop_na() %>% 
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "position",
      values_to = "coordinate"
    ) %>%
  tidyr::separate(
      col = "position",
      into = base::c("strand","position"),
      sep = "_"
    ) %>%
  dplyr::select(
      "loc","strand","position","coordinate"
    ) %>% 
  tidyr::pivot_wider(
      names_from = "position",
      values_from = "coordinate"
    ) %>%
  dplyr::mutate(
      sample ="N_10M",
      type = "K562"
    ) %>%
  dplyr::filter(
      loc == l_
    )

probes
```

```{r}
cage_sel <- cage_sel %>%
  dplyr::filter(!(type == "HCT116" & sample == "N"))

plot <- PRIME::plotTracks(cage_sel, ran) +
  ggplot2::facet_wrap(
      facets = . ~ type + sample,
      ncol = 1,
      # scales = "free_y",
      strip.position = "right"
    ) +
  coord_cartesian(
      ylim = base::c(-0.25,0.25)
    ) +
  scale_y_continuous(
      labels = base::c(-0.2,0,0.2),
      breaks = base::c(-0.2,0,0.2)
    ) +
  geom_rect(
      inherit.aes = FALSE,
      data = probes,
      mapping = aes(
          xmin = start,
          xmax = end
        ),
      ymin = -0.2,
      ymax = -0.1,
      fill = "grey",
      alpha = 0.5
    )

ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

## Print data
```{r eval=FALSE}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("trackHubFiles/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        # "_GATA1", "_MYC"
        "_probeSites",
        # "_K562CTSSPool", "_AllCellCTSSsPool"
        "_AllCellCTSSsPool",
        # "_enhancer", "_gene", "_e1", "_e2", "_e3_e4", "_e5", "_e6_e7",
        # "_PC91", "_HT292", "_PC92", "_PC93", "_PC94", "_PC95", "_PVT1",
        # "_A3751", "_A3752", "_A3753", "_A3754", "_A3755", "_MDA1", "_MDA2",
        # "_A5493", "_HepG21", "_HCT1161"
        out,
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    device = "pdf",
    dpi = 300,
    width = 2.5, # 2.5 # 8
    height = 6
  )
```

```{r}
probes_seqs <- enhs %>%
  dplyr::select(-comment) %>% 
  drop_na() %>% 
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "position",
      values_to = "coordinate"
    ) %>%
  tidyr::separate(
      col = "position",
      into = base::c("strand","position"),
      sep = "_"
    ) %>% 
  dplyr::select(
      "sample","loc","position","coordinate",
      "type","reference","seqname","strand"
    ) %>% 
  tidyr::pivot_wider(
      names_from = "position",
      values_from = "coordinate"
    ) %>%
  dplyr::rename(
      names = "seqname",
      enhancer_coordinates = "loc",
      enhancer_type = "type"
    ) %>%
  dplyr::mutate(
    strand = dplyr::case_when(
        strand == "minus" ~ "-",
        strand == "plus" ~ "+",
        TRUE ~ strand
      ),
    probe_coordinates = base::paste0(names,":",start,"-",end),
    probe_seq = purrr::pmap(
        .l = base::list(
            names, start, end, strand
          ),
       .f = Biostrings::getSeq,
        x = Hsapiens,
        width = NA,
        as.character = TRUE
      ),
    probe_seq = base::unlist(probe_seq)
  )
probes_seqs
```

#### Evaluate
```{r}
probes_seqs %>%
  dplyr::filter(row_number() == 46) %>%
  dplyr::pull(probe_seq)
```

#### Write out
```{r}
utils::write.table(
    x = probes_seqs,
    file = base::paste0(
        here::here("0.External_resources/"),
        base::format(base::Sys.Date(), "%Y%m%d"),
        "_enhancers_K562_v1",
        ".tsv"
      ),
    quote = FALSE,
    sep = "\t",
    eol = "\n",
    row.names = FALSE,
    col.names = TRUE
  )
```


```{r}
# plot <- cage_sel %>%
#   dplyr::filter(
#       sample != "GM12878_10M_Cap_align"
#     ) %>% 
#   dplyr::mutate(
#     type = base::factor(
#         x = type,
#         levels = base::c("500K_C","500K_N","5M_N","1M_trip","1M_ctrl","groCap","10M_Cap_align"),
#         ordered = TRUE
#       ),
#     score = dplyr::case_when(
#         strand == "minus" & type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ score,
#         strand == "minus" & type %in% base::c("500K_N","5M_N","500K_C") ~ score*200,
#         TRUE ~ score
#       )
#     ) %>%
#   PRIME::plotTracks(., ran) +
#     ggplot2::facet_wrap(
#         facets = . ~ type,
#         ncol = 1,
#         strip.position = "right",
#         scales = "free_y"
#       ) +
#     ggh4x::facetted_pos_scales(
#       y = base::list(
#         type %in% base::c("10M_Cap_align","1M_trip","1M_ctrl","groCap") ~ scale_y_continuous(
#             limits = base::c(-3,3)
#           ),
#         type %in% base::c("500K_N","5M_N","500K_C") ~ scale_y_continuous(
#             limits = base::c(-300,300),
#             breaks = base::c(-200,0,200),
#             labels = base::c(-1,0,200)
#           )
#         )
#       )
# ggrastr::rasterise(
#     input = plot,
#     layers = "Bar",
#     dpi = 300
#   )
```

## [5.1.] Print data
```{r eval=FALSE}
# for(ext in c("pdf", "png")){
#     ggplot2::ggsave(
#       filename = base::paste0(
#           here::here("trackHubFiles/"),
#           format(Sys.time(), "%Y%m%d"),
#           out,
#           # "_enh2",
#           "_500k5M",
#           "_TmmTpm",
#           ".", ext
#         ),
#       plot = ggplot2::last_plot(),
#       device = ext,
#       dpi = 300,
#       width = 4,
#       height = 8
#     )
#   }
```

## [5.2.] E.g., APOC1 example (ABO)
#### see [4.10.] in "10.3.Cell_line_vs_trait_enrichment.Rmd"
```{r}
sites_ <- base::readRDS(
  file = base::list.files(
        path = here::here("10.External_validation"),
        full.names = TRUE
      ) %>%
    stringr::str_subset("ApoB.rds")
    # stringr::str_subset("MCHC.rds")
  )
```

## E.g., Filter out ApoB-related genetic variants in the vicinity of APOC1 gene (ABO)
#### see also [2.3.] for 'start_pos' and 'end_pos' objects
```{r warning=FALSE}
chr <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

sites <- sites_ %>% 
  tidyr::pivot_wider( # 218
      names_from = "cell.line",
      values_from = "score"
    ) %>%
  # dplyr::filter(
  #     stringr::str_detect(variant, "chr19:44")
  #   ) %>%
  tidyr::separate(
      col = variant,
      sep = ":",
      into = base::c("chrom","start","end","gv"),
      convert = TRUE
    ) %>%
  dplyr::filter(
      chrom == chr & start > start_pos & end < end_pos,
    ) %>%
  dplyr::mutate(
      end = start
    ) %>% 
  # dplyr::select(chrom, start, gv) %>% 
  GenomicRanges::makeGRangesFromDataFrame()
  # dplyr::mutate(
  #     start = base::as.double(start) - extend,
  #     end = base::as.double(end) + extend
  #   ) %>% 
  # dplyr::select(start, end)
```

#### see [1.0.-1.4.], especially [1.2.1.] in "10.3.Cell_line_vs_trait_enrichment.Rmd" for 'unfiltered.PRIME'-object
#### Alternative: Use IRanges::reduce() on olap from unfiltered.PRIME and filter for the highest PRIME score
#### evaluate with 'n = "HepG2_N"' # 61,589

#### Pooled PRIME data
```{r}
sites_scores_ <- purrr::map_dfr(
  .x = base::names(pooled.PRIME),
  .f = function(n){
    pd <- pooled.PRIME[[n]]
    olap <- IRanges::findOverlaps(
        query = sites,
        subject = pd,
        minoverlap = 1
      )
    return(
      pd[S4Vectors::subjectHits(olap),] %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = stringr::str_replace(n,"(.*?)_.*","\\1"),
            sample = stringr::str_replace(n,"(.*?)_(.*)","\\2")
          )
      )
    }
  )
```

#### Unfiltered PRIME data
```{r}
sites_scores_ <- purrr::map_dfr(
  .x = base::names(unfiltered.PRIME),
  .f = function(n){
    pd <- unfiltered.PRIME[[n]]
    olap <- IRanges::findOverlaps(
        query = sites,
        subject = pd,
        minoverlap = 1
      )
    
    gr <- pd[S4Vectors::subjectHits(olap),]
    out <- IRanges::reduce(x = gr, with.revmap = TRUE)
    S4Vectors::mcols(out)$score <- base::do.call(
          what = rbind,
          args = base::lapply(
            X = S4Vectors::mcols(out)$revmap,
            FUN = function(i) {
              base::max(S4Vectors::mcols(gr)$score[i])
            }
          )
        )
    
    return(
      out %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = stringr::str_replace(n,"(.*?)_.*","\\1"),
            sample = stringr::str_replace(n,"(.*?)_(.*)","\\2")
          )
      )
    }
  )
```

#### Cut sites to plotted range
```{r}
sites_scores <- sites_scores_ %>%
  dplyr::mutate(
    start = dplyr::case_when(
        start <= start_pos ~ start_pos,
        .default = start
      ),
    end = dplyr::case_when(
        end >= end_pos ~ end_pos,
        .default = end
      )
    )
```

#### Plot data
```{r}
# base::range(cage_sel$score)
minY = base::range(cage_sel$score)[1] # -2.1 # -0.5 # -0.15
maxY = base::range(cage_sel$score)[2] # 2.1 # 0.5 # 0.15
coef = 0.75
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

plot_ <- PRIME::plotTracks(cage_sel, ran) +
  geom_rect(
      inherit.aes = FALSE,
      data = sites_scores,
      mapping = aes(
          xmin = start,
          xmax = end,
          fill = score
        ),
      ymin = minY,
      ymax = maxY,
      # fill = "grey",
      alpha = 0.5
    ) +
  facet_wrap(
      facets = . ~ type + sample,
      ncol = 1,
      strip.position = "right",
      scales = "fixed" #"fixed" "free_y"
    ) +
  scale_fill_gradientn(
      limits = base::c(0,1),
      colours = cols
    ) +
  ylim(minY,maxY)  +
  guides(
      fill = guide_colourbar(
        label.position = "right",
        title = NULL,
        title.position = "left",
        title.theme = element_text(
            angle = 90,
            hjust = 0.5
          ),
        ticks=T,
        nbin=100,
        barheight=10,
        label=T,
        barwidth=0.5
      )
    ) # +
  # theme(
  #     aspect.ratio = 0.8
  #   )
  
plot <- ggrastr::rasterise(
    input = plot_,
    layers = "Bar",
    dpi = 300
  )
plot
```

## Print data
```{r eval=FALSE}
for(ext in c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("trackHubFiles/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          # "_ABOprimelocus",
          # "_LPLprimelocus",
          # "_LIPCprimeloci",
          "_ABCA7primelocus",
          # "_TMPRSS6primelocus",
          "_unfilteredPRIMEloci",
          out,
          ".", ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 4
    )
  }
```

## [5.3.] Plot c-KIT & KLF6 (Fig6)
```{r}
varPL <- base::readRDS(
  file  = base::list.files(
        path = here::here("11.LDSC_FANTOM5_copy"),
        full.names = TRUE
      ) %>%
    stringr::str_subset("rds") %>%
    stringr::str_subset("varPL") %>%
    stringr::str_subset("_AllImmuneCells")
  )
```

```{r}
icPL <- base::readRDS(
  file  = base::list.files(
        path = here::here("11.LDSC_FANTOM5_copy"),
        full.names = TRUE
      ) %>%
    stringr::str_subset("rds") %>%
    stringr::str_subset("20260305_icPL_AllImmuneCells") # 2,056,763
    # stringr::str_subset("20260305_icPL_KidneyRelatedCells") # 2,772,597
    # stringr::str_subset("20260311_icPL_AllImuneNeutrophilFABM6") # 2,258,436
    # stringr::str_subset("20260311_icPL_ACOT11") # 3,762,040
    # stringr::str_subset("20260311_icPL_NCOR2") # 2,381,852
    # stringr::str_subset("20260311_icPL_TRDMT1") # 3,493,705
    # stringr::str_subset("20260312_icPL_IKFZ3") # 2,243,878
  )
```

#### Evaluate
```{r}
# chrom = "chr15"
# start_pos = 67121091
# end_pos = 67121491
# 
# icPL %>%
#   dplyr::filter(
#       chr == chrom & start >= start_pos & end <= end_pos
#     )
```

#### Evaluate with 'dplyr::filter(varPL, Gene_symbol == "CMA1")'
#### 'dplyr::distinct(seqnames,start,end,PL,.keep_all = TRUE)'
```{r}
chrom <- GenomeInfoDb::seqnames(ran) %>%
  base::as.character()
start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

sites_scores <- icPL %>% # varPL # icPL
  dplyr::filter(
      chr == chrom & start >= start_pos & end <= end_pos &
      PL >= 0.75
      # base::abs(Z) > 2
    ) %>%
  # dplyr::mutate(
  #     cell = case_when(
  #       cell == "acute.myeloid.leukemia.FAB.M6" ~ "acute_myeloid_leukemia_(FAB_M6)",
  #       cell == "B.lymphoblastoid" ~ "B_lymphoblastoid",
  #       cell == "chronic.myelogenous.leukemia" ~ "chronic_myelogenous_leukemia",
  #       cell == "epitheloid.carcinoma" ~ "epitheloid_carcinoma",
  #       cell == "Hep.2" ~ "Hep-2",
  #       cell == "kidney.epithelial.cell" ~ "kidney epithelial cell",
  #       cell == "mast.cell" ~ "mast cell",
  #       cell == "mesenchymal.cell" ~ "mesenchymal cell",
  #       cell == "natural.killer.cell" ~ "natural killer cell",
  #       cell == "skin.fibroblast" ~ "skin fibroblast",
  #       cell == "small.cell.lung.carcinoma" ~ "small_cell_lung_carcinoma",
  #       cell == "vascular.associated.smooth.muscle.cell" ~ "vascular associated smooth muscle cell",
  #       cell == "cervical.cancer" ~ "cervical_cancer",
  #       cell == "Hep.2" ~ "Hep-2",
  #       TRUE ~ cell
  #     )
  #   ) %>%
  dplyr::mutate(
      sample = stringr::str_replace_all(cell, "\\.", " ")
    )
```

#### cage_sel <- cage_sel_
#### facets_sel <- base::c("dendritic cell", "mast cell", "monocyte", "reticulocyte", "T cell", "neutrophil", "acute_myeloid_leukemia_(FAB_M6)")
```{r}
minY = -20 # -50 # -500 # base::range(cage_sel_$score)[1]
maxY = 20 # 50 # 500 # base::range(cage_sel_$score)[2]

coef = 0.75
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

plot <- cage_sel %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = facets_sel, # facets_sel_ # facets_sel
        ordered = TRUE
      )
    ) %>%
  PRIME::plotTracks(., ran) +
  geom_rect(
      inherit.aes = FALSE,
      data = sites_scores,
      mapping = aes(
          xmin = start,
          xmax = end,
          fill = PL
        ),
      ymin = minY,
      ymax = maxY,
      # fill = "grey",
      alpha = 0.5
    ) +
  # facet_wrap(
  #     facets = . ~ sample,
  #     ncol = 1,
  #     strip.position = "right",
  #     scales = "fixed" #"fixed" "free_y"
  #   ) +
  scale_y_continuous(
      limits = base::c(minY,maxY),
      breaks = base::seq(minY,maxY,(maxY-minY)/4)
    ) +
  scale_fill_gradientn(
      limits = base::c(0.0,1.0),
      colours = cols
    )
   
ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

## Print data
```{r eval=FALSE}
ext  = "pdf" # "png"
ggplot2::ggsave(
    filename = base::paste0(
        here::here("trackHubFiles/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        # "_KLF6", "_cKit", "_CCL1", "_OR2AT4", "_CMA1", "_SMAD3", "_GATA3",
        # "_SLC18A2", "_PAX8", "_SPTBN1",
        # "_NCOR2", "_ACOT11", "_CD82", "_TRDMT1", "_IKFZ3",
        "_CMA1",
        # "_gene", "_enhancer", "_geneWider", "_enhancerMore",
        "_gene",
        out,
        # "_v1", "_v2",
        "_v3",
        ".", ext
      ),
    plot = ggplot2::last_plot(),
    device = ext,
    dpi = 300,
    width = 8, # 8, # 2.5,
    height = 4 # 4 # 8
  )
```

## [5.4.] Plot MYC example (20260506)
#### see [1.0.-1.4.], especially [1.2.1.] in "10.3.Cell_line_vs_trait_enrichment.Rmd" for 'unfiltered.PRIME'-object
#### Alternative: Use IRanges::reduce() on olap from unfiltered.PRIME and filter for the highest PRIME score
#### evaluate with 'n = "HepG2_N"' # 61,589

#### Pooled PRIME data
```{r}
sites_scores_ <- purrr::map_dfr(
  .x = base::names(pooled.PRIME),
  .f = function(n){
    pd <- pooled.PRIME[[n]]
    olap <- IRanges::findOverlaps(
        query = ran,
        subject = pd,
        minoverlap = 1
      )
    return(
      pd[S4Vectors::subjectHits(olap),] %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
            type = stringr::str_replace(n,"(.*?)_.*","\\1"),
            sample = stringr::str_replace(n,"(.*?)_(.*)","\\2")
          )
      )
    }
  )
sites_scores_
```

#### Cut sites to plotted range
```{r}
sites_scores <- sites_scores_ %>%
  dplyr::mutate(
    start = dplyr::case_when(
        start <= start_pos ~ start_pos,
        .default = start
      ),
    end = dplyr::case_when(
        end >= end_pos ~ end_pos,
        .default = end
      )
    )
sites_scores
```

```{r}
minY = -0.5  # -0.3 # -5 # base::range(cage_sel$score)[1]
maxY = 0.5  # 0.3 # 5 # base::range(cage_sel$score)[2]

coef = 0.75
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

cage_sel <- cage_sel %>%
  dplyr::filter(!(type == "HCT116" & sample == "N"))

plot <- PRIME::plotTracks(cage_sel, ran) +
  ggplot2::facet_wrap(
      facets = . ~ type + sample,
      ncol = 1,
      # scales = "free_y",
      strip.position = "right"
    ) +
  coord_cartesian(
      ylim = base::c(minY,maxY)
    ) +
  # scale_y_continuous(
  #     labels = base::c(-0.2,0,0.2),
  #     breaks = base::c(-0.2,0,0.2)
  #   ) +
  # geom_rect(
  #     inherit.aes = FALSE,
  #     data = mycs, # <- MYC enhancers
  #     mapping = aes(
  #         xmin = start,
  #         xmax = end
  #       ),
  #     ymin = minY,
  #     ymax = maxY,
  #     fill = "grey",
  #     alpha = 0.5
  #   ) +
  geom_rect(
      inherit.aes = FALSE,
      data = sites_scores,  # <- PRIME loci
      mapping = aes(
          xmin = start,
          xmax = end,
          fill = score
        ),
      ymin = minY,
      ymax = maxY,
      alpha = 0.5
    ) +
  scale_fill_gradientn(
      limits = base::c(0,1),
      colours = cols
    ) +
  theme(
      aspect.ratio = 1,
      legend.position = "none"
    )

ggrastr::rasterise(
    input = plot,
    layers = "Bar",
    dpi = 300
  )
```

## Print data
```{r eval=FALSE}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("trackHubFiles/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_MYC",
        "_AllCellCTSSsPool",
        # "_enhancer", "_gene", "_e1", "_e2", "_e3_e4", "_e5", "_e6_e7",
        # "_PC91", "_HT292", "_PC92", "_PC93", "_PC94", "_PC95", "_PVT1",
        # "_A3751", "_A3752", "_A3753", "_A3754", "_A3755", "_MDA1", "_MDA2",
        # "_A5493", "_HepG21", "_HCT1161"
        "_PC91_subset",
        out,
        "_v2",
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    device = "pdf",
    dpi = 300,
    width = 3, # 2.5 # 8
    height = 6
  )
```

## [6.] Genomic_landscape -------------------------

## [6.0.] Plot genomic environment
#### evaluate with 'columns(txdb)'
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

start_pos <- BiocGenerics::start(ran)
end_pos <- BiocGenerics::end(ran)

# BiocManager::install("ggbio", version = '3.16')

plot_land <- ggbio::autoplot(
      object = txdb,
      which = ran,
      truncate.gaps = FALSE,
      truncate.fun = NA,
      stat = "identity",
      label = FALSE
    ) +
  scale_x_continuous(
      limits = base::c(start_pos,end_pos)
    ) +
  theme_bw() +
  theme(
      plot.margin = grid::unit(base::c(0.05,0,0,0), "lines")
    )
base::print(plot_land)
```

## [6.1.] Align genomic landscape and score plot
```{r}
g <- ggplot2::ggplotGrob(plot)
l <- ggplot2::ggplotGrob(plot_land@ggplot)

l_add <- gtable::gtable_add_cols(
    x = l,
    widths = grid::unit(1, "lines"),
    pos = -1
  )

l_add$widths <- g$widths

gridExtra::grid.arrange(
    g,l_add,
    heights = grid::unit(base::c(0.8, 0.2),"npc")
  )
```

## [6.2.] Print data
```{r}
# outfile <- base::paste0(
#     here::here("trackHubFiles/plots/"),
#     base::format(base::Sys.time(), "%Y%m%d"),
#     "_proximalTripGro",
#     out,
#     "_RAPGEF2",
#     "_GM12878",
#     "_poolsOnly",
#     ".pdf"
#   )
# 
# grDevices::pdf(
#     file = outfile,
#     width = 8,
#     height = 4
#   )
# base::print(
#     gridExtra::grid.arrange(
#       g,l_add,
#       heights = grid::unit(base::c(0.8, 0.2),"npc")
#     )
#   )
# grDevices::dev.off()
```

```{r}
outfile <- base::paste0(
    here::here("trackHubFiles/plots/"),
    base::format(base::Sys.time(), "%Y%m%d"),
    "_APOBgene",
    # "_ABCA7gene", # "_ABCA7enhancer",
    # "_TMPRSS6gene", # "_TMPRSS6promoter",
    # "_ABOgene", # "_ABOprimelocus", # "_ABOgene", # "_ABOenhancer_2",
    # "_LPLgene", # "_LPLenhancer",
    # "_LIPCgene",
    # "_APOC1intergenic",
    "_unfilteredPRIMEloci",
    # "_pooledPRIMEloci",
    # "_rescaled",
    out,
    ".pdf"
  )

grDevices::pdf(
    file = outfile,
    width = 8, # 4 # 8
    height = 4 # 6 # 4
  )
base::print(
    gridExtra::grid.arrange(
      g,l_add,
      heights = grid::unit(base::c(0.8, 0.2),"npc")
    )
  )
grDevices::dev.off()
```

