---
title: "2.1 Quantifying nascent enrichment within DHSs"
output: html_document
date: "2024-02-29"
---
## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(dplyr)
base::library(PRIME)
```

## [0.1.] Annotations
```{r}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```


## [0.2.] Set working directory
```{r}
#base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/2.NascentTranscription/")
```


# [1.] GM12878 replicate subsetted  -------------------------

## [1.1.] Read in GM12878 replicate subsetted
```{r}
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
  )
)
```


## [1.2.] Read in GM12878 ENCODE DHSs 
```{r}
GM12878.dhs <- utils::read.table(
file = base::paste0(
        here::here("0.External_resources/"),
        "E116-DNase.macs2.hg38.narrowPeak"
)
)

BiocGenerics::colnames(GM12878.dhs) <- c("chr","start","end")
GM12878.dhs <- GenomicRanges::GRanges(GM12878.dhs[,1:3])

GM12878.dhs$thick <- IRanges::IRanges(
  start=start(GM12878.dhs)+((end(GM12878.dhs)-start(GM12878.dhs))/2))

keep <- GenomeInfoDb::seqlevels(GM12878.dhs)[1:23]
GM12878.dhs <- GenomeInfoDb::keepSeqlevels(
  GM12878.dhs, keep, pruning.mode = "coarse")

GM12878.dhs <- GM12878.dhs %>%
  CAGEfightR::swapRanges() %>% 
  GenomicRanges::promoters(upstream=200, downstream = 200)
```


## [1.3.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
GM12878.dhs.s <- subsetByOverlaps(
  GM12878.dhs, rowRanges(GM12878.s.CTSSs))

GM12878.dhs.s<- .GlobalEnv$quantifyClustersOlap(
  GM12878.s.CTSSs, GM12878.dhs.s, inputAssay="counts")
```


## [1.4.] Normalize and annotate DHSs
```{r}
GM12878.dhs.s <- CAGEfightR::calcTPM(GM12878.dhs.s)

GM12878.dhs.s <- CAGEfightR::assignTxType(GM12878.dhs.s, txModels=txdb)
```

## [1.5.] Save output
```{r}
base::saveRDS(GM12878.dhs, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "repSub.DHSs.quantified.GM12878.rds")
)
```


# [2.] GM12878 pooled subsetted  -------------------------

## [2.1.] Read in GM12878 pooled subsetted
```{r}
GM12878.p.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
  )
)
```

#continue from here

## [2.2.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
GM12878.dhs.p <- subsetByOverlaps(
  GM12878.dhs, rowRanges(GM12878.p.CTSSs))

GM12878.dhs.p <- .GlobalEnv$quantifyClustersOlap(
  GM12878.p.CTSSs, GM12878.dhs.p, inputAssay="counts")
```


## [2.3.] Normalize and annotate DHSs
```{r}
GM12878.dhs.p <- CAGEfightR::calcTPM(GM12878.dhs.p)

GM12878.dhs.p <- CAGEfightR::assignTxType(GM12878.dhs.p, txModels=txdb)
```

## [2.4.] Save output
```{r}
base::saveRDS(GM12878.dhs.p, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "poolSub.DHSs.quantified.GM12878.rds")
) 
```





# [3.] k562 replicate subsetted  -------------------------

## [3.1.] Read in k562 replicate subsetted
```{r}
k562.s.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "repSub.combined.k562.CTSSs.rmSingletons.rds")
)
```


## [3.2.] Read in k562 ENCODE DHSs 
```{r}
k562.dhs <- utils::read.table(
  file = base::paste0(
        here::here("0.External_resources/"),
        "E123-DNase.macs2.hg38.narrowPeak")
)

BiocGenerics::colnames(k562.dhs) <- c("chr","start","end")
k562.dhs <- GenomicRanges::GRanges(k562.dhs[,1:3])

k562.dhs$thick <- IRanges::IRanges(
  start=start(k562.dhs)+((end(k562.dhs)-start(k562.dhs))/2))

keep <- GenomeInfoDb::seqlevels(k562.dhs)[1:23]
k562.dhs <- GenomeInfoDb::keepSeqlevels(
  k562.dhs, keep, pruning.mode = "coarse")

k562.dhs <- k562.dhs %>%
  CAGEfightR::swapRanges() %>% 
  GenomicRanges::promoters(upstream=200, downstream = 200)
```


## [3.3.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
k562.dhs.s <- subsetByOverlaps(
  k562.dhs, rowRanges(k562.s.CTSSs))

k562.dhs.s<- .GlobalEnv$quantifyClustersOlap(
  k562.s.CTSSs, k562.dhs.s, inputAssay="counts")
```


## [3.4.] Normalize and annotate DHSs
```{r}
k562.dhs.s <- CAGEfightR::calcTPM(k562.dhs.s)

k562.dhs.s <- CAGEfightR::assignTxType(k562.dhs.s, txModels=txdb)
```

## [3.5.] Save output
```{r}
base::saveRDS(k562.dhs,
file = base::paste0(
        here::here("2.NascentTranscription/"), 
        "repSub.DHSs.quantified.k562.rds")
)
```


# [4.] k562 pooled subsetted  -------------------------

## [4.1.] Read in k562 pooled subsetted
```{r}
k562.p.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "poolSub.combined.k562.CTSSs.rmSingletons.rds")
)
```


## [4.2.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
k562.dhs.p <- subsetByOverlaps(
  k562.dhs, rowRanges(k562.p.CTSSs))

k562.dhs.p <- .GlobalEnv$quantifyClustersOlap(
  k562.p.CTSSs, k562.dhs.p, inputAssay="counts")
```


## [4.3.] Normalize and annotate DHSs
```{r}
k562.dhs.p <- CAGEfightR::calcTPM(k562.dhs.p)

k562.dhs.p <- CAGEfightR::assignTxType(k562.dhs.p, txModels=txdb)
```

## [4.4.] Save output
```{r}
base::saveRDS(k562.dhs.p, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "poolSub.DHSs.quantified.K562.rds")
)
```
