---
title: "2.1 Quantifying nascent enrichment within DHSs"
output: html_document
date: "2024-02-29"
---
## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(dplyr)
base::library(PRIME)
```

## [0.1.] Annotations
```{r}
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```


## [0.2.] Set working directory
```{r}
#base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/2.Cellular_localization/")
```


# [1.] GM12878 replicate subsetted  -------------------------

## [1.1.] Read in GM12878 replicate subsetted
```{r}
GM12878.s.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "repSub.combined.GM12878.CTSSs.rmSingletons.rds"
  )
)
```


## [1.2.] Read in GM12878 ENCODE DHSs 
```{r}
GM12878.dhs <- utils::read.table(
file = base::paste0(
        here::here("0.External_resources/"),
        "E116-DNase.macs2.hg38.narrowPeak"
)
)

BiocGenerics::colnames(GM12878.dhs) <- c("chr","start","end")
GM12878.dhs <- GenomicRanges::GRanges(GM12878.dhs[,1:3])

GM12878.dhs$thick <- IRanges::IRanges(
  start=start(GM12878.dhs)+((end(GM12878.dhs)-start(GM12878.dhs))/2))

keep <- GenomeInfoDb::seqlevels(GM12878.dhs)[1:23]
GM12878.dhs <- GenomeInfoDb::keepSeqlevels(
  GM12878.dhs, keep, pruning.mode = "coarse")

GM12878.dhs <- GM12878.dhs %>%
  CAGEfightR::swapRanges() %>% 
  GenomicRanges::promoters(upstream=200, downstream = 200)
```


## [1.3.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
GM12878.dhs.s <- subsetByOverlaps(
  GM12878.dhs, rowRanges(GM12878.s.CTSSs))

GM12878.dhs.s<- .GlobalEnv$quantifyClustersOlap(
  GM12878.s.CTSSs, GM12878.dhs.s, inputAssay="counts")
```


## [1.4.] Normalize and annotate DHSs
```{r}
GM12878.dhs.s <- CAGEfightR::calcTPM(GM12878.dhs.s)

GM12878.dhs.s <- CAGEfightR::assignTxType(GM12878.dhs.s, txModels=txdb)
```

## [1.5.] Save output
```{r}
base::saveRDS(GM12878.dhs, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "repSub.DHSs.quantified.GM12878.rds")
)
```


# [2.] GM12878 pooled subsetted  -------------------------

## [2.1.] Read in GM12878 pooled subsetted
```{r}
GM12878.p.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "poolSub.combined.GM12878.CTSSs.rmSingletons.rds"
  )
)
```

#continue from here

## [2.2.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
GM12878.dhs.p <- subsetByOverlaps(
  GM12878.dhs, rowRanges(GM12878.p.CTSSs))

GM12878.dhs.p <- .GlobalEnv$quantifyClustersOlap(
  GM12878.p.CTSSs, GM12878.dhs.p, inputAssay="counts")
```


## [2.3.] Normalize and annotate DHSs
```{r}
GM12878.dhs.p <- CAGEfightR::calcTPM(GM12878.dhs.p)

GM12878.dhs.p <- CAGEfightR::assignTxType(GM12878.dhs.p, txModels=txdb)
```

## [2.4.] Save output
```{r}
base::saveRDS(GM12878.dhs.p, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "poolSub.DHSs.quantified.GM12878.rds")
) 
```





# [3.] k562 replicate subsetted  -------------------------

## [3.1.] Read in k562 replicate subsetted
```{r}
k562.s.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "repSub.combined.k562.CTSSs.rmSingletons.rds")
)
```


## [3.2.] Read in k562 ENCODE DHSs 
```{r}
k562.dhs <- utils::read.table(
  file = base::paste0(
        here::here("0.External_resources/"),
        "E123-DNase.macs2.hg38.narrowPeak")
)

BiocGenerics::colnames(k562.dhs) <- c("chr","start","end")
k562.dhs <- GenomicRanges::GRanges(k562.dhs[,1:3])

k562.dhs$thick <- IRanges::IRanges(
  start=start(k562.dhs)+((end(k562.dhs)-start(k562.dhs))/2))

keep <- GenomeInfoDb::seqlevels(k562.dhs)[1:23]
k562.dhs <- GenomeInfoDb::keepSeqlevels(
  k562.dhs, keep, pruning.mode = "coarse")

k562.dhs <- k562.dhs %>%
  CAGEfightR::swapRanges() %>% 
  GenomicRanges::promoters(upstream=200, downstream = 200)
```


## [3.3.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
k562.dhs.s <- subsetByOverlaps(
  k562.dhs, rowRanges(k562.s.CTSSs))

k562.dhs.s<- .GlobalEnv$quantifyClustersOlap(
  k562.s.CTSSs, k562.dhs.s, inputAssay="counts")
```


## [3.4.] Normalize and annotate DHSs
```{r}
k562.dhs.s <- CAGEfightR::calcTPM(k562.dhs.s)

k562.dhs.s <- CAGEfightR::assignTxType(k562.dhs.s, txModels=txdb)
```

## [3.5.] Save output
```{r}
base::saveRDS(k562.dhs,
file = base::paste0(
        here::here("2.NascentTranscription/"), 
        "repSub.DHSs.quantified.k562.rds")
)
```


# [4.] k562 pooled subsetted  -------------------------

## [4.1.] Read in k562 pooled subsetted
```{r}
k562.p.CTSSs <- base::readRDS(
  file = base::paste0(
        here::here("1.CTSSs/"),
        "poolSub.combined.k562.CTSSs.rmSingletons.rds")
)
```


## [4.2.] Quantify CTSS signal within +/- 200bp from DHS center 
```{r}
k562.dhs.p <- subsetByOverlaps(
  k562.dhs, rowRanges(k562.p.CTSSs))

k562.dhs.p <- .GlobalEnv$quantifyClustersOlap(
  k562.p.CTSSs, k562.dhs.p, inputAssay="counts")
```


## [4.3.] Normalize and annotate DHSs
```{r}
k562.dhs.p <- CAGEfightR::calcTPM(k562.dhs.p)

k562.dhs.p <- CAGEfightR::assignTxType(k562.dhs.p, txModels=txdb)
```

## [4.4.] Save output
```{r}
base::saveRDS(k562.dhs.p, 
file = base::paste0(
        here::here("2.NascentTranscription/"),
        "poolSub.DHSs.quantified.K562.rds")
)
```



























##Things to move to plotting part
```{r}
#Packages for plotting
#library(epistack)
#library(BiocParallel)
#library(dplyr)
#library(reshape2)
#library(limma)
#library(ggplot2)

custom.col <- c("#FFDB6D", "#C4961A", "#F4EDCA", 
                "#D16103", "#C3D7A4", "#52854C", "#4E84C4", "#293352","grey") 
                

#tpm.df <- as.data.frame(assay(GM12878.dhs, "TPM"))
#tpm.df$txType <- rowData(GM12878.dhs)$txType
```





```{r}
tpm.df$C_1M <- rowMeans(tpm.df[,c("GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3")])
tpm.df$N_1M <- rowMeans(tpm.df[,c("GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3")])

tpm.df$C_500k <- rowMeans(tpm.df[,c("GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3")])
tpm.df$N_500k <- rowMeans(tpm.df[,c("GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3")])

tpm.df$N_5M <- rowMeans(tpm.df[,c("GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3")])
tpm.df$N_10M <- rowMeans(tpm.df[,c("GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3")])


tpm.df$Ctrl <- rowMeans(tpm.df[,c("GM12878_ctrl_C1","GM12878_ctrl_C2","GM12878_ctrl_C3")])
tpm.df$Trip <- rowMeans(tpm.df[,c("GM12878_triptolide_C1","GM12878_triptolide_C2","GM12878_triptolide_C3","GM12878_triptolide_C4")])

tpm.df$m <- rowMeans(tpm.df[,c("C_1M","N_1M")])
tpm.df$m2 <- rowMeans(tpm.df[,c("C_500k","N_500k")])

tpm.df$m51 <- rowMeans(tpm.df[,c("C_1M","N_5M")])


pdf("./plots.2/repSub.C500K.N500K.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_500k, N_500k, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 500k cellular CAGE [TPM]",y="average DHS signal in 500k nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/repSub.volcano500k.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m2, log2(N_500k/C_500k), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2()+ labs(x="Average expression across 500k CAGE [TPM] ",y="log2(500k nuclei/500k Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.C1M.N1M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_1M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 1M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/repSub.volcano1M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m, log2(N_1M/C_1M), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2()+ labs(x="Average expression across 1M CAGE [TPM] ",y="log2(1M nuclei/1M Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.C1M.N10M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_10M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/repSub.C1M.N5M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_5M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 5M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()



pdf("./plots.2/repSub.volcano5M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m51, log2(N_5M/C_1M), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.volcano5M.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes(m51, log2(N_5M/C_1M), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()

```




```{r}
pdf("./plots.2/repSub.N1M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_1M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M nuclei CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/repSub.C1M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/repSub.N5M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_5M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 5M nuclei CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()


pdf("./plots.2/repSub.N10M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_10M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()
```


```{r}

pdf("./plots.2/repSub.volcano5M.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_5M_N+GM12878_1M_C)/2, log2(GM12878_5M_N/GM12878_1M_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 5MN & 1MC CAGE [TPM] ",y="log2(5M pooled nuclei/1M pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.volcano5M.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_5M_N+GM12878_1M_C)/2, log2(GM12878_5M_N/GM12878_1M_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()


pdf("./plots.2/repSub.volcano1M.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_1M_N+GM12878_1M_C)/2, log2(GM12878_1M_N/GM12878_1M_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 1MN & 1MC CAGE [TPM] ",y="log2(1M pooled nuclei/1M pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.volcano1M.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_1M_N+GM12878_1M_C)/2, log2(GM12878_1M_N/GM12878_1M_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 1MN & 1MC CAGE [TPM] ",y="log2(1M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()




pdf("./plots.2/repSub.volcano500K.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_500K_N+GM12878_500K_C)/2, log2(GM12878_500K_N/GM12878_500K_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 500KN & 500KC CAGE [TPM] ",y="log2(500K pooled nuclei/500K pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/repSub.volcano500K.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_500K_N+GM12878_500K_C)/2, log2(GM12878_500K_N/GM12878_500K_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 500KN & 500KC CAGE [TPM] ",y="log2(500K nuclei/500K Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()






```




#Same now for pooled subsampled



```{r}
#Load in pre computed CTSSs for GM12878
GM12878.p.CTSSs <- readRDS("../1.CTSSs/poolSub.combined.GM12878.CTSSs.rds")
```

```{r}
#Load in DHSs 
GM12878.dhs <- read.table("../0.External_resources/E116-DNase.macs2.hg38.narrowPeak")
colnames(GM12878.dhs) <- c("chr","start","end")
GM12878.dhs <- GRanges(GM12878.dhs[,1:3])
GM12878.dhs$thick <- IRanges(start=start(GM12878.dhs)+((end(GM12878.dhs)-start(GM12878.dhs))/2))
```


```{r}
#Quantify DHS clusters
GM12878.CTSSs.stand <- keepStandardChromosomes(GM12878.p.CTSSs,pruning.mode = "coarse")
GM12878.dhs <- keepStandardChromosomes(GM12878.dhs, pruning.mode="coarse")
GM12878.dhs <- GM12878.dhs %>% swapRanges() %>% promoters(upstream=200, downstream = 200)

seqlevels(GM12878.dhs) <- seqlevels(GM12878.CTSSs.stand)
seqinfo(GM12878.dhs) <- seqinfo(GM12878.CTSSs.stand)
GM12878.dhs<- subsetByOverlaps(GM12878.dhs,rowRanges(GM12878.CTSSs.stand))

GM12878.dhs <- quantifyClustersOlap(GM12878.CTSSs.stand, GM12878.dhs,inputAssay="counts")

```


```{r}
GM12878.dhs <- calcTPM(GM12878.dhs)
tpm.df <- as.data.frame(assay(GM12878.dhs, "TPM"))
GM12878.dhs <- assignTxType(GM12878.dhs, txModels=txdb)
tpm.df$txType <- rowData(GM12878.dhs)$txType
```








```{r}
tpm.df$C_1M <- rowMeans(tpm.df[,c("GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3")])
tpm.df$N_1M <- rowMeans(tpm.df[,c("GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3")])

tpm.df$C_500k <- rowMeans(tpm.df[,c("GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3")])
tpm.df$N_500k <- rowMeans(tpm.df[,c("GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3")])

tpm.df$N_5M <- rowMeans(tpm.df[,c("GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3")])
tpm.df$N_10M <- rowMeans(tpm.df[,c("GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3")])


tpm.df$Ctrl <- rowMeans(tpm.df[,c("GM12878_ctrl_C1","GM12878_ctrl_C2","GM12878_ctrl_C3")])
tpm.df$Trip <- rowMeans(tpm.df[,c("GM12878_triptolide_C1","GM12878_triptolide_C2","GM12878_triptolide_C3","GM12878_triptolide_C4")])

tpm.df$m <- rowMeans(tpm.df[,c("C_1M","N_1M")])
tpm.df$m2 <- rowMeans(tpm.df[,c("C_500k","N_500k")])

tpm.df$m51 <- rowMeans(tpm.df[,c("C_1M","N_5M")])


pdf("./plots.2/poolSub.C500K.N500K.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_500k, N_500k, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 500k cellular CAGE [TPM]",y="average DHS signal in 500k nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/poolSub.volcano500k.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m2, log2(N_500k/C_500k), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2()+ labs(x="Average expression across 500k CAGE [TPM] ",y="log2(500k nuclei/500k Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.C1M.N1M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_1M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 1M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/poolSub.volcano1M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m, log2(N_1M/C_1M), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2()+ labs(x="Average expression across 1M CAGE [TPM] ",y="log2(1M nuclei/1M Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.C1M.N10M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_10M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/poolSub.C1M.N5M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, N_5M, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 5M nuclei CAGE [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()



pdf("./plots.2/poolSub.volcano5M.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(m51, log2(N_5M/C_1M), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.volcano5M.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes(m51, log2(N_5M/C_1M), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()

```



```{r}
pdf("./plots.2/poolSub.N1M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_1M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M nuclei CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/poolSub.C1M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(C_1M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()

pdf("./plots.2/poolSub.N5M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_5M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 5M nuclei CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()


pdf("./plots.2/poolSub.N10M.GroCap.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes(N_10M, GM12878_groCap_1, color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_y_log10() + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 1, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2() + labs(x="average DHS signal in 1M cellular CAGE [TPM]",y="average DHS signal in 10M GroCap [TPM]", color="log2(Triptolide/Control) [TPM]")
dev.off()
```



```{r}

pdf("./plots.2/poolSub.volcano5M.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_5M_N+GM12878_1M_C)/2, log2(GM12878_5M_N/GM12878_1M_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 5MN & 1MC CAGE [TPM] ",y="log2(5M pooled nuclei/1M pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.volcano5M.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_5M_N+GM12878_1M_C)/2, log2(GM12878_5M_N/GM12878_1M_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 5MN & 1MC CAGE [TPM] ",y="log2(5M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()


pdf("./plots.2/poolSub.volcano1M.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_1M_N+GM12878_1M_C)/2, log2(GM12878_1M_N/GM12878_1M_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 1MN & 1MC CAGE [TPM] ",y="log2(1M pooled nuclei/1M pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.volcano1M.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_1M_N+GM12878_1M_C)/2, log2(GM12878_1M_N/GM12878_1M_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 1MN & 1MC CAGE [TPM] ",y="log2(1M nuclei/1M Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()




pdf("./plots.2/poolSub.volcano500K.pooled.trpFC.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_500K_N+GM12878_500K_C)/2, log2(GM12878_500K_N/GM12878_500K_C), color=log2(Trip/Ctrl))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_gradient2(low="#01665E",high="#8E540C")+ labs(x="average expression across pooled 500KN & 500KC CAGE [TPM] ",y="log2(500K pooled nuclei/500K pooled Cellular) [TPM]", color="log2(Triptolide/Control)")
dev.off()

pdf("./plots.2/poolSub.volcano500K.pooled.groCap.pdf", height=10, width=15)
ggplot(tpm.df, aes((GM12878_500K_N+GM12878_500K_C)/2, log2(GM12878_500K_N/GM12878_500K_C), color=log10(1+GM12878_groCap_1))) + geom_point(size=0.5) + scale_x_log10()  +
  theme_bw() +  geom_abline(intercept = 0, slope = 0, color="darkorange") + facet_wrap(~txType) + scale_color_distiller(palette = "BrBG", values = c(0,0.05,0.1, 0.5,1))+ labs(x="Average expression across 500KN & 500KC CAGE [TPM] ",y="log2(500K nuclei/500K Cellular) [TPM]", color="log10(GRO-cap)")
dev.off()






```



