---
title: "10.1.K562_external_validation"
output: html_document
date: "2024-07-01"
---

## Evaluation of predicted regulatory elements using ClinVar non-benign and non-coding variants, fine-mapped GWAS variants, fine-mapped eQTLs, and CRISPRi data.


## [0.0.] Load necessary packages
```{r warning=TRUE, message=TRUE}
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)

# base::library(CAGEfightR)
# base::library(GenomicRanges)
# base::library(readr)
# base::library(rtracklayer)
# base::library(parallel)
# base::library(ggplot2)
# base::library(GenomicFeatures)
# base::library(SummarizedExperiment)
# base::library(tibble)
# base::library(ggrepel)
# base::library(tidyr)
# base::library(ggpattern)

# theme_set(theme_bw(base_size = 10))
```

## [0.1.] Add custom color palettes
```{r}
col.scheme <- base::c(
    "#377eb8","#d39200","#93aa00",
    "#4daf4a","#984ea3","#ff7f01",
    "#e41b1c","#256866","#f8bc24",
    "grey","black","hotpink"
  )
```

## [0.2.] Define all necessary parameters and file paths
```{r}
PRIMEloci.thres = 0.75

pooled.PRIMEloci.files <- base::c(
    "K562_sld_C" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_C.bed"),
    "K562_sld_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N.bed"),
    "K562_sld_10M_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N_10M.bed")
  )

pooled.PRIMEloci.names <- base::c(
    "K562_sld_C" = "PRIMEloci CAGE cells pooled",
    "K562_sld_N" = "PRIMEloci CAGE nuclei pooled",
    "K562_sld_10M_N" = "PRIMEloci CAGE nuclei 10M pooled"
  )

pooled.DL.names <- base::c(
    "K562_C" = "Divergent loci CAGE cells",
    "K562_N" = "Divergent loci CAGE nuclei"
  )

#cCRE.file <- "../0.External_resources/ENCFF464BRU.bed"
cCRE.file <- here::here("0.External_resources/ENCFF414OGC_ENCFF806YEZ_ENCFF849TDM_ENCFF736UDR.bed")

DHS.file <- here::here("0.External_resources/E123-DNase.macs2.hg38.narrowPeak")

FANTOM5.file <- here::here("0.External_resources/F5.hg38.enhancers.expression.matrix.gz")
FANTOM5.enhancers.file <- here::here("0.External_resources/F5.hg38.enhancers.bed.gz")
FANTOM5.id = "CNhs12334"

CRISPR.pos.file <- here::here("5.EPCrisprBenchmark/epc_pos_gr.rds")
CRISPR.neg.file <- here::here("5.EPCrisprBenchmark/epc_neg_gr.rds")

k27ac.file <- here::here("0.External_resources/ENCFF038DDS.bed")

```

## [1.] Read in data -------------------------

## [1.0.] Read in CDSs from txdb
```{r}
## Annotated CDS
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
cds <- base::unlist(GenomicFeatures::cdsBy(txdb,"gene"))
cds <- BiocGenerics::sort(cds)
```

## [1.1.] Overlap CDSs with background SNPs (9,913,974 unique autosomal SNP locations)
Remove SNPs overlapping with CDSs
```{r}
SNP.data <- rtracklayer::import.bed( # 9.991.229
   con = base::paste0(
       here::here("0.External_resources/"),
       "all.bg.SNPs.hg38.baseline.v1.1.bed.sorted"
     ) ## SNP.file
  )
SNP.data <- BiocGenerics::sort(SNP.data)
olap <- GenomicRanges::findOverlaps( ## 93.742
    query = SNP.data,
    subject = cds,
    minoverlap = 1L
  )
SNP.data$name <- NULL
SNP.data <- base::unique( # 9.914.129
    SNP.data[-S4Vectors::queryHits(olap)]
  )
GenomeInfoDb::seqlevels(txdb, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.2.] Read in promoters from txdb
```{r}
promoters <- IRanges::promoters(
    x = txdb,
    upstream = 500,
    downstream = 500
  )
```

## [1.3.] Annotate identified PRIMElocis by txdb annotated promoters into proximal (overlapping with promoter) and distal
## PRIMEloci, pooled replicates; evaluate with n = "K562_sld_C" # 4.049.578
```{r}
resize.width = 200
pooled.PRIMEloci <- base::lapply(
  X = base::names(pooled.PRIMEloci.files),
  FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIMEloci.files[n],
          show_col_types = FALSE
        )
      score <- gr$score # # 4.049.578
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3]) # 4.049.578
      gr$score <- score
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 3.915.085
      gr <- IRanges::resize(
          x = gr,
          width = resize.width,
          fix = "center"
        )
      pc <- base::rep("distal", base::length(gr))
      olap <- GenomicRanges::findOverlaps(
          query = gr,
          subject = promoters,
          minoverlap = 1
        )
      pc[S4Vectors::queryHits(olap)] <- "proximal"
      gr$class <- pc
      return(gr)
    }
  )
base::names(pooled.PRIMEloci) <- base::names(pooled.PRIMEloci.files)
```

## [1.4.] Annotate identified Divergent loci by txdb annotated promoters into proximal (overlapping with promoter) and distal
```{r}
pooled.DLs <- base::readRDS(
    file = base::paste0(
      here::here("3.DivergentLoci/"),
      "poolSub.k562.summitDLs.rmSingletons.rds"
    )
  )
if (base::is.list(pooled.DLs)) {
  pooled.DLs <- base::do.call(
      what = "c",
      args = pooled.DLs
    )
  }
GenomeInfoDb::seqlevels(pooled.DLs, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
pooled.DLs <- pooled.DLs[pooled.DLs$sample %in% base::c("K562_C","K562_N")] # 104.105

resize.width = 200
pooled.DLs <- IRanges::resize(
    x = pooled.DLs,
    width = resize.width,
    fix = "center"
  )
dlc <- base::rep("distal",base::length(pooled.DLs))
olap <- GenomicRanges::findOverlaps(
    query = pooled.DLs,
    subject = promoters,
    minoverlap = 1
  )
dlc[S4Vectors::queryHits(olap)] <- "proximal"
pooled.DLs$class <- dlc # 104.105
```

## [1.5.] Read in noise estimates
```{r}
pooled.noise.thresholds <- readr::read_csv(
    file = base::paste0(
        here::here("9.Noise/"),
        "K562.poolSub.divnoise.csv"
      ),
    col_names = TRUE,
    show_col_types = FALSE
  )
pooled.noise.thresholds <- pooled.noise.thresholds[
    pooled.noise.thresholds[,1]=="95%",
    base::c("K562_C","K562_N","K562_C","K562_N")
  ]
```

## [1.5.] Read in, prepare and subset eQTL data
```{r}
eQTL.pip.thres = 0.5
eQTL.data <- readr::read_tsv(
      file = base::paste0(
        here::here("0.External_resources/"),
        "GTEx_30tissues_release1_pip_01_Whole_Blood.tsv"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
eQTL.data <- eQTL.data[,base::c(1,3)]
base::colnames(eQTL.data) <- base::c("variant","pip")
pos <- base::as.numeric(
  base::sapply(
      X = eQTL.data$variant,
      FUN = function(n){
          base::strsplit(n,"_")[[1]][2]
      }
    )
  ) + 1
eQTL.data <- GenomicRanges::GRanges(
    seqnames = base::as.character(
      base::sapply(
        X = eQTL.data$variant,
        FUN = function(n){
            base::strsplit(n,"_")[[1]][1]
          }
        )
      ),
    score = eQTL.data$pip,
    ranges = IRanges::IRanges(pos, pos)
  )
eQTL.data <- base::sort(eQTL.data)
GenomeInfoDb::seqlevels(eQTL.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
eQTL.data <- base::subset(eQTL.data, score >= eQTL.pip.thres)
eQTL.data$score <- NULL
```

## [1.6.] Subset eQTLs for sites outside of CDSs
```{r}
olap <- GenomicRanges::findOverlaps(
    query = eQTL.data,
    subject = cds,
    minoverlap = 1
  )
eQTL.data <- base::unique(
    eQTL.data[-S4Vectors::queryHits(olap)]
  )
eQTL.n <- base::length(eQTL.data)
```

## [1.7.] Read in, prepare and subset GWAS data
```{r}
GWAS.pip.thres = 0.2

GWAS.data <- readr::read_tsv( # 66.638
    file = base::paste0(
        here::here("0.External_resources/"),
        "UKBB_94traits_release1_pip01_clean_uniq_hg38.bed"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
base::colnames(GWAS.data) <- base::c(
    "chr","start","end",
    "name","score","strand",
    "RS_id","trait","pip"
  )
GWAS.data <- GenomicRanges::makeGRangesFromDataFrame( # 66.638
    df = GWAS.data,
    keep.extra.columns = TRUE,
    starts.in.df.are.0based = TRUE
  )
GWAS.data <- base::subset( # 30.743
    GWAS.data, pip >= GWAS.pip.thres
  )
GenomeInfoDb::seqlevels(GWAS.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.8.] Remove GWAS hits overlapping with CDSs & unwanted traits
```{r}
traits <- base::c(
    "HbA1c","Hb","RBC",
    "MCV","MCH","MCHC"
  )

olap <- GenomicRanges::findOverlaps(
    query = GWAS.data,
    subject = cds,
    minoverlap = 1
  )
GWAS.data <- GWAS.data[-S4Vectors::queryHits(olap)]
GWAS.data <- base::subset(GWAS.data, trait %in% traits)
trait.count.df <- base::do.call(
  what = "rbind",
  args = base::lapply(
    X = traits,
    FUN = function(t) {
      dat <- base::subset(GWAS.data, trait == t)
      return(
          base::as.data.frame(
            base::table(dat$trait)
          )
        )
      }
    )
  )

GWAS.data$pip <- NULL
GWAS.data$trait <- NULL
GWAS.data <- base::unique(GWAS.data) # 2.877
GWAS.n <- base::length(GWAS.data)
```

## [1.9.] Read in and subset ClinVar data for sites not overlapping with CDSs
```{r}
ClinVar.data <- readr::read_tsv(
    file = base::paste0(
        here::here("0.External_resources/"),
        "ClinVar_GRCh38_nonbenign_noncoding_SNV.tsv"
      ),
    show_col_types = FALSE
  )
ClinVar.data <- GenomicRanges::makeGRangesFromDataFrame(
    df = ClinVar.data
  )
GenomeInfoDb::seqlevels(ClinVar.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
olap <- GenomicRanges::findOverlaps(
    query = ClinVar.data,
    subject = cds,
    minoverlap = 1
  )
ClinVar.data <- ClinVar.data[-S4Vectors::queryHits(olap)]
ClinVar.data <- base::sort(ClinVar.data)
ClinVar.data <- base::unique(ClinVar.data)
ClinVar.n <- base::length(ClinVar.data)
```

## [1.10.] Read in all PINTS point estimate data
```{r}
PINTS.files <- base::c(
    "PINTS GRO/PRO-cap distal" = here::here("0.External_resources/K562_Distal_PINTS_GRO_PRO_cap.bed"),
    "PINTS GRO/PRO-cap proximal" = here::here("0.External_resources/K562_Proximal_PINTS_GRO_PRO_cap.bed")
  )

point.data <- base::lapply(
  X = base::names(PINTS.files),
  FUN = function(n){
      rtracklayer::import.bed(PINTS.files[[n]])
    }
  )
base::names(point.data) <- base::names(PINTS.files)
point.data[["PINTS GRO/PRO-cap"]] <- base::c(point.data[[1]],point.data[[2]])
```


```{r}
dREG.files <- base::c(
    "dREG PRO-seq" = here::here("0.External_resources/GSE150625_PROseq_1M_K562_dREGpeaks_clean.bed"),
    "dREG qPRO-seq" = here::here("0.External_resources/GSE150625_qPROseq_1M_K562_dREGpeaks_clean.bed")
  )

for (n in base::names(dREG.files)) {
    gr <- rtracklayer::import.bed(con = dREG.files[n])
    GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
    olap <- GenomicRanges::findOverlaps(
        query = gr,
        subject = promoters,
        minoverlap = 1
      )
    gr.proximal <- gr[base::unique(S4Vectors::queryHits(olap))]
    gr.distal <- gr[base::setdiff(1:base::length(gr), S4Vectors::queryHits(olap))]
    point.data[[n]] <- gr
    point.data[[base::paste0(n," distal")]] <- gr.distal
    point.data[[base::paste0(n," proximal")]] <- gr.proximal
  }
```

## [1.11.] Read in all pooled PRIMEloci files
```{r}
for (n in names(pooled.PRIMEloci.files)) {
  gr <- read_tsv(pooled.PRIMEloci.files[n],show_col_types = FALSE)
  gr <- subset(gr,score>=PRIMEloci.thres)
  gr <- makeGRangesFromDataFrame(gr[,1:3])
  seqlevels(gr,pruning.mode="coarse") <- seqlevels(SNP.data)
  olap <- findOverlaps(gr, promoters, minoverlap = 1)
  gr.proximal <- gr[unique(queryHits(olap))]
  gr.distal <- gr[setdiff(1:length(gr), queryHits(olap))]
  point.data[[pooled.PRIMEloci.names[n]]] <- gr
  point.data[[paste0(pooled.PRIMEloci.names[n]," distal")]] <- gr.distal
  point.data[[paste0(pooled.PRIMEloci.names[n]," proximal")]] <- gr.proximal
}
```


```{r}
cCRE <- read_tsv(cCRE.file,col_names=FALSE,show_col_types=FALSE)
df <- cCRE[,1:3]
colnames(df) <- c("chrom","start","end")
cCRE.data <- makeGRangesFromDataFrame(df)
cCRE.data$class <- cCRE[,10,drop=TRUE]
for (n in c("dELS","PLS")) {
 x <- subset(cCRE.data,class==n)
 x$class <- NULL
 seqlevels(x,pruning.mode="coarse") <- seqlevels(SNP.data)
 olap <- findOverlaps(x, promoters, minoverlap = 1)
 x.proximal <- x[unique(queryHits(olap))]
 x.distal <- x[setdiff(1:length(x), queryHits(olap))]
 point.data[[paste0("ENCODE cCRE ",n," distal")]] <- x.distal
 point.data[[paste0("ENCODE cCRE ",n," proximal")]] <- x.proximal
}
point.data[["ENCODE cCRE"]] <- c(point.data[["ENCODE cCRE dELS distal"]],point.data[["ENCODE cCRE dELS proximal"]],point.data[["ENCODE cCRE PLS distal"]],point.data[["ENCODE cCRE PLS proximal"]])
```


```{r}
DHS <- import.bed(DHS.file)
olap <- findOverlaps(DHS, promoters, minoverlap = 1)
DHS.proximal <- DHS[unique(queryHits(olap))]
DHS.distal <- DHS[setdiff(1:length(DHS), queryHits(olap))]
point.data[["DHS"]] <- DHS
point.data[["DHS distal"]] <- DHS.distal
point.data[["DHS proximal"]] <- DHS.proximal
```


```{r}
k27ac <- read.table(k27ac.file)
colnames(k27ac) <- c("chr","start","end","name","score","strand","signal_value","pvalue","qvalue","peak")
k27ac <- GRanges(k27ac)
k27ac.dhs <- subsetByOverlaps(DHS, k27ac)
start(k27ac) <- start(k27ac)+k27ac$peak
end(k27ac) <- start(k27ac)+1

point.data[["K27ac DHS"]] <- k27ac.dhs
point.data[["K27ac"]] <- k27ac



header <- colnames(read_tsv(FANTOM5.file,n_max=1,show_col_types=FALSE))
FANTOM5.data <- read_tsv(FANTOM5.file,skip=1,col_names=FALSE,show_col_types=FALSE)[,c(1,match(FANTOM5.id,header)+1)]
FANTOM5.enhancers <- flank(swapRanges(import.bed(FANTOM5.enhancers.file)),200,both=TRUE)
FANTOM5.enhancers <- FANTOM5.enhancers[FANTOM5.enhancers$name %in% deframe(FANTOM5.data[FANTOM5.data[2]>0,1])]
point.data[["FANTOM5 enhancers"]] <- FANTOM5.enhancers

for (sample in c("K562_C","K562_N")) {
  gr <- pooled.DLs[pooled.DLs$sample==sample]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[pooled.DL.names[sample]]] <- gr

  gr <- pooled.DLs[pooled.DLs$sample==sample & pooled.DLs$class=="distal"]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[paste0(pooled.DL.names[sample]," distal")]] <- gr

  gr <- pooled.DLs[pooled.DLs$sample==sample & pooled.DLs$class=="proximal"]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[paste0(pooled.DL.names[sample]," proximal")]] <- gr
}

for (n in names(point.data)) {
  point.data[[n]] <- sort(point.data[[n]])
  seqlevels(point.data[[n]],pruning.mode="coarse") <- seqlevels(SNP.data)
  point.data[[n]] <- resize(point.data[[n]], width=resize.width, fix="center")
  point.data[[n]] <- reduce(point.data[[n]],min.gapwidth=0)
}

point.data.assay <- names(point.data)
point.data.loc <- rep("genome",length(point.data.assay))
point.data.loc[grep("distal",point.data.assay)] <- "distal"
point.data.loc[grep("proximal",point.data.assay)] <- "proximal"
point.data.loc[which(point.data.assay == "FANTOM5 enhancers")] <- "distal"
point.data.loc[which(point.data.assay == "ENCODE cCRE dELS")] <- "distal"
point.data.loc[which(point.data.assay == "ENCODE cCRE PLS")] <- "proximal"
point.data.assay <- gsub(" distal", "", gsub(" proximal", "", gsub(" PLS", "", gsub(" dELS","",point.data.assay))))
point.data.assay <- gsub(" distal", "", gsub(" proximal", "", point.data.assay))
names(point.data.assay) <- names(point.data.loc) <- names(point.data)

## CRISPR data
CRISPR.pos.data <- readRDS(CRISPR.pos.file)
CRISPR.neg.data <- readRDS(CRISPR.neg.file)
CRISPR.n <- length(CRISPR.pos.data)

```

Point estimates
```{r}

point.enrichment.precision.recall <- function(data, fg.data, bg.data, assay, location) {
  
  do.call("rbind", lapply(names(data), function(n) {
    
    fg.olap <- findOverlaps(fg.data, data[[n]], minoverlap=1)
    fg.count <- length(unique(queryHits(fg.olap)))
    region.count <- length(unique(subjectHits(fg.olap)))
    bg.olap <- findOverlaps(bg.data, data[[n]], minoverlap=1)
    bg.count <- length(unique(queryHits(bg.olap)))
    
    data.frame(enrichment=(fg.count/length(fg.data)) / (bg.count/length(bg.data)),
               precision=fg.count/(fg.count+bg.count),
               recall=fg.count/length(fg.data),
               fg.count=fg.count,
               bg.count=bg.count,
               region.count=region.count,
               sample=n,
               assay=assay[n],
               loc=location[n])
  }))
}

point.ClinVar.df <- point.enrichment.precision.recall(point.data, ClinVar.data, SNP.data, point.data.assay, point.data.loc)
point.ClinVar.df

point.eQTL.df <- point.enrichment.precision.recall(point.data, eQTL.data, SNP.data, point.data.assay, point.data.loc)
point.eQTL.df

point.GWAS.df <- point.enrichment.precision.recall(point.data, GWAS.data, SNP.data, point.data.assay, point.data.loc)
point.GWAS.df

point.CRISPR.df <- point.enrichment.precision.recall(point.data, CRISPR.pos.data, CRISPR.neg.data, point.data.assay, point.data.loc)
point.CRISPR.df


```

Plot ClinVar point estimate results
```{r}
point.ClinVar.genome.df <- subset(point.ClinVar.df, loc=="genome")
point.ClinVar.non.genome.df <- subset(point.ClinVar.df, loc!="genome")

col.scheme <- setNames(col.scheme, c("DHS","Divergent loci CAGE cells","Divergent loci CAGE nuclei",
                                     "dREG PRO-seq","dREG qPRO-seq","ENCODE cCRE",
                                     "PINTS GRO/PRO-cap","PRIMEloci CAGE cells pooled","PRIMEloci CAGE nuclei pooled","K27ac", "K27ac DHS","PRIMEloci CAGE nuclei 10M pooled"))


ggplot(point.ClinVar.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",ClinVar.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_ClinVar_enrichment_vs_recall_point_estimates.pdf"),width=8,height=7)

## Supplementary Figure
ggplot(point.ClinVar.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",ClinVar.n,")"))+ scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_ClinVar_enrichment_vs_recall_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

## Figure 1 E bar plots
point.ClinVar.genome.df$sample <- factor(point.ClinVar.genome.df$sample, levels=point.ClinVar.genome.df$sample[order(point.ClinVar.genome.df$recall, decreasing = TRUE)])
ggplot(point.ClinVar.genome.df, aes(sample, recall,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",ClinVar.n,")"))+ scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_ClinVar_recall_point_estimates_bar.pdf"),width=3,height=3)

##  Figure 1 E
point.ClinVar.genome.df$sample <- factor(point.ClinVar.genome.df$sample, levels=point.ClinVar.genome.df$sample[order(point.ClinVar.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.ClinVar.genome.df, aes(sample, enrichment,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_ClinVar_enrichment_point_estimates_bar.pdf"),width=3,height=3)


#BarPlots for point estimates
scaling_factor <- max(point.ClinVar.genome.df$recall) / max(point.ClinVar.genome.df$enrichment)

# Create a new column for scaled enrichment
point.ClinVar.genome.df$enrichment_scaled <- point.ClinVar.genome.df$enrichment * scaling_factor

point.ClinVar.genome.df_long <- point.ClinVar.genome.df %>%
  pivot_longer(cols = c(recall, enrichment_scaled), names_to = "metric", values_to = "value") %>%
  mutate(metric = factor(metric, levels = c("recall","enrichment_scaled"))) 

# Plot using ggplot with adjusted spacing
ggplot(data = point.ClinVar.genome.df_long) +
  geom_col(
    aes(
      x = sample,
      y = value,
      fill = sample,
      group = metric,
      alpha = metric  
    ),
    position = position_dodge(width = 0.6),  
    color = "black",
    width = 0.6  
  ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled enrichment
  scale_y_continuous(
    name = "Recall",
    limits = c(0, max(point.ClinVar.genome.df$recall) * 1.1),
    sec.axis = sec_axis(
      ~ . / scaling_factor,  
      name = "Enrichment"
    )
  ) +
  scale_fill_manual(values = col.scheme) +
  scale_alpha_manual(values = c("recall" = 1, "enrichment_scaled" = 0.5)) +  # Set alpha levels for recall and enrichment
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "black"),
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "Sample", fill = "Sample") +
  theme(legend.position = "none")

ggsave(paste0("plots/K562_ClinVar_enrichment_recall_point_estimates_bar.pdf"),width=14,height=6)


```

Plot eQTL point estimate results
```{r}

point.eQTL.genome.df <- subset(point.eQTL.df, loc=="genome")
point.eQTL.non.genome.df <- subset(point.eQTL.df, loc!="genome")

ggplot(point.eQTL.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",eQTL.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_point_estimates.pdf"),width=8,height=7)

## Supplementary Figure
ggplot(point.eQTL.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",eQTL.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

#  Figure 1 F
point.eQTL.genome.df$sample <- factor(point.eQTL.genome.df$sample, levels=point.eQTL.genome.df$sample[order(point.eQTL.genome.df$recall, decreasing = TRUE)])
ggplot(point.eQTL.genome.df, aes(sample, recall,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",eQTL.n,")")) + scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_eQTL_recall_PIP_",eQTL.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

#  Figure 1 F
point.eQTL.genome.df$sample <- factor(point.eQTL.genome.df$sample, levels=point.eQTL.genome.df$sample[order(point.eQTL.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.eQTL.genome.df, aes(sample, enrichment,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+ scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_eQTL_enrichment_PIP_",eQTL.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)


#BarPlots for point estimates
scaling_factor <- max(point.eQTL.genome.df$recall) / max(point.eQTL.genome.df$enrichment)

# Create a new column for scaled enrichment
point.eQTL.genome.df$enrichment_scaled <- point.eQTL.genome.df$enrichment * scaling_factor

point.eQTL.genome.df_long <- point.eQTL.genome.df %>%
  pivot_longer(cols = c(recall, enrichment_scaled), names_to = "metric", values_to = "value") %>%
  mutate(metric = factor(metric, levels = c("recall","enrichment_scaled"))) 

# Plot using ggplot with adjusted spacing
ggplot(data = point.eQTL.genome.df_long) +
  geom_col(
    aes(
      x = sample,
      y = value,
      fill = sample,
      group = metric,
      alpha = metric  
    ),
    position = position_dodge(width = 0.6),  
    color = "black",
    width = 0.6  
  ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled enrichment
  scale_y_continuous(
    name = "Recall",
    limits = c(0, max(point.eQTL.genome.df$recall) * 1.1),
    sec.axis = sec_axis(
      ~ . / scaling_factor,  
      name = "Enrichment"
    )
  ) +
  scale_fill_manual(values = col.scheme) +
  scale_alpha_manual(values = c("recall" = 1, "enrichment_scaled" = 0.5)) +  # Set alpha levels for recall and enrichment
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "black"),
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "Sample", fill = "Sample") +
  theme(legend.position = "none")

ggsave(paste0("plots/K562_eQTL_enrichment_recall_point_estimates_bar.pdf"),width=7,height=5)
```

Plot GWAS point estimate results
```{r}

point.GWAS.genome.df <- subset(point.GWAS.df, loc=="genome")
point.GWAS.non.genome.df <- subset(point.GWAS.df, loc!="genome")

ggplot(point.GWAS.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",GWAS.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_point_estimates.pdf"),width=8,height=7)

ggplot(point.GWAS.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",GWAS.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

point.GWAS.genome.df$sample <- factor(point.GWAS.genome.df$sample, levels=point.GWAS.genome.df$sample[order(point.GWAS.genome.df$recall, decreasing = TRUE)])
ggplot(point.GWAS.genome.df, aes(sample, recall,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",GWAS.n,")")) + scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_GWAS_recall_PIP_",GWAS.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

point.GWAS.genome.df$sample <- factor(point.GWAS.genome.df$sample, levels=point.GWAS.genome.df$sample[order(point.GWAS.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.GWAS.genome.df, aes(sample, enrichment, fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+ scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_GWAS_enrichment_PIP_",GWAS.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

```

Plot CRISPR point estimate results
```{r}

point.CRISPR.genome.df <- subset(point.CRISPR.df, loc=="genome")
point.CRISPR.non.genome.df <- subset(point.CRISPR.df, loc!="genome")

ggplot(point.CRISPR.genome.df, aes(recall, precision, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",CRISPR.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_CRISPR_precision_vs_recall_point_estimates.pdf"),width=8,height=7)

ggplot(point.CRISPR.non.genome.df, aes(recall, precision, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",CRISPR.n,")")) + scale_color_manual(values=col.scheme) + facet_wrap(~loc, scales="free")
ggsave(paste0("plots/K562_CRISPR_precision_vs_recall_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

point.CRISPR.genome.df$sample <- factor(point.CRISPR.genome.df$sample, levels=point.CRISPR.genome.df$sample[order(point.CRISPR.genome.df$recall, decreasing = TRUE)])
ggplot(point.CRISPR.genome.df, aes(sample, recall,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",CRISPR.n,")")) + scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_CRISPR_recall_point_estimates_bar.pdf"),width=3,height=3)

point.CRISPR.genome.df$sample <- factor(point.CRISPR.genome.df$sample, levels=point.CRISPR.genome.df$sample[order(point.CRISPR.genome.df$precision, decreasing = TRUE)])
ggplot(point.CRISPR.genome.df, aes(sample, precision,fill=sample)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + scale_fill_manual(values=col.scheme)
ggsave(paste0("plots/K562_CRISPR_precision_point_estimates_bar.pdf"),width=3,height=3)





#BarPlots for point estimates
scaling_factor <- max(point.CRISPR.genome.df$recall) / max(point.CRISPR.genome.df$precision)

# Create a new column for scaled precision
point.CRISPR.genome.df$precision_scaled <- point.CRISPR.genome.df$precision * scaling_factor

point.CRISPR.genome.df_long <- point.CRISPR.genome.df %>%
  pivot_longer(cols = c(recall, precision_scaled), names_to = "metric", values_to = "value") %>%
  mutate(metric = factor(metric, levels = c("recall","precision_scaled"))) 

# Plot using ggplot with adjusted spacing
ggplot(data = point.CRISPR.genome.df_long) +
  geom_col(
    aes(
      x = sample,
      y = value,
      fill = sample,
      group = metric,
      alpha = metric  
    ),
    position = position_dodge(width = 0.6),  
    color = "black",
    width = 0.6  
  ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled precision
  scale_y_continuous(
    name = "Recall",
    limits = c(0, max(point.CRISPR.genome.df$recall) * 1.1),
    sec.axis = sec_axis(
      ~ . / scaling_factor,  
      name = "Precision"
    )
  ) +
  scale_fill_manual(values = col.scheme) +
  scale_alpha_manual(values = c("recall" = 1, "precision_scaled" = 0.5)) +  # Set alpha levels for recall and precision
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "black"),
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "Sample", fill = "Sample") +
  theme(legend.position = "none")

ggsave(paste0("plots/K562_CRISPR_precision_recall_point_estimates_bar.pdf"),width=7,height=5)


```


Enrichment, precision, and recall versus score threshold
```{r}

enrichment.precision.recall <- function(data, fg.data, bg.data, assay) {
  
  do.call("rbind", lapply(names(data), function(n) {
    
    gr <- data[[n]]
    scores <- seq(0,1,by=0.01)
    cnt <- sapply(scores, function(sco) sum(gr$score>=sco))
    scores <- scores[cnt>=10]
    
    gr.fg.olap <- findOverlaps(fg.data, gr, minoverlap=1)
    gr.bg.olap <- findOverlaps(bg.data, gr, minoverlap=1)
    
    do.call("rbind", lapply(scores, function(sco) {
      
      selection <- which(gr$score>=sco)
      fg.count <- length(unique(queryHits(gr.fg.olap)[which(subjectHits(gr.fg.olap) %in% selection)]))
      bg.count <- length(unique(queryHits(gr.bg.olap)[which(subjectHits(gr.bg.olap) %in% selection)]))
      
      data.frame(enrichment=(fg.count/length(fg.data))/(bg.count/length(bg.data)),
                 precision=fg.count/(fg.count+bg.count),
                 recall=fg.count/length(fg.data),
                 fg.count=fg.count,
                 bg.count=bg.count,
                 score=sco,
                 sample=n,
                 assay=assay[n])
    }))
    
  }))
}

pooled.PRIMEloci.ClinVar.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, ClinVar.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.ClinVar.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, ClinVar.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.eQTL.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, eQTL.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.eQTL.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, eQTL.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.GWAS.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, GWAS.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.GWAS.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, GWAS.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.CRISPR.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, CRISPR.pos.data, CRISPR.neg.data, pooled.PRIMEloci.names)
#PRIMEloci.CRISPR.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, CRISPR.pos.data, CRISPR.neg.data, PRIMEloci.names)

```


Plot results
```{r}

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data[pooled.PRIMEloci.CRISPR.enrichment.recall.data$recall>0.01,], aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=subset(point.ClinVar.df,loc=="genome")) + xlab(paste0("recall (n=",ClinVar.n,")")) + scale_color_manual(values=col.scheme)
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_enrichment_vs_recall_and_point_estimates_genome.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data[pooled.PRIMEloci.CRISPR.enrichment.recall.data$recall>0.01,], aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.eQTL.genome.df) + xlab(paste0("recall (n=",eQTL.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_and_point_estimates_genome.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data[pooled.PRIMEloci.CRISPR.enrichment.recall.data$recall>0.01,], aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.GWAS.genome.df) + xlab(paste0("recall (n=",GWAS.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_and_point_estimates_genome.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data[pooled.PRIMEloci.CRISPR.enrichment.recall.data$recall>0.01,], aes(recall, precision, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.CRISPR.genome.df) + xlab(paste0("recall (n=",CRISPR.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_precision_vs_recall_and_point_estimates_genome.pdf"),width=7,height=5)

```


```{r}

#Specifically for distal
pooled.PRIMEloci.distal <- lapply(pooled.PRIMEloci, function(gr) gr[gr$class == "distal"])
pooled.PRIMEloci.proximal <- lapply(pooled.PRIMEloci, function(gr) gr[gr$class == "proximal"])

pooled.PRIMEloci.GWAS.enrichment.recall.data.distal <- enrichment.precision.recall(pooled.PRIMEloci.distal, GWAS.data, SNP.data, pooled.PRIMEloci.names)

pooled.PRIMEloci.GWAS.enrichment.recall.data.proximal <- enrichment.precision.recall(pooled.PRIMEloci.proximal, GWAS.data, SNP.data, pooled.PRIMEloci.names)

point.GWAS.genome.distal <- subset(point.GWAS.df, loc=="distal")

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data.distal[pooled.PRIMEloci.GWAS.enrichment.recall.data.distal$recall>0.001,], aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.GWAS.genome.distal) + xlab(paste0("recall (n=",GWAS.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_recall_and_point_estimates_distal.pdf"),width=7,height=5)

point.GWAS.genome.proximal <- subset(point.GWAS.df, loc=="proximal")

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data.proximal[pooled.PRIMEloci.GWAS.enrichment.recall.data.proximal$recall>0.001,], aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.GWAS.genome.proximal) + xlab(paste0("recall (n=",GWAS.n,")")) + scale_color_manual(values=col.scheme)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_recall_and_point_estimates_proximal.pdf"),width=7,height=5)

```


Enrichment, precision, recall vs score

```{r}

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_enrichment_vs_score.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_recall_vs_score.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_enrichment_vs_score_PIP_",eQTL.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_recall_vs_score_PIP_",eQTL.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_score_PIP_",GWAS.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_recall_vs_score_PIP_",GWAS.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data, aes(score, precision, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_precision_vs_score.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_recall_vs_score.pdf"),width=7,height=5)

```



```{r}

base::library(UpSetR) 
gr_list <- GRangesList(point.data)

gr_list[["GWAS.data"]] <- GWAS.data
gr_list[["eQTL.data"]] <- eQTL.data
gr_list[["ClinVar.data"]] <- ClinVar.data
gr_list[["CRISPR.data"]] <- CRISPR.pos.data
gr_list[["CRISPR.neg.data"]] <- CRISPR.neg.data




create_upset_plot <- function(subset_name, gr_list, sets_to_include) {
  # Get the specific subset's GRanges
  subset_gr <- gr_list[[subset_name]]

  # Create a binary matrix for overlaps, restricted to the subset
  overlap_matrix <- sapply(gr_list, function(gr) {
    overlaps <- findOverlaps(subset_gr, gr)
    overlap_vector <- rep(0, length(subset_gr))
    overlap_vector[unique(queryHits(overlaps))] <- 1
    return(overlap_vector)
  })
  colnames(overlap_matrix) <- names(gr_list)
  
  # Convert to a data frame
  overlap_df <- as.data.frame(overlap_matrix)
  rownames(overlap_df) <- NULL
  
  # Filter to include only relevant sets
  subset_overlap_df <- overlap_df[, sets_to_include]
  
  # Generate the UpSet plot
  base::library(UpSetR)
  p<- upset(
    subset_overlap_df,
    sets = names(gr_list)[sets_to_include],
    order.by = "freq",
    sets.bar.color=col.scheme[names(col.scheme) %in% names(gr_list)[sets_to_include]]
  )
  
    p1 <-upset(
    subset_overlap_df,
    sets = names(gr_list)[sets_to_include],
    order.by = "freq",
    sets.bar.color=col.scheme[match(p$labels,names(col.scheme))],
    line.size=0
  )
    p1
}





# Define sets to include for intersection analysis
sets_to_include <-  names(gr_list) %in% c("PRIMEloci CAGE cells pooled", "PRIMEloci CAGE nuclei pooled","DHS")

# Generate plots for PRIMEloci/DHS focus
s <- create_upset_plot("GWAS.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_GWAS_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("ClinVar.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_ClinVar_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("eQTL.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_eQTL_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("CRISPR.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_CRISPR.pos_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("CRISPR.neg.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_CRISPR.neg_upset.pdf",height=6, width=8)
s
dev.off()


# Generate plots for full panel
sets_to_include <-  names(gr_list) %in% names(col.scheme)

s <- create_upset_plot("GWAS.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_full_GWAS_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("ClinVar.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_full_ClinVar_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("eQTL.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_full_eQTL_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("CRISPR.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_full_CRISPR.pos_upset.pdf",height=6, width=8)
s
dev.off()

s <- create_upset_plot("CRISPR.neg.data", gr_list, sets_to_include)
pdf("plots/K562_PRIMEloci_full_CRISPR.neg_upset.pdf",height=6, width=8)
s
dev.off()





```


