---
title: "10.1.K562_external_validation"
output: html_document
date: "2024-07-01"
---

Evaluation of predicted regulatory elements using ClinVar non-benign and non-coding variants, fine-mapped GWAS variants, 
fine-mapped eQTLs, and CRISPRi data.

```{r}

library(CAGEfightR)
library(GenomicRanges)
library(readr)
library(rtracklayer)
library(parallel)
library(ggplot2)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(SummarizedExperiment)
library(tibble)
library(ggrepel)

theme_set(theme_bw(base_size = 10))

```

Parameters

```{r}

pooled.focus.samples <- c("K562_C","K562_N")
focus.samples <- c("K562_C1","K562_C2","K562_C3","K562_N4","K562_N5","K562_N6")

PRIMEloci.thres <- 0.2

PRIMEloci.files <- paste("../8.Genomewide_prediction/pred_all_repSub_tcs_",focus.samples,".bed",sep="")
names(PRIMEloci.files) <- focus.samples

PRIMEloci.names <- c("K562_C1" = "PRIMEloci CAGE cells replicates",
                     "K562_C2" = "PRIMEloci CAGE cells replicates",
                     "K562_C3" = "PRIMEloci CAGE cells replicates",
                     "K562_N4" = "PRIMEloci CAGE nuclei replicates",
                     "K562_N5" = "PRIMEloci CAGE nuclei replicates",
                     "K562_N6" = "PRIMEloci CAGE nuclei replicates")

pooled.PRIMEloci.files <- c("K562_C" = "../8.Genomewide_prediction/prediction_all-cell-types_with_Meena-GM12878-model/K562-on-Meena-GM12878-model_pred_all_profiles_subtnorm_tcs_K562_C.bed",
                            "K562_N" = "../8.Genomewide_prediction/prediction_all-cell-types_with_Meena-GM12878-model/K562-on-Meena-GM12878-model_pred_all_profiles_subtnorm_tcs_K562_N.bed")

pooled.PRIMEloci.names <- c("K562_C" = "PRIMEloci CAGE cells pooled",
                            "K562_N" = "PRIMEloci CAGE nuclei pooled")

pooled.DL.file <- "../3.DivergentLoci/poolSub.k562.summitDLs.rmSingletons.rds"
pooled.noise.file <- "../9.Noise/K562.poolSub.divnoise.csv"

pooled.DL.names <- c("K562_C" = "Divergent loci CAGE cells",
                     "K562_N" = "Divergent loci CAGE nuclei")

resize.width <- 200

SNP.file <- "../0.External_resources/all.bg.SNPs.hg38.baseline.v1.1.bed.sorted"
eQTL.file <- "../0.External_resources/GTEx_30tissues_release1_pip_01_Whole_Blood.tsv"
GWAS.file <- "../0.External_resources/UKBB_94traits_release1_pip01_clean_uniq_hg38.bed"
ClinVar.file <- "../0.External_resources/ClinVar_GRCh38_nonbenign_noncoding_SNV.tsv"

eQTL.pip.thres <- 0.5

GWAS.pip.thres <- 0.2
traits <- c("HbA1c","Hb","RBC","MCV","MCH","MCHC")

cCRE.file <- "../0.External_resources/ENCFF464BRU.bed"

DHS.file <- "../0.External_resources/E123-DNase.macs2.hg38.narrowPeak"

PINTS.files <- c("PINTS GRO/PRO-cap distal" = "../0.External_resources/K562_Distal_PINTS_GRO_PRO_cap.bed",
                 "PINTS GRO/PRO-cap proximal" = "../0.External_resources/K562_Proximal_PINTS_GRO_PRO_cap.bed")

dREG.files <- c("dREG PRO-seq" = "../0.External_resources/GSE150625_PROseq_1M_K562_dREGpeaks_clean.bed",
                "dREG qPRO-seq" = "../0.External_resources/GSE150625_qPROseq_1M_K562_dREGpeaks_clean.bed")

FANTOM5.file <- "../0.External_resources/F5.hg38.enhancers.expression.matrix.gz"
FANTOM5.enhancers.file <- "../0.External_resources/F5.hg38.enhancers.bed.gz"
FANTOM5.id <- "CNhs12334"

CRISPR.pos.file <- "../5.EPCrisprBenchmark/1.EPC.pos.gr.rds"
CRISPR.neg.file <- "../5.EPCrisprBenchmark/1.EPC.neg.notsig.gr.rds"

```


Read in data

```{r}

## Annotated CDS
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
cds <- unlist(cdsBy(txdb,"gene"))

## Background SNPs (9,913,974 unique autosomal SNP locations)
SNP.data <- import.bed(SNP.file)
SNP.data <- sort(SNP.data)
olap <- findOverlaps(SNP.data,cds,minoverlap=1)
SNP.data$name <- NULL
SNP.data <- unique(SNP.data[-queryHits(olap)])

## Annotated promoters
seqlevels(txdb,pruning.mode="coarse") <- seqlevels(SNP.data)
promoters <- promoters(txdb,upstream=500,downstream=500)

## PRIMEloci, replicates
PRIMEloci <- lapply(names(PRIMEloci.files), function(n) {
  gr <- read_tsv(PRIMEloci.files[n],show_col_types = FALSE)
  score <- gr$score
  gr <- makeGRangesFromDataFrame(gr[,1:3])
  gr$score <- score
  seqlevels(gr,pruning.mode="coarse") <- seqlevels(SNP.data)
  gr <- resize(gr, width=resize.width, fix="center")
  pc <- rep("distal",length(gr))
  olap <- findOverlaps(gr,promoters,minoverlap = 1)
  pc[queryHits(olap)] <- "proximal"
  gr$class <- pc
  gr
})
names(PRIMEloci) <- names(PRIMEloci.files)

## PRIMEloci, pooled replicates
pooled.PRIMEloci <- lapply(names(pooled.PRIMEloci.files), function(n) {
  gr <- read_tsv(pooled.PRIMEloci.files[n],show_col_types = FALSE)
  score <- gr$score
  gr <- makeGRangesFromDataFrame(gr[,1:3])
  gr$score <- score
  seqlevels(gr,pruning.mode="coarse") <- seqlevels(SNP.data)
  gr <- resize(gr, width=resize.width, fix="center")
  pc <- rep("distal",length(gr))
  olap <- findOverlaps(gr,promoters,minoverlap = 1)
  pc[queryHits(olap)] <- "proximal"
  gr$class <- pc
  gr
})
names(pooled.PRIMEloci) <- names(pooled.PRIMEloci.files)

## Divergent loci from pooled samples
pooled.DLs <- readRDS(pooled.DL.file)
if (is.list(pooled.DLs))
  pooled.DLs <- do.call("c",pooled.DLs)
seqlevels(pooled.DLs, pruning.mode="coarse") <- seqlevels(SNP.data)
pooled.DLs <- pooled.DLs[pooled.DLs$sample %in% pooled.focus.samples]
pooled.DLs <- resize(pooled.DLs, width=resize.width, fix="center")
dlc <- rep("distal",length(pooled.DLs))
olap <- findOverlaps(pooled.DLs,promoters,minoverlap = 1)
dlc[queryHits(olap)] <- "proximal"
pooled.DLs$class <- dlc

## noise estimates
pooled.noise.thresholds <- read_csv(pooled.noise.file,col_names=TRUE,show_col_types=FALSE)
pooled.noise.thresholds <- pooled.noise.thresholds[pooled.noise.thresholds[,1]=="95%",pooled.focus.samples]

## eQTL data
eQTL.data <- read_tsv(eQTL.file,col_names=FALSE,show_col_types=FALSE)[,c(1,3)]
colnames(eQTL.data) <- c("variant","pip")
pos <- as.numeric(sapply(eQTL.data$variant, function(n) strsplit(n,"_")[[1]][2]))+1
eQTL.data <- GRanges(seqnames=as.character(sapply(eQTL.data$variant, function(n) strsplit(n,"_")[[1]][1])),
                     score=eQTL.data$pip,
                     ranges=IRanges(pos, pos))
eQTL.data <- sort(eQTL.data)
seqlevels(eQTL.data,pruning.mode="coarse") <- seqlevels(SNP.data)
eQTL.data <- subset(eQTL.data, score >= eQTL.pip.thres)
eQTL.data$score <- NULL
olap <- findOverlaps(eQTL.data,cds,minoverlap=1)
eQTL.data <- unique(eQTL.data[-queryHits(olap)])
eQTL.n <- length(eQTL.data)

## GWAS data
GWAS.data <- read_tsv(GWAS.file,col_names=FALSE,show_col_types=FALSE)
colnames(GWAS.data) <- c("chr","start","end","name","score","strand","RS_id","trait","pip")
GWAS.data <- makeGRangesFromDataFrame(GWAS.data,keep.extra.columns = TRUE, starts.in.df.are.0based=TRUE)
GWAS.data <- subset(GWAS.data, pip >= GWAS.pip.thres)
seqlevels(GWAS.data,pruning.mode="coarse") <- seqlevels(SNP.data)
olap <- findOverlaps(GWAS.data,cds,minoverlap=1)
GWAS.data <- GWAS.data[-queryHits(olap)]
GWAS.data <- subset(GWAS.data, trait %in% traits)
trait.count.df <- do.call("rbind", lapply(traits, function(t) {
  dat <- subset(GWAS.data, trait==t)
  as.data.frame(table(dat$trait))
}))
trait.count.df
GWAS.data$pip <- NULL
GWAS.data$trait <- NULL
GWAS.data <- unique(GWAS.data)
GWAS.n <- length(GWAS.data)

## ClinVar data
ClinVar.data <- read_tsv(ClinVar.file,show_col_types=FALSE)
ClinVar.data <- makeGRangesFromDataFrame(ClinVar.data)
seqlevels(ClinVar.data,pruning.mode="coarse") <- seqlevels(SNP.data)
olap <- findOverlaps(ClinVar.data,cds,minoverlap=1)
ClinVar.data <- ClinVar.data[-queryHits(olap)]
ClinVar.data <- sort(ClinVar.data)
ClinVar.data <- unique(ClinVar.data)
ClinVar.n <- length(ClinVar.data)

## Point estimate data
point.data <- lapply(names(PINTS.files), function(n) import.bed(PINTS.files[[n]]))
names(point.data) <- names(PINTS.files)
point.data[["PINTS GRO/PRO-cap"]] <- c(point.data[[1]],point.data[[2]])

for (n in names(dREG.files)) {
  gr <- import.bed(dREG.files[n])
  seqlevels(gr,pruning.mode="coarse") <- seqlevels(SNP.data)
  olap <- findOverlaps(gr, promoters, minoverlap = 1)
  gr.proximal <- gr[unique(queryHits(olap))]
  gr.distal <- gr[setdiff(1:length(gr), queryHits(olap))]
  point.data[[n]] <- gr
  point.data[[paste0(n," distal")]] <- gr.distal
  point.data[[paste0(n," proximal")]] <- gr.proximal
}

for (n in names(pooled.PRIMEloci.files)) {
  gr <- read_tsv(pooled.PRIMEloci.files[n],show_col_types = FALSE)
  gr <- subset(gr,score>=PRIMEloci.thres)
  gr <- makeGRangesFromDataFrame(gr[,1:3])
  seqlevels(gr,pruning.mode="coarse") <- seqlevels(SNP.data)
  olap <- findOverlaps(gr, promoters, minoverlap = 1)
  gr.proximal <- gr[unique(queryHits(olap))]
  gr.distal <- gr[setdiff(1:length(gr), queryHits(olap))]
  point.data[[pooled.PRIMEloci.names[n]]] <- gr
  point.data[[paste0(pooled.PRIMEloci.names[n]," distal")]] <- gr.distal
  point.data[[paste0(pooled.PRIMEloci.names[n]," proximal")]] <- gr.proximal
}

cCRE <- read_tsv(cCRE.file,col_names=FALSE,show_col_types=FALSE)
df <- cCRE[,1:3]
colnames(df) <- c("chrom","start","end")
cCRE.data <- makeGRangesFromDataFrame(df)
cCRE.data$class <- cCRE[,10,drop=TRUE]
for (n in c("dELS","PLS")) {
 x <- subset(cCRE.data,class==n)
 x$class <- NULL
 point.data[[paste0("ENCODE cCRE ",n)]] <- x
}
point.data[["ENCODE cCRE"]] <- c(point.data[["ENCODE cCRE dELS"]],point.data[["ENCODE cCRE PLS"]])

DHS <- import.bed(DHS.file)
olap <- findOverlaps(DHS, promoters, minoverlap = 1)
DHS.proximal <- DHS[unique(queryHits(olap))]
DHS.distal <- DHS[setdiff(1:length(DHS), queryHits(olap))]
point.data[["DHS"]] <- DHS
point.data[["DHS distal"]] <- DHS.distal
point.data[["DHS proximal"]] <- DHS.proximal

header <- colnames(read_tsv(FANTOM5.file,n_max=1,show_col_types=FALSE))
FANTOM5.data <- read_tsv(FANTOM5.file,skip=1,col_names=FALSE,show_col_types=FALSE)[,c(1,match(FANTOM5.id,header)+1)]
FANTOM5.enhancers <- flank(swapRanges(import.bed(FANTOM5.enhancers.file)),200,both=TRUE)
FANTOM5.enhancers <- FANTOM5.enhancers[FANTOM5.enhancers$name %in% deframe(FANTOM5.data[FANTOM5.data[2]>0,1])]
point.data[["FANTOM5 enhancers"]] <- FANTOM5.enhancers

for (sample in pooled.focus.samples) {
  gr <- pooled.DLs[pooled.DLs$sample==sample]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[pooled.DL.names[sample]]] <- gr

  gr <- pooled.DLs[pooled.DLs$sample==sample & pooled.DLs$class=="distal"]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[paste0(pooled.DL.names[sample]," distal")]] <- gr

  gr <- pooled.DLs[pooled.DLs$sample==sample & pooled.DLs$class=="proximal"]
  gr <- gr[which(gr$score > deframe(pooled.noise.thresholds[,sample]))]
  point.data[[paste0(pooled.DL.names[sample]," proximal")]] <- gr
}

for (n in names(point.data)) {
  point.data[[n]] <- sort(point.data[[n]])
  seqlevels(point.data[[n]],pruning.mode="coarse") <- seqlevels(SNP.data)
  point.data[[n]] <- resize(point.data[[n]], width=resize.width, fix="center")
  point.data[[n]] <- reduce(point.data[[n]],min.gapwidth=0)
}

point.data.assay <- names(point.data)
point.data.loc <- rep("genome",length(point.data.assay))
point.data.loc[grep("distal",point.data.assay)] <- "distal"
point.data.loc[grep("proximal",point.data.assay)] <- "proximal"
point.data.loc[which(point.data.assay == "FANTOM5 enhancers")] <- "distal"
point.data.loc[which(point.data.assay == "ENCODE cCRE dELS")] <- "distal"
point.data.loc[which(point.data.assay == "ENCODE cCRE PLS")] <- "proximal"
point.data.assay <- gsub(" distal", "", gsub(" proximal", "", gsub(" PLS", "", gsub(" dELS","",point.data.assay))))
point.data.assay <- gsub(" distal", "", gsub(" proximal", "", point.data.assay))
names(point.data.assay) <- names(point.data.loc) <- names(point.data)

## CRISPR data
CRISPR.pos.data <- readRDS(CRISPR.pos.file)
CRISPR.neg.data <- readRDS(CRISPR.neg.file)
CRISPR.n <- length(CRISPR.pos.data)

```

Point estimates

```{r}

point.enrichment.precision.recall <- function(data, fg.data, bg.data, assay, location) {
  
  do.call("rbind", lapply(names(data), function(n) {
    
    fg.olap <- findOverlaps(fg.data, data[[n]], minoverlap=1)
    fg.count <- length(unique(queryHits(fg.olap)))
    region.count <- length(unique(subjectHits(olap)))
    bg.olap <- findOverlaps(bg.data, data[[n]], minoverlap=1)
    bg.count <- length(unique(queryHits(bg.olap)))
    
    data.frame(enrichment=(fg.count/length(fg.data)) / (bg.count/length(bg.data)),
               precision=fg.count/(fg.count+bg.count),
               recall=fg.count/length(fg.data),
               fg.count=fg.count,
               bg.count=bg.count,
               region.count=region.count,
               sample=n,
               assay=assay[n],
               loc=location[n])
  }))
}

point.ClinVar.df <- point.enrichment.precision.recall(point.data, ClinVar.data, SNP.data, point.data.assay, point.data.loc)
point.ClinVar.df

point.eQTL.df <- point.enrichment.precision.recall(point.data, eQTL.data, SNP.data, point.data.assay, point.data.loc)
point.eQTL.df

point.GWAS.df <- point.enrichment.precision.recall(point.data, GWAS.data, SNP.data, point.data.assay, point.data.loc)
point.GWAS.df

point.CRISPR.df <- point.enrichment.precision.recall(point.data, CRISPR.pos.data, CRISPR.neg.data, point.data.assay, point.data.loc)
point.CRISPR.df


```

Plot ClinVar point estimate results

```{r}

point.ClinVar.genome.df <- subset(point.ClinVar.df, loc=="genome")
point.ClinVar.non.genome.df <- subset(point.ClinVar.df, loc!="genome")

ggplot(point.ClinVar.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",ClinVar.n,")"))
ggsave(paste0("plots/K562_ClinVar_enrichment_vs_recall_point_estimates.pdf"),width=8,height=7)

## Supplementary Figure
#ggplot(point.ClinVar.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
#  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
#  xlab(paste0("recall (n=",ClinVar.n,")"))
#ggsave(paste0("plots/K562_ClinVar_enrichment_vs_recall_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

## Figure 1 E bar plots
point.ClinVar.genome.df$sample <- factor(point.ClinVar.genome.df$sample, levels=point.ClinVar.genome.df$sample[order(point.ClinVar.genome.df$recall, decreasing = TRUE)])
ggplot(point.ClinVar.genome.df, aes(sample, recall)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",ClinVar.n,")"))
ggsave(paste0("plots/K562_ClinVar_recall_point_estimates_bar.pdf"),width=3,height=3)

##  Figure 1 E
point.ClinVar.genome.df$sample <- factor(point.ClinVar.genome.df$sample, levels=point.ClinVar.genome.df$sample[order(point.ClinVar.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.ClinVar.genome.df, aes(sample, enrichment)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))
ggsave(paste0("plots/K562_ClinVar_enrichment_point_estimates_bar.pdf"),width=3,height=3)

```

Plot eQTL point estimate results

```{r}

point.eQTL.genome.df <- subset(point.eQTL.df, loc=="genome")
point.eQTL.non.genome.df <- subset(point.eQTL.df, loc!="genome")

ggplot(point.eQTL.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",eQTL.n,")"))
ggsave(paste0("plots/K562_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_point_estimates.pdf"),width=8,height=7)

## Supplementary Figure
#ggplot(point.eQTL.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
#  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
#  xlab(paste0("recall (n=",eQTL.n,")"))
#ggsave(paste0("plots/K562_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

#  Figure 1 F
point.eQTL.genome.df$sample <- factor(point.eQTL.genome.df$sample, levels=point.eQTL.genome.df$sample[order(point.eQTL.genome.df$recall, decreasing = TRUE)])
ggplot(point.eQTL.genome.df, aes(sample, recall)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",eQTL.n,")"))
ggsave(paste0("plots/K562_eQTL_recall_PIP_",eQTL.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

#  Figure 1 F
point.eQTL.genome.df$sample <- factor(point.eQTL.genome.df$sample, levels=point.eQTL.genome.df$sample[order(point.eQTL.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.eQTL.genome.df, aes(sample, enrichment)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))
ggsave(paste0("plots/K562_eQTL_enrichment_PIP_",eQTL.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

```

Plot GWAS point estimate results

```{r}

point.GWAS.genome.df <- subset(point.GWAS.df, loc=="genome")
point.GWAS.non.genome.df <- subset(point.GWAS.df, loc!="genome")

ggplot(point.GWAS.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",GWAS.n,")"))
ggsave(paste0("plots/K562_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_point_estimates.pdf"),width=8,height=7)

#ggplot(point.GWAS.non.genome.df, aes(recall, enrichment, color=assay, label=sample)) + 
#  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
#  xlab(paste0("recall (n=",GWAS.n,")"))
#ggsave(paste0("plots/K562_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

point.GWAS.genome.df$sample <- factor(point.GWAS.genome.df$sample, levels=point.GWAS.genome.df$sample[order(point.GWAS.genome.df$recall, decreasing = TRUE)])
ggplot(point.GWAS.genome.df, aes(sample, recall)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",GWAS.n,")"))
ggsave(paste0("plots/K562_GWAS_recall_PIP_",GWAS.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

point.GWAS.genome.df$sample <- factor(point.GWAS.genome.df$sample, levels=point.GWAS.genome.df$sample[order(point.GWAS.genome.df$enrichment, decreasing = TRUE)])
ggplot(point.GWAS.genome.df, aes(sample, enrichment)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))
ggsave(paste0("plots/K562_GWAS_enrichment_PIP_",GWAS.pip.thres,"_point_estimates_bar.pdf"),width=3,height=3)

```

Plot CRISPR point estimate results

```{r}

point.CRISPR.genome.df <- subset(point.CRISPR.df, loc=="genome")
point.CRISPR.non.genome.df <- subset(point.CRISPR.df, loc!="genome")

ggplot(point.CRISPR.genome.df, aes(recall, precision, color=assay, label=sample)) + 
  geom_point() + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
  xlab(paste0("recall (n=",CRISPR.n,")"))
ggsave(paste0("plots/K562_CRISPR_precision_vs_recall_point_estimates.pdf"),width=8,height=7)

#ggplot(point.CRISPR.non.genome.df, aes(recall, precision, color=assay, label=sample)) + 
#  geom_point(aes(shape=loc)) + theme(aspect.ratio=1) + geom_text_repel(size=2,min.segment.length = 0,color="black") + 
#  xlab(paste0("recall (n=",CRISPR.n,")"))
#ggsave(paste0("plots/K562_CRISPR_precision_vs_recall_point_estimates_proximal_vs_distal.pdf"),width=8,height=7)

point.CRISPR.genome.df$sample <- factor(point.CRISPR.genome.df$sample, levels=point.CRISPR.genome.df$sample[order(point.CRISPR.genome.df$recall, decreasing = TRUE)])
ggplot(point.CRISPR.genome.df, aes(sample, recall)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) + ylab(paste0("recall (n=",CRISPR.n,")"))
ggsave(paste0("plots/K562_CRISPR_recall_point_estimates_bar.pdf"),width=3,height=3)

point.CRISPR.genome.df$sample <- factor(point.CRISPR.genome.df$sample, levels=point.CRISPR.genome.df$sample[order(point.CRISPR.genome.df$precision, decreasing = TRUE)])
ggplot(point.CRISPR.genome.df, aes(sample, precision)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))
ggsave(paste0("plots/K562_CRISPR_precision_point_estimates_bar.pdf"),width=3,height=3)

```


Enrichment, precision, and recall versus score threshold

```{r}

enrichment.precision.recall <- function(data, fg.data, bg.data, assay) {
  
  do.call("rbind", lapply(names(data), function(n) {
    
    gr <- data[[n]]
    scores <- seq(0,1,by=0.01)
    cnt <- sapply(scores, function(sco) sum(gr$score>=sco))
    scores <- scores[cnt>=10]
    
    gr.fg.olap <- findOverlaps(fg.data, gr, minoverlap=1)
    gr.bg.olap <- findOverlaps(bg.data, gr, minoverlap=1)
    
    do.call("rbind", lapply(scores, function(sco) {
      
      selection <- which(gr$score>=sco)
      fg.count <- length(unique(queryHits(gr.fg.olap)[which(subjectHits(gr.fg.olap) %in% selection)]))
      bg.count <- length(unique(queryHits(gr.bg.olap)[which(subjectHits(gr.bg.olap) %in% selection)]))
      
      data.frame(enrichment=(fg.count/length(fg.data))/(bg.count/length(bg.data)),
                 precision=fg.count/(fg.count+bg.count),
                 recall=fg.count/length(fg.data),
                 fg.count=fg.count,
                 bg.count=bg.count,
                 score=sco,
                 sample=n,
                 assay=assay[n])
    }))
    
  }))
}

pooled.PRIMEloci.ClinVar.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, ClinVar.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.ClinVar.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, ClinVar.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.eQTL.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, eQTL.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.eQTL.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, eQTL.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.GWAS.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, GWAS.data, SNP.data, pooled.PRIMEloci.names)
#PRIMEloci.GWAS.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, GWAS.data, SNP.data, PRIMEloci.names)

pooled.PRIMEloci.CRISPR.enrichment.recall.data <- enrichment.precision.recall(pooled.PRIMEloci, CRISPR.pos.data, CRISPR.neg.data, pooled.PRIMEloci.names)
#PRIMEloci.CRISPR.enrichment.recall.data <- enrichment.precision.recall(PRIMEloci, CRISPR.pos.data, CRISPR.neg.data, PRIMEloci.names)

```


Plot results

```{r}

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data, aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=subset(point.ClinVar.df,loc=="genome")) + xlab(paste0("recall (n=",ClinVar.n,")"))
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_enrichment_vs_recall_and_point_estimates_genome.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data, aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.eQTL.genome.df) + xlab(paste0("recall (n=",eQTL.n,")"))
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_enrichment_vs_recall_PIP_",eQTL.pip.thres,"_and_point_estimates_genome.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data, aes(recall, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.GWAS.genome.df) + xlab(paste0("recall (n=",GWAS.n,")"))
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_recall_PIP_",GWAS.pip.thres,"_and_point_estimates_genome.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data, aes(recall, precision, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1) +
  geom_point(data=point.CRISPR.genome.df) + xlab(paste0("recall (n=",CRISPR.n,")"))
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_precision_vs_recall_and_point_estimates_genome.pdf"),width=7,height=5)

```

Enrichment, precision, recall vs score

```{r}

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_enrichment_vs_score.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.ClinVar.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave("plots/K562_pooled_PRIMEloci_ClinVar_recall_vs_score.pdf",width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_enrichment_vs_score_PIP_",eQTL.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.eQTL.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_eQTL_recall_vs_score_PIP_",eQTL.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data, aes(score, enrichment, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_enrichment_vs_score_PIP_",GWAS.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.GWAS.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_GWAS_recall_vs_score_PIP_",GWAS.pip.thres,".pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data, aes(score, precision, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_precision_vs_score.pdf"),width=7,height=5)

ggplot(pooled.PRIMEloci.CRISPR.enrichment.recall.data, aes(score, recall, color=assay)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)
ggsave(paste0("plots/K562_pooled_PRIMEloci_CRISPR_recall_vs_score.pdf"),width=7,height=5)

```
