---
title: "10.1. K562 external validation"
output: html_document
date: "2024-07-01"
---

## Evaluation of predicted regulatory elements using ClinVar non-benign and non-coding variants, fine-mapped GWAS variants, fine-mapped eQTLs, and CRISPRi data.


## [0.0.] Load necessary packages
```{r warning=TRUE, message=TRUE}
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(UpSetR)
```

## [0.1.] Add custom color palettes
```{r}
col.scheme <- base::c(
    "#377eb8","#d39200","#93aa00",
    "#4daf4a","#984ea3","#ff7f01",
    "#e41b1c","#256866","#f8bc24",
    "grey","black","hotpink"
  )
```

## [1.] Read in & prepare data -------------------------

## [1.0.] Read in CDSs from txdb
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
cds <- base::unlist(GenomicFeatures::cdsBy(txdb,"gene"))
cds <- BiocGenerics::sort(cds)
```

## [1.1.] Overlap CDSs with background SNPs (9,913,974 unique autosomal SNP locations)
#### Remove SNPs overlapping with CDSs
```{r}
## Read in SNP.file
SNP.data <- rtracklayer::import.bed( # 9.991.229
   con = base::paste0(
       here::here("0.External_resources/"),
       "all.bg.SNPs.hg38.baseline.v1.1.bed.sorted"
     )
  )
SNP.data <- BiocGenerics::sort(SNP.data)
olap <- GenomicRanges::findOverlaps( ## 93.742
    query = SNP.data,
    subject = cds,
    minoverlap = 1L
  )
SNP.data$name <- NULL
SNP.data <- base::unique( # 9.914.129
    SNP.data[-S4Vectors::queryHits(olap)]
  )
GenomeInfoDb::seqlevels(txdb, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.2.] Read in promoters from txdb
```{r}
promoters <- IRanges::promoters(
    x = txdb,
    upstream = 500,
    downstream = 500
  )
```

## [1.3.] Annotate identified PRIME loci by txdb annotated promoters into proximal (overlapping with promoter) or distal
## PRIME loci, pooled replicates; evaluate with n = "K562_sld_C" # 4.049.578
```{r}
resize.width = 200
pooled.PRIME.files <- base::c(
    "K562_sld_C" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_C.bed"),
    "K562_sld_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N.bed"),
    "K562_sld_10M_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N_10M.bed")
  )

pooled.PRIME <- base::lapply(
  X = base::names(pooled.PRIME.files),
  FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIME.files[n],
          show_col_types = FALSE
        )
      score <- gr$score # # 4.049.578
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3]) # 4.049.578
      gr$score <- score
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 3.915.085
      gr <- IRanges::resize(
          x = gr,
          width = resize.width,
          fix = "center"
        )
      pc <- base::rep("distal", base::length(gr))
      olap <- GenomicRanges::findOverlaps(
          query = gr,
          subject = promoters,
          minoverlap = 1
        )
      pc[S4Vectors::queryHits(olap)] <- "proximal"
      gr$class <- pc
      return(gr)
    }
  )
base::names(pooled.PRIME) <- base::names(pooled.PRIME.files)
```

## [1.4.] Annotate identified Divergent loci by txdb annotated promoters into proximal (overlapping with promoter) or distal
```{r}
pooled.DLs <- base::readRDS(
    file = base::paste0(
      here::here("3.DivergentLoci/"),
      #"poolSub.k562.summitDLs.rmSingletons.rds"
      "poolSub.K562.DLs.FANTOM5style.rds"
    )
  )
pooled.DLs[[1]]$sample <- "K562_C"
pooled.DLs[[2]]$sample <- "K562_N"
pooled.DLs[[3]]$sample <- "K562_N_10M"
if (base::is.list(pooled.DLs)) {
  pooled.DLs <- base::do.call(
      what = "c",
      args = pooled.DLs
    )
  }
GenomeInfoDb::seqlevels(pooled.DLs, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
pooled.DLs <- pooled.DLs[pooled.DLs$sample %in% base::c("K562_C","K562_N")] # 104.105

resize.width = 200
pooled.DLs <- IRanges::resize(
    x = pooled.DLs,
    width = resize.width,
    fix = "center"
  )
dlc <- base::rep("distal",base::length(pooled.DLs))
olap <- GenomicRanges::findOverlaps(
    query = pooled.DLs,
    subject = promoters,
    minoverlap = 1
  )
dlc[S4Vectors::queryHits(olap)] <- "proximal"
pooled.DLs$class <- dlc # 104.105
```

## [1.5.] Read in, prepare and subset eQTL data
```{r}
eQTL.pip.thres = 0.5
eQTL.data <- readr::read_tsv(
      file = base::paste0(
        here::here("0.External_resources/"),
        "GTEx_30tissues_release1_pip_01_Whole_Blood.tsv"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
eQTL.data <- eQTL.data[,base::c(1,3)]
base::colnames(eQTL.data) <- base::c("variant","pip")
pos <- base::as.numeric(
  base::sapply(
      X = eQTL.data$variant,
      FUN = function(n){
          base::strsplit(n,"_")[[1]][2]
      }
    )
  ) + 1
eQTL.data <- GenomicRanges::GRanges(
    seqnames = base::as.character(
      base::sapply(
        X = eQTL.data$variant,
        FUN = function(n){
            base::strsplit(n,"_")[[1]][1]
          }
        )
      ),
    score = eQTL.data$pip,
    ranges = IRanges::IRanges(pos, pos)
  )
eQTL.data <- base::sort(eQTL.data)
GenomeInfoDb::seqlevels(eQTL.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
eQTL.data <- base::subset(eQTL.data, score >= eQTL.pip.thres)
eQTL.data$score <- NULL
```

## [1.6.] Subset eQTLs for sites outside of CDSs
```{r}
olap <- GenomicRanges::findOverlaps(
    query = eQTL.data,
    subject = cds,
    minoverlap = 1
  )
eQTL.data <- base::unique(
    eQTL.data[-S4Vectors::queryHits(olap)]
  )
eQTL.n <- base::length(eQTL.data)
```

## [1.7.] Read in, prepare and subset GWAS data
```{r}
GWAS.pip.thres = 0.2

GWAS.data <- readr::read_tsv( # 66.638
    file = base::paste0(
        here::here("0.External_resources/"),
        "UKBB_94traits_release1_pip01_clean_uniq_hg38.bed"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
base::colnames(GWAS.data) <- base::c(
    "chr","start","end",
    "name","score","strand",
    "RS_id","trait","pip"
  )
GWAS.data <- GenomicRanges::makeGRangesFromDataFrame( # 66.638
    df = GWAS.data,
    keep.extra.columns = TRUE,
    starts.in.df.are.0based = TRUE
  )
GWAS.data <- base::subset( # 30.743
    GWAS.data, pip >= GWAS.pip.thres
  )
GenomeInfoDb::seqlevels(GWAS.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.8.] Remove GWAS hits overlapping with CDSs & unwanted traits
```{r}
traits <- base::c(
    "HbA1c","Hb","RBC",
    "MCV","MCH","MCHC"
  )

olap <- GenomicRanges::findOverlaps(
    query = GWAS.data,
    subject = cds,
    minoverlap = 1
  )
GWAS.data <- GWAS.data[-S4Vectors::queryHits(olap)]
GWAS.data <- base::subset(GWAS.data, trait %in% traits)
trait.count.df <- base::do.call(
  what = "rbind",
  args = base::lapply(
    X = traits,
    FUN = function(t) {
      dat <- base::subset(GWAS.data, trait == t)
      return(
          base::as.data.frame(
            base::table(dat$trait)
          )
        )
      }
    )
  )

GWAS.data$pip <- NULL
GWAS.data$trait <- NULL
GWAS.data <- base::unique(GWAS.data) # 2.877
GWAS.n <- base::length(GWAS.data)
```

## [1.9.] Read in and subset ClinVar data for sites not overlapping with CDSs
```{r}
ClinVar.data <- readr::read_tsv(
    file = base::paste0(
        here::here("0.External_resources/"),
        #"ClinVar_GRCh38_nonbenign_noncoding_SNV.tsv"
        "clinvar/ClinVar_GRCh38_allCategories_noncoding_SNV.tsv"
      ),
    show_col_types = FALSE
  )
ClinVar.data <- GenomicRanges::makeGRangesFromDataFrame(
    df = ClinVar.data
  )
GenomeInfoDb::seqlevels(ClinVar.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
olap <- GenomicRanges::findOverlaps(
    query = ClinVar.data,
    subject = cds,
    minoverlap = 1
  )
if (length(S4Vectors::queryHits(olap))>0)
  ClinVar.data <- ClinVar.data[-S4Vectors::queryHits(olap)]
ClinVar.data <- base::sort(ClinVar.data)
ClinVar.data <- base::unique(ClinVar.data)
ClinVar.n <- base::length(ClinVar.data)
```

## [1.10.] Read in all PINTS data for point estimates
```{r}
PINTS.files <- base::c(
    "PINTS GRO/PRO-cap distal" = here::here("0.External_resources/K562_Distal_PINTS_GRO_PRO_cap.bed"),
    "PINTS GRO/PRO-cap proximal" = here::here("0.External_resources/K562_Proximal_PINTS_GRO_PRO_cap.bed")
  )

point.data <- base::lapply(
  X = base::names(PINTS.files),
  FUN = function(n){
      rtracklayer::import.bed(PINTS.files[[n]])
    }
  )
base::names(point.data) <- base::names(PINTS.files)

point.data[["PINTS GRO/PRO-cap"]] <- base::c(point.data[[1]],point.data[[2]])
```

## [1.11.] Read in all dREG data for points estimates
```{r}
dREG.files <- base::c(
    "dREG PRO-seq" = here::here("0.External_resources/GSE150625_PROseq_1M_K562_dREGpeaks_clean.bed"),
    "dREG qPRO-seq" = here::here("0.External_resources/GSE150625_qPROseq_1M_K562_dREGpeaks_clean.bed")
  )

for (n in base::names(dREG.files)) {
    gr <- rtracklayer::import.bed(con = dREG.files[n])
    GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
    olap <- GenomicRanges::findOverlaps(
        query = gr,
        subject = promoters,
        minoverlap = 1
      )
    gr.proximal <- gr[base::unique(S4Vectors::queryHits(olap))]
    gr.distal <- gr[base::setdiff(1:base::length(gr), S4Vectors::queryHits(olap))]
    point.data[[n]] <- gr
    point.data[[base::paste0(n," distal")]] <- gr.distal
    point.data[[base::paste0(n," proximal")]] <- gr.proximal
  }
```

## [1.11.] Read in all pooled PRIME files
#### see [1.3.] for øpooled.PRIME.filesø object
```{r}
PRIME.thres = 0.75
pooled.PRIME.names <- base::c(
    "K562_sld_C" = "PRIME CAGE cells pooled",
    "K562_sld_N" = "PRIME CAGE nuclei pooled",
    "K562_sld_10M_N" = "PRIME CAGE nuclei 10M pooled"
  )

for (n in base::names(pooled.PRIME.files)) {
    gr <- readr::read_tsv(
        file = pooled.PRIME.files[n],
        show_col_types = FALSE
      )
    gr <- base::subset(gr, score >= PRIME.thres)
    gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3])
    GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
    olap <- IRanges::findOverlaps(
        query = gr,
        subject = promoters,
        minoverlap = 1
      )
    gr.proximal <- gr[base::unique(S4Vectors::queryHits(olap))]
    gr.distal <- gr[base::setdiff(1:base::length(gr), S4Vectors::queryHits(olap))]
    point.data[[pooled.PRIME.names[n]]] <- gr
    point.data[[base::paste0(pooled.PRIME.names[n]," distal")]] <- gr.distal
    point.data[[base::paste0(pooled.PRIME.names[n]," proximal")]] <- gr.proximal
  }
```

## [1.11.] Read in cCRE data for point estimates
#### cCRE.file <- "../0.External_resources/ENCFF464BRU.bed"
```{r}
cCRE <- readr::read_tsv(
    file = base::paste0(
        here::here("0.External_resources/"),
        "ENCFF414OGC_ENCFF806YEZ_ENCFF849TDM_ENCFF736UDR.bed"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
df <- cCRE[,1:3]
base::colnames(df) <- base::c("chrom","start","end")
cCRE.data <- GenomicRanges::makeGRangesFromDataFrame(df)
cCRE.data$class <- cCRE[,10, drop = TRUE]

for (n in base::c("dELS","PLS")) {
   x <- base::subset(cCRE.data, class == n)
   x$class <- NULL
   GenomeInfoDb::seqlevels(x,pruning.mode="coarse") <- GenomeInfoDb::seqlevels(SNP.data)
   olap <- GenomicRanges::findOverlaps(
       query = x,
       subject = promoters,
       minoverlap = 1
     )
   x.proximal <- x[base::unique(S4Vectors::queryHits(olap))]
   x.distal <- x[base::setdiff(1:length(x), S4Vectors::queryHits(olap))]
   point.data[[paste0("ENCODE cCRE ",n," distal")]] <- x.distal
   point.data[[paste0("ENCODE cCRE ",n," proximal")]] <- x.proximal
  }

point.data[["ENCODE cCRE"]] <- base::c(
    point.data[["ENCODE cCRE dELS distal"]],
    point.data[["ENCODE cCRE dELS proximal"]],
    point.data[["ENCODE cCRE PLS distal"]],
    point.data[["ENCODE cCRE PLS proximal"]]
  )
```

## [1.12.] Read in DHS data for point estimates for point estimates
#### seqlevel adjustment step is missing, creating a warning in IRanges::subsetByOverlaps of [1.13.]
```{r}
DHS <- rtracklayer::import.bed(
  con = base::paste0(
      here::here("0.External_resources/"),
      "E123-DNase.macs2.hg38.narrowPeak"
    )
  )
olap <- GenomicRanges::findOverlaps(
    query = DHS,
    subject = promoters,
    minoverlap = 1
  )
DHS.proximal <- DHS[base::unique(S4Vectors::queryHits(olap))]
DHS.distal <- DHS[base::setdiff(1:length(DHS), S4Vectors::queryHits(olap))]

point.data[["DHS"]] <- DHS
point.data[["DHS distal"]] <- DHS.distal
point.data[["DHS proximal"]] <- DHS.proximal
```

## [1.13.] Read in all H3K27ac data and combine with DHSs for point estimates
#### seqlevel adjustment step is missing, creating a warning in IRanges::subsetByOverlaps
```{r}
k27ac <- utils::read.table(
  file = base::paste0(
      here::here("0.External_resources/"),
      "ENCFF038DDS.bed"
    )
  )
base::colnames(k27ac) <- base::c(
    "chr","start","end",
    "name","score","strand",
    "signal_value","pvalue",
    "qvalue","peak"
  )
k27ac <- GenomicRanges::GRanges(k27ac)
GenomeInfoDb::seqlevels(k27ac,pruning.mode="coarse") <- GenomeInfoDb::seqlevels(SNP.data)
k27ac.dhs <- IRanges::subsetByOverlaps(
    x = DHS,
    ranges = k27ac
  )
GenomicRanges::start(k27ac) <- GenomicRanges::start(k27ac) + k27ac$peak
GenomicRanges::end(k27ac) <- GenomicRanges::start(k27ac) + 1

point.data[["K27ac DHS"]] <- k27ac.dhs
point.data[["K27ac"]] <- k27ac
```

## [1.13.] Read in FANTOM5 data for point estimates
```{r}
FANTOM5.id = "CNhs12334"

header <- base::colnames(
  readr::read_tsv(
      file = base::paste0(
          here::here("0.External_resources/"),
          "F5.hg38.enhancers.expression.matrix.gz"
        ),
      n_max = 1,
      show_col_types = FALSE
    )
  )
FANTOM5.data <- readr::read_tsv(
    file = base::paste0(
        here::here("0.External_resources/"),
        "F5.hg38.enhancers.expression.matrix.gz"
      ),
    skip = 0,
    col_names = TRUE,
    show_col_types = FALSE
  )
FANTOM5.data <- FANTOM5.data[,base::c(1,base::match(FANTOM5.id,header)+1)]

FANTOM5.enhancers <- IRanges::flank(
  x = CAGEfightR::swapRanges(
      rtracklayer::import.bed(
          con = base::paste0(
            here::here("0.External_resources/"),
            "F5.hg38.enhancers.bed.gz"
          )
        )
      ),
    width = 200,
    both = TRUE
  )
FANTOM5.enhancers <- FANTOM5.enhancers[
    FANTOM5.enhancers$name %in% tibble::deframe(FANTOM5.data[FANTOM5.data[2]>0,1])
  ]
point.data[["FANTOM5 enhancers"]] <- FANTOM5.enhancers
```

## [1.14.] Read in noise estimates
```{r}
pooled.noise.thresholds <- readr::read_csv(
    file = base::paste0(
        here::here("9.Noise/"),
        "K562.poolSub.divnoise.csv"
      ),
    col_names = TRUE,
    show_col_types = FALSE
  )
pooled.noise.thresholds <- pooled.noise.thresholds[
    pooled.noise.thresholds[,1]=="95%",
    base::c("K562_C","K562_N","K562_C","K562_N")
  ]
```

## [1.15.] Subset DLs by library-specific noise estimates 
#### see [1.4.] for pooled.DLs object
```{r}
pooled.DL.names <- base::c(
    "K562_C" = "Divergent loci CAGE cells",
    "K562_N" = "Divergent loci CAGE nuclei"
  )

for (sample in base::c("K562_C","K562_N")) {
    gr <- pooled.DLs[pooled.DLs$sample == sample]
    gr <- gr[base::which(gr$score > tibble::deframe(pooled.noise.thresholds[,sample]))]
    point.data[[pooled.DL.names[sample]]] <- gr
  
    gr <- pooled.DLs[pooled.DLs$sample == sample & pooled.DLs$class == "distal"]
    gr <- gr[base::which(gr$score > tibble::deframe(pooled.noise.thresholds[,sample]))]
    point.data[[base::paste0(pooled.DL.names[sample]," distal")]] <- gr
  
    gr <- pooled.DLs[pooled.DLs$sample==sample & pooled.DLs$class == "proximal"]
    gr <- gr[base::which(gr$score > tibble::deframe(pooled.noise.thresholds[,sample]))]
    point.data[[base::paste0(pooled.DL.names[sample]," proximal")]] <- gr
  }
```

## [1.16.] Resize all sites across all datasets
```{r}
resize.width = 200
for (n in base::names(point.data)) {
    point.data[[n]] <- base::sort(point.data[[n]])
    GenomeInfoDb::seqlevels(point.data[[n]],pruning.mode="coarse") <- GenomeInfoDb::seqlevels(SNP.data)
    point.data[[n]] <- IRanges::resize(
        x = point.data[[n]],
        width = resize.width,
        fix = "center"
      )
    point.data[[n]] <- IRanges::reduce(
        x = point.data[[n]],
        min.gapwidth = 0
      )
  }
```

## [1.17.] Rename datasets
```{r}
point.data.assay <- base::names(point.data)
point.data.loc <- base::rep("genome", base::length(point.data.assay))
point.data.loc[base::grep("distal", point.data.assay)] <- "distal"
point.data.loc[base::grep("proximal", point.data.assay)] <- "proximal"
point.data.loc[base::which(point.data.assay == "FANTOM5 enhancers")] <- "distal"
point.data.loc[base::which(point.data.assay == "ENCODE cCRE dELS")] <- "distal"
point.data.loc[base::which(point.data.assay == "ENCODE cCRE PLS")] <- "proximal"
point.data.assay <- base::gsub(
  " distal", "", base::gsub(
    " proximal", "", base::gsub(
      " PLS", "", base::gsub(
        " dELS", "", point.data.assay
        )
      )
    )
  )
point.data.assay <- base::gsub(
  " distal", "", base::gsub(
    " proximal", "", point.data.assay
    )
  )
base::names(point.data.assay) <- base::names(point.data.loc) <- base::names(point.data)
```

## [1.18.] Read in CRISPR data
```{r}
CRISPR.pos.data <- base::readRDS(
  file = base::paste0(
      here::here("5.EPCrisprBenchmark/"),
      "epc_pos_gr.rds"
    )
  )
CRISPR.neg.data <- base::readRDS(
  file = base::paste0(
      here::here("5.EPCrisprBenchmark/"),
      "epc_neg_gr.rds"
    )
  )
CRISPR.n <- base::length(CRISPR.pos.data)
```

## [1.19.] Read in DC-TAPseq data
#### Add dc-tap-seq validations
```{r}
dc_tap <- readr::read_delim(
    file = base::paste0(
        here::here("0.External_resources/PL_dcTapSeq/"),
        "media-3.tsv"
      ),
    delim = "\t"
  )

dc_tap_hg38 <- dc_tap[,base::c(2:4,9:80)]
base::colnames(dc_tap_hg38)[1:3] <- base::c("chr","start","end")
dc_tap_hg38 <- GenomicRanges::makeGRangesFromDataFrame(
    df = dc_tap_hg38,
    keep.extra.columns = TRUE
  )

dc.tap.enhancers.pos <- dc_tap_hg38[
    dc_tap_hg38$significant == TRUE & dc_tap_hg38$cell_type == "K562",
  ]
dc.tap.enhancers.neg <- dc_tap_hg38[
    dc_tap_hg38$significant == FALSE & dc_tap_hg38$cell_type == "K562",
  ]

S4Vectors::mcols(dc.tap.enhancers.pos) <- NULL 
S4Vectors::mcols(dc.tap.enhancers.neg) <- NULL

dc.tap.enhancers.pos <- base::unique(dc.tap.enhancers.pos) # 258 -> 223
dc.tap.enhancers.neg <- base::unique(dc.tap.enhancers.neg) # 7217 -> 923

dc.tap.n <- base::length(dc.tap.enhancers.pos)

# For some reason a lot of them overlap
dc.tap.enhancers.neg <- IRanges::subsetByOverlaps( # 923 -> 725
    dc.tap.enhancers.neg,
    dc.tap.enhancers.pos,
    invert = TRUE
  )
```

# [2.] Calculate precision, enrichment and recall -------------------------

## [2.0.] Helper function for preparing enrichment-recall curves
```{r}
point.enrichment.precision.recall <- function(data, fg.data, bg.data, assay, location) {
  
    base::do.call(
      what = "rbind",
      args = base::lapply(
        X = base::names(data),
        FUN = function(n) {
    
          fg.olap <- GenomicRanges::findOverlaps(
              query = fg.data,
              subject = data[[n]],
              minoverlap = 1
            )
          fg.count <- base::length(
            base::unique(
                S4Vectors::queryHits(fg.olap)
              )
            )
          region.count <- base::length(
            base::unique(
                S4Vectors::subjectHits(fg.olap)
              )
            )
          bg.olap <- GenomicRanges::findOverlaps(
              query = bg.data,
              subject = data[[n]],
              minoverlap = 1
            )
          bg.count <- base::length(
            base::unique(
                S4Vectors::queryHits(bg.olap)
              )
            )
          return(
            base::data.frame(
              enrichment = (fg.count/base::length(fg.data)) / (bg.count/base::length(bg.data)),
              precision = fg.count/(fg.count + bg.count),
              recall = fg.count/base::length(fg.data),
              fg.count = fg.count,
              bg.count = bg.count,
              region.count = region.count,
              sample = n,
              assay = assay[n],
              loc = location[n]
            )
          )
        }
      )
    )
  }
```

## [2.1.] Run helper function across all data modality combinations
#### evaluate with n = "DHS proximal"
```{r}
point.ClinVar.df <- .GlobalEnv$point.enrichment.precision.recall(
    data = point.data,
    fg.data = ClinVar.data,
    bg.data = SNP.data,
    assay = point.data.assay,
    location = point.data.loc
  )

point.eQTL.df <- .GlobalEnv$point.enrichment.precision.recall(
    data = point.data,
    fg.data = eQTL.data,
    bg.data = SNP.data,
    assay = point.data.assay,
    location = point.data.loc
  )

point.GWAS.df <- .GlobalEnv$point.enrichment.precision.recall(
    data = point.data,
    fg.data = GWAS.data,
    bg.data = SNP.data,
    assay = point.data.assay,
    location = point.data.loc
  )

point.CRISPR.df <- .GlobalEnv$point.enrichment.precision.recall(
    data = point.data,
    fg.data = CRISPR.pos.data,
    bg.data = CRISPR.neg.data,
    assay = point.data.assay,
    location = point.data.loc
  )

point.dctap.df <- .GlobalEnv$point.enrichment.precision.recall(
    data = point.data,
    fg.data = dc.tap.enhancers.pos,
    bg.data = dc.tap.enhancers.neg,
    assay = point.data.assay,
    location = point.data.loc
  )
```

## [3.] Visualize ClinVar results -------------------------

## [3.0.] Prepare ClinVar plotting
```{r}
point.ClinVar.genome.df <- base::subset(point.ClinVar.df, loc == "genome")
point.ClinVar.non.genome.df <- base::subset(point.ClinVar.df, loc != "genome")

col.scheme <- stats::setNames(
  object = col.scheme,
  nm = base::c(
      "DHS","Divergent loci CAGE cells","Divergent loci CAGE nuclei",
      "dREG PRO-seq","dREG qPRO-seq","ENCODE cCRE",
      "PINTS GRO/PRO-cap","PRIME CAGE cells pooled","PRIME CAGE nuclei pooled",
      "K27ac", "K27ac DHS","PRIME CAGE nuclei 10M pooled"
    )
  )
```

## [3.1.] Visualization
```{r}
ggplot2::ggplot(
    data = point.ClinVar.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point() +
  theme_bw() +
  theme(
      aspect.ratio = 1
    ) +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
    base::paste0("recall (n=",ClinVar.n,")")) +
  scale_color_manual(
      values = col.scheme
    )
```

## [3.2.] Printint plots
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_ClinVar_enrichment_vs_recall_point_estimates.pdf"
      ),
    width = 8,
    height = 7,
    device = "pdf"
  )
```

## [3.3.] Supplementary Figure
```{r}
ggplot2::ggplot(
    data = point.ClinVar.non.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point(
    mapping = aes(
        shape = loc
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    ) +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
    base::paste0("recall (n=",ClinVar.n,")")) +
  scale_color_manual(
      values = col.scheme
    )
```

## [3.4.] Printint plots
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_ClinVar_enrichment_vs_recall_point_estimates_proximal_vs_distal.pdf"
      ),
    width = 8,
    height = 7,
    device = "pdf"
  )
```

## [3.5.] Prepare Figure 1 E bar plots
```{r}
point.ClinVar.genome.df$sample <- base::factor(
  x = point.ClinVar.genome.df$sample,
  levels = point.ClinVar.genome.df$sample[
      base::order(point.ClinVar.genome.df$recall, decreasing = FALSE)
    ]
  )
```

## [3.6.] Visualize Figure 1 E bar plots
```{r}
ggplot2::ggplot(
    data = point.ClinVar.genome.df,
    mapping = aes(
        x = recall,
        y = sample,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    # axis.text.x = element_text(
    #     angle = 45,
    #     vjust = 1,
    #     hjust = 1
    #   )
    axis.text.y = element_blank()
    ) +
  ylab(
    base::paste0("recall (n=",ClinVar.n,")")) +
  scale_fill_manual(
      values = col.scheme
    )
```

## [3.7.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_ClinVar_recall_point_estimates_bar.pdf"
      ),
    width = 4,
    height = 3,
    device = "pdf"
  )
```

## [3.8.] Prepare Figure 1 E
```{r}
point.ClinVar.genome.df$sample <- base::factor(
  x = point.ClinVar.genome.df$sample,
  levels = point.ClinVar.genome.df$sample[
      order(point.ClinVar.genome.df$enrichment, decreasing = FALSE)
    ]
  )
```

## [3.9.] Plot Figure 1 E
```{r}
ggplot2::ggplot(
    data = point.ClinVar.genome.df,
    mapping = aes(
        x = enrichment,
        y = sample,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    # axis.text.x = element_text(
    #     angle = 45,
    #     vjust = 1,
    #     hjust = 1
    #   )
    axis.text.y = element_blank()
    ) +
  ylab(
    base::paste0("recall (n=",ClinVar.n,")")) +
  scale_fill_manual(
      values = col.scheme
    )
```

## [3.10.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_ClinVar_enrichment_point_estimates_bar.pdf"
      ),
    width = 4,
    height = 3,
    device = "pdf"
  )
```

## [3.11] Prepare bar plots for point estimates
#### Create a new column for scaled enrichment
```{r}
scaling_factor <- base::max(point.ClinVar.genome.df$recall) / base::max(point.ClinVar.genome.df$enrichment)

point.ClinVar.genome.df$enrichment_scaled <- point.ClinVar.genome.df$enrichment * scaling_factor

point.ClinVar.genome.df_long <- point.ClinVar.genome.df %>%
  tidyr::pivot_longer(
      cols = base::c("recall", "enrichment_scaled"),
      names_to = "metric",
      values_to = "value"
    ) %>%
  mutate(
    metric = base::factor(
        x = metric,
        levels = base::c("recall","enrichment_scaled")
      )
    )
```

## [3.12.] Plot using ggplot with adjusted spacing
```{r}
ggplot2::ggplot(
      data = point.ClinVar.genome.df_long
    ) +
  geom_col(
      mapping = aes(
          x = sample,
          y = value,
          fill = sample,
          group = metric,
          alpha = metric  
        ),
      position = position_dodge(width = 0.6),  
      color = "black",
      width = 0.6  
    ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled enrichment
  scale_y_continuous(
    name = "Recall",
    limits = base::c(0, base::max(point.ClinVar.genome.df$recall) * 1.1),
    sec.axis = ggplot2::sec_axis(
        ~ . / scaling_factor,  
        name = "Enrichment"
      )
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  scale_alpha_manual(
      values = base::c("recall" = 1, "enrichment_scaled" = 0.5)
    ) +  # Set alpha levels for recall and enrichment
  theme_bw() +
  theme(
      aspect.ratio = 1/2,
      axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
      axis.title.x = element_blank(),
      axis.title.y.left = element_text(
          color = "black"
        ),
      axis.title.y.right = element_text(
          color = "black"
        ),
      legend.title = element_blank(),
      legend.position = "bottom"
    ) +
  labs(x = "Sample", fill = "Sample") +
  theme(legend.position = "none")
```

## [3.13.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_ClinVar_enrichment_recall_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [4.] Visualize ClinVar results -------------------------

## [4.1.] Plot eQTL point estimate results
```{r}
point.eQTL.genome.df <- subset(point.eQTL.df, loc == "genome")
point.eQTL.non.genome.df <- subset(point.eQTL.df, loc != "genome")
```

## [4.2.] Visualization
```{r}
ggplot2::ggplot(
    data = point.eQTL.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point() +
  theme_bw() +
  theme(
      aspect.ratio = 1
    ) +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
      base::paste0("recall (n=",eQTL.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    )
```

## [4.3.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_eQTL_enrichment_vs_recall_PIP_",
        eQTL.pip.thres,
        "_point_estimates.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [4.4.] Supplementary Figure
#### <<important>>
```{r}
ggplot2::ggplot(
    data = point.eQTL.non.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point(
    mapping = aes(
        shape = loc
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    ) +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) +
  xlab(
      base::paste0("recall (n=",eQTL.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    )
```

## [4.5.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_eQTL_enrichment_vs_recall_PIP_",
        eQTL.pip.thres,
        "_point_estimates_proximal_vs_distal.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [4.6.] Prepare Figure 1 F
```{r}
point.eQTL.genome.df$sample <- base::factor(
  x = point.eQTL.genome.df$sample,
  levels = point.eQTL.genome.df$sample[
      base::order(point.eQTL.genome.df$recall, decreasing = FALSE)
    ]
  )
```

## [4.7.] Plot Figure 1 F
```{r}
ggplot2::ggplot(
  data = point.eQTL.genome.df,
  mapping = aes(
      y = sample,
      x = recall,
      fill = sample
    )
  ) +
  geom_bar(
      stat = "identity"
    ) + 
  ylab(
      base::paste0("recall (n=",eQTL.n,")")
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
    axis.text.y = element_blank()
    # axis.text.x = element_text(
    #     angle = 45,
    #     vjust = 1,
    #     hjust = 1
    #   )
    )
```

## [4.8.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_eQTL_recall_PIP_",
        eQTL.pip.thres,
        "_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [4.9.] Prepae Figure 1 F
```{r}
point.eQTL.genome.df$sample <- base::factor(
  x = point.eQTL.genome.df$sample,
  levels = point.eQTL.genome.df$sample[
      base::order(point.eQTL.genome.df$enrichment, decreasing = FALSE)
    ]
  )
```

## [4.10.] Plot Figure 1F
```{r}
ggplot2::ggplot(
    data = point.eQTL.genome.df,
    mapping = aes(
        x = enrichment,
        y = sample,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  theme_bw() +
  theme(
      # axis.text.x = element_text(
      #     angle = 90,
      #     vjust = 1,
      #     hjust = 1
      #   )
      axis.text.y = element_blank(),
      axis.title.y = element_blank()
    ) +
  scale_fill_manual(
      values = col.scheme
    )
```

## [4.11.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_eQTL_enrichment_PIP_",
        eQTL.pip.thres,
        "_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [4.12.] Prepare bar plots for point estimates
#### Create a new column for scaled enrichment
```{r}
scaling_factor <- base::max(point.eQTL.genome.df$recall) / base::max(point.eQTL.genome.df$enrichment)

point.eQTL.genome.df$enrichment_scaled <- point.eQTL.genome.df$enrichment * scaling_factor

point.eQTL.genome.df_long <- point.eQTL.genome.df %>%
  tidyr::pivot_longer(
      cols = base::c("recall", "enrichment_scaled"),
      names_to = "metric",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    metric = base::factor(
        x = metric,
        levels = base::c("recall","enrichment_scaled")
      )
    )
```

## [4.13.] Plot using ggplot with adjusted spacing
```{r}
ggplot2::ggplot(
      data = point.eQTL.genome.df_long
    ) +
  geom_col(
      mapping = aes(
          x = sample,
          y = value,
          fill = sample,
          group = metric,
          alpha = metric  
        ),
      position = position_dodge(width = 0.6),  
      color = "black",
      width = 0.6  
    ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled enrichment
  scale_y_continuous(
      name = "Recall",
      limits = base::c(0, base::max(point.eQTL.genome.df$recall) * 1.1),
      sec.axis = ggplot2::sec_axis(
        transform = ~ . / scaling_factor,
        name = "Enrichment"
      )
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  scale_alpha_manual(
      values = base::c("recall" = 1, "enrichment_scaled" = 0.5)
    ) + 
  labs(x = "Sample", fill = "Sample") +
  # Set alpha levels for recall and enrichment
  theme_bw() +
  theme(
      aspect.ratio = 1/2,
      axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
      axis.title.x = element_blank(),
      axis.title.y.left = element_text(
          color = "black"
        ),
      axis.title.y.right = element_text(
          color = "black"
        ),
      legend.title = element_blank(),
      legend.position = "none"
    )
```

## [4.14.] Plot using ggplot with adjusted spacing
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_eQTL_enrichment_recall_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [5.] Visualize GWAS results -------------------------

## [5.0.] Plot GWAS point estimate results
```{r}
point.GWAS.genome.df <- base::subset(point.GWAS.df, loc == "genome")
point.GWAS.non.genome.df <- base::subset(point.GWAS.df, loc != "genome")
```

## [5.1.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point() +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [5.2.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
      here::here("10.External_validation/plots/"),
        "K562_GWAS_enrichment_vs_recall_PIP_",
        GWAS.pip.thres,
        "_point_estimates.pdf"
      ),
    width = 6,
    height = 5,
    device = "pdf"
  )
```

## [5.3.] Print plot
```{r}
ggplot2::ggplot(
    data = point.GWAS.non.genome.df,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay,
        label = sample
      )
    ) + 
  geom_point(
    mapping = aes(
        shape = loc
      )
    ) +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [5.4.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_GWAS_enrichment_vs_recall_PIP_",
        GWAS.pip.thres,
        "_point_estimates_proximal_vs_distal.pdf"
      ),
    width = 6,
    height = 5,
    device = "pdf"
  )
```

## [5.5.] Prepare data for plotting
```{r}
point.GWAS.genome.df$sample <- base::factor(
  x = point.GWAS.genome.df$sample,
  levels = point.GWAS.genome.df$sample[
      base::order(point.GWAS.genome.df$recall, decreasing = FALSE)
    ]
  )
```

## [5.6.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.genome.df,
    mapping = aes(
        x = recall,
        y = sample,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  theme_bw() +
  theme(
      # axis.text.x = element_text(
      #     angle = 90,
      #     vjust = 1,
      #     hjust = 1
      #   )
      axis.text.y = element_blank()
    ) +
  ylab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_fill_manual(
      values = col.scheme
    )
```

## [5.7.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_GWAS_recall_PIP_",
        GWAS.pip.thres,
        "_point_estimates_bar.pdf"
      ),
    width = 4,
    height = 3,
    device = "pdf"
  )
```

## [5.8.] Prepare plotting
```{r}
point.GWAS.genome.df$sample <- base::factor(
  x = point.GWAS.genome.df$sample,
  levels = point.GWAS.genome.df$sample[
      base::order(point.GWAS.genome.df$enrichment, decreasing = FALSE)
    ]
  )
```

## [5.9.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.genome.df,
    mapping = aes(
        x = enrichment,
        y = sample,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  theme_bw() +
  theme(
      # axis.text.x = element_text(
      #     angle = 90,
      #     vjust = 1,
      #     hjust = 1
      #   )
      axis.text.y = element_blank()
    ) +
  scale_fill_manual(
      values = col.scheme
    )
```

## [5.10.] Plot data
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_GWAS_enrichment_PIP_",
        GWAS.pip.thres,
        "_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 4,
    device = "pdf"
  )
```

## [6.] Visualize CRISPR results -------------------------

## [6.0.] Prepare data for plotting
```{r}
point.CRISPR.genome.df <- base::subset(point.CRISPR.df, loc == "genome")
point.CRISPR.non.genome.df <- base::subset(point.CRISPR.df, loc != "genome")
```

## [6.1.] Plot CRISPR point estimate results
```{r}
ggplot2::ggplot(
    data = point.CRISPR.genome.df,
    mapping = aes(
        x = recall,
        y = precision,
        color = assay,
        label = sample
      )
    ) + 
  geom_point() +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) + 
  xlab(
      base::paste0("recall (n=",CRISPR.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() + 
  theme(
      aspect.ratio = 1
    )
```

## [6.2.] Print plot
```{r}
ggplot2::ggsave(
    base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_CRISPR_precision_vs_recall_point_estimates.pdf"
      ),
    width = 8,
    height = 7,
    device = "pdf"
  )
```

## [6.3.] Plot data
```{r}
ggplot2::ggplot(
    data = point.CRISPR.non.genome.df,
    mapping = aes(
        x = recall,
        y = precision,
        color = assay,
        label = sample
      )
    ) + 
  geom_point() +
  ggrepel::geom_text_repel(
      size = 2,
      min.segment.length = 0,
      color = "black"
    ) +
  facet_wrap(
      facets = ~loc,
      scales = "free"
    ) +
  xlab(
      base::paste0("recall (n=",CRISPR.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) + 
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [6.4.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_CRISPR_precision_vs_recall_point_estimates_proximal_vs_distal.pdf"
      ),
    width = 6,
    height = 3,
    device = "pdf"
  )
```

## [6.5.] Prepare plotting
```{r}
point.CRISPR.genome.df$sample <- base::factor(
  x = point.CRISPR.genome.df$sample,
  levels = point.CRISPR.genome.df$sample[
      base::order(point.CRISPR.genome.df$recall, decreasing = TRUE)
    ]
  )
```

## [6.6.] Plot data
```{r}
ggplot2::ggplot(
    data = point.CRISPR.genome.df,
    mapping = aes(
        x = sample,
        y = recall,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) + 
  ylab(
      base::paste0("recall (n=",CRISPR.n,")")
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      )
    )
```

## [6.7.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_CRISPR_recall_point_estimates_bar.pdf"
      ),
    width = 3,
    height = 3,
    device = "pdf"
  )
```

## [6.8.] Prepare plotting
```{r}
point.CRISPR.genome.df$sample <- base::factor(
  x = point.CRISPR.genome.df$sample,
  levels = point.CRISPR.genome.df$sample[
      base::order(point.CRISPR.genome.df$precision, decreasing = TRUE)
    ]
  )
```

## [6.9.] Plot data
```{r}
ggplot2::ggplot(
  data = point.CRISPR.genome.df,
  mapping = aes(
      x = sample,
      y = precision,
      fill = sample
    )
  ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
        angle = 45,
        vjust = 1,
        hjust = 1
      )
    )
```

## [6.10.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_CRISPR_precision_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [6.11.] Prepare bar plots for point estimates
#### Create a new column for scaled precision
```{r}
scaling_factor <- base::max(point.CRISPR.genome.df$recall) / base::max(point.CRISPR.genome.df$precision)

point.CRISPR.genome.df$precision_scaled <- point.CRISPR.genome.df$precision * scaling_factor

point.CRISPR.genome.df_long <- point.CRISPR.genome.df %>%
  tidyr::pivot_longer(
      cols = base::c("recall", "precision_scaled"),
      names_to = "metric",
      values_to = "value"
    ) %>%
  dplyr::mutate(
    metric = base::factor(
        x = metric,
        levels = base::c("recall","precision_scaled")
      )
    )
```

## [6.12.] Plot data
```{r}
ggplot2::ggplot(
      data = point.CRISPR.genome.df_long
    ) +
  geom_col(
      mapping = aes(
          x = sample,
          y = value,
          fill = sample,
          group = metric,
          alpha = metric  
        ),
      position = position_dodge(width = 0.6),  
      color = "black",
      width = 0.6  
    ) +
  # Primary Y-axis for recall and secondary Y-axis for scaled precision
  scale_y_continuous(
      name = "Recall",
      limits = base::c(0, base::max(point.CRISPR.genome.df$recall) * 1.1),
      sec.axis = ggplot2::sec_axis(
        transform = ~ . / scaling_factor,  
        name = "Precision"
      )
    ) +
  scale_fill_manual(
      values = col.scheme
    ) +
  # Set alpha levels for recall and precision
  scale_alpha_manual(
      values = base::c("recall" = 1, "precision_scaled" = 0.5)
    ) + 
  theme_bw() +
  theme(
      aspect.ratio = 1/2,
      axis.text.x = element_text(
          angle = 90,
          vjust = 0.5,
          hjust = 1
        ),
      axis.title.y.left = element_text(
          color = "black"
        ),
      axis.title.y.right = element_text(
          color = "black"
        ),
      legend.title = element_blank(),
      legend.position = "bottom"
    ) +
  labs(
      x = "Sample", fill = "Sample"
    ) +
  theme(
      legend.position = "none"
    )
```

## [6.13.] Plot data
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_CRISPR_precision_recall_point_estimates_bar.pdf"
      ),
    width = 5,
    height = 3,
    device = "pdf"
  )
```

## [7.] Calculate enrichment, precision, and recall versus score threshold -------------------------

## [7.0.] Helper function for preparing enrichment-recall curves
```{r}
enrichment.precision.recall <- function(data, fg.data, bg.data, assay) {
  
  base::do.call(
    what = "rbind",
    args = base::lapply(
      X = base::names(data),
      FUN = function(n) {
    
        gr <- data[[n]]
        scores <- base::seq(0, 1, by = 0.01)
        cnt <- base::sapply(
          X = scores,
          FUN = function(sco){
              base::sum(gr$score>=sco)
            }
          )
        scores <- scores[cnt>=10]
        
        gr.fg.olap <- GenomicRanges::findOverlaps(
            query = fg.data,
            subject = gr,
            minoverlap = 1
          )
        gr.bg.olap <- GenomicRanges::findOverlaps(
            query = bg.data,
            subject = gr,
            minoverlap = 1
          )
    
        base::do.call(
          what = "rbind",
          args = base::lapply(
            X = scores,
            FUN = function(sco){
      
              selection <- base::which(gr$score>=sco)
              fg.count <- base::length(
                base::unique(
                  S4Vectors::queryHits(gr.fg.olap)[
                      base::which(S4Vectors::subjectHits(gr.fg.olap) %in% selection)
                    ]
                  )
                )
              bg.count <- base::length(
                base::unique(
                  S4Vectors::queryHits(gr.bg.olap)[
                      base::which(S4Vectors::subjectHits(gr.bg.olap) %in% selection)
                    ]
                  )
                )
              
              base::data.frame(
                    enrichment = (fg.count/length(fg.data))/(bg.count/length(bg.data)),
                    precision = fg.count/(fg.count+bg.count),
                    recall = fg.count/length(fg.data),
                    fg.count = fg.count,
                    bg.count = bg.count,
                    score = sco,
                    sample = n,
                    assay = assay[n]
                )
              }
            )
          )
        }
      )
    )
  }
```

## [7.1.] Apply helper function
```{r}
pooled.PRIME.ClinVar.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME, # PRIME
    fg.data = ClinVar.data,
    bg.data = SNP.data,
    assay =  pooled.PRIME.names # PRIME.names
  )

pooled.PRIME.eQTL.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME, # PRIME
    fg.data = eQTL.data,
    bg.data = SNP.data,
    assay = pooled.PRIME.names # PRIME.names
  )

pooled.PRIME.GWAS.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME, # PRIME
    fg.data = GWAS.data,
    bg.data = SNP.data,
    assay = pooled.PRIME.names # PRIME.names
  )

pooled.PRIME.CRISPR.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME, # PRIME
    fg.data = CRISPR.pos.data,
    bg.data = CRISPR.neg.data,
    assay = pooled.PRIME.names # PRIME.names
  )

pooled.PRIME.dctap.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME,
    fg.data = dc.tap.enhancers.pos,
    bg.data = dc.tap.enhancers.neg,
    assay = pooled.PRIME.names # PRIME.names
  )
```

## [8.] Visualize CLinVar, eQTL & CRISPR results -------------------------

## [8.0.] Plot results
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.ClinVar.enrichment.recall.data[
        pooled.PRIME.CRISPR.enrichment.recall.data$recall > 0.01,
      ],
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = base::subset(point.ClinVar.df, loc == "genome")
    ) +
  xlab(
      base::paste0("recall (n=",ClinVar.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [8.1.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_ClinVar_enrichment_vs_recall_and_point_estimates_genome.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.2.] Print plot
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.eQTL.enrichment.recall.data[
        pooled.PRIME.CRISPR.enrichment.recall.data$recall > 0.01,
      ],
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = point.eQTL.genome.df
    ) +
  xlab(
      base::paste0("recall (n=",eQTL.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [8.3.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_eQTL_enrichment_vs_recall_PIP_",
        eQTL.pip.thres,
        "_and_point_estimates_genome.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.4.] Plot data
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.GWAS.enrichment.recall.data[
        pooled.PRIME.CRISPR.enrichment.recall.data$recall > 0.01,
      ],
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = point.GWAS.genome.df
    ) +
  xlab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [8.5.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_GWAS_enrichment_vs_recall_PIP_",
        GWAS.pip.thres,
        "_and_point_estimates_genome.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.6.] Plot data
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.CRISPR.enrichment.recall.data[
        pooled.PRIME.CRISPR.enrichment.recall.data$recall > 0.01,
      ],
    mapping = aes(
        x = recall,
        y = precision,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = point.CRISPR.genome.df
    ) +
  xlab(
      base::paste0("recall (n=",CRISPR.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [8.7.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_CRISPR_precision_vs_recall_and_point_estimates_genome.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.8.] Plot data
```{r}
point.dctap.genome.df <- base::subset(
    x = point.dctap.df, loc == "genome"
  )
ggplot2::ggplot(
    data = pooled.PRIME.dctap.enrichment.recall.data,
    mapping = aes(
        x = recall,
        y = precision,
        color = assay
      )
    ) +
  geom_line(
      mapping = aes(
          group = sample
        ),
      size = 1
    ) +
  geom_point(
      data = point.dctap.genome.df,
      size = 3
    ) +
  xlab(
    base::paste0(
        "recall (n=",dc.tap.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```
## [8.9.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/"),
        "K562_pooled_PRIME_dctap_precision_vs_recall_and_point_estimates_genome.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.10.] Prepare data for running distal-proximal analysis again
#### Specifically for distal
```{r}
pooled.PRIME.distal <- base::lapply(
  X = pooled.PRIME,
  FUN = function(gr){
      gr[gr$class == "distal"]
    }
  )
pooled.PRIME.proximal <- base::lapply(
  X = pooled.PRIME,
  FUN = function(gr){
      gr[gr$class == "proximal"]
    }
  )
```

## [8.11.] Run separate analysis for distal and proximal
```{r}
pooled.PRIME.GWAS.enrichment.recall.data.distal <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME.distal,
    fg.data = GWAS.data,
    bg.data = SNP.data,
    assay = pooled.PRIME.names
  )

pooled.PRIME.GWAS.enrichment.recall.data.proximal <- .GlobalEnv$enrichment.precision.recall(
    data = pooled.PRIME.proximal,
    fg.data = GWAS.data,
    bg.data = SNP.data,
    assay = pooled.PRIME.names
  )
```

## [8.12.] Subset data for plotting
```{r}
point.GWAS.genome.distal <- base::subset(point.GWAS.df, loc == "distal")
```

## [8.13.] Plot data
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.GWAS.enrichment.recall.data.distal[
        pooled.PRIME.GWAS.enrichment.recall.data.distal$recall > 0.001,
      ],
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = point.GWAS.genome.distal
    ) + 
  theme_bw() +
  theme(
      aspect.ratio = 1
    ) +
  xlab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    )
```

## [8.14.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_GWAS_enrichment_vs_recall_and_point_estimates_distal.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [8.15.] Subset data for plotting
```{r}
point.GWAS.genome.proximal <- base::subset(point.GWAS.df, loc == "proximal")
```

## [8.16.] Plot data
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.GWAS.enrichment.recall.data.proximal[
        pooled.PRIME.GWAS.enrichment.recall.data.proximal$recall > 0.001,
      ],
    mapping = aes(
        x = recall,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
      data = point.GWAS.genome.proximal
    ) +
  xlab(
      base::paste0("recall (n=",GWAS.n,")")
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [8.17.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_GWAS_enrichment_vs_recall_and_point_estimates_proximal.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```


## [9.] Enrichment, precision, recall vs score -------------------------

## [9.0.] Plot data
```{r}
ggplot2::ggplot(
  data = pooled.PRIME.ClinVar.enrichment.recall.data,
  mapping = aes(
      x = score,
      y = enrichment,
      color = assay
    )
  ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.1.] Print data
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_ClinVar_enrichment_vs_score.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.2.] Plot data
```{r}
ggplot2::ggplot(
  data = pooled.PRIME.ClinVar.enrichment.recall.data,
  mapping = aes(
      x = score,
      y = recall,
      color = assay
    )
  ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.3.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_ClinVar_recall_vs_score.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.4.] Plot data
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.eQTL.enrichment.recall.data,
    mapping = aes(
        x = score,
        y = enrichment,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.5.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_eQTL_enrichment_vs_score_PIP_",
        eQTL.pip.thres,
        ".pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.6.] Plot data
```{r}
ggplot2::ggplot(
  data = pooled.PRIME.eQTL.enrichment.recall.data,
  mapping = aes(
      x = score,
      y = recall,
      color = assay
    )
  ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() + 
  theme(
      aspect.ratio = 1
    )
```

## [9.7.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_eQTL_recall_vs_score_PIP_",
        eQTL.pip.thres,
        ".pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.8.] Plot data
```{r}
ggplot2::ggplot(
  data = pooled.PRIME.GWAS.enrichment.recall.data,
  mapping = aes(
      x = score,
      y = enrichment,
      color = assay
    )
  ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.9.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_GWAS_enrichment_vs_score_PIP_",
        GWAS.pip.thres,
        ".pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.10.] Plot data
```{r}
ggplot2::ggplot(
  data = pooled.PRIME.GWAS.enrichment.recall.data,
  mapping = aes(
      x = score,
      y = recall,
      color = assay
    )
  ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme(
    aspect.ratio=1)
```

## [9.11.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_GWAS_recall_vs_score_PIP_",
        GWAS.pip.thres,
        ".pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.12.] Plot data 
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.CRISPR.enrichment.recall.data,
    mapping = aes(
        x = score,
        y = precision,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.13.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_CRISPR_precision_vs_score.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [9.14.] Plot data 
```{r}
ggplot2::ggplot(
    data = pooled.PRIME.CRISPR.enrichment.recall.data,
    mapping = aes(
        x = score,
        y = recall,
        color = assay
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [9.15.] Print data
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "K562_pooled_PRIME_CRISPR_recall_vs_score.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [10.] Upset plots -------------------------

```{r}
gr_list <- GRangesList(point.data)

gr_list[["GWAS.data"]] <- GWAS.data
gr_list[["eQTL.data"]] <- eQTL.data
gr_list[["ClinVar.data"]] <- ClinVar.data
gr_list[["CRISPR.data"]] <- CRISPR.pos.data
gr_list[["CRISPR.neg.data"]] <- CRISPR.neg.data
gr_list[["dcTap.pos.data"]] <- dc.tap.enhancers.pos
gr_list[["dcTap.neg.data"]] <- dc.tap.enhancers.neg
```

```{r}
create_upset_plot <- function(subset_name, gr_list, sets_to_include) {
    # Get the specific subset's GRanges
    subset_gr <- gr_list[[subset_name]]
  
    # Create a binary matrix for overlaps, restricted to the subset
    overlap_matrix <- sapply(gr_list, function(gr) {
      overlaps <- findOverlaps(subset_gr, gr)
      overlap_vector <- rep(0, length(subset_gr))
      overlap_vector[unique(queryHits(overlaps))] <- 1
      return(overlap_vector)
    })
    colnames(overlap_matrix) <- names(gr_list)
    
    # Convert to a data frame
    overlap_df <- as.data.frame(overlap_matrix)
    rownames(overlap_df) <- NULL
    
    # Filter to include only relevant sets
    subset_overlap_df <- overlap_df[, sets_to_include]
    
    # Generate the UpSet plot
    base::library(UpSetR)
    p<- upset(
      subset_overlap_df,
      sets = names(gr_list)[sets_to_include],
      order.by = "freq",
      sets.bar.color=col.scheme[names(col.scheme) %in% names(gr_list)[sets_to_include]]
    )
    
      p1 <-upset(
      subset_overlap_df,
      sets = names(gr_list)[sets_to_include],
      order.by = "freq",
      sets.bar.color=col.scheme[match(p$labels,names(col.scheme))],
      line.size=0
    )
      p1
  }
```

## Define sets to include for intersection analysis
```{r}
sets_to_include <-  names(gr_list) %in% base::c("PRIME CAGE cells pooled", "PRIME CAGE nuclei pooled","DHS")
```

## Generate plots for PRIME/DHS focus
```{r}
s <- create_upset_plot("GWAS.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_GWAS_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("ClinVar.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_ClinVar_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("eQTL.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_eQTL_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("CRISPR.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_CRISPR.pos_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("CRISPR.neg.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_CRISPR.neg_upset.pdf",height=6, width=8)
s
dev.off()
```

# Generate plots for full panel
```{r}
sets_to_include <-  names(gr_list) %in% names(col.scheme)
```

```{r}
s <- create_upset_plot("GWAS.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_full_GWAS_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("ClinVar.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_full_ClinVar_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("eQTL.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_full_eQTL_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("CRISPR.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_full_CRISPR.pos_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("CRISPR.neg.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_full_CRISPR.neg_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("dcTap.pos.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_dcTap.pos_upset.pdf",height=6, width=8)
s
dev.off()
```

```{r}
s <- create_upset_plot("dcTap.neg.data", gr_list, sets_to_include)
pdf("plots/K562_PRIME_dcTap.neg_upset.pdf",height=6, width=8)
s
dev.off()
```
