---
title: "10.3.Cell_line_vs_trait_enrichment"
output: html_document
date: "2024-07-01"
---

## Evaluation of PRIMEloci using fine-mapped GWAS variants

## [0.0.] Load necessary packages
```{r warning=TRUE, message=TRUE}
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(UpSetR)
```

## [0.1.] Add custom color palettes
```{r}
col.scheme <- base::c(
    "#377eb8","#d39200","#93aa00",
    "#4daf4a","#984ea3","#ff7f01",
    "#e41b1c","#256866","#f8bc24",
    "grey","black","hotpink"
  )
```

## [1.] Read in & prepare data -------------------------

## [1.0.] Read in CDSs from txdb
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
cds <- base::unlist(GenomicFeatures::cdsBy(txdb,"gene"))
cds <- BiocGenerics::sort(cds)
```

## [1.1.] Overlap CDSs with background SNPs (9,913,974 unique autosomal SNP locations)
#### Remove SNPs overlapping with CDSs
```{r}
SNP.data <- rtracklayer::import.bed( # 9.991.229
   con = base::paste0(
       here::here("0.External_resources/"),
       "all.bg.SNPs.hg38.baseline.v1.1.bed.sorted"
     ) ## SNP.file
  )
SNP.data <- BiocGenerics::sort(SNP.data)
olap <- GenomicRanges::findOverlaps( ## 93.742
    query = SNP.data,
    subject = cds,
    minoverlap = 1L
  )
SNP.data$name <- NULL
SNP.data <- base::unique( # 9.914.129
    SNP.data[-S4Vectors::queryHits(olap)]
  )
# GenomeInfoDb::seqlevels(txdb, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.2.] Subset PRIMEloci from pooled replicates by threshold
#### evaluate with  n = "K562_C" # 4,049,568
```{r}
PRIMEloci.thres = 0.75
resize.width = 200
pooled.PRIMEloci.files <- base::c(
    "A549_C" = here::here("8.Genomewide_prediction/A549_sld_on_PRIMEloci1.0_C.bed"),
    "A549_N" = here::here("8.Genomewide_prediction/A549_sld_on_PRIMEloci1.0_N.bed"),
    "GM12878_C" = here::here("8.Genomewide_prediction/GM12878_sld_on_PRIMEloci1.0_C_1M.bed"),
    "GM12878_N" = here::here("8.Genomewide_prediction/GM12878_sld_on_PRIMEloci1.0_N_1M.bed"),
    "HCT116_C" = here::here("8.Genomewide_prediction/HCT116_sld_on_PRIMEloci1.0_C.bed"),
    "HepG2_C" = here::here("8.Genomewide_prediction/HepG2_sld_on_PRIMEloci1.0_C.bed"),
    "HepG2_N" = here::here("8.Genomewide_prediction/HepG2_sld_on_PRIMEloci1.0_N.bed"),
    "K562_C" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_C.bed"),
    "K562_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N.bed")
  )

pooled.PRIMEloci <- base::lapply(
  X = base::names(pooled.PRIMEloci.files),
  FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIMEloci.files[n],
          show_col_types = FALSE
        )
      
      gr <- gr[base::which(gr$score >= PRIMEloci.thres),] # 245.709
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3])
      
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 237.454
      
      gr <- IRanges::resize(
          x = gr,
          width = resize.width,
          fix = "center"
        )
      
      return(IRanges::reduce(x = gr))
    }
  )
base::names(pooled.PRIMEloci) <- base::names(pooled.PRIMEloci.files)
```

## [1.3.] Prepare non-subset PRIMEloci set
```{r}
unfiltered.PRIMEloci <- base::lapply(
    X = base::names(pooled.PRIMEloci.files),
    FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIMEloci.files[n],
          show_col_types = FALSE
        )
      score <- gr$score
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3])
      gr$score <- score
      
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 237.454
      #gr <- resize(gr, width=resize.width, fix="center")
      return(gr)
    }
  )
base::names(unfiltered.PRIMEloci) <- base::names(pooled.PRIMEloci.files)
```

## [1.4.] Read in, prepare and subset GWAS data to contain only intergenic GWAS
#### Read out traits associated with these intergenic GWAS
```{r}
GWAS.pip.thres = 0.2
GWAS.data <- readr::read_tsv( # 66.638
    file = base::paste0(
        here::here("0.External_resources/"),
        "UKBB_94traits_release1_pip01_clean_uniq_hg38.bed"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
base::colnames(GWAS.data) <- base::c(
    "chr","start","end",
    "name","score","strand",
    "RS_id","trait","pip"
  )
GWAS.data <- GenomicRanges::makeGRangesFromDataFrame( # 66.638
    df = GWAS.data,
    keep.extra.columns = TRUE,
    starts.in.df.are.0based = TRUE
  )
GWAS.data <- base::subset( # 30.743
    GWAS.data, pip >= GWAS.pip.thres
  )
GenomeInfoDb::seqlevels(GWAS.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)

olap <- GenomicRanges::findOverlaps(
    query = GWAS.data,
    subject = cds,
    minoverlap = 1
  )
GWAS.data <- GWAS.data[-S4Vectors::queryHits(olap)] ## 28.182
traits <- base::unique(GWAS.data$trait) # 94
```

## [1.5.] Overlap PRIMElocis with GWAS in a trait-specific manner to subset traits
#### evaluate with t = "MCV", n = "K562_C", x = trait.loci.olap[[1]]
##  'trait.loci.olap' includes all GWAS hits for a particular trait that overlaps with any of the PRIMElocis from any PRIMEloci dataset
## 'trait.olap.count' summarises the overall overlaps of traits across all PRIMEloci datasets
```{r}
trait.loci.olap <- base::sapply(
    X = traits,
    FUN = function(t) {
      ids <- base::which(GWAS.data$trait == t)
      
      base::unique(
        base::unlist(
          base::lapply(
            X = base::names(pooled.PRIMEloci),
            FUN = function(n) {
              pd <- pooled.PRIMEloci[[n]] # 32.407
              GWAS.olap <- GenomicRanges::findOverlaps(
                  query = GWAS.data[ids,],
                  subject = pd,
                  minoverlap = 1
                )
              return(ids[S4Vectors::queryHits(GWAS.olap)])
            }
          )
        )
      )
    }
  )

trait.olap.count <- base::sapply(
  X = trait.loci.olap,
  FUN = function(x){
    base::length(
        base::unique(GWAS.data[x,])
      )
    }
  )
filtered.traits <- base::names(
    base::which(trait.olap.count >= 10)
  )
```

## [1.6.] Aggregate PRIMEloci-GWAS-overlaps into a matrix
#### evaluate with t = "eGFRcys", n = "K562_C"
```{r}
trait.matrices <- base::lapply(
    X = filtered.traits,
    FUN = function(t) {
    
      #g <- unique(GWAS.data[trait.loci.olap[[t]],])
      g <- GWAS.data[trait.loci.olap[[t]],]
      name <- base::paste0(
          GenomeInfoDb::seqnames(g),":",
          GenomicRanges::start(g) + 1,":",
          GenomicRanges::end(g),":",
          g$RS_id
        )
      m <- base::matrix(
          data = 0,
          ncol = base::length(pooled.PRIMEloci),
          nrow = base::length(g),
          dimnames = base::list(name, base::names(pooled.PRIMEloci))
        )
      
      for (n in base::names(pooled.PRIMEloci)) {
          pd <- unfiltered.PRIMEloci[[n]]
          GWAS.olap <- IRanges::findOverlaps(
              query = g,
              subject = pd,
              minoverlap = 1
            )
          m[base::unique(S4Vectors::queryHits(GWAS.olap)),n] <- base::as.numeric(
            base::by(
                data = pd$score[S4Vectors::subjectHits(GWAS.olap)],
                INDICES = S4Vectors::queryHits(GWAS.olap),
                FUN = base::max
              )
            )
        }
      return(m)
    }
  )
base::names(trait.matrices) <- filtered.traits
```

## [1.7.] Calculate GWAS enrichment and recall for every trait separately
```{r}
point.GWAS.df <- base::do.call(
          what = "rbind",
          args = base::lapply(
            X = base::names(pooled.PRIMEloci),
            FUN = function(n) {
              pd <- pooled.PRIMEloci[[n]]
              bg.olap <- IRanges::findOverlaps(
                  query = SNP.data,
                  subject = pd,
                  minoverlap = 1
                )
              bg.count <- base::length(
                base::unique(
                    S4Vectors::queryHits(bg.olap)
                  )
                )
              GWAS.olap <- IRanges::findOverlaps(
                  query = GWAS.data,
                  subject = pd,
                  minoverlap = 1
                )
          return(
            base::do.call(
              what = "rbind",
              args = base::lapply(
                X = filtered.traits,
                FUN = function(t){
            
                  fg <- base::which(GWAS.data$trait == t)
                  fg.olap <- GWAS.olap[S4Vectors::queryHits(GWAS.olap) %in% fg,]
                  fg.count <- base::length(
                    base::unique(
                        S4Vectors::queryHits(fg.olap)
                      )
                    )
            
              base::data.frame(
                  enrichment = (fg.count/base::length(fg)) / (bg.count/base::length(SNP.data)),
                  recall = fg.count/length(fg),
                  fg.count = fg.count,
                  bg.count = bg.count,
                  fg.n = base::length(fg),
                  bg.n = base::length(SNP.data),
                  sample = n,
                  trait = t
                )
              }
            )
          )
        )
      }
    )
  )
```

# [2.] Visualize GWAS results -------------------------

## [2.0.] Prepare plotting of trait enrichment*recall across all samples
```{r}
point.GWAS.df <- point.GWAS.df %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(
      scaled_enrichment = scales::rescale(enrichment)
    ) %>%
  dplyr::ungroup()
point.GWAS.df <- base::as.data.frame(point.GWAS.df)

enr.mat <- base::matrix(
  data = 0,
  nrow = base::length(filtered.traits),
  ncol = base::length(base::unique(point.GWAS.df$sample)),
  dimnames = base::list(
      filtered.traits,
      base::unique(point.GWAS.df$sample)
    )
  )
for (i in 1:base::nrow(point.GWAS.df)){ 
    enr.mat[point.GWAS.df[i,"trait"],point.GWAS.df[i,"sample"]] <- point.GWAS.df[i,"scaled_enrichment"]*point.GWAS.df[i,"recall"]
  }

ro <- stats::hclust(
    d = stats::dist(base::log2(enr.mat + 1))
  )$order

point.GWAS.df$sample <- base::factor(
  x = point.GWAS.df$sample,
  levels = base::c(
      "K562_C","K562_N","GM12878_C",
      "GM12878_N","HepG2_C","HepG2_N",
      "A549_C","A549_N","HCT116_C"
    )
  )
point.GWAS.df$trait <- base::factor(
    x = point.GWAS.df$trait,
    levels = base::rownames(enr.mat)[ro]
  )
```

## [2.1.] Plot enrichment*recall matrix across samples & traits
```{r}
cols <- grDevices::colorRampPalette(
    base::rev(RColorBrewer::brewer.pal(11,"RdBu"))
  )(100)

ggplot2::ggplot(
    data = point.GWAS.df,
    mapping = aes(
        x = sample,
        y = trait,
        fill = scaled_enrichment
      )
    ) +
  geom_tile(
      height = 0.9,
      width = 0.9
    ) +
  coord_equal() +
  # scale_fill_distiller(
  #     direction = 1
  #   ) +
  scale_fill_gradientn(
      colors = .GlobalEnv$cols,
      limits = base::c(0L,1L)
    ) +
  theme_bw() +
  theme(
    # aspect.ratio = 1,
    axis.text.x = element_text(
        angle = 90,
        hjust = 1
      ),
    axis.title = element_blank(),
    legend.title = element_blank()
  )
```

## [2.2.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "GWAS_trait_enrichment_vs_cell_line.pdf"
      ),
    width = 3,
    height = 5
  )
```

## [2.3.] Plot all trait matrices across all GWAS & all cell lines
#### evaluate with 'trait = "MCH"'
```{r}
focus.traits <- base::as.character(
  base::unique(
      base::subset(x = point.GWAS.df, enrichment >= 10)$trait
    )
  )

for (trait in focus.traits) {
  
  m <- trait.matrices[[trait]]
  # base::print(base::dim(m))
  
  ro <- stats::hclust(stats::dist(m))$order
  co <- stats::hclust(stats::dist(base::t(m)))$order
  
  df <- base::data.frame(
      score = base::as.numeric(m),
      cell.line = base::rep(
          base::colnames(m),
          each = base::nrow(m)
        ),
      variant = base::rownames(m)
    )
  df$variant <- base::factor(
      x = df$variant,
      levels = base::rownames(m)[ro]
    )
  df$cell.line <- base::factor(
      x = df$cell.line,
      levels = base::colnames(m)[co]
    )
  
  ggplot2::ggplot(
      data = df,
      mapping = aes(
          x = cell.line,
          y = variant,
          fill = score
        )
      ) +
    scale_fill_distiller(
        direction = 1
      ) +
    ggtitle(trait) +
    geom_tile() +
    theme(
        axis.text.x = element_text(
            angle = 90,
            hjust = 1
          ),
        axis.title = element_blank()
      )
    
  # ggplot2::ggsave(
  #   filename = base::paste0(
  #         here::here("10.External_validation/plots/"),
  #         trait,
  #         "_variant_PRIMEloci_score_vs_cell_line.pdf"
  #       ),
  #     width = 5,
  #     height = 5,
  #     device = "pdf"
  #   )
  }
```


## [3.] Calculate enrichment, precision, and recall versus score threshold -------------------------

## [3.0.] Helper function for preparing enrichment-recall curves
```{r}
enrichment.precision.recall <- function(data, fg.data, bg.data, assay) {
  
  base::do.call(
    what = "rbind",
    args = base::lapply(
      X = base::names(data),
      FUN = function(n) {
    
        gr <- data[[n]]
        scores <- base::seq(0, 1, by = 0.01)
        cnt <- base::sapply(
          X = scores,
          FUN = function(sco){
              base::sum(gr$score>=sco)
            }
          )
        scores <- scores[cnt>=10]
        
        gr.fg.olap <- GenomicRanges::findOverlaps(
            query = fg.data,
            subject = gr,
            minoverlap = 1
          )
        gr.bg.olap <- GenomicRanges::findOverlaps(
            query = bg.data,
            subject = gr,
            minoverlap = 1
          )
    
        base::do.call(
          what = "rbind",
          args = base::lapply(
            X = scores,
            FUN = function(sco){
      
              selection <- base::which(gr$score>=sco)
              fg.count <- base::length(
                base::unique(
                  S4Vectors::queryHits(gr.fg.olap)[
                      base::which(S4Vectors::subjectHits(gr.fg.olap) %in% selection)
                    ]
                  )
                )
              bg.count <- base::length(
                base::unique(
                  S4Vectors::queryHits(gr.bg.olap)[
                      base::which(S4Vectors::subjectHits(gr.bg.olap) %in% selection)
                    ]
                  )
                )
              
              base::data.frame(
                    enrichment = (fg.count/length(fg.data))/(bg.count/length(bg.data)),
                    precision = fg.count/(fg.count+bg.count),
                    recall = fg.count/length(fg.data),
                    fg.count = fg.count,
                    bg.count = bg.count,
                    score = sco,
                    sample = n,
                    assay = assay[n]
                )
              }
            )
          )
        }
      )
    )
  }
```

## [3.1.] Run enrichment recall analysis for all traits separately
#### run_enrichment_for_trait as a wrapper function of enrichment.precision.recall
```{r}
run_enrichment_for_trait <- function(trait) {
    .GlobalEnv$enrichment.precision.recall(
      data = unfiltered.PRIMEloci,
      fg.data = GWAS.data[GWAS.data$trait == trait,], 
      bg.data = SNP.data, 
      assay = names(unfiltered.PRIMEloci)
    )
  }

per.trait.results <- parallel::mclapply(
    X = filtered.traits,
    FUN = run_enrichment_for_trait,
    mc.cores = 30
  )
base::names(per.trait.results) <- focus.traits
```

## [3.2.] Plot all trait-specific enrichment-recall curves
```{r}
base::lapply(
    X = base::seq_along(per.trait.results),
    FUN = function(i) {
      trait_name <- filtered.traits[i]
      data <- per.trait.results[[i]]
    
      # Create plots
      pr <- ggplot2::ggplot(
          data = data,
          mapping = aes(
              x = recall,
              y = enrichment,
              color = sample
            )
          ) +
        geom_line(
          mapping = aes(
              group = sample
            )
          ) +
        ggtitle(
          base::paste(
              "Recall vs Enrichment for", trait_name
            )
          ) +
        theme_bw() +
        theme(
            aspect.ratio = 1
          )
    
      # ggplot2::ggsave(
      #     filename = base::paste0(
      #         here::here("10.External_validation/plots/"),
      #         "cellLinePlots/",
      #         trait_name, "_enrichment_recall_plot.pdf"
      #       ),
      #     plot = pr,
      #     device = "pdf"
      #   )
    }
  )
```

## [3.3.] Subset traits
```{r}
# Over pooled traits
#blood.traits <- c("HbA1c","Hb","RBC","MCV","MCH","MCHC")
#hep.traits <- c("ApoB", "TC", "TBil", "SHBG", "LDLC", "TG","Alb")

# Blood-related traits (erythrocytes, platelets, hematology)
blood.traits <- base::c(
    "HbA1c",  # Glycated hemoglobin, diabetes marker
    "Hb",      # Hemoglobin, oxygen transport
    "RBC",     # Red Blood Cell count
    "MCV",     # Mean Corpuscular Volume, RBC size
    "MCH",     # Mean Corpuscular Hemoglobin, RBC hemoglobin content
    "MCHC"    # Mean Corpuscular Hemoglobin Concentration
    # "Plt"      # Platelet count, clotting function
  )
# Immune-related traits (white blood cells, inflammation, autoimmunity)
immune.traits <- base::c(
    "WBC",    # Total White Blood Cell count
    "Neutro", # Neutrophils, first-line immune defense
    "Lym",    # Lymphocytes, adaptive immunity
    "Mono",   # Monocytes, innate immune response
    "Eosino", # Eosinophils, allergy and parasite defense
    "Baso",   # Basophils, involved in histamine response
    "CRP",    # C-reactive protein, systemic inflammation marker
    "IBD",    # Inflammatory Bowel Disease, chronic immune disorder
    "AID_Combined", # Autoimmune diseases combined category
    "Asthma", # Chronic inflammatory airway disease
    "Hypothyroidism", # Can be autoimmune (e.g., Hashimoto's)
    "Fibroblastic_Disorders" # Potential immune involvement
  )
# Liver/metabolic traits (hepatic function, lipids, glucose metabolism)
hep.traits <- base::c(
    "ApoB",     # Apolipoprotein B, LDL-associated
    "TC",       # Total Cholesterol
    "LDLC",     # Low-Density Lipoprotein Cholesterol (bad cholesterol)
    "TG",       # Triglycerides, fat storage
    "Alb",      # Albumin, major liver protein
    "SHBG",     # Sex Hormone Binding Globulin, linked to liver metabolism
    "TBil",     # Total Bilirubin, liver function indicator
    "ALT",      # Alanine Aminotransferase, liver enzyme
    "AST",      # Aspartate Aminotransferase, liver enzyme
    "GGT",      # Gamma-Glutamyl Transferase, liver and bile function
    "Urea",     # Metabolic waste, partly liver-related
    "Glucose",  # Blood glucose levels, diabetes-related
    "IGF1"      # Insulin-like Growth Factor 1, growth hormone-related
  )

blood.GWAS.data <- base::subset(GWAS.data, trait %in% blood.traits) # 4.848 # 3.838 w/o Plt
hep.GWAS.data <- base::subset(GWAS.data, trait %in% hep.traits) # 5.005
immune.GWAS.data <- base::subset(GWAS.data, trait %in% immune.traits) # 4.093
```

## [3.4.] Run enrichment recall analysis for subset traits
```{r}
cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = unfiltered.PRIMEloci,
    fg.data = blood.GWAS.data,
    bg.data = SNP.data,
    assay = base::names(unfiltered.PRIMEloci)
  )

cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = unfiltered.PRIMEloci,
    fg.data = hep.GWAS.data,
    bg.data = SNP.data,
    assay = base::names(unfiltered.PRIMEloci)
  )

cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data <- .GlobalEnv$enrichment.precision.recall(
    data = unfiltered.PRIMEloci,
    fg.data = immune.GWAS.data,
    bg.data = SNP.data,
    assay = base::names(unfiltered.PRIMEloci)
  )
```

## [3.5.] Plot blood GWAS data
```{r}
ggplot2::ggplot(
    data = cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = sample
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
    data = cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data[
        cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data$score==0.75,
      ]
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  scale_y_continuous(
      limits = base::c(0,13)
    ) +
  xlab(
    base::paste0(
        "recall (n=", base::length(blood.GWAS.data$name),")"
      )
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.justification = base::c(1,1),
      legend.position = base::c(0.95,0.95),
      legend.background = element_rect(
          fill = "grey"
        ),
      legend.title = element_blank(),
      legend.spacing.y = grid::unit(3, 'lines')
    )
```

## [3.6.] Print blood GWAS plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_bloodTraits",
        "_enrichment_recall_plot.pdf"
      ),
    dpi = 300,
    height = 4,
    width = 4,
    device = "pdf"
  )
```

## [3.7.] Plot hepatocyte / liver GWAS
```{r}
ggplot2::ggplot(
    data = cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = sample
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
    data = cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data[
        cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data$score == 0.75,
      ]
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  scale_y_continuous(
      limits = base::c(0,13)
    ) +
  xlab(
    base::paste0(
        "recall (n=", base::length(hep.GWAS.data$name), ")"
      )
    ) +
  theme_bw() +
  theme(
        aspect.ratio = 1,
        legend.justification = base::c(1,1),
        legend.position = base::c(0.95,0.95),
        legend.background = element_rect(
            fill = "grey"
          ),
        legend.title = element_blank(),
        legend.spacing.y = grid::unit(3, 'lines')
      )
```

## [3.8.] Print hepatocyte / liver GWAS plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_hepTraits",
        "_enrichment_recall_plot.pdf"
      ),
    dpi = 300,
    height = 4,
    width = 4,
    device = "pdf"
  )
```

## [3.9.] Plot immune GWAS
```{r}
ggplot2::ggplot(
    data = cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data,
    mapping = aes(
        x = recall,
        y = enrichment,
        color = sample
      )
    ) +
  geom_line(
    mapping = aes(
        group = sample
      )
    ) +
  geom_point(
    data = cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data[
        cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data$score == 0.75,
      ]
    ) +
  scale_color_manual(
      values = col.scheme
    ) +
  scale_y_continuous(
      limits = base::c(0,13)
    ) +
  xlab(
    base::paste0(
        "recall (n=", base::length(hep.GWAS.data$name), ")"
      )
    ) +
  theme_bw() +
  theme(
        aspect.ratio = 1,
        legend.justification = base::c(1,1),
        legend.position = base::c(0.95,0.95),
        legend.background = element_rect(
            fill = "grey"
          ),
        legend.title = element_blank(),
        legend.spacing.y = grid::unit(3, 'lines')
      )
```

## [3.10.] Print immune GWAS plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "immuneTraits",
        "_enrichment_recall_plot.pdf"
      ),
    device = "pdf"
  )
```

## [3.11.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%blood.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = enrichment
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 7
    ) +
  theme_bw() +
  theme(
      axis.text.x = element_text(
          angle = 90,
          hjust = 1
        ),
      axis.title = element_blank()
    )
```

## [3.12.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "bloodTraits",
        "_enrichment_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [3.13.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%blood.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = recall
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 2
    ) +
  theme(
      axis.text.x = element_text(
          angle = 90,
          hjust = 1
        ),
      axis.title = element_blank()
    )
```

## [3.14.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "bloodTraits",
        "_recall_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [3.15.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%hep.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = enrichment
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 7
    ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
        angle = 90,
        hjust = 1
      )
    )
```

## [3.16.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "hepTraits",
        "_enrichment_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [3.17.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%hep.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = recall
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 2
    ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
        angle = 90,
        hjust = 1
      )
    )
```

## [3.18.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "hepTraits",
        "_recall_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [3.19.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%immune.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = enrichment
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 7
    ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
        angle = 90,
        hjust = 1
      )
    )
```

## [3.20.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "immuneTraits",
        "_enrichment_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

## [3.21.] Plot data
```{r}
ggplot2::ggplot(
    data = point.GWAS.df[point.GWAS.df$trait%in%immune.traits,],
    mapping = aes(
        x = sample,
        y = trait,
        fill = recall
      )
    ) +
  geom_tile() +
  scale_fill_distiller(
      direction = 1,
      palette = 2
    ) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
        angle = 90,
        hjust = 1
      )
    )
```

## [3.22.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/plots/"),
        "cellLinePlots/",
        "immuneTraits",
        "_recall_hm.pdf"
      ),
    width = 7,
    height = 5,
    device = "pdf"
  )
```

# [4.]

## [4.0.] Annotate GWAS hits as promoter (proximal) or enhancer (distal)
```{r}
promoters <- IRanges::promoters(
    x = txdb,
    upstream = 500,
    downstream = 500
  )
GWAS.data$type <- base::rep("distal",base::length(GWAS.data))
olap <- IRanges::findOverlaps(
    query = GWAS.data,
    subject = promoters,
    minoverlap = 1
  )
GWAS.data$type[S4Vectors::queryHits(olap)] <- "proximal"
```

## [4.1.] Convert data
```{r}
GWAS.df <- base::as.data.frame(GWAS.data)
GWAS.df$name <- base::paste0(
    GenomeInfoDb::seqnames(GWAS.data),":",
    GenomicRanges::start(GWAS.data) + 1,":",
    GenomicRanges::end(GWAS.data),":",
    GWAS.data$RS_id
  )
```

## [4.2.] Overlap non-subset PRIMElocis with GWAS in a trait-specific manner to subset traits
#### see [1.5.] above for previous instalment of 'trait.loci.olap' object generation
#### difference to [1.5.] is that here overlaps are calculated with 'unfiltered.PRIMEloci'
```{r}
trait.loci.olap <- base::sapply(
  X = traits,
  FUN = function(t) {
    ids <- base::which(GWAS.data$trait == t)
    base::unique(
      base::unlist(
        base::lapply(
          X = base::names(unfiltered.PRIMEloci),
          FUN = function(n) {
              pd <- unfiltered.PRIMEloci[[n]]
              GWAS.olap <- IRanges::findOverlaps(
                  query = GWAS.data[ids,],
                  subject = pd,
                  minoverlap = 1
                )
              return(ids[S4Vectors::queryHits(GWAS.olap)])
            }
          )
        )
      )
    }
  )
```

## [4.3.] Aggregate PRIMEloci-GWAS-overlaps into a matrix
```{r}
trait.matrices <- base::lapply(
  X = base::c(
      blood.traits, hep.traits, immune.traits
    ),
  FUN = function(t) {
    g <- GWAS.data[trait.loci.olap[[t]],]
    name <- base::paste0(
        GenomeInfoDb::seqnames(g),":",
        GenomicRanges::start(g) + 1,":",
        GenomicRanges::end(g),":",
        g$RS_id
      )
    m <- base::matrix(
        data = 0,
        ncol = base::length(unfiltered.PRIMEloci),
        nrow = base::length(g),
        dimnames = base::list(name, base::names(unfiltered.PRIMEloci))
      )
  
    for (n in base::names(unfiltered.PRIMEloci)){
      pd <- unfiltered.PRIMEloci[[n]]
      GWAS.olap <- IRanges::findOverlaps(
          query = g,
          subject = pd,
          minoverlap = 1
        )
      m[base::unique(S4Vectors::queryHits(GWAS.olap)),n] <- base::as.numeric(
        base::by(
            data = pd$score[subjectHits(GWAS.olap)],
            INDICES = S4Vectors::queryHits(GWAS.olap),
            FUN = base::max
          )
        )
      }
    return(m)
    }
  )
base::names(trait.matrices) <- base::c(blood.traits,hep.traits,immune.traits)
utils::str(trait.matrices)
```

## [4.4.] Plot traits
#### evaluate with trait = "HbA1c"
```{r}
for (trait in base::c(blood.traits,hep.traits,immune.traits)) {
  
  m <- trait.matrices[[trait]]
  base::print(base::dim(m))
  
  ro <- stats::hclust(stats::dist(m))$order
  #co <- hclust(dist(t(m)))$order
  
  df <- base::data.frame(
      score = base::as.numeric(m),
      cell.line = base::rep(base::colnames(m), each = base::nrow(m)),
      variant = base::rownames(m)
    )
  df$variant <- base::factor(
      x = df$variant,
      levels = base::rownames(m)[ro]
    )
  #df$cell.line <- factor(df$cell.line, levels=colnames(m)[co])
  m <- base::mean(df$score[df$score>0])
  df$score[df$score==0] <- NA
  
  df <- base::merge(
      x = df,
      y = GWAS.df[,base::c("name","type")],
      by.x = "variant",
      by.y = "name"
    )
  
  ggplot2::ggplot(
      data = df,
      mapping = aes(
          x = cell.line,
          y = variant
        )
      ) + 
    geom_tile(
        mapping = aes(
            fill = score
          ),
        color = "grey90",
        size = 0.1
      ) +  # Heatmap
    scale_fill_gradient2(
        low = "#256866",
        mid = "#FFFFCC",
        high = "orange",
        na.value = NA,
        midpoint = m
      ) +
    ggnewscale::new_scale_fill() +
    # Allows a second independent fill scale
    geom_tile(
        mapping = aes(
            x = base::length(base::unique(df$cell.line)) + 0.5,
            fill = type
          ),
        width = 0.2
      ) +  # Annotation strip
    ggtitle(trait) +
    scale_x_discrete(
        expand = base::c(0, 0)
      ) +
    scale_y_discrete(
        expand = base::c(0, 0)
      ) +  
    scale_fill_manual(
      values = base::c(
          "proximal"="#4daf4a",
          "distal" = "#984ea3"
        )
      ) + # Discrete for annotation
    theme_bw() +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_text(
          angle = 90,
          hjust = 1
        ),
      axis.text.y = element_text(
          size = 3
        ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(
            color = "black",
            fill = NA,
            linewidth = 1
          )
        )
     
  # ggplot2::ggsave(
  #     filename = base::paste0(
  #         here::here("10.External_validation/plots/"),
  #         "cellLinePlots/",
  #         trait,
  #         "_variant_PRIMEloci_score_vs_cell_line.pdf"
  #       ),
  #     width = 5,
  #     height = 5,
  #     device = "pdf"
  #   )
}
```

## [4.5.] Prepare data for MCHC trait specific plotting
```{r}
trait = "ApoB" # "MCHC" # "ApoB"
m <- trait.matrices[[trait]]

ro <- stats::hclust(stats::dist(m))$order
#co <- hclust(dist(t(m)))$order
  
df <- base::data.frame(
    score = base::as.numeric(m),
    cell.line = base::rep(base::colnames(m), each = base::nrow(m)),
    variant = base::rownames(m)
  )
df$variant <- base::factor(
    x = df$variant,
    levels = base::rownames(m)[ro]
  )

midpoint <- base::mean(df$score[df$score>0])
# # df$score[df$score==0] <- NA

df <- base::merge(
      x = df,
      y = GWAS.df[,base::c("name","type")],
      by.x = "variant",
      by.y = "name"
    ) %>%
  dplyr::distinct(
      variant, score, cell.line, type
    )

# annos <- GWAS.df %>% 
#   dplyr::select(name,type) %>%
#   dplyr::distinct(name,type)
# 
# df %>%
#   dplyr::rename(name = "variant") %>% 
#    dplyr::left_join(
#        x = .,
#        y = annos,
#        by = name
#      )
```

## [4.6.] MCHC trait specific plotting
```{r}
coef = midpoint
REtype = "proximal" # "distal" # "proximal"
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#ffffff","#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc")
    )(coef*100),
  grDevices::colorRampPalette(
      base::c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e")
    )((1-coef)*100)
  )

df %>%
  dplyr::filter(type == REtype) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          x = variant,
          y = cell.line
        )
      ) + 
    geom_tile(
        mapping = aes(
            fill = score
          ),
        color = "grey90",
        size = 0.1
      ) +  # Heatmap
    scale_fill_gradientn(
        colors = .GlobalEnv$cols,
        limits = base::c(0L,1L)
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 5/3, #5/3, #1/3,
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(
              color = "black",
              fill = NA,
              linewidth = 1
            ),
        legend.title = element_blank()
      ) +
    guides(
        fill = guide_colorbar(
          label.position = "right",
          title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5
            ),
          ticks = T,
          nbin = 100,
          barheight = 5,
          label = T,
          barwidth = 0.5
        ),
        color = "none"
      )
```

## [4.7.] Print plot to file
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_", trait,
        "_", REtype,
        "_variant_PRIMEloci_score_vs_cell_line.pdf"
      ),
    plot = ggplot2::last_plot(),
    dpi = 300,
    width = 3, #1.6, #8,
    height = 3
  )
```

## [4.8.] Summary plot
```{r}
df %>%
  dplyr::filter(
      type == REtype
    ) %>%
  dplyr::group_by(cell.line) %>% 
  dplyr::summarise(
      nr_0 = base::sum(score > 0),
      nr_75 = base::sum(score > 0.75)
    ) %>%
  tidyr::pivot_longer(
      cols = base::c("nr_0","nr_75"),
      values_to = "count",
      names_to = "cutoff"
    ) %>% 
  ggplot2::ggplot(
        data = .,
        mapping = aes(
          x = count,
          y = cell.line,
          fill = cutoff
        )
      ) +
    geom_bar(
        stat = "identity",
        colour = "black",
        position = position_identity()
      ) +
    scale_fill_manual(
        values = base::c("white","grey50")
      ) +
    # geom_jitter(
    #     height = 0.3,
    #     width = 0
    #   ) +
    theme_bw() +
    theme(
        aspect.ratio = 1,
        panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        legend.justification = base::c(1,0),
        legend.position = base::c(0.95,0.05),
        legend.background = element_rect(
          fill = "grey"
        )
      )
```

## [4.9.] Print summary plot to file
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("10.External_validation/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_", trait,
        "_", REtype,
        "_variant_PRIMEloci_score_vs_cell_line",
        "_summary",
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    dpi = 300,
    width = 3,
    height = 3
  )
```

## [4.10.] Alternative plotting
```{r}
df %>%
  dplyr::filter(type == "distal") %>%
  tidyr::pivot_wider(
      names_from = "cell.line",
      values_from = "score"
    ) %>%
  dplyr::mutate(
      diff = 2*K562_N - base::rowSums(dplyr::across(tidyselect::contains("_")))
    ) %>%
  dplyr::arrange(
      K562_N, dplyr::desc(diff)
    ) %>%
  dplyr::mutate(
      variant = forcats::fct_inorder(f = variant)
    ) %>%
  tidyr::pivot_longer(
      cols = tidyselect::contains("_"),
      names_to = "cell.line",
      values_to = "score"
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
          x = variant,
          y = cell.line
        )
      ) + 
    geom_tile(
        mapping = aes(
            fill = score
          ),
        color = "grey90",
        size = 0.1
      ) +  # Heatmap
    scale_fill_gradientn(
        colors = .GlobalEnv$cols,
        limits = base::c(0L,1L)
      ) +
    theme_bw() +
    theme(
        aspect.ratio = 1/3,
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(
              color = "black",
              fill = NA,
              linewidth = 1
            ),
        legend.title = element_blank()
      ) +
    guides(
        fill = guide_colorbar(
          label.position = "right",
          title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5
            ),
          ticks=T,
          nbin=100,
          barheight=5,
          label=T,
          barwidth=0.5
        ),
        color = "none"
      )
```

