---
title: "10.3.Cell_line_vs_trait_enrichment"
output: html_document
date: "2024-07-01"
---

## Evaluation of PRIMEloci using fine-mapped GWAS variants

## [0.0.] Load necessary packages
```{r warning=TRUE, message=TRUE}
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(UpSetR)
```

## [0.1.] Add custom color palettes
```{r}
col.scheme <- base::c(
    "#377eb8","#d39200","#93aa00",
    "#4daf4a","#984ea3","#ff7f01",
    "#e41b1c","#256866","#f8bc24",
    "grey","black","hotpink"
  )
```

## [0.2.] Parameters
```{r}

```

## [1.] Read in & prepare data -------------------------

## [1.0.] Read in CDSs from txdb
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
cds <- base::unlist(GenomicFeatures::cdsBy(txdb,"gene"))
cds <- BiocGenerics::sort(cds)
```

## [1.1.] Overlap CDSs with background SNPs (9,913,974 unique autosomal SNP locations)
#### Remove SNPs overlapping with CDSs
```{r}
SNP.data <- rtracklayer::import.bed( # 9.991.229
   con = base::paste0(
       here::here("0.External_resources/"),
       "all.bg.SNPs.hg38.baseline.v1.1.bed.sorted"
     ) ## SNP.file
  )
SNP.data <- BiocGenerics::sort(SNP.data)
olap <- GenomicRanges::findOverlaps( ## 93.742
    query = SNP.data,
    subject = cds,
    minoverlap = 1L
  )
SNP.data$name <- NULL
SNP.data <- base::unique( # 9.914.129
    SNP.data[-S4Vectors::queryHits(olap)]
  )
# GenomeInfoDb::seqlevels(txdb, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)
```

## [1.2.] Subset PRIMEloci from pooled replicates by threshold
#### evaluate with  n = "K562_C" # 4,049,568
```{r}
PRIMEloci.thres = 0.75
resize.width = 200
pooled.PRIMEloci.files <- base::c(
    "A549_C" = here::here("8.Genomewide_prediction/A549_sld_on_PRIMEloci1.0_C.bed"),
    "A549_N" = here::here("8.Genomewide_prediction/A549_sld_on_PRIMEloci1.0_N.bed"),
    "GM12878_C" = here::here("8.Genomewide_prediction/GM12878_sld_on_PRIMEloci1.0_C_1M.bed"),
    "GM12878_N" = here::here("8.Genomewide_prediction/GM12878_sld_on_PRIMEloci1.0_N_1M.bed"),
    "HCT116_C" = here::here("8.Genomewide_prediction/HCT116_sld_on_PRIMEloci1.0_C.bed"),
    "HepG2_C" = here::here("8.Genomewide_prediction/HepG2_sld_on_PRIMEloci1.0_C.bed"),
    "HepG2_N" = here::here("8.Genomewide_prediction/HepG2_sld_on_PRIMEloci1.0_N.bed"),
    "K562_C" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_C.bed"),
    "K562_N" = here::here("8.Genomewide_prediction/K562_sld_on_PRIMEloci1.0_N.bed")
  )

pooled.PRIMEloci <- base::lapply(
  X = base::names(pooled.PRIMEloci.files),
  FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIMEloci.files[n],
          show_col_types = FALSE
        )
      
      gr <- gr[base::which(gr$score >= PRIMEloci.thres),] # 245.709
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3])
      
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 237.454
      
      gr <- IRanges::resize(
          x = gr,
          width = resize.width,
          fix = "center"
        )
      
      return(IRanges::reduce(x = gr))
    }
  )
base::names(pooled.PRIMEloci) <- base::names(pooled.PRIMEloci.files)
```

## [1.3.] Prepare non-subset PRIMEloci set
```{r}
unfiltered.PRIMEloci <- base::lapply(
    X = base::names(pooled.PRIMEloci.files),
    FUN = function(n) {
      gr <- readr::read_tsv( # # 4.049.578
          file = pooled.PRIMEloci.files[n],
          show_col_types = FALSE
        )
      score <- gr$score
      gr <- GenomicRanges::makeGRangesFromDataFrame(gr[,1:3])
      gr$score <- score
      
      GenomeInfoDb::seqlevels(gr, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data) # 237.454
      #gr <- resize(gr, width=resize.width, fix="center")
      return(gr)
    }
  )
base::names(pooled.PRIMEloci) <- base::names(pooled.PRIMEloci.files)
```

## [1.4.] Read in, prepare and subset GWAS data to contain only intergenic GWAS
#### Read out traits associated with these intergenic GWAS
```{r}
GWAS.pip.thres = 0.2
GWAS.data <- readr::read_tsv( # 66.638
    file = base::paste0(
        here::here("0.External_resources/"),
        "UKBB_94traits_release1_pip01_clean_uniq_hg38.bed"
      ),
    col_names = FALSE,
    show_col_types = FALSE
  )
base::colnames(GWAS.data) <- base::c(
    "chr","start","end",
    "name","score","strand",
    "RS_id","trait","pip"
  )
GWAS.data <- GenomicRanges::makeGRangesFromDataFrame( # 66.638
    df = GWAS.data,
    keep.extra.columns = TRUE,
    starts.in.df.are.0based = TRUE
  )
GWAS.data <- base::subset( # 30.743
    GWAS.data, pip >= GWAS.pip.thres
  )
GenomeInfoDb::seqlevels(GWAS.data, pruning.mode = "coarse") <- GenomeInfoDb::seqlevels(SNP.data)

olap <- GenomicRanges::findOverlaps(
    query = GWAS.data,
    subject = cds,
    minoverlap = 1
  )
GWAS.data <- GWAS.data[-S4Vectors::queryHits(olap)] ## 28.182
traits <- base::unique(GWAS.data$trait) # 94
```

## [1.5.] Overlap PRIMElocis with GWAS in a trait-specific manner & aggregate overlaps into a matrix
#### evaluate with t = "MCV", n = "K562_C", x = trait.loci.olap[[1]]
##  'trait.loci.olap' includes all GWAS hits for a particular trait that overlaps with any of the PRIMElocis from any PRIMEloci dataset
## 'trait.olap.count' summarises the overall overlaps of traits across all PRIMEloci datasets
```{r}
trait.loci.olap <- base::sapply(
    X = traits,
    FUN = function(t) {
      ids <- base::which(GWAS.data$trait == t)
      
      base::unique(
        base::unlist(
          base::lapply(
            X = base::names(pooled.PRIMEloci),
            FUN = function(n) {
              pd <- pooled.PRIMEloci[[n]] # 32.407
              GWAS.olap <- GenomicRanges::findOverlaps(
                  query = GWAS.data[ids,],
                  subject = pd,
                  minoverlap = 1
                )
              return(ids[S4Vectors::queryHits(GWAS.olap)])
            }
          )
        )
      )
    }
  )

trait.olap.count <- base::sapply(
  X = trait.loci.olap,
  FUN = function(x){
    base::length(
        base::unique(GWAS.data[x,])
      )
    }
  )
filtered.traits <- base::names(
    base::which(trait.olap.count >= 10)
  )
```


```{r}
trait.matrices <- base::lapply(
  X = filtered.traits,
  FUN = function(t) {
  
    #g <- unique(GWAS.data[trait.loci.olap[[t]],])
    g <- GWAS.data[trait.loci.olap[[t]],]
    name <- paste0(seqnames(g),":",start(g)+1,":",end(g),":",g$RS_id)
    m <- matrix(0,ncol=length(pooled.PRIMEloci),nrow=length(g),dimnames=list(name, names(pooled.PRIMEloci)))
    
    for (n in names(pooled.PRIMEloci)) {
      pd <- unfiltered.PRIMEloci[[n]]
      GWAS.olap <- findOverlaps(g, pd, minoverlap=1)
      m[unique(queryHits(GWAS.olap)),n] <- as.numeric(by(pd$score[subjectHits(GWAS.olap)], queryHits(GWAS.olap), max))
    }
    m
  })
names(trait.matrices) <- filtered.traits

```

Calculate GWAS enrichment and recall

```{r}

point.GWAS.df <- do.call("rbind", lapply(names(pooled.PRIMEloci), function(n) {
  pd <- pooled.PRIMEloci[[n]]
  bg.olap <- findOverlaps(SNP.data, pd, minoverlap=1)
  bg.count <- length(unique(queryHits(bg.olap)))
  GWAS.olap <- findOverlaps(GWAS.data, pd, minoverlap=1)
  
  do.call("rbind", lapply(filtered.traits, function(t) {
    
    fg <- which(GWAS.data$trait==t)
    fg.olap <- GWAS.olap[queryHits(GWAS.olap) %in% fg,]
    fg.count <- length(unique(queryHits(fg.olap)))
    
    data.frame(enrichment=(fg.count/length(fg)) / (bg.count/length(SNP.data)),
               recall=fg.count/length(fg),
               fg.count=fg.count,
               bg.count=bg.count,
               fg.n=length(fg),
               bg.n=length(SNP.data),
               sample=n,
               trait=t)
    }))
}))

point.GWAS.df


```


Plot GWAS results

```{r}
point.GWAS.df <- point.GWAS.df %>%
  group_by(trait) %>%
  mutate(scaled_enrichment = rescale(enrichment)) %>%
  ungroup()

point.GWAS.df <- as.data.frame(point.GWAS.df)
enr.mat <- matrix(0,nrow=length(filtered.traits),ncol=length(unique(point.GWAS.df$sample)), dimnames=list(filtered.traits,unique(point.GWAS.df$sample)))
for (i in 1:nrow(point.GWAS.df)) 
  enr.mat[point.GWAS.df[i,"trait"],point.GWAS.df[i,"sample"]] <- point.GWAS.df[i,"scaled_enrichment"]*point.GWAS.df[i,"recall"]

ro <- hclust(dist(log2(enr.mat+1)))$order

point.GWAS.df$sample <- factor(point.GWAS.df$sample, levels=c("K562_C","K562_N","GM12878_C","GM12878_N","HepG2_C","HepG2_N","A549_C","A549_N","HCT116_C"))
point.GWAS.df$trait <- factor(point.GWAS.df$trait, levels=rownames(enr.mat)[ro])

ggplot(point.GWAS.df,aes(sample,trait,fill=scaled_enrichment)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1) +
  xlab("") + ylab("")
ggsave("plots/GWAS_trait_enrichment_vs_cell_line.pdf",width=3,height=5)

```

```{r}

focus.traits <- as.character(unique(subset(point.GWAS.df, enrichment>=10)$trait))

for (trait in focus.traits) {
  
  m <- trait.matrices[[trait]]
  
  print(dim(m))
  
  ro <- hclust(dist(m))$order
  co <- hclust(dist(t(m)))$order
  
  df <- data.frame(score=as.numeric(m),
                   cell.line=rep(colnames(m),each=nrow(m)),
                   variant=rownames(m))
  df$variant <- factor(df$variant,levels=rownames(m)[ro])
  df$cell.line <- factor(df$cell.line, levels=colnames(m)[co])
  
  ggplot(df, aes(cell.line, variant, fill=score)) + geom_tile() + xlab("") + ylab("") + 
    theme(axis.text.x = element_text(angle=90,hjust=1)) + 
    #theme(axis.text.y = element_blank()) + 
    scale_fill_distiller(direction=1) +
    ggtitle(trait)
  ggsave(paste0("plots/",trait,"_variant_PRIMEloci_score_vs_cell_line.pdf"),width=5,height=5)
  
}

```




```{r}
enrichment.precision.recall <- function(data, fg.data, bg.data, assay) {
  
  do.call("rbind", lapply(names(data), function(n) {

    gr <- data[[n]]
    scores <- seq(0,1,by=0.05)
    cnt <- sapply(scores, function(sco) sum(gr$score>=sco))
    scores <- scores[cnt>=10]
    
    gr.fg.olap <- findOverlaps(fg.data, gr, minoverlap=1)
    gr.bg.olap <- findOverlaps(bg.data, gr, minoverlap=1)
    
    do.call("rbind", lapply(scores, function(sco) {
      
      selection <- which(gr$score>=sco)
      fg.count <- length(unique(queryHits(gr.fg.olap)[which(subjectHits(gr.fg.olap) %in% selection)]))
      bg.count <- length(unique(queryHits(gr.bg.olap)[which(subjectHits(gr.bg.olap) %in% selection)]))
      
      data.frame(enrichment=(fg.count/length(fg.data))/(bg.count/length(bg.data)),
                 precision=fg.count/(fg.count+bg.count),
                 recall=fg.count/length(fg.data),
                 fg.count=fg.count,
                 bg.count=bg.count,
                 score=sco,
                 sample=n,
                 assay=assay[n])
    }))
    
  }))
}


run_enrichment_for_trait <- function(trait) {
  enrichment.precision.recall(
    unfiltered.PRIMEloci,
    GWAS.data[GWAS.data$trait == trait,], 
    SNP.data, 
    names(unfiltered.PRIMEloci)
  )
}

per.trait.results <- mclapply(filtered.traits, run_enrichment_for_trait,mc.cores=30)
names(per.trait.results) <- focus.traits 

lapply(seq_along(per.trait.results), function(i) {
  trait_name <- filtered.traits[i]
  data <- per.trait.results[[i]]
  
  # Create plots
  pr <- ggplot(data, aes(recall, enrichment, color = sample)) +
    geom_line(aes(group = sample)) +
    theme(aspect.ratio = 1) +
    xlab(paste0("recall")) +
    ggtitle(paste("Recall vs Enrichment for", trait_name))
  
  
  # Save arranged plots
  ggsave(filename = paste0("plots/cellLinePlots/", trait_name, "_enrichment_recall_plot.pdf"),
         plot = pr, device = "pdf")
})
```







```{r}
# Over pooled traits
#blood.traits <- c("HbA1c","Hb","RBC","MCV","MCH","MCHC")
#hep.traits <- c("ApoB", "TC", "TBil", "SHBG", "LDLC", "TG","Alb")

# Blood-related traits (erythrocytes, platelets, hematology)
blood.traits <- c("HbA1c",  # Glycated hemoglobin, diabetes marker
                  "Hb",      # Hemoglobin, oxygen transport
                  "RBC",     # Red Blood Cell count
                  "MCV",     # Mean Corpuscular Volume, RBC size
                  "MCH",     # Mean Corpuscular Hemoglobin, RBC hemoglobin content
                  "MCHC",    # Mean Corpuscular Hemoglobin Concentration
                  "Plt")     # Platelet count, clotting function

# Immune-related traits (white blood cells, inflammation, autoimmunity)
immune.traits <- c("WBC",    # Total White Blood Cell count
                   "Neutro", # Neutrophils, first-line immune defense
                   "Lym",    # Lymphocytes, adaptive immunity
                   "Mono",   # Monocytes, innate immune response
                   "Eosino", # Eosinophils, allergy and parasite defense
                   "Baso",   # Basophils, involved in histamine response
                   "CRP",    # C-reactive protein, systemic inflammation marker
                   "IBD",    # Inflammatory Bowel Disease, chronic immune disorder
                   "AID_Combined", # Autoimmune diseases combined category
                   "Asthma", # Chronic inflammatory airway disease
                   "Hypothyroidism", # Can be autoimmune (e.g., Hashimoto's)
                   "Fibroblastic_Disorders") # Potential immune involvement

# Liver/metabolic traits (hepatic function, lipids, glucose metabolism)
hep.traits <- c("ApoB",     # Apolipoprotein B, LDL-associated
                "TC",       # Total Cholesterol
                "LDLC",     # Low-Density Lipoprotein Cholesterol (bad cholesterol)
                "TG",       # Triglycerides, fat storage
                "Alb",      # Albumin, major liver protein
                "SHBG",     # Sex Hormone Binding Globulin, linked to liver metabolism
                "TBil",     # Total Bilirubin, liver function indicator
                "ALT",      # Alanine Aminotransferase, liver enzyme
                "AST",      # Aspartate Aminotransferase, liver enzyme
                "GGT",      # Gamma-Glutamyl Transferase, liver and bile function
                "Urea",     # Metabolic waste, partly liver-related
                "Glucose",  # Blood glucose levels, diabetes-related
                "IGF1")     # Insulin-like Growth Factor 1, growth hormone-related

blood.GWAS.data <- subset(GWAS.data, trait %in% blood.traits)
hep.GWAS.data <- subset(GWAS.data, trait %in% hep.traits)
immune.GWAS.data <- subset(GWAS.data, trait %in% immune.traits)


cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data <- enrichment.precision.recall(unfiltered.PRIMEloci, blood.GWAS.data, SNP.data, names(unfiltered.PRIMEloci))

cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data <- enrichment.precision.recall(unfiltered.PRIMEloci, hep.GWAS.data, SNP.data, names(unfiltered.PRIMEloci))

cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data <- enrichment.precision.recall(unfiltered.PRIMEloci, immune.GWAS.data, SNP.data, names(unfiltered.PRIMEloci))


ggplot(cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data, aes(recall, enrichment, color=sample)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)  + xlab(paste0("recall (n=",length(blood.GWAS.data$name),")")) + scale_color_manual(values=col.scheme) + 
  geom_point(data=cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data[cellLines.PRIMEloci.blood.GWAS.enrichment.recall.data$score==0.75,])

ggsave(filename = paste0("plots/cellLinePlots/", "bloodTraits", "_enrichment_recall_plot.pdf"), device = "pdf")


ggplot(cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data, aes(recall, enrichment, color=sample)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)  + xlab(paste0("recall (n=",length(hep.GWAS.data$name),")")) + scale_color_manual(values=col.scheme)+ 
  geom_point(data=cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data[cellLines.PRIMEloci.hep.GWAS.enrichment.recall.data$score==0.75,])


ggsave(filename = paste0("plots/cellLinePlots/", "hepTraits", "_enrichment_recall_plot.pdf"), device = "pdf")


ggplot(cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data, aes(recall, enrichment, color=sample)) + geom_line(aes(group=sample)) +
  theme(aspect.ratio=1)  + xlab(paste0("recall (n=",length(immune.GWAS.data$name),")")) + scale_color_manual(values=col.scheme) + 
  geom_point(data=cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data[cellLines.PRIMEloci.immune.GWAS.enrichment.recall.data$score==0.75,])

ggsave(filename = paste0("plots/cellLinePlots/", "immuneTraits", "_enrichment_recall_plot.pdf"), device = "pdf")

```



```{r}
ggplot(point.GWAS.df[point.GWAS.df$trait%in%blood.traits,],aes(sample,trait,fill=enrichment)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=7) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "bloodTraits", "_enrichment_hm.pdf"), device = "pdf",width=7,height=5)

ggplot(point.GWAS.df[point.GWAS.df$trait%in%blood.traits,],aes(sample,trait,fill=recall)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=2) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "bloodTraits", "_recall_hm.pdf"), device = "pdf",width=7,height=5)


ggplot(point.GWAS.df[point.GWAS.df$trait%in%hep.traits,],aes(sample,trait,fill=enrichment)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=7) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "hepTraits", "_enrichment_hm.pdf"), device = "pdf",width=7,height=5)

ggplot(point.GWAS.df[point.GWAS.df$trait%in%hep.traits,],aes(sample,trait,fill=recall)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=2) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "hepTraits", "_recall_hm.pdf"), device = "pdf",width=7,height=5)


ggplot(point.GWAS.df[point.GWAS.df$trait%in%immune.traits,],aes(sample,trait,fill=enrichment)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=7) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "immuneTraits", "_enrichment_hm.pdf"), device = "pdf",width=7,height=5)

ggplot(point.GWAS.df[point.GWAS.df$trait%in%immune.traits,],aes(sample,trait,fill=recall)) + geom_tile() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) + scale_fill_distiller(direction=1,palette=2) +
  xlab("") + ylab("")

ggsave(filename = paste0("plots/cellLinePlots/", "immuneTraits", "_recall_hm.pdf"), device = "pdf",width=7,height=5)

```


```{r}
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
promoters <- promoters(txdb,upstream=500,downstream=500)
GWAS.data$type <- rep("distal",length(GWAS.data))
olap <- findOverlaps(GWAS.data,promoters,minoverlap = 1)
GWAS.data$type[queryHits(olap)] <- "proximal"
GWAS.df <- as.data.frame(GWAS.data)
GWAS.df$name <- paste0(seqnames(GWAS.data),":",start(GWAS.data)+1,":",end(GWAS.data),":",GWAS.data$RS_id)

trait.loci.olap <- sapply(traits, function(t) {
  ids <- which(GWAS.data$trait==t)
  unique(unlist(lapply(names(unfiltered.PRIMEloci), function(n) {
    pd <- unfiltered.PRIMEloci[[n]]
    GWAS.olap <- findOverlaps(GWAS.data[ids,], pd, minoverlap=1)
    ids[queryHits(GWAS.olap)]
})))
})

trait.matrices <- lapply(c(blood.traits,hep.traits,immune.traits), function(t) {
  
  g <- GWAS.data[trait.loci.olap[[t]],]
  name <- paste0(seqnames(g),":",start(g)+1,":",end(g),":",g$RS_id)
  m <- matrix(0,ncol=length(unfiltered.PRIMEloci),nrow=length(g),dimnames=list(name, names(unfiltered.PRIMEloci)))
  
  for (n in names(unfiltered.PRIMEloci)) {
    pd <- unfiltered.PRIMEloci[[n]]
    GWAS.olap <- findOverlaps(g, pd, minoverlap=1)
    m[unique(queryHits(GWAS.olap)),n] <- as.numeric(by(pd$score[subjectHits(GWAS.olap)], queryHits(GWAS.olap), max))
  }
  m
})
names(trait.matrices) <- c(blood.traits,hep.traits,immune.traits)

for (trait in c(blood.traits,hep.traits,immune.traits)) {
  
  m <- trait.matrices[[trait]]
  
  print(dim(m))
  
  ro <- hclust(dist(m))$order
  #co <- hclust(dist(t(m)))$order
  
  df <- data.frame(score=as.numeric(m),
                   cell.line=rep(colnames(m),each=nrow(m)),
                   variant=rownames(m))
  df$variant <- factor(df$variant,levels=rownames(m)[ro])
  #df$cell.line <- factor(df$cell.line, levels=colnames(m)[co])
  m <- mean(df$score[df$score>0])
  df$score[df$score==0] <- NA
  
  df <- merge(df, GWAS.df[,c("name","type")], by.x="variant",by.y="name")
  
  ggplot(df, aes(cell.line, variant)) + 
      geom_tile(aes(fill=score), color="grey90", size=0.1) +  # Heatmap
      scale_fill_gradient2(low = "#256866", mid = "#FFFFCC", high = "orange", na.value=NA, midpoint=m) +  
      ggnewscale::new_scale_fill() +  # Allows a second independent fill scale
      geom_tile(aes(x=length(unique(df$cell.line))+0.5, fill=type), width=0.2) +  # Annotation strip
      scale_fill_manual(values=c("proximal"="#4daf4a", "distal"="#984ea3")) +  # Discrete for annotation
      xlab("") + 
      ylab("") + 
      theme_bw() +
      theme(axis.text.x = element_text(angle=90, hjust=1),
            axis.text.y = element_text(size=3),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.border = element_rect(color="black", fill=NA, linewidth=1)) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_discrete(expand = c(0, 0)) +  
      ggtitle(trait)

  ggsave(paste0("plots/cellLinePlots/",trait,"_variant_PRIMEloci_score_vs_cell_line.pdf"),width=5,height=5)
  
}

```





