---
title: "20250815_workflow_SNPtracing"
output: html_document
date: "2025-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [1.] Read in data -------------------------

## [1.0.] Identify LD score files 
```{r}
fls <- base::list.files(
      path = base::paste0(
          here::here("11.LDSC_FANTOM5/ldsc_files/"),
          "weights"
        ),
      full.names = TRUE
  ) %>%
  stringr::str_subset("ldscore.gz")
```

## [1.0.] Read in all SNP positions
```{r}
pos <- purrr::map_dfr( # 1,186,879
  .x = fls,
  .f = function(f){
    readr::read_delim(
      file = f,
      delim = "\t",
      col_names = TRUE,
      col_types = readr::cols(
          .default = readr::col_character(),
          BP = readr::col_integer(),
          L2 = readr::col_double()
        )
      )
    }
  ) %>%
  dplyr::rename(
      chr = "CHR",
      start = "BP"
    )
```

## [1.1.] Read in sumstat object to identify anemia trait IDs
#### see [2.0.] in 'ldsc_plotting.Rmd'
#### crucial information: ldsc_results/D61_score0.5.log 
```{r}
sumstat_annot <- readr::read_tsv( # 12,098
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "ukb31063_ldsc_sumstat_manifest_aws_sep2022.tsv"
    )
  )
sumstat_annot <- sumstat_annot %>% #4,170
  dplyr::filter(
      is_primary_gwas == TRUE &
      phenotype %in% base::unique(ldsc_score_$Trait)
    )
```

## [1.2.] Subset for anemia trait IDs
#### D3_OTHERANAEMIA, D3_ANAEMIANAS, D61, D64
```{r}
sumstat_annot %>%
  dplyr::filter(
      stringr::str_detect(description,"anaemia")
    )
```

## [1.3.] Read in anemia related SNPs
#### Diagnoses - main ICD10: D61 Other aplastic anaemias
```{r}
snps <- readr::read_delim( # 1,096,648
  file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "sumstats/D61.ldsc.imputed_v3.both_sexes.tsv.bgz"
    ),
  delim = "\t",
  col_names = TRUE,
  col_types = readr::cols(
      .default = readr::col_character(),
      N = readr::col_integer(),
      Z = readr::col_double()
    )
  )
```

## [1.4.] Merge the data
```{r}
mrg <- dplyr::left_join(
    x = snps,
    y = pos,
    by = "SNP"
  )
```

#### Evaluate
```{r}
dplyr::arrange(snps,dplyr::desc(Z))
dplyr::arrange(mrg,dplyr::desc(Z))
```
##<<bookmark>>
#### Potentially crucial PL scores for mast cells?
```{r}
"/projects/ralab/data/projects/nucleiCAGEproject/0.External_resources/FANTOM5_cellFacet_bedFiles/mast.cell_distal_qn.PL.score0.5.bed"
```

