---
title: "20250815_workflow_SNPtracing"
output: html_document
date: "2025-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [0.] Prerequisites -------------------------

## [0.0.] Load necessary packages
```{r message=FALSE}
base::library("org.Hs.eg.db")
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

# [1.] Read in data -------------------------

## [1.0.] Identify LD score files 
```{r}
fls <- base::list.files(
      path = base::paste0(
          here::here("11.LDSC_FANTOM5/ldsc_files/"),
          "weights"
        ),
      full.names = TRUE
  ) %>%
  stringr::str_subset("ldscore.gz")
```

## [1.0.] Read in all SNP positions
```{r}
pos <- purrr::map_dfr( # 1,186,879
  .x = fls,
  .f = function(f){
    readr::read_delim(
      file = f,
      delim = "\t",
      col_names = TRUE,
      col_types = readr::cols(
          .default = readr::col_character(),
          BP = readr::col_integer(),
          L2 = readr::col_double()
        )
      )
    }
  ) %>%
  dplyr::rename(
      chr = "CHR",
      start = "BP"
    )
```

## [1.1.] Read in sumstat object to identify anemia trait IDs
#### see [2.0.] in 'ldsc_plotting.Rmd'
#### crucial information: ldsc_results/D61_score0.5.log 
```{r}
# Load the data
ldsc_score_ <- readr::read_tsv( # 3,577,860
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "combined_results.tsv"
    )
  )
ldsc_score_$Trait<- base::sub(pattern = "_score[0-9.]+$", "", ldsc_score_$Trait) 

sumstat_annot <- readr::read_tsv( # 12,098
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "ukb31063_ldsc_sumstat_manifest_aws_sep2022.tsv"
    )
  )
sumstat_annot <- sumstat_annot %>% #4,170
  dplyr::filter(
      is_primary_gwas == TRUE &
      phenotype %in% base::unique(ldsc_score_$Trait)
    )
```

## [1.2.] Subset for anemia trait IDs
#### D3_OTHERANAEMIA, D3_ANAEMIANAS, D61, D64
```{r}
sumstat_annot %>%
  dplyr::filter(
      stringr::str_detect(description,"anaemia")
    )
```

#### Identify all four files
```{r}
snps_sel_ <- sumstat_annot %>%
  dplyr::filter(
      stringr::str_detect(phenotype,"D3_OTHERANAEMIA|D3_ANAEMIANAS|D61|D64")
    ) %>%
  dplyr::pull(ldsc_sumstat_file)
```

## [1.3.] Read in anemia related SNPs
#### Diagnoses - main ICD10: D61 Other aplastic anaemias
```{r}
snps <- readr::read_delim( # 1,096,648
  file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "sumstats/D61.ldsc.imputed_v3.both_sexes.tsv.bgz"
    ),
  delim = "\t",
  col_names = TRUE,
  col_types = readr::cols(
      .default = readr::col_character(),
      N = readr::col_integer(),
      Z = readr::col_double()
    )
  )
```

#### Read in anemia related SNPs across all 4 SNP sets
```{r}
snps <- purrr::map_dfr( # 4,386,592 (=4*1,096,648)
  .x = snps_sel_,
  .f = function(f){
    SNPset <- stringr::str_replace(f, "^(.*?)\\.(.*)$", "\\1")
    out <- readr::read_delim( # 1,096,648
      file = base::paste0(
          here::here("11.LDSC_FANTOM5/ldsc_files/"),
          "sumstats/", f
        ),
      delim = "\t",
      col_names = TRUE,
      col_types = readr::cols(
          .default = readr::col_character(),
          N = readr::col_integer(),
          Z = readr::col_double()
        )
      )
    return(
      out %>%
        dplyr::mutate(
          SNPset = SNPset
        )
      )
    }
  )
```

#### Evaluate
```{r}
snps %>%
  distinct(SNP,A1,A2)

snps %>%
  dplyr::group_by(SNPset) %>%
  dplyr::summarise(n = dplyr::n())
```

## [1.4.] Merge the data
```{r}
mrg <- dplyr::left_join(
      x = snps,
      y = pos,
      by = "SNP"
    ) %>%
  dplyr::mutate(
      end = start
    ) %>% 
  dplyr::select(
      chr,start, end, SNPset, SNP, Z, L2
    ) %>%
  tidyr::drop_na()
  # dplyr::filter(
  #   dplyr::if_any(
  #     .cols = tidyselect::everything(),
  #     .fns =  ~ !base::is.na(.x)
  #   )
  # )
```

#### Evaluate
```{r}
dplyr::arrange(snps,dplyr::desc(Z))
dplyr::arrange(mrg,dplyr::desc(Z))
# dplyr::distinct(mrg,N) # N not selected in [1.4.]

GenomicRanges::makeGRangesFromDataFrame(mrg)

mrg %>%
  dplyr::filter(SNP == "rs1000000")
```

## [1.5.] Read in mast cell PRIMEloci
#### "/projects/ralab/data/projects/nucleiCAGEproject/0.External_resources/FANTOM5_cellFacet_bedFiles/mast.cell_distal_qn.PL.score0.5.bed"
```{r}
base::list.files(
      path = here::here("0.External_resources/FANTOM5_cellFacet_bedFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("mast.cell")

mcPL <- readr::read_delim( # 325,054
  file = base::paste0(
      here::here("0.External_resources/"),
      "FANTOM5_cellFacet_bedFiles/",
      "mast.cell_distal_qn.PL.score0.5.bed"
    ),
  delim = "\t",
  col_names = base::c("chr","start","end","PL"),
  col_types = readr::cols(
      .default = readr::col_integer(),
      chr = readr::col_character(),
      PL = readr::col_double()
    )
  )
```

#### Evaluate
```{r}
mcPL %>%
  dplyr::summarise(
      ran = base::range(PL)
    )
utils::tail(mcPL)
utils::tail(mrg)
```

## [1.5.] Read in PRIMEloci for all 5 immune effector cell types (20260227)
```{r}
facets_sel_ <- base::c(
    "mast cell", "dendritic cell", "monocyte",
    "T cell", "reticulocyte"
  )
facets_sel <- stringr::str_replace(facets_sel," ", ".") %>%
  base::paste0(collapse = "|")

fs <- base::list.files(
      path = here::here("0.External_resources/FANTOM5_cellFacet_bedFiles"),
      full.names = TRUE
    ) %>%
  stringr::str_subset(facets_sel) %>%
  stringr::str_subset("qn.PL.score0.5.bed")

icPL <- purrr::map_dfr( # 2,056,763
    .x = fs,
    .f = function(f){
      cell <- stringr::str_replace(f, "^(.*)/(.*?)_(.*)$", "\\2")
      site <- stringr::str_replace(f, "^(.*)/(.*?)_(.*?)_(.*)$", "\\3")
      out <- readr::read_delim(
        file = f,
        delim = "\t",
        col_names = base::c("chr","start","end","PL"),
        col_types = readr::cols(
            .default = readr::col_integer(),
            chr = readr::col_character(),
            PL = readr::col_double()
          )
        )
      return(
        out %>%
          dplyr::mutate(
            cell = cell,
            site = site
          )
      )
    }
  )
```

# [2.] Mast cell PRIMEloci with genetic variants -------------------------

## [2.0.] Identify variants overlapping mast cell PRIMEloci
```{r}
mcPLgr <- GenomicRanges::makeGRangesFromDataFrame(mcPL)
GenomeInfoDb::seqlevelsStyle(mcPLgr) <- "UCSC"
S4Vectors::mcols(mcPLgr)[base::c("PL")] <- mcPL$PL
mcPLgr <- base::sort(mcPLgr)

mrggr <- GenomicRanges::makeGRangesFromDataFrame(mrg)
GenomeInfoDb::seqlevelsStyle(mrggr) <- "UCSC"
S4Vectors::mcols(mrggr)[base::c("SNP","Z","L2")] <- mrg[,base::c("SNP","Z","L2")]
mrggr <- base::sort(mrggr)

ovl <- IRanges::findOverlaps(
    query = mcPLgr,
    subject = mrggr
  )

ovl_1 <- mcPLgr[S4Vectors::queryHits(ovl)] %>%
  tibble::as_tibble() %>%
  dplyr::select(
      seqnames,start,end,PL
    )

ovl_2 <- mrggr[S4Vectors::subjectHits(ovl)] %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      var_pos = base::paste(seqnames,start,sep = "_")
    ) %>% 
  dplyr::select(
      var_pos,SNP,Z,L2
    )

varPL_mc <- dplyr::bind_cols( # 346,340
    ovl_1, ovl_2
  )
```

```{r}
icPLgr <- GenomicRanges::makeGRangesFromDataFrame(icPL) # 2,056,763
GenomeInfoDb::seqlevelsStyle(icPLgr) <- "UCSC"
S4Vectors::mcols(icPLgr)[base::c("PL","cell","site")] <- icPL[,base::c("PL","cell","site")]
icPLgr <- base::sort(icPLgr)

mrggr <- GenomicRanges::makeGRangesFromDataFrame(mrg) # 4,284,612 
GenomeInfoDb::seqlevelsStyle(mrggr) <- "UCSC"
S4Vectors::mcols(mrggr)[base::c("SNP","SNPset","Z","L2")] <- mrg[,base::c("SNP","SNPset","Z","L2")]
mrggr <- base::sort(mrggr)

ovl <- IRanges::findOverlaps(
    query = icPLgr,
    subject = mrggr
  )

ovl_1 <- icPLgr[S4Vectors::queryHits(ovl)] %>%
  tibble::as_tibble() %>%
  dplyr::select(
      seqnames,start,end,PL,cell,site
    )

ovl_2 <- mrggr[S4Vectors::subjectHits(ovl)] %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
      var_pos = base::paste(seqnames,start,sep = "_")
    ) %>% 
  dplyr::select(
      var_pos,SNP,SNPset,Z,L2
    )

varPL <- dplyr::bind_cols( # 1,385,360
    ovl_1, ovl_2
  )
```


```{r}
varPL %>%
  # arrange(desc(Z)) %>%
  dplyr::filter(
      Z > 2 & cell == "mast.cell" & SNPset != "D61" & site == "distal"
    ) %>%
  dplyr::arrange(dplyr::desc(PL))
```

## [2.1.] Assign nearest neighbouring genes
#### CAGEfightR::assignGeneID(varPLgr,txdb)
```{r}
varPLgr <- GenomicRanges::makeGRangesFromDataFrame(varPL)
GenomeInfoDb::seqlevelsStyle(varPLgr) <- "UCSC"
S4Vectors::mcols(varPLgr)[base::c("PL","var_pos","SNP","Z","L2")] <- varPL[,base::c("PL","var_pos","SNP","Z","L2")]
varPLgr <- base::sort(varPLgr)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
orgDb <- org.Hs.eg.db

extGenes_ <- GenomicFeatures::genes(x = txdb) %>%
  IRanges::trim() %>%
  base::sort()
# extGenes <- IRanges::promoters(
#         x = extGenes_,
#         upstream = 0,
#         downstream = 0
#       ) %>%
#   IRanges::trim() %>%
#   base::sort()

TSSs <- IRanges::resize(
    x = extGenes_,
    width = 1,
    fix = "start"
  )
geneNames <- AnnotationDbi::select(
    x = orgDb,
    keys = TSSs$gene_id,
    keytype = "ENTREZID",
    columns = "SYMBOL"
  )
# S4Vectors::mcols(extGenes_) <- geneNames$SYMBOL
S4Vectors::mcols(TSSs)["Gene_symbol"] <- geneNames$SYMBOL

# extGenes_[100:200,]
TSSs[100:200,]

nG <- IRanges::nearest(
    x = varPLgr,
    subject = TSSs
  )

# TSSs[nG,]
varPL[,"Gene_symbol"] <- TSSs[nG,]$Gene_symbol
varPL[,"ENTREZID"] <- TSSs[nG,]$gene_id
dplyr::arrange(varPL,dplyr::desc(PL))
```

```{r}
varPL %>%
  # arrange(desc(Z)) %>%
  dplyr::filter(
      Z > 2 & cell == "mast.cell" & SNPset != "D61" & site == "distal"
    ) %>%
  dplyr::arrange(dplyr::desc(PL))
```

## [2.2.] Subset & evaluate
#### 'selec' object includes 'open targets' related to anemia
#### super enhancer upstream of KLF6: chr10:3,704,497-4,144,825, chr10:3806827-3807227
#### enhancer upstream of KIT: chr4:54635436-54635836, intronic enhancer in KIT: chr4:54667216-54667616
#### intronic enhancer in SOX13
#### intronic enhancer in IL4R and IL1RL2, enhancer upstream of IL6R
```{r}
selec <- base::c(
    "SLC4A1", "HBB", "FANCA",     # not expressed
    "SPTB", "PKLR", "G6PD",       # SptB, G6PD expr
    "RPS19", "FANCI", "SPTA1",    # RPS19 & FANCI expr
    "FANCC", "BRCA2", "FANCD2",   # FANCC, BRCA2, FANCD2
    "BRIP1", "TMPRSS6", "MMACHC", # BRIP1, TMPRSS6 with intronic enhancer(?), MMACHC with downstream enhancer
    "GATA1", "RPL5", "FANCG",     # GATA1 expressed and intronic enhancer, RPL5 expres, FANCG perhaps downstream enhancer
    "FANCE", "SLC46A1", "GPI",    # FANCE, SLC46A1, GPI all expres.
    "RPS26", "PALB2", "MTRR",     # RPS26, PALB2, MTRR all expres.
    "BRCA1"                       # BRCA1 expres.
  )
selec = "KIT"
selec = "BRCA2"

varPL %>%
  dplyr::mutate(
      pos = base::paste0(seqnames,":",start,"-",end)
    ) %>%
  # dplyr::filter(
  #     # seqnames == "chr4" 
  #     Gene_symbol %in% selec
  #   ) %>% 
  dplyr::select(
      pos,PL,var_pos,SNP,Z,L2,Gene_symbol
    ) %>% 
  dplyr::arrange(dplyr::desc(PL))
```

## [2.3.] Save & load varPL
```{r}
base::saveRDS(
    object = varPL,
    file = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/"),
        base::format(base::Sys.time(),"%Y%m%d"),
        "_varPL",
        "_AllImmuneCells","_4AnemiaTraits",
        ".rds"
      ),
    compress = TRUE
  )
```

```{r}
varPL <- base::readRDS(
  file  = base::list.files(
        path = here::here("11.LDSC_FANTOM5_copy"),
        full.names = TRUE
      ) %>%
    stringr::str_subset("rds") %>%
    stringr::str_subset("varPL")
  )
```

# [3.] GO term analysis -------------------------

## [3.0.] Identify mart
#### listMarts() -> listDatasets()
```{r}
biomaRt::useMart(
      biomart = "ENSEMBL_MART_ENSEMBL"
    ) %>% 
  biomaRt::listDatasets() %>%
  dplyr::filter(
      base::grepl('.*sapiens.*',dataset)
    )
GRCh38.p14.mart <- biomaRt::useMart(
    biomart = "ENSEMBL_MART_ENSEMBL",
    dataset = "hsapiens_gene_ensembl"
  )
```

## [3.1.] Retrieve data for all human genes
#### dplyr::distinct(GRCh38.p14.bm,chromosome_name)
```{r}
attr <- base::c(
    # "cdna",
    "ensembl_gene_id",
    "external_gene_name",
    "chromosome_name",
    "start_position",
    "end_position",
    "gene_biotype",
    "ensembl_transcript_id",
    "transcription_start_site",
    "strand"
  )
GRCh38.p14.bm <- biomaRt::getBM(
    mart = GRCh38.p14.mart,
    attributes = attr
  )
```

## [3.2.] Retrieve Entrez-Ensembl-conversions
```{r}
attr <- base::c(
    "ensembl_gene_id",
    "entrezgene_id",
    "gene_biotype"
  )
GRCh38.p14.EnsToEnt <- biomaRt::getBM(
    mart = GRCh38.p14.mart,
    attributes = attr
  )
```

## [3.3.] Data wrangling
```{r}
sel_chrom <- base::c(base::as.character(1:22),"MT","Y","X")

GRCh38.p14.bm.mod <- GRCh38.p14.bm %>%
  dplyr::filter(
      chromosome_name %in% sel_chrom &
      gene_biotype == "protein_coding"
    ) %>%
  dplyr::mutate(
    chromosome_name = base::factor(
        x = chromosome_name,
        levels = sel_chrom,
        ordered = TRUE
      ),
    strand = dplyr::case_when(
        strand == "1" ~ "+",
        strand == "-1" ~ "-"
      ),
    score = "."
  ) %>%
  dplyr::select(
      chr = "chromosome_name",
      start = "start_position",
      end = "end_position",
      ensembl_transcript_id,
      score,
      strand,
      ensembl_gene_id,
      external_gene_name,
      transcription_start_site
    ) %>% 
  dplyr::arrange(chr,start)
```

## [3.4.] Add entrezIds to BDGP6.28.bm.mod
#### Evaluate with dplyr::filter(GRCh38.p14.bm.mod.ent, external_gene_name == "SCNN1D"), should be "6339"
```{r}
GRCh38.p14.bm.mod.ent <- GRCh38.p14.EnsToEnt %>%
  dplyr::filter(
      gene_biotype == "protein_coding"
    ) %>% 
  dplyr::select(-gene_biotype) %>% 
  dplyr::left_join(
      x = GRCh38.p14.bm.mod,
      y = .,
      by = "ensembl_gene_id"
    )
```

## [3.5.] Read out Ids associated with protein coding genes
```{r}
ensUniv <- GRCh38.p14.bm.mod.ent %>%
  dplyr::distinct(ensembl_gene_id) %>% # 20,096
  dplyr::pull(ensembl_gene_id)

entUniv <- GRCh38.p14.bm.mod.ent %>% 
  dplyr::distinct(entrezgene_id) %>% # 19,659
  dplyr::pull(entrezgene_id) %>%
  base::as.character()
```

#### Evaluate with AnnotationDbi::keytypes(org.Hs.eg.db)
```{r}
curOnt = base::c("CC", "MF", "BP")[2]
curGenes <- varPL %>% 
  dplyr::filter(PL >= 0.75) %>%
  dplyr::pull("ENTREZID") # %>%
  # base::as.integer()

allGO <- purrr::map_dfr(
  .x = base::c("CC", "MF", "BP"),
  .f = function(curOnt){
     postGO <- clusterProfiler::enrichGO(
        gene = curGenes,
        keyType = "ENTREZID", # "GENENAME",
        ont = curOnt,
        pAdjustMethod = "BH",
        pvalueCutoff = 1,
        qvalueCutoff = 1,
        OrgDb = orgDb,
        universe = entUniv,
        minGSSize = 0,
        maxGSSize = Inf,
        pool = FALSE
      )
     return(
       postGO@result %>%
          dplyr::mutate(
              ont = curOnt
            )
      )
    }
  )
```

```{r}
sigCutOff = 0.01
allGO %>%
  dplyr::filter(qvalue <= sigCutOff) %>%
  dplyr::arrange(qvalue)
```
