---
title: "FANTOM5 plotting"
date: "2025-06-03"
output: html_document
---

## [-1.0.] Install necessary packages
```{r}
utils::install.packages("factoextra")
utils::install.packages("tsne")
```

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
# base::library(CAGEfightR)
# base::library(GenomicRanges)
# base::library(readr)
# base::library(rtracklayer)
# base::library(parallel)
# base::library(ggplot2)
# base::library(GenomicFeatures)
# base::library(SummarizedExperiment)
# base::library(tibble)
# base::library(ggrepel)
# base::library(dplyr)
# base::library(scales)
# base::library(gridExtra)
# base::library(BuenColors)
# base::library(reshape2)
# base::library(vegan)
# base::library(RColorBrewer)
# base::library(pheatmap)
# base::library(dendsort)
# base::library(tidyr)
# base::library(patchwork)
# base::library(ggdendro)
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.1.] Load pooled cellFacets data (PL score > 0.75)
#### 141 cellFacets of 700 donors
```{r}
cellFacet.rse <- base::readRDS( # cellFacet.rse_1
  file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "cellFacet.rds"
    )
  )
```

## [0.2.] Read in FANTOM5 predictions
```{r}
F5 <- rtracklayer::import.bed(
  con = base::paste0(
      here::here("0.External_resources/"),
      "F5.hg38.enhancers.bed.gz"
    )
  )
```

## [0.3.] Define categories for all cellFacets
```{r}
# cell_type_categories <- base::c(
#   "acute.lymphoblastic.leukemia.T.ALL" = "Cancer / Tumor",
#   "heart" = "Muscle",
#   "acute.myeloid.leukemia.FAB.M6" = "Cancer / Tumor",
#   "embryonic.kidney" = "Stem / Progenitor",
#   "hair.follicle.cell" = "Epithelial",
#   "kidney" = "Epithelial",
#   "melanoma" = "Cancer / Tumor",
#   "giant.cell.carcinoma" = "Cancer / Tumor",
#   "fibroblast.of.gingiva" = "Fibroblast/Connective",
#   "T.cell" = "Immune",
#   "fibroblast.of.choroid.plexus" = "Fibroblast/Connective",
#   "cord.blood.derived" = "Stem / Progenitor",
#   "preadipocyte" = "Fibroblast/Connective",
#   "small.cell.lung.carcinoma" = "Cancer / Tumor",
#   "gingival.epithelial.cell" = "Epithelial",
#   "fibroblast.of.tunica.adventitia.of.artery" = "Fibroblast/Connective",
#   "thyroid.carcinoma" = "Cancer / Tumor",
#   "placental.epithelial.cell" = "Epithelial",
#   "tendon.cell" = "Fibroblast/Connective",
#   "Wilms.tumor" = "Cancer / Tumor",
#   "carcinoid" = "Cancer / Tumor",
#   "fat.cell" = "Fibroblast/Connective",
#   "thymus" = "Immune",
#   "neuroblastoma" = "Cancer / Tumor",
#   "mesothelioma" = "Cancer / Tumor",
#   "meninx" = "CNS/Neural",
#   "eye" = "CNS/Neural",
#   "clear.cell.carcinoma" = "Cancer / Tumor",
#   "lung.adenocarcinoma" = "Cancer / Tumor",
#   "schwannoma" = "CNS/Neural",
#   "choriocarcinoma" = "Cancer / Tumor",
#   "acute.myeloid.leukemia.FAB.M7" = "Cancer / Tumor",
#   "endothelial.cell.of.hepatic.sinusoid" = "Endothelial",
#   "small.intestine" = "Epithelial",
#   "acute.myeloid.leukemia.FAB.M4eo" = "Cancer / Tumor",
#   "smooth.muscle.cell.of.prostate" = "Muscle",
#   "monocyte" = "Immune",
#   "testicular.germ.cell.embryonal.carcinoma" = "Cancer / Tumor",
#   "chondrocyte" = "Fibroblast/Connective",
#   "teratocarcinoma" = "Cancer / Tumor",
#   "urothelial.cell" = "Epithelial",
#   "merkel.cell.carcinoma" = "Cancer / Tumor",
#   "mammary.epithelial.cell" = "Epithelial",
#   "retinal.pigment.epithelial.cell" = "Epithelial",
#   "internal.male.genitalia" = "Other / Mixed",
#   "breast.carcinoma" = "Cancer / Tumor",
#   "hepatocellular.carcinoma" = "Cancer / Tumor",
#   "blood.vessel" = "Endothelial",
#   "colon.carcinoma" = "Cancer / Tumor",
#   "mesothelial.cell" = "Epithelial",
#   "epithelial.cell.of.esophagus" = "Epithelial",
#   "oral.squamous.cell.carcinoma" = "Cancer / Tumor",
#   "hepatocyte" = "Epithelial",
#   "spinal.cord" = "CNS/Neural",
#   "melanocyte" = "CNS/Neural",
#   "neuron" = "CNS/Neural",
#   "neuroectodermal.tumor" = "Cancer / Tumor",
#   "spleen" = "Immune",
#   "rhabdomyosarcoma" = "Cancer / Tumor",
#   "medulloblastoma" = "Cancer / Tumor",
#   "Hep.2" = "Cancer / Tumor",
#   "lung" = "Epithelial",
#   "epidermoid.carcinoma" = "Cancer / Tumor",
#   "large.cell.lung.carcinoma" = "Cancer / Tumor",
#   "testis" = "Other / Mixed",
#   "enteric.smooth.muscle.cell" = "Muscle",
#   "transitional.cell.carcinoma" = "Cancer / Tumor",
#   "tongue" = "Epithelial",
#   "cardiac.myocyte" = "Muscle",
#   "myxofibrosarcoma" = "Cancer / Tumor",
#   "endothelial.cell.of.lymphatic.vessel" = "Endothelial",
#   "neuronal.stem.cell" = "Stem / Progenitor",
#   "acute.myeloid.leukemia.FAB.M5" = "Cancer / Tumor",
#   "leiomyoma" = "Cancer / Tumor",
#   "ductal.cell.carcinoma" = "Cancer / Tumor",
#   "keratinocyte" = "Epithelial",
#   "myoblast" = "Muscle",
#   "epithelioid.sarcoma" = "Cancer / Tumor",
#   "osteosarcoma" = "Cancer / Tumor",
#   "squamous.cell.lung.carcinoma" = "Cancer / Tumor",
#   "brain" = "CNS/Neural",
#   "bronchioalveolar.carcinoma" = "Cancer / Tumor",
#   "acute.myeloid.leukemia.FAB.M4" = "Cancer / Tumor",
#   "epithelial.cell.of.prostate" = "Epithelial",
#   "fibroblast.of.lymphatic.vessel" = "Fibroblast/Connective",
#   "amniotic.epithelial.cell" = "Epithelial",
#   "liposarcoma" = "Cancer / Tumor",
#   "epithelial.cell.of.Malassez" = "Epithelial",
#   "hepatic.stellate.cell" = "Fibroblast/Connective",
#   "corneal.epithelial.cell" = "Epithelial",
#   "gastric.cancer" = "Cancer / Tumor",
#   "squamous.cell.carcinoma" = "Cancer / Tumor",
#   "liver" = "Epithelial",
#   "stromal.cell" = "Fibroblast/Connective",
#   "astrocyte" = "CNS/Neural",
#   "cardiac.fibroblast" = "Fibroblast/Connective",
#   "acute.lymphoblastic.leukemia.B.ALL" = "Cancer / Tumor",
#   "thyroid.gland" = "Epithelial",
#   "gastric.adenocarcinoma" = "Cancer / Tumor",
#   "renal.cell.carcinoma" = "Cancer / Tumor",
#   "lens.epithelial.cell" = "Epithelial",
#   "dendritic.cell" = "Immune",
#   "smooth.muscle.cell.of.trachea" = "Muscle",
#   "macrophage" = "Immune",
#   "pericyte.cell" = "Endothelial",
#   "mast.cell" = "Immune",
#   "gall.bladder.carcinoma" = "Cancer / Tumor",
#   "skin.fibroblast" = "Fibroblast/Connective",
#   "skeletal.muscle.cell" = "Muscle",
#   "respiratory.epithelial.cell" = "Epithelial",
#   "uterus" = "Epithelial",
#   "trabecular.meshwork.cell" = "Fibroblast/Connective",
#   "mesenchymal.cell" = "Fibroblast/Connective",
#   "ciliated.epithelial.cell" = "Epithelial",
#   "epitheloid.carcinoma" = "Cancer / Tumor",
#   "sensory.epithelial.cell" = "Epithelial",
#   "fibroblast.of.periodontium" = "Fibroblast/Connective",
#   "bile.duct.carcinoma" = "Cancer / Tumor",
#   "prostate.cancer" = "Cancer / Tumor",
#   "neutrophil" = "Immune",
#   "kidney.epithelial.cell" = "Epithelial",
#   "glioblastoma" = "Cancer / Tumor",
#   "lymphocyte.of.B.lineage" = "Immune",
#   "chronic.myelogenous.leukemia" = "Cancer / Tumor",
#   "acute.myeloid.leukemia.FAB.M2" = "Cancer / Tumor",
#   "throat" = "Epithelial",
#   "large.intestine" = "Epithelial",
#   "osteoblast" = "Fibroblast/Connective",
#   "serous.adenocarcinoma" = "Cancer / Tumor",
#   "cervical.cancer" = "Cancer / Tumor",
#   "Burkitts.lymphoma" = "Cancer / Tumor",
#   "reticulocyte" = "Immune",
#   "skeletal.muscle.tissue" = "Muscle",
#   "natural.killer.cell" = "Immune",
#   "signet.ring.carcinoma" = "Cancer / Tumor",
#   "vascular.associated.smooth.muscle.cell" = "Muscle",
#   "embryonic.pancreas" = "Stem / Progenitor",
#   "acinar.cell" = "Epithelial",
#   "blood.vessel.endothelial.cell" = "Endothelial",
#   "keratocyte" = "Fibroblast/Connective",
#   "B.lymphoblastoid" = "Immune"
# )
```

## [1.] PCA & k means clustering -------------------------

## [1.0.] Perform PCA
```{r}
SummarizedExperiment::colData(cellFacet.rse)
mat <- SummarizedExperiment::assay(cellFacet.rse,"PL.score") %>%
  base::replace(base::is.na(.), 0) %>%
  base::as.matrix() %>%
  base::t()
mat[1:5,1:5]
base::dim(mat)
pr <- stats::prcomp(mat)
utils::str(pr)
```

## [1.1.] Visualize PCA results
```{r}
factoextra::fviz_pca_ind(
    pr,
    col.ind = "cos2",
    gradient.cols = base::c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE
  )
```

## [1.2.] Visualize PCA results
```{r}
pr$x %>%
  tibble::as_tibble(rownames = "var") %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = PC1,
        y = PC2,
        label = base::ifelse(PC2  > 50, var, "")
      )
    ) +
  geom_point() +
  ggrepel::geom_text_repel(
      box.padding = 0.5,
      max.overlaps = Inf,
      size = 2
    ) +
  scale_x_continuous(
      limits = base::c(-700,700)
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [1.3.] PCA-dependent tSNE analysis
```{r}
tmat <- tsne::tsne(X = pr$x, initial_dims = 2)
base::colnames(tmat) <- base::c("tsne_1","tsne_2")
base::rownames(tmat) <- base::rownames(pr$x)
```

```{r}
factoextra::fviz_nbclust(mat, kmeans, method = "wss")
```

## [1.4.] k means clustering
```{r}
kmat <- stats::kmeans(
    x = mat,
    centers = 7,
    nstart = 25
  )
utils::str(kmat)
```

## [1.5.] Save 'kmat' object
```{r}
base::saveRDS(
  object = kmat,
  file = base::paste0(
      here::here("11.LDSC_FANTOM5_copy/"),
      base::format(base::Sys.Date(), "%Y%m%d"),
      "_F5_clusters.rds"
    )
  )
```

## [1.6.] Data wrangling & combination
```{r}
f5cl <- tibble::tibble(
    cat = base::names(kmat$cluster),
    clus = kmat$cluster
  )
f5mat <- tmat %>%
  tibble::as.tibble(
      rownames = "cat"
    ) %>%
  dplyr::left_join(
      x = ., y = f5cl, by = "cat"
    ) %>%
  dplyr::mutate(
      clus = base::factor(
        x = clus,
        levels = 1:7,
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(clus)
```

## [1.6.] Data visualization
```{r}
ggplot2::ggplot(
      data = f5mat,
      mapping = aes(
        x = tsne_1,
        y = tsne_2,
        colour = clus
      )
    ) +
  geom_point() +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [1.] Data wrangling -------------------------

## [1.0.] Add annotations
```{r}
SummarizedExperiment::rowData(cellFacet.rse)$location <- NA
SummarizedExperiment::rowData(cellFacet.rse)$location[SummarizedExperiment::rowData(cellFacet.rse)$txType %in% base::c("promoter","proximal","fiveUTR")] <- "proximal"
SummarizedExperiment::rowData(cellFacet.rse)$location[SummarizedExperiment::rowData(cellFacet.rse)$txType %in% base::c("intron","intergenic")] <- "distal"
keep <- !base::is.na(SummarizedExperiment::rowData(cellFacet.rse)$location)
cellFacet.rse <- cellFacet.rse[keep,]

F5 <- CAGEfightR::assignTxType(
    object = F5,
    txModels = txdb
  )
F5$location[F5$txType %in% base::c("promoter","proximal","fiveUTR")] <- "proximal"
F5$location[F5$txType %in% base::c("intron","intergenic")] <- "distal"
```

## [1.1.] Calculate support across cellFacets at PL score 0.5
```{r}
SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score")[
    base::is.na(SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score"))
  ] <- 0
cellFacet.rse <- CAGEfightR::calcSupport(
    object = cellFacet.rse,
    inputAssay = "qn.PL.score",
    unexpressed = 0.5,
    outputColumn = "support0.5"
  )
cellFacet.rse <- CAGEfightR::calcSupport(
    object = cellFacet.rse,
    inputAssay = "qn.PL.score",
    unexpressed = 0.75,
    outputColumn = "support0.75"
  )
```

## [1.2.] Calculate number of proximal and distal regions at PL score 0.5 and 0.75
```{r}
SummarizedExperiment::colData(cellFacet.rse)$distal0.5 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse,location == "distal"),
      i = "qn.PL.score"
    ) > 0.5
  )
SummarizedExperiment::colData(cellFacet.rse)$distal0.75 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "distal"),
      i = "qn.PL.score"
    ) > 0.75
  )
SummarizedExperiment::colData(cellFacet.rse)$proximal0.5 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "proximal"),
      i = "qn.PL.score"
    ) > 0.5
  )
SummarizedExperiment::colData(cellFacet.rse)$proximal0.75 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "proximal"),
      i = "qn.PL.score"
    ) > 0.75
  )
```

## [1.2.] Calculate binary jaccard distance between cellFacets at PL score 0.5
```{r}
binary_mat <- SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score") > 0.5

jaccard_dist <- vegan::vegdist(
    x = base::t(binary_mat),
    method = "jaccard"
  )
jaccard_mat <- base::as.matrix(jaccard_dist)

jaccard_sim <- 1 - jaccard_mat
```

## [1.3.] Selecting cellFacets for main figure
#### __Deprecated__
```{r}
# main <- base::names(cell_type_categories)[cell_type_categories != "Cancer / Tumor"]
# 
# jaccard_sim_main <- jaccard_sim[
#     base::row.names(jaccard_sim) %in% main,
#     base::colnames(jaccard_sim) %in% main
#   ]
```

## [1.4.] Cluster and sort
```{r}
hc <- stats::hclust(
    stats::as.dist(1 - jaccard_sim),
    method = "average"
  )
hc_sorted <- dendsort::dendsort(hc)

# hc_main <- stats::hclust(
#     stats::as.dist(1 - jaccard_sim_main),
#     method = "average"
#   )
# hc_main_sorted <- dendsort::dendsort(hc_main)
```

## [2.] Descriptive plots -------------------------

## [2.0.] Plot support across cellFacets at PL score 0.5
#### ggsave("11.LDSC_FANTOM5/plots/support0.5.pdf", height=10,width=7)
```{r}
sps <- cellFacet.rse %>%
  SummarizedExperiment::rowData() %>%
  base::as.data.frame() %>%
  tidyr::pivot_longer(
      cols = base::c("support0.5", "support0.75"),
      names_to = "support",
      values_to = "qn.PL.score"
    ) %>% 
  dplyr::mutate(
    location = base::factor(
        x = location,
        levels = base::c("proximal", "distal"),
        ordered = TRUE
      ),
    support = base::factor(
        x = support,
        levels = base::c("support0.5", "support0.75"),
        ordered = TRUE
      )
    )
```

```{r}
ggplot2::ggplot(
    data = sps,
    mapping = aes(
        x = qn.PL.score,
        fill = location
      )
    )  +
  geom_bar() +
  scale_fill_manual(
      values = base::c("proximal" = "#1b9e77", "distal" = "#d95f02")
    ) +
  facet_wrap(
      facets = support ~ location,
      # ncol = 1,
      scales = "free"
    )

ggplot2::ggplot(
    data = sps,
    mapping = aes(
        x = qn.PL.score,
        # fill = location,
        colour = location
      )
    )  +
  # geom_bar() +
  geom_density() +
  scale_fill_manual(
      values = base::c("proximal" = "#1b9e77", "distal" = "#d95f02")
    ) +
  facet_wrap(
      # facets = support ~ location,
      facets = . ~ support,
      # ncol = 1,
      scales = "free"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.1.] Number of RE at PL score 0.5 clustered based on jaccard distance
#### see [1.3.] above for 'hc_sorted' object
#### Curated clusters: 1:29; 30:73; 74:85; 86:91; 92:107; 108:130; 131:141
```{r}
ordered_samples <- hc_sorted$labels[hc_sorted$order]

# Filter colData to include only those in clustering
df <- SummarizedExperiment::colData(cellFacet.rse) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      sample %in% ordered_samples
    )

# Get dendrogram data
dend <- stats::as.dendrogram(hc_sorted)
dend_data <- ggdendro::dendro_data(dend)
label_map <- stats::setNames(dend_data$labels$x, dend_data$labels$label)
seg_data <- dend_data$segments

# Reshape data to long format for plotting and assign dendrogram x positions
df_long <- df %>%
  tidyr::pivot_longer(
      cols = base::c("distal0.5", "proximal0.5"),
      names_to = "region_type",
      values_to = "count"
    ) %>%
  dplyr::mutate(
      region_type = base::factor(
          x = region_type,
          levels = base::c("proximal0.5", "distal0.5")
        ),
      x = label_map[as.character(sample)]
    )
```

## [2.2.] Curate clusters manually based on hierachical clustering
#### Wilms tumor - kidney cancer
#### clus1: pancreas, fibroblasts, endothelial, bones, lung, liver, retina & cornea, epithelial, epidermis, smooth muscle
#### clus2: kidney, colon, stomach, liver, fibroblast, epithelia, smooth & sceletal muscle, mesothelium, fat, thyroid, gums & mouth, epidermis, ovary, urinary tract, gall bladder cervics, lung, prostate, glia, nerves, pancreas, bones
#### clus3: blood vessel, skeletal & smooth muscle, skin fibroblast,
```{r}
type <- "PPPPPPPPPPPPPPPCPPPPPCCPCPCCCPCPCPPPPPPPPPCCPPCCCCCCCCCCCCCCCCCCPCCCCCCPPPPCPPPPPPPPPCCCCCCPPPPPPPPPPPPPPPPCCCCPCCCCCCPCPPPPCPPPPPPPPPPPPPCPP"
stringr::str_length(type)

curclus <- tibble::tibble(
    cat = hc_sorted$labels[hc_sorted$order],
    comb = base::c(
        base::rep("1_P",29-1+1),
        base::rep("2_P",46-30+1),
        base::rep("3_C",71-47+1),
        base::rep("4_P",85-72+1),
        base::rep("5_C",91-86+1),
        base::rep("6_P",107-92+1),
        base::rep("7_C",125-108+1),
        base::rep("8_P",141-126+1)
      )
    ) %>%
  tidyr::separate(
      col = comb, into = base::c("clus","bio")
    )

# curclus <- tibble::tibble(
#   cat = hc_sorted$labels[hc_sorted$order],
#   clus = base::c(
#       base::rep("1",26-1+1),
#       base::rep("2",73-27+1),
#       base::rep("3",91-74+1),
#       # base::rep("3",82-74+1),
#       # base::rep("4",91-83+1),
#       base::rep("4",107-92+1),
#       base::rep("5",125-108+1),
#       base::rep("6",141-126+1)
#     )
#   )

curmat <- tmat %>%
  tibble::as_tibble(
      rownames = "cat"
    ) %>%
  dplyr::left_join(
      x = ., y = curclus, by = "cat"
    ) %>%
  dplyr::mutate(
      cp = stringr::str_split_1(type,"")
    )

curpca <- pr$x %>%
  tibble::as_tibble(rownames = "cat") %>%
   dplyr::left_join(
      x = ., y = curclus, by = "cat"
    ) %>%
  dplyr::mutate(
      cp = stringr::str_split_1(type,"")
    )
```

```{r}
ggplot2::ggplot(
      data = curmat,
      mapping = aes(
        x = tsne_1,
        y = tsne_2,
        colour = cp, #bio, #clus,
        label = base::ifelse(clus == "4", cat, "")
      )
    ) +
  geom_point(
      size = 3,
      alpha = 0.5
    ) +
  # scale_color_manual(
  #     limits = base::as.character(1:8),
  #     values = base::c("#588b8b", "#f94144" , "#f28f3b" ,"#95e214" ,"#2d3047" ,"#c86bfa", "#90be6d", "#ffb703")
  #   ) +
  scale_color_manual(
      values = base::c("#2d3047" ,"#c86bfa", "#90be6d", "#ffb703")
    ) +
  # ggrepel::geom_text_repel(
  #     box.padding = 0.5,
  #     max.overlaps = Inf,
  #     size = 2
  #   ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_tSNE",
        # "_TumPhysClusters",
        "_TumorPhysio",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 5,
    height = 4
  )
```

#### 1:6 clusters: base::c("#588b8b", "#f94144" , "#f28f3b" ,"#95e214" ,"#2d3047" ,"#c86bfa")
```{r}
ggplot2::ggplot(
      data = curpca,
      mapping = aes(
        x = PC1,
        y = PC2,
        color = cp, #bio, #clus,
        label = base::ifelse(clus == "4", cat, "")
      )
    ) +
  geom_point(
      size = 3,
      alpha = 0.5
    ) +
  # ggrepel::geom_text_repel(
  #     box.padding = 0.5,
  #     max.overlaps = Inf,
  #     size = 2
  #   ) +
  # scale_color_manual(
  #     limits = base::as.character(1:8),
  #     values = base::c("#588b8b", "#f94144" , "#f28f3b" ,"#95e214" ,"#2d3047" ,"#c86bfa", "#90be6d", "#ffb703")
  #   ) +
  scale_color_manual(
      values = base::c("#2d3047" ,"#c86bfa")
    ) +
  # scale_x_continuous(
  #     limits = base::c(-700,700)
  #   ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```


```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_PCA",
        # "_TumPhysClusters",
        "_TumorPhysio",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 5,
    height = 4
  )
```

## [2.1.] Combined plot
#### ggsave("11.LDSC_FANTOM5/plots/nBarplot.pdf", combined_plot, width = 7, height = 15)
```{r}
bar_plot <- ggplot2::ggplot(
    data = df_long,
    mapping = aes(
        x = x,
        y = count,
        fill = region_type
      )
    ) +
  geom_bar(
      stat = "identity", position = "dodge"
    ) +
  geom_hline(
      yintercept = base::length(F5),
      linetype = "dashed",
      color = "darkred",
      linewidth = 0.7
    ) +
  geom_hline(
      yintercept = 100000,
      linetype = "dashed",
      color = "darkgrey",
      linewidth = 0.7
    ) +
  scale_x_continuous(
      breaks = df_long$x,
      labels = df_long$sample
    ) +
  scale_fill_manual(
      values = base::c("proximal0.5" = "#1b9e77", "distal0.5" = "#d95f02")
    ) +
  coord_flip() +
  labs(
      x = NULL,
      y = "Region count",
      fill = "Region Type"
    ) +
  theme_minimal() +
  theme(
      axis.text.y = element_text(size = 6, hjust = 1),
      axis.title.y = element_blank(),
      panel.grid.major.y = element_blank(),  # remove horizontal lines
      panel.grid.minor.y = element_blank()   # just in case
    )

# Dendrogram plot 
dend_plot <- ggplot2::ggplot(
      data = seg_data,
      mapping = aes(
        x = x,
        y = y,
        xend = xend,
        yend = yend
      )
    ) +
  geom_segment() +
  scale_y_reverse() +
  coord_flip() +
  theme_void() +
  theme(
      plot.margin = margin(r = -10)
    )  # Reduce spacing to snugly align with barplot

# Combine and plot
library(patchwork)
combined_plot <- dend_plot + bar_plot + patchwork::plot_layout(widths = c(2, 1))
combined_plot
```


```{r}
ggplot2::ggplot(
    data = df_long,
    mapping = aes(
        x = x,
        y = count,
        fill = region_type
      )
    ) +
  geom_bar(
      stat = "identity", position = "dodge"
    ) +
  geom_hline(
      yintercept = base::length(F5),
      linetype = "dashed",
      color = "darkred",
      linewidth = 0.7
    ) +
  geom_hline(
      yintercept = 100000,
      linetype = "dashed",
      color = "darkgrey",
      linewidth = 0.7
    ) +
  scale_x_continuous(
      breaks = df_long$x,
      labels = df_long$sample
    ) +
  scale_fill_manual(
      values = base::c("proximal0.5" = "#1b9e77", "distal0.5" = "#d95f02")
    ) +
  coord_flip() +
  labs(
      x = NULL,
      y = "Region count",
      fill = "Region Type"
    ) +
  theme_minimal() +
  theme(
      axis.text.y = element_text(size = 6, hjust = 1),
      axis.title.y = element_blank(),
      panel.grid.major.y = element_blank(),  # remove horizontal lines
      panel.grid.minor.y = element_blank()   # just in case
    )
```

```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_nBarplot.pdf"
      ),
    device = "pdf",
    plot = combined_plot,
    width = 7,
    height = 15
  )

ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_dendogramOnly.pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 10,
    height = 15
  )
```

#### __Deprecated__
```{r}
# # Same for main figure
# ordered_samples <- hc_main_sorted$labels[hc_main_sorted$order]
# 
# # Filter colData to include only those in clustering
# df <- SummarizedExperiment::colData(cellFacet.rse) %>% 
#   base::as.data.frame() %>%
#   tibble::as_tibble() %>%
#   dplyr::filter(sample %in% ordered_samples)
# 
# # Get dendrogram data
# dend <- stats::as.dendrogram(hc_main_sorted)
# dend_data <- ggdendro::dendro_data(dend)
# label_map <- stats::setNames(dend_data$labels$x, dend_data$labels$label)
# seg_data <- dend_data$segments
# 
# # Reshape data to long format for plotting and assign dendrogram x positions
# df_long <- df %>%
#   tidyr::pivot_longer(
#       cols = base::c("distal0.5", "proximal0.5"),
#       names_to = "region_type",
#       values_to = "count"
#     ) %>%
#   dplyr::mutate(
#       region_type = base::factor(
#           x = region_type,
#           levels = base::c("proximal0.5", "distal0.5"),
#           ordered = TRUE
#         ),
#       x = label_map[base::as.character(sample)]
#     )
```

#### __Deprecated__
```{r}
# # Bar plot 
# bar_plot <- ggplot2::ggplot(
#     data = df_long,
#     mapping = aes(
#         x = x,
#         y = count,
#         fill = region_type
#       )
#     ) +
#   geom_bar(
#       stat = "identity",
#       position = "dodge"
#     ) +
#   geom_hline(
#       yintercept = base::length(F5),
#       linetype = "dashed",
#       color = "darkred",
#       linewidth = 0.7
#     ) +
#   geom_hline(
#       yintercept = 100000,
#       linetype = "dashed",
#       color = "darkgrey",
#       linewidth = 0.7
#     ) +
#   scale_fill_manual(
#       values = base::c("proximal0.5" = "#1b9e77", "distal0.5" = "#d95f02")
#     ) +
#   scale_x_continuous(
#       breaks = df_long$x,
#       labels = df_long$sample
#     ) +
#   coord_flip() +
#   labs(
#       x = NULL,
#       y = "Region count",
#       fill = "Region Type"
#     ) +
#   theme_minimal() +
#   theme(
#         axis.text.y = element_text(size = 6, hjust = 1),
#         axis.title.y = element_blank(),
#       panel.grid.major.y = element_blank(),  # remove horizontal lines
#       panel.grid.minor.y = element_blank()   # just in case
#     )
# 
# # Dendrogram plot 
# dend_plot <- ggplot2::ggplot(
#       data = seg_data
#     ) +
#   geom_segment(
#     mapping = aes(
#         x = x,y = y, xend = xend, yend = yend)
#     ) +
#   scale_y_reverse() +
#   coord_flip() +
#   theme_void() +
#   theme(
#       plot.margin = margin(r = -10)
#     )  # Reduce spacing to snugly align with barplot
# 
# # Combine and plot
# combined_plot <- dend_plot + bar_plot + plot_layout(widths = c(1, 3))
# combined_plot
```

## [2.] LDSC analysis plots -------------------------

## [2.0.] Read in main LDSC results
```{r}
# Load the data
ldsc_score_ <- readr::read_tsv( # 3,577,860
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "combined_results.tsv"
    )
  )
ldsc_score_$Trait<- base::sub(pattern = "_score[0-9.]+$", "", ldsc_score_$Trait) 

sumstat_annot <- readr::read_tsv( # 12,098
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "ukb31063_ldsc_sumstat_manifest_aws_sep2022.tsv"
    )
  )
sumstat_annot <- sumstat_annot %>% #4,170
  dplyr::filter(
      is_primary_gwas == TRUE &
      phenotype %in% base::unique(ldsc_score_$Trait)
    )
```

#### D3_OTHERANAEMIA, D3_ANAEMIANAS, D61, D64
```{r}
sumstat_annot %>%
  dplyr::filter(
    stringr::str_detect(description,"anaemia")
  )
```

#### Evaluate scaling
```{r}
ldsc_score_ %>%
  dplyr::arrange(dplyr::desc(Prop._h2)) %>%
  dplyr::reframe(
      ran = base::range(Prop._h2),
      ran_SNP = base::range(Prop._SNPs)
    )

ldsc_score_ %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = Prop._SNPs
      )
    ) +
  geom_density() +
  theme_bw() +
  theme(
    aspect.ratio = 1
  )
```

## [2.1.] Add description to data
#### 25020 "unknown"  locations
```{r}
ldsc_score <- dplyr::left_join(
    x = ldsc_score_,
    y = dplyr::select(sumstat_annot, phenotype, description),
    by = base::c("Trait" = "phenotype")
  )

ldsc_score <- ldsc_score %>%
  dplyr::mutate(
      loc = stringr::str_extract(Category, "distal|proximal"),
      loc = tidyr::replace_na(loc, "unknown")
    )

ldsc_score$score <- base::sub(".*_(\\d+\\.\\d+)L2_0$", "\\1", ldsc_score$Category)

ldsc_score$sample <- base::sub("_.*", "", ldsc_score$Category)

ldsc_score <- ldsc_score[ldsc_score$Category != "baseL2_0",] # 3,552,840
# ldsc_data <- ldsc_score[ldsc_score$score == "0.5",] # 592,140
```

## [2.2.] Read out minimum scores
#### For cellFacets with over 100K regions
```{r}
min_top_scores <- base::apply(
  X = SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score")[
      SummarizedExperiment::rowData(cellFacet.rse)$location == "distal",
    ],
  MARGIN = 2,
  FUN = function(x) {
    sorted_vals <- sort(x, decreasing = TRUE)
    if (length(sorted_vals) < 100000) {
        min(sorted_vals)  # If fewer than 100k rows, take the smallest of all
      } else {
        min(sorted_vals[1:100000])  # Min of top 100,000
      }
    }
  )
```

## [2.3.] Prepare description-sample combinations for plotting
```{r}
sub <- tibble::tibble(
  description = base::c(
      "Reticulocyte count",
      "Reticulocyte count",
      "Monocyte count",
      "Monocyte count",
      "Basophill count",
      "Basophill count",
      "Urate (umol/L)",
      "Urate (umol/L)"
    ),
  sample = base::c(
      "reticulocyte",
      "F5",
      "monocyte",
      "F5",
      "mast.cell",
      "F5",
      "kidney",
      "F5"
    )
  )
```

#### evaluate overlap: ldsc_score %>% dplyr::filter(sample %in%  ordered_samples[108:125]) %>% distinct(sample)
```{r}
des <- base::factor(
      x = base::rep(trai,2),
      levels = trai,
      ordered = TRUE
    ) %>%
  base::sort()

sps <- base::c(
  base::rbind(
      ordered_samples[108:125],
      base::rep("F5",base::length(ordered_samples[108:125]))
    )
  )

sub <- tibble::tibble(
    description = des,
    sample = sps
  )
```

#### see [2.6.] further down for trai
#### base::c("reticulocyte", "mast.cell")
```{r}
# cur_trai <- trai[4:17] # all plotted traits
# cur_trai <- trai[5:17] # reticulocyte without # Nucleated red blood cell count
cur_trai <- trai[27:30]
des <- base::factor(
      x = base::rep(cur_trai,2),
      levels = cur_trai,
      ordered = TRUE
    ) %>%
  base::sort()

# ct <- "reticulocyte"
ct <- "mast.cell"
sps <- base::rep(base::c(ct,"F5"),base::length(cur_trai))

sub <- tibble::tibble(
    description = des,
    sample = sps
  )
```

## [2.3.] Prepare proportion of heritability and enrichment data for description-sample combinations
```{r}
thresh <- function(score, min_score){
  # base::cat(score," ", min_score)
  if (!base::is.na(min_score) && min_score > 0.5) {
      threshold <- base::floor((min_score - 1e-8) * 20) / 20  # round down to nearest 0.05
      base::as.numeric(base::as.character(score)) >= threshold
    } else {
      TRUE
    }
  }

df <- dplyr::inner_join(
    x = ldsc_score,
    y = sub,
    by = base::colnames(sub)
  )

f5sum <- df %>%
  dplyr::filter(sample == "F5") %>%
  dplyr::group_by(description) %>%
  dplyr::summarize(
      Prop._h2 = base::mean(Prop._h2, na.rm = TRUE),
      Enrichment = base::mean(Enrichment, na.rm = TRUE),
      Enrichment_p = base::min(Enrichment_p, na.rm = TRUE),
      sample = "F5",
      .groups = "drop"
    )

dfsum <- df %>%
  dplyr::filter(sample != "F5") %>%
  dplyr::group_by(description,sample) %>%
  dplyr::mutate(
      min_score = min_top_scores[sample],
      fil_score = purrr::pmap_lgl(
        .l = base::list(score, min_score),
        .f = .GlobalEnv$thresh
      )
    ) %>%
  dplyr::filter(fil_score == TRUE)

comb <- dplyr::bind_rows(f5sum, dfsum)
```

## [2.4.] Visualize data
```{r}
 ggplot2::ggplot(
    data = comb,
    mapping = aes(
        x = Prop._h2,
        y = Enrichment,
        color = sample,
        size = (-1)*base::log10(Enrichment_p),
        label = score
      )
    ) +
  geom_point(
      alpha = 0.5
    ) +
  ggrepel::geom_text_repel(
      size = 3,
      show.legend = FALSE
    ) +
  scale_color_manual(
      values = base::c("#1f78b4" ,"#c86bfa", "#90be6d", "#ffb703", "#2d3047")
    ) +
  facet_wrap(
      facets = ~ description
      # nrow = 1
    ) +
  theme_bw() +
  theme(
      strip.text.x = element_text(size = 5),
      aspect.ratio = 1,
      panel.spacing = grid::unit(0.1, "lines")
    )
```

```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_enrichment_proportion",
        "_mastcell",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 8,
    height = 6
  )
```

## [2.5.] Scale enrichment and prepare visualisation
#### Subset data for PRIMEloci threshold of 0.5
```{r}
scaling_individual <- function(Enrichment,max_enr,min_enr){
  base::ifelse(
      test = max_enr != min_enr,
      yes = (Enrichment - min_enr) / (max_enr - min_enr),
      no = 0
    )
  }

scaling_group <- function(df){
  max_enr <- base::max(df$Enrichment, na.rm = TRUE)
  min_enr <- base::min(df$Enrichment, na.rm = TRUE)
  
  max_prop <- base::max(df$Prop._h2, na.rm = TRUE)
  min_prop <- base::min(df$Prop._h2, na.rm = TRUE)
  
  return(
    df %>%   
      dplyr::mutate(
        Enrichment_scaled = purrr::map_dbl(
            .x = Enrichment,
            .f = scaling_individual,
            max_enr = max_enr,
            min_enr = min_enr
          ),
        # Proportion_scaled = (Prop._SNPs - min_prop) / (max_prop - min_prop)
        Proportion_scaled = (Prop._h2 - min_prop) / (max_prop - min_prop)
      )
    )
  }

ldsc_data <- ldsc_score %>% # 592,140
  dplyr::mutate(
      Category = stringr::str_replace(Category, "(.*?)_.*$", "\\1")
    ) %>%
  dplyr::filter(
      score == "0.5"
    ) %>%
  dplyr::group_by(Trait) %>%
  tidyr::nest() %>% 
  dplyr::mutate(
      newData = purrr::map(
        .x = data,
        .f = scaling_group
      )
    ) %>%
  dplyr::select(-data) %>% 
  tidyr::unnest(cols = newData) %>%
  dplyr::ungroup()
```

#### Global scaling of 'Prop._h2'
```{r}
rangG <- ldsc_score %>% # 592,140
  dplyr::mutate(
      Category = stringr::str_replace(Category, "(.*?)_.*$", "\\1")
    ) %>%
  dplyr::filter(
      score == "0.5"
    ) %>%
  dplyr::pull(Prop._h2) %>%
  base::range()

scaling_individual <- function(Enrichment,max_enr,min_enr){
  base::ifelse(
      test = max_enr != min_enr,
      yes = (Enrichment - min_enr) / (max_enr - min_enr),
      no = 0
    )
  }

scaling_group <- function(df){
  max_enr <- base::max(df$Enrichment, na.rm = TRUE)
  min_enr <- base::min(df$Enrichment, na.rm = TRUE)
  
  return(
    df %>%   
      dplyr::mutate(
        Enrichment_scaled = purrr::map_dbl(
            .x = Enrichment,
            .f = scaling_individual,
            max_enr = max_enr,
            min_enr = min_enr
          ),
        Proportion_scaled = (Prop._h2 - rangG[1]) / (rangG[2] - rangG[1])
      )
    )
  }

ldsc_data <- ldsc_score %>% # 592,140
  dplyr::mutate(
      Category = stringr::str_replace(Category, "(.*?)_.*$", "\\1")
    ) %>%
  dplyr::filter(
      score == "0.5"
    ) %>%
  dplyr::group_by(Trait) %>%
  tidyr::nest() %>% 
  dplyr::mutate(
      newData = purrr::map(
        .x = data,
        .f = scaling_group
      )
    ) %>%
  dplyr::select(-data) %>% 
  tidyr::unnest(cols = newData) %>%
  dplyr::ungroup()
```


## [2.6.] Define groups of cellFacets to include into the visualization
```{r}
# cat <- base::c(
#     "B.lymphoblastoid", "blood.vessel.endothelial.cell", "brain",
#     "F5", "kidney", "liver",
#     "monocyte", "neutrophil", "reticulocyte",
#     "skeletal.muscle.tissue", "spleen", "T.cell",
#     "thymus"
#   )

# cat <- base::c(
#     # Immune cells
#     "B.lymphoblastoid", "cord.blood.derived", "dendritic.cell",
#     "lymphocyte.of.B.lineage", "macrophage", "mast.cell",
#     "monocyte", "natural.killer.cell", "neutrophil",
#     "reticulocyte", "T.cell",
#     # Immune-related tissues
#     "spleen", "thymus", "blood.vessel", "blood.vessel.endothelial.cell",
#     # Include FANTOM5 for comparison
#     "F5"
#   )

# cat <- base::c(
#     ordered_samples[108:125], "cord.blood.derived", "dendritic.cell",
#     "mast.cell", "monocyte", "reticulocyte",
#     "T.cell", "spleen", "thymus",
#     "blood.vessel", "blood.vessel.endothelial.cell", "F5"
#   )

cat <- base::c(
    "mast.cell", "dendritic.cell", "monocyte",
    "T.cell", "reticulocyte", base::rev(ordered_samples[108:125]),
    "F5"
  )
```

## [2.7.] Define groups of traits to include into the visualization
#### keywords for determined search: "inflam","haemorrhage","lymph","rheuma"
#### 'sub' includes all relevant immune / disease traits, 'selec' includes a selected group of traits showing strong differences between samples
```{r}
# trai <- base::c(
#     "Alanine aminotransferase (U/L)", "Albumin (g/L)", "Apoliprotein A (g/L)",
#     "Aspartate aminotransferase (U/L)", "Body mass index (BMI)", "Creatinine (umol/L)",
#     "Diastolic blood pressure, automated reading", "Eosinophill count",
#     "Ever depressed for a whole week", "Getting up in morning", "High light scatter reticulocyte count",
#     "Mean corpuscular volume", "Monocyte count", "Reticulocyte count",
#     "SHBG (nmol/L)", "Urate (umol/L)", "White blood cell (leukocyte) count"
#   )

# trai <- base::c(
#     "White blood cell (leukocyte) count", "Red blood cell (erythrocyte) count", "Haemoglobin concentration",
#     "Mean corpuscular volume",
#     "Mean corpuscular haemoglobin", "Mean corpuscular haemoglobin concentration",
#     "Red blood cell (erythrocyte) distribution width",
#     "Platelet count", "Mean platelet (thrombocyte) volume", "Platelet distribution width",
#     "Lymphocyte count", "Monocyte count",
#     "Neutrophill count",
#     "Eosinophill count", "Basophill count",
#     "Mean reticulocyte volume", "Mean sphered cell volume", "Immature reticulocyte fraction",
#     "Reticulocyte count", "High light scatter reticulocyte count"
#   )

# sub <- base::c(
#     2386:2391, 2393:2417, 2529:2530,
#     3002:3016, 3020:3022, 3024:3033,
#     3034:3192, 3292, 3453:3473,
#     3512, 3514, 3516, 3521:3551,
#     3708:3709, 3774:3775, 3986
#   )
# trai <- ldsc_score %>%
#   dplyr::distinct(description) %>%
#   dplyr::filter(
#       dplyr::row_number() %in% sub
#     ) %>%
#   dplyr::pull(description)
  
selec <- base::c(
    "White blood cell (leukocyte) count",
    "Lymphocyte count",
    "Lymphocyte percentage",
    
    "Mean corpuscular haemoglobin",
    "Mean corpuscular haemoglobin concentration",
    "Mean corpuscular volume",
    "Mean sphered cell volume",
    "Nucleated red blood cell count",
    "Red blood cell (erythrocyte) count",
    "Red blood cell (erythrocyte) distribution width",
    "Reticulocyte count",
    "Reticulocyte percentage",
    "Mean reticulocyte volume",
    "Immature reticulocyte fraction",
    "High light scatter reticulocyte count",
    "High light scatter reticulocyte percentage",
    
    "Mean platelet (thrombocyte) volume",
    "Platelet count",
    "Platelet crit",
    "Platelet distribution width",
    "Neutrophill count",
    "Neutrophill percentage",
    "Monocyte count",
    "Monocyte percentage",
    "Eosinophill count",
    "Eosinophill percentage",
   
    "Other anaemias",
    "Other and unspecified anaemias",
    "Diagnoses - main ICD10: D61 Other aplastic anaemias",
    "Diagnoses - main ICD10: D64 Other anaemias"
    
    # "Secondary polycythaemia",
    # "Suggestive for eosinophilic asthma",
    # "Thrombocytopenia, unspecified",
    # 
    # "Aortic aneurysm",
    # "Diagnoses - main ICD10: I71 Aortic aneurysm and dissection",
    # "Diagnoses - main ICD10: I77 Other disorders of arteries and arterioles",
    # "Diagnoses - main ICD10: I78 Diseases of capillaries",
    # 
    # "Lymphomas",
    # "Hodgkin lymphoma",
    # "Diagnoses - main ICD10: C81 Hodgkin's disease",
    # "Diagnoses - main ICD10: C91 Lymphoid leukaemia",
    # "Lymphoid leukaemia",
    # "Primary lymphoid and hematopoietic malignant neoplasms",
    # "Multiple myeloma and malignant plasma cell neoplasms",
    # "Secondary and unspecified malignant neoplasm of lymph nodes",
    # "Diagnoses - main ICD10: C77 Secondary and unspecified malignant neoplasm of lymph nodes",
    # "Underlying (primary) cause of death: ICD10: C90.0 Multiple myeloma",
    # "Diagnoses - main ICD10: C90 Multiple myeloma and malignant plasma cell neoplasms",
    # "Underlying (primary) cause of death: ICD10: C92.0 Acute myeloid leukaemia",
    # 
    # "Arterial embolism and thrombosis",
    # "Diagnoses - main ICD10: I74 Arterial embolism and thrombosis",
    # 
    # "Crohns disease of large intestine",
    # 
    # "Bronchitis",
    # "Certain infectious and parasitic diseases",
    # "Diagnoses - main ICD10: A87 Viral meningitis",
    # "Diagnoses - main ICD10: J22 Unspecified acute lower respiratory infection",
    # "Diagnoses - main ICD10: J31 Chronic rhinitis, nasopharyngitis and pharyngitis",
    # 
    # "Diagnoses - main ICD10: J46 Status asthmaticus",
    # "Diagnoses - main ICD10: M05 Seropositive rheumatoid arthritis",
    # "Diagnoses - main ICD10: R59 Enlarged lymph nodes",
    # 
    # "Diagnoses - main ICD10: J33 Nasal polyp",
    # 
    # "Diagnoses - main ICD10: D16 Benign neoplasm of bone and articular cartilage",
    # "Diagnoses - main ICD10: D24 Benign neoplasm of breast",
    # "Diagnoses - main ICD10: C02 Malignant neoplasm of other and unspecified parts of tongue",
    # "Diagnoses - main ICD10: C09 Malignant neoplasm of tonsil",
    # "Diagnoses - main ICD10: C22 Malignant neoplasm of liver and intrahepatic bile ducts",
    # "Diagnoses - main ICD10: C44 Other malignant neoplasms of skin",
    # "Diagnoses - main ICD10: C45 Mesothelioma",
    # "Diagnoses - main ICD10: D25 Leiomyoma of uterus",
    # "Diagnoses - main ICD10: C53 Malignant neoplasm of cervix uteri",
    # "Diagnoses - main ICD10: C67 Malignant neoplasm of bladder",
    # "Secondary malignant neoplasm of respiratory and digestive organs",
    # "Malignant neoplasm of stomach",
    # "Malignant neoplasm of colon",
    # "Diagnoses - main ICD10: C20 Malignant neoplasm of rectum",
    # "Diagnoses - main ICD10: C69 Malignant neoplasm of eye and adnexa",
    # "Diagnoses - main ICD10: C71 Malignant neoplasm of brain",
    # "Malignant neoplasm of brain",
    # "Malignant neoplasm of eye, brain and central nervous system",
    # "Diagnoses - main ICD10: C73 Malignant neoplasm of thyroid gland",
    # "Malignant neoplasm of endocrine gland",
    # "Melignant neoplasm of mesothelium and soft tissue",
    # "Malignant neoplasm of other connective and soft tissue"
  )

trai <- ldsc_score %>%
  dplyr::distinct(description) %>%
  dplyr::filter(
      description %in% selec
    ) %>%
  dplyr::mutate(
      description = base::factor(
        x = description,
        levels = selec,
        ordered = TRUE
      )
    ) %>% 
  dplyr::arrange(description) %>%
  dplyr::pull(description)
```

## [2.8.] Heatmap for visualization of enrichment and proportion of heritability
```{r}
df_sub <- ldsc_data %>%
  dplyr::filter(
      Category %in% cat &
      description %in% trai
    ) %>%
  dplyr::mutate(
    Category = base::factor(
        x = Category,
        levels = base::rev(cat),
        ordered = TRUE
      ),
      description = base::factor(
        x = description,
        levels = trai,
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(dplyr::desc(Prop._h2))
```

#### Evaluation & normalization
```{r}
dplyr::filter(ldsc_data, sample == "F5")
dplyr::filter(df_sub, sample == "F5")

base::range(ldsc_score_$Prop._SNPs)
base::range(df_sub$Prop._SNPs)

base::range(ldsc_score_$Prop._h2)
base::range(df_sub$Prop._h2)

# ma <- base::max(df_sub$Prop._SNPs)
# df_sub <- df_sub %>% dplyr::mutate(NormProp = Prop._SNPs/ma)
# df_sub %>% dplyr::summarise(ranNP = base::range(NormProp))
```

## [2.9.] Plot data
```{r}
m <- base::mean(df_sub$Enrichment_scaled)
ggplot2::ggplot(
    data = df_sub,
    mapping = aes(
        x = description,
        y = Category
      )
    ) +
  geom_tile(
      mapping = aes(
          fill = Enrichment_scaled,
          height = Proportion_scaled, # NormProp, # Prop._h2, # Prop._h2 + 0.25, # Prop._SNPs
          width  = Proportion_scaled # NormProp # Prop._h2, # Prop._h2 + 0.25, # Prop._SNPs
        ),
      color = NA
    ) +
  geom_tile(
      fill = NA,
      color = "black",
      width = 1,
      height = 1
    ) +
  scale_fill_gradient2(
      low = "#0000FF", # "#256866", 
      mid = "#FFFFFF", # "#FFFFCC", 
      high = "#FF0000",  # "orange",
      midpoint = m, 
      na.value = "grey50", 
      name = "Scaled Enrichment"
    ) +
  # labs(
  #     x = "Trait",
  #     y = "Category"
  #   ) +
  coord_fixed(ratio = 1) +
  theme_bw() +
  theme(
      axis.text.x = element_text(
          angle = 90,
          hjust = 1,
          vjust = 0.5,
          size = 8
        ),
      axis.text.y = element_text(
          size = 8
        ),
      # axis.text = element_blank(),
      legend.position = "right",
      axis.title = element_blank(),
      panel.grid = element_blank()
    )
```

## [2.10.] Print plot to file
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_heatmap",
        "_Immune",
        # "_NormalisedProportion",
        "_PropH2ScaledGlobal",
        # "_unnormalized",
        "_ordered",
        "_v4",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 10,
    height = 10
  )
```

## [2.11.] Remove doublets
#### Doublets are included, e.g., dplyr::filter(sample == "B.lymphoblastoid" & description == "Body mass index (BMI)")
#### Solved by taking the average of the enrichment scores for duplicated sample-description combinations
```{r}
pr <- ldsc_score %>% # 3,552,840 
  dplyr::filter(score == 0.5) %>% # 592,140
  dplyr::select(
      sample, description, Enrichment # Prop._h2
    ) %>%
  dplyr::group_by(sample,description) %>%
  tidyr::pivot_wider(
      names_from = description,
      values_from = Enrichment, # Prop._h2,
      values_fn = base::mean
    ) %>%
  dplyr::ungroup() %>%
  tibble::column_to_rownames(var = "sample") %>% 
  base::as.matrix() %>%
  stats::prcomp(.)
```

## Visualize PCA results
```{r}
pr$x %>%
  tibble::as_tibble(rownames = "var") %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = PC1,
        y = PC2,
        label = base::ifelse(PC2  > 50, var, "")
      )
    ) +
  geom_point() +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [3.] Plotting additional trait-sample combinations -------------------------

## [3.0.] Define trait combinations
#### Kidney/Liver function
```{r}
kidney_liver_traits <- base::c(
    "Creatinine (enzymatic) in urine",
    "Albumin (g/L)",
    "Alkaline phosphatase (U/L)", 
    "Alanine aminotransferase (U/L)",
    "Aspartate aminotransferase (U/L)",
    "Direct bilirubin (umol/L)",
    "Urea (mmol/L)",
    "Calcium (mmol/L)",
    "Cholesterol (mmol/L)", 
    "Creatinine (umol/L)",
    "C-reactive protein (mg/L)",
    "Cystatin C (mg/L)", 
    "Gamma glutamyltransferase (U/L)",
    "Glucose (mmol/L)",
    "Glycated haemoglobin (mmol/mol)", 
    "HDL cholesterol (mmol/L)",
    "IGF-1 (nmol/L)",
    "LDL direct (mmol/L)",
    "Phosphate (mmol/L)", 
    "SHBG (nmol/L)",
    "Total bilirubin (umol/L)",
    "Total protein (g/L)",
    "Triglycerides (mmol/L)", 
    "Urate (umol/L)",
    "Vitamin D (nmol/L)",
    "Apoliprotein A (g/L)"
  )
```

#### Body composition
```{r}
body_composition_traits <- base::c(
    "Body mass index (BMI)",
    "Weight",
    "Waist circumference",
    "Whole body fat mass",
    "Whole body fat-free mass",
    "Whole body water mass",
    "Basal metabolic rate",
    "Impedance of whole body",
    #"Impedance of leg (right)", 
    #"Impedance of leg (left)",
    #"Impedance of arm (right)", 
    #"Impedance of arm (left)", 
    #"Leg fat percentage (right)", 
    #"Leg fat mass (right)",
    #"Leg fat-free mass (right)", 
    #"Leg predicted mass (right)", 
    #"Leg fat percentage (left)", 
    #"Leg fat mass (left)",
    #"Leg fat-free mass (left)", 
    #"Leg predicted mass (left)", 
    #"Arm fat percentage (right)", 
    #"Arm fat mass (right)",
    #"Arm fat-free mass (right)", 
    #"Arm predicted mass (right)", 
    #"Arm fat percentage (left)", 
    #"Arm fat mass (left)",
    #"Arm fat-free mass (left)", 
    #"Arm predicted mass (left)", 
    "Trunk fat percentage",
    "Trunk fat mass",
    "Trunk fat-free mass",
    "Trunk predicted mass",
    "Sitting height",
    "Standing height"
  )
```

#### Behaviour-related traits
```{r}
behavioral_traits <- base::c(
    "Sleep duration", 
    "Getting up in morning", 
    "Morning/evening person (chronotype)", 
    "Nap during day",
    "Sleeplessness / insomnia", 
    "Daytime dozing / sleeping (narcolepsy)",
    #"Ever had prolonged feelings of sadness or depression",
    "Ever thought that life not worth living", 
    #"Felt very upset when reminded of stressful experience in past month", 
    "Frequency of depressed mood in last 2 weeks",
    "Recent feelings of depression", 
    "Recent feelings of tiredness or low energy", 
    "Frequency of unenthusiasm / disinterest in last 2 weeks", 
    "Frequency of tenseness / restlessness in last 2 weeks",
    "Frequency of tiredness / lethargy in last 2 weeks", 
    #"Seen doctor (GP) for nerves, anxiety, tension or depression",
    "Seen a psychiatrist for nerves, anxiety, tension or depression", 
    "Able to confide", 
    "Ever depressed for a whole week", 
    "Ever unenthusiastic/disinterested for a whole week"
  )
```

#### Heart measures
```{r}
heart_traits <- base::c(
    "Pulse rate",
    "Diastolic blood pressure, automated reading",
    "Systolic blood pressure, automated reading",
    "Diastolic blood pressure, manual reading",
    "Systolic blood pressure, manual reading",
    "ECG, heart rate"
  )
```

#### Eye measures
```{r}
eye_traits <- base::c(
    "Spherical power (left)",
    #"Spherical power (right)", 
    "Cylindrical power (left)", 
    "3mm weak meridian (left)", 
    "6mm weak meridian (left)", 
    #"6mm weak meridian (right)", 
    #"3mm weak meridian (right)", 
    #"3mm strong meridian angle (left)",
    #"3mm strong meridian angle (right)",
    "3mm cylindrical power (right)", 
    "6mm cylindrical power (right)", 
    #"6mm cylindrical power (left)", 
    #"3mm cylindrical power (left)", 
    "3mm strong meridian (right)", 
    #"6mm strong meridian (right)", 
    "6mm strong meridian (left)", 
    #"3mm strong meridian (left)", 
    #"6mm asymmetry index (left)",
    "6mm asymmetry index (right)", 
    #"3mm regularity index (right)", 
    "Intra-ocular pressure, corneal-compensated (right)", 
    "Intra-ocular pressure, Goldmann-correlated (right)", 
    "Corneal hysteresis (right)", 
    "Corneal resistance factor (right)" 
    #"Intra-ocular pressure, corneal-compensated (left)", 
    #"Intra-ocular pressure, Goldmann-correlated (left)", 
    #"Corneal hysteresis (left)", 
    #"Corneal resistance factor (left)"
  )
```

#### Blood traits
```{r}
blood_traits <- base::c(
    "Lymphocyte count",
    "Platelet count",
    "Eosinophill count",
    "Mean corpuscular haemoglobin concentration",
    "Mean corpuscular volume",
    "Neutrophill percentage",
    "Monocyte count",
    "Reticulocyte count",
    "Red blood cell (erythrocyte) distribution width",
    "Platelet distribution width"
  )
```

## [3.1.] Define FANTOM5 samples to be included
```{r}
relevant_samples <- base::c(
    "astrocyte",                                # Neural / CNS
    "brain",
    "neuron",
    #"spinal.cord",
    "eye",                                      # Eye
    "lens.epithelial.cell",
    "retinal.pigment.epithelial.cell",
    #"corneal.epithelial.cell",
    "heart",                                    # Heart / vasculature
    "cardiac.myocyte",
    "cardiac.fibroblast",
    "blood.vessel",
    "blood.vessel.endothelial.cell",
    "vascular.associated.smooth.muscle.cell",
    #"endothelial.cell.of.hepatic.sinusoid",
    #"endothelial.cell.of.lymphatic.vessel",
    "skeletal.muscle.cell",                    # Muscle
    #"skeletal.muscle.tissue",
    "liver",                                   # Liver / fat / digestive
    "hepatocyte",
    #"hepatic.stellate.cell",
    "fat.cell",
    "preadipocyte",
    "acinar.cell",
    "embryonic.pancreas",
    "embryonic.kidney",
    "kidney",
    "kidney.epithelial.cell",
    "small.intestine",
    "large.intestine",
    #"epithelial.cell.of.esophagus",
    "skin.fibroblast",                        # Connective / fibroblasts
    #"fibroblast.of.lymphatic.vessel",
    "mesenchymal.cell",
    #"stromal.cell",
    #"tendon.cell",
    "cord.blood.derived",                     # Blood / immune cells
    "T.cell",
    "neutrophil",
    "monocyte",
    "natural.killer.cell",
    "lymphocyte.of.B.lineage",
    "reticulocyte",
    "testis",                                # Reproductive / endocrine
    "thyroid.gland",
    #"uterus",
    "ciliated.epithelial.cell"               # Misc / less obvious
    #"keratocyte"
    #"hair.follicle.cell",
    #"tongue",
    #"trabecular.meshwork.cell",
    #"urothelial.cell",
    #"smooth.muscle.cell.of.prostate",
    #"smooth.muscle.cell.of.trachea",
    #"fibroblast.of.choroid.plexus",
    #"fibroblast.of.gingiva",
    #"fibroblast.of.periodontium",
    #"fibroblast.of.tunica.adventitia.of.artery",
    #"gingival.epithelial.cell",
    #"epithelial.cell.of.Malassez",
    #"epithelial.cell.of.prostate",
    #"sensory.epithelial.cell",
    #"amniotic.epithelial.cell",
    #"chondrocyte",
    #"keratinocyte"
  )
```

## [3.2.] Combine traits in order
```{r}
all_traits <- base::c(
    eye_traits,
    kidney_liver_traits,
    body_composition_traits,
    behavioral_traits,
    heart_traits,
    blood_traits
  )
```

## [3.3.] Heatmap for visualization of enrichment and proportion of heritability
```{r}
df_sub <- ldsc_data %>%
  dplyr::filter(
      sample %in% relevant_samples &
      description %in% all_traits
    ) %>%
  dplyr::mutate(
    sample = base::factor(
        x = sample,
        levels = relevant_samples,
        ordered = TRUE
      ),
    description = base::factor(
        x = description,
        levels = all_traits,
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(dplyr::desc(Prop._h2))
```

## [3.4.] Plot data
```{r}
m <- base::mean(df_sub$Enrichment_scaled)
ggplot2::ggplot(
    data = df_sub,
    mapping = aes(
        x = description,
        y = sample # Category
      )
    ) +
  geom_tile(
      mapping = aes(
          fill = Enrichment_scaled,
          height = Proportion_scaled, # NormProp, # Prop._h2, # Prop._h2 + 0.25, # Prop._SNPs
          width  = Proportion_scaled # NormProp # Prop._h2, # Prop._h2 + 0.25, # Prop._SNPs
        ),
      color = NA
    ) +
  geom_tile(
      fill = NA,
      color = "black",
      width = 1,
      height = 1
    ) +
  scale_fill_gradient2(
      low = "#0000FF", # "#256866", 
      mid = "#FFFFFF", # "#FFFFCC", 
      high = "#FF0000",  # "orange",
      midpoint = m, 
      na.value = "grey50", 
      name = "Scaled Enrichment"
    ) +
  # labs(
  #     x = "Trait",
  #     y = "Category"
  #   ) +
  coord_fixed(ratio = 1) +
  theme_bw() +
  theme(
      axis.text.x = element_text(
          angle = 90,
          hjust = 1,
          vjust = 0.5,
          size = 8
        ),
      axis.text.y = element_text(
          size = 8
        ),
      # axis.text = element_blank(),
      legend.position = "right",
      axis.title = element_blank(),
      panel.grid = element_blank()
    )
```

## [3.5.] Print plot to file
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_heatmap",
        "_AllTraits",
        "_PropH2Scaled",
        # "_NormalisedProportion",
        # "_PropH2ScaledGlobal",
        # "_unnormalized",
        "_ordered",
        "_v1",
        ".pdf"
      ),
    device = "pdf",
    plot = ggplot2::last_plot(),
    width = 15,
    height = 8
  )
```

## [4.] Additional analyses -------------------------

# Grouped dotplot for enrichments specifically in blood!
```{r}
blood_traits <- base::c(
  #"White blood cell (leukocyte) count",
  #"Red blood cell (erythrocyte) count",
  #"Haemoglobin concentration",
  "Mean corpuscular volume",
  #"Mean corpuscular haemoglobin",
  #"Mean corpuscular haemoglobin concentration",
  "Red blood cell (erythrocyte) distribution width",
  #"Platelet count",
  #"Mean platelet (thrombocyte) volume",
  #"Platelet distribution width",
  "Lymphocyte count",
  "Monocyte count",
  #"Neutrophill count",
  "Eosinophill count",
  "Basophill count",
  #"Mean reticulocyte volume",
  #"Mean sphered cell volume",
  #"Immature reticulocyte fraction",
  #"High light scatter reticulocyte count",
  "Reticulocyte count"
)

immune_relevant_celltypes <- base::c(
  # Immune cells
  "B.lymphoblastoid",
  "cord.blood.derived",
  "dendritic.cell",
  "lymphocyte.of.B.lineage",
  "macrophage",
  "mast.cell",
  "monocyte",
  "natural.killer.cell",
  "neutrophil",
  "reticulocyte",
  "T.cell",
  # Immune-related tissues
  "spleen",
  "thymus",
  "blood.vessel",
  "blood.vessel.endothelial.cell",
  # Include FANTOM5 for comparison
  "F5"
)

df <- filter(ldsc_data, Category %in% immune_relevant_celltypes & description %in% blood_traits)



df_top3plusF5 <- df %>%
  group_by(description) %>%
  # get top 3 cell types excluding F5_distal
  filter(Category != "F5_distal") %>%
  slice_max(order_by = Enrichment, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  bind_rows(
    df %>%
      filter(Category == "F5_distal")
  ) %>%
  distinct(description, Category, .keep_all = TRUE)


ggplot(df_top3plusF5, aes(x = Enrichment, y = description, color = Category)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = Enrichment - Enrichment_std_error,
                     xmax = Enrichment + Enrichment_std_error),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  scale_color_manual(values = brewer.pal(12, "Set3") %>% c(brewer.pal(8, "Dark2")[1:4])) +  # Combine palettes for 16 colors
  theme_minimal() +
  labs(x = "Enrichment", y = "Trait") +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 9),
    panel.spacing = unit(1, "lines")
  ) +
  xlim(2.5,65)

ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/blood.enrichments.pdf",height=10, width=10)
```




```{r}
# ldsc_data <- ldsc_data %>%
#   group_by(Trait) %>%
#   mutate(
#     Enrichment_scaled = if (max(Enrichment, na.rm = TRUE) != min(Enrichment, na.rm = TRUE)) {
#       (Enrichment - min(Enrichment, na.rm = TRUE)) /
#         (max(Enrichment, na.rm = TRUE) - min(Enrichment, na.rm = TRUE))
#     } else {
#       0
#     }
#   ) %>%
#   ungroup()

# Selected traits/cellFacets for figures
cat <- c("B.lymphoblastoid_distal",
         "blood.vessel.endothelial.cell_distal",
         "brain_distal",
         "F5_distal",
         "kidney_distal",
         "liver_distal",
         "monocyte_distal",
         "neutrophil_distal",
         "reticulocyte_distal",
         "skeletal.muscle.tissue_distal",
         "spleen_distal",
         "T.cell_distal",
         "thymus_distal"
  
)


trai <- c("Alanine aminotransferase (U/L)",
          "Albumin (g/L)",
          "Apoliprotein A (g/L)",
          "Aspartate aminotransferase (U/L)",
          "Body mass index (BMI)",
          "Creatinine (umol/L)",
          "Diastolic blood pressure, automated reading",
          "Eosinophill count",
          "Ever depressed for a whole week",
          "Getting up in morning",
          "High light scatter reticulocyte count",
          "Mean corpuscular volume",
          "Monocyte count",
          "Reticulocyte count",
          "SHBG (nmol/L)",
          "Urate (umol/L)",
          "White blood cell (leukocyte) count"
  
)

#  Heatmap

df_sub <- filter(ldsc_data, Category %in% cat & description %in% trai)
m <- mean(df_sub$Enrichment_scaled)
 p <- ggplot(df_sub, aes(x = description, y = Category)) +
    geom_tile(aes(fill = Enrichment_scaled, 
                  height = Prop._h2 + 0.4, 
                  width  = Prop._h2 + 0.4),
              color = NA) +
    geom_tile(fill = NA, color = "black", width = 1, height = 1) +
    scale_fill_gradient2(
      low      = "#256866", 
      mid      = "#FFFFCC", 
      high     = "orange",
      midpoint = m, 
      na.value = "grey50", 
      name     = "Enrichment"
    ) +
    labs(
      x     = "Trait",
      y     = "Category"
    ) +
    theme_minimal() +
    theme(
      axis.text.x     = element_text(angle = 45, hjust = 1),
      axis.text.y     = element_text(size = 3),
      legend.position = "right",
      panel.grid      = element_blank()
    ) +
    coord_fixed(ratio = 1)

ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/heatmap.Scaledenrichments.pdf",p,height=10, width=10)

```

