---
title: "FANTOM5 plotting"
date: "2025-06-03"
output: html_document
---

## [-1.0.] Install necessary packages
```{r}
utils::install.packages("factoextra")
utils::install.packages("tsne")
```

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r}
# base::library(CAGEfightR)
# base::library(GenomicRanges)
# base::library(readr)
# base::library(rtracklayer)
# base::library(parallel)
# base::library(ggplot2)
# base::library(GenomicFeatures)
# base::library(SummarizedExperiment)
# base::library(tibble)
# base::library(ggrepel)
# base::library(dplyr)
# base::library(scales)
# base::library(gridExtra)
# base::library(BuenColors)
# base::library(reshape2)
# base::library(vegan)
# base::library(RColorBrewer)
# base::library(pheatmap)
# base::library(dendsort)
# base::library(tidyr)
# base::library(patchwork)
# base::library(ggdendro)
base::library(tidyverse)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.1.] Load pooled cellFacets data (PL score > 0.75)
```{r}
cellFacet.rse_1 <- base::readRDS(
  file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "cellFacet.rds"
    )
  )
```

## [0.2.] Read in FANTOM5 predictions
```{r}
F5 <- rtracklayer::import.bed(
  con = base::paste0(
      here::here("0.External_resources/"),
      "F5.hg38.enhancers.bed.gz"
    )
  )
```

## [0.3.] Define categories for all cellFacets
```{r}
cell_type_categories <- base::c(
  "acute.lymphoblastic.leukemia.T.ALL" = "Cancer / Tumor",
  "heart" = "Muscle",
  "acute.myeloid.leukemia.FAB.M6" = "Cancer / Tumor",
  "embryonic.kidney" = "Stem / Progenitor",
  "hair.follicle.cell" = "Epithelial",
  "kidney" = "Epithelial",
  "melanoma" = "Cancer / Tumor",
  "giant.cell.carcinoma" = "Cancer / Tumor",
  "fibroblast.of.gingiva" = "Fibroblast/Connective",
  "T.cell" = "Immune",
  "fibroblast.of.choroid.plexus" = "Fibroblast/Connective",
  "cord.blood.derived" = "Stem / Progenitor",
  "preadipocyte" = "Fibroblast/Connective",
  "small.cell.lung.carcinoma" = "Cancer / Tumor",
  "gingival.epithelial.cell" = "Epithelial",
  "fibroblast.of.tunica.adventitia.of.artery" = "Fibroblast/Connective",
  "thyroid.carcinoma" = "Cancer / Tumor",
  "placental.epithelial.cell" = "Epithelial",
  "tendon.cell" = "Fibroblast/Connective",
  "Wilms.tumor" = "Cancer / Tumor",
  "carcinoid" = "Cancer / Tumor",
  "fat.cell" = "Fibroblast/Connective",
  "thymus" = "Immune",
  "neuroblastoma" = "Cancer / Tumor",
  "mesothelioma" = "Cancer / Tumor",
  "meninx" = "CNS/Neural",
  "eye" = "CNS/Neural",
  "clear.cell.carcinoma" = "Cancer / Tumor",
  "lung.adenocarcinoma" = "Cancer / Tumor",
  "schwannoma" = "CNS/Neural",
  "choriocarcinoma" = "Cancer / Tumor",
  "acute.myeloid.leukemia.FAB.M7" = "Cancer / Tumor",
  "endothelial.cell.of.hepatic.sinusoid" = "Endothelial",
  "small.intestine" = "Epithelial",
  "acute.myeloid.leukemia.FAB.M4eo" = "Cancer / Tumor",
  "smooth.muscle.cell.of.prostate" = "Muscle",
  "monocyte" = "Immune",
  "testicular.germ.cell.embryonal.carcinoma" = "Cancer / Tumor",
  "chondrocyte" = "Fibroblast/Connective",
  "teratocarcinoma" = "Cancer / Tumor",
  "urothelial.cell" = "Epithelial",
  "merkel.cell.carcinoma" = "Cancer / Tumor",
  "mammary.epithelial.cell" = "Epithelial",
  "retinal.pigment.epithelial.cell" = "Epithelial",
  "internal.male.genitalia" = "Other / Mixed",
  "breast.carcinoma" = "Cancer / Tumor",
  "hepatocellular.carcinoma" = "Cancer / Tumor",
  "blood.vessel" = "Endothelial",
  "colon.carcinoma" = "Cancer / Tumor",
  "mesothelial.cell" = "Epithelial",
  "epithelial.cell.of.esophagus" = "Epithelial",
  "oral.squamous.cell.carcinoma" = "Cancer / Tumor",
  "hepatocyte" = "Epithelial",
  "spinal.cord" = "CNS/Neural",
  "melanocyte" = "CNS/Neural",
  "neuron" = "CNS/Neural",
  "neuroectodermal.tumor" = "Cancer / Tumor",
  "spleen" = "Immune",
  "rhabdomyosarcoma" = "Cancer / Tumor",
  "medulloblastoma" = "Cancer / Tumor",
  "Hep.2" = "Cancer / Tumor",
  "lung" = "Epithelial",
  "epidermoid.carcinoma" = "Cancer / Tumor",
  "large.cell.lung.carcinoma" = "Cancer / Tumor",
  "testis" = "Other / Mixed",
  "enteric.smooth.muscle.cell" = "Muscle",
  "transitional.cell.carcinoma" = "Cancer / Tumor",
  "tongue" = "Epithelial",
  "cardiac.myocyte" = "Muscle",
  "myxofibrosarcoma" = "Cancer / Tumor",
  "endothelial.cell.of.lymphatic.vessel" = "Endothelial",
  "neuronal.stem.cell" = "Stem / Progenitor",
  "acute.myeloid.leukemia.FAB.M5" = "Cancer / Tumor",
  "leiomyoma" = "Cancer / Tumor",
  "ductal.cell.carcinoma" = "Cancer / Tumor",
  "keratinocyte" = "Epithelial",
  "myoblast" = "Muscle",
  "epithelioid.sarcoma" = "Cancer / Tumor",
  "osteosarcoma" = "Cancer / Tumor",
  "squamous.cell.lung.carcinoma" = "Cancer / Tumor",
  "brain" = "CNS/Neural",
  "bronchioalveolar.carcinoma" = "Cancer / Tumor",
  "acute.myeloid.leukemia.FAB.M4" = "Cancer / Tumor",
  "epithelial.cell.of.prostate" = "Epithelial",
  "fibroblast.of.lymphatic.vessel" = "Fibroblast/Connective",
  "amniotic.epithelial.cell" = "Epithelial",
  "liposarcoma" = "Cancer / Tumor",
  "epithelial.cell.of.Malassez" = "Epithelial",
  "hepatic.stellate.cell" = "Fibroblast/Connective",
  "corneal.epithelial.cell" = "Epithelial",
  "gastric.cancer" = "Cancer / Tumor",
  "squamous.cell.carcinoma" = "Cancer / Tumor",
  "liver" = "Epithelial",
  "stromal.cell" = "Fibroblast/Connective",
  "astrocyte" = "CNS/Neural",
  "cardiac.fibroblast" = "Fibroblast/Connective",
  "acute.lymphoblastic.leukemia.B.ALL" = "Cancer / Tumor",
  "thyroid.gland" = "Epithelial",
  "gastric.adenocarcinoma" = "Cancer / Tumor",
  "renal.cell.carcinoma" = "Cancer / Tumor",
  "lens.epithelial.cell" = "Epithelial",
  "dendritic.cell" = "Immune",
  "smooth.muscle.cell.of.trachea" = "Muscle",
  "macrophage" = "Immune",
  "pericyte.cell" = "Endothelial",
  "mast.cell" = "Immune",
  "gall.bladder.carcinoma" = "Cancer / Tumor",
  "skin.fibroblast" = "Fibroblast/Connective",
  "skeletal.muscle.cell" = "Muscle",
  "respiratory.epithelial.cell" = "Epithelial",
  "uterus" = "Epithelial",
  "trabecular.meshwork.cell" = "Fibroblast/Connective",
  "mesenchymal.cell" = "Fibroblast/Connective",
  "ciliated.epithelial.cell" = "Epithelial",
  "epitheloid.carcinoma" = "Cancer / Tumor",
  "sensory.epithelial.cell" = "Epithelial",
  "fibroblast.of.periodontium" = "Fibroblast/Connective",
  "bile.duct.carcinoma" = "Cancer / Tumor",
  "prostate.cancer" = "Cancer / Tumor",
  "neutrophil" = "Immune",
  "kidney.epithelial.cell" = "Epithelial",
  "glioblastoma" = "Cancer / Tumor",
  "lymphocyte.of.B.lineage" = "Immune",
  "chronic.myelogenous.leukemia" = "Cancer / Tumor",
  "acute.myeloid.leukemia.FAB.M2" = "Cancer / Tumor",
  "throat" = "Epithelial",
  "large.intestine" = "Epithelial",
  "osteoblast" = "Fibroblast/Connective",
  "serous.adenocarcinoma" = "Cancer / Tumor",
  "cervical.cancer" = "Cancer / Tumor",
  "Burkitts.lymphoma" = "Cancer / Tumor",
  "reticulocyte" = "Immune",
  "skeletal.muscle.tissue" = "Muscle",
  "natural.killer.cell" = "Immune",
  "signet.ring.carcinoma" = "Cancer / Tumor",
  "vascular.associated.smooth.muscle.cell" = "Muscle",
  "embryonic.pancreas" = "Stem / Progenitor",
  "acinar.cell" = "Epithelial",
  "blood.vessel.endothelial.cell" = "Endothelial",
  "keratocyte" = "Fibroblast/Connective",
  "B.lymphoblastoid" = "Immune"
)
```

## [1.] PCA & k means clustering -------------------------

## [1.0.] Perform PCA
```{r}
SummarizedExperiment::colData(cellFacet.rse)
mat <- SummarizedExperiment::assay(cellFacet.rse,"PL.score") %>%
  base::replace(base::is.na(.), 0) %>%
  base::as.matrix() %>%
  base::t()
mat[1:5,1:5]
base::dim(mat)
pr <- stats::prcomp(mat)
utils::str(pr)
```

## [1.1.] Visualize PCA results
```{r}
factoextra::fviz_pca_ind(
    pr,
    col.ind = "cos2",
    gradient.cols = base::c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE
  )
```

## [1.2.] Visualize PCA results
```{r}
pr$x %>%
  tibble::as_tibble(rownames = "var") %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = PC1,
        y = PC2,
        label = base::ifelse(PC2  > 50, var, "")
      )
    ) +
  geom_point() +
  ggrepel::geom_text_repel(
      box.padding = 0.5,
      max.overlaps = Inf,
      size = 2
    ) +
  scale_x_continuous(
      limits = base::c(-700,700)
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [1.3.] PCA-dependent tSNE analysis
```{r}
tmat <- tsne::tsne(X = pr$x, initial_dims = 2)
base::colnames(tmat) <- base::c("tsne_1","tsne_2")
base::rownames(tmat) <- base::rownames(pr$x)
```

```{r}
factoextra::fviz_nbclust(mat, kmeans, method = "wss")
```

## [1.4.] k means clustering
```{r}
kmat <- stats::kmeans(
    x = mat,
    centers = 7,
    nstart = 25
  )
utils::str(kmat)
```

## [1.5.] Save 'kmat' object
```{r}
base::saveRDS(
  object = kmat,
  file = base::paste0(
      here::here("11.LDSC_FANTOM5_copy/"),
      base::format(base::Sys.Date(), "%Y%m%d"),
      "_F5_clusters.rds"
    )
  )
```

## [1.6.] Data wrangling & combination
```{r}
f5cl <- tibble::tibble(
    cat = base::names(kmat$cluster),
    clus = kmat$cluster
  )
f5mat <- tmat %>%
  tibble::as.tibble(
      rownames = "cat"
    ) %>%
  dplyr::left_join(
      x = ., y = f5cl, by = "cat"
    ) %>%
  dplyr::mutate(
      clus = base::factor(
        x = clus,
        levels = 1:7,
        ordered = TRUE
      )
    ) %>%
  dplyr::arrange(clus)
```

## [1.6.] Data visualization
```{r}
ggplot2::ggplot(
      data = f5mat,
      mapping = aes(
        x = tsne_1,
        y = tsne_2,
        colour = clus
      )
    ) +
  geom_point() +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [1.] Data wrangling -------------------------

## [1.0.] Add annotations
```{r}
SummarizedExperiment::rowData(cellFacet.rse)$location <- NA
SummarizedExperiment::rowData(cellFacet.rse)$location[SummarizedExperiment::rowData(cellFacet.rse)$txType %in% base::c("promoter","proximal","fiveUTR")] <- "proximal"
SummarizedExperiment::rowData(cellFacet.rse)$location[SummarizedExperiment::rowData(cellFacet.rse)$txType %in% base::c("intron","intergenic")] <- "distal"
keep <- !base::is.na(SummarizedExperiment::rowData(cellFacet.rse)$location)
cellFacet.rse <- cellFacet.rse[keep,]

F5 <- CAGEfightR::assignTxType(
    object = F5,
    txModels = txdb
  )
F5$location[F5$txType %in% base::c("promoter","proximal","fiveUTR")] <- "proximal"
F5$location[F5$txType %in% base::c("intron","intergenic")] <- "distal"
```

## [1.1.] Calculate support across cellFacets at PL score 0.5
```{r}
SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score")[
    base::is.na(SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score"))
  ] <- 0
cellFacet.rse <- CAGEfightR::calcSupport(
    object = cellFacet.rse,
    inputAssay = "qn.PL.score",
    unexpressed = 0.5,
    outputColumn = "support0.5"
  )
cellFacet.rse <- CAGEfightR::calcSupport(
    object = cellFacet.rse,
    inputAssay = "qn.PL.score",
    unexpressed = 0.75,
    outputColumn = "support0.75"
  )
```

## [1.2.] Calculate number of proximal and distal regions at PL score 0.5 and 0.75
```{r}
SummarizedExperiment::colData(cellFacet.rse)$distal0.5 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse,location == "distal"),
      i = "qn.PL.score"
    ) > 0.5
  )
SummarizedExperiment::colData(cellFacet.rse)$distal0.75 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "distal"),
      i = "qn.PL.score"
    ) > 0.75
  )
SummarizedExperiment::colData(cellFacet.rse)$proximal0.5 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "proximal"),
      i = "qn.PL.score"
    ) > 0.5
  )
SummarizedExperiment::colData(cellFacet.rse)$proximal0.75 <- base::colSums(
  SummarizedExperiment::assay(
      x = base::subset(cellFacet.rse, location == "proximal"),
      i = "qn.PL.score"
    ) > 0.75
  )
```

## [1.2.] Calculate binary jaccard distance between cellFacets at PL score 0.5
```{r}
binary_mat <- SummarizedExperiment::assay(cellFacet.rse, "qn.PL.score") > 0.5

jaccard_dist <- vegan::vegdist(
    x = base::t(binary_mat),
    method = "jaccard"
  )
jaccard_mat <- base::as.matrix(jaccard_dist)

jaccard_sim <- 1 - jaccard_mat
```

## [1.3.] Selecting cellFacets for main figure
```{r}
main <- base::names(cell_type_categories)[cell_type_categories != "Cancer / Tumor"]

jaccard_sim_main <- jaccard_sim[
    base::row.names(jaccard_sim) %in% main,
    base::colnames(jaccard_sim) %in% main
  ]
```

## [1.4.] Cluster and sort
```{r}
hc <- stats::hclust(
    stats::as.dist(1 - jaccard_sim),
    method = "average"
  )
hc_sorted <- dendsort::dendsort(hc)

hc_main <- stats::hclust(
    stats::as.dist(1 - jaccard_sim_main),
    method = "average"
  )
hc_main_sorted <- dendsort::dendsort(hc_main)
```

## [2.] Descriptive plots -------------------------

## [2.0.] Plot support across cellFacets at PL score 0.5
#### ggsave("11.LDSC_FANTOM5/plots/support0.5.pdf", height=10,width=7)
```{r}
sps <- cellFacet.rse %>%
  SummarizedExperiment::rowData() %>%
  base::as.data.frame() %>%
  tidyr::pivot_longer(
      cols = base::c("support0.5", "support0.75"),
      names_to = "support",
      values_to = "qn.PL.score"
    ) %>% 
  dplyr::mutate(
    location = base::factor(
        x = location,
        levels = base::c("proximal", "distal"),
        ordered = TRUE
      ),
    support = base::factor(
        x = support,
        levels = base::c("support0.5", "support0.75"),
        ordered = TRUE
      )
    )
```

```{r}
ggplot2::ggplot(
    data = sps,
    mapping = aes(
        x = qn.PL.score,
        fill = location
      )
    )  +
  geom_bar() +
  scale_fill_manual(
      values = base::c("proximal" = "#1b9e77", "distal" = "#d95f02")
    ) +
  facet_wrap(
      facets = support ~ location,
      # ncol = 1,
      scales = "free"
    )

ggplot2::ggplot(
    data = sps,
    mapping = aes(
        x = qn.PL.score,
        # fill = location,
        colour = location
      )
    )  +
  # geom_bar() +
  geom_density() +
  scale_fill_manual(
      values = base::c("proximal" = "#1b9e77", "distal" = "#d95f02")
    ) +
  facet_wrap(
      # facets = support ~ location,
      facets = . ~ support,
      # ncol = 1,
      scales = "free"
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.1.] Number of RE at PL score 0.5 clustered based on jaccard distance
#### see [1.3.] above for 'hc_sorted' object
#### Curated clusters: 1:29; 30:73; 74:85; 86:91; 92:107; 108:130; 131:141
```{r}
ordered_samples <- hc_sorted$labels[hc_sorted$order]

# Filter colData to include only those in clustering
df <- SummarizedExperiment::colData(cellFacet.rse) %>%
  tibble::as_tibble() %>%
  dplyr::filter(
      sample %in% ordered_samples
    )

# Get dendrogram data
dend <- stats::as.dendrogram(hc_sorted)
dend_data <- ggdendro::dendro_data(dend)
label_map <- stats::setNames(dend_data$labels$x, dend_data$labels$label)
seg_data <- dend_data$segments

# Reshape data to long format for plotting and assign dendrogram x positions
df_long <- df %>%
  tidyr::pivot_longer(
      cols = base::c("distal0.5", "proximal0.5"),
      names_to = "region_type",
      values_to = "count"
    ) %>%
  dplyr::mutate(
      region_type = base::factor(
          x = region_type,
          levels = base::c("proximal0.5", "distal0.5")
        ),
      x = label_map[as.character(sample)]
    )
```

```{r}
curclus <- tibble::tibble(
  cat = hc_sorted$labels[hc_sorted$order],
  clus = base::c(
      base::rep("1",29-1+1),
      base::rep("2",73-30+1), 
      base::rep("3",85-74+1), 
      base::rep("4",91-86+1), 
      base::rep("5",107-92+1), 
      base::rep("6",130-108+1), 
      base::rep("7",141-131+1)
    )
  )

curmat <- tmat %>%
  tibble::as.tibble(
      rownames = "cat"
    ) %>%
  dplyr::left_join(
      x = ., y = curclus, by = "cat"
    )

ggplot2::ggplot(
      data = curmat,
      mapping = aes(
        x = tsne_1,
        y = tsne_2,
        colour = clus,
        label = base::ifelse(clus == 2, cat, "")
      )
    ) +
  geom_point() +
  ggrepel::geom_text_repel(
      box.padding = 0.5,
      max.overlaps = Inf,
      size = 2
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1
    )
```

## [2.1.] Combined plot
#### ggsave("11.LDSC_FANTOM5/plots/nBarplot.pdf", combined_plot, width = 7, height = 15)
```{r}
bar_plot <- ggplot2::ggplot(
    data = df_long,
    mapping = aes(
        x = x,
        y = count,
        fill = region_type
      )
    ) +
  geom_bar(
      stat = "identity", position = "dodge"
    ) +
  geom_hline(
      yintercept = base::length(F5),
      linetype = "dashed",
      color = "darkred",
      linewidth = 0.7
    ) +
  geom_hline(
      yintercept = 100000,
      linetype = "dashed",
      color = "darkgrey",
      linewidth = 0.7
    ) +
  scale_x_continuous(
      breaks = df_long$x,
      labels = df_long$sample
    ) +
  scale_fill_manual(
      values = base::c("proximal0.5" = "#1b9e77", "distal0.5" = "#d95f02")
    ) +
  coord_flip() +
  labs(
      x = NULL,
      y = "Region count",
      fill = "Region Type"
    ) +
  theme_minimal() +
  theme(
      axis.text.y = element_text(size = 6, hjust = 1),
      axis.title.y = element_blank(),
      panel.grid.major.y = element_blank(),  # remove horizontal lines
      panel.grid.minor.y = element_blank()   # just in case
    )

# Dendrogram plot 
dend_plot <- ggplot2::ggplot(
      data = seg_data,
      mapping = aes(
        x = x,
        y = y,
        xend = xend,
        yend = yend
      )
    ) +
  geom_segment() +
  scale_y_reverse() +
  coord_flip() +
  theme_void() +
  theme(
      plot.margin = margin(r = -10)
    )  # Reduce spacing to snugly align with barplot

# Combine and plot
library(patchwork)
combined_plot <- dend_plot + bar_plot + patchwork::plot_layout(widths = c(1, 3))
combined_plot
```

```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("11.LDSC_FANTOM5_copy/plots/"),
        base::format(base::Sys.Date(),"%Y%m%d"),
        "_nBarplot.pdf"
      ),
    device = "pdf",
    plot = combined_plot,
    width = 7,
    height = 15
  )
```


```{r}
# Same for main figure
ordered_samples <- hc_main_sorted$labels[hc_main_sorted$order]

# Filter colData to include only those in clustering
df <- as.data.frame(colData(cellFacet.rse)) %>%
  as_tibble() %>%
  filter(sample %in% ordered_samples)

# Get dendrogram data
dend <- as.dendrogram(hc_main_sorted)
dend_data <- dendro_data(dend)
label_map <- setNames(dend_data$labels$x, dend_data$labels$label)
seg_data <- dend_data$segments

# Reshape data to long format for plotting and assign dendrogram x positions
df_long <- df %>%
  pivot_longer(cols = c(distal0.5, proximal0.5),
               names_to = "region_type",
               values_to = "count") %>%
  mutate(region_type = factor(region_type, levels = c("proximal0.5", "distal0.5")),
         x = label_map[as.character(sample)])

# Bar plot 
bar_plot <- ggplot(df_long, aes(x = x, y = count, fill = region_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("proximal0.5" = "#1b9e77", "distal0.5" = "#d95f02")) +
  scale_x_continuous(breaks = df_long$x,
                     labels = df_long$sample) +
  coord_flip() +
  labs(x = NULL, y = "Region count", fill = "Region Type") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6, hjust = 1),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = length(F5), linetype = "dashed", color = "darkred", linewidth = 0.7) +
  geom_hline(yintercept = 100000, linetype = "dashed", color = "darkgrey", linewidth = 0.7)+
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6, hjust = 1),
    panel.grid.major.y = element_blank(),  # remove horizontal lines
    panel.grid.minor.y = element_blank()   # just in case
  )



# Dendrogram plot 
dend_plot <- ggplot(seg_data) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  scale_y_reverse() +
  coord_flip() +
  theme_void() +
  theme(plot.margin = margin(r = -10))  # Reduce spacing to snugly align with barplot

# Combine and plot
combined_plot <- dend_plot + bar_plot + plot_layout(widths = c(1, 3))
combined_plot

ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/nBarplot.main.pdf", combined_plot, width = 15, height = 15)
```

## [2.] LDSC analysis plots -------------------------

## [2.0] Read in main LDSC results
```{r}
# Load the data
ldsc_score <- readr::read_tsv(
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "combined_results.tsv"
    )
  )
ldsc_score$Trait<- base::sub(pattern = "_score[0-9.]+$", "", ldsc_score$Trait) 

sumstat_annot <- readr::read_tsv(
    file = base::paste0(
      here::here("11.LDSC_FANTOM5/ldsc_files/"),
      "ukb31063_ldsc_sumstat_manifest_aws_sep2022.tsv"
    )
  )
sumstat_annot <- filter(sumstat_annot, is_primary_gwas==TRUE & phenotype %in% unique(ldsc_score$Trait))

# Add description to data
ldsc_score <- dplyr::left_join(
  ldsc_score,
  sumstat_annot %>% dplyr::select(phenotype, description),
  by = c("Trait" = "phenotype")
)

ldsc_score <- ldsc_score %>%
  mutate(loc = case_when(
    grepl("distal", Category) ~ "distal",  
    grepl("proximal", Category) ~ "proximal",  
    TRUE ~ "unknown"  
  ))

ldsc_score$score <- sub(".*_(\\d+\\.\\d+)L2_0$", "\\1", ldsc_score$Category)

ldsc_score$sample <- sub("_.*", "", ldsc_score$Category)


ldsc_score <- ldsc_score[ldsc_score$Category!="baseL2_0",]
ldsc_data <- ldsc_score[ldsc_score$score=="0.5",]
```






```{r}
plot_trait_scores <- function(ldsc_score, description, samples, min_top_scores) {
  # Subset relevant data
  df <- unique(ldsc_score[ldsc_score$description == description &
                            ldsc_score$sample %in% c("F5", samples), ])
  
  # Average F5 per category
  f5_summary <- df %>%
    filter(sample == "F5") %>%
    group_by(sample) %>%
    summarize(
      Prop._h2 = mean(Prop._h2, na.rm = TRUE),
      Enrichment = mean(Enrichment, na.rm = TRUE),
      Enrichment_p = min(Enrichment_p, na.rm = TRUE),
      score = "F5",
      Category = "F5",
      sample = "F5")

  # Other samples as-is
  other_samples <- df %>%
    filter(sample %in% samples) %>%
    group_by(sample) %>%
    filter({
      min_score <- min_top_scores[sample[1]]
      if (!is.na(min_score) && min_score > 0.5) {
        threshold <- floor((min_score - 1e-8) * 20) / 20  # round down to nearest 0.05
        as.numeric(as.character(score)) >= threshold
      } else {
        TRUE
      }
    }) %>%
    ungroup()
  
  # Combine
  plot_data <- bind_rows(f5_summary, other_samples)
  
  # Generate dynamic color mapping
  other_levels <- unique(plot_data$sample)
  other_levels <- setdiff(other_levels, "F5")
  sample_colors <- setNames(c("#1f78b4", RColorBrewer::brewer.pal(max(3, length(other_levels)), "Set1")[1:length(other_levels)]),
                            c("F5", other_levels))
  
  # Plot
  ggplot(plot_data, aes(Prop._h2, Enrichment, color = sample, size = -log10(Enrichment_p))) +
    geom_point() +
    geom_text(aes(label = score), vjust = -1.2, size = 4, show.legend = FALSE) +
    scale_color_manual(values = sample_colors) +
    labs(title = description) +
    theme_minimal() 
}

# For cellFacets with over 100K regions
min_top_scores <- apply(assay(cellFacet.rse, "qn.PL.score")[rowData(cellFacet.rse)$location=="distal",], 2, function(x) {
  sorted_vals <- sort(x, decreasing = TRUE)
  if (length(sorted_vals) < 100000) {
    min(sorted_vals)  # If fewer than 100k rows, take the smallest of all
  } else {
    min(sorted_vals[1:100000])  # Min of top 100,000
  }
})


plot_trait_scores(ldsc_score, "Reticulocyte count", c("reticulocyte"), min_top_scores)
ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/her.enrichment.reticulocyte.pdf",width=7.5, height=5)

plot_trait_scores(ldsc_score, "Monocyte count", c("monocyte"), min_top_scores)
ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/her.enrichment.monocyte.pdf",width=7.5, height=5)

plot_trait_scores(ldsc_score, "Basophill count", c("mast.cell"), min_top_scores)
ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/her.enrichment.basophill.pdf",width=7.5, height=5)


plot_trait_scores(ldsc_score, "Urate (umol/L)", c("kidney"), min_top_scores)
ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/her.enrichment.urate.pdf",width=7.5, height=5)
```




```{r}

ldsc_data$Category <- gsub("_top1e\\+05L2_0","",ldsc_data$Category)
ldsc_data$Category <- gsub("_0.5L2_0","",ldsc_data$Category)
cat <- c("B.lymphoblastoid_distal",
         "blood.vessel.endothelial.cell_distal",
         "brain_distal",
         "F5_distal",
         "kidney_distal",
         "liver_distal",
         "monocyte_distal",
         "neutrophil_distal",
         "reticulocyte_distal",
         "skeletal.muscle.tissue_distal",
         "spleen_distal",
         "T.cell_distal",
         "thymus_distal"
  
)


trai <- c("Alanine aminotransferase (U/L)",
          "Albumin (g/L)",
          "Apoliprotein A (g/L)",
          "Aspartate aminotransferase (U/L)",
          "Body mass index (BMI)",
          "Creatinine (umol/L)",
          "Diastolic blood pressure, automated reading",
          "Eosinophill count",
          "Ever depressed for a whole week",
          "Getting up in morning",
          "High light scatter reticulocyte count",
          "Mean corpuscular volume",
          "Monocyte count",
          "Reticulocyte count",
          "SHBG (nmol/L)",
          "Urate (umol/L)",
          "White blood cell (leukocyte) count"
  
)

#  Heatmap
df_sub <- filter(ldsc_data, Category %in% cat & description %in% trai)
m <- mean(df_sub$Enrichment_scaled)
 p <- ggplot(df_sub, aes(x = description, y = Category)) +
    geom_tile(aes(fill = Enrichment_scaled, 
                  height = Prop._h2 + 0.4, 
                  width  = Prop._h2 + 0.4),
              color = NA) +
    geom_tile(fill = NA, color = "black", width = 1, height = 1) +
    scale_fill_gradient2(
      low      = "#256866", 
      mid      = "#FFFFCC", 
      high     = "orange",
      midpoint = m, 
      na.value = "grey50", 
      name     = "Enrichment"
    ) +
    labs(
      x     = "Trait",
      y     = "Category"
    ) +
    theme_minimal() +
    theme(
      axis.text.x     = element_text(angle = 45, hjust = 1),
      axis.text.y     = element_text(size = 3),
      legend.position = "right",
      panel.grid      = element_blank()
    ) +
    coord_fixed(ratio = 1)
  
# Grouped dotplot for enrichments specifically in blood!

blood_traits <- c(
  #"White blood cell (leukocyte) count",
  #"Red blood cell (erythrocyte) count",
  #"Haemoglobin concentration",
  "Mean corpuscular volume",
  #"Mean corpuscular haemoglobin",
  #"Mean corpuscular haemoglobin concentration",
  "Red blood cell (erythrocyte) distribution width",
  #"Platelet count",
  #"Mean platelet (thrombocyte) volume",
  #"Platelet distribution width",
  "Lymphocyte count",
  "Monocyte count",
  #"Neutrophill count",
  "Eosinophill count",
  "Basophill count",
  #"Mean reticulocyte volume",
  #"Mean sphered cell volume",
  #"Immature reticulocyte fraction",
  #"High light scatter reticulocyte count",
  "Reticulocyte count"
)


immune_relevant_celltypes <- c(
  # Immune cells
  "B.lymphoblastoid_distal",
  "cord.blood.derived_distal",
  "dendritic.cell_distal",
  "lymphocyte.of.B.lineage_distal",
  "macrophage_distal",
  "mast.cell_distal",
  "monocyte_distal",
  "natural.killer.cell_distal",
  "neutrophil_distal",
  "reticulocyte_distal",
  "T.cell_distal",

  # Immune-related tissues
  "spleen_distal",
  "thymus_distal",
  "blood.vessel_distal",
  "blood.vessel.endothelial.cell_distal",
  
  # Include FANTOM5 for comparison
  "F5_distal"
)

df <- filter(ldsc_data, Category %in% immune_relevant_celltypes & description %in% blood_traits)



df_top3plusF5 <- df %>%
  group_by(description) %>%
  # get top 3 cell types excluding F5_distal
  filter(Category != "F5_distal") %>%
  slice_max(order_by = Enrichment, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  bind_rows(
    df %>%
      filter(Category == "F5_distal")
  ) %>%
  distinct(description, Category, .keep_all = TRUE)


ggplot(df_top3plusF5, aes(x = Enrichment, y = description, color = Category)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = Enrichment - Enrichment_std_error,
                     xmax = Enrichment + Enrichment_std_error),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  scale_color_manual(values = brewer.pal(12, "Set3") %>% c(brewer.pal(8, "Dark2")[1:4])) +  # Combine palettes for 16 colors
  theme_minimal() +
  labs(x = "Enrichment", y = "Trait") +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 9),
    panel.spacing = unit(1, "lines")
  ) +
  xlim(2.5,65)

ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/blood.enrichments.pdf",height=10, width=10)
```




```{r}
ldsc_data <- ldsc_data %>%
  group_by(Trait) %>%
  mutate(
    Enrichment_scaled = if (max(Enrichment, na.rm = TRUE) != min(Enrichment, na.rm = TRUE)) {
      (Enrichment - min(Enrichment, na.rm = TRUE)) /
        (max(Enrichment, na.rm = TRUE) - min(Enrichment, na.rm = TRUE))
    } else {
      0
    }
  ) %>%
  ungroup()

# Selected traits/cellFacets for figures
cat <- c("B.lymphoblastoid_distal",
         "blood.vessel.endothelial.cell_distal",
         "brain_distal",
         "F5_distal",
         "kidney_distal",
         "liver_distal",
         "monocyte_distal",
         "neutrophil_distal",
         "reticulocyte_distal",
         "skeletal.muscle.tissue_distal",
         "spleen_distal",
         "T.cell_distal",
         "thymus_distal"
  
)


trai <- c("Alanine aminotransferase (U/L)",
          "Albumin (g/L)",
          "Apoliprotein A (g/L)",
          "Aspartate aminotransferase (U/L)",
          "Body mass index (BMI)",
          "Creatinine (umol/L)",
          "Diastolic blood pressure, automated reading",
          "Eosinophill count",
          "Ever depressed for a whole week",
          "Getting up in morning",
          "High light scatter reticulocyte count",
          "Mean corpuscular volume",
          "Monocyte count",
          "Reticulocyte count",
          "SHBG (nmol/L)",
          "Urate (umol/L)",
          "White blood cell (leukocyte) count"
  
)

#  Heatmap

df_sub <- filter(ldsc_data, Category %in% cat & description %in% trai)
m <- mean(df_sub$Enrichment_scaled)
 p <- ggplot(df_sub, aes(x = description, y = Category)) +
    geom_tile(aes(fill = Enrichment_scaled, 
                  height = Prop._h2 + 0.4, 
                  width  = Prop._h2 + 0.4),
              color = NA) +
    geom_tile(fill = NA, color = "black", width = 1, height = 1) +
    scale_fill_gradient2(
      low      = "#256866", 
      mid      = "#FFFFCC", 
      high     = "orange",
      midpoint = m, 
      na.value = "grey50", 
      name     = "Enrichment"
    ) +
    labs(
      x     = "Trait",
      y     = "Category"
    ) +
    theme_minimal() +
    theme(
      axis.text.x     = element_text(angle = 45, hjust = 1),
      axis.text.y     = element_text(size = 3),
      legend.position = "right",
      panel.grid      = element_blank()
    ) +
    coord_fixed(ratio = 1)

ggsave("/projects/ralab/data/projects/nucleiCAGEproject/11.LDSC_FANTOM5/plots/heatmap.Scaledenrichments.pdf",p,height=10, width=10)

```

