---
title: "4.1.TC.perSample"
output: html_document
date: "2024-03-05"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r}
library(CAGEfightR)

#CAGEfighR extensions (Replace with library(PRIME))
setwd("/home/mhf817")
sep <- .Platform$file.sep
cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
source(paste(cfr.extensions.dir, "utils.R", sep = sep))
source(paste(cfr.extensions.dir, "support.R", sep = sep))
source(paste(cfr.extensions.dir, "cumulative.R", sep = sep))
source(paste(cfr.extensions.dir, "decompose.R", sep = sep))
source(paste(cfr.extensions.dir, "enhancers.R", sep = sep))
source(paste(cfr.extensions.dir, "noise.R", sep = sep))
source(paste(cfr.extensions.dir, "coverage.R", sep = sep))
source(paste(cfr.extensions.dir, "complexity.R", sep = sep))
source(paste(cfr.extensions.dir, "subsample.R", sep = sep))
source(paste(cfr.extensions.dir, "dispersion.R", sep = sep))


#Annotations
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

#Additional packages
library(BiocParallel)
library(dplyr)
```


## [0.2.] Set working directory
```{r}
base::setwd("/projects/ralab/data/projects/nucleiCAGEproject/4.Promoters/")
```


## [0.3.] wrapper function for calling TCs per sample
```{r}
#Function to create per sample clustering objects
perSampleTCs<- function(ctss, inputAssay="counts") {
  
  
  base::message(
    base::paste("Running TC calling for",
                base::colnames(ctss)))
  
  ## CalcPooled signal for sample and remove CTSSs with 0 score
  ctss <- CAGEfightR::calcPooled(ctss, inputAssay = inputAssay)
  
  ctss <- subset(ctss,score>0)
  
  #Call TCs for each sample
  object <- CAGEfightR::clusterUnidirectionally(ctss)
  object$sample <- base::colnames(ctss)
  
  #Return TC as GRanges
  object
}
```


# [1.] K562 replicate subsetted  -------------------------

## [1.0.] Read in K562 replicate subsetted CTSSs
```{r}
k562.s.CTSSs <- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "repSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
```

## [1.1.] Load and resize K562 DNase regions
```{r}
k562.dhs <- readr::read_delim(
      file = base::paste0(
          here::here("0.External_resources/"),
          "E123-DNase.macs2.hg38.narrowPeak"
        ),
      col_names = base::c("chr","start","end","rank"),
      col_types = readr::cols(
        .default = readr::col_integer(),
        chr =  readr::col_character(),
        rank = readr::col_skip()
      )
    ) %>%
  GenomicRanges::GRanges()

k562.dhs$thick <- IRanges::IRanges(
    start = BiocGenerics::start(k562.dhs) + (
        (BiocGenerics::end(k562.dhs) - BiocGenerics::start(k562.dhs))/2
      )
  )


k562.dhs <- k562.dhs %>% 
  CAGEfightR::swapRanges() %>% 
  GenomicRanges::promoters(upstream=200, downstream=200) 
```


##[1.2.] Per sample TC calling
```{r}
bpparam <- BiocParallel::MulticoreParam(20)

k562.TCs <- BiocParallel::bplapply(
  base::colnames(k562.s.CTSSs),function(x)
    .GlobalEnv$perSampleTCs(k562.s.CTSSs[,x]), 
  BPPARAM = bpparam)
```


##[1.3.] Annotate called TCs
```{r}
olaps <- base::lapply(
  k562.TCs, function(x) 
    GenomicRanges::countOverlaps(x,k562.dhs)) 

k562.TCs <- base::suppressWarnings(
  base::lapply(k562.TCs, function(x)
    CAGEfightR::assignTxType(x,txModels=txdb))) 
```


##[1.4.] Calculate promoter shape metrics
```{r}
k562.TCs <- BiocParallel::bplapply(
  1:base::length(k562.TCs),
  function(x) CAGEfightR::calcShape(k562.TCs[[x]],
                                    pooled=CAGEfightR::calcPooled(k562.s.CTSSs[,x])))

k562.TCs <- BiocParallel::bplapply(
  1:base::length(k562.TCs),
  function(x) CAGEfightR::calcShape(object=k562.TCs[[x]],
                                    pooled=CAGEfightR::calcPooled(k562.s.CTSSs[,x]),
                                    shapeFunction=shapeEntropy, outputColumn="shapeEntropy"))
```


##[1.5.] Format and save output
```{r}
k562.TCs <- GenomicRanges::GRangesList(k562.TCs)

k562.TCs <- BiocGenerics::unlist(k562.TCs)

k562.TCs$dhs.olap <- BiocGenerics::unlist(olaps)

k562.TCs <- k562.TCs[k562.TCs$sample!="K562_Cap_align",]

k562.TCs$width <- BiocGenerics::width(k562.TCs)

base::saveRDS(k562.TCs, file="repSub.TCs.k562.rds")
```


# [2.] K562 pooled subsetted  -------------------------

## [2.1.] Read in K562 pooled subsetted CTSSs
```{r}
k562.p.CTSSs<- base::readRDS(
    file = base::paste0(
      here::here("1.CTSSs/"),
      "poolSub.combined.k562.CTSSs.rmSingletons.rds"
    )
  )
```



##[2.2.] Per sample TC calling
```{r}
k562.p.TCs <- BiocParallel::bplapply(
  base::colnames(k562.p.CTSSs),function(x)
    .GlobalEnv$perSampleTCs(k562.p.CTSSs[,x]), 
  BPPARAM = bpparam)
```


##[2.3.] Annotate called TCs
```{r}
olaps <- base::lapply(
  k562.p.TCs, function(x) 
    GenomicRanges::countOverlaps(x,k562.dhs)) 

k562.p.TCs <- base::suppressWarnings(
  base::lapply(k562.p.TCs, function(x)
    CAGEfightR::assignTxType(x,txModels=txdb))) 
```


##[2.4.] Calculate promoter shape metrics
```{r}
k562.p.TCs <- BiocParallel::bplapply(
  1:base::length(k562.p.TCs),
  function(x) CAGEfightR::calcShape(k562.p.TCs[[x]],
                                    pooled=CAGEfightR::calcPooled(k562.p.CTSSs[,x])))

k562.p.TCs <- BiocParallel::bplapply(
  1:base::length(k562.p.TCs),
  function(x) CAGEfightR::calcShape(object=k562.p.TCs[[x]],
                                    pooled=CAGEfightR::calcPooled(k562.p.CTSSs[,x]),
                                    shapeFunction=shapeEntropy, outputColumn="shapeEntropy"))
```


##[2.5.] Format and save output
```{r}
k562.p.TCs <- GenomicRanges::GRangesList(k562.p.TCs) 

k562.p.TCs <- BiocGenerics::unlist(k562.p.TCs)

k562.p.TCs$dhs.olap <- BiocGenerics::unlist(olaps)

k562.p.TCs <- k562.p.TCs[k562.p.TCs$sample!="K562_Cap_align",]

k562.p.TCs$width <- BiocGenerics::width(k562.p.TCs)

k562.s.TCs$sample <- base::factor(
  k562.s.TCs$sample,
  levels=c("K562_C1","K562_C2","K562_C3",
           "K562_C","K562_N4","K562_N5",
           "K562_N6","K562_N","K562_10M_N1",
           "K562_10M_N2","K562_10M_N3","K562_N_10M",
           "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1",
           "K562_pcr_align"))


k562.p.TCs$sample <- base::factor(
  k562.p.TCs$sample, 
  levels=c("K562_C1","K562_C2","K562_C3",
           "K562_C","K562_N4","K562_N5",
           "K562_N6","K562_N","K562_10M_N1",
           "K562_10M_N2","K562_10M_N3","K562_N_10M",
           "K562_groCap_1","K562_ProCap_pcr_2","K562_ProCap_pcr_1",
           "K562_pcr_align"))

base::saveRDS(k562.p.TCs, file="poolSub.TCs.k562.rds")
```





#Move this part to another script



## [0.2.] Add custom color palette
```{r}
custom.col <- base::c(
  "#FFDB6D","#C4961A","#F4EDCA", 
  "#D16103","#C3D7A4","#52854C",
  "#4E84C4","#293352","grey"
  )


gm.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#DF613F","#DF5643","#DD4425",
  "#92B9C9","#76A5B6","#004352",
  "#29A957","#008E37","#007A30",
  "#4F5D5C","#444E4C","#3D4138",
  "#AC97C2","#977FB4","#7D62A3"
  )
                     
gm.pooled.pal <- base::c(
  "#FF8D00","#FF4425","#00439B",
  "#009072","#878784","#7D62D0"
  )

gm.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#92B9C9",
  "#76A5B6","#004352","#00439B",
  "#29A957","#008E37","#007A30",
  "#009072","#4F5D5C","#444E4C",
  "#3D4138","#878784","#AC97C2",
  "#977FB4","#7D62A3","#7D62D0"
  )



k.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#4F5D5C",
  "#92B9C9","#76A5B6","#004352",
  "#00439B","#29A957","#008E37",
  "#007A30"
  )
```


```{r}
#Look at shapes 

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=100) %>%
  ggplot(aes(sample,fill=sample)) + geom_bar() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3") 

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(score>=3) %>%
  ggplot(aes(txType,log10(score),fill=txType)) + geom_boxplot() + coord_flip() + theme_bw() + scale_fill_manual(values=custom.col) + labs(title="DLs detected score>=3") + facet_wrap(~sample) 

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0) %>%
  ggplot(aes(sample,log10(score),fill=sample)) + geom_boxplot() + coord_flip() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3") + facet_wrap(~txType) 

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0 & width>1 & txType=="promoter") %>%
  ggplot(aes(shapeEntropy,fill=sample)) + geom_density() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3") +facet_wrap(~sample)

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0 & score>=10  & txType=="promoter") %>%
  ggplot(aes(log10(1+width),fill=sample)) + geom_density() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="DLs detected score>=3") +facet_wrap(~sample)


k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0 & score>=3  & txType=="promoter") %>%
  ggplot(aes(log10(1+width),fill=sample)) + geom_histogram() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="promoter width [bp]") +facet_wrap(~sample)



k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0 & score>=3  & txType=="promoter") %>%
  ggplot(aes(shapeEntropy,fill=sample)) + geom_histogram() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="promoter shapeEntropy") +facet_wrap(~sample) + xlim(-1,8)

k562.TCs %>% as.data.frame(row.names=NULL) %>%  subset(dhs.olap>0 & score>=3  & txType=="promoter") %>%
  ggplot(aes(IQR,fill=sample)) + geom_histogram() + theme_bw() + scale_fill_manual(values=k.full.pal) + labs(title="promoter 10-90IQR") +facet_wrap(~sample) + scale_x_log10()



```




