---
title: "2.3.Initiator"
output: html_document
date: "2024-03-06"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
library(CAGEfightR)
library(PRIME)
library(BSgenome.Hsapiens.UCSC.hg38)
library(BiocParallel)
library(dplyr)
```


## [0.1.] Set working directory
```{r}
setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/4.Promoters/")
```


## [0.2.] Initiator classification function (now part of PRIME)
```{r}
#Specialized funtions  (loaded with PRIME)
initiatorScore <- function(object, bsg) {
  # Ensure the input is of class GRanges or StitchedGPos
  if (!inherits(object, c("GRanges", "StitchedGPos"))) {
    stop("object must be of class GRanges or StitchedGPos")
  }
  
  # Remove out-of-bound indices
  idx <- GenomicRanges:::get_out_of_bound_index(object)
  if (length(idx) > 0) {
    object <- object[-idx]
  }
  
  # Create promoters based on the class of object
  if (class(object) == "GRanges") {
    swapped_object <- swapRanges(object)
    x <- promoters(swapped_object, upstream = 1, downstream = 1)
  } else if (class(object)== "StitchedGPos") {
    x <- promoters(object, upstream = 1, downstream = 1)
  }
    
  # Get sequences
  x$seq <- getSeq(bsg, x, as.character = TRUE)
  
  # Define INR categories
  YC <- c("CC", "TC")
  YR <- c("CA", "TA", "CG", "TG")
  
  # Assign INR category
  x$INR <- "other"
  x$INR[x$seq %in% YC] <- "YC"
  x$INR[x$seq %in% YR] <- "YR"
  
  return(x)
}
```

# [1.] K562 replicate subsetted  -------------------------

## [1.0.] Read in data run classification and create output
```{r}
k562.TCs <- base::readRDS("./repSub.TCs.k562.rds")
k562.TCs<- .GlobalEnv$initiatorScore(k562.TCs,Hsapiens)
base::saveRDS(k562.TCs, file="./repSub.TCs.k562.rds")
```


# [2.] K562 pool subsetted  -------------------------

## [2.0.] Read in data run classification and create output
```{r}
k562.p.TCs <- base::readRDS("./poolSub.TCs.k562.rds")
k562.p.TCs<- .GlobalEnv$initiatorScore(k562.p.TCs[,-8],Hsapiens)
base::saveRDS(k562.p.TCs, file="./poolSub.TCs.k562.rds")
```


# [3.] GM12878 replicate subsetted  -------------------------

## [3.0.] Read in data run classification and create output
```{r}
gm12878.TCs <- base::readRDS("./repSub.TCs.GM12878.rds")
gm12878.TCs<- .GlobalEnv$initiatorScore(gm12878.TCs,Hsapiens)
base::saveRDS(gm12878.TCs, file="./repSub.TCs.GM12878.rds")
```


# [4.] GM12878 pool subsetted  -------------------------

## [4.0.] Read in data run classification and create output
```{r}
gm12878.p.TCs <- base::readRDS("./poolSub.TCs.GM12878.rds")
gm12878.p.TCs<- .GlobalEnv$initiatorScore(gm12878.p.TCs,Hsapiens)
base::saveRDS(gm12878.p.TCs, file="./poolSub.TCs.GM12878.rds")
```