---
title: "4.2.GeneExpression"
output: html_document
date: "2024-03-11"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r message=FALSE,warning=FALSE}
base::library(tidyverse)
base::library(magrittr)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(org.Hs.eg.db)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

# [1.] K562 replicate subsetted  -------------------------

## [1.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.k562.rds"
    )
  )

K562.TCs <- CAGEfightR::assignGeneID(
    object = K562.TCs,
    geneModels = txdb
  )

K562.cageExpr <- K562.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    ) # %>%
  # dplyr::filter(is.na(geneID))
```

## [1.1.] Read in and format ENCODE gene expression
```{r}
K562.geneExpr <- utils::read.table(
      file = base::paste0(
          here::here("0.External_resources/"),
          "ENCFF928NYA.tsv"
        ),
      header = TRUE
    ) %>% 
  dplyr::mutate(
    ensembl = stringr::str_replace(
        gene_id,"\\..*",""
      )
    )
```

## [1.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = K562.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
K562.cageExpr <- K562.cageExpr[
    K562.cageExpr$geneID %in% base::names(found_genes),
  ]
K562.cageExpr$ensembl <- base::as.character(found_genes)

K562.comb.expr <- dplyr::left_join(
    x = K562.cageExpr,
    y = K562.geneExpr,
    by = "ensembl"
  )
```

## [1.3.] Prepare correlation analysis
```{r}
K562.cor.data <- K562.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(K562.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% #dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [1.4.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(K562.cor.data),
      y = base::colnames(K562.cor.data)
    ) %>% 
  # dplyr::filter(row_number() %in% 1:10) %>% 
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- K562.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

## [1.5.] Visualization of pearson correlation coefficients
```{r}
cols <- grDevices::colorRampPalette(
    base::rev(RColorBrewer::brewer.pal(11,"RdBu"))
  )(100)

ggplot2::ggplot(
      data = corrs,
      mapping = aes(
        x = x,
        y = y,
        fill = cor
      )
    ) +
  geom_tile(
      stat = "identity",
      width = 0.9,
      height = 0.9
    ) +
  geom_text(
    mapping = aes(
          label = base::round(cor,2),
          color = base::ifelse(cor>0.8,"on","off")
        ),
      size = 3
    ) +
  scale_fill_gradientn(
      colors = .GlobalEnv$cols,
      limits = base::c(0L,1L)
    ) +
  scale_color_manual(
      values = base::c(on="white",off="black")
    ) +
  theme(
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5,
        size = 8
      ),
      axis.text.y = element_text(
        size = 8
      ),
      aspect.ratio = 1,
      axis.title = element_blank()
    ) +
  guides(
      fill = guide_colorbar(
        label.position = "right",
        title = NULL,
        title.position = "left",
        title.theme = element_text(
          angle = 90,
          hjust = 0.5
        ),
        ticks=T,
        nbin=100,
        barheight=5,
        label=T,
        barwidth=0.5
      ),
      color = "none"
    )
```

## [1.6.] Calculate correlations
#### Evaluate with 'x = "K562_C1"'
```{r}
corel <- base::lapply(
  X = base::unique(K562.comb.expr$sample),
  FUN = function(x) {
    samp <- K562.comb.expr[
        K562.comb.expr$sample == x &
        K562.comb.expr$TPM > 0 &
        K562.comb.expr$gene_counts > 0,
      ]
    cor <- stats::cor(
        x = base::log10(samp$gene_counts), ## includes NAs
        y = base::log10(samp$TPM) ## includes NAs
      )
    }
  )

K562.cor.df <- data.frame(
    samp = base::unique(K562.comb.expr$sample),
    correlation = base::unlist(corel)
  )
```

## [1.7.] Save files in RData container
```{r}
base::save(
  k562.cor.df,k562.comb.expr,
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.k562.GeneLevel.RData"
    )
  )
```

# [2.] K562 pool subsetted  -------------------------

## [2.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "poolSub.TCs.k562.rds"
    )
  )

K562.TCs <- CAGEfightR::assignGeneID( # 47.99 %, 29227
    object = K562.TCs,
    geneModels = txdb
  )

K562.cageExpr <- K562.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    )

dplyr::distinct(K562.cageExpr,sample)
```

## [2.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = K562.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
K562.cageExpr <- K562.cageExpr[
    K562.cageExpr$geneID %in% base::names(found_genes),
  ]
K562.cageExpr$ensembl <- base::as.character(found_genes)

K562.comb.expr <- dplyr::left_join(
    x = K562.cageExpr,
    y = K562.geneExpr,
    by = "ensembl"
  )
```

## [2.2.] Prepare correlation analysis
```{r}
K562.cor.data <- K562.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(K562.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% #dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [2.3.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(K562.cor.data),
      y = base::colnames(K562.cor.data)
    ) %>%
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- K562.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

## [2.4.] Calculate correlations
```{r}
corel <- base::lapply(
    X = base::unique(K562.comb.expr$sample),
    FUN = function(x) {
    samp <- K562.comb.expr[
        K562.comb.expr$sample == x &
        K562.comb.expr$TPM > 0 &
        K562.comb.expr$gene_counts > 0,
      ]
    cor <- stats::cor(
        base::log10(samp$gene_counts),
        base::log10(samp$TPM)
      )
    }
  )

K562.cor.df <- base::data.frame(
    samp = base::unique(k562.comb.expr$sample),
    correlation = base::unlist(corel)
  )
```

## [2.5.] Save data
```{r}
k562.p.cor.df <- k562.cor.df
k562.p.comb.expr <- k562.comb.expr

base::save(
    k562.p.cor.df,
    k562.p.comb.expr,
    file = base::paste0(
      here::here("4.Promoters/"),
      "poolSub.k562.GeneLevel.RData"
    )
  )
```

# [3.] GM12878 replicate subsetted  -------------------------

## [3.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.GM12878.rds"
    )
  )

GM12878.TCs <- CAGEfightR::assignGeneID(
    object = GM12878.TCs,
    geneModels = txdb
  )

GM12878.cageExpr <- GM12878.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    )

dplyr::distinct(GM12878.cageExpr, sample)
```

## [3.1.] Read in and format ENCODE gene expression
```{r}
GM12878.geneExpr <- utils::read.table(
      file = base::paste0(
          here::here("0.External_resources/"),
          "ENCFF345SHY.tsv"
        ),
      header = TRUE
    ) %>% 
  dplyr::mutate(
    ensembl = stringr::str_replace(
        gene_id,"\\..*",""
      )
    )
```

## [3.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
# map <- mapIds(org.Hs.eg.db,
#               keys = gm12878.cageExpr$geneID,
#               keytype="ENTREZID",
#               column = "ENSEMBL",
#               multiVals="first")
# 
# inds <- which(!is.na(map))
# found_genes <- map[inds]
# 
# gm12878.cageExpr <- gm12878.cageExpr[
#   gm12878.cageExpr$geneID %in% names(found_genes),]
# 
# gm12878.cageExpr$ensembl <- as.character(found_genes)
# 
# gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```

```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = GM12878.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
GM12878.cageExpr <- GM12878.cageExpr[
    GM12878.cageExpr$geneID %in% base::names(found_genes),
  ]
GM12878.cageExpr$ensembl <- base::as.character(found_genes)

GM12878.comb.expr <- dplyr::left_join(
    x = GM12878.cageExpr,
    y = GM12878.geneExpr,
    by = "ensembl"
  )
```

##<<bookmark>>

## [3.3.] Prepare correlation analysis
```{r}
GM12878.cor.data <- GM12878.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(GM12878.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% # dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [3.4.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(GM12878.cor.data),
      y = base::colnames(GM12878.cor.data)
    ) %>%
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- K562.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

## [3.5.] Calculate correlations
```{r}
corel <- lapply(unique(gm12878.comb.expr$sample),function(x) {
  samp <- gm12878.comb.expr[gm12878.comb.expr$sample==x & gm12878.comb.expr$TPM>0 & gm12878.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

gm12878.cor.df <- data.frame(samp=unique(gm12878.comb.expr$sample), correlation=unlist(corel))
```


## [3.4.] Write output
```{r}
save(gm12878.cor.df,gm12878.comb.expr,file="./repSub.GM12878.GeneLevel.RData")
```


# [4.] GM12878 pool subsetted  -------------------------

## [4.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS("./poolSub.TCs.GM12878.rds")

GM12878.TCs <- CAGEfightR::assignGeneID(GM12878.TCs,geneModels=txdb)

gm12878.cageExpr <- GM12878.TCs %>%
  as.data.frame(row.names=NULL) %>%
  group_by(sample, geneID) %>% 
  summarize(gene_counts=sum(score))
```



## [4.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = gm12878.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

gm12878.cageExpr <- gm12878.cageExpr[
  gm12878.cageExpr$geneID %in% names(found_genes),]

gm12878.cageExpr$ensembl <- as.character(found_genes)

gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```


## [4.2.] Calculate correlations
```{r}
corel <- lapply(unique(gm12878.comb.expr$sample),function(x) {
  samp <- gm12878.comb.expr[gm12878.comb.expr$sample==x & gm12878.comb.expr$TPM>0 & gm12878.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

gm12878.cor.df <- data.frame(samp=unique(gm12878.comb.expr$sample), correlation=unlist(corel))
```


## [4.3.] Write output
```{r}
gm12878.p.cor.df <- gm12878.cor.df
gm12878.p.comb.expr <- gm12878.comb.expr
save(gm12878.p.cor.df,gm12878.p.comb.expr,file="./poolSub.GM12878.GeneLevel.RData")
```

