---
title: "4.2.GeneExpression"
output: html_document
date: "2024-03-11"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r message=FALSE,warning=FALSE}
base::library(tidyverse)
base::library(magrittr)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(org.Hs.eg.db)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.2.] Add custom color palettes
```{r}
gm.pooled.pal <- base::c(
    "#FF8D00","#00439B","#FF4425",
    "#009072","#878784","#7D62D0",
    "hotpink"
  )
```

# [1.] K562 replicate subsetted  -------------------------

## [1.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.k562.rds"
    )
  )

K562.TCs <- CAGEfightR::assignGeneID(
    object = K562.TCs,
    geneModels = txdb
  )

K562.cageExpr <- K562.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    ) # %>%
  # dplyr::filter(is.na(geneID))
```

## [1.1.] Read in and format ENCODE gene expression
```{r}
K562.geneExpr <- utils::read.table(
      file = base::paste0(
          here::here("0.External_resources/"),
          "ENCFF928NYA.tsv"
        ),
      header = TRUE
    ) %>% 
  dplyr::mutate(
    ensembl = stringr::str_replace(
        gene_id,"\\..*",""
      )
    )
```

## [1.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = K562.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
K562.cageExpr <- K562.cageExpr[
    K562.cageExpr$geneID %in% base::names(found_genes),
  ]
K562.cageExpr$ensembl <- base::as.character(found_genes)

K562.comb.expr <- dplyr::left_join(
    x = K562.cageExpr,
    y = K562.geneExpr,
    by = "ensembl"
  )
```

## [1.3.] Prepare correlation analysis
```{r}
K562.cor.data <- K562.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(K562.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% #dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [1.4.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(K562.cor.data),
      y = base::colnames(K562.cor.data)
    ) %>% 
  # dplyr::filter(row_number() %in% 1:10) %>% 
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- K562.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

## [1.5.] Visualization of pearson correlation coefficients
```{r}
cols <- grDevices::colorRampPalette(
    base::rev(RColorBrewer::brewer.pal(11,"RdBu"))
  )(100)

ggplot2::ggplot(
      data = corrs,
      mapping = aes(
        x = x,
        y = y,
        fill = cor
      )
    ) +
  geom_tile(
      stat = "identity",
      width = 0.9,
      height = 0.9
    ) +
  geom_text(
    mapping = aes(
          label = base::round(cor,2),
          color = base::ifelse(cor>0.8,"on","off")
        ),
      size = 3
    ) +
  scale_fill_gradientn(
      colors = .GlobalEnv$cols,
      limits = base::c(0L,1L)
    ) +
  scale_color_manual(
      values = base::c(on="white",off="black")
    ) +
  theme(
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5,
        size = 8
      ),
      axis.text.y = element_text(
        size = 8
      ),
      aspect.ratio = 1,
      axis.title = element_blank()
    ) +
  guides(
      fill = guide_colorbar(
        label.position = "right",
        title = NULL,
        title.position = "left",
        title.theme = element_text(
          angle = 90,
          hjust = 0.5
        ),
        ticks=T,
        nbin=100,
        barheight=5,
        label=T,
        barwidth=0.5
      ),
      color = "none"
    )
```

## [1.6.] Calculate correlations
#### Evaluate with 'x = "K562_C1"'
```{r}
corel <- base::lapply(
  X = base::unique(K562.comb.expr$sample),
  FUN = function(x) {
    samp <- K562.comb.expr[
        K562.comb.expr$sample == x &
        K562.comb.expr$TPM > 0 &
        K562.comb.expr$gene_counts > 0,
      ]
    cor <- stats::cor(
        x = base::log10(samp$gene_counts), ## includes NAs
        y = base::log10(samp$TPM) ## includes NAs
      )
    }
  )

K562.cor.df <- data.frame(
    samp = base::unique(K562.comb.expr$sample),
    correlation = base::unlist(corel)
  )
```

## [1.7.] Save files in RData container
```{r}
base::save(
  k562.cor.df,k562.comb.expr,
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.k562.GeneLevel.RData"
    )
  )
```

# [2.] K562 pool subsetted  -------------------------

## [2.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "poolSub.TCs.k562.rds"
    )
  )

K562.TCs <- CAGEfightR::assignGeneID( # 47.99 %, 29227
    object = K562.TCs,
    geneModels = txdb
  )

K562.cageExpr <- K562.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    )

dplyr::distinct(K562.cageExpr,sample)
```

## [2.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = K562.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
K562.cageExpr <- K562.cageExpr[
    K562.cageExpr$geneID %in% base::names(found_genes),
  ]
K562.cageExpr$ensembl <- base::as.character(found_genes)

K562.comb.expr <- dplyr::left_join(
    x = K562.cageExpr,
    y = K562.geneExpr,
    by = "ensembl"
  )
```

## [2.2.] Prepare correlation analysis
```{r}
K562.cor.data <- K562.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(K562.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% #dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [2.3.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(K562.cor.data),
      y = base::colnames(K562.cor.data)
    ) %>%
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- K562.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

## [2.4.] Calculate correlations
```{r}
corel <- base::lapply(
    X = base::unique(K562.comb.expr$sample),
    FUN = function(x) {
    samp <- K562.comb.expr[
        K562.comb.expr$sample == x &
        K562.comb.expr$TPM > 0 &
        K562.comb.expr$gene_counts > 0,
      ]
    cor <- stats::cor(
        base::log10(samp$gene_counts),
        base::log10(samp$TPM)
      )
    }
  )

K562.cor.df <- base::data.frame(
    samp = base::unique(k562.comb.expr$sample),
    correlation = base::unlist(corel)
  )
```

## [2.5.] Save data
```{r}
k562.p.cor.df <- k562.cor.df
k562.p.comb.expr <- k562.comb.expr

base::save(
    k562.p.cor.df,
    k562.p.comb.expr,
    file = base::paste0(
      here::here("4.Promoters/"),
      "poolSub.k562.GeneLevel.RData"
    )
  )
```

# [3.] GM12878 replicate subsetted  -------------------------

## [3.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.GM12878.rds"
    )
  )

GM12878.TCs <- CAGEfightR::assignGeneID(
    object = GM12878.TCs,
    geneModels = txdb
  )

GM12878.cageExpr <- GM12878.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(
      gene_counts = base::sum(score),
      .groups = "drop"
    ) %>%
  dplyr::mutate(
      sample = base::as.character(sample)
    )

dplyr::distinct(GM12878.cageExpr, sample)
```

## [3.1.] Read in and format ENCODE gene expression
```{r}
GM12878.geneExpr <- utils::read.table(
      file = base::paste0(
          here::here("0.External_resources/"),
          "ENCFF345SHY.tsv"
        ),
      header = TRUE
    ) %>% 
  dplyr::mutate(
    ensembl = stringr::str_replace(
        gene_id,"\\..*",""
      )
    )
```

## [3.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
# map <- mapIds(org.Hs.eg.db,
#               keys = gm12878.cageExpr$geneID,
#               keytype="ENTREZID",
#               column = "ENSEMBL",
#               multiVals="first")
# 
# inds <- which(!is.na(map))
# found_genes <- map[inds]
# 
# gm12878.cageExpr <- gm12878.cageExpr[
#   gm12878.cageExpr$geneID %in% names(found_genes),]
# 
# gm12878.cageExpr$ensembl <- as.character(found_genes)
# 
# gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```

```{r}
map <- AnnotationDbi::mapIds(
    x = org.Hs.eg.db,
    keys = GM12878.cageExpr$geneID,
    keytype = "ENTREZID",
    column = "ENSEMBL",
    multiVals = "first"
  )

found_genes <- map[base::which(!base::is.na(map))]
GM12878.cageExpr <- GM12878.cageExpr[
    GM12878.cageExpr$geneID %in% base::names(found_genes),
  ]
GM12878.cageExpr$ensembl <- base::as.character(found_genes)

GM12878.comb.expr <- dplyr::left_join(
    x = GM12878.cageExpr,
    y = GM12878.geneExpr,
    by = "ensembl"
  )
```

## [3.3.] Prepare correlation analysis
```{r}
GM12878.cor.data <- GM12878.cageExpr %>%
  tidyr::pivot_wider(
      names_from = "sample",
      values_from = "gene_counts"
    ) %>%
  tidyr::drop_na(geneID) %>% 
  dplyr::left_join(
      x = .,
      y = dplyr::select(GM12878.geneExpr,"ensembl", "TPM"),
      by = "ensembl"
    ) %>% # dplyr::filter(is.na(geneID))
  dplyr::mutate(
    dplyr::across(
        .cols = tidyselect::everything(),
        .fns = ~ tidyr::replace_na(.,0)
      )
    ) %>%
  dplyr::select(-geneID,-ensembl)
```

## [3.4.] Pearson correlation analysis
```{r}
corrs <- base::expand.grid(
      x = base::colnames(GM12878.cor.data),
      y = base::colnames(GM12878.cor.data)
    ) %>%
  dplyr::mutate(
    cor = purrr::pmap_dbl(
      .l = base::list(x,y),
      .f = function(x,y){
        d <- GM12878.cor.data %>%
          dplyr::select(x,y) %>% 
          dplyr::filter(
              dplyr::across(
                .cols = tidyselect::everything(),
                .fns = ~ . > 0
              )
            )
        c <- stats::cor(
            x = base::log10(d[,x]),
            y = base::log10(d[,y])
          )
        return(
            base::unlist(base::unname(c))
          )
        }
      )
    )
```

#### Include also mutually exclusive signal
## To be discarded
```{r}
# corrs <- base::expand.grid(
#       x = base::colnames(GM12878.cor.data),
#       y = base::colnames(GM12878.cor.data)
#     ) %>%
#   dplyr::mutate(
#     cor = purrr::pmap_dbl(
#       .l = base::list(x,y),
#       .f = function(x,y){
#         d <- GM12878.cor.data %>%
#           dplyr::select(x,y) %>% 
#           dplyr::filter(
#               dplyr::if_any(
#                 .cols = tidyselect::everything(),
#                 .fns = ~ . > 0
#               )
#             )
#         c <- stats::cor(
#             x = d[,x], y = d[,y]
#           )
#         return(
#             base::unlist(base::unname(c))
#           )
#         }
#       )
#     )
```

#### Evaluation
## To be discarded
```{r}
# GM12878.cor.data %>%
#   dplyr::select("GM12878_10M_N","TPM") %>% 
#   dplyr::filter(
#     dplyr::if_any(
#           .cols = tidyselect::everything(),
#           .fns = ~ base::is.nan(.)
#         ) &
#       dplyr::if_any(
#         .cols = tidyselect::everything(),
#         .fns = ~ base::is.na(.)
#       )
#     )
# 
# GM12878.cor.data %>%
#   dplyr::select("GM12878_10M_N","TPM") %>% 
#   dplyr::filter(
#       dplyr::if_any(
#         .cols = tidyselect::everything(),
#         .fns = ~ . > 0
#       )
#     )
# 
# GM12878.cor.data %>%
#   dplyr::select("GM12878_10M_N","TPM") %>% 
#   dplyr::filter(
#       dplyr::if_all(
#         .cols = tidyselect::everything(),
#         .fns = ~ . > 0
#       )
#     )
# 
# 
# d <- GM12878.cor.data %>%
#   dplyr::select("GM12878_10M_N","TPM") %>% 
#   dplyr::filter(
#           dplyr::if_any(
#         .cols = tidyselect::everything(),
#         .fns = ~ . > 0
#       )
#     )
# c <- stats::cor(
#     x = d[,1],
#     y = d[,2]
#   )
```

## [3.5.] Visualization of pearson correlation coefficients
```{r}
cols <- grDevices::colorRampPalette(
    base::rev(RColorBrewer::brewer.pal(11,"RdBu"))
  )(100)
samp_use <- base::c(
    "GM12878_500K_C", "GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3",
    "GM12878_1M_C", "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3",
    "GM12878_500K_N", "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3",
    "GM12878_1M_N", "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3",
    "GM12878_5M_N", "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3",
    "GM12878_10M_N", "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3",
    "GM12878_1M_ctrl", "GM12878_ctrl_C1", "GM12878_ctrl_C2", "GM12878_ctrl_C3",
    "GM12878_1M_trip", "GM12878_triptolide_C1", "GM12878_triptolide_C2", "GM12878_triptolide_C3", "GM12878_triptolide_C4",
    "GM12878_groCap_1", "TPM"
  )
 
corrs %>%
  dplyr::filter(
      x %in% samp_use & y %in% samp_use
    ) %>% 
  dplyr::mutate(
    x = base::factor(
        x = x,
        levels = samp_use,
        ordered = TRUE
      ),
    y = base::factor(
        x = y,
        levels = samp_use,
        ordered = TRUE
      )
    ) %>%
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = x,
        y = y,
        fill = cor
      )
    ) +
  geom_tile(
      stat = "identity",
      width = 0.9,
      height = 0.9
    ) +
  geom_text(
    mapping = aes(
          label = base::round(cor,2),
          color = base::ifelse(cor>0.8,"on","off")
        ),
      size = 2
    ) +
  scale_fill_gradientn(
      colors = .GlobalEnv$cols,
      limits = base::c(0L,1L)
    ) +
  scale_color_manual(
      values = base::c(on="white",off="black")
    ) +
  theme(
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5,
        size = 8
      ),
      axis.text.y = element_text(
        size = 8
      ),
      aspect.ratio = 1,
      axis.title = element_blank()
    ) +
  guides(
      fill = guide_colorbar(
          label.position = "right",
          title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5
            ),
          ticks = T,
          nbin = 100,
          barheight = 5,
          label = TRUE,
          barwidth = 0.5
        ),
      color = "none"
    )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.TCs.GM12878",
          "_correlationPlot",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```

## [3.6.] Visualization of pearson correlation coefficients for main samples only
```{r}
main_samples <- base::c(
    "GM12878_500K_C","GM12878_1M_C","GM12878_500K_N","GM12878_1M_N",
    "GM12878_5M_N","GM12878_10M_N","GM12878_groCap_1", "TPM"
  )

corrs %>%
  dplyr::filter(
      x %in% main_samples & y %in% main_samples
    ) %>%
  dplyr::mutate(
      x = base::factor(
          x = x,
          levels = main_samples,
          ordered = TRUE
        ),
      y = base::factor(
        x = y,
        levels = main_samples,
        ordered = TRUE
      )
    ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = x,
        y = y,
        fill = cor
      )
    ) +
  geom_tile(
      stat = "identity",
      width = 0.9,
      height = 0.9
    ) +
  geom_text(
    mapping = aes(
          label = base::round(cor,2),
          color = base::ifelse(cor>0.8,"on","off")
        ),
      size = 3
    ) +
  scale_fill_gradientn(
      colors = .GlobalEnv$cols,
      limits = base::c(0L,1L)
    ) +
  scale_color_manual(
      values = base::c(on="white",off="black")
    ) +
  theme_bw() +
  theme(
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5,
        size = 8
      ),
      axis.text.y = element_text(
        size = 8
      ),
      aspect.ratio = 1,
      axis.title = element_blank()
    ) +
  guides(
      fill = guide_colorbar(
          label.position = "right",
          title = NULL,
          title.position = "left",
          title.theme = element_text(
              angle = 90,
              hjust = 0.5
            ),
          ticks = T,
          nbin = 100,
          barheight = 5,
          label = TRUE,
          barwidth = 0.5
        ),
      color = "none"
    )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.TCs.GM12878",
          "_correlationPlot",
          "_subset",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [3.7.] Calculate correlations
## To be discarded
```{r}
# corel <- base::lapply(
#     X = base::unique(GM12878.comb.expr$sample),
#     FUN = function(x) {
#       samp <- GM12878.comb.expr[
#           GM12878.comb.expr$sample == x &
#           GM12878.comb.expr$TPM > 0 &
#           GM12878.comb.expr$gene_counts > 0,
#         ]
#     cor <- stats::cor(
#         base::log10(samp$gene_counts),
#         base::log10(samp$TPM)
#       )
#     }
#   )
# 
# GM12878.cor.df <- base::data.frame(
#     samp = base::unique(GM12878.comb.expr$sample),
#     correlation = base::unlist(corel)
#   )
```

## [3.8.] Write output
## To be discarded
```{r}
# base::save(
#     gm12878.cor.df,gm12878.comb.expr,
#     file = base::paste0(
#       here::here("4.Promoters/"),
#       "repSub.GM12878.GeneLevel.RData"
#     )
#   )
```

```{r}
base::saveRDS(
    object = corrs,
    file = base::paste0(
        here::here("4.Promoters/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_repSub.TCs.GM12878",
        "_correlations",
        ".rds"
      ),
    compress = TRUE
  )
```

```{r}
corrs <- base::list.files(
      path = here::here("4.Promoters"),
      full.names = TRUE
    ) %>%
  stringr::str_subset("correlations") %>% 
  base::readRDS(
      file = .
    )
```

## [3.9.]
```{r}
coef = 0.1 # 0.3
cols <- base::c(
  grDevices::colorRampPalette(
      base::c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e")
    )(coef*100),
  grDevices::colorRampPalette(
      c("#eec73a", "#e29421", "#e29421", "#f05336","#ce472e"),
      bias=2
    )((1-coef)*100)
  )

GM12878.comb.expr %>%
    dplyr::filter(
        sample %in% main_samples
      ) %>%
    dplyr::mutate(
      sample = base::factor(
          x = sample,
          levels = main_samples,
          ordered = TRUE
        )
      ) %>% 
  ggplot2::ggplot(
      data = .,
      mapping = aes(
        x = gene_counts,
        y = TPM
      )
    ) +
  geom_hex(
      bins = 30
    ) +
  scale_x_continuous(
      trans = 'log10'
    ) +
  scale_y_continuous(
      trans = 'log10'
    ) +
  scale_fill_gradientn(
      colours = cols
    ) +
  facet_wrap(
      facets = sample ~ .,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
    )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.TCs.GM12878",
          "_scatterPlot",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 8,
      height = 3
    )
  }
```

# [4.] GM12878 pool subsetted  -------------------------

## [4.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS("./poolSub.TCs.GM12878.rds")

GM12878.TCs <- CAGEfightR::assignGeneID(GM12878.TCs,geneModels=txdb)

gm12878.cageExpr <- GM12878.TCs %>%
  as.data.frame(row.names=NULL) %>%
  group_by(sample, geneID) %>% 
  summarize(gene_counts=sum(score))
```

## [4.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = gm12878.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

gm12878.cageExpr <- gm12878.cageExpr[
  gm12878.cageExpr$geneID %in% names(found_genes),]

gm12878.cageExpr$ensembl <- as.character(found_genes)

gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```

## [4.2.] Calculate correlations
```{r}
corel <- lapply(unique(gm12878.comb.expr$sample),function(x) {
  samp <- gm12878.comb.expr[gm12878.comb.expr$sample==x & gm12878.comb.expr$TPM>0 & gm12878.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

gm12878.cor.df <- data.frame(samp=unique(gm12878.comb.expr$sample), correlation=unlist(corel))
```

## [4.3.] Write output
```{r}
gm12878.p.cor.df <- gm12878.cor.df
gm12878.p.comb.expr <- gm12878.comb.expr
save(gm12878.p.cor.df,gm12878.p.comb.expr,file="./poolSub.GM12878.GeneLevel.RData")
```

