---
title: "4.2.GeneExpression"
output: html_document
date: "2024-03-11"
---


## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
base::library(BSgenome.Hsapiens.UCSC.hg38)
base::library(org.Hs.eg.db)
base::library(AnnotationDbi)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## To be discarded
```{r}
# #CAGEfighR extensions replace with PRIME
# setwd("/home/mhf817")
# sep <- .Platform$file.sep
# cfr.extensions.dir <- "/home/mhf817/CAGEfightR_extensions"
# source(paste(cfr.extensions.dir, "utils.R", sep = sep))
# source(paste(cfr.extensions.dir, "support.R", sep = sep))
# source(paste(cfr.extensions.dir, "cumulative.R", sep = sep))
# source(paste(cfr.extensions.dir, "decompose.R", sep = sep))
# source(paste(cfr.extensions.dir, "enhancers.R", sep = sep))
# source(paste(cfr.extensions.dir, "noise.R", sep = sep))
# source(paste(cfr.extensions.dir, "coverage.R", sep = sep))
# source(paste(cfr.extensions.dir, "complexity.R", sep = sep))
# source(paste(cfr.extensions.dir, "subsample.R", sep = sep))
# source(paste(cfr.extensions.dir, "dispersion.R", sep = sep))
# library(dplyr)
# library(reshape2)
```

## [0.1.] Load data
#### Load environment: 'base::load("./repSub.GM12878.GeneLevel.RData")'; 'base::load("./repSub.K562.GeneLevel.RData")'
```{r}
GM12878.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.GM12878.rds"
    )
  )
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.k562.rds"
    )
  )
```

## [0.1.] Set working directory
## To be discarded
```{r}
# base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/4.Promoters/")
```

# [1.] K562 replicate subsetted  -------------------------

## [1.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS("./repSub.TCs.k562.rds")

K562.TCs <- CAGEfightR::assignGeneID(
    object = K562.TCs,
    geneModels = txdb
  )

k562.cageExpr <- K562.TCs %>%
  base::as.data.frame(row.names = NULL) %>%
  dplyr::group_by(sample, geneID) %>% 
  dplyr::summarize(gene_counts = base::sum(score))
```


## [1.1.] Read in and format ENCODE gene expression
```{r}
k562.geneExpr <- utils::read.table(
  "../0.External_resources/ENCFF928NYA.tsv",
  header=T)

k562.geneExpr$ensembl <- base::gsub(
  "\\..*","",
  k562.geneExpr$gene_id)
```


## [1.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = k562.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

k562.cageExpr <- k562.cageExpr[
  k562.cageExpr$geneID %in% names(found_genes),]

k562.cageExpr$ensembl <- as.character(found_genes)

k562.comb.expr <- merge(k562.cageExpr,k562.geneExpr, by.y="ensembl",by.x="ensembl")
```


## [1.3.] Calculate correlations
```{r}
corel <- lapply(unique(k562.comb.expr$sample),function(x) {
  samp <- k562.comb.expr[k562.comb.expr$sample==x & k562.comb.expr$TPM>0 & k562.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

k562.cor.df <- data.frame(samp=unique(k562.comb.expr$sample), correlation=unlist(corel))
```


## [1.4.] Write output
```{r}
save(k562.cor.df,k562.comb.expr,file="./repSub.k562.GeneLevel.RData")
```



# [2.] K562 pool subsetted  -------------------------

## [2.0.] Read in predefined TCs and quantify gene level expression
```{r}
K562.TCs <- base::readRDS("./poolSub.TCs.k562.rds")

K562.TCs <- CAGEfightR::assignGeneID(K562.TCs,geneModels=txdb)

k562.cageExpr <- K562.TCs %>%
  as.data.frame(row.names=NULL) %>%
  group_by(sample, geneID) %>% 
  summarize(gene_counts=sum(score))
```



## [2.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = k562.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

k562.cageExpr <- k562.cageExpr[
  k562.cageExpr$geneID %in% names(found_genes),]

k562.cageExpr$ensembl <- as.character(found_genes)

k562.comb.expr <- merge(k562.cageExpr,k562.geneExpr, by.y="ensembl",by.x="ensembl")
```


## [2.2.] Calculate correlations
```{r}
corel <- lapply(unique(k562.comb.expr$sample),function(x) {
  samp <- k562.comb.expr[k562.comb.expr$sample==x & k562.comb.expr$TPM>0 & k562.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

k562.cor.df <- data.frame(samp=unique(k562.comb.expr$sample), correlation=unlist(corel))
```


## [2.3.] Write output
```{r}
k562.p.cor.df <- k562.cor.df
k562.p.comb.expr <- k562.comb.expr
save(k562.p.cor.df,k562.p.comb.expr,file="./poolSub.k562.GeneLevel.RData")
```





# [3.] GM12878 replicate subsetted  -------------------------

## [3.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS("./repSub.TCs.GM12878.rds")

GM12878.TCs <- CAGEfightR::assignGeneID(GM12878.TCs,geneModels=txdb)

gm12878.cageExpr <- GM12878.TCs %>%
  as.data.frame(row.names=NULL) %>%
  group_by(sample, geneID) %>% 
  summarize(gene_counts=sum(score))
```


## [3.1.] Read in and format ENCODE gene expression
```{r}
gm12878.geneExpr <- utils::read.table(
  "../0.External_resources/ENCFF345SHY.tsv",
  header=T)

gm12878.geneExpr$ensembl <- base::gsub(
  "\\..*","",
  gm12878.geneExpr$gene_id)
```


## [3.2.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = gm12878.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

gm12878.cageExpr <- gm12878.cageExpr[
  gm12878.cageExpr$geneID %in% names(found_genes),]

gm12878.cageExpr$ensembl <- as.character(found_genes)

gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```


## [3.3.] Calculate correlations
```{r}
corel <- lapply(unique(gm12878.comb.expr$sample),function(x) {
  samp <- gm12878.comb.expr[gm12878.comb.expr$sample==x & gm12878.comb.expr$TPM>0 & gm12878.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

gm12878.cor.df <- data.frame(samp=unique(gm12878.comb.expr$sample), correlation=unlist(corel))
```


## [3.4.] Write output
```{r}
save(gm12878.cor.df,gm12878.comb.expr,file="./repSub.GM12878.GeneLevel.RData")
```



# [4.] GM12878 pool subsetted  -------------------------

## [4.0.] Read in predefined TCs and quantify gene level expression
```{r}
GM12878.TCs <- base::readRDS("./poolSub.TCs.GM12878.rds")

GM12878.TCs <- CAGEfightR::assignGeneID(GM12878.TCs,geneModels=txdb)

gm12878.cageExpr <- GM12878.TCs %>%
  as.data.frame(row.names=NULL) %>%
  group_by(sample, geneID) %>% 
  summarize(gene_counts=sum(score))
```



## [4.1.] Map IDs between CAGE and ENCODE gene expression
```{r}
map <- mapIds(org.Hs.eg.db,
              keys = gm12878.cageExpr$geneID,
              keytype="ENTREZID",
              column = "ENSEMBL",
              multiVals="first")

inds <- which(!is.na(map))
found_genes <- map[inds]

gm12878.cageExpr <- gm12878.cageExpr[
  gm12878.cageExpr$geneID %in% names(found_genes),]

gm12878.cageExpr$ensembl <- as.character(found_genes)

gm12878.comb.expr <- merge(gm12878.cageExpr,gm12878.geneExpr, by.y="ensembl",by.x="ensembl")
```


## [4.2.] Calculate correlations
```{r}
corel <- lapply(unique(gm12878.comb.expr$sample),function(x) {
  samp <- gm12878.comb.expr[gm12878.comb.expr$sample==x & gm12878.comb.expr$TPM>0 & gm12878.comb.expr$gene_counts>0,]
  cor <- cor(log10(samp$gene_counts), log10(samp$TPM))
})

gm12878.cor.df <- data.frame(samp=unique(gm12878.comb.expr$sample), correlation=unlist(corel))
```


## [4.3.] Write output
```{r}
gm12878.p.cor.df <- gm12878.cor.df
gm12878.p.comb.expr <- gm12878.comb.expr
save(gm12878.p.cor.df,gm12878.p.comb.expr,file="./poolSub.GM12878.GeneLevel.RData")
```

