---
title: "4.4.promoterPlotting"
output: html_document
date: "2024-06-27"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(dplyr)
base::library(ggplot2)
base::library(forcats)
```

## [0.1.] Load data
#### Load environment: 'base::load("./repSub.GM12878.GeneLevel.RData")'; 'base::load("./repSub.K562.GeneLevel.RData")'
```{r}
GM12878.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.GM12878.rds"
    )
  )
K562.TCs <- base::readRDS(
  file = base::paste0(
      here::here("4.Promoters/"),
      "repSub.TCs.k562.rds"
    )
  )
```

## [0.2.] Add custom color palettes
```{r}
gm.pooled.pal <- base::c(
    "#FF8D00","#00439B","#FF4425",
    "#009072","#878784","#7D62D0",
    "hotpink"
  )

gm.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#92B9C9",
    "#76A5B6","#004352","#00439B",
    "#29A957","#008E37","#007A30",
    "#009072","#4F5D5C","#444E4C",
    "#3D4138","#878784","#AC97C2",
    "#977FB4","#7D62A3","#7D62D0",
    "hotpink"
  )

k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#4F5D5C",
    "#92B9C9","#76A5B6","#004352",
    "#00439B","#29A957"
  )
```

# [1.] Plotting K562 replicate subsetted metrics -------------------------

## [1.0.] Read in TC and gene level metrics
```{r}
use <- base::levels(K562.TCs$sample)[-base::c(14,16)]
K562.TCs <- K562.TCs[K562.TCs$sample %in% use,]
```

## [1.1.] Plot number and activity metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

## [1.2.] Plot shape metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample, nrow=2) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
}
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=2) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```

## [1.3.] Plot initator metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=sample)) +
  scale_fill_manual(values=k.full.pal) +
  geom_boxplot() +
  labs(x="Sample",title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(k.full.pal)) +
  labs(x="Sample",title="promoter summit initiator >=10 counts") + 
  coord_flip() + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```

## [1.4.] Plot gene level metrics
```{r}
K562.cor.df$sample <- K562.cor.df$samp
K562.comb.expr <- K562.comb.expr[K562.comb.expr$sample %in% use,]
K562.cor.df <- K562.cor.df[K562.cor.df$sample %in% use,]
```

```{r}
K562.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=K562.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=2) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=k.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.K562.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 20,
      height = 5
    )
  }
```

# [2.] Plotting K562 pool subsetted metrics -------------------------

## [2.0.] Read in TC and gene level metrics
```{r}
rm(K562.comb.expr,K562.cor.df,K562.TCs)
base::load("./poolSub.K562.GeneLevel.RData")
K562.TCs <- base::readRDS("poolSub.TCs.K562.rds")

use <- base::c(
    "K562_C1","K562_C2","K562_C3","K562_C",
    "K562_N4","K562_N5","K562_N6","K562_N",
    "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
    "K562_groCap_1","K562_ProCap_pcr_2"
  )

K562.TCs <- K562.TCs[K562.TCs$sample %in% use,]
```

## [2.1.] Plot number and activity metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}

```

## [2.2.] Plot shape metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample, nrow=2) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=2) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```

## [2.3.] Plot initator metrics of TCs
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=sample)) +
  scale_fill_manual(values=k.full.pal) +
  geom_boxplot() +
  labs(x="Sample",title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
  }
```

```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(k.full.pal)) +
  labs(x="Sample",title="promoter summit initiator >=10 counts") + 
  coord_flip() + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```

## [2.4.] Plot gene level metrics
```{r}
K562.p.cor.df$sample <- K562.p.cor.df$samp
K562.p.comb.expr <- K562.p.comb.expr[K562.p.comb.expr$sample %in% use,]
K562.p.cor.df <- K562.p.cor.df[K562.p.cor.df$sample %in% use,]
K562.p.comb.expr$sample <- factor(K562.p.comb.expr$sample, levels=use)
K562.p.cor.df$sample <- factor(K562.p.cor.df$sample, levels=use)
```

```{r}
K562.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=K562.p.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+3,y=2.5e+4,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample, nrow=2) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=k.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```

```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.K562.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 20,
      height = 5
    )
  }
```

# [3.] Plotting GM12878 replicate subsetted metrics -------------------------

## [3.0.] Read in TC and gene level metrics
```{r}
samp_use <- base::c(
    "GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3","GM12878_500K_C",
    "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3","GM12878_1M_C",
    "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3","GM12878_500K_N",
    "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3","GM12878_1M_N",
    "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3","GM12878_5M_N",
    "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3","GM12878_10M_N",
    "GM12878_groCap_1"
  )

main_samples <- base::c(
    "GM12878_500K_C","GM12878_500K_N","GM12878_1M_C","GM12878_1M_N",
    "GM12878_5M_N","GM12878_10M_N","GM12878_groCap_1"
  )

GM12878.TCs <- GM12878.TCs[GM12878.TCs$sample %in% samp_use,]
GM12878.TCs$sample <- base::factor(
    x = GM12878.TCs$sample,
    levels = samp_use,
    ordered = TRUE
  )
GM12878.TCs.tbl <- tibble::as_tibble(
    x = GM12878.TCs,
    rownames = "ID"
  )
```

## [3.1.] Plot number of TCs
#### To be dicarded
```{r}
# GM12878.TCs.tbl %>% 
#   dplyr::filter(score >= 10) %>%
#   ggplot2::ggplot(
#     data = .,
#     mapping = aes(
#         x = sample,
#         fill = sample
#       )
#     ) +
#   geom_bar() +
#   coord_flip() +
#   scale_fill_manual(
#       values = gm.full.pal
#     ) +
#   labs(
#       title = "TCs detected score>=10"
#     ) +
#   facet_wrap(
#       facets = . ~ txType,
#       nrow = 1
#     ) +
#   theme_bw() +
#   theme(
#       aspect.ratio = 1,
#       legend.position = "none",
#       panel.spacing = grid::unit(0.1, "lines")
#       # legend.title = element_blank(),
#       # legend.justification = base::c(1,0),
#       # legend.background = element_rect(
#       #     fill = "grey95"
#       #   ),
#       # legend.text = element_text(size = 8),
#       # legend.key.size = grid::unit(0.9, "lines")
#     )
```

```{r}
cnts <- GM12878.TCs.tbl %>% 
  dplyr::mutate(
      txType = base::factor(
        x = txType,
        levels = base::c(
          "intergenic", "proximal", "promoter",
          "fiveUTR", "exon", "intron",
          "CDS", "threeUTR", "antisense"
        )
      )
    ) %>%
  dplyr::filter(sample %in% main_samples) %>% 
  dplyr::filter(score >= 10) %>%
  dplyr::group_by(sample, txType) %>% 
  dplyr::summarise(
      cnt = dplyr::n(),
      .groups = "drop"
    )
```

```{r}
ggplot2::ggplot(
    data = cnts,
    mapping = aes(
        x = sample,
        y = cnt,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  # labs(
  #     title = "TCs detected score>=10"
  #   ) +
  facet_wrap(
      facets = .~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.text.x = element_blank(),
      # axis.text = element_text(
      #     angle = 90
      #   ),
      axis.title = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1, "lines")
    )
```

## To be discarded
```{r}
# GM12878.TCs.tbl %>% 
#   dplyr::filter(sample %in% main_samples) %>% 
#   dplyr::filter(score >= 10) %>%
#   ggplot2::ggplot(
#     data = .,
#     mapping = aes(
#         x = score,
#         y = sample,
#         fill = sample
#       )
#     ) +
#   ggridges::geom_density_ridges(
#       scale = 1,
#       alpha = 0.8
#     ) +
#   scale_fill_manual(
#       values = gm.pooled.pal
#     ) +
#   scale_x_continuous(
#       limits = base::c(0,100)
#     ) +
#   # labs(
#   #     title = "TCs detected score>=10"
#   #   ) +
#   facet_wrap(
#       facets = . ~ txType,
#       nrow = 1
#     ) +
#   theme_bw() +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines"),
#       panel.spacing = grid::unit(0.1, "lines")
#     )
```

## [3.2.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.nTCs.score10",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```

## [3.3.] Number of TCs overlapping DHSs
#### To be dicarded
```{r}
# GM12878.TCs.tbl %>%
#   dplyr::filter(dhs.olap > 0) %>%
#   ggplot2::ggplot(
#     data = .,
#     mapping = aes(
#         x = sample,
#         fill = sample
#       )
#     ) +
#   geom_bar() +
#   coord_flip() +
#   scale_fill_manual(values = gm.full.pal) +
#   labs(title = "TCs detected overlapping DHSs") +
#   facet_wrap(~txType) +
#   theme_bw() +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

```{r}
cnts <- GM12878.TCs.tbl %>% 
  dplyr::filter(sample %in% main_samples) %>% 
  dplyr::filter(dhs.olap > 0) %>%
  dplyr::mutate(
      txType = base::factor(
        x = txType,
        levels = base::c(
          "intergenic", "proximal", "promoter",
          "fiveUTR", "exon", "intron",
          "CDS", "threeUTR", "antisense"
        )
      )
    ) %>%
  dplyr::group_by(sample, txType) %>%
  dplyr::summarise(
      cnt = dplyr::n(),
      .groups = "drop"
    )
```

```{r}
ggplot2::ggplot(
    data = cnts,
    mapping = aes(
        x = sample,
        y = cnt,
        fill = sample
      )
    ) +
  geom_bar(
      stat = "identity"
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  facet_wrap(
      facets = . ~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.text.x = element_blank(),
      axis.title = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1, "lines")
    )
```

## [3.4.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.nTCs.dhs",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```

## [3.5.] Number of TCs with expression (above 10)
```{r}
plot <- GM12878.TCs.tbl %>% 
  dplyr::filter(sample %in% main_samples) %>% 
  dplyr::filter(score >= 10) %>%
   dplyr::mutate(
      txType = base::factor(
        x = txType,
        levels = base::c(
          "intergenic", "proximal", "promoter",
          "fiveUTR", "exon", "intron",
          "CDS", "threeUTR", "antisense"
        )
      )
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = sample,
        y = base::log10(score),
        fill = sample
      )
    ) +
  geom_boxplot() +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  facet_wrap(
      facets = . ~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      legend.title = element_blank(),
      legend.position = "bottom",
      # legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1, "lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Boxplot",
    dpi = 300
  )
```

## [3.6.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.TCs.expression.score10",
          "_alternative",
          "_rasterise",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```

## [3.7.] Expression of only those TCs overlapping with DHSs
```{r}
plot <- GM12878.TCs.tbl %>% 
  dplyr::filter(sample %in% main_samples) %>% 
   dplyr::mutate(
      txType = base::factor(
        x = txType,
        levels = base::c(
          "intergenic", "proximal", "promoter",
          "fiveUTR", "exon", "intron",
          "CDS", "threeUTR", "antisense"
        )
      )
    ) %>%
  dplyr::filter(dhs.olap > 0) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = sample,
        y = base::log10(score),
        fill = sample
      )
    ) +
  geom_boxplot() +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  facet_wrap(
      facets = . ~ txType,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      legend.title = element_blank(),
      legend.position = "bottom",
      # legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines"),
      panel.spacing = grid::unit(0.1, "lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Boxplot",
    dpi = 300
  )
```

## [3.8.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.TCs.expression.dhs",
          "_alternative",
          "_rasterise",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 9,
      height = 4
    )
  }
```

## [3.9.] Plot shape metrics of promoters above score 10 for all samples
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(
      score >= 10 & txType == "promoter"
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = shapeEntropy,
        fill = sample
      )
    ) + 
  geom_density() +
  scale_fill_manual(
      values = gm.full.pal
    ) + 
  labs(
      title = "promoter shapeEntropy"
    ) +
  facet_wrap(~sample) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(size = 8),
      # legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.10.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

## [3.11.] Plot IQR as a measure of dispersion for promoters of all samples
```{r}
GM12878.TCs.tbl %>%
    dplyr::filter(
      score >= 10 & txType == "promoter"
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = base::log10(1+IQR),
        fill = sample
      )
    ) + 
  geom_density() + 
  scale_fill_manual(
      values = gm.full.pal
    ) + 
  labs(
      title = "promoter 10-90IQR"
    ) +
  facet_wrap(~sample) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(size = 8),
      # legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.12.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```

## [3.13.] Plot shape entopies for promoter with expression above 10
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(
      score >= 10 & txType == "promoter" & sample %in% main_samples
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = shapeEntropy,
        fill = sample
      )
    ) + 
  geom_density() +
  scale_fill_manual(
      values = gm.full.pal[base::c(4,8,12,16,20,24,25)]
    ) + 
  labs(
      title = "promoter shapeEntropy"
    ) +
  facet_wrap(
      facets = ~ sample,
      nrow = 1
    ) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(size = 8),
      # legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.14.] Alternative plotting of shape entopies for promoter with expression above 10
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(
      score >= 10 &
      txType == "promoter" &
      sample %in% main_samples
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = shapeEntropy,
        y = sample,
        fill = sample
      )
    ) + 
  ggridges::geom_density_ridges(
      scale = 3,
      alpha = 0.8
    ) +
  geom_vline(
      xintercept = base::c(2,3.3),
      linetype = "dashed",
      alpha = 0.5
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.title.y = element_blank()
    )
```

## [3.15.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.shapeEntopy.10Score.main",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [3.16.] Plot IQR for pooled samples only
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(
      score >= 10 & txType == "promoter" & sample %in% main_samples
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = base::log10(1+IQR),
        fill = sample
      )
    ) + 
  geom_density() + 
  scale_fill_manual(
      values = gm.pooled.pal
    ) + 
  labs(
      title = "promoter 10-90IQR"
    ) +
  facet_wrap(
      facets = ~sample,
      nrow = 1
    ) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(size = 8),
      # legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.17.] Alternative plotting of IQRs for pooled samples only
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(
      score >= 10 & txType == "promoter" & sample %in% main_samples
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = base::log10(IQR + 1),
        y = sample,
        fill = sample
      )
    ) + 
  ggridges::geom_density_ridges(
      scale = 3,
      alpha = 0.8
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      axis.title.y = element_blank()
    )
```

## [3.18.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0(
          here::here("4.Promoters/plots/"),
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.IQR.10Score.main",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 5,
      height = 5
    )
  }
```

## [3.19.] Plot initator metrics of TCs
```{r}
plot <- GM12878.TCs.tbl %>% 
  dplyr::filter(sample %in% main_samples) %>% 
   dplyr::mutate(
      txType = base::factor(
        x = txType,
        levels = base::c(
          "intergenic", "proximal", "promoter",
          "fiveUTR", "exon", "intron",
          "CDS", "threeUTR", "antisense"
        )
      )
    ) %>%
  dplyr::filter(
      score >= 10 & txType == "promoter"
    ) %>% 
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = sample,
        y = base::log10(score),
        fill = sample
      )
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  geom_boxplot() +
  facet_wrap(
      facets = . ~ INR,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.title.x = element_blank(),
      axis.text = element_text(
          angle = 90,
          hjust = 1,
          vjust = -1
        ),
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
    )
ggrastr::rasterise(
    input = plot,
    layers = "Boxplot",
    dpi = 300
  )
```

## To be discarded
```{r}
# GM12878.TCs.tbl %>%
#   dplyr::filter(
#       score >= 10 & txType == "promoter"
#     ) %>% 
#   ggplot2::ggplot(
#     data = .,
#     mapping = aes(
#         x = sample,
#         y = base::log10(score),
#         fill = sample
#       )
#     ) +
#   scale_fill_manual(
#       values = gm.full.pal
#     ) +
#   geom_boxplot() +
#   scale_x_discrete(
#       limits = rev(levels(GM12878.TCs$sample))
#     ) +
#   labs(
#       title = "promoter summit initiator expression >=10 counts"
#     ) +
#   facet_wrap(
#       facets = . ~ INR,
#       ncol = 1
#     ) + 
#   coord_flip() +
#   theme_bw() +
#   theme(
#       aspect.ratio = 1,
#       legend.title = element_blank(),
#       legend.justification = base::c(1,0),
#       legend.background = element_rect(
#           fill = "grey95"
#         ),
#       legend.text = element_text(size = 8),
#       legend.key.size = grid::unit(0.9, "lines")
#     )
```

## [3.20.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.initatorExpression.score10",
          "_alternative",
          "_rasterise",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 4,
      height = 3
    )
  }
```

## [3.21.] Plot number of promoter initiators with a score above 10
```{r}
GM12878.TCs.tbl %>%
  dplyr::filter(sample %in% main_samples) %>% 
  dplyr::filter(
      score >= 10 & txType == "promoter"
    ) %>%
  dplyr::group_by(sample, INR) %>% 
  dplyr::summarise(
      count = dplyr::n(),
      .groups = "drop_last"
    ) %>% 
  dplyr::mutate(
      cnt_rel = count/base::sum(count)
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        y = cnt_rel*100,
        x = sample,
        fill = sample
      )
    ) + 
  geom_bar(
      stat = "identity"
    ) +
  scale_fill_manual(
      values = gm.pooled.pal
    ) +
  facet_wrap(
      facets = . ~ INR,
      nrow = 1
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      axis.title.x = element_blank(),
      axis.text.x = element_text(
          angle = 90,
          hjust = 1,
          vjust = 0.5
        ),
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
    )
```

## [3.22.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.nInitiators.score10",
          "_alternative",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 4,
      height = 3
    )
  }
```

## [3.23.] Promoter summit expression for scores above 10
```{r}
GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names = NULL) %>%
  BiocGenerics::subset(
      score >= 10 &
      txType == "promoter" &
      sample %in% main_samples
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = fct_rev(sample),
        y = log10(score),
        fill = fct_rev(sample)
      )
    ) +
  scale_fill_manual(
      values = rev(gm.full.pal[c(4,8,12,16,20,24,25)])
    ) +
  geom_boxplot() +
  labs(
      title="promoter summit initiator expression >=10 counts"
    ) +
  facet_wrap(
      facets = . ~ INR,
      nrow = 1
    ) + 
  coord_flip() +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.position = "none",
      panel.spacing = grid::unit(0.1,"lines")
      # legend.title = element_blank(),
      # legend.justification = base::c(1,0),
      # legend.background = element_rect(
      #     fill = "grey95"
      #   ),
      # legend.text = element_text(size = 8),
      # legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.24.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.initatorExpression.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```

## [3.25.] Promoter summit initiator counts with score above 10
```{r}
GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names = NULL) %>%
  BiocGenerics::subset(
      score >= 10 &
      txType == "promoter" &
      sample %in% main_samples
    ) %>%
  dplyr::group_by(sample, INR) %>% 
  dplyr::summarise(count = dplyr::n()) %>% 
  dplyr::mutate(
      perc = count/base::sum(count)
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = INR,
        y = perc*100,
        fill = fct_rev(sample)
      )
    ) + 
  geom_col(
      position = "dodge"
    ) + 
  scale_fill_manual(
      values = base::rev(gm.full.pal[c(4,8,12,16,20,24,25)])
    ) +
  labs(
      title = "promoter summit initiator >=10 counts"
    ) + 
  coord_flip()+ 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [3.26.] Print plot
```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.promoter.nInitiators.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
  }
```

## [3.4.] Plot gene level metrics
```{r}
GM12878.cor.df$sample <- GM12878.cor.df$samp
GM12878.cor.df <- GM12878.cor.df[GM12878.cor.df$sample %in% samp_use,]
GM12878.cor.df$sample <- factor(GM12878.cor.df$sample, levels=samp_use)

GM12878.comb.expr <- GM12878.comb.expr[GM12878.comb.expr$sample %in% samp_use,]
GM12878.comb.expr$sample <- factor(GM12878.comb.expr$sample, levels=samp_use)
```


```{r}
GM12878.comb.expr %>%
  base::subset(TPM > 0 & gene_counts > 0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=GM12878.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 25
    )
  }
```


```{r}
GM12878.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=GM12878.cor.df[GM12878.cor.df$sample %in% main_samples,],
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=1) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.GM12878.geneExpression.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
  }
```


# [4.] Plotting GM12878 pooled subsetted metrics  

## [4.0.] Read in TC and gene level metrics
```{r}
rm(GM12878.TCs)
base::load("./poolSub.GM12878.GeneLevel.RData")
GM12878.TCs <- base::readRDS("poolSub.TCs.GM12878.rds")

samp_use <- c("GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3","GM12878_500K_C",
              "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3","GM12878_500K_N",
              "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3","GM12878_1M_C",
              "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3","GM12878_1M_N",
              "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3","GM12878_5M_N",
              "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3","GM12878_10M_N",
              "GM12878_groCap_1")

main_samples <- c("GM12878_500K_C","GM12878_500K_N","GM12878_1M_C","GM12878_1M_N",
                  "GM12878_5M_N","GM12878_10M_N","GM12878_groCap_1")

GM12878.TCs <- GM12878.TCs[GM12878.TCs$sample %in% samp_use,]
GM12878.TCs$sample <- factor(GM12878.TCs$sample, levels=samp_use)
```


## [4.1.] Plot number and activity metrics of TCs
```{r}
GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}

```


## [4.2.] Plot shape metrics of TCs
```{r}
GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=20 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )



for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.shapeEntopy.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.IQR.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

```




## [4.3.] Plot initator metrics of TCs
```{r}
GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(sample, log10(score),fill=sample)) +
  scale_fill_manual(values=gm.full.pal) +
  geom_boxplot() +
  scale_x_discrete(limits = rev(levels(GM12878.TCs$sample))) +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}



GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal)) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=fct_rev(sample))) +
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  geom_boxplot() +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.initatorExpression.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}



GM12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.promoter.nInitiators.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```



## [4.4.] Plot gene level metrics
```{r}
GM12878.p.cor.df$sample <- GM12878.p.cor.df$samp
GM12878.p.cor.df <- GM12878.p.cor.df[GM12878.p.cor.df$sample %in%samp_use,]
GM12878.p.cor.df$sample <- factor(GM12878.p.cor.df$sample, levels=samp_use)

GM12878.p.comb.expr <- GM12878.p.comb.expr[GM12878.p.comb.expr$sample %in% samp_use,]
GM12878.p.comb.expr$sample <- factor(GM12878.p.comb.expr$sample, levels=samp_use)


GM12878.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=GM12878.p.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 25
    )
}





GM12878.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=GM12878.p.cor.df[GM12878.p.cor.df$sample %in% main_samples,],
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=1) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.GM12878.geneExpression.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
  }
```







