---
title: "4.4.promoterPlotting"
output: html_document
date: "2024-06-27"
---

## [0.] Prerequsites -------------------------

## [0.0.] Load necessary packages
```{r}
base::library(CAGEfightR)
base::library(dplyr)
base::library(ggplot2)
base::library(forcats)
```


## [0.1.] Add custom color palettes
```{r}
custom.col <- base::c(
  "#FFDB6D","#C4961A","#F4EDCA", 
  "#D16103","#C3D7A4","#52854C",
  "#4E84C4","#293352","grey"
  )

gm.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#DF613F","#DF5643","#DD4425",
  "#92B9C9","#76A5B6","#004352",
  "#29A957","#008E37","#007A30",
  "#4F5D5C","#444E4C","#3D4138",
  "#AC97C2","#977FB4","#7D62A3"
  )
                     
gm.pooled.pal <- base::c(
  "#FF8D00","#FF4425","#00439B",
  "#009072","#878784","#7D62D0"
  )

gm.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#92B9C9",
  "#76A5B6","#004352","#00439B",
  "#29A957","#008E37","#007A30",
  "#009072","#4F5D5C","#444E4C",
  "#3D4138","#878784","#AC97C2",
  "#977FB4","#7D62A3","#7D62D0",
  "hotpink"
  )

k.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#4F5D5C",
  "#92B9C9","#76A5B6","#004352",
  "#00439B","#29A957"
  )
```


# [1.] Plotting K562 replicate subsetted metrics -------------------------

## [1.0.] Read in TC and gene level metrics
```{r}
base::load("./repSub.k562.GeneLevel.RData")
k562.TCs <- base::readRDS("repSub.TCs.k562.rds")

use <- base::levels(k562.TCs$sample)[-c(14,16)]
k562.TCs <- k562.TCs[k562.TCs$sample %in% use,]
```

## [1.1.] Plot number and activity metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```


## [1.2.] Plot shape metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample, nrow=2) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
}
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=2) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```




## [1.3.] Plot initator metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=sample)) +
  scale_fill_manual(values=k.full.pal) +
  geom_boxplot() +
  labs(x="Sample",title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(k.full.pal)) +
  labs(x="Sample",title="promoter summit initiator >=10 counts") + 
  coord_flip() + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```


## [1.4.] Plot gene level metrics
```{r}
k562.cor.df$sample <- k562.cor.df$samp
k562.comb.expr <- k562.comb.expr[k562.comb.expr$sample %in% use,]
k562.cor.df <- k562.cor.df[k562.cor.df$sample %in% use,]
```


```{r}
k562.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=k562.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=2) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=k.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.k562.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 20,
      height = 5
    )
  }
```

# [2.] Plotting K562 pool subsetted metrics -------------------------

## [2.0.] Read in TC and gene level metrics
```{r}
rm(k562.comb.expr,k562.cor.df,k562.TCs)
base::load("./poolSub.k562.GeneLevel.RData")
k562.TCs <- base::readRDS("poolSub.TCs.k562.rds")

use <- base::c(
    "K562_C1","K562_C2","K562_C3","K562_C",
    "K562_N4","K562_N5","K562_N6","K562_N",
    "K562_10M_N1","K562_10M_N2","K562_10M_N3","K562_N_10M",
    "K562_groCap_1","K562_ProCap_pcr_2"
  )

k562.TCs <- k562.TCs[k562.TCs$sample %in% use,]
```

## [2.1.] Plot number and activity metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) + 
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
  }
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=fct_rev(sample),y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=k.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    ) +
  labs(x="Sample")
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}

```

## [2.2.] Plot shape metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample, nrow=2) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=k.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=2) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 5
    )
  }
```




## [2.3.] Plot initator metrics of TCs
```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=sample)) +
  scale_fill_manual(values=k.full.pal) +
  geom_boxplot() +
  labs(x="Sample",title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
  }
```


```{r}
k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(k.full.pal)) +
  labs(x="Sample",title="promoter summit initiator >=10 counts") + 
  coord_flip() + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 10
    )
  }
```


## [2.4.] Plot gene level metrics
```{r}
k562.p.cor.df$sample <- k562.p.cor.df$samp
k562.p.comb.expr <- k562.p.comb.expr[k562.p.comb.expr$sample %in% use,]
k562.p.cor.df <- k562.p.cor.df[k562.p.cor.df$sample %in% use,]
k562.p.comb.expr$sample <- factor(k562.p.comb.expr$sample, levels=use)
k562.p.cor.df$sample <- factor(k562.p.cor.df$sample, levels=use)
```


```{r}
k562.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=k562.p.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+3,y=2.5e+4,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample, nrow=2) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=k.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```


```{r}
for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.k562.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 20,
      height = 5
    )
  }
```





# [3.] Plotting GM12878 replicate subsetted metrics  

## [3.0.] Read in TC and gene level metrics
```{r}
base::load("./repSub.GM12878.GeneLevel.RData")
gm12878.TCs <- base::readRDS("repSub.TCs.GM12878.rds")
samp_use <- c("GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3","GM12878_500K_C",
              "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3","GM12878_500K_N",
              "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3","GM12878_1M_C",
              "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3","GM12878_1M_N",
              "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3","GM12878_5M_N",
              "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3","GM12878_10M_N",
              "GM12878_groCap_1")

main_samples <- c("GM12878_500K_C","GM12878_500K_N","GM12878_1M_C","GM12878_1M_N",
                  "GM12878_5M_N","GM12878_10M_N","GM12878_groCap_1")

gm12878.TCs <- gm12878.TCs[gm12878.TCs$sample %in% samp_use,]
gm12878.TCs$sample <- factor(gm12878.TCs$sample, levels=samp_use)
```


## [3.1.] Plot number and activity metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}

```


## [3.2.] Plot shape metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )



for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.shapeEntopy.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.IQR.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

```




## [3.3.] Plot initator metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(sample, log10(score),fill=sample)) +
  scale_fill_manual(values=gm.full.pal) +
  geom_boxplot() +
  scale_x_discrete(limits = rev(levels(gm12878.TCs$sample))) +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}



gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal)) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=fct_rev(sample))) +
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  geom_boxplot() +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.initatorExpression.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}



gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.promoter.nInitiators.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```


## [3.4.] Plot gene level metrics
```{r}
gm12878.cor.df$sample <- gm12878.cor.df$samp
gm12878.cor.df <- gm12878.cor.df[gm12878.cor.df$sample %in% samp_use,]
gm12878.cor.df$sample <- factor(gm12878.cor.df$sample, levels=samp_use)

gm12878.comb.expr <- gm12878.comb.expr[gm12878.comb.expr$sample %in% samp_use,]
gm12878.comb.expr$sample <- factor(gm12878.comb.expr$sample, levels=samp_use)

gm12878.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=gm12878.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 25
    )
}




gm12878.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=gm12878.cor.df[gm12878.cor.df$sample %in% main_samples,],
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=1) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_repSub.gm12878.geneExpression.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
  }

```


# [4.] Plotting GM12878 pooled subsetted metrics  

## [4.0.] Read in TC and gene level metrics
```{r}
rm(gm12878.TCs)
base::load("./poolSub.GM12878.GeneLevel.RData")
gm12878.TCs <- base::readRDS("poolSub.TCs.GM12878.rds")

samp_use <- c("GM12878_500K_C1","GM12878_500K_C2","GM12878_500K_C3","GM12878_500K_C",
              "GM12878_500K_N1","GM12878_500K_N2","GM12878_500K_N3","GM12878_500K_N",
              "GM12878_1M_C1","GM12878_1M_C2","GM12878_1M_C3","GM12878_1M_C",
              "GM12878_1M_N1","GM12878_1M_N2","GM12878_1M_N3","GM12878_1M_N",
              "GM12878_5M_N1","GM12878_5M_N2","GM12878_5M_N3","GM12878_5M_N",
              "GM12878_10M_N1","GM12878_10M_N2","GM12878_10M_N3","GM12878_10M_N",
              "GM12878_groCap_1")

main_samples <- c("GM12878_500K_C","GM12878_500K_N","GM12878_1M_C","GM12878_1M_N",
                  "GM12878_5M_N","GM12878_10M_N","GM12878_groCap_1")

gm12878.TCs <- gm12878.TCs[gm12878.TCs$sample %in% samp_use,]
gm12878.TCs$sample <- factor(gm12878.TCs$sample, levels=samp_use)
```


## [4.1.] Plot number and activity metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected score>=10") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.nTCs.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,fill=sample)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs detected overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.nTCs.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression with >=10 score") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.TCs.expression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(dhs.olap>0) %>%
  ggplot2::ggplot(aes(x=sample,y=log10(score),fill=sample)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_manual(values=gm.full.pal) +
  labs(title="TCs expression overlapping DHSs") +
  facet_wrap(~txType) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      legend.key.size = grid::unit(0.9, "lines")
    )


for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.TCs.expression.dhs",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}

```


## [4.2.] Plot shape metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=20 & txType=="promoter") %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.shapeEntopy.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )



for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.IQR.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(x=shapeEntropy,fill=sample)) + 
  geom_density() +
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter shapeEntropy") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,7.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.shapeEntopy.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter"& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(log10(1+IQR),fill=sample)) + 
  geom_density() + 
  scale_fill_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) + 
  labs(title="promoter 10-90IQR") +
  facet_wrap(~sample,nrow=1) +
  xlim(0,2.5) + 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.IQR.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
}

```




## [4.3.] Plot initator metrics of TCs
```{r}
gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(sample, log10(score),fill=sample)) +
  scale_fill_manual(values=gm.full.pal) +
  geom_boxplot() +
  scale_x_discrete(limits = rev(levels(gm12878.TCs$sample))) +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.initatorExpression.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}



gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal)) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.nInitiators.10Score",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 15,
      height = 15
    )
}


gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  ggplot2::ggplot(aes(fct_rev(sample), log10(score),fill=fct_rev(sample))) +
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  geom_boxplot() +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR,ncol=1) + 
  coord_flip() +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.initatorExpression.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}



gm12878.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% main_samples) %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = fct_rev(sample))) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=rev(gm.full.pal[c(4,8,12,16,20,24,25)])) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.promoter.nInitiators.10Score.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 10,
      height = 15
    )
}
```



## [4.4.] Plot gene level metrics
```{r}
gm12878.p.cor.df$sample <- gm12878.p.cor.df$samp
gm12878.p.cor.df <- gm12878.p.cor.df[gm12878.p.cor.df$sample %in%samp_use,]
gm12878.p.cor.df$sample <- factor(gm12878.p.cor.df$sample, levels=samp_use)

gm12878.p.comb.expr <- gm12878.p.comb.expr[gm12878.p.comb.expr$sample %in% samp_use,]
gm12878.p.comb.expr$sample <- factor(gm12878.p.comb.expr$sample, levels=samp_use)


gm12878.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=gm12878.p.cor.df,
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.geneExpression",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 25
    )
}





gm12878.p.comb.expr %>%
  base::subset(TPM>0 & gene_counts>0& sample %in% main_samples) %>%
  ggplot2::ggplot(aes(gene_counts,TPM,color=sample,group=sample))+
  geom_point(size=0.5) +
  geom_text(data=gm12878.p.cor.df[gm12878.p.cor.df$sample %in% main_samples,],
            aes(label=paste0("r=",round(correlation,3)),x=2.5e+5,y=2.5e+5,group=sample),
            inherit.aes = FALSE,size=4) +
  facet_wrap(~sample,nrow=1) + 
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values=gm.full.pal[c(4,8,12,16,20,24,25)]) +
  theme_bw() +
    theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  ) 

for(ext in base::c("pdf", "png")){
    ggplot2::ggsave(
      filename = base::paste0("./plots/",
          base::format(base::Sys.time(), "%Y%m%d"),
          "_poolSub.gm12878.geneExpression.main",
          ".",ext
        ),
      plot = ggplot2::last_plot(),
      device = ext,
      dpi = 300,
      width = 25,
      height = 5
    )
  }
```







