---
title: "4.5. G correction evaluation"
output: html_document
date: "2024-07-01"
---


## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(PRIME)
base::library(BSgenome.Hsapiens.UCSC.hg38)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.1.] Register server cores / threads
```{r}
base::library(BiocParallel)
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 10
    )
  )
```

## [0.2.] Set color palette
```{r}
k.full.pal <- base::c(
    "#F5CE00","#E2AA00","#B68D00",
    "#FF8D00","#DF613F","#DF5643",
    "#DD4425","#FF4425","#4F5D5C",
    "#92B9C9","#76A5B6","#004352",
    "#00439B","#29A957","#008E37",
    "#007A30"
  )
```

## [0.3.] wrapper function for calling TCs per sample
#### Function to create per sample clustering objects.
```{r}
perSampleTCs<- function(ctss, inputAssay="counts") {
    
    base::message(
      base::paste0(
          "Running TC calling for ",
          base::colnames(ctss)
        )
      )
    
    ## CalcPooled signal for sample and remove CTSSs with 0 score
    ctss <- CAGEfightR::calcPooled(ctss, inputAssay = inputAssay)
    
    ctss <- subset(ctss,score>0)
    
    #Call TCs for each sample
    object <- CAGEfightR::clusterUnidirectionally(ctss)
    object$sample <- base::colnames(ctss)
    
    #Return TC as GRanges
    return(object)
  }
```

# [1.] Process K562 datasets with and without 5' G correction -------------------------

## [1.0.] Read in K562 design matrix
```{r}
fn <- base::paste0(
    here::here("1.CTSSs/"),
    "K562.design.txt"
  )
clns <- base::c( 
    "nr","ID", "ENCODE_ID",
    "bw_plus", "bw_minus","cell_count",
    "Cell.line", "Type", "replicate",
    "method", "use"
  )

design_ <- readr::read_delim(
   file =  base::paste0(
        here::here("1.CTSSs/"),
        "K562.design.txt"
      ),
   delim = "\t",
   quote = "\"",
   skip = 1,
   col_names = clns,
   col_types = readr::cols(
     .default = readr::col_character(),
     nr = readr::col_skip(),
     replicate = readr::col_integer()
   )
  )
```

## [1.1.] Data wrangling
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
design <- design_ %>%
  dplyr::mutate(
    mapping = stringr::str_extract(ID,"old"),
    mapping = dplyr::case_when(
        base::is.na(mapping) ~ "new",
        TRUE ~ "old"
      )
    ) %>% 
  tibble::column_to_rownames(var = "ID")
```

## [1.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [1.3.] Quantify CTSSs with no reference genome set
```{r}
K562.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )
```

## [1.4.] Remove non-standard chromosomes & pool replicates
```{r}
keep <- GenomeInfoDb::seqlevels(K562.CTSSs)[1:23]

K562.CTSSs <- GenomeInfoDb::keepSeqlevels(
    x = K562.CTSSs, value = keep, pruning.mode = "coarse"
  )

K562.CTSSs <- K562.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()

K562.p.CTSSs <- PRIME::poolReplicates(
  object = K562.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(K562.CTSSs)$Cell.line,
      SummarizedExperiment::colData(K562.CTSSs)$Type,
      SummarizedExperiment::colData(K562.CTSSs)$mapping,
      sep = "_"
    )
  )

K562.p.CTSSs <- CAGEfightR::calcTPM(K562.p.CTSSs)
```

## [1.5.] Subsample to min depth
```{r}
min <- base::min(SummarizedExperiment::colData(K562.p.CTSSs)$totalTags)

K562.p.s.CTSSs <- PRIME::subsampleTarget(
    object = K562.p.CTSSs, target = min
  )
```

## [1.6.] Clustering per sample
```{r}
K562.TCs <- base::lapply(
  X = base::colnames(K562.p.s.CTSSs),
  FUN = function(x){
      .GlobalEnv$perSampleTCs(K562.p.s.CTSSs[,x])
    }
  )
```

## [1.7.] Add annotation
```{r}
K562.TCs <- base::suppressWarnings(
  base::lapply(
    X = K562.TCs,
    FUN = function(x){
        CAGEfightR::assignTxType(x, txModels = txdb)
      }
    )
  )
```

## [1.8.] Classify initiator sequence
```{r}
K562.TCs <- GenomicRanges::GRangesList(K562.TCs)

K562.TCs <- BiocGenerics::unlist(K562.TCs)

K562.TCs <- PRIME::initiatorScore(object = K562.TCs, bsg = Hsapiens)
```

## [1.9.] Plot classified initiator sequence
#### 'Promoter summit initiator expression >=10 counts'
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names = NULL) %>%
  BiocGenerics::subset(
      score >= 10 & txType == "promoter"
    ) %>%
  dplyr::mutate(
    INR = base::factor(
        x = INR,
        levels = base::c("YC","YR","other"),
        ordered = TRUE
      ),
    sample = base::factor(
        x = sample,
        levels = base::c("K562_C_new","K562_N_new","K562_C_old","K562_N_old"),
        ordered = TRUE
      )
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = base::log10(score),
        y = sample,
        fill = sample
      )
    ) +
  geom_boxplot() +
  scale_fill_manual(
      values = k.full.pal[c(1,5,9,10)]
    ) +
  facet_wrap(
      facets = . ~INR,
      ncol = 1,
      strip.position = "right"
    ) + 
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      panel.spacing = grid::unit(0.1, "lines"),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.10.] Print plot
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        # here::here("4.Promoters/plots/"),
        here::here(""),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_mappingComp",
        "_K562",
        "_promoter",
        "_initatorExpression",
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    width = 4,
    height = 6
  )
```

## [1.11.] Plot classified initiator sequence
#### 'Promoter summit initiator >=10 counts'
```{r}
K562.TCs %>% 
  BiocGenerics::as.data.frame(row.names = NULL) %>%
  BiocGenerics::subset(
      score >= 10 & txType == "promoter"
    ) %>%
  dplyr::group_by(sample, INR) %>% 
  dplyr::summarise(
      count = dplyr::n(),
      .groups = "drop"
    ) %>% 
  dplyr::mutate(
    perc = count/base::sum(count),
    INR = base::factor(
        x = INR,
        levels = base::c("YC","YR","other"),
        ordered = TRUE
      ),
    sample = base::factor(
        x = sample,
        levels = base::c("K562_C_new","K562_N_new","K562_C_old","K562_N_old"),
        ordered = TRUE
      )
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = perc*100,
        y = sample,
        fill = sample
      )
    ) + 
  geom_col() +
  facet_wrap(
      facets = . ~ INR,
      ncol = 1,
      strip.position = "right"
    ) + 
  scale_fill_manual(
      values = k.full.pal[base::c(1,5,9,10)]
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      panel.spacing = grid::unit(0.1, "lines"),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

## [1.12.] Plot classified initiator sequence
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        # here::here("4.Promoters/plots/"),
        here::here(""),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_mappingComp",
        "_K562",
        "_promoter",
        "_numberInitiators",
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    width = 4,
    height = 6
  )
```

## [1.13.] Plot seqlogos
```{r}
promoters_old <- K562.TCs %>%
  BiocGenerics::subset(
      score >= 10 & txType == "promoter" &
      sample %in% base::c("K562_C_old","K562_N_old")
    ) %>%               
  CAGEfightR::swapRanges() %>%                         
  GenomicRanges::promoters(
      upstream = 5,
      downstream = 5
    )

promoter_seqs_old <- promoters_old %>%
  BSgenome::getSeq(Hsapiens, .)

promoter_seqs_old %>%                      
    base::as.character() %>%                   
    base::split(
        base::as.factor(promoters_old$INR)
      ) %>% 
    ggseqlogo::ggseqlogo(
        data = .,
        ncol = 1
      ) +
    ggseqlogo::theme_logo() +                     
    theme(
        axis.title.x = element_blank(),
        aspect.ratio = 1/2
      )
```

## [1.14.] Print seqlogos
```{r}
ggplot2::ggsave(
  filename = base::paste0(
        here::here("4.Promoters/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_mappingComp",
        "_K562",
        "_promoter",
        "_seqlogoOLD",
        ".pdf"
      ),
    plot = ggplot2::last_plot(),
    width = 4,
    height = 6
  )
```

## [1.15.] Plot seqLogos
```{r}
promoters_new <- K562.TCs %>%
  BiocGenerics::subset(
      score >= 10 & txType == "promoter" &
      sample %in% base::c("K562_C_new","K562_N_new")
    ) %>%               
  CAGEfightR::swapRanges() %>%                         
  GenomicRanges::promoters(
      upstream = 5,
      downstream = 5
    )

promoter_seqs_new <- promoters_new %>%
  BSgenome::getSeq(Hsapiens, .)

promoter_seqs_new %>%
    base::as.character() %>%
    base::split(
        base::as.factor(promoters_new$INR)
      ) %>%
    ggseqlogo::ggseqlogo(
        data = .,
        ncol = 1
      ) +
    # ggseqlogo::theme_logo() +              
    theme(
        axis.title.x = element_blank(),
        aspect.ratio = 1/2
      )
```

## [1.16.] Print seqlogos
```{r}
ggplot2::ggsave(
    filename = base::paste0(
        here::here("4.Promoters/plots/"),
        base::format(base::Sys.time(), "%Y%m%d"),
        "_mappingComp",
        "_K562",
        "_promoter",
        "_seqlogoNEW",
        ".pdf"
      ), 
    plot = ggplot2::last_plot(),
    width = 4,
    height = 6
  )
```

## [1.17.] Shift positions in old to correct for trimming
```{r}
shiftStranded <- function(x, shift = 0L,...){
    GenomicRanges::shift(x , shift = shift*base::ifelse('-' == BiocGenerics::strand(x),-1,1),...)
  }

shifted_promoters_old <- promoters_old %>%             
  CAGEfightR::swapRanges() %>%
  .GlobalEnv$shiftStranded(-1)

shifted_promoters_old$thick <- IRanges::ranges(shifted_promoters_old)

shifted_promoters_old <- PRIME::initiatorScore(
    object = shifted_promoters_old, bsg = Hsapiens
  )
shifted_promoters_old$sample <- base::paste(
    shifted_promoters_old$sample, "shifted", sep = "_"
  )

K562.TCs.2 <- base::c(K562.TCs, shifted_promoters_old)
```

## [1.18.] Plot seqLogos
#### 'Promoter summit initiator >=10 counts'
```{r}
K562.TCs.2 %>% 
  BiocGenerics::as.data.frame(row.names = NULL) %>%
  BiocGenerics::subset(
      score >= 10 & txType == "promoter"
    ) %>%
  dplyr::group_by(sample, INR) %>% 
  dplyr::summarise(count = dplyr::n()) %>% 
  dplyr::mutate(
      perc = count/base::sum(count),
      INR = base::factor(
        x = INR,
        levels = base::c("YC","YR","other"),
        ordered = TRUE
      )
    ) %>%
  ggplot2::ggplot(
    data = .,
    mapping = aes(
        x = perc*100,
        y = sample,
        fill = sample
      )
    ) + 
  geom_col() +
  facet_wrap(
      facets = . ~INR,
      ncol = 1,
      strip.position = "right"
    ) + 
  scale_fill_manual(
      values = k.full.pal[base::c(1,5,9,10,2,6,11,15)]
    ) +
  theme_bw() +
  theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.justification = base::c(1,0),
      legend.background = element_rect(
          fill = "grey95"
        ),
      legend.text = element_text(size = 8),
      panel.spacing = grid::unit(0.1, "lines"),
      legend.key.size = grid::unit(0.9, "lines")
    )
```

