---
title: "4.5.Mapping.INR"
output: html_document
date: "2024-07-01"
---


## [0.] Prerequsites -------------------------

## [0.0.] Load necesarray packages
```{r message=FALSE}
base::library(tidyverse)
base::library(CAGEfightR)
#base::library(PRIME)
base::library(ggplot2)
base::library(ggseqlogo)
base::library(BSgenome.Hsapiens.UCSC.hg38)
base::library(BiocParallel)
base::library(dplyr)
base::library(TxDb.Hsapiens.UCSC.hg38.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## [0.1.] Register server cores / threads
```{r}
BiocParallel::register(
  BPPARAM = BiocParallel::MulticoreParam(
      workers = 10
    )
  )
```

## [0.2.] Set working directory
```{r eval=FALSE}
base::setwd("/maps/projects/ralab/data/projects/nucleiCAGEproject/1.CTSSs/")
```

## [0.3.] Function for pooling replicates
```{r}
#In future included in PRIME so can be removed
poolReplicates <- function(object, replicates, inputAssay="counts") {
    
    ## dimensions
    assertthat::assert_that(
        base::ncol(object) == base::length(replicates),
        base::length(base::unique(replicates)) > 1
      )

    base::message("Calculating pooled counts...")
    a <- base::do.call(
        what = "cbind",
        args = base::lapply(
          X = base::unique(replicates),
          FUN = function(b){
            Matrix::rowSums(
                SummarizedExperiment::assay(object,inputAssay)[,replicates==b,drop=FALSE]
              )
            }
          )
        )
    a <- methods::as(a, "sparseMatrix")
    base::colnames(a) <- base::unique(replicates)

    base::message("Create a pasted design matrix...")
    c <- base::do.call(
      what = "rbind",
      args = base::lapply(
        X = base::unique(replicates),
        FUN = function(b){ 
          base::apply(
              X = SummarizedExperiment::colData(object)[replicates==b,],
              MARGIN = 2 ,
              FUN = base::paste ,
              collapse = ","
            )
          }
        )
      )

    rse <- SummarizedExperiment::SummarizedExperiment(
        assays = S4Vectors::SimpleList(counts = a),
        rowRanges = SummarizedExperiment::rowRanges(object),
        colData = base::as.data.frame(c)
      )
    rse <- CAGEfightR::calcTotalTags(rse)
    return(rse)
  }
```

## [0.4.] wrapper function for calling TCs per sample
```{r}
#Function to create per sample clustering objects
perSampleTCs<- function(ctss, inputAssay="counts") {
  
  
  base::message(
    base::paste("Running TC calling for",
                base::colnames(ctss)))
  
  ## CalcPooled signal for sample and remove CTSSs with 0 score
  ctss <- CAGEfightR::calcPooled(ctss, inputAssay = inputAssay)
  
  ctss <- subset(ctss,score>0)
  
  #Call TCs for each sample
  object <- CAGEfightR::clusterUnidirectionally(ctss)
  object$sample <- base::colnames(ctss)
  
  #Return TC as GRanges
  object
}
```

## [0.5.] wrapper function for calling TCs per sample
```{r}
#Specialized funtions  (loaded with PRIME)
initiatorScore <- function(object, bsg) {
  # Ensure the input is of class GRanges or StitchedGPos
  if (!inherits(object, c("GRanges", "StitchedGPos"))) {
    stop("object must be of class GRanges or StitchedGPos")
  }
  
  # Remove out-of-bound indices
  idx <- GenomicRanges:::get_out_of_bound_index(object)
  if (length(idx) > 0) {
    object <- object[-idx]
  }
  
  # Create promoters based on the class of object
  if (class(object) == "GRanges") {
    swapped_object <- swapRanges(object)
    x <- promoters(swapped_object, upstream = 1, downstream = 1)
  } else if (class(object)== "StitchedGPos") {
    x <- promoters(object, upstream = 1, downstream = 1)
  }
    
  # Get sequences
  x$seq <- getSeq(bsg, x, as.character = TRUE)
  
  # Define INR categories
  YC <- c("CC", "TC")
  YR <- c("CA", "TA", "CG", "TG")
  
  # Assign INR category
  x$INR <- "other"
  x$INR[x$seq %in% YC] <- "YC"
  x$INR[x$seq %in% YR] <- "YR"
  
  return(x)
}
```



# [1.] Process K562 datasets with and without 5' G correction

## [1.0.] Read in K562 design matrix
```{r}
design <- utils::read.table(
    file =  base::paste0(
        here::here("1.CTSSs/"),
        "k562.mappingCompare.design.txt"
      ),
    sep = "\t",
    header = T
  )
```

## [1.1.] Filter and format for use
#### CAGEfightR::quantifyCTSSs() requires rownames to be specified
```{r}
row.names(design) <- design$ID 

design$mapping <- c(rep("new",6),rep("old",6))
```

## [1.2.] Prepare CAGEFightR input
```{r}
plus_files <- design$bw_plus
minus_files <- design$bw_minus

base::names(plus_files) <- base::rownames(design)
base::names(minus_files) <- base::rownames(design)

bw_plus <- rtracklayer::BigWigFileList(plus_files)
bw_minus <- rtracklayer::BigWigFileList(minus_files)
```

## [1.3.] Quantify CTSSs with no reference genome set
```{r}
k562.CTSSs <- CAGEfightR::quantifyCTSSs(
    plusStrand = bw_plus,
    minusStrand = bw_minus,
    design = design
  )
```


## [1.4.] Keep only standard chromosomes and pool replicates
```{r}
keep <- GenomeInfoDb::seqlevels(k562.CTSSs)[1:23]

k562.CTSSs <- GenomeInfoDb::keepSeqlevels(k562.CTSSs, keep, pruning.mode = "coarse")


k562.CTSSs <- k562.CTSSs %>%
  CAGEfightR::calcTPM() %>%
  CAGEfightR::calcPooled()

k562.p.CTSSs <- .GlobalEnv$poolReplicates(
  object = k562.CTSSs,
  replicates = base::paste(
      SummarizedExperiment::colData(k562.CTSSs)$Cell.line,
      SummarizedExperiment::colData(k562.CTSSs)$Type,
      SummarizedExperiment::colData(k562.CTSSs)$mapping,
      sep="_"
    )
  )

k562.p.CTSSs <- CAGEfightR::calcTPM(k562.p.CTSSs)
```


## [1.5.] Subsample to min depth
```{r}
min <- base::min(colData(k562.p.CTSSs)$totalTags)

k562.p.s.CTSSs <- PRIME::subsampleTarget(k562.p.CTSSs,target=min)
```


## [1.5.] Clustering per sample
```{r}
k562.TCs <- lapply(
  base::colnames(k562.p.s.CTSSs),function(x)
    .GlobalEnv$perSampleTCs(k562.p.s.CTSSs[,x]))
```

## [1.6.] Add annotation
```{r}
k562.TCs <- base::suppressWarnings(
  base::lapply(k562.TCs, function(x)
    CAGEfightR::assignTxType(x,txModels=txdb))) 
```

## [1.7.] Classify initiator sequence
```{r}
k562.TCs <- GenomicRanges::GRangesList(k562.TCs)

k562.TCs <- BiocGenerics::unlist(k562.TCs)

k562.TCs<- .GlobalEnv$initiatorScore(k562.TCs,Hsapiens)
```

## [1.8.] Classify initiator sequence
```{r}
k.full.pal <- base::c(
  "#F5CE00","#E2AA00","#B68D00",
  "#FF8D00","#DF613F","#DF5643",
  "#DD4425","#FF4425","#4F5D5C",
  "#92B9C9","#76A5B6","#004352",
  "#00439B","#29A957","#008E37",
  "#007A30"
  )

k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  ggplot2::ggplot(aes(sample, log10(score),fill=sample)) +
  scale_fill_manual(values=k.full.pal[c(1,5,9,10)]) +
  geom_boxplot() +
  labs(title="promoter summit initiator expression >=10 counts") +
  facet_wrap(~INR) + 
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
  
ggplot2::ggsave(
  filename="../4.Promoters/plots/mappingComp.k562.promoter.initatorExpression.10Score.pdf",
  plot=last_plot(),
  width=20,
  height=5)


k562.TCs %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = sample)) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=k.full.pal[c(1,5,9,10)]) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )

ggplot2::ggsave(
  filename="../4.Promoters/plots/mappingComp.k562.promoter.nInitiators.10Score.pdf",
  plot=last_plot(),
  width=10,
  height=10)
```

## [1.9.] seqlogos
```{r}
promoters_old <- k562.TCs %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% c("K562_C_old","K562_N_old")) %>%               
  CAGEfightR::swapRanges() %>%                         
  GenomicRanges::promoters(upstream=10, downstream=10)

promoter_seqs_old <- promoters_old %>%
  BSgenome::getSeq(Hsapiens, .)

promoter_seqs_old %>%                      
    as.character %>%                   
    split(as.factor(promoters_old$INR)) %>% 
    ggseqlogo(data=., ncol=1) +
    theme_logo() +                     
    theme(axis.title.x=element_blank())

ggplot2::ggsave(
  filename="../4.Promoters/plots/mappingComp.k562.promoter.seqlogoOLD.10Score.pdf",
  plot=last_plot(),
  width=10,
  height=10)




promoters_new <- k562.TCs %>%
  BiocGenerics::subset(score>=10 & txType=="promoter" & sample %in% c("K562_C_new","K562_N_new")) %>%               
  CAGEfightR::swapRanges() %>%                         
  GenomicRanges::promoters(upstream=10, downstream=10)

promoter_seqs_new <- promoters_new %>%
  BSgenome::getSeq(Hsapiens, .)

promoter_seqs_new %>%                      
    as.character %>%                   
    split(as.factor(promoters_new$INR)) %>% 
    ggseqlogo(data=., ncol=1) +
    theme_logo() +                     
    theme(axis.title.x=element_blank())

ggplot2::ggsave(
  filename="../4.Promoters/plots/mappingComp.k562.promoter.seqlogoNEW.10Score.pdf",
  plot=last_plot(),
  width=10,
  height=10)
```


## [1.10.] Shift positions in old to correct for trimming
```{r}
shiftStranded<-function(x, shift=0L,...) GenomicRanges::shift(x ,shift=shift*ifelse('-'==strand(x),-1,1),...)
shifted_promoters_old <- promoters_old %>%             
  CAGEfightR::swapRanges() %>%
  shiftStranded(-1)

shifted_promoters_old$thick <- ranges(shifted_promoters_old)

shifted_promoters_old<- .GlobalEnv$initiatorScore(shifted_promoters_old,Hsapiens)
shifted_promoters_old$sample <- paste(shifted_promoters_old$sample,"shifted", sep="_")

k562.TCs.2 <- c(k562.TCs, shifted_promoters_old)


k562.TCs.2 %>% 
  BiocGenerics::as.data.frame(row.names=NULL) %>%
  BiocGenerics::subset(score>=10 & txType=="promoter") %>%
  group_by(sample, INR) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count)) %>%
  ggplot2::ggplot(aes(x = INR, y = perc*100, fill = sample)) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=k.full.pal[c(1,5,9,10,2,6,11,15)]) +
  labs(title="promoter summit initiator >=10 counts") + 
  coord_flip()+ 
  theme(
    aspect.ratio = 1,
    legend.title = element_blank(),
    legend.justification = base::c(1,0),
    legend.background = element_rect(
        fill = "grey95"
      ),
    legend.text = element_text(size = 8),
    legend.key.size = grid::unit(0.9, "lines")
  )
```







